### 2016-09-27 Manuscripts_GMD.tex
first commit

%% Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
%% ---------------------------------
%% This template should be used for copernicus.cls
%% The class file and some style files are bundled in the Copernicus Latex Package, which can be downloaded from the different journal webpages.
%% For further assistance please contact Copernicus Publications at: production@copernicus.org
%% http://publications.copernicus.org/for_authors/manuscript_preparation.html

%% Please use the following documentclass and journal abbreviations for discussion papers and final revised papers.

%% 2-column papers and discussion papers
\documentclass[journal abbreviation, manuscript]{copernicus}

%% Journal abbreviations (please use the same for discussion papers and final revised papers)

% Archives Animal Breeding (aab)
% Atmospheric Chemistry and Physics (acp)
% Advances in Geosciences (adgeo)
% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
% Annales Geophysicae (angeo)
% ASTRA Proceedings (ap)
% Atmospheric Measurement Techniques (amt)
% Advances in Radio Science (ars)
% Advances in Science and Research (asr)
% Biogeosciences (bg)
% Climate of the Past (cp)
% Drinking Water Engineering and Science (dwes)
% Earth System Dynamics (esd)
% Earth Surface Dynamics (esurf)
% Earth System Science Data (essd)
% Fossil Record (fr)
% Geographica Helvetica (gh)
% Geoscientific Instrumentation, Methods and Data Systems (gi)
% Geoscientific Model Development (gmd)
% Geothermal Energy Science (gtes)
% Hydrology and Earth System Sciences (hess)
% History of Geo- and Space Sciences (hgss)
% Journal of Sensors and Sensor Systems (jsss)
% Mechanical Sciences (ms)
% Natural Hazards and Earth System Sciences (nhess)
% Nonlinear Processes in Geophysics (npg)
% Ocean Science (os)
% Proceedings of the International Association of Hydrological Sciences (piahs)
% Primate Biology (pb)
% Scientific Drilling (sd)
% SOIL (soil)
% Solid Earth (se)
% The Cryosphere (tc)
% Web Ecology (we)
% Wind Energy Science (wes)

%% \usepackage commands included in the copernicus.cls:
%\usepackage[german, english]{babel}
%\usepackage{tabularx}
%\usepackage{cancel}
%\usepackage{multirow}
%\usepackage{supertabular}
%\usepackage{algorithmic}
%\usepackage{algorithm}
%\usepackage{amsthm}
%\usepackage{float}
%\usepackage{subfig}
%\usepackage{rotating}

\begin{document}

\title{Plume-SPH 1.0: A 3D dust-gas volcano plume model based on smoothed partical hydrodynamics}

% \Author[affil]{given_name}{surname}

\Author[1]{Zhixuan}{Cao}
\Author[1]{Abani}{Patra}
\Author[2]{Marcus}{Bursik}
\Author[3]{E. Bruce}{Pitman}
\Author[4]{Matthew}{Jones}
\affil[1]{Department of Mechanical and Aerospace Engineering, University at Buffalo, SUNY, New York, United State}
\affil[2]{Department of Geology, University at Buffalo, SUNY, New York, United State}
\affil[3]{Department of Mathematics, University at Buffalo, SUNY, New York, United State}
\affil[4]{Center for Computational Research,
University at Buffalo, SUNY, New York, United State}

%% The [] brackets identify the author with the corresponding affiliation. 1, 2, 3, etc. should be inserted.

\runningtitle{Plume-SPH 1.0}

\runningauthor{Z. Cao et al}

\correspondence{Abani Patra (abani.patra@gmail.com)}

\received{}
\pubdiscuss{} %% only important for two-stage journals
\revised{}
\accepted{}
\published{}

%% These dates will be inserted by Copernicus Publications during the typesetting process.

\firstpage{1}

\maketitle

\begin{abstract}
Volcanic ash, if persists in the atmosphere, will pose great threat on aircrafts and communities. VATDs (Volcanic ash transport and dispersion models) usually take output of plume models as source terms. The accuracy of source terms is crucial for VATDs. All of existing 3D (three dimensional) plume models are using mesh-based methods. SPH (Smoothed particle hydrodynamics), as a mesh free method has several advantages over mesh-based methods in modeling of multiple phase free boundary flow. As an initial effort on exploiting the feasibility and advantage of SPH in volcano plume modeling we adopt a relative simpler model, a 3D dusty-gas dynamic model, targeting at capturing the more basic features of volcano plume. Erupted material is represented by a single phase while air is another phase. Immediate dynamic and thermal dynamic equilibrium between air and erupted material are assumed in this model. Affect of wind is also ignored. Several newly developed techniques in SPH are adopted and trimmed to address numerical challenges in simulating multiphase compressible turbulent flow with classical SPH method. The $SPH-\varepsilon$ turbulence model is adopted to capture sub-particle mixing. Heat exchange due to turbulence is then calculated with Reynolds analogy. Corrected formulism of SPH is used to handle tensile instability and deficiency of particle distribution near the boundaries. We also propose feasible ways to impose velocity inlet and pressure outlet boundary conditions. The solver of this model is parallelized with MPI (message passing interface) obtaining good weak scalability and strong scalability. The model is first verified by a comparing velocity and concentration distribution along axis and cross section with experimental results of JPUE (jet or plume which is ejected from a nozzle into a uniform environment). The top height of Pinatubo eruption (15 June 1991) simulated by our model is consistent with both observation and existing 3D plume models. Profiles of several integrated variables are compared with these by existing 3D plum models and further verifies the correctness of our model. Analysis on the plume evolution process illustrated that this model is able to reproduce the basic physics of plume development. The comparison also implies that turbulence model plays a significant role in 3D volcano plume modeling.
\end{abstract}

\introduction  %% \introduction[modified heading if necessary]
The dynamics of explosive eruption clouds has been the central issue of volcanology science for a long time. Primary hazards associated with explosive grey volcanoes include pyroclastic density currents (flows and surges), the widespread deposition of airfall tephra, and the threats posed by volcanic ash. In addition, there are also other secondary hazards associated with volcanoes \citep{lockwood2013volcanoes}.
Simulation of all possible hazards with one model is difficult due to diversity among these hazards. Our concern in current work is the risks that volcanic ash posed on aircrafts and communities. During volcanic eruptions, VATDs (volcanic ash transport and dispersion models) are used to forecast the location and movement of ash clouds over hours to days. Those VATDs use eruption source parameters, such as plume height, mass eruption rate, duration, and the mass fraction distribution of erupted debris finer than about $4 \Phi$ (or $63  \mu$m), which can remain in the cloud for many hours or days. Observational data for such parameters are usually unavailable in the first minutes or hours after an eruption is detected. Moreover, these input parameters are subject to change during an eruption, requiring rapid re-assignment of new parameters. Usually, plume models are used to provide sources terms for VATDs. This paper reports a new 3D (three dimensional) volcano plume model exploiting advantage of mesh-free methods in multiple phase free boundary flow modeling.\\
%
The basic physics behind of volcanic plume has been understood for more than half a century \citep{morton1956turbulent, settle1978volcanic, wilson1978control}. Several volcano plume models have been developed in the past few decades. Specific aspects of a plume have been studied by 1D (one dimensional) plume models which describe the steady state solution of a plume under idealized boundary conditions.
The effects of magma type and vent conditions \citep{woods1988fluid, woods1995decompression}, atmospheric conditions \citep{ woods1993moist, sparks1997volcanic, bursik2001effect}, external surface water \citep{koyaguchi1996formation}, thermal disequilibrium, and particle fallout \citep{woods1991particle} have been assessed in increasing details through 1D numerical simulations. Recently developed 1D models (eg. \citep{mastin2007user, woodhouse2013interaction}) tend to accounting for more physics effects in one model. FPLUME-1.0 \citep{folch2016fplume} accounts for wind affect, entrainment of moisture, water phase change, particle fallout and re-entrainment and even wet aggregation of ash. However, In these 1D models, the entrainment of air is evaluated based two coefficients: entrainment coefficient for the vertical plume and the entrainment coefficient that describes the effect of wind. Different 1D models adopt different entrainment coefficients based on specific formulation or calibration against well-documented case studies. The feedback from plume to atmosphere is ignored in 1D models. Even though determination of essential parameters such as the entrainment is not based on first principles, such simple models nevertheless allow us to investigate the importance of physical mechanisms in a volcanic plume. In addition, these simplified models requires little computational sources and can run on standard personal computers or on web sites in very short time. As a result, 1D software (such as \citep{267, 1194, 3541}) for volcano plume development, combined with VATDs (such as \citep{114, draxler2015hysplit}) are widely used in research and practice. While these 1D models can generate well-matched results with 3D (three dimensional) models for weak plumes, much greater variability is observed for strong plume scenarios, especially for local variables \citep{costa2016results}. In addition, the need for greater skills in associated hazards forecasts with 1D models requires 2D (two dimensional) and 3D models.\\
The development of 2D and 3D, time-dependent, and multiphase numerical models for volcano plume have provided new explanations for many features of explosive volcanism. One of these 3D models is PDAC (Pyroclastic Dispersion Analysis Code) \citep{neri2003multiparticle}  which is a non-equilibrium, multiphase, 3D compressible flow model. Conservation equations for each phase is solved separately with finite volume method. And a parallel computing version of PDAC was developed lately \citep{ongaro2007parallel}. Advanced numerical techniques, such as second order scheme and semi-explicit time upgrading, was also adopted afterwards to improve the accuracy of PDAC \citep{carcano2013semi}. 
Another 3D model, SK-3D \citep{suzuki2005numerical} is a three-dimensional time-dependent fluid dynamics model that can reproduce the entrainment process of eruption clouds with relative simpler physics but high order of accuracy and high spatial resolution, which is applicable to time-dependent phenomena in actual volcanological simulations. 
A serials of simulations based on SK-3D was conducted, including determination of the entrainment coefficients of eruption columns as a function of height \citep{suzuki2010numerical}, study the effect of wind field on entrainment coefficient \citep{suzuki20133d}, establishment of the relationship between the observable quantities of the eruption clouds and the eruption conditions at the vent \citep{suzuki2009three} (previously, such relationship is based on 1D models and greatly depends on empirical constants), investigation of the effect of the intensity of turbulence in the umbrella cloud on dispersion and sedimentation of tephra \citep{koyaguchi2009effect}. 
While the SK-3D focuses on accurately capturing the entrainment caused by turbulent mixture with higher resolution and numerical method of higher order, PADC takes the disequilibrium between different phases into account and hence is a true multiphase model. Another 3D model, ATHAM ( Active Tracer High-Resolution Atmospheric Model) \citep{oberhuber1998volcanic} focuses more on microphysics within the plume. As pyroclastic flow is not the initial concern of ATHAM, immediately dynamics and thermodynamics equilibrium is assumed in ATHAM. The dynamic core of ATHAM solves the compressible Euler equations for momentum, pressure and temperature of the gas particle mixture \citep{oberhuber1998volcanic}. The subgrid turbulence closure scheme differentiates between the horizontal and vertical directions \citep{herzog2003prognostic} to capture turbulent mixing. The cloud microphysics predicts the mass of hydrometeors in liquid and ice phase \citep{herzog1998effect}. Additional modules include radiation \citep{langmann1997radiative}, gas phase chemistry \citep{trentmann2002simulation} and gas scavenging by hydrometeors \citep{textor2003injection} were added lately. An further extension was made to include particle aggregation in 2006 \citep{textor2006volcanic1, textor2006volcanic2}. However, the resolution of ATHAM is still pretty coarse compared with SK-3D and PDAC.
Besides strengthening their strength (ATHAM has been adding more and more microphysics, PDAC was extended to consider more phases), these models are learning from each other. PDAC is trying to include effect of microphysics into the model while ATHAM wants to extend its ability to modeling pyroclastic flow. Both are using finer and finer resolution. Recently, A first order, non-equilibrium compressible 3D model, ASHEE \citep{cerminara2016ashee}, is derived based on three dimensional N phases Eulerian transportation equation. ASHEE is valid for low concentration and low Strokes number region and much faster than N-phase Eulerian model. The model is based an open source numerical solver OpenFOAM adopting unstructured finite volume method.\\
To summarize, each model has its own focus based on problem of interest. Different simplification was made based on their scope. Improvements have been making since their birth. Accuracy (depends on comprehensiveness of the model, resolution of discretization, numerical error, and order of accuracy) and simulation time (depends on number of governing equations, resolution, numerical methods and parallel techniques) are always a couple of conflict considerations in 3D plume simulations.\\
%
All of these existing 3D models are based on mesh-based methods (Eulerian methods). And there is no 3D plume model based on mesh free method (Lagrange methods) yet. Actually, Lagrangian methods have several features that are suitable for volcano plume simulation. Among such Lagrangian methods, Smoothed Particle Hydrodynamics \citep{gingold1977smoothed,lucy1977numerical}, even though relatively newer compared to well designed mesh based methods such as Finite Difference, has shown good agreement with experiments by many applications in fluid dynamic area. And it currently is, by far the most widely used scheme. Specifically, the reason we choose SPH as the numerical method for volcanic plume simulation are as following:
\begin{itemize}
\item For mesh based method, either interface tracking (Lagrangian) \citep{harlow1965numerical, wrobel1991computational, cheng1995simplified} or interface capturing (Eulerian) \citep{hirt1981volume, youngs1982time, gerlach2006comparison, gopala2008volume} methods are used to reconstruct the flow interface of free boundary flow. High computational cost, a tendency to form numerical instabilities and the inability to track complex topological changes are the significant drawbacks of Lagrangian techniques \citep{hirt1981volume, unverdi1992front, anderson1998diffuse}. For interface capturing (Eulerian) method, the surrendering of surface detail before the phase transport calculation means that interface reconstruction is required between time steps to recover the interface information, which need additional numerical effort \citep{hirt1981volume, youngs1982time}. Being able to adaptively adjust the discretization and automatically construct the interface, SPH does not require additional numerical effort for interface construction and therefor more suitable for volcano plume simulation.
\item Advection term in the Navier-Stokes equations does not appear explicitly in discretized formula of SPH (as illustrated in Eq. (\ref{eq:gov-nc-rho}) to Eq. (\ref{eq:gov-nc-e})). So advection term will be treated exactly with SPH \citep{monaghan2005smoothed}.
\item It is easy to include various physics effect (like self gravity, radiative cooling and chemical reaction) into the model. It does not require a major overhaul and re-tooling every time new physics is introduced \citep{monaghan1995sph}. This implies that accounting for more physics is easier for SPH model.
\item With more than one material, each described by its own set of particles, interface problems between phases are often trivial for SPH but difficult for mesh based schemes. So multiple phase flow can be easily handled by SPH. Adding of new phases to the model also does not require a major overhaul and re-tooling. As will be shown in later paragraph, adding of new phases will lead to adding of several lines into the source code for new phases and additional interaction terms between existing phases and newly added phases.
\end{itemize}
As discussed in the previous paragraph, existing 3D plume models focus on one or several specific aspects of plume and have been extended to be more comprehensive by accounting for more physics or more phases. Easy extensibility and capability of handling multiples phase flow without additional numerical effort greatly facilitate future extension of SPH models. As volcanic plume are in nature multiphase and without pre-defined boundary in the atmosphere, SPH is a suitable numerical method for plume modeling. The more basic physics, such as entrainment of air and thermal expansion, are essential for all plume modeling while some other physics, such as water condense and aggregation ect., are important in specific scenario. As an initial effort on exploiting advantages of SPH in volcano plume modeling, we focus on capturing more basic features in plume development. So our current model is far away from comprehensive. Fortunately, the easy extension feature of SPH facilitate developing of a more comprehensive model in the future.\\
%
As SPH is a relative new method in computational fluid dynamics, implementation of SPH for compressible multiphase turbulent flow is scarce. Several remedies of classical SPH are adopted in our implementation. One of the most popular implementations of SPH is simulating of free surface flow, such as breaking-waves and flood. Less attention was paid to velocity inlet and pressure outlet boundary conditions which are required in plume modeling. We impose pressure boundary condition by adding extra layers of static ghost particles. Additional constrain on time step is used to avoid growing up of numerical fluctuations near the pressure boundary. We impose velocity inlet boundary condition by placing several layers of ghost particles moving with eruption velocity. Turbulence model is crucial for reproducing the entrainment of air. There are several turbulence models proposed for SPH method \citep{issa2005numerical,violeau2007numerical}. We adopt a LANS (Lagrangian Averaged Navier-Stokes) turbulence model \citep{monaghan2011turbulence}, which was originally proposed for incompressible, and extend it for compressible flow accounting for turbulent heat exchange. Corrected formulism of SPH is adopted to bypass the well-known tensile instability issues of classical SPH.\\
%
Simulation of volcano plume with acceptable accuracy requires fine resolutions (very high particle counts) that cannot be accomplished without parallel computing within given time window. The core solver of our model is parallelized by MPI (message passing interface standard) parallelism. In addition, a dynamic load balancing strategy is also adopted. What's more, imposing of some types of boundary conditions (such as eruption boundary condition) requires dynamically adding and removing of particles during simulation. To address this issue, we adopt an efficient data management scheme based on SFC (space filling curve) and hash table.\\
The organization of this paper is as following:
The physics model of the plume is first presented in section \ref{sec:physics-model} and leads to a complete mathematical description of the volcano plume (governing equations and boundary conditions). In section \ref{sec:SPH_method} we will briefly introduce the numerical tool, SPH method. Both the fundamental discretization formulism and techniques that used to handle specific issues involved in plume modeling will be discussed. Validation and verification with numerical tests is presented in section \ref{sec:verification-validation}. In section \ref{sec:conclusion}, a discussion on future work is given following a brief summary.
\section{Physical Model} \label{sec:physics-model}
\subsection{Description of the model}
During an explosive eruption, a volcanic jet erupts out from a vent with a speed of several tens to more than 150 meters per second, driven by expanding gas. The jet is initially denser than the surrounding atmosphere and begins to decelerate through negative buoyancy and turbulent interaction with surrounding air. Cauliflower-like vortices are generated along jet margins, within which process, air is entrained and heated up, reducing the bulk density of the entire jet, in many cases, to less than that of the surrounding atmosphere. Once it becomes buoyant, such a jet develops into a plinian or subplinian plume, rising kilometers to tens of kilometers until its heat is exhausted that buoyancy is lost. Jets that lose their momentum before becoming buoyant collapse back onto the ground and transform into pyroclastic flows, surges and ignimbrites. Simutaneously, relative coarse particle might separate with main stream of the plume falling down onto the ground and may be re-entrained into the plume at a lower height \citep{ernst1996sedimentation}. Within this process, erupted vapor will condense to liquid (droplet) and even further to ice. Latent heat released from phase change of erupted vapor will further heat up entrained air and make the bulk density further dilute. Even entrained vapor might also experience similar process and impose impact on plume development. Particle aggregation process \citep{carey1982influence,taddeucci2011aggregation}, either due to presence of liquid water, resulting from particle collision or driven by electrostatic forces might occur inside plume and thereby affect the sedimentation. 
This is essentially a multiple phases turbulent mixing process coupled with heat transfer and many other microphysics. As an initial effort on exploiting the feasibility and advantage of SPH in plume modeling, our model is designed to describe an injection of well mixed solid and volcanic gas from a circular vent above a flat surface into a stratified stationary atmosphere following SK-3D \citep{suzuki2005numerical}. In this model, molecular viscosity and physics heat conduction is neglected since turbulent exchange coefficients are dominant. Erupted material consist of solid with different size and mixture of gases (various constituent) is assumed to be well mixed and behave like a single phase fluid (phase 2) which is valid for eruptions with fine particles and ash. Air (also a mixture of different gases) is assumed to be another phase (phase 1). Immediate thermodynamics equilibrium is assumed so that no separate energy equation is needed for each phase. As a result, there is only one energy equation for both phases (heat exchange term between different phases will not show up under this assumption). Immediate dynamics equilibrium is assumed so that no separate momentum equation is needed for each phase. As a result, there is only one vector momentum equation for both phases (drag force term will not show up with this assumption). Because of above assumptions, all other micro-physics process (like phase change of  \texorpdfstring{H\textsubscript{2}O}, aggregation, decomposition, absorption of gas on the surface of solid, solution of gas into the liquid) and chemical process will not be considered in this model. These microphysics factors will come to play critical roles under particular eruption conditions. Capturing of these processes need a more comprehensive model. One critical element in plume development, the affect of wind, is also not considered in this model yet. Introducing of wind effect in our model will need imposing of pressure boundary conditions in a dynamic way, which requires more numerical effort and algorithm design. To summarize, our model is not valid for eruptions where wind affect plays a significant role in its development, usually refereed as weak plume. Our model is also lack ability in modeling plumes with large particles or in which microphysics plays non-ignorable roles, such as a eruption of \textit{El Chich{\'o}n} volcano at April 4th \citep{sigurdsson19841982, folch2016fplume}. At present stage, we are more interested in more basic (and more critical) aspects of volcano plume and therefor devote our effort to this relative simpler model. It is worthwhile to mention here that because SPH is adopted as our numerical method, adding of these physics into our model would require much less work in terms of programming compared to mesh based methods.\\
\subsection{Governing equations}
Based above assumptions, the governing equation of our model is given as following, which is the same as governing equations of SK-3D \citep{suzuki2005numerical}.
\begin{eqnarray}
\dfrac{\partial \rho}{\partial t} + \nabla \cdot (\rho \textbf{v}) &=& 0 \label{eq:gov-cs-rho} \\
\dfrac{\partial \rho \xi}{\partial t} + \nabla \cdot (\rho \xi \textbf{v}) &=& 0 \label{eq:gov-cs-ks}\\
\dfrac{\partial \rho \textbf{v}}{\partial t} + \nabla \cdot (\rho \textbf{v} \textbf{v} + p\textbf{I}) &=& \rho \textbf{g} \label{eq:gov-cs-v} \\
\dfrac{\partial \rho E}{\partial t} + \nabla \cdot [(\rho E + p )\textbf{v}] &=& \rho \textbf{g} \cdot\textbf{v} \label{eq:gov-cs-e}
\end{eqnarray}
$\xi$ is the mass fraction of ejected material.
$E = e + K $ is total energy which is summation of kinetic energy $K$ and internal energy $e$.
An additional equation is required to close the system. In this model, the equation for closing the system is the EOS (equation of state) for ideal gas.
\begin{equation}
p = (\gamma_m - 1)\rho e \label{eq:EOS}
\end{equation}
Where 
\begin{equation}
\gamma_m = R_m/C_{vm} + 1 \label{eq:gov-gm}
\end{equation}
\begin{equation}
R_m = n_gR_g + n_aR_a  \label{eq:gov-Rm}
\end{equation}
\begin{equation}
C_{vm} = n_s C_{vs} + n_g C_{vg} + n_a C_{va} \label{eq:gov-Cvm}
\end{equation}
\begin{equation}
n_a = 1 - \xi \label{eq:gov-na}
\end{equation}
\begin{equation}
n_g = \xi n_{g0} \label{eq:gov-ng}
\end{equation}
\begin{equation}
n_s = \xi - n_g \label{eq:gov-ns}
\end{equation}
Where, $C_v$ is specific heat with constant volume, $n$ is mass fraction, $R$ is gas constant. The subscript 
$m$ represents mixture of ejected material and air, $s$ represents solid portion in ejected material, $g$ represents gas portion in the ejected material and $a$ represents air.
In mesh based method, governing equations in Eulerian form, Eq. (\ref{eq:gov-cs-rho}) to Eq. (\ref{eq:gov-cs-e}), will be directly discretized. For SPH, governing equations in Lagrange form is needed. By deducting Kinetic energy from energy equation, subtracting mass conservation from momentum equation, combining transient term and advection term into material derivative term, the governing equation is put into the final formulism, in which advection term does not show up explicitly.
\begin{eqnarray}
\dfrac{D \rho}{D t} + \rho \nabla \cdot \textbf{v} = 0 \label{eq:gov-nc-rho}\\
\dfrac{D \rho \xi}{D t} + \rho \xi \nabla \cdot \textbf{v} = 0 \label{eq:gov-nc-ks}\\
\dfrac{D \textbf{v}}{D t} + \dfrac{\nabla P}{\rho} =\textbf{g} \label{eq:gov-nc-v}\\
\dfrac{D e}{D t} + \dfrac{P \nabla \cdot \textbf{v}}{\rho} = 0 \label{eq:gov-nc-e}
\end{eqnarray}
Governing equations in Lagrange form, Eq. (\ref{eq:gov-nc-rho}) to Eq.(\ref{eq:gov-nc-e}), and Eulerian form, Eq. (\ref{eq:gov-cs-rho}) to Eq. (\ref{eq:gov-cs-e}), are actually equivalent as deriving of Lagrange form from Eulerian form is purely based on mathematics without any additional physics assumption.
\subsection{Boundary conditions}
In current model the initial domain is a box. The boundaries are categorized into velocity inlet (a circle area at the center of the bottom of the box), wall boundary (box bottom), pressure outlet of boundary (other faces of the box).
\subsubsection{Velocity inlet}
At the vent, temperature of erupted material $T$, eruption velocity $\textbf{v}$, the mass fraction of vapor in erupted material $n_{g0}$ and mass discharge rate $\dot M$ is given. The pressure of erupted material $p$ is assumed to be the same as ambient pressure for pressure-balanced eruption. The radius of vent is determined from $\rho$, $\dot M$ and $\textbf{v}$. To write them in terms of primitive variables. 
\begin{eqnarray}
\rho =const = p/(R_m T) \label{eq:erupt_bc_rho} \\
\xi=const=1 \label{eq:erupt_bc_xi}\\
\textbf{v} = const =\{u,v,w\}^T \label{eq:erupt_bc_v}\\
\dfrac{\partial e}{\partial n}=\dot M e /(\pi r^2) \label{eq:erupt_bc_e}
\end{eqnarray} 
Where $r$ is the radius of the vent.
\subsubsection{Non-slip wall boundary}
Velocity is zero for non-slip wall boundary. If assume the boundary to be adiabatic, heat flux should be zero on the boundary. The flux of mass should also be zero. As a result, internal energy flux, which consists of heat flux and energy flux carried by mass flux, will vanish on the wall boundary. So the wall boundary condition in terms of primitive variables are:
\begin{eqnarray}
\dfrac{\partial \rho}{\partial n} = const = 0\label{eq:wall_bc_rho} \\
\dfrac{\partial \xi}{\partial n} = const = 0 \label{eq:wall_bc_xi}\\ 
\textbf{v} = const =\{0,0,0\}^T \label{eq:wall_bc_v}\\
\dfrac{\partial e }{\partial n} = 0\label{eq:wall_bc_e}
\end{eqnarray} 
\subsubsection{Open outlet pressure boundary condition}
The pressure of the surrounding atmosphere is given. Except for the pressure, boundary conditions for density, velocity, and energy on the outlet should depend on the solution. As we ignore the viscosity, the shear stress is gone and norm stress (whose magnitude equals to pressure) will balance with ambient pressure.
\begin{equation}
p = p_a(h)\label{eq:pressure_bc_p} 
\end{equation} 
%
\section{SPH Method} \label{sec:SPH_method}
SPH is a meshfree Lagragian method. In SPH, the domain is discretized as a set of particles or discretization points and the position of each particle (discretization point) is updated at every time step. Approximation of all field variables (velocity, density and pressure, ect.) are obtained by interpolation based on discretization points. The physical laws (such as conservation laws of mass, momentum and energy) written in the form of PDEs (partial differential equations) need to be transformed into the Lagrangian particle formalism of SPH. Using a kernel function that provides the weighted estimation of the field variables at any point, the integral equations are evaluated as sums over neighbor particles. Thus, physical properties (density, velocity, internal energy, ect.) associated to the particle are updated based on its neighbors. The kernel function has a compact support determined by smoothing length of each particle. There are several review papers \citep{monaghan1992smoothed, monaghan2005smoothed, rosswog2009astrophysical, price2012smoothed, monaghan2012smoothed}, giving a pretty comprehensive view over SPH, we will only refer here to the representation of the constitutive equations in SPH and focus on numerical techniques for plume modeling.
\subsection{Fundametal principle}
There are several routines for discretizing governing equations (partitial differential equations or ordernary differential equations) with SPH. There is slight difference between discretized formulism obtained by different routines. We present here one of them following JJ Monaghan \citep{monaghan1992smoothed, monaghan2005smoothed}. The start point of approximating a function with SPH is one translation property of Dirac function.

\begin{equation}
A(\textbf{x})=\int_{-\infty}^{\infty} A(\textbf{x} \prime) \delta (\textbf{x} \prime - \textbf{x}) d\textbf{x} \prime
\label{eq:Dirac-translation}
\end{equation}
The Dirac function in Eq. (\ref{eq:Dirac-translation}) can be approximated by a weighting funtion $w(\textbf{x}-\textbf{x}\prime, h)$ (or $w(\textbf{x}\prime-\textbf{x}, h)$) which will tend to a Dirac function when $h \rightarrow 0$ :
\begin{equation}
\lim _{h \rightarrow 0} w(\textbf{x} \prime-\textbf{x}, h) =  \delta (\textbf{x} \prime - \textbf{x})
\label{eq:SPH_kernel_delta}
\end{equation}
The function $A(\textbf{x})$ then can be approximated by:
\begin{equation}
A(\textbf{x}) \approx <A(\textbf{x})> = \int_{\Omega} A(\textbf{x} \prime) w(\textbf{x}-\textbf{x}\prime, h) d\textbf{x}\prime + O(h^2)
\label{eq:SPH-fundamental-principle}
\end{equation}
Where $h$ is the smoothing length, determining the interaction distance. According to Eq. (\ref{eq:SPH-fundamental-principle}), the order of accuracy of SPH method is second order. However, in practice, second order can not be achieved because there is no guarantee on the symmetricity of particle distribution in real simulation \citep{price2012smoothed}.
Recall that $d\textbf{x}\prime = \dfrac{dm \prime}{\rho \prime}$, the integration equation, Eq. (\ref{eq:SPH-fundamental-principle}), can be approximated by summation and lead to an approximation of the function at a particle (idiscretization point) a:

\begin{equation}
<A(\textbf{x})> \approx \sum_b m_b \dfrac{A_b}{\rho_b} w(\textbf{x}-\textbf{x}_b, h)
\label{eq:SPH-approximation-sum}
\end{equation}

Where the summation is over all the particles b within the region of compact support of the weighting function. 
Gradient terms may be straightforwardly calculated by taking the derivative of Eq. (\ref{eq:SPH-approximation-sum}), giving

\begin{equation}
\begin{split}
<\nabla A(\textbf{x})> & = \dfrac{\partial }{\partial \textbf{x}} \int_{\Omega} A(\textbf{x} \prime) w(\textbf{x}-\textbf{x}\prime, h) d\textbf{x}\prime + O(h^2) \\
& \approx \sum_b m_b \dfrac{A_b}{\rho_b} \nabla w(\textbf{x} - \textbf{x}_b, h)
\end{split} 
\label{eq:SPH-scalar-function-gradient}
\end{equation}

For vector quantities the expressions are similar, simply replacing $A$ with $\textbf{A}$ in Eq. (\ref{eq:SPH-approximation-sum}), giving
\begin{eqnarray}
<\textbf{A}(\textbf{x})> \approx \sum_b m_b \dfrac{\textbf{A}_b}{\rho_b} w(\textbf{x}-\textbf{x}_b, h) \\
<\nabla \cdot \textbf{A}(\textbf{x})> \approx \sum_b m_b \dfrac{\textbf{A}_b}{\rho_b} \cdot \nabla w(\textbf{x} - \textbf{x}_b, h) \\
<\nabla \times \textbf{A}(\textbf{x})> \approx \sum_b m_b \dfrac{\textbf{A}_b}{\rho_b} \times \nabla w(\textbf{x} - \textbf{x}_b, h) \\
<\nabla^j \textbf{A}^i(\textbf{x})> \approx \sum_b m_b \dfrac{\textbf{A}_b^i}{\rho_b} \nabla^j w(\textbf{x} - \textbf{x}_b, h) 
\label{eq:SPH-vecctor-function}
\end{eqnarray}
%$just Italic for math symbol A$ \\
%$\textbf{bold Italic for vector A}$\\
%$\textbf{\text{bold norm for matrix A}}$ \\
%\textit{ Italic \textbf{bold Italic, A for vector}}
%\textbf{For matrix, Just bold A}
\subsection{Weighting function}
As described in previous section, the weighting function (or kernel function) is used to replace Dirac function and should satisfy Eq.  (\ref{eq:SPH_kernel_delta}). So it can be viewed as approximation form of Dirac function and satisfy normalization condition:
\begin{equation}
\int	 w(\textbf{x}-\textbf{x}\prime, h) d\textbf{x}\prime = 1
\label{eq:SPH-kernel-normalization-prop}
\end{equation}
Besides normalization, weighting function of particle $a$ has to be symmetric with respect to $a$ to ensure that neighbor particles locate at the same distance away will contribute equally to SPH summation equation, see Eq. (\ref{eq:SPH-kernel-symmetric}). Weighting function also need to satisfy conditions such as positivity and compact support. In addition, kernel function must be monotonically decreasing with the distance between particles.\\
\begin{equation}
w(\textbf{x}- \textbf{x} \prime, h) = w(\textbf{x} \prime - \textbf{x}, h)
\label{eq:SPH-kernel-symmetric}
\end{equation}
There is a wide variety of possible weighting functions, such as spline functions (with different orders) and Gaussian function. Generally, the accuracy increases with the order of the polynomials of the kernel function, but the computational time also increases as number of interactions increase. 
We are adopting a truncated Gaussian function as the weighting function in our simulation.
\begin{equation}
w(\textbf{x} - \textbf{x}_b) = 
\begin{cases} 
      \dfrac{1}{(h \sqrt{\pi})^d} exp [- (\dfrac{\textbf{x} - \textbf{x} \prime}{h})^2 ] &  \vert \textbf{x} - \textbf{x} \prime \vert \leq 3h\\
      0 & \text{Otherwise}
\end{cases}
\label{eq:SPH-kernel}
\end{equation}
Where $d$ is number of dimensions.
The derivative of weighting function:
\begin{equation}
\nabla w(\textbf{x} - \textbf{x}_b) = 
\begin{cases} 
      -2(\dfrac{\textbf{x} - \textbf{x} \prime}{h^2}) \dfrac{1}{(h \sqrt{\pi})^d} exp [- (\dfrac{\textbf{x} - \textbf{x} \prime}{h})^2 ] &  \vert \textbf{x} - \textbf{x} \prime \vert \leq 3h\\
      0 & \text{Otherwise}
\end{cases}
\end{equation}
\subsection{Discretization of governing equations}
The basic interpolants given in Eq. (\ref{eq:SPH-approximation-sum}) to Eq. (\ref{eq:SPH-vecctor-function}) provide a general way of obtaining SPH expressions of governing equations. The problem is that using these expressions ``{as is}" in general leads to quite poor gradient estimates. Various tricks are used to conserve linear and angular momentum and thermal energy \citep{monaghan1992smoothed}. Special treatments are also need for second order derivative terms. We only refer here one of these possible discretizations of compressible Euler equations with SPH:
\begin{eqnarray}
<\rho_a> = \sum m_b w_{ab} (h_a) \label{eq:ns-sph-d} \\
<\dfrac{d \textbf{v}_a}{d t}>= \sum_b m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}) \nabla_a w_{a b}(h) +\textbf{g} \label{eq:ns-sph-v} \\
<\dfrac{d e_a}{d t}>=
 0.5\sum_b m_b \textbf{v}_{a b}(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}) \cdot \nabla_a w_{a b}(h) \label{eq:ns-sph-e}
\end{eqnarray}
Where, $\textbf{v}_{a b} = \textbf{v}_a - \textbf{v}_b$. $\Pi$ is artificial viscosity term, will be discussed in section \ref{sec:artificial-viscosity}.
However, these discretized formulations will not guarantee conservation of entropy. We can obtain conservation of entropy by giving up conservation of mass or thermal energy adopting an alternative discretized formulations \citep{monaghan1992smoothed}.

As a Lagrangian method, particle position will also been update at every time step.
\begin{equation}
<\dfrac{d \textbf{x}_a}{dt}> = \textbf{v}_a \label{eq:SPH-update-pos}
\end{equation}
Adding new physics and new phases into the model is trivial in terms of discretization. For example, add of new source (or sink) into Eq.  (\ref{eq:ns-sph-d}), add a drag force into Eq. (\ref{eq:ns-sph-v})  and a heat exchange term into Eq. (\ref{eq:ns-sph-e}) will lead to the new discretization form:
\begin{eqnarray}
<\rho_a> = \sum m_b w_{ab} (h_a) + \dot{\rho}(\textbf{x},t)\label{eq:ns-source-sph-d} \\
<\dfrac{d \textbf{v}_a}{d t}>= \sum_b m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}) \nabla_a w_{a b}(h) +\textbf{g} + D \sum	_b m_b \dfrac{\textbf{v}_b - \textbf{v}_a}{\rho_b} \label{eq:ns-drag-sph-v} \\
<\dfrac{d e_a}{d t}>=
 0.5\sum_b m_b \textbf{v}_{a b}(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}) \cdot \nabla_a w_{a b}(h) + \sum_b \dfrac{m_b}{\rho_b}(\kappa_a + \kappa_b) \dfrac{(T_a - T_b)}{\textbf{r}_a - \textbf{r}_b} w_{ab}(h) \label{eq:ns-conduction-sph-e}
\end{eqnarray}
Where the source term $\dot{\rho}$ can be a "sink" of erupted vapor due to phase change which is considered in ATHAM \citep{oberhuber1998volcanic}. $D$ is drag force coefficient. $\kappa$ is heat conduction coefficient. Other physics can be added easily in a similar way. Adding of these new terms will lead to modification at several lines in the source code. The drag force term should show up only when dynamics dis-equilibrium between different phases is considered. In that case, each phase will need one set of governing equations of Navier-Strokes type. Adding of new phases into SPH code will only need add several new lines for the new phase besides adding interaction terms.

\subsection{Artificial viscosity} \label{sec:artificial-viscosity}
In classic SPH, shock waves are handled by introducing artificial viscosity to smear out discontinuities. $\Pi$, in discretized momentum equation, Eq. (\ref{eq:ns-sph-v}), and energy equation, Eq. (\ref{eq:ns-sph-e}), represents artificial viscosity term which is essentially a second order derivative term. As in the case of first derivatives, second derivatives can be estimated by differentiating a SPH interpolant twice. However, such formulism has two disadvantages: First, it is very sensitive to particle disorder. Second, the second derivative of the kernel can change sign and will lead to unphysical representation (viscous dissipation cause decrease of the entropy). 
One of the most commonly used form of artificial viscosity is:
\begin{equation}
\Pi_{ab}=- \frac{\nu}{\bar{\rho}_{ab}} \dfrac{ \textbf{v}_{ab} \cdot \textbf{x}_{ab}}{\textbf{x}_{ab}^2 + (\eta h)^2}
\label{eq:art-vis-original}
\end{equation}
Absolute(kinematic) viscous coefficient $\nu$ is defined as:
\begin{equation}
\nu = \alpha \bar{h}_{ab} \bar{c}_{ab}
\end{equation}
Where 
\begin{eqnarray}
\bar{c}_{ab} = \dfrac{c_a + c_b}{2} \\
\bar{\rho}_{ab} = \dfrac{\rho_a + \rho_b}{2} \\
\textbf{v}_{ab}=\textbf{v}_a-\textbf{v}_b \\
\textbf{x}_{ab}=\textbf{x}_a-\textbf{x}_b
\end{eqnarray}
The artificial viscosity term $\Pi_{ab}$ is a Galilean invariant and vanishes for rigid rotation. It will produce a repulsive force between two particles when they are approaching each other and an attractive force when they are receding from each other. \\
The SPH viscosity can be related to a continuum viscosity by converting the summation to integrals \citep{monaghan2005smoothed}. It has been shown that the shear viscosity coefficient $\eta = \frac{\rho \alpha h c}{8} $ and the bulk viscosity coefficient $ \zeta = \frac{5 \eta}{3}$ for two dimensional and $\eta = \frac{\rho \alpha h c}{10} $ , $ \zeta = \frac{5 \eta}{3}$ for three dimensional.
An extra term was added to $\nu$ considering aspects of the dissipative term in shock solutions based on Riemann solvers and lead to:
\begin{equation}
\Pi_{ab} = 
\begin{cases} 
      \dfrac{- \alpha \mu_{ab} \bar{c}_{ab} + \beta \mu_{ab}^2} {\bar{\rho}_{ab}} & \textbf{v}_{ab} \cdot \textbf{x}_{ab} < 0\\
      0 & \textbf{v}_{ab} \cdot \textbf{x}_{ab} > 0
\end{cases}
\label{eq:art-vis-shock}
\end{equation}
Where
\begin{equation}
\mu_{ab} = \dfrac{h \textbf{v}_{ab} \cdot \textbf{x}_{ab}}{\textbf{x}_{ab}^2 + (\eta h)^2} 
\end{equation}
$\alpha$ and $\beta$ are free parameters that must be adjusted for each case. $\alpha = 1$ and $\beta = 2$ are  recommended by Monaghan for best results. In our simulation, these two parameters are calibrated to  $\alpha = 0.3$ and $\beta = 0.6$. $\eta$ is usually taken as 0.1 for preventing singularities when $\textbf{x}_{ab} = 0$.
\subsection{Timestep}
The physical properties (velocity, density, position and pressure) change every time step due to particle interactions. We have to consider the Courant condition, which is in spirit similar to the Courant condition for the grid-based methods, to determine the amount of the time step $\Delta t$.
\begin{equation}
\Delta t = CFL \min_a \bigg \lbrace \dfrac{[\frac{m_a}{\rho_a}]^{\frac{1}{d}}}{c_a} \bigg \rbrace
\end{equation}
Where $c_a$ is sound speed at particle a, $d$ is number of dimensions. First order Euler time integration, with $CFL = 0.2$ has been used in three dimensional calculations.
\subsection{Tensile instability and corrected derivatives}
The classical SPH method suffered from tensile instability and boundary deficiency. To address these difficulties, Chen \citep{chen1999improvement} proposed a corrective SPH method. For 1D case, by employing a Taylor expansion for $A(x)$ about $x_a$, multiplying both side by kernel function and then do intergration over the domain, gives
\begin{equation}
\int_{\Omega} A(x) w(x- x_a, h) dx = 
A_a \int_{\Omega} w(x - x_a, h) dx +A_{xa} \int_{\Omega} (x-x_a) w(x - x_a, h) dx +...
\end{equation}
Ignore derivative terms higher than first order, and write the intergral in particle approximation form:
\begin{equation}
A_a = \frac{\sum_b m_b \dfrac{A_b}{\rho_b} w(x_a-x_b, h)}{\sum_b m_b \dfrac{1}{\rho_b} w(x_a-x_b, h)}
\label{eq:CSP-function-approximation-1d}
\end{equation}
Equation \ref{eq:CSP-function-approximation-1d} implies that $\sum_b m_b \dfrac{1}{\rho_b} w(x-x_b, h) = 1$, which can be viewd as the approximation form of Eq. \ref{eq:SPH-kernel-normalization-prop}. The first order derivative term can be obtained in a similar way and lead to:
\begin{equation}
\nabla A_a = \frac{\sum_b m_b \dfrac{A_b - A_a}{\rho_b} \nabla_a w(x_a-x_b, h)}{\sum_b m_b \dfrac{x_b - x_a}{\rho_b} \nabla_a w(x_a-x_b, h)}
\end{equation}
For higher dimesions problem, expression for function approximation are exactly the same as Eq. (\ref{eq:CSP-function-approximation-1d}) even though the derivation is different. The first order derivative can be solved explicitly or numerically from:
\begin{equation}
\text{\textbf{M}}_{\alpha \beta a} \textbf{A}_{\alpha a} = \textbf{F}_{\beta a}
\end{equation}
where
\begin{eqnarray}
\text{\textbf{M}}_{\alpha \beta a} = \sum_b m_b \dfrac{\textbf{x}_b^{\alpha} - \textbf{x}_a^{\alpha}}{\rho_b} \nabla_a^{\beta} w(\textbf{x}_a -\textbf{x}_b, h) \\
\textbf{A}_{\alpha a} = \dfrac{\partial A}{\partial x^{\alpha}} (\textbf{x}_a)\\
\textbf{F}_{\beta a} = \sum_b m_b \dfrac{A_b - A_a}{\rho_b} \nabla_a^{\beta} w(\textbf{x}_a -\textbf{x}_b, h)
\end{eqnarray}
$\alpha$ and $\beta$ are direcion indices, ranging from 1 to 3 for 3D problem. 
\subsection{multiple phase SPH}
Numerical simulation of multiphase flows is usually difficult due to the existence of interface between two fluids. And the interface is usually moving or deformable. This present severe challenges to conventional Eulerian grid-based numerical methods. As we discussed in the introduction section, special algorithms to treat and track the interface between different phases is required for mesh-based method to deal with multiple phases problems. What's worse, the engulfment process during plume development makes geometry of interface very complicated. SPH is able to handle multiple phase problem without without additional numerical effort for dealing with interface, which makes it suitable for volcano plume simulation. Take the governing equation adopted in our model as an example. Different phases of fluid (air and erupted material) will be represented by two different sets of SPH particles(or discretization points). Based on assumption we made in section \ref{sec:physics-model}, only density need to be updated respectively for each phase. The updating of density is exactly the same as Eq. (\ref{eq:ns-sph-d}) in spirit.  Particles of phase 1 will not be counted while evaluate density of phase 2 and vice verse. Updating of density is based on the following discretized equations.
\begin{equation}
<\rho_{\alpha}^a>=\frac{\sum m_b w_{\alpha b} (h_{\alpha})}{\sum \frac{m_b}{\rho_b} w_{\alpha b} (h_{\alpha}) +\sum \frac{m_j}{\rho_j} w_{\alpha j} (h_\alpha)} \label{eq:gov-sph-d1}
\end{equation}
\begin{equation}
<\rho_\alpha^{sg}>=\frac{\sum_j m_j w_{\alpha j} (h_\alpha)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} (h_{\alpha}) +\sum \frac{m_j}{\rho_j} w_{\alpha j} (h_\alpha)} \label{eq:gov-sph-d2}
\end{equation}
Where the subscript $a$ and $b$ represents air particles (phase 1) while $i$ and $j$ represents erupted material. $\beta$ and $\alpha$ represents either erupted material or air.
$\rho_a^a$ is density of phase 1 (air). 
 $\rho_i^{sg}$ is density of phase 2 (erupted material).
$\rho=\rho^a + \rho^{sg}$ is density of mixture of phase 1 and phase 2. In the areas far away from interface, updating of density is exactly the same as that for single phase flow. For example, in the right side and left side portition (or blue) of Fig. \ref{fig:SPH-multiple-density}, where there are only air particles, Eq.  (\ref{eq:gov-sph-d2}) will lead to zero and total density will be:
\begin{equation}
\rho_{\alpha}=\rho_{\alpha}^a=\frac{\sum m_b w_{\alpha b} (h_{\alpha})}{\sum \frac{m_b}{\rho_b} w_{\alpha b}}
\end{equation}
Which is a special case of Eq. (\ref{eq:CSP-function-approximation-1d}). For these areas occupied by only particles of phase 2 and far away from interface, similarly, the equation for density update will become: 
\begin{equation}
\rho_{\alpha}=\rho_{\alpha}^{sg}=\frac{\sum m_j w_{\alpha j} (h_{\alpha})}{\sum \frac{m_j}{\rho_j} w_{\alpha j}}
\end{equation}
That is to say, the same density updating equation can be applied for both phases and no additional numerical treatment needed to locate where is the interface.

\begin{figure}
\includegraphics[width=12cm]{Interface.png}
\caption{In the left figure, the blue particles (phaseID=1) represent air particles the red ones (phaseID=2) represent erupted material. The right figure shows corresponding mass fraction. Mass fraction are evaluated based on Eq. (\ref{eq:gov-sph-d1}) and Eq. (\ref{eq:gov-sph-d1}) without any interface track or capture method.}
\label{fig:SPH-multiple-density}
\end{figure}

\subsection{Turbulence modeling with SPH}
Typically, turbulence fluctuations in a volcanic column occur on very different length scales, ranging from millimeters up to hundreds of meters. The entrainment of air and heat tranfer are dominated by such turbulent fluctuations as turbulent exchange coefficient would be several magnitudes larger than molecular viscosity (and heat conduction) coefficient. One way to include enough turbulence in the model is to use SPH-SPS (SPH sub-particle-scale) method.
%\citep {gotoh2004sph, shao2006sph}. 
Another way is to use extremely fine resolution (The direct numerical simulation method) at the expense of much higher computational cost that results from both increase in number of particles and decrease in time steps constrained by CFL condition. Here we choose the SPH-SPS method to capture  momentum and energy exchange caused by smaller scale turbulence.\\
Even though free surface flow are in nature turbulent, detailed research in turbulence using the SPH method is rather scarce. Among the first papers on turbulence are the implementations of the $\alpha$ turbulence model \citep{holm1999fluctuation} by Monaghan \citep{monaghan2002sph} and the Reynolds averaged Navier?Stokes (RANS) approach by Violeau and Issa \citep{violeau2007numerical}. A first wall-bounded large eddy simulation (LES) was performed by Issa \citep{issa2005numerical}. An improved version of $SPH-\alpha$ method is proposed in 2011 by Monaghan \citep{monaghan2011turbulence}, which we refer to as $SPH-\varepsilon$ method. We will adopt $SPH-\varepsilon$ method in simulation of volcano plume. As most of these papers constrain their focus on incompressible or weakly compressible fluid. To simulate turbulence in plume, we need to extend existing incompressible turbulent modeling methods to compressible flows.\\
$SPH-\alpha$ model \citep{monaghan2002sph} is a powerful extension of the XSPH algorithm which reduces disorder at short length scales and retains the constants of the motion. But in that paper \citep{monaghan2002sph}, energy conservation equation is discretized in a way that is exactly the same as usual SPH. There is no special treatment on energy conservation equation. And the recent improved version \citep{monaghan2011turbulence} of $SPH-\alpha$ method focus on incompressible flow only. Actually, there are several additional terms in the averaged compressible Navier-Stokes equations besides the Reynolds stress term \citep{NASACompressibleTurbulence}. The averaged momentum equation for compressible flow are in the same form as that for incompressible flow, all of other additional terms except for turbulent stress term are in the energy equation). However, most turbulence modeling focuses on the Reynolds stress terms which are either solved directly or defined via a constitutive relation. Less attention is typically given to the other terms that need to be modeled. Most commonly, a Reynolds analogy is used to model the turbulent heat flux. Simulations of heat transfer or other scalar transfer in turbulent flow simply involve adding transport terms for thermal energy or species concentration, at the expense of greater storage and longer computing times but without other difficulties \citep{cebeci2013analysis}. We adpot this strategy in our simulation of turbulent heat transfer with $SPH-\varepsilon$. PDAC \citep{neri2003multiparticle} also adopted a Reynolds analogy in its turbulence modeling. The additional terms associated with molecular diffusion and turbulent transport in the energy equation are either modeled in different ways or neglected sometimes \citep{NASACompressibleTurbulence}. We neglect these terms in our simulation.\\
\subsubsection{$SPH-\varepsilon$ turbulence model}
In $SPH-\varepsilon$ method \citep{monaghan2011turbulence}, a turbulence model is constructed within the framework of SPH in such a way that general principles such as conservation of energy, momentum and circulation are satisfied using the ideas associated with the LANS (Lagrangian averaged Navier?Stokes equations). The basic idea of $SPH-\varepsilon$ is that determine a smoothed velocity $\widehat{\textbf{v}}$ by a linear operation on the unsmoothed velocity $\textbf{v}$. The SPH particles move with this smoothed velocity and the average motion of the fluid is determined by Eq. (\ref{eq:gov-update-pos-turbulence}):
\begin{equation}
\dfrac{d \textbf{x}_a}{dt} = \widehat{\textbf{v}}_a \label{eq:gov-update-pos-turbulence}
\end{equation}
Lagrange?s equations contain extra velocity dependent term which represents the stresses induced by the smoothing. Once the form of the smoothing is chosen these stresses are determined. 
%The overall effect is to produce a redistribution of energy without dissipation. 
The typical LANS model uses a smoothed velocity $\widehat{\textbf{v}}$ 
defined in terms of the unsmoothed velocity $\textbf{v}$ by:
\begin{equation}
\widehat{\textbf{v}}(\textbf{x})=\int \textbf{v}(\textbf{x} \prime)G(\vert \textbf{x} \prime - \textbf{x} \vert, l) d\textbf{x} \prime
\end{equation}
Where $G$ satisfies:
\begin{equation}
\int G(\vert \textbf{x} \prime - \textbf{x} \vert, l) d\textbf{x} \prime =1
\end{equation}
and is a member of a sequence of functions which tends to the $\delta$ function in the limit where $ l\rightarrow 0$. A typical example is Gaussian.
The length scale $l$ determines the characteristic width of the kernel and the distance over which the velocity is smoothed.\\
It is common practice in $LANS-\alpha$ to use a differential equation for the smoothing rather than the integral form and finally reach to system of equations that need to be solved implicitly. In $SPH-\varepsilon$ method, a XSPH smoothing is adopted which will conserve linear and angular momentum. This furtherly avoid solving of system of equations and make the method simple to implement and cheap for computation. 
The discretized form of momentum equation is obtained through lengthy derivation based on Lagrange?s equations. Derivation details and other discussions, such as conservation of momentum and energy, satisfactory of circulation theory, energy/momentum spectrum analysis based on Founrier transfer are detailed in Monaghan's paper \citep{monaghan2011turbulence}. Here we just have a brief summary on key steps.\\
The smooth that Monaghan adopted is:
\begin{equation}
\widehat{\textbf{v}}(\textbf{x})=\textbf{v}(\textbf{x})+ \epsilon \int (\textbf{v}(\textbf{x} \prime)-\textbf{v}(\textbf{x}))G(\vert \textbf{x} \prime - \textbf{x} \vert, l) d\textbf{x} \prime
\end{equation}
As function $G$ has the same feature as kernel function $w$, SPH approximation of the intergration will lead to:
\begin{equation} \label{eq:SPH-epsilon-filtering}
\widehat{\textbf{v}}(\textbf{x})=\textbf{v}(\textbf{x})+\epsilon \sum_b m_b \dfrac{(\textbf{v}_b -\textbf{v})}{\rho _b} G(\vert \textbf{x} _b - \textbf{x} \vert, l)
\end{equation}
By making the replacement:
\begin{equation}
\label{eq:replacement-in-turb-derive}
\dfrac{G(\vert \textbf{x} _b - \textbf{x} _a \vert, l_a)}{\rho _b} \rightarrow \dfrac{K_{ab}}{M}
\end{equation}
Where $K_{ab} = l^d G_{ab}$, $M = \rho_0 l^d$ in which $d$ is the dimension and $\rho_0$ is initial density. $SPH-\varepsilon$ turbulence model is obtained after longthy derivation:
\begin{equation}
\label{eq:monaghan-mom-turb}
\dfrac{d \textbf{v}_a}{dt} = -\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}) \bigtriangledown_aw_{a b}(h_a)] + \sum_b m_b \dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{M} \bigtriangledown_a K_{ab}
\end{equation}
Instead of using Eq. (\ref{eq:monaghan-mom-turb}) which is from Monaghan's paper, we make inverse replacement of Eq. (\ref{eq:replacement-in-turb-derive}) and notice that if $l$ is uniform and keep constant: 
%Please notice that the above equation is slightly different from the original equation in Monaghan's paper (Eq. (2.17)). The original equation is based on the assumption that density is uniform initially. This assumption is not general and atmosphere density is stratified in our plume simulation. And the slight  modification will not influence following derivation if $l$ is uniform and keep constant becasue:
\begin{equation}
\nabla K_{ab} = \nabla (l^d G_{ab}) = l^d \nabla G_{ab}
\end{equation}
So the discretized momentum equation with $SPH-\varepsilon$ turbulence model in our simulation will be:
\begin{equation}
\label{eq:SPH-mom-epsilon-turb}
\dfrac{d \textbf{v}_a}{dt} = -\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}) \bigtriangledown_aw_{a b}(h_a)] + \sum_b m_b \Phi_{ab}\bigtriangledown_aG_{ab}(l_a)
\end{equation}
where 
\begin{equation}
\Phi_{ab}=\dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\rho_b} 
\end{equation}
and take $\varepsilon = 0.8$ as recommended by JJ Monaghan.
\subsubsection{Turbulence heat transfer}
We adopt Reynolds analogy to get heat transfer coefficient due to turbulence.
Prandtl number is defined as:
\begin{equation}
Pr=\dfrac{C_p \mu}{\kappa}
\end{equation}
Where, $\mu$ is dynamic viscosity, $\kappa$ is thermal conductivity. And $\mu$  can be written in term of abslolute viscosity (kinematic viscosity) as:
\begin{align}
\mu=\rho \nu
\end{align}
Then
\begin{equation}
\kappa=\dfrac{C_p \mu}{Pr}
\end{equation}
Typical value of $Pr_t$ for air is 0.7 $\sim$ 0.9 . We take $Pr_t=0.85$ for gases as recommended by Kays, William M \citep{kays1994turbulent} from summarizing of experimental results. 
JJ Monaghan \citep{monaghan2005smoothed} summarized the simulation of viscosity and heat conduction in his review on SPH. We will refer to his summary in our following discussion. The additional term in discretized momentum equation, Eq. (\ref{eq:SPH-mom-epsilon-turb}), is the turbulent shear stress term. 
Recall that viscosity term can be discretized with SPH as shown in Eq. (\ref{eq:art-vis-original}). 
Comparing of this term to physical viscous term for compressible gas shows that it has both bulk viscosity and shear viscosity, where shear viscosity coefficient is \citep{monaghan2005smoothed}:\\
\begin{equation}
\nu_t = S \nu
\end{equation}
with
\begin{equation}
S= 
\begin{cases} 
      \dfrac{1}{6} & for  \quad d=3 \\
      \\
     \dfrac{5}{24}  & for  \quad d=2 
\end{cases}
\end{equation}
The turbulent viscosity coefficient can be inferred from that form if we can reformulate the "turbulence stress" term in a form which is similar to molecular shear viscous term.
Reformulate the "turbulence stress" term:
\begin{equation}
 \sum_b \dfrac{\varepsilon}{2} \dfrac{m_b}{\rho_b} \textbf{v}_{ab} \cdot \textbf{v}_{ab} \nabla_a G_{ab}(l_a)= \sum_b \dfrac{\varepsilon}{2S} m_b \dfrac{\textbf{v}_{ab}}{\rho_b} \dfrac{S \textbf{v}_{ab} \cdot \textbf{x}_{ab}}{x_{ab}^2} \dfrac{x_{ab}^2}{\textbf{x}_{ab}} \nabla_a G_{ab}(l_a) 
\end{equation}
Then the turbulent viscosity coefficient will be
\begin{equation}
\nu_t = \dfrac{\varepsilon}{2S} \dfrac{\textbf{v}_{ab} \cdot \textbf{x}_{ab}}{\rho_b}
\end{equation}
Please notice that turbulent viscosity term has opposite sigh with molecular viscosity term in discretized momentum equation and there is a minus sign in the expression of $\Pi_{ab}$, they will cancelled out.
However, the above equation is correct only for 1D situation. For 2D or 3D, it is not easy to get the explicit expression of viscosity. \\
We adopt an alternative way: obtaining a value for each pair of particles instead of persist on getting an anlytical expression. Choosing smoothing function as the same as SPH kernel and the smoothing scale $l$ as the same as smoothing length $h$, the ratio between turbulent shear stress and physical shear stress is: 
\begin{equation}
\begin{split}
\Upsilon_{ab} &= \dfrac{\dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\rho_b}}{\dfrac{S \nu}{\rho_{ab}} \frac{\textbf{v}_{ab} \cdot \textbf{x}_{ab}}{x_{ab}^2 + \eta^2 h_{ab}^2}} \\
 & = \dfrac{\varepsilon (x_{ab}^2 + \eta^2 h_{ab}^2)}{2 S \nu} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\textbf{v}_{ab} \cdot \textbf{x}_{ab}}
\end{split}
\end{equation}
$\Upsilon_{ab}$ is essentially equivalent to ratio between turbulent viscous effect attributed by particle b on particle a and molecular viscous effect attributed by particle b on particle a. Turbulent viscosity can be easily obtained by:
\begin{equation}
\begin{split}
\nu_{t,ab} &= \nu \Upsilon_{ab} \\
&= \dfrac{\varepsilon (x_{ab}^2 + \eta^2 h_{ab}^2)}{2 S} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\textbf{v}_{ab} \cdot \textbf{x}_{ab}}
\end{split}
\end{equation}
The corresponding turbulent thermal conductivity should be
\begin{equation}
\kappa_{t,ab}=\dfrac{\varepsilon \overline{C_{p,ab}} \overline{\rho_{ab}} (x_{ab}^2 + \eta^2 h_{ab}^2) \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 S Pr_t\textbf{v}_{ab} \cdot \textbf{x}_{ab}}
\end{equation}
The term that used to prevent singularity now can be removed. 
\begin{equation}
\kappa_{t,ab}=\dfrac{\varepsilon \overline{C_{p,ab}} \overline{\rho_{ab}} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 S Pr_t\textbf{v}_{ab} \cdot \textbf{x}_{ab} }
\end{equation}
We also need to prevent singularity, so: 
\begin{equation}
\kappa_{t,ab}= 
\begin{cases} 
      0 & if  \quad \textbf{v}_{ab}=0 \quad or \quad \textbf{x}_{ab}=0 \\
      \dfrac{\varepsilon \overline{C_{p,ab}} \overline{\rho_{ab}} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 S Pr_t\textbf{v}_{ab} \cdot \textbf{x}_{ab} } & \text{otherwise}
\end{cases}
\label{eq:SPH-LANS-heat-conductivity}
\end{equation}
The heat conduction equation without source term is:
\begin{equation}
C_p \dfrac{dT}{dt} = \dfrac{1}{\rho} \nabla (\kappa \nabla T)
\end{equation}
Second spatial derivative can be approximated with SPH by following JJ Monaghan \citep{monaghan2005smoothed}. 
\begin{equation}
C_p \dfrac{dT}{dt} = \sum_b \dfrac{m_b}{\rho_a \rho_b} (\kappa_a + \kappa_b) (T_a - T_b) F_{ab} (h_a)
\end{equation}
Where
\begin{equation}
F_{ab} \textbf{x}_{ab} = \nabla _a w_{ab}
\end{equation}
And  $F_{ab} \leq 0$, which guarantees that heat flux will flow from hot to cold.
%JJ Monaghan \citep{cleary1999conduction} showed a simple alteration to the standard SPH formulation and ensures continuity of heat flux across discontinuities in material properties. 
Plug the turbulent thermal conductivity into the heat conduction equation:
\begin{equation}
\begin{split}
C_p \dfrac{dT}{dt} 
& = \sum_b \dfrac{m_b}{\rho_a \rho_b} (\kappa_a + \kappa_b) (T_a - T_b) F_{ab} (h_a) \\
 &= 2 \sum_b \dfrac{m_b}{\rho_a \rho_b} \dfrac{\overline{C_{p,ab}} \overline{\rho_{ab}} \varepsilon x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 Pr_t  S \textbf{v}_{ab} \cdot \textbf{x}_{ab} } (T_a - T_b) F_{ab} (h_a)
\end{split}
\end{equation}
Notice that the number 2 at the front of the equation comes from intergration approximation of second order derivative \citep {cleary1999conduction}. By further simplification, we get:
\begin{equation}
C_p \dfrac{dT}{dt}
 =\dfrac{\varepsilon}{S  Pr_t}  \sum_b \dfrac{m_b}{\rho_a \rho_b} \dfrac{\overline{C_{p,ab}} \overline{\rho_{ab}} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\textbf{v}_{ab} \cdot \textbf{x}_{ab}} (T_a - T_b) F_{ab} (h_a)
\end{equation}
\subsubsection{Discretized governing Equation with $SPH-\varepsilon$ turbulence model}
Plug discretized turbulent stress term and turbulent heat transfer term into the momentum and energy equation, we got new discretized governing equation:
\begin{equation}
<\dfrac{d \textbf{v}_{\alpha}}{d t}>= 
-\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}) \bigtriangledown_{\alpha}w_{\alpha b}(h_{\alpha})]
-\sum_j [m_j (\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j} - \Phi_{\alpha j}) \bigtriangledown_{\alpha}w_{\alpha j}(h_{\alpha})]
+\textbf{g} \label{eq:gov-sph-v}
\end{equation}
With 
\begin{equation}
\Phi_{\alpha \beta}=\dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{\alpha \beta} \cdot \textbf{v}_{\alpha \beta}} {\rho_{\beta}} 
\end{equation}

\begin{equation}
\begin{split}
<\dfrac{d e_{\alpha}}{d t}>
& = 0.5\sum_b [m_b \widehat{\textbf{v}_{\alpha b}} (\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}) \bigtriangledown_{\alpha}w_{\alpha b}(h_{\alpha}] 
 + 2 \sum_b \dfrac{m_b}{\rho_{\alpha} \rho_b} \kappa_{t,\alpha b} (T_{\alpha} - T_b) F_{\alpha b} (h_{\alpha}) \\
 & +0.5\sum_j [m_j \widehat{\textbf{v}_{\alpha b}}(\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j} - \Phi_{\alpha j}) \bigtriangledown_{\alpha}w_{\alpha j}(h_{\alpha})]
 + 2 \sum_j \dfrac{m_j}{\rho_{\alpha} \rho_j} \kappa_{t,\alpha j} (T_{\alpha} - T_j) F_{\alpha j} (h_{\alpha})
\end{split}
\label{eq:gov-sph-e}
\end{equation}
with $\kappa_{t,\alpha \beta}$ given by Eq. (\ref{eq:SPH-LANS-heat-conductivity}). 
As the particle-scale movement of flow is based on smoothed velocity, the velocity in the energy equation should also be smoothed.
The filtering process will be done accroding to Eq. (\ref{eq:SPH-epsilon-filtering}). Position of particles will be updated according to Eq. (\ref{eq:gov-update-pos-turbulence}). Smoothed velocity is also used while computing artificial viscosity $\Pi_{ab}$.\\
\subsection{Boundary conditions in SPH}
\subsubsection{Wall boundary condtion}
Traditionally either ghost particles that mirror SPH particles across the boundary \citep {ferrari2009new}, or boundary forces \citep {monaghan2009sph} have been used to impose the wall boundary conditions. One disadvantage of the later one is that the boundary forces tend to corrupt the solution in the neighborhood of such boundary forces. In addition, the natural way of imposing eruption boundary condition is using eruption ghost particles. To impose boundary conditions in a consistent way we adopt a modified version of ghost particle method \citep {kumar2013parallel} for wall boundary conditions. Stationary ghosts particles are deployed the same way as initial deployment of real particles. Instead of enforcing the symmetry particle by particle, symmetric field across the boundary is explicitly enforced. Ghost positions are reflected into the domain and physical properties are calculated at these reflected positions by SPH interpolations. It should be noted that ghost particles are not used for these interpolations. Assign all properties at the corresponding reflect position to that ghost particle except for velocity. We assign a velocity to each wall ghost particles that is the negative of the interpolated velocity of its corresponding reflection inside the fluid domain. By this way, the no-slip wall boundary condition (Eq. (\ref{eq:wall_bc_v})) is imposed naturally. These wall ghost particles will serve as neighbors in momentum and energy update. Implementation details about this method can be found in \citep {kumar2013parallel}. As these wall ghost particles keep stationary, there will be no mass flux on the boundary (Eq. (\ref{eq:wall_bc_rho}) and (\ref{eq:wall_bc_xi})). In addition, as temperature is also symmetric with respect to the boundary, the gradient of temperature will vanish on the wall boundary and hence no internal energy flux on the wall boundary (Eq. (\ref{eq:wall_bc_e})). 
\subsubsection{Eruption boundary condition}
The natural way of imposing eruption boundary condition is using ghost particles which moves with eruption velocity and bears the temperature of erupt material. A parabolic velocity profile which represents a fully developed Hagen-Poiseuille flow is used to determine the particle velocity. The detailed shape of the parabolic profile will determined based on averaged eruption velocity (Eq. \ref{eq:erupt_bc_v}). The correctness of eruption boundary condtion for density (Eq. (\ref{eq:erupt_bc_rho})) can be guaranteed by choosing mass of erupted ghost particles given distribution of these particles. The internal energy associated with these particles are set to a value so that Eq. (\ref{eq:erupt_bc_e}) is satisfied. The mass fraction of erupted material (Eq. (\ref{eq:erupt_bc_xi})) is automatically satisfied as all particles in the eruption conduit are of phase 2. The density, momentum, and internal energy of these eruption ghost particles will not be updated before they move above ground. As soon as they enter the atmosphere, these ghost particle will be shift to real particles and their physics properties and position will be updated based on discretized governing equations. New ghost particle will be added at the bottom of the eruption conduit as these exising ghost particles move upwards.\\
\subsubsection{Pressure boundary condition}
An other boundary in this model is pressure outlet boundary. For flow in a straight channel, it is possible to treat the exit the same as the entry with a prescribed velocity profile. For flow with more complex channel, an exit far downstream of the flow disturbance is also feasible. However, the natural boundary (Eq.  (\ref{eq:pressure_bc_p})) condition is more suitable for plume simulation where the outlet is open atmosphere. The way we impose pressure boundary conditon is simple: adding several layers of pressure ghost particles sourounding the real atmosphere particles. Physics properties except for velocity (pressure, density, temperature) are determined based on the elevation of particles. Velocity is set to zero for static atmosphere. The physics properties for pressure ghost particles will not be updated while these for real particles will be updated at every time step. As position of all pressure ghost particles are keep constant, we are essentially imposing a static pressure boundary condition. Real particles that moves out the the pressure boundary will be removed.\\
As simulation progress, change in position and physical properties of real particles near pressure boundary boundaries might corrupt pressure boundary condition that was established initially. This shortcoming is relieved by choosing a larger computational domain so that these might-be-corrupted boundary will be far away from turbulent mixing area. In addition, to avoid enlarging fluctuation, we add another constrain on time step: 
\begin{equation}
\Delta t \leq CFL_p \dfrac{h}{v}
\end{equation}
Where $CFL_p$ is a safty coefficient which has similar functionalality as the normal $CFL$ number. Too small $CFL_p$ will slow down simulation while too large $CFL_p$ will lose its ability of mitigate numerical fluctuation near the boundary. The proper $CFL_p$ is determined by series of simulation tests. 
\subsection{Parallellism and performance}
One disadvantage of 3D model is that it usually takes much longer time than 1D models to complete one simulation. This disadvantage further preventing doing simulation with finer resolution and accounting for more physics in one model. Non-intrusive uncertainty analysis which is commonly adopted in hazard forcasting requires finishing multiple simulations winthin a given time window. High performance computing provides a powerful routine for relieving the misery. Among existing CPU parallel SPH schemes, most of them focus on neighbors searching algorithm and dynamic load balancing (eg. \citep {ferrari2009new, crespo2015dualsphysics}). Less attention has been paid to developing of more flexible data management schemes for more complicated problems. Motivated by techniques developed for mesh based methods, we develop a complete framwork for parallelizing SPH program with MPI standard model allowing flexible and efficient data access in this paper.\\
Any implementation of SPH code requires efficient searching and updating of neighbors during simulation. Of the many choices possible for this we adopted a background grid which was proposed by Monaghan and Lattanzio \citep {monaghan1985refined} and is quite popular in parallel SPH. The background grid is also used for domain decomposition in SPH. We refer to the elements of background grid, namely squares for two dimension and cubics for three dimension, as buckets. 
As for the actual storage of data representing the physical quantities associated to each particle, different strategies have been adopted in existing implementations of SPH.
In both SPHysics and DualSPHysics \citep {crespo2015dualsphysics}, the physical quantities of each particle (position, velocity, density...) are stored in arrays, and the particles (and the arrays with particle data) are reordered following the order of the cells. This has two advantages: 1) access pattern is more regular and more efficient, 2) it is ease to identify the particles that belong to a cell by using a range since the first particle of each cell is known. But adding, deleting and especially accessing of particles are cumbersome. Ferrari \citep {ferrari2009new} adopted linked lists using pointers so that particles can be deleted or added during the simulation. Storage problems caused by fix-size arrays are thereby also eliminated. We define C++ classes which contain all data of particles and buckets. As for the management of data, we adopt hash tables to store pointers to particles and buckets, which give us not only flexibility of deleting and adding elements, but also quicker access compared with linked list. Instead of using the "nature manner" to number particles, we adopt SFC based index to give each particle and background bucket an unique identifier -- a strategy known to preserve some locality at minimal cost. The SFC based numbering strategy is further extended to include time step information so that particles added at the same position but different time will have different identifiers. 
As for domain decomposition, even though more complicated graph-based partition tools \citep {biswas1999experiments} might get higher quality decomposition, they requires much more effort in programming and computation. So we adopt an easy-programming scheme based on SFC \citep {patra1999efficient}.\\
We will provide a supplementary documentation describing details about the data strucutre, domain decomposition, load balancing and domain adjusting.
\begin{figure}[!t]
\centering
\includegraphics[scale=0.35]{strong}
\caption{Strong scalability test results}
\label{fig:strong_scale}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[scale=0.35]{weak_scale}
\caption{Weak scalability test results}
\label{fig:weak_scale}
\end{figure}

\begin{figure}[!t]
\centering
\includegraphics[scale=0.35]{adj_vs_no}
\caption{The effect of domain adjusting on simulation time. The figure on the top shows simultion time without domain adjusting, the figure on the bottom shows simultion time with domain adjusting. Different bins represent simulation time up to specific physics time indicated by $x$ axis.}
\label{fig:adj_vs_no}
\end{figure}

Experiments have been carried out on CCR (the CCRcomputational cluster of Center for Computational Research) at Buffalo. E5645 CPUs running at 2.40GHz clock rate with 4GB memory per core on a Q-Logic Infiniband is used in these tests. Each node comprise of two sockets with 6 of these cores. Memory and level 3 cache are shared on each node. The initial domain is $[-4.8km,4.8km]\times [-4.8km,4.8km] \times [0km, 6km]$. Almost linear speed up is observed in our strong scalability test (Fig. \ref{fig:strong_scale}).
The weak scalability test is conducted with the same initial domain and various smoothing lengths. Each simulation is run for 400 steps. The average number of real particles of each process is kept constant at $25900$. As shown in Fig. \ref{fig:weak_scale}, simulation times increase around $1/3$ when number of cores increase from 16 to 4208. For the test problem in this section, the volcanic plume will finally reach to a region of $[-30km \,\,\, 30km] \times [-30km\,\,\,30km] \times [1.5km\,\,\,40km]$ after around 400 seconds of eruption. When numerical simulation goes up to 90 seconds, the plume is still within a region of $[-10km\,\,\,10km] \times [-10km\,\,\,10km] \times [0km\,\,\,25km]$. This implies that adjusting of domain can avoid computing large number of uninfluenced air particles, especially for the beginning stage of simulation. A domain adjusting algorithm is desinged and implemented in our code. Figure \ref{fig:adj_vs_no} shows that simulation time of the test problem is greatly reduced when we adopt the domain adjusting strategy in our code.

\section{Verification and validation} \label{sec:verification-validation}
We will now present a series of numerical simulations that were used to verify and validate the model. The SPH plume model is first verified by a JPUE simulation. Velocity and mass fraction distribution both along the central axis and cross transverse are compared with experimental results. The parten of  ambient particles entrainment is also clearly shown. Then a simulation of representative strong volcano plume is conducted. Both global variables and local variable are comparable with simulation results from exisitng 3D plume models. Analysis on evolution of plume shows detailed air entrainment mechanism. 
\subsection{Simulation of JPUE}
JPUE can be considered as a simplified volcanic plume. While the effect of stratified atmosphere and the effect of expansion due to high temperature in volcanic plume are not represented, JPUE reproduces the entrainment due to turbulent mixing which is one of the key elements in volcanic plume development. There exist consistently good experimental data \citep { list1982turbulent,dimotakis1983structure, papanicolaou1988investigations} that describe the JPUE flow field giving an insight into detailes of JPUE, such as transverse velocity and concetration profile. In this section, we will verify that our code and the $SPH-\varepsilon$ turbulence model is able to reproduce feature of turbulent entrainment by a JPUE simulation.\\
As many of these experiments were conducted with liquid, we replace the original equation of state (Eq. (\ref{eq:EOS})) with a weakly compressible Tait equation of state \citep {becker2007weakly} (see Eq. (\ref{eq:EOS-Tait})) to avoid solving the Poisson equation.
\begin{equation}
p=B[(\dfrac{\rho}{\rho_0})^{\gamma}-1]
\label{eq:EOS-Tait}
\end{equation}
with $\gamma=7$ and $B$ is evaluated by:
\begin{equation}
B=\dfrac{\rho_0 c^2}{\gamma}
\end{equation}
Where $c$ is speed of sound in the liquid. Energy equation is actually decoupled from momentum and mass conservation equation by using this EOS. In addition, the "atmosphere" is assumed to be uniform and gravity is set to be zero. We set the temperature and density of ejected material from the vent the same as its surrounding ambient. This will further simplify the scenario for the convenience of studying turbulent mixing. One overall feature of JPUE is "self-similarity" which means that the evolution of the JPUE is determined soley by the local scale of length and velocity which theoretically account for the fact that the rate of entrainment at the edge of JPUE is proportional to a characteristic velociy at each height. As a results, physical and numerical experiments do not necessary to have exactly the same setups and are compared on a non-dimensional basis.\\

\begin{figure}
\includegraphics[width=8cm]{vel_cross}
\caption{Dimensionless velocity distribution across the cross-section.}
\label{fig:JPUE_cross-section_vel}
\end{figure}
\begin{figure}
\includegraphics[width=8cm]{conc_cross}
\caption{Dimensionless concetration distribution across the cross-section.}
\label{fig:JPUE_cross-section_conc}
\end{figure}

\begin{figure}
\includegraphics[width=8cm]{velo_along_axis}
\caption{Dimensionless velocity distribution along centraline.}
\label{fig:JPUE_along-axis_vel}
\end{figure}

\begin{figure}
\includegraphics[width=8cm]{conc_along_axis}
\caption{Dimensionless concetration distribution along centraline.}
\label{fig:JPUE_along-axis_conc}
\end{figure}

A three dimensional axisymmetric JPUE which ejects from a round vent is simulated and results are compared with experiments \citep {george1977turbulence, papanicolaou1988investigations} for verification purpose. Experimental data of concentration and velocity distribution across the cross-section was fit into a Gaussian profile: $\dfrac{\varphi}{\varphi_c}=exp[-coef (\dfrac{r}{z})^2)]$ (solid line) even though there is no priori reason. Panos \citep{papanicolaou1988investigations} also fit concentration distribution and jet width based on velocity along central line into a straight line : $\dfrac{\varphi_0}{\varphi_c}=slope (\dfrac{z}{D} + intercept)$ (dash line). Where $\varphi$ is either velocity or concetration, the subscript $c$ represents the centraline, subscript $0$ represents the cross-sectionally averaged exit value. $r$ is the distance from the centraline on any cross-section. $z$ is the axial distance from the origin of the jet transverse section underconsideration. $D$ is the diameter of vent. 
The coefficient $coef$ for concetration is 80 and 50 respectively according to Panos and George. $coef$ for velocity is 90 and 55 respectively according to Panos and George. 
$slope$ for jet width based on velocity is 0.104 and for concetration is 0.157. $intercept$ for jet width based on velocity is 2.58 while that for the concetration is 4.35.\\
Although both velocity and concetration are found to be well matched with experimental results, a small disparity in both velocity and concetration are observed near the boundary of the jet. There are several factors that would attribute to such disparity. Reynolds number are not reported in many experiments assuming a high enough Reynolds number. In addition, some details of the of the experiments, such as exit velocity profile, viscosity of the experimental liquid, are not reported, neither. These factors prevent us from numerically reproducing these experiments in an exact way as they were conducted. However, the features of JPUE is correctly reproduced with our code.\\
We also investigated the mixing due to turbulence in JPUE simulation by checking the mixture of two phases. It is shown in Fig. \ref{fig:Turb_mixing} that the ejected material and ambient fluids are mixed through eddies at the outer shear region. And the inner dense core dispersed gradually due to erosion of the outer shear region. Hence, our confidence in the numerical correctness of our code  is greatly reinforced.
%What need to be noticed is that the mass of particle of phase 1 (blue) is 8 time of that of phase 2 (red). It is a trade off between resolution and computational expense to give particle of two phases different particles mass. Both advanced SPH techniques like adaptive particles size and better data management and access strategies need to be exploited to address this issue.

\begin{figure}
\includegraphics[width=8cm]{JPUE_entrainment.png}
\caption{Left figure shows shows particle distribution. The right figure shows the mass fraction of erupted material at the moment corresponding to left figure. Particles of phase 1 (blue) are gradually entrained and mixed with erupted particles (red) as jet flows down stream. What need to be noticed is that the mass of particle of phase 1 (blue) is 8 time of that of phase 2 (red). It is a trade off between resolution and computational expense to give particle of two phases different particles mass.}
\label{fig:Turb_mixing}
\end{figure}

\subsection{Simulation on Plume}
The development of volcano plume is more complicated than JPUE in several aspects. Besides turbulent entrainment of ambient fluids, development of volcano plume also involves heating up of entrained air and expanding in a stratified atmosphere. A strong eruption column without wind is tested in this section for the purpose of further verification and validation. Input parameters, material properties and atmosphere are choosen to be the same as the strong plume no wind case in an inter-comparison study on eruptive column models \citep{costa2016results}. The strong plume scenario was based on climactic phase of the Pinatubo eruption (Philippines, 15 June 1991). Material properties are selected based on properties of Pinatubo and Shinmoe-dake eruption.Both global variable and local variables are compared with existing models. Then then plume evolution process are investigated.

\subsubsection{Global and local variables}
One of the key global quantities of great interest is the altitude to which the plume rises. The top height predicted by our model is around 40 km which agrees well with other plume models. For example, the height predicted by PDAC is 42500 m, by SK-3D is 39920 m, by ATHAM is 33392 m, by AHSEE is 36700 m. As for local variables, the profiles of integrated temperature, density, mass fraction of entrained air, gas mass fraction, mass fraction of solid and the radius of the plume as a function of height are compared with existing 1D and 3D models in Fig. \ref{fig:strong_local_temp} ~ \ref{fig:strong_local_radius}. To get rid of significant fluctuations in time and space we did time averaging and spatial integration of the dynamic 3D flow fields by following Cerminara \citep {cerminara2016large}.
As particles distribute in the space in a disordered manner. We first project simulation results (on disordered particles) onto a pre-defined grid before doing time average and spatial intergration. The interpolation method is the basic SPH kernel based interpolation.
\\
\begin{figure}
\center
\includegraphics[width=5cm]{Temp}
\caption{Temperature as a function of height}
\label{fig:strong_local_temp}
\end{figure}
\begin{figure}
\center
\includegraphics[width=5cm]{density_strong}
\caption{Density of the strong plume without wind after reaching its top height}
\label{fig:strong_local_density}
\end{figure}

%\begin{figure}
%\center
%\includegraphics[width=5cm]{velocity_strong}
%\caption{Velocity of the strong plume without wind after reaching its top height}
%\label{fig:strong_local_velocity}
%\end{figure}

The profiles of local variables matche well with simulation results of exisitng 3D models in general sense. The basic phenomena in volcano plume development is correctly captured by our model.
%: the plume first rises up to its top height and then falls down to the netural buoyancy height and starts expanding.\\
%
\begin{figure}
\includegraphics[width=15cm]{msfrac}
\caption{The mass fraction of entrained air, gas, and solid as a function of height.}
\label{fig:strong_plume_mass_fraction}
\end{figure}
%
As the height increase, the amount of entrained air will increase. Around the netural height, where the umbrella expanding, the entrainment of air shows a slight decrease due to lack of air surrounding the column at that height. The profile for gas, which account for both air and vapor, shows a very similar tendency as that of entrained air. First of all, vapor condese is not considered in our model. In addition, as we assume that erupted material behaves like a single phase water, the mass fraction of gas will simply be a function of entrained air (Eq. (\ref{eq:gas-frac-express})). Among these 3D models, ATHAM takes vapor condense into account and Eq. (\ref{eq:gas-frac-express}) does not hold for ATHAM. However, the profile of entrained air and profile of gas predicted by ATHAM are still very close to each other which implies that ignoring of water phase change is a valid assumption for eruptions similar to this test case (strong plume with erupted water fraction in erupted material less than 5\%). Because air occupies a larger portition of the gas and plays a more significant role. As for mass fraction of solid, similarly, we have Eq. (\ref{eq:solid-frac-express1}) or Eq. (\ref{eq:solid-frac-express2}). PDAC which treat partices of two different size range as two separate phases got similar mass fraction profile. Which implies that our model is at least valid for eruptions similar to the test case. 
\begin{equation}
n_a + n_g = n_a + (1-n_a) n_{g0}
\label{eq:gas-frac-express}
\end{equation}
\begin{equation}
n_s = (1 - n_a) + (1- n_{g0})
\label{eq:solid-frac-express1}
\end{equation}
\begin{equation}
n_s = 1 - (n_a + n_g)
\label{eq:solid-frac-express1}
\end{equation}
%
\begin{figure}
\center
\includegraphics[width=5cm]{radius_strong}
\caption{Radius of the strong plume without wind after reaching its top height}
\label{fig:strong_local_radius}
\end{figure}
%
With more cool air entrained into the plume and mixing with hot erupted material, the temperature of the plume decreases as the height increase as shown in Fig. \ref{fig:strong_local_temp}. In the meanwhile, bulk density decrease due to entrainment and expansion (Fig. \ref{fig:strong_local_density}).\\
Our model adopts the same assumptions and governing equations as SK-3D. However, there is still a big disparity between the profiles of local variables of our model and 3K-3D. One of the big difference between our model and 3K-3D is that we adopt a LANS type of turbulence modle while 3K-3D adopts a LES (large eddy simulation) turbulence model. This implies that turbulence model plays a critical role in plume simulation.
\subsubsection{Evolution of plume} 
\begin{figure}
\center
\includegraphics[width=5cm]{t15}
\caption{Contour of mass fraction and velocity quiver at 15 seconds, column rises up quickly due to intial momentum.}
\label{fig:strong_t15}
\end{figure}

\begin{figure}
\includegraphics[width=10cm]{t50}
\caption{Contour of mass fraction and velocity quiver at 50 seconds, column keeps rising up and at the same time entraining air into the plume at the outer shear region. The entrainment of air is shown in a zoomed view in which the velocity quiver clearly shows entrainment of air into the plume at the margin of plume.}
\label{fig:strong_plume_t50}
\end{figure}

\begin{figure}
\center
{
\includegraphics[width=10cm]{t100}
\caption{Contour of mass fraction and velocity quiver at 100 seconds, it keeps rising up and entraining air. The entrainment of air is clearly shown in zoomed view}
\label{fig:strong_plume_t100}
}
\end{figure}

\begin{figure}
\center
\includegraphics[width=5cm]{t200}
\caption{Contour of mass fraction at 200 seconds, plume reaches its top height}
\label{fig:strong_t200}
\end{figure}

\begin{figure}
\center
\includegraphics[width=5cm]{t300}
\caption{Contour of mass fraction at 300 seconds, the cloud is spreading}
\label{fig:strong_t300}
\end{figure}

Figure \ref{fig:strong_t15}~\ref{fig:strong_t300} shows the dynamic evolution process of volcano plume. The contour level of mass fraction are 0.7, 0.5, 0.3, 0.1, $1\times 10 ^{-1}$, $1\times 10 ^{-2}$, $1\times 10 ^{-3}$, $1\times 10^{-4}$ and $1\times 10^{-5}$. The plume keeps rising up and entraining air into the plume untill it reaches its top height(around t=200 seconds). Afterwards the mixtures of erupted material and air then falls back to neutral height and starts spreading radially. Detailed velocity quiver plots show entrainning of air into the plume at the edge of column. The evlolution of the plume further verifies our model.
%
\conclusions  \label{sec:conclusion}%% \conclusions[modified heading if necessary]
A new plume model was developed based on SPH method. Advanced numerical techniques in SPH were exploited and talored for this model. High performance computing was used to mitigate confliction between accuracy (depends on comprehensiveness of the model, resolution, order of accuracy of numerical methods, shceme for time upgrading ...) and simulation time (depends on, comprehensiveness, resolution, order of accuracy of numerical methods, shceme for time upgrading ... and computational techniques). The correctness of the code and model was verified and validated by a series of test simulations.\\
Currently existing 3D models focus on certain aspect of the volcano plume (PDAC on pyroclastic flow, ATHAM on microphysics, and SK-3D on entrainment with higher accuracy and higer oder of accuracy) and hence, naturely, different assumptions were made in these models. However, these different aspects of volcano plume are not independent and actually are coupled. In addtion, there is no absolute boundary to determine which kind of hazard is dominant in certain eruption. So it is necessary to simulate all associated hazards in one model. Actually, effort has already been put on developing more comprehensive plume models. For example, a large-particle module (LPM) was added to ATHAM to track the paths of rocky particles (pyroclasts or tephra) within the plume and predicts where these particles fall \citep{kobs2009modeling}. We were also motivated by such a tendency of plume modeling to choose SPH as our numerical tool besides its ability in dealing with interface for multiple phase flows. As mentioned in the introduction section, SPH method has good extension features and adding of new physics and phases requires tiny modification on the code compared with mesh-based methods. The last but not the least, the dramatisc development of computational power makes it possible to establish a comprehensvie model. Even though right now, current computational capacity does not allow us to have too comprehensive model, the easy-extension feature of SPH makes it feasible to keep adding new physics into the model when necessary. What presented in this paper is actually our initial effort and results towards developing a first priciple based plume model with comprehensive physics, adopting proper numerical tools and power of high performance computing. More advanced numerical techniques, such as adaptive particle size, Godnuv-SPH, semi-explicit time advancing scheme and better data management strategies and algorithms are on our list to exploit. In the near future, affect of wind field will be take into account. 
%In all existing 3D plume models, numerical calculations are performed on a non-uniform (vertically and horizontally stretched) grid, with different grid resolutions and domain sizes. However, implementation of different resolution in SPH methods is not straight forward and requires more work.

\section{Code availability}
The Plume-SPH code with a user manual providing instructions for installation, running and visualization is archive in github (https://github.com/Plume-SPH/plume-sph.git). The input data for all simulations presented in this work can also be found in the repository.
\section{Data availability}
Data for running the test cases, including meterological data, properties of erupted material and air, and eruption condition data can be found in source code repository. Raw data of simulation results is archived in google drive. Access permission will be given up request. Post-processed data for generating figures is provided as supplyment along with the manuscripts. Please note that some of these pictures are generating by visualizing the raw data with visualization tools, there is no separate data for these figures.

%\appendix
%\section{}    %% Appendix A
%
%\subsection{}     %% Appendix A1, A2, etc.
%
%\appendixfigures  %% needs to be added in front of appendix figures in one-column style (\documentclass[acp, manuscript]{copernicus})
%
%\appendixtables   %% needs to be added in front of appendix tables in one-column style (\documentclass[acp, manuscript]{copernicus})

\authorcontribution{TEXT}

\competinginterests{Authors have the standard conflict with employers and listed sponsors}

%\disclaimer{TEXT}

\begin{acknowledgements}
All developers of Titan-2D, especially Dinesh Kumar who developed the GSPH version of Titan-2D, are greatly appreciated as Plume-SPH is based on their code. Advice and data (which is not shown in this paper) provided by Suzuki Yujiro gave us great help at the initial stage of model establishment and therefor are greatly appreciated. We appreciate Tomaso Esposti Ongaro for providing simulation data of PDAC (were also provided by Antonio Costa later together with other data) which helped us in doing early verification and improvement. We appreciate Antonio Costa for providing simulation results of existing 3D models, which are used in verification and validation section of this paper. We thank Matteo Cerminara for his helping on post-processing of plume simulation results. Computational results reported here were performed at the Center for Computational Research at the University at Buffalo. This project is supported by Grants No. NSF 1131074 from the National Science Foundation.
\end{acknowledgements}

%% REFERENCES

%% The reference list is compiled as follows:
\bibliographystyle{copernicus}
\bibliography{Reference.bib}

%% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,
%% authors experienced with BibTeX only have to include the following two lines:
%%
%% \bibliographystyle{copernicus}
%% \bibliography{example.bib}
%%
%% URLs and DOIs can be entered in your BibTeX file as:
%%
%% URL = {http://www.xyz.org/~jones/idx_g.htm}
%% DOI = {10.5194/xyz}

%% LITERATURE CITATIONS
%%
%% command                        & example result
%% \citet{jones90}|               & Jones et al. (1990)
%% \citep{jones90}|               & (Jones et al., 1990)
%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
%% \citeauthor{jones90}|          & Jones et al.
%% \citeyear{jones90}|            & 1990

%% FIGURES

%% When figures and tables are placed at the end of the MS (article in one-column style), please add \clearpage
%% between bibliography and first table and/or figure as well as between each table and/or figure.

%% ONE-COLUMN FIGURES

%%f
%\begin{figure}[t]
%\includegraphics[width=8.3cm]{FILE NAME}
%\caption{TEXT}
%\end{figure}
%
%%% TWO-COLUMN FIGURES
%
%%f
%\begin{figure*}[t]
%\includegraphics[width=12cm]{FILE NAME}
%\caption{TEXT}
%\end{figure*}
%
%
%%% TABLES
%%%
%%% The different columns must be seperated with a & command and should
%%% end with \\ to identify the column brake.
%
%%% ONE-COLUMN TABLE
%
%%t
%\begin{table}[t]
%\caption{TEXT}
%\begin{tabular}{column = lcr}
%\tophline
%
%\middlehline
%
%\bottomhline
%\end{tabular}
%\belowtable{} % Table Footnotes
%\end{table}
%
%%% TWO-COLUMN TABLE
%
%%t
%\begin{table*}[t]
%\caption{TEXT}
%\begin{tabular}{column = lcr}
%\tophline
%
%\middlehline
%
%\bottomhline
%\end{tabular}
%\belowtable{} % Table Footnotes
%\end{table*}
%
%
%%% MATHEMATICAL EXPRESSIONS
%
%%% All papers typeset by Copernicus Publications follow the math typesetting regulations
%%% given by the IUPAC Green Book (IUPAC: Quantities, Units and Symbols in Physical Chemistry,
%%% 2nd Edn., Blackwell Science, available at: http://old.iupac.org/publications/books/gbook/green_book_2ed.pdf, 1993).
%%%
%%% Physical quantities/variables are typeset in italic font (t for time, T for Temperature)
%%% Indices which are not defined are typeset in italic font (x, y, z, a, b, c)
%%% Items/objects which are defined are typeset in roman font (Car A, Car B)
%%% Descriptions/specifications which are defined by itself are typeset in roman font (abs, rel, ref, tot, net, ice)
%%% Abbreviations from 2 letters are typeset in roman font (RH, LAI)
%%% Vectors are identified in bold italic font using \vec{x}
%%% Matrices are identified in bold roman font
%%% Multiplication signs are typeset using the LaTeX commands \times (for vector products, grids, and exponential notations) or \cdot
%%% The character * should not be applied as mutliplication sign
%
%
%%% EQUATIONS
%
%%% Single-row equation
%
%\begin{equation}
%
%\end{equation}
%
%%% Multiline equation
%
%\begin{align}
%& 3 + 5 = 8\\
%& 3 + 5 = 8\\
%& 3 + 5 = 8
%\end{align}
%
%
%%% MATRICES
%
%\begin{matrix}
%x & y & z\\
%x & y & z\\
%x & y & z\\
%\end{matrix}
%
%
%%% ALGORITHM
%
%\begin{algorithm}
%\caption{}
%\label{a1}
%\begin{algorithmic}
%
%\end{algorithmic}
%\end{algorithm}
%
%
%%% CHEMICAL FORMULAS AND REACTIONS
%
%%% For formulas embedded in the text, please use \chem{}
%
%%% The reaction environment creates labels including the letter R, i.e. (R1), (R2), etc.
%
%\begin{reaction}
%%% \rightarrow should be used for normal (one-way) chemical reactions
%%% \rightleftharpoons should be used for equilibria
%%% \leftrightarrow should be used for resonance structures
%\end{reaction}
%
%
%%% PHYSICAL UNITS
%%%
%%% Please use \unit{} and apply the exponential notation

\end{document}

### 2016-09-29 Manuscripts_GMD.tex
second commit

@@
 %% http://publications.copernicus.org/for_authors/manuscript_preparation.html
-
 %% Please use the following documentclass and journal abbreviations for discussion papers and final revised papers.
-
 %% 2-column papers and discussion papers
 \documentclass[journal abbreviation, manuscript]{copernicus}
-
 %% Journal abbreviations (please use the same for discussion papers and final revised papers)
-
 % Archives Animal Breeding (aab)
@@
 % Wind Energy Science (wes)
-
 %% \usepackage commands included in the copernicus.cls:
@@
 %\usepackage{rotating}
-
 \begin{document}
-
-\title{Plume-SPH 1.0: A 3D dust-gas volcano plume model based on smoothed partical hydrodynamics}
-
+\title{Plume-SPH 1.0: A 3D dust-gas volcano plume model based on smoothed particle hydrodynamics}
 % \Author[affil]{given_name}{surname}
-
 \Author[1]{Zhixuan}{Cao}
@@
 University at Buffalo, SUNY, New York, United State}
-
 %% The [] brackets identify the author with the corresponding affiliation. 1, 2, 3, etc. should be inserted.
-
 \runningtitle{Plume-SPH 1.0}
-
 \runningauthor{Z. Cao et al}
-
 \correspondence{Abani Patra (abani.patra@gmail.com)}
-
 \received{}
@@
 \published{}
-
 %% These dates will be inserted by Copernicus Publications during the typesetting process.
-
 \firstpage{1}
-
 \maketitle
-
 \begin{abstract}
@@
 \end{abstract}
-
 \introduction  %% \introduction[modified heading if necessary]
@@
 In addition, the need for greater skills in associated hazards forecasts with 1D models requires 2D (two dimensional) and 3D models.\\
-The development of 2D and 3D, time-dependent, and multiphase numerical models for volcano plume have provided new explanations for many features of explosive volcanism.
+The development of 2D and 3D, time-dependent, and multiphase numerical models for volcano plume has provided new explanations for many features of explosive volcanism.
 One of these 3D models is PDAC (Pyroclastic Dispersion Analysis Code) \citep{neri2003multiparticle}  which is a non-equilibrium, multiphase, 3D compressible flow model.
-Conservation equations for each phase is solved separately with finite volume method.
+Conservation equations for each phase are solved separately with finite volume method.
 And a parallel computing version of PDAC was developed lately \citep{ongaro2007parallel}.
@@
 The dynamic core of ATHAM solves the compressible Euler equations for momentum, pressure and temperature of the gas particle mixture \citep{oberhuber1998volcanic}.
-The subgrid turbulence closure scheme differentiates between the horizontal and vertical directions \citep{herzog2003prognostic} to capture turbulent mixing.
+The subgrid-scale turbulence closure scheme differentiates between the horizontal and vertical directions \citep{herzog2003prognostic} to capture turbulent mixing.
 The cloud microphysics predicts the mass of hydrometeors in liquid and ice phase \citep{herzog1998effect}.
@@
 Based above assumptions, the governing equation of our model is given as following, which is the same as governing equations of SK-3D \citep{suzuki2005numerical}.
-\begin{eqnarray}
-\dfrac{\partial \rho}{\partial t} + \nabla \cdot (\rho \textbf{v}) &=& 0 \label{eq:gov-cs-rho} \\
-\dfrac{\partial \rho \xi}{\partial t} + \nabla \cdot (\rho \xi \textbf{v}) &=& 0 \label{eq:gov-cs-ks}\\
-\dfrac{\partial \rho \textbf{v}}{\partial t} + \nabla \cdot (\rho \textbf{v} \textbf{v} + p\textbf{I}) &=& \rho \textbf{g} \label{eq:gov-cs-v} \\
-\dfrac{\partial \rho E}{\partial t} + \nabla \cdot [(\rho E + p )\textbf{v}] &=& \rho \textbf{g} \cdot\textbf{v} \label{eq:gov-cs-e}
-\end{eqnarray}
+\begin{align}
+\dfrac{\partial \rho}{\partial t} + \nabla \cdot (\rho \textbf{v}) = 0 \label{eq:gov-cs-rho} \\
+\dfrac{\partial \rho \xi}{\partial t} + \nabla \cdot (\rho \xi \textbf{v}) = 0 \label{eq:gov-cs-ks}\\
+\dfrac{\partial \rho \textbf{v}}{\partial t} + \nabla \cdot (\rho \textbf{v} \textbf{v} + p\textbf{I}) = \rho \textbf{g} \label{eq:gov-cs-v} \\
+\dfrac{\partial \rho E}{\partial t} + \nabla \cdot [(\rho E + p )\textbf{v}] = \rho \textbf{g} \cdot\textbf{v} \label{eq:gov-cs-e}
+\end{align}
 $\xi$ is the mass fraction of ejected material.
@@
 By deducting Kinetic energy from energy equation, subtracting mass conservation from momentum equation, combining transient term and advection term into material derivative term, the governing equation is put into the final formulism, in which advection term does not show up explicitly.
-\begin{eqnarray}
+\begin{align}
 \dfrac{D \rho}{D t} + \rho \nabla \cdot \textbf{v} = 0 \label{eq:gov-nc-rho}\\
@@
 \dfrac{D e}{D t} + \dfrac{P \nabla \cdot \textbf{v}}{\rho} = 0 \label{eq:gov-nc-e}
-\end{eqnarray}
+\end{align}
 Governing equations in Lagrange form, Eq.
@@
 In current model the initial domain is a box.
-The boundaries are categorized into velocity inlet (a circle area at the center of the bottom of the box), wall boundary (box bottom), pressure outlet of boundary (other faces of the box).
+The boundaries are categorized into velocity inlet (a circle area at the center of the bottom of the box), wall boundary (box bottom), pressure outlet (other faces of the box).
 \subsubsection{Velocity inlet}
@@
 The radius of vent is determined from $\rho$, $\dot M$ and $\textbf{v}$.
-To write them in terms of primitive variables.
-\begin{eqnarray}
+Equation (\ref{eq:erupt_bc_rho}) ~ (\ref{eq:erupt_bc_e}) gives velocity outlet boundary condition wrote in terms of primitive variables.
+\begin{align}
 \rho =const = p/(R_m T) \label{eq:erupt_bc_rho} \\
@@
 \dfrac{\partial e}{\partial n}=\dot M e /(\pi r^2) \label{eq:erupt_bc_e}
-\end{eqnarray} 
+\end{align} 
 Where $r$ is the radius of the vent.
@@
 As a result, internal energy flux, which consists of heat flux and energy flux carried by mass flux, will vanish on the wall boundary.
-So the wall boundary condition in terms of primitive variables are:
-\begin{eqnarray}
+Equation (\ref{eq:wall_bc_rho}) ~ (\ref{eq:wall_bc_e}) gives no-slip wall boundary condition wrote in terms of primitive variables.
+\begin{align}
 \dfrac{\partial \rho}{\partial n} = const = 0\label{eq:wall_bc_rho} \\
@@
 \dfrac{\partial e }{\partial n} = 0\label{eq:wall_bc_e}
-\end{eqnarray} 
+\end{align} 
 \subsubsection{Open outlet pressure boundary condition}
@@
 \section{SPH Method} \label{sec:SPH_method}
-SPH is a meshfree Lagragian method.
+SPH is a mesh free Lagragian method.
 In SPH, the domain is discretized as a set of particles or discretization points and the position of each particle (discretization point) is updated at every time step.
-Approximation of all field variables (velocity, density and pressure, ect.) are obtained by interpolation based on discretization points.
+Approximation of all field variables (velocity, density and pressure, ect.) is obtained by interpolation based on discretization points.
 The physical laws (such as conservation laws of mass, momentum and energy) written in the form of PDEs (partial differential equations) need to be transformed into the Lagrangian particle formalism of SPH.
@@
 There are several review papers \citep{monaghan1992smoothed, monaghan2005smoothed, rosswog2009astrophysical, price2012smoothed, monaghan2012smoothed}, giving a pretty comprehensive view over SPH, we will only refer here to the representation of the constitutive equations in SPH and focus on numerical techniques for plume modeling.
-\subsection{Fundametal principle}
-There are several routines for discretizing governing equations (partitial differential equations or ordernary differential equations) with SPH.
+\subsection{Fundamental principle}
+There are several routines for discretizing governing equations (partial differential equations or ordinary differential equations) with SPH.
 There is slight difference between discretized formulism obtained by different routines.
@@
 The start point of approximating a function with SPH is one translation property of Dirac function.
-
 \begin{equation}
@@
 The Dirac function in Eq.
-(\ref{eq:Dirac-translation}) can be approximated by a weighting funtion $w(\textbf{x}-\textbf{x}\prime, h)$ (or $w(\textbf{x}\prime-\textbf{x}, h)$) which will tend to a Dirac function when $h \rightarrow 0$ :
+(\ref{eq:Dirac-translation}) can be approximated by a weighting function $w(\textbf{x}-\textbf{x}\prime, h)$ (or $w(\textbf{x}\prime-\textbf{x}, h)$) which will tend to a Dirac function when $h \rightarrow 0$ :
 \begin{equation}
@@
 (\ref{eq:SPH-fundamental-principle}), the order of accuracy of SPH method is second order.
-However, in practice, second order can not be achieved because there is no guarantee on the symmetricity of particle distribution in real simulation \citep{price2012smoothed}.
+However, in practice, second order can not be achieved because there is no guarantee on the symmetry of particle distribution in real simulation \citep{price2012smoothed}.
 Recall that $d\textbf{x}\prime = \dfrac{dm \prime}{\rho \prime}$, the integration equation, Eq.
-(\ref{eq:SPH-fundamental-principle}), can be approximated by summation and lead to an approximation of the function at a particle (idiscretization point) a:
-
+(\ref{eq:SPH-fundamental-principle}), can be approximated by summation and lead to an approximation of the function at a particle (discretization point) a:
 \begin{equation}
@@
 \end{equation}
-
 Where the summation is over all the particles b within the region of compact support of the weighting function. 
 Gradient terms may be straightforwardly calculated by taking the derivative of Eq. (\ref{eq:SPH-approximation-sum}), giving
-
 \begin{equation}
@@
 \end{equation}
-
 For vector quantities the expressions are similar, simply replacing $A$ with $\textbf{A}$ in Eq. (\ref{eq:SPH-approximation-sum}), giving
-\begin{eqnarray}
+\begin{align}
 <\textbf{A}(\textbf{x})> \approx \sum_b m_b \dfrac{\textbf{A}_b}{\rho_b} w(\textbf{x}-\textbf{x}_b, h) \\
@@
 \label{eq:SPH-vecctor-function}
-\end{eqnarray}
+\end{align}
 %$just Italic for math symbol A$ \\
@@
 (\ref{eq:SPH-kernel-symmetric}).
-Weighting function also need to satisfy conditions such as positivity and compact support.
+Weighting function also needs to satisfy conditions such as positivity and compact support.
 In addition, kernel function must be monotonically decreasing with the distance between particles.\\
@@
 \subsection{Discretization of governing equations}
-The basic interpolants given in Eq.
+The basic interpolation given in Eq.
 (\ref{eq:SPH-approximation-sum}) to Eq.
@@
 Special treatments are also need for second order derivative terms.
-We only refer here one of these possible discretizations of compressible Euler equations with SPH:
-\begin{eqnarray}
+We only refer here one of these possible discretization forms of compressible Euler equations with SPH:
+\begin{align}
 <\rho_a> = \sum m_b w_{ab} (h_a) \label{eq:ns-sph-d} \\
@@
  0.5\sum_b m_b \textbf{v}_{a b}(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}) \cdot \nabla_a w_{a b}(h) \label{eq:ns-sph-e}
-\end{eqnarray}
+\end{align}
 Where, $\textbf{v}_{a b} = \textbf{v}_a - \textbf{v}_b$. $\Pi$ is artificial viscosity term, will be discussed in section \ref{sec:artificial-viscosity}.
@@
 We can obtain conservation of entropy by giving up conservation of mass or thermal energy adopting an alternative discretized formulations \citep{monaghan1992smoothed}.
-
 As a Lagrangian method, particle position will also been update at every time step.
@@
 (\ref{eq:ns-sph-e}) will lead to the new discretization form:
-\begin{eqnarray}
+\begin{align}
 <\rho_a> = \sum m_b w_{ab} (h_a) + \dot{\rho}(\textbf{x},t)\label{eq:ns-source-sph-d} \\
@@
  0.5\sum_b m_b \textbf{v}_{a b}(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}) \cdot \nabla_a w_{a b}(h) + \sum_b \dfrac{m_b}{\rho_b}(\kappa_a + \kappa_b) \dfrac{(T_a - T_b)}{\textbf{r}_a - \textbf{r}_b} w_{ab}(h) \label{eq:ns-conduction-sph-e}
-\end{eqnarray}
+\end{align}
 Where the source term $\dot{\rho}$ can be a "sink" of erupted vapor due to phase change which is considered in ATHAM \citep{oberhuber1998volcanic}.
@@
 Adding of new phases into SPH code will only need add several new lines for the new phase besides adding interaction terms.
-
 \subsection{Artificial viscosity} \label{sec:artificial-viscosity}
@@
 (\ref{eq:ns-sph-e}), represents artificial viscosity term which is essentially a second order derivative term.
-As in the case of first derivatives, second derivatives can be estimated by differentiating a SPH interpolant twice.
+As in the case of first derivatives, second derivatives can be estimated by differentiating a SPH interpolation twice.
 However, such formulism has two disadvantages: First, it is very sensitive to particle disorder.
@@
 Where 
-\begin{eqnarray}
+\begin{align}
 \bar{c}_{ab} = \dfrac{c_a + c_b}{2} \\
@@
 \textbf{x}_{ab}=\textbf{x}_a-\textbf{x}_b
-\end{eqnarray}
+\end{align}
 The artificial viscosity term $\Pi_{ab}$ is a Galilean invariant and vanishes for rigid rotation.
@@
 $\eta$ is usually taken as 0.1 for preventing singularities when $\textbf{x}_{ab} = 0$.
-\subsection{Timestep}
+\subsection{Time step}
 The physical properties (velocity, density, position and pressure) change every time step due to particle interactions.
@@
 \end{equation}
-For higher dimesions problem, expression for function approximation are exactly the same as Eq.
+For higher dimensions problem, expression for function approximation are exactly the same as Eq.
 (\ref{eq:CSP-function-approximation-1d}) even though the derivation is different.
@@
 where
-\begin{eqnarray}
+\begin{align}
 \text{\textbf{M}}_{\alpha \beta a} = \sum_b m_b \dfrac{\textbf{x}_b^{\alpha} - \textbf{x}_a^{\alpha}}{\rho_b} \nabla_a^{\beta} w(\textbf{x}_a -\textbf{x}_b, h) \\
@@
 \textbf{F}_{\beta a} = \sum_b m_b \dfrac{A_b - A_a}{\rho_b} \nabla_a^{\beta} w(\textbf{x}_a -\textbf{x}_b, h)
-\end{eqnarray}
+\end{align}
 $\alpha$ and $\beta$ are direcion indices, ranging from 1 to 3 for 3D problem. 
@@
 This present severe challenges to conventional Eulerian grid-based numerical methods.
-As we discussed in the introduction section, special algorithms to treat and track the interface between different phases is required for mesh-based method to deal with multiple phases problems.
+As we discussed in the introduction section, special algorithms to treat and track the interface between different phases is required for mesh-based method to deal with multiple phase problems.
 What's worse, the engulfment process during plume development makes geometry of interface very complicated.
-SPH is able to handle multiple phase problem without without additional numerical effort for dealing with interface, which makes it suitable for volcano plume simulation.
+SPH is able to handle multiple phase problems without additional numerical effort for dealing with interface, which makes it suitable for volcano plume simulation.
 Take the governing equation adopted in our model as an example.
-Different phases of fluid (air and erupted material) will be represented by two different sets of SPH particles(or discretization points).
+Different phases of fluid (air and erupted material) will be represented by two different sets of SPH particles (or discretization points).
 Based on assumption we made in section \ref{sec:physics-model}, only density need to be updated respectively for each phase.
@@
 In the areas far away from interface, updating of density is exactly the same as that for single phase flow.
-For example, in the right side and left side portition (or blue) of Fig.
+For example, in the right side and left side (or blue areas) in Fig.
 \ref{fig:SPH-multiple-density}, where there are only air particles, Eq.
@@
 That is to say, the same density updating equation can be applied for both phases and no additional numerical treatment needed to locate where is the interface.
-
 \begin{figure}
@@
 \end{figure}
-
 \subsection{Turbulence modeling with SPH}
 Typically, turbulence fluctuations in a volcanic column occur on very different length scales, ranging from millimeters up to hundreds of meters.
-The entrainment of air and heat tranfer are dominated by such turbulent fluctuations as turbulent exchange coefficient would be several magnitudes larger than molecular viscosity (and heat conduction) coefficient.
+The entrainment of air and heat transfer are dominated by such turbulent fluctuations as turbulent exchange coefficient would be several magnitudes larger than molecular viscosity (and heat conduction) coefficient.
 One way to include enough turbulence in the model is to use SPH-SPS (SPH sub-particle-scale) method.
@@
 Even though free surface flow are in nature turbulent, detailed research in turbulence using the SPH method is rather scarce.
-Among the first papers on turbulence are the implementations of the $\alpha$ turbulence model \citep{holm1999fluctuation} by Monaghan \citep{monaghan2002sph} and the Reynolds averaged Navier?Stokes (RANS) approach by Violeau and Issa \citep{violeau2007numerical}.
+Among the first papers on turbulence are the implementations of the $\alpha$ turbulence model \citep{holm1999fluctuation} by Monaghan \citep{monaghan2002sph} and the Reynolds averaged Navier-Stokes (RANS) approach by Violeau and Issa \citep{violeau2007numerical}.
 A first wall-bounded large eddy simulation (LES) was performed by Issa \citep{issa2005numerical}.
@@
 Simulations of heat transfer or other scalar transfer in turbulent flow simply involve adding transport terms for thermal energy or species concentration, at the expense of greater storage and longer computing times but without other difficulties \citep{cebeci2013analysis}.
-We adpot this strategy in our simulation of turbulent heat transfer with $SPH-\varepsilon$.
+We adopt this strategy in our simulation of turbulent heat transfer with $SPH-\varepsilon$.
 PDAC \citep{neri2003multiparticle} also adopted a Reynolds analogy in its turbulence modeling.
@@
 \subsubsection{$SPH-\varepsilon$ turbulence model}
-In $SPH-\varepsilon$ method \citep{monaghan2011turbulence}, a turbulence model is constructed within the framework of SPH in such a way that general principles such as conservation of energy, momentum and circulation are satisfied using the ideas associated with the LANS (Lagrangian averaged Navier?Stokes equations).
+In $SPH-\varepsilon$ method \citep{monaghan2011turbulence}, a turbulence model is constructed within the framework of SPH in such a way that general principles such as conservation of energy, momentum and circulation are satisfied using the ideas associated with the LANS (Lagrangian averaged Navier-Stokes equations).
 The basic idea of $SPH-\varepsilon$ is that determine a smoothed velocity $\widehat{\textbf{v}}$ by a linear operation on the unsmoothed velocity $\textbf{v}$.
@@
 \end{equation}
-Lagrange?s equations contain extra velocity dependent term which represents the stresses induced by the smoothing. Once the form of the smoothing is chosen these stresses are determined. 
+Lagrangian equations contain extra term which represents the stresses induced by the smoothing. Once the form of the smoothing is chosen these stresses are determined. 
 %The overall effect is to produce a redistribution of energy without dissipation. 
@@
 In $SPH-\varepsilon$ method, a XSPH smoothing is adopted which will conserve linear and angular momentum.
-This furtherly avoid solving of system of equations and make the method simple to implement and cheap for computation.
-The discretized form of momentum equation is obtained through lengthy derivation based on Lagrange?s equations.
+In this way solving of system of equations is avoided and it also makes the method simple to implement and cheap for computation.
+The discretized form of momentum equation is obtained through lengthy derivation based on Lagrangian equations.
 Derivation details and other discussions, such as conservation of momentum and energy, satisfactory of circulation theory, energy/momentum spectrum analysis based on Founrier transfer are detailed in Monaghan's paper \citep{monaghan2011turbulence}.
@@
 \end{equation}
-As function $G$ has the same feature as kernel function $w$, SPH approximation of the intergration will lead to:
+As function $G$ has the same feature as kernel function $w$, SPH approximation of the integration will lead to:
 \begin{equation} \label{eq:SPH-epsilon-filtering}
@@
 \end{equation}
-Where $K_{ab} = l^d G_{ab}$, $M = \rho_0 l^d$ in which $d$ is the dimension and $\rho_0$ is initial density. $SPH-\varepsilon$ turbulence model is obtained after longthy derivation:
+Where $K_{ab} = l^d G_{ab}$, $M = \rho_0 l^d$ in which $d$ is the dimension and $\rho_0$ is initial density. $SPH-\varepsilon$ turbulence model is obtained after lengthy derivation:
 \begin{equation}
@@
 This assumption is not general and atmosphere density is stratified in our plume simulation.
-And the slight  modification will not influence following derivation if $l$ is uniform and keep constant becasue:
+And the slight modification will not influence following derivation if $l$ is uniform and keep constant because:
 \begin{equation}
@@
 \end{equation}
-Where, $\mu$ is dynamic viscosity, $\kappa$ is thermal conductivity. And $\mu$  can be written in term of abslolute viscosity (kinematic viscosity) as:
+Where, $\mu$ is dynamic viscosity, $\kappa$ is thermal conductivity. And $\mu$  can be written in term of absolute viscosity (kinematic viscosity) as:
 \begin{align}
@@
 However, the above equation is correct only for 1D situation. For 2D or 3D, it is not easy to get the explicit expression of viscosity. \\
-We adopt an alternative way: obtaining a value for each pair of particles instead of persist on getting an anlytical expression.
+We adopt an alternative way: obtaining a value for each pair of particles instead of persisting on getting an analytical expression.
 Choosing smoothing function as the same as SPH kernel and the smoothing scale $l$ as the same as smoothing length $h$, the ratio between turbulent shear stress and physical shear stress is:
@@
 \end{equation}
-Notice that the number 2 at the front of the equation comes from intergration approximation of second order derivative \citep {cleary1999conduction}. By further simplification, we get:
+Notice that the number 2 at the front of the equation comes from integration approximation of second order derivative \citep {cleary1999conduction}. By further simplification, we get:
 \begin{equation}
@@
 \end{equation}
-
 \begin{equation}
@@
 As the particle-scale movement of flow is based on smoothed velocity, the velocity in the energy equation should also be smoothed.
-The filtering process will be done accroding to Eq.
+The filtering process will be done according to Eq.
 (\ref{eq:SPH-epsilon-filtering}).
@@
 \ref{eq:erupt_bc_v}).
-The correctness of eruption boundary condtion for density (Eq.
+The correctness of eruption boundary condition for density (Eq.
 (\ref{eq:erupt_bc_rho})) can be guaranteed by choosing mass of erupted ghost particles given distribution of these particles.
@@
 As soon as they enter the atmosphere, these ghost particle will be shift to real particles and their physics properties and position will be updated based on discretized governing equations.
-New ghost particle will be added at the bottom of the eruption conduit as these exising ghost particles move upwards.\\
+New ghost particle will be added at the bottom of the eruption conduit as these existing ghost particles move upwards.\\
 \subsubsection{Pressure boundary condition}
@@
 For flow with more complex channel, an exit far downstream of the flow disturbance is also feasible.
-However, the natural boundary (Eq.
-(\ref{eq:pressure_bc_p})) condition is more suitable for plume simulation where the outlet is open atmosphere.
-The way we impose pressure boundary conditon is simple: adding several layers of pressure ghost particles sourounding the real atmosphere particles.
+However, the natural boundary condition (Eq.
+(\ref{eq:pressure_bc_p})) is more suitable for plume simulation where the outlet is open atmosphere.
+The way we impose pressure boundary condition is simple: adding several layers of pressure ghost particles surrounding the real atmosphere particles.
 Physics properties except for velocity (pressure, density, temperature) are determined based on the elevation of particles.
@@
 As position of all pressure ghost particles are keep constant, we are essentially imposing a static pressure boundary condition.
-Real particles that moves out the the pressure boundary will be removed.\\
+Real particles that moves out the pressure boundary will be removed.\\
 As simulation progress, change in position and physical properties of real particles near pressure boundary boundaries might corrupt pressure boundary condition that was established initially.
-This shortcoming is relieved by choosing a larger computational domain so that these might-be-corrupted boundary will be far away from turbulent mixing area.
+This shortcoming is relieved by choosing a larger computational domain so that boundaries that might be corrupted will locate far away from turbulent mixing area.
 In addition, to avoid enlarging fluctuation, we add another constrain on time step:
@@
 \end{equation}
-Where $CFL_p$ is a safty coefficient which has similar functionalality as the normal $CFL$ number.
+Where $CFL_p$ is a safety coefficient which has similar function as the normal $CFL$ number.
 Too small $CFL_p$ will slow down simulation while too large $CFL_p$ will lose its ability of mitigate numerical fluctuation near the boundary.
@@
 This disadvantage further preventing doing simulation with finer resolution and accounting for more physics in one model.
-Non-intrusive uncertainty analysis which is commonly adopted in hazard forcasting requires finishing multiple simulations winthin a given time window.
+Non-intrusive uncertainty analysis which is commonly adopted in hazard forecasting requires finishing multiple simulations within a given time window.
 High performance computing provides a powerful routine for relieving the misery.
@@
 Less attention has been paid to developing of more flexible data management schemes for more complicated problems.
-Motivated by techniques developed for mesh based methods, we develop a complete framwork for parallelizing SPH program with MPI standard model allowing flexible and efficient data access in this paper.\\
+Motivated by techniques developed for mesh based methods, we develop a complete framework for parallelizing SPH program with MPI standard model allowing flexible and efficient data access in this paper.\\
 Any implementation of SPH code requires efficient searching and updating of neighbors during simulation.
@@
 The background grid is also used for domain decomposition in SPH.
-We refer to the elements of background grid, namely squares for two dimension and cubics for three dimension, as buckets.
+We refer to the elements of background grid, namely squares for two dimensions and cubic for three dimension, as buckets.
 As for the actual storage of data representing the physical quantities associated to each particle, different strategies have been adopted in existing implementations of SPH.
@@
 The SFC based numbering strategy is further extended to include time step information so that particles added at the same position but different time will have different identifiers.
-As for domain decomposition, even though more complicated graph-based partition tools \citep {biswas1999experiments} might get higher quality decomposition, they requires much more effort in programming and computation.
+As for domain decomposition, even though more complicated graph-based partitioning tools \citep {biswas1999experiments} might get higher quality decomposition, they requires much more effort in programming and computation.
 So we adopt an easy-programming scheme based on SFC \citep {patra1999efficient}.\\
-We will provide a supplementary documentation describing details about the data strucutre, domain decomposition, load balancing and domain adjusting.
+We will provide a supplementary documentation describing details about the data structure, domain decomposition, load balancing and domain adjusting.
 \begin{figure}[!t]
@@
 \end{figure}
-
 \begin{figure}[!t]
@@
 \caption{The effect of domain adjusting on simulation time.
-The figure on the top shows simultion time without domain adjusting, the figure on the bottom shows simultion time with domain adjusting.
+The figure on the top shows simulation time without domain adjusting, the figure on the bottom shows simulation time with domain adjusting.
 Different bins represent simulation time up to specific physics time indicated by $x$ axis.}
@@
 \end{figure}
-
-Experiments have been carried out on CCR (the CCRcomputational cluster of Center for Computational Research) at Buffalo.
+Experiments have been carried out on CCR (the computational cluster of Center for Computational Research) at Buffalo.
 E5645 CPUs running at 2.40GHz clock rate with 4GB memory per core on a Q-Logic Infiniband is used in these tests.
@@
 Memory and level 3 cache are shared on each node.
-The initial domain is $[-4.8km,4.8km]\times [-4.8km,4.8km] \times [0km, 6km]$.
+The initial domain is $[-4.8km,4.8km] \times [-4.8km,4.8km] \times [0km, 6km]$.
 Almost linear speed up is observed in our strong scalability test (Fig.
@@
 This implies that adjusting of domain can avoid computing large number of uninfluenced air particles, especially for the beginning stage of simulation.
-A domain adjusting algorithm is desinged and implemented in our code.
+A domain adjusting algorithm is designed and implemented in our code.
 Figure \ref{fig:adj_vs_no} shows that simulation time of the test problem is greatly reduced when we adopt the domain adjusting strategy in our code.
-
 \section{Verification and validation} \label{sec:verification-validation}
-We will now present a series of numerical simulations that were used to verify and validate the model.
-The SPH plume model is first verified by a JPUE simulation.
+We will now present a series of numerical simulations to verify and validate the model.
+Plume-SPH is first verified by a JPUE simulation.
 Velocity and mass fraction distribution both along the central axis and cross transverse are compared with experimental results.
-The parten of  ambient particles entrainment is also clearly shown.
+The pattern of ambient particles entrainment is also clearly shown.
 Then a simulation of representative strong volcano plume is conducted.
-Both global variables and local variable are comparable with simulation results from exisitng 3D plume models.
+Both global variables and local variable are comparable with simulation results from existing 3D plume models.
 Analysis on evolution of plume shows detailed air entrainment mechanism.
@@
 While the effect of stratified atmosphere and the effect of expansion due to high temperature in volcanic plume are not represented, JPUE reproduces the entrainment due to turbulent mixing which is one of the key elements in volcanic plume development.
-There exist consistently good experimental data \citep { list1982turbulent,dimotakis1983structure, papanicolaou1988investigations} that describe the JPUE flow field giving an insight into detailes of JPUE, such as transverse velocity and concetration profile.
+There exist consistently good experimental data \citep { list1982turbulent,dimotakis1983structure, papanicolaou1988investigations} that describe the JPUE flow field giving an insight into details of JPUE, such as transverse velocity and concentration profile.
 In this section, we will verify that our code and the $SPH-\varepsilon$ turbulence model is able to reproduce feature of turbulent entrainment by a JPUE simulation.\\
@@
 This will further simplify the scenario for the convenience of studying turbulent mixing.
-One overall feature of JPUE is "self-similarity" which means that the evolution of the JPUE is determined soley by the local scale of length and velocity which theoretically account for the fact that the rate of entrainment at the edge of JPUE is proportional to a characteristic velociy at each height.
+One overall feature of JPUE is "self-similarity" which means that the evolution of the JPUE is determined solely by the local scale of length and velocity which theoretically account for the fact that the rate of entrainment at the edge of JPUE is proportional to a characteristic velocity at each height.
 As a results, physical and numerical experiments do not necessary to have exactly the same setups and are compared on a non-dimensional basis.\\
-
 \begin{figure}
@@
 \includegraphics[width=8cm]{conc_cross}
-\caption{Dimensionless concetration distribution across the cross-section.}
+\caption{Dimensionless concentration distribution across the cross-section.}
 \label{fig:JPUE_cross-section_conc}
 \end{figure}
-
 \begin{figure}
@@
 \end{figure}
-
 \begin{figure}
@@
 \end{figure}
-
 A three dimensional axisymmetric JPUE which ejects from a round vent is simulated and results are compared with experiments \citep {george1977turbulence, papanicolaou1988investigations} for verification purpose.
@@
 Panos \citep{papanicolaou1988investigations} also fit concentration distribution and jet width based on velocity along central line into a straight line : $\dfrac{\varphi_0}{\varphi_c}=slope (\dfrac{z}{D} + intercept)$ (dash line).
-Where $\varphi$ is either velocity or concetration, the subscript $c$ represents the centraline, subscript $0$ represents the cross-sectionally averaged exit value.
-$r$ is the distance from the centraline on any cross-section.
-$z$ is the axial distance from the origin of the jet transverse section underconsideration.
+Where $\varphi$ is either velocity or concentration, the subscript $c$ represents the centreline, subscript $0$ represents the cross-sectionally averaged exit value.
+$r$ is the distance from the centreline on any cross-section.
+$z$ is the axial distance from the origin of the jet transverse section under consideration.
 $D$ is the diameter of vent.
-The coefficient $coef$ for concetration is 80 and 50 respectively according to Panos and George. $coef$ for velocity is 90 and 55 respectively according to Panos and George. 
-$slope$ for jet width based on velocity is 0.104 and for concetration is 0.157. $intercept$ for jet width based on velocity is 2.58 while that for the concetration is 4.35.\\
-Although both velocity and concetration are found to be well matched with experimental results, a small disparity in both velocity and concetration are observed near the boundary of the jet.
+The coefficient $coef$ for concentration is 80 and 50 respectively according to Panos and George. $coef$ for velocity is 90 and 55 respectively according to Panos and George. 
+$slope$ for jet width based on velocity is 0.104 and for concentration is 0.157. $intercept$ for jet width based on velocity is 2.58 while that for the concentration is 4.35.\\
+Although both velocity and concentration are found to be well matched with experimental results, a small disparity in both velocity and concentration are observed near the boundary of the jet.
 There are several factors that would attribute to such disparity.
-Reynolds number are not reported in many experiments assuming a high enough Reynolds number.
-In addition, some details of the of the experiments, such as exit velocity profile, viscosity of the experimental liquid, are not reported, neither.
+Reynolds number is not reported in many experiments assuming a high enough Reynolds number.
+In addition, some details of the experiments, such as exit velocity profile, viscosity of the experimental liquid, are not reported, neither.
 These factors prevent us from numerically reproducing these experiments in an exact way as they were conducted.
@@
 Both advanced SPH techniques like adaptive particles size and better data management and access strategies need to be exploited to address this issue.
-
 \begin{figure}
 \includegraphics[width=8cm]{JPUE_entrainment.png}
-\caption{Left figure shows shows particle distribution.
+\caption{Left figure shows particle distribution.
 The right figure shows the mass fraction of erupted material at the moment corresponding to left figure.
@@
 \end{figure}
-
 \subsection{Simulation on Plume}
@@
 A strong eruption column without wind is tested in this section for the purpose of further verification and validation.
-Input parameters, material properties and atmosphere are choosen to be the same as the strong plume no wind case in an inter-comparison study on eruptive column models \citep{costa2016results}.
+Input parameters, material properties and atmosphere are chosen to be the same as the strong plume no wind case in an inter-comparison study on eruptive column models \citep{costa2016results}.
 The strong plume scenario was based on climactic phase of the Pinatubo eruption (Philippines, 15 June 1991).
-Material properties are selected based on properties of Pinatubo and Shinmoe-dake eruption.Both global variable and local variables are compared with existing models.
-Then then plume evolution process are investigated.
-
+Material properties are selected based on properties of Pinatubo and Shinmoe-dake eruption.
+Both global variable and local variables are compared with existing models.
+Then plume evolution process is investigated.
 \subsubsection{Global and local variables}
@@
 As particles distribute in the space in a disordered manner.
-We first project simulation results (on disordered particles) onto a pre-defined grid before doing time average and spatial intergration.
+We first project simulation results (on disordered particles) onto a pre-defined grid before doing time average and spatial integration.
 The interpolation method is the basic SPH kernel based interpolation.
@@
 \end{figure}
-
 %\begin{figure}
@@
 %\end{figure}
-
-The profiles of local variables matche well with simulation results of exisitng 3D models in general sense. The basic phenomena in volcano plume development is correctly captured by our model.
-%: the plume first rises up to its top height and then falls down to the netural buoyancy height and starts expanding.\\
+The profiles of local variables match well with simulation results of existing 3D models in general sense. The basic phenomena in volcano plume development is correctly captured by our model.
+%: the plume first rises up to its top height and then falls down to the neutral buoyancy height and starts expanding.\\
 %
@@
 As the height increase, the amount of entrained air will increase.
-Around the netural height, where the umbrella expanding, the entrainment of air shows a slight decrease due to lack of air surrounding the column at that height.
+Around the neutral height, where the umbrella expanding, the entrainment of air shows a slight decrease due to lack of air surrounding the column at that height.
 The profile for gas, which account for both air and vapor, shows a very similar tendency as that of entrained air.
-First of all, vapor condese is not considered in our model.
+First of all, vapor condense is not considered in our model.
 In addition, as we assume that erupted material behaves like a single phase water, the mass fraction of gas will simply be a function of entrained air (Eq.
@@
 However, the profile of entrained air and profile of gas predicted by ATHAM are still very close to each other which implies that ignoring of water phase change is a valid assumption for eruptions similar to this test case (strong plume with erupted water fraction in erupted material less than 5\%).
-Because air occupies a larger portition of the gas and plays a more significant role.
+Because air occupies a larger portion of the gas and plays a more significant role.
 As for mass fraction of solid, similarly, we have Eq.
@@
 (\ref{eq:solid-frac-express2}).
-PDAC which treat partices of two different size range as two separate phases got similar mass fraction profile.
+PDAC which treat particles of two different sizes as two separate phases got similar mass fraction profile.
 Which implies that our model is at least valid for eruptions similar to the test case.
@@
 However, there is still a big disparity between the profiles of local variables of our model and 3K-3D.
-One of the big difference between our model and 3K-3D is that we adopt a LANS type of turbulence modle while 3K-3D adopts a LES (large eddy simulation) turbulence model.
+One of the big differences between our model and 3K-3D is that we adopt a LANS type of turbulence model while 3K-3D adopts a LES (large eddy simulation) turbulence model.
 This implies that turbulence model plays a critical role in plume simulation.
@@
 \includegraphics[width=5cm]{t15}
-\caption{Contour of mass fraction and velocity quiver at 15 seconds, column rises up quickly due to intial momentum.}
+\caption{Contour of mass fraction and velocity quiver at 15 seconds, column rises up quickly due to initial momentum.}
 \label{fig:strong_t15}
 \end{figure}
-
 \begin{figure}
@@
 \end{figure}
-
 Figure \ref{fig:strong_t15}~\ref{fig:strong_t300} shows the dynamic evolution process of volcano plume.
 The contour level of mass fraction are 0.7, 0.5, 0.3, 0.1, $1\times 10 ^{-1}$, $1\times 10 ^{-2}$, $1\times 10 ^{-3}$, $1\times 10^{-4}$ and $1\times 10^{-5}$.
-The plume keeps rising up and entraining air into the plume untill it reaches its top height(around t=200 seconds).
-Afterwards the mixtures of erupted material and air then falls back to neutral height and starts spreading radially.
-Detailed velocity quiver plots show entrainning of air into the plume at the edge of column.
-The evlolution of the plume further verifies our model.
+The plume keeps rising up and entraining air into the plume until it reaches its top height(around t=200 seconds).
+Afterwards the mixtures of erupted material and air then fall back to neutral height and starts spreading radially.
+Detailed velocity quiver plots show entraining of air into the plume at the edge of column.
+The evolution of the plume further verifies our model.
 %
@@
 A new plume model was developed based on SPH method.
-Advanced numerical techniques in SPH were exploited and talored for this model.
-High performance computing was used to mitigate confliction between accuracy (depends on comprehensiveness of the model, resolution, order of accuracy of numerical methods, shceme for time upgrading ...) and simulation time (depends on, comprehensiveness, resolution, order of accuracy of numerical methods, shceme for time upgrading ...
+Advanced numerical techniques in SPH were exploited and tailored for this model.
+High performance computing was used to mitigate confliction between accuracy (depends on comprehensiveness of the model, resolution, order of accuracy of numerical methods, scheme for time upgrading.) and simulation time (depends on, comprehensiveness, resolution, order of accuracy of numerical methods, scheme for time upgrading ...
 and computational techniques).
 The correctness of the code and model was verified and validated by a series of test simulations.\\
-Currently existing 3D models focus on certain aspect of the volcano plume (PDAC on pyroclastic flow, ATHAM on microphysics, and SK-3D on entrainment with higher accuracy and higer oder of accuracy) and hence, naturely, different assumptions were made in these models.
+Currently existing 3D models focus on certain aspect of the volcano plume (PDAC on pyroclastic flow, ATHAM on microphysics, and SK-3D on entrainment with higher accuracy and higher order of accuracy) and hence, naturally, different assumptions were made in these models.
 However, these different aspects of volcano plume are not independent and actually are coupled.
-In addtion, there is no absolute boundary to determine which kind of hazard is dominant in certain eruption.
+In addition, there is no absolute boundary to determine which kind of hazard is dominant in certain eruption.
 So it is necessary to simulate all associated hazards in one model.
 Actually, effort has already been put on developing more comprehensive plume models.
-For example, a large-particle module (LPM) was added to ATHAM to track the paths of rocky particles (pyroclasts or tephra) within the plume and predicts where these particles fall \citep{kobs2009modeling}.
+For example, a large-particle module (LPM) was added to ATHAM to track the paths of rocky particles (pyroclastic or tephra) within the plume and predicts where these particles fall \citep{kobs2009modeling}.
 We were also motivated by such a tendency of plume modeling to choose SPH as our numerical tool besides its ability in dealing with interface for multiple phase flows.
 As mentioned in the introduction section, SPH method has good extension features and adding of new physics and phases requires tiny modification on the code compared with mesh-based methods.
-The last but not the least, the dramatisc development of computational power makes it possible to establish a comprehensvie model.
+The last but not the least, the dramatic development of computational power makes it possible to establish a comprehensive model.
 Even though right now, current computational capacity does not allow us to have too comprehensive model, the easy-extension feature of SPH makes it feasible to keep adding new physics into the model when necessary.
-What presented in this paper is actually our initial effort and results towards developing a first priciple based plume model with comprehensive physics, adopting proper numerical tools and power of high performance computing.
+What presented in this paper is actually our initial effort and results towards developing a first principle based plume model with comprehensive physics, adopting proper numerical tools and power of high performance computing.
 More advanced numerical techniques, such as adaptive particle size, Godnuv-SPH, semi-explicit time advancing scheme and better data management strategies and algorithms are on our list to exploit.
@@
 However, implementation of different resolution in SPH methods is not straight forward and requires more work.
-
 \section{Code availability}
@@
 \section{Data availability}
-Data for running the test cases, including meterological data, properties of erupted material and air, and eruption condition data can be found in source code repository.
+Data for running the test cases, including meteorological data, properties of erupted material and air, and eruption condition data can be found in source code repository.
 Raw data of simulation results is archived in google drive.
 Access permission will be given up request.
-Post-processed data for generating figures is provided as supplyment along with the manuscripts.
+Post-processed data for generating figures is provided as supplement along with the manuscripts.
 Please note that some of these pictures are generating by visualizing the raw data with visualization tools, there is no separate data for these figures.
-
 %\appendix
@@
 %\appendixtables   %% needs to be added in front of appendix tables in one-column style (\documentclass[acp, manuscript]{copernicus})
-
 \authorcontribution{TEXT}
-
 \competinginterests{Authors have the standard conflict with employers and listed sponsors}
-
 %\disclaimer{TEXT}
-
 \begin{acknowledgements}
@@
 \end{acknowledgements}
-
 %% REFERENCES
-
 %% The reference list is compiled as follows:
@@
 \bibliography{Reference.bib}
-
 %% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,
@@
 %% DOI = {10.5194/xyz}
-
 %% LITERATURE CITATIONS
@@
 %% \citeyear{jones90}|            & 1990
-
 %% FIGURES
-
 %% When figures and tables are placed at the end of the MS (article in one-column style), please add \clearpage
 %% between bibliography and first table and/or figure as well as between each table and/or figure.
-
 %% ONE-COLUMN FIGURES
-
 %%f
@@
 %\begin{algorithm}
-%\caption{}
+%\caption{_}
 %\label{a1}
 %\begin{algorithmic}
-%
+%_
 %\end{algorithmic}
@@
 %%% Please use \unit{} and apply the exponential notation
-
 \end{document}

### 2016-10-01 Manuscripts_GMD.tex
add a pic

@@
 All of existing 3D (three dimensional) plume models are using mesh-based methods.
-SPH (Smoothed particle hydrodynamics), as a mesh free method has several advantages over mesh-based methods in modeling of multiple phase free boundary flow.
+SPH (Smoothed particle hydrodynamics), as a mesh free method has several advantages over mesh-based methods in modeling of multiphase free boundary flow.
 As an initial effort on exploiting the feasibility and advantage of SPH in volcano plume modeling we adopt a relative simpler model, a 3D dusty-gas dynamic model, targeting at capturing the more basic features of volcano plume.
-Erupted material is represented by a single phase while air is another phase.
-Immediate dynamic and thermal dynamic equilibrium between air and erupted material are assumed in this model.
-Affect of wind is also ignored.
+In this model, erupted material is assumed to be well mixed and represented by one phase while air is another phase.
+Immediate dynamic and thermal dynamic equilibrium between air and erupted material are assumed.
+Affect of wind is not considered.
 Several newly developed techniques in SPH are adopted and trimmed to address numerical challenges in simulating multiphase compressible turbulent flow with classical SPH method.
@@
 Corrected formulism of SPH is used to handle tensile instability and deficiency of particle distribution near the boundaries.
-We also propose feasible ways to impose velocity inlet and pressure outlet boundary conditions.
-The solver of this model is parallelized with MPI (message passing interface) obtaining good weak scalability and strong scalability.
-The model is first verified by a comparing velocity and concentration distribution along axis and cross section with experimental results of JPUE (jet or plume which is ejected from a nozzle into a uniform environment).
+We also propose feasible ways to impose velocity inlet and pressure outlet boundary conditions, both of which are scarce in traditional implementation of SPH.
+Core solver of our model is parallelized with MPI (message passing interface) obtaining good weak scalability and strong scalability.
+The model is first verified by comparing velocity and concentration distribution along axis and cross section with experimental results of JPUE (jet or plume which is ejected from a nozzle into a uniform environment).
 The top height of Pinatubo eruption (15 June 1991) simulated by our model is consistent with both observation and existing 3D plume models.
@@
 <\dfrac{d e_{\alpha}}{d t}>
-& = 0.5\sum_b [m_b \widehat{\textbf{v}_{\alpha b}} (\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}) \bigtriangledown_{\alpha}w_{\alpha b}(h_{\alpha}] 
+& = 0.5\sum_b [m_b \widehat{\textbf{v}_{\alpha b}} (\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}) \bigtriangledown_{\alpha}w_{\alpha b}(h_{\alpha})] 
  + 2 \sum_b \dfrac{m_b}{\rho_{\alpha} \rho_b} \kappa_{t,\alpha b} (T_{\alpha} - T_b) F_{\alpha b} (h_{\alpha}) \\
@@
 The proper $CFL_p$ is determined by series of simulation tests.
+\begin{figure}
+\includegraphics[width=18cm]{t120_bc_proc.png}
+\caption{A cross section view of the simulation domain.
+Subfigure a shows the mass fraction of a plume simulation.
+Subfigure b shows all boundary conditions: the dark blue is for velocity inlet boundary, the light blue is for pressure outlet boundary, the gray is for no-slip wall boundary condition.
+Ghost particle ID of All real particles are set to 100, so is 100 in b.
+Subfigure c shows the cross section view of domain decomposition based on SFC.
+The simulation is conducted on 144 processors (12 nodes with 12 cores each node), so there are 144 subdomains in total.}
+\label{fig:bc_and_domain_decomp}
+\end{figure}
 \subsection{Parallellism and performance}
@@
 \begin{equation}
-n_s = (1 - n_a) + (1- n_{g0})
+n_s = (1 - n_a) (1- n_{g0})
 \label{eq:solid-frac-express1}
@@
 n_s = 1 - (n_a + n_g)
-\label{eq:solid-frac-express1}
+\label{eq:solid-frac-express2}
 \end{equation}
@@
 \section{Code availability}
-The Plume-SPH code with a user manual providing instructions for installation, running and visualization is archive in github (https://github.com/Plume-SPH/plume-sph.git).
+The Plume-SPH code with a user manual providing instructions for installation, running and visualization are archived in github (https://github.com/Plume-SPH/plume-sph.git).
 The input data for all simulations presented in this work can also be found in the repository.
@@
 Data for running the test cases, including meteorological data, properties of erupted material and air, and eruption condition data can be found in source code repository.
-Raw data of simulation results is archived in google drive.
-Access permission will be given up request.
-Post-processed data for generating figures is provided as supplement along with the manuscripts.
-Please note that some of these pictures are generating by visualizing the raw data with visualization tools, there is no separate data for these figures.
+Raw data of simulation results are archived in google drive.
+Access permission will be given upon request.
+Post-processed data for plotting figures is provided as supplement along with the manuscript.
+Please note that some of these pictures are generated by visualizing the raw data with visualization tools, there is no separate data for these figures.
 %\appendix

### 2016-10-03 Manuscripts_GMD.tex
add subsections in introduction

@@
 \introduction  %% \introduction[modified heading if necessary]
+\subsection{Volcanic ash hazards}
 The dynamics of explosive eruption clouds has been the central issue of volcanology science for a long time.
@@
 %
+\subsection{Existing plume models}
 The basic physics behind of volcanic plume has been understood for more than half a century \citep{morton1956turbulent, settle1978volcanic, wilson1978control}.
@@
 %
+\subsection{Features of SPH}
 All of these existing 3D models are based on mesh-based methods (Eulerian methods).
@@
 %
+\subsection{Our contribution}
 As SPH is a relative new method in computational fluid dynamics, implementation of SPH for compressible multiphase turbulent flow is scarce.
@@
 The core solver of our model is parallelized by MPI (message passing interface standard) parallelism.
-In addition, a dynamic load balancing strategy is also adopted.
+In addition, a dynamic load balancing strategy is also developed.
 What's more, imposing of some types of boundary conditions (such as eruption boundary condition) requires dynamically adding and removing of particles during simulation.
-To address this issue, we adopt an efficient data management scheme based on SFC (space filling curve) and hash table.\\
+To address this issue, we adopt an efficient data management scheme based on time dependent SFC (space filling curve) and hash table.
+The computational cost is further reduced by adjusting simulation domain adaptively.\\
 The organization of this paper is as following:
@@
 The proper $CFL_p$ is determined by series of simulation tests.
+\\
+In subfigure b of Fig. \ref{fig:bc_and_domain_decomp}, different type of boundary conditons are marked with different color. The red portion are main computational domain. 
 \begin{figure}
@@
 As for domain decomposition, even though more complicated graph-based partitioning tools \citep {biswas1999experiments} might get higher quality decomposition, they requires much more effort in programming and computation.
-So we adopt an easy-programming scheme based on SFC \citep {patra1999efficient}.\\
+So we adopt an easy-programming scheme based on SFC \citep {patra1999efficient}.
+Subfigure c of Fig.
+\ref{fig:bc_and_domain_decomp} shows a cross section view of domian decomposition.\\
 We will provide a supplementary documentation describing details about the data structure, domain decomposition, load balancing and domain adjusting.

### 2016-10-04 Manuscripts_GMD.tex
slight modification

@@
 The accuracy of source terms is crucial for VATDs.
-All of existing 3D (three dimensional) plume models are using mesh-based methods.
-SPH (Smoothed particle hydrodynamics), as a mesh free method has several advantages over mesh-based methods in modeling of multiphase free boundary flow.
+All of existing 3D (three dimensional) plume models are using mesh based methods.
+SPH (Smoothed particle hydrodynamics), as a mesh free method has several advantages over mesh based methods in modeling of multiphase free boundary flow.
 As an initial effort on exploiting the feasibility and advantage of SPH in volcano plume modeling we adopt a relative simpler model, a 3D dusty-gas dynamic model, targeting at capturing the more basic features of volcano plume.
@@
 Immediate dynamic and thermal dynamic equilibrium between air and erupted material are assumed.
-Affect of wind is not considered.
+Effect of wind is not considered.
 Several newly developed techniques in SPH are adopted and trimmed to address numerical challenges in simulating multiphase compressible turbulent flow with classical SPH method.
@@
 Usually, plume models are used to provide sources terms for VATDs.
-This paper reports a new 3D (three dimensional) volcano plume model exploiting advantage of mesh-free methods in multiple phase free boundary flow modeling.\\
+This paper reports a new 3D (three dimensional) volcano plume model exploiting advantage of mesh-free methods in multiphase free boundary flow modeling.\\
 %
 \subsection{Existing plume models}
-The basic physics behind of volcanic plume has been understood for more than half a century \citep{morton1956turbulent, settle1978volcanic, wilson1978control}.
+The basic physics of volcanic plume has been understood for more than half a century \citep{morton1956turbulent, settle1978volcanic, wilson1978control}.
 Several volcano plume models have been developed in the past few decades.
@@
 Recently developed 1D models (eg.
-\citep{mastin2007user, woodhouse2013interaction}) tend to accounting for more physics effects in one model.
-FPLUME-1.0 \citep{folch2016fplume} accounts for wind affect, entrainment of moisture, water phase change, particle fallout and re-entrainment and even wet aggregation of ash.
-However, In these 1D models, the entrainment of air is evaluated based two coefficients: entrainment coefficient for the vertical plume and the entrainment coefficient that describes the effect of wind.
+\citep{mastin2007user, woodhouse2013interaction}) tend to account for more physics effects in one model.
+FPLUME-1.0 \citep{folch2016fplume} accounts for wind effect, entrainment of moisture, water phase change, particle fallout and re-entrainment and even wet aggregation of ash.
+However, In these 1D models, the entrainment of air is evaluated based on two coefficients: entrainment coefficient for the vertical plume and the entrainment coefficient that describes the effect of wind.
 Different 1D models adopt different entrainment coefficients based on specific formulation or calibration against well-documented case studies.
@@
 In addition, these simplified models requires little computational sources and can run on standard personal computers or on web sites in very short time.
-As a result, 1D software (such as \citep{267, 1194, 3541}) for volcano plume development, combined with VATDs (such as \citep{114, draxler2015hysplit}) are widely used in research and practice.
+As a result, 1D software for volcano plume development (such as \citep{267, 1194, 3541}), combined with VATDs (such as \citep{114, draxler2015hysplit}) are widely used in research and practice.
 While these 1D models can generate well-matched results with 3D (three dimensional) models for weak plumes, much greater variability is observed for strong plume scenarios, especially for local variables \citep{costa2016results}.
@@
 Additional modules include radiation \citep{langmann1997radiative}, gas phase chemistry \citep{trentmann2002simulation} and gas scavenging by hydrometeors \citep{textor2003injection} were added lately.
-An further extension was made to include particle aggregation in 2006 \citep{textor2006volcanic1, textor2006volcanic2}.
+A further extension was made to include particle aggregation in 2006 \citep{textor2006volcanic1, textor2006volcanic2}.
 However, the resolution of ATHAM is still pretty coarse compared with SK-3D and PDAC.
@@
 Both are using finer and finer resolution.
-Recently, A first order, non-equilibrium compressible 3D model, ASHEE \citep{cerminara2016ashee}, is derived based on three dimensional N phases Eulerian transportation equation.
+Recently, a first order, non-equilibrium compressible 3D model, ASHEE \citep{cerminara2016ashee}, is derived based on three dimensional N phases Eulerian transportation equation.
 ASHEE is valid for low concentration and low Strokes number region and much faster than N-phase Eulerian model.
-The model is based an open source numerical solver OpenFOAM adopting unstructured finite volume method.\\
-To summarize, each model has its own focus based on problem of interest.
+The model is based an open source numerical solver OpenFOAM adopting unstructured Finite Volume method.\\
+To summarize, each 3D model has its own focus based on problem of interest.
 Different simplification was made based on their scope.
@@
 \subsection{Features of SPH}
-All of these existing 3D models are based on mesh-based methods (Eulerian methods).
+All of these existing 3D models are based on mesh based methods (Eulerian methods).
 And there is no 3D plume model based on mesh free method (Lagrange methods) yet.
@@
 \item With more than one material, each described by its own set of particles, interface problems between phases are often trivial for SPH but difficult for mesh based schemes.
-So multiple phase flow can be easily handled by SPH.
+So multiphase flow can be easily handled by SPH.
 Adding of new phases to the model also does not require a major overhaul and re-tooling.
@@
 Both the fundamental discretization formulism and techniques that used to handle specific issues involved in plume modeling will be discussed.
-Validation and verification with numerical tests is presented in section \ref{sec:verification-validation}.
+Verification and validation with numerical tests is presented in section \ref{sec:verification-validation}.
 In section \ref{sec:conclusion}, a discussion on future work is given following a brief summary.
@@
 Latent heat released from phase change of erupted vapor will further heat up entrained air and make the bulk density further dilute.
-Even entrained vapor might also experience similar process and impose impact on plume development.
+Even entrained vapor might also experience a similar process and impose impact on plume development.
 Particle aggregation process \citep{carey1982influence,taddeucci2011aggregation}, either due to presence of liquid water, resulting from particle collision or driven by electrostatic forces might occur inside plume and thereby affect the sedimentation.
-This is essentially a multiple phases turbulent mixing process coupled with heat transfer and many other microphysics.
+All in all, the process of plume development is  essentially a multiphase turbulent mixing process coupled with heat transfer and many other microphysics and chemical reaction.
 As an initial effort on exploiting the feasibility and advantage of SPH in plume modeling, our model is designed to describe an injection of well mixed solid and volcanic gas from a circular vent above a flat surface into a stratified stationary atmosphere following SK-3D \citep{suzuki2005numerical}.
@@
 Because of above assumptions, all other micro-physics process (like phase change of  \texorpdfstring{H\textsubscript{2}O}, aggregation, decomposition, absorption of gas on the surface of solid, solution of gas into the liquid) and chemical process will not be considered in this model.
-These microphysics factors will come to play critical roles under particular eruption conditions.
-Capturing of these processes need a more comprehensive model.
-One critical element in plume development, the affect of wind, is also not considered in this model yet.
+These ignored microphysics factors will come to play critical roles under particular eruption conditions.
+Capturing of these processes needs a more comprehensive model.
+One critical element in plume development, the effect of wind, is also not considered in this model yet.
 Introducing of wind effect in our model will need imposing of pressure boundary conditions in a dynamic way, which requires more numerical effort and algorithm design.
-To summarize, our model is not valid for eruptions where wind affect plays a significant role in its development, usually refereed as weak plume.
-Our model is also lack ability in modeling plumes with large particles or in which microphysics plays non-ignorable roles, such as a eruption of \textit{El Chich{\'o}n} volcano at April 4th \citep{sigurdsson19841982, folch2016fplume}.
+To summarize, our model is not valid for eruptions where wind effect plays a significant role in its development, usually refereed as weak plume.
+Our model is also lack ability in modeling plumes with large particles or eruptions in which microphysics plays non-ignorable roles, such as a eruption of \textit{El Chich{\'o}n} volcano at April 4th \citep{sigurdsson19841982, folch2016fplume}.
 At present stage, we are more interested in more basic (and more critical) aspects of volcano plume and therefor devote our effort to this relative simpler model.
@@
 $\xi$ is the mass fraction of ejected material.
-$E = e + K $ is total energy which is summation of kinetic energy $K$ and internal energy $e$.
+$E = e + K $ is total energy which is a summation of kinetic energy $K$ and internal energy $e$.
 An additional equation is required to close the system. In this model, the equation for closing the system is the EOS (equation of state) for ideal gas.
@@
 %
-\section{SPH Method} \label{sec:SPH_method}
+\section{SPH method} \label{sec:SPH_method}
 SPH is a mesh free Lagragian method.
@@
 Approximation of all field variables (velocity, density and pressure, ect.) is obtained by interpolation based on discretization points.
-The physical laws (such as conservation laws of mass, momentum and energy) written in the form of PDEs (partial differential equations) need to be transformed into the Lagrangian particle formalism of SPH.
+The physical laws (such as conservation laws of mass, momentum and energy) written in the form of PDEs (partial differential equations)  or ODEs (ordinary differential equations) need to be transformed into the Lagrangian particle formalism of SPH.
 Using a kernel function that provides the weighted estimation of the field variables at any point, the integral equations are evaluated as sums over neighbor particles.
-Thus, physical properties (density, velocity, internal energy, ect.) associated to the particle are updated based on its neighbors.
+Thus, field variables associated to the particle are updated based on its neighbors.
 The kernel function has a compact support determined by smoothing length of each particle.
-There are several review papers \citep{monaghan1992smoothed, monaghan2005smoothed, rosswog2009astrophysical, price2012smoothed, monaghan2012smoothed}, giving a pretty comprehensive view over SPH, we will only refer here to the representation of the constitutive equations in SPH and focus on numerical techniques for plume modeling.
+There are several review papers \citep{monaghan1992smoothed, monaghan2005smoothed, rosswog2009astrophysical, price2012smoothed, monaghan2012smoothed}, giving a pretty comprehensive view over SPH, we will only refer here to the representation of the constitutive equations in SPH and put more focus on specific numerical techniques for plume modeling.
 \subsection{Fundamental principle}
-There are several routines for discretizing governing equations (partial differential equations or ordinary differential equations) with SPH.
+There are several routines for discretizing governing equations (PDEs or ODEs) with SPH.
 There is slight difference between discretized formulism obtained by different routines.
@@
 Recall that $d\textbf{x}\prime = \dfrac{dm \prime}{\rho \prime}$, the integration equation, Eq.
-(\ref{eq:SPH-fundamental-principle}), can be approximated by summation and lead to an approximation of the function at a particle (discretization point) a:
+(\ref{eq:SPH-fundamental-principle}), can be approximated by summation and lead to an approximation of the function at particle (discretization point) a:
 \begin{equation}
@@
 \end{equation}
-Where the summation is over all the particles b within the region of compact support of the weighting function. 
+Where the summation is over all the particles within the region of compact support of the weighting function. 
 Gradient terms may be straightforwardly calculated by taking the derivative of Eq. (\ref{eq:SPH-approximation-sum}), giving
@@
 \end{equation}
-For vector quantities the expressions are similar, simply replacing $A$ with $\textbf{A}$ in Eq. (\ref{eq:SPH-approximation-sum}), giving
+For vector quantities the expressions are similar, simply replacing $A$ with $\textbf{A}$ in Eq. (\ref{eq:SPH-approximation-sum}) and and Eq. (\ref{eq:SPH-scalar-function-gradient}), giving
 \begin{align}
@@
 \end{align}
-%$just Italic for math symbol A$ \\
-%$\textbf{bold Italic for vector A}$\\
-%$\textbf{\text{bold norm for matrix A}}$ \\
-%\textit{ Italic \textbf{bold Italic, A for vector}}
-%\textbf{For matrix, Just bold A}
+%
 \subsection{Weighting function}
@@
 \end{align}
-Where, $\textbf{v}_{a b} = \textbf{v}_a - \textbf{v}_b$. $\Pi$ is artificial viscosity term, will be discussed in section \ref{sec:artificial-viscosity}.
+Where, $\textbf{v}_{a b} = \textbf{v}_a - \textbf{v}_b$. $\Pi$ is artificial viscosity term, which will be discussed in section \ref{sec:artificial-viscosity}.
 However, these discretized formulations will not guarantee conservation of entropy.
@@
 Adding new physics and new phases into the model is trivial in terms of discretization.
-For example, add of new source (or sink) into Eq.
-(\ref{eq:ns-sph-d}), add a drag force into Eq.
-(\ref{eq:ns-sph-v})  and a heat exchange term into Eq.
+For example, adding of new source (or sink) into Eq.
+(\ref{eq:ns-sph-d}), adding a drag force into Eq.
+(\ref{eq:ns-sph-v})  anding a heat exchange term into Eq.
 (\ref{eq:ns-sph-e}) will lead to the new discretization form:
@@
 In that case, each phase will need one set of governing equations of Navier-Strokes type.
-Adding of new phases into SPH code will only need add several new lines for the new phase besides adding interaction terms.
+Adding of new phases into SPH code only needs to add several new lines for the new phase besides adding interaction terms.
 \subsection{Artificial viscosity} \label{sec:artificial-viscosity}
-In classic SPH, shock waves are handled by introducing artificial viscosity to smear out discontinuities.
+In classical SPH, shock waves are handled by introducing artificial viscosity to smear out discontinuities.
 $\Pi$, in discretized momentum equation, Eq.
@@
 $\alpha$ and $\beta$ are direcion indices, ranging from 1 to 3 for 3D problem. 
-\subsection{multiple phase SPH}
+\subsection{Multiphase SPH}
 Numerical simulation of multiphase flows is usually difficult due to the existence of interface between two fluids.
@@
 This present severe challenges to conventional Eulerian grid-based numerical methods.
-As we discussed in the introduction section, special algorithms to treat and track the interface between different phases is required for mesh-based method to deal with multiple phase problems.
+As we discussed in the introduction section, special algorithms to treat and track the interface between different phases is required for mesh based method to deal with multiphase problems.
 What's worse, the engulfment process during plume development makes geometry of interface very complicated.
-SPH is able to handle multiple phase problems without additional numerical effort for dealing with interface, which makes it suitable for volcano plume simulation.
+SPH is able to handle multiphase problems without additional numerical effort for dealing with interface, which makes it suitable for volcano plume simulation.
 Take the governing equation adopted in our model as an example.
@@
 \begin{equation}
-<\rho_{\alpha}^a>=\frac{\sum m_b w_{\alpha b} (h_{\alpha})}{\sum \frac{m_b}{\rho_b} w_{\alpha b} (h_{\alpha}) +\sum \frac{m_j}{\rho_j} w_{\alpha j} (h_\alpha)} \label{eq:gov-sph-d1}
-\end{equation}
-\begin{equation}
-<\rho_\alpha^{sg}>=\frac{\sum_j m_j w_{\alpha j} (h_\alpha)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} (h_{\alpha}) +\sum \frac{m_j}{\rho_j} w_{\alpha j} (h_\alpha)} \label{eq:gov-sph-d2}
+<\rho_{\alpha}^a>=\frac{\sum m_b w_{\alpha b} (h_{\alpha})}{\sum \frac{m_b}{\rho_b} w_{\alpha b} (h_{\alpha}) +\sum \frac{m_j}{\rho_j} w_{\alpha j} (h_{\alpha})} \label{eq:gov-sph-d1}
+\end{equation}
+\begin{equation}
+<\rho_\alpha^{sg}>=\frac{\sum_j m_j w_{\alpha j} (h_\alpha)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} (h_{\alpha}) +\sum \frac{m_j}{\rho_j} w_{\alpha j} (h_{\alpha})} \label{eq:gov-sph-d2}
 \end{equation}
@@
  $\rho_i^{sg}$ is density of phase 2 (erupted material).
-$\rho=\rho^a + \rho^{sg}$ is density of mixture of phase 1 and phase 2.
+$\rho=\rho^a + \rho^{sg}$ is density of mixture of air and erupted material. By definition, the mass fraction is updated according to Eq. (\ref{eq:gov-sph-xi}).
+\begin{equation}
+<\xi_{\alpha}> = \dfrac{\rho^{sg}_{\alpha}}{\rho_{\alpha}}
+\label{eq:gov-sph-xi}
+\end{equation}
 In the areas far away from interface, updating of density is exactly the same as that for single phase flow.
@@
 \end{equation}
-That is to say, the same density updating equation can be applied for both phases and no additional numerical treatment needed to locate where is the interface.
+That is to say, the same density updating equation can be applied for both phases and no additional numerical treatment needed to locate where the interface is.
 \begin{figure}
@@
 Another way is to use extremely fine resolution (The direct numerical simulation method) at the expense of much higher computational cost that results from both increase in number of particles and decrease in time steps constrained by CFL condition.
-Here we choose the SPH-SPS method to capture  momentum and energy exchange caused by smaller scale turbulence.\\
-Even though free surface flow are in nature turbulent, detailed research in turbulence using the SPH method is rather scarce.
-Among the first papers on turbulence are the implementations of the $\alpha$ turbulence model \citep{holm1999fluctuation} by Monaghan \citep{monaghan2002sph} and the Reynolds averaged Navier-Stokes (RANS) approach by Violeau and Issa \citep{violeau2007numerical}.
-A first wall-bounded large eddy simulation (LES) was performed by Issa \citep{issa2005numerical}.
+Here we choose the SPH-SPS method.\\
+Even though free surface flow are in nature turbulent, detailed research in turbulence using SPH method is rather scarce.
+Among the first papers on turbulence are the implementations of the $\alpha$ turbulence model \citep{holm1999fluctuation} by Monaghan \citep{monaghan2002sph} and the RANS (Reynolds averaged Navier-Stokes) approach by Violeau \citep{violeau2007numerical}.
+A first wall-bounded LES (large eddy simulation) simulation with SPH was performed by Issa \citep{issa2005numerical}.
 An improved version of $SPH-\alpha$ method is proposed in 2011 by Monaghan \citep{monaghan2011turbulence}, which we refer to as $SPH-\varepsilon$ method.
-We will adopt $SPH-\varepsilon$ method in simulation of volcano plume.
-As most of these papers constrain their focus on incompressible or weakly compressible fluid.
-To simulate turbulence in plume, we need to extend existing incompressible turbulent modeling methods to compressible flows.\\
+We will adopt $SPH-\varepsilon$ method in our model.
+As most of these papers constrain their focus on incompressible or weakly compressible flow.
+To simulate turbulence in plume, we need to extend existing incompressible turbulent modeling methods to compressible flow.\\
 $SPH-\alpha$ model \citep{monaghan2002sph} is a powerful extension of the XSPH algorithm which reduces disorder at short length scales and retains the constants of the motion.
-But in that paper \citep{monaghan2002sph}, energy conservation equation is discretized in a way that is exactly the same as usual SPH.
+But in that paper, energy conservation equation is discretized in a way that is exactly the same as usual SPH.
 There is no special treatment on energy conservation equation.
@@
 Actually, there are several additional terms in the averaged compressible Navier-Stokes equations besides the Reynolds stress term \citep{NASACompressibleTurbulence}.
-The averaged momentum equation for compressible flow are in the same form as that for incompressible flow, all of other additional terms except for turbulent stress term are in the energy equation).
+The averaged momentum equation for compressible flow are in the same form as that for incompressible flow, all of other additional terms except for turbulent stress term are in the energy equation.
 However, most turbulence modeling focuses on the Reynolds stress terms which are either solved directly or defined via a constitutive relation.
-Less attention is typically given to the other terms that need to be modeled.
+Less attention is typically given to the other terms.
 Most commonly, a Reynolds analogy is used to model the turbulent heat flux.
@@
 \end{equation}
-\subsubsection{Discretized governing Equation with $SPH-\varepsilon$ turbulence model}
+\subsubsection{Discretized governing equations with $SPH-\varepsilon$ turbulence model}
 Plug discretized turbulent stress term and turbulent heat transfer term into the momentum and energy equation, we got new discretized governing equation:
@@
 It should be noted that ghost particles are not used for these interpolations.
-Assign all properties at the corresponding reflect position to that ghost particle except for velocity.
-We assign a velocity to each wall ghost particles that is the negative of the interpolated velocity of its corresponding reflection inside the fluid domain.
+Assign all properties at the corresponding reflect position to the ghost particle except for velocity.
+We assign a velocity to each wall ghost particles that is the negative of the interpolated velocity at its corresponding reflection inside the fluid domain.
 By this way, the no-slip wall boundary condition (Eq.
@@
 These wall ghost particles will serve as neighbors in momentum and energy update.
-Implementation details about this method can be found in \citep {kumar2013parallel}.
+More implementation details about this method can be found in \citep {kumar2013parallel}.
 As these wall ghost particles keep stationary, there will be no mass flux on the boundary (Eq.
@@
 The density, momentum, and internal energy of these eruption ghost particles will not be updated before they move above ground.
-As soon as they enter the atmosphere, these ghost particle will be shift to real particles and their physics properties and position will be updated based on discretized governing equations.
+As soon as they move out from erupt conduit, these ghost particle will be shift to real particles and their physics properties and position will be updated based on discretized governing equations.
 New ghost particle will be added at the bottom of the eruption conduit as these existing ghost particles move upwards.\\
 \subsubsection{Pressure boundary condition}
-An other boundary in this model is pressure outlet boundary.
+Another boundary in this model is pressure outlet boundary.
 For flow in a straight channel, it is possible to treat the exit the same as the entry with a prescribed velocity profile.
@@
 (\ref{eq:pressure_bc_p})) is more suitable for plume simulation where the outlet is open atmosphere.
-The way we impose pressure boundary condition is simple: adding several layers of pressure ghost particles surrounding the real atmosphere particles.
+The way we impose pressure boundary condition is: adding several layers of pressure ghost particles surrounding the real atmosphere particles.
 Physics properties except for velocity (pressure, density, temperature) are determined based on the elevation of particles.
@@
 Less attention has been paid to developing of more flexible data management schemes for more complicated problems.
-Motivated by techniques developed for mesh based methods, we develop a complete framework for parallelizing SPH program with MPI standard model allowing flexible and efficient data access in this paper.\\
+Motivated by techniques developed for mesh based methods, we develop a complete framework for parallelizing SPH program with MPI standard model allowing flexible and efficient data access.\\
 Any implementation of SPH code requires efficient searching and updating of neighbors during simulation.
@@
 The background grid is also used for domain decomposition in SPH.
-We refer to the elements of background grid, namely squares for two dimensions and cubic for three dimension, as buckets.
+We refer to the elements of background grid, namely squares for two dimensions and cubic for three dimensions, as buckets.
 As for the actual storage of data representing the physical quantities associated to each particle, different strategies have been adopted in existing implementations of SPH.
@@
 \ref{fig:bc_and_domain_decomp} shows a cross section view of domian decomposition.\\
-We will provide a supplementary documentation describing details about the data structure, domain decomposition, load balancing and domain adjusting.
+We will provide a supplementary documentation describing details about the data structure, domain decomposition, load balancing and domain adjusting.\\
 \begin{figure}[!t]
@@
 \end{figure}
+%
 \begin{figure}[!t]
@@
 \end{figure}
+%
 Experiments have been carried out on CCR (the computational cluster of Center for Computational Research) at Buffalo.
@@
 For example, a large-particle module (LPM) was added to ATHAM to track the paths of rocky particles (pyroclastic or tephra) within the plume and predicts where these particles fall \citep{kobs2009modeling}.
-We were also motivated by such a tendency of plume modeling to choose SPH as our numerical tool besides its ability in dealing with interface for multiple phase flows.
-As mentioned in the introduction section, SPH method has good extension features and adding of new physics and phases requires tiny modification on the code compared with mesh-based methods.
+We were also motivated by such a tendency of plume modeling to choose SPH as our numerical tool besides its ability in dealing with interface for multiphase flows.
+As mentioned in the introduction section, SPH method has good extension features and adding of new physics and phases requires tiny modification on the code compared with mesh based methods.
 The last but not the least, the dramatic development of computational power makes it possible to establish a comprehensive model.
@@
 More advanced numerical techniques, such as adaptive particle size, Godnuv-SPH, semi-explicit time advancing scheme and better data management strategies and algorithms are on our list to exploit.
-In the near future, affect of wind field will be take into account.
+In the near future, effect of wind field will be take into account.
 %In all existing 3D plume models, numerical calculations are performed on a non-uniform (vertically and horizontally stretched) grid, with different grid resolutions and domain sizes.

### 2016-10-15 Manuscripts_GMD.tex
update plume evolution section.

@@
 Corrected formulism of SPH is used to handle tensile instability and deficiency of particle distribution near the boundaries.
-We also propose feasible ways to impose velocity inlet and pressure outlet boundary conditions, both of which are scarce in traditional implementation of SPH.
-Core solver of our model is parallelized with MPI (message passing interface) obtaining good weak scalability and strong scalability.
+We also propose feasible ways to impose velocity inlet and pressure outlet boundary conditions, both of which are scarce in traditional implementations of SPH.
+The core solver of our model is parallelized with MPI (message passing interface) obtaining good weak scalability and strong scalability.
 The model is first verified by comparing velocity and concentration distribution along axis and cross section with experimental results of JPUE (jet or plume which is ejected from a nozzle into a uniform environment).
@@
 Simulation of all possible hazards with one model is difficult due to diversity among these hazards.
-Our concern in current work is the risks that volcanic ash posed on aircrafts and communities.
+Our concern in current work is the risks that volcanic ash poses on aircrafts and communities.
 During volcanic eruptions, VATDs (volcanic ash transport and dispersion models) are used to forecast the location and movement of ash clouds over hours to days.
@@
 Moreover, these input parameters are subject to change during an eruption, requiring rapid re-assignment of new parameters.
-Usually, plume models are used to provide sources terms for VATDs.
-This paper reports a new 3D (three dimensional) volcano plume model exploiting advantage of mesh-free methods in multiphase free boundary flow modeling.\\
+Usually, plume models are used to provide source terms for VATDs.
+%This paper reports a new 3D (three dimensional) volcano plume model exploiting advantage of mesh-free methods in multiphase free boundary flow modeling.\\
 %
@@
 Even though determination of essential parameters such as the entrainment is not based on first principles, such simple models nevertheless allow us to investigate the importance of physical mechanisms in a volcanic plume.
-In addition, these simplified models requires little computational sources and can run on standard personal computers or on web sites in very short time.
+In addition, these simplified models requires little computational resource and can run on standard personal computers or on web sites in very short time.
 As a result, 1D software for volcano plume development (such as \citep{267, 1194, 3541}), combined with VATDs (such as \citep{114, draxler2015hysplit}) are widely used in research and practice.
@@
 Jets that lose their momentum before becoming buoyant collapse back onto the ground and transform into pyroclastic flows, surges and ignimbrites.
-Simutaneously, relative coarse particle might separate with main stream of the plume falling down onto the ground and may be re-entrained into the plume at a lower height \citep{ernst1996sedimentation}.
+When plume rises up, relative coarse particle might separate with main stream of the plume falling down onto the ground and probably be re-entrained into the plume at a lower height \citep{ernst1996sedimentation}.
 Within this process, erupted vapor will condense to liquid (droplet) and even further to ice.
 Latent heat released from phase change of erupted vapor will further heat up entrained air and make the bulk density further dilute.
-Even entrained vapor might also experience a similar process and impose impact on plume development.
+The entrained vapor might also experience a similar process and impose impact on plume development.
 Particle aggregation process \citep{carey1982influence,taddeucci2011aggregation}, either due to presence of liquid water, resulting from particle collision or driven by electrostatic forces might occur inside plume and thereby affect the sedimentation.
-All in all, the process of plume development is  essentially a multiphase turbulent mixing process coupled with heat transfer and many other microphysics and chemical reaction.
+All in all, the process of plume development is  essentially a multiphase turbulent mixing process coupled with heat transfer and many other microphysics and chemical reaction.\\
 As an initial effort on exploiting the feasibility and advantage of SPH in plume modeling, our model is designed to describe an injection of well mixed solid and volcanic gas from a circular vent above a flat surface into a stratified stationary atmosphere following SK-3D \citep{suzuki2005numerical}.
@@
 As a result, there is only one vector momentum equation for both phases (drag force term will not show up with this assumption).
-Because of above assumptions, all other micro-physics process (like phase change of  \texorpdfstring{H\textsubscript{2}O}, aggregation, decomposition, absorption of gas on the surface of solid, solution of gas into the liquid) and chemical process will not be considered in this model.
+\\
+Because of above assumptions, all other micro-physics process (like phase change of  \texorpdfstring{H\textsubscript{2}O}, aggregation, decomposition, absorption of gas on the surface of solid, solution of gas into the liquid) and chemical process is not be considered in this model.
 These ignored microphysics factors will come to play critical roles under particular eruption conditions.
@@
 \subsubsection{Global and local variables}
+\begin{figure}
+\center
+\includegraphics[width=5cm]{Temp}
+\caption{Temperature as a function of height}
+\label{fig:strong_local_temp}
+\end{figure}
+\begin{figure}
+\center
+\includegraphics[width=5cm]{density_strong}
+\caption{Density of the strong plume without wind after reaching its top height}
+\label{fig:strong_local_density}
+\end{figure}
+%\begin{figure}
+%\center
+%\includegraphics[width=5cm]{velocity_strong}
+%\caption{Velocity of the strong plume without wind after reaching its top height}
+%\label{fig:strong_local_velocity}
+%\end{figure}
 One of the key global quantities of great interest is the altitude to which the plume rises.
@@
 We first project simulation results (on disordered particles) onto a pre-defined grid before doing time average and spatial integration.
-The interpolation method is the basic SPH kernel based interpolation.
-\\
-\begin{figure}
-\center
-\includegraphics[width=5cm]{Temp}
-\caption{Temperature as a function of height}
-\label{fig:strong_local_temp}
-\end{figure}
-\begin{figure}
-\center
-\includegraphics[width=5cm]{density_strong}
-\caption{Density of the strong plume without wind after reaching its top height}
-\label{fig:strong_local_density}
-\end{figure}
-%\begin{figure}
-%\center
-%\includegraphics[width=5cm]{velocity_strong}
-%\caption{Velocity of the strong plume without wind after reaching its top height}
-%\label{fig:strong_local_velocity}
-%\end{figure}
+The interpolation method is the basic SPH kernel based interpolation.\\
 The profiles of local variables match well with simulation results of existing 3D models in general sense. The basic phenomena in volcano plume development is correctly captured by our model.
@@
 \label{fig:strong_plume_mass_fraction}
+\end{figure}
+%
+%
+\begin{figure}
+\center
+\includegraphics[width=5cm]{radius_strong}
+\caption{Radius of the strong plume without wind after reaching its top height}
+\label{fig:strong_local_radius}
 \end{figure}
@@
 \end{equation}
-%
-\begin{figure}
-\center
-\includegraphics[width=5cm]{radius_strong}
-\caption{Radius of the strong plume without wind after reaching its top height}
-\label{fig:strong_local_radius}
-\end{figure}
-%
 With more cool air entrained into the plume and mixing with hot erupted material, the temperature of the plume decreases as the height increase as shown in Fig.
@@
 \center
-\includegraphics[width=5cm]{t15}
-\caption{Contour of mass fraction and velocity quiver at 15 seconds, column rises up quickly due to initial momentum.}
-\label{fig:strong_t15}
-\end{figure}
+\includegraphics[width=15cm]{strong_elevation}
+\caption{This figure shows the dynamic evolution process of the Pinatubo with time.
+At the very beginning, it is driven by initial momentum and rises up quickly, mixing and entraining surround air.
+Around 100s, the jet stage finishes and it becomes a plume as there were enough air entrained.
+The plume keeps rising up until reaches the top height then starts spreading.}
+\label{fig:strong_elevation}
+\end{figure}
+
+%\begin{figure}
+%\center
+%\includegraphics[width=5cm]{t15}
+%\caption{Contour of mass fraction and velocity quiver at 15 seconds, column rises up quickly due to initial momentum.}
+%\label{fig:strong_t15}
+%\end{figure}
+
 \begin{figure}
 \includegraphics[width=10cm]{t50}
-\caption{Contour of mass fraction and velocity quiver at 50 seconds, column keeps rising up and at the same time entraining air into the plume at the outer shear region.
-The entrainment of air is shown in a zoomed view in which the velocity quiver clearly shows entrainment of air into the plume at the margin of plume.}
+\caption{Contour of mass fraction and velocity quiver at 50 seconds in plane x=0, the column keeps rising up and at the same time entraining air into the plume.
+The zoomed view (right figure) of the quiver plot clearly shows mechanism of air entrainment.
+The contour levels of mass fraction are 0.7, 0.5, 0.3, 0.1, $1\times 10 ^{-1}$, $1\times 10 ^{-2}$, $1\times 10 ^{-3}$, $1\times 10^{-4}$ and $1\times 10^{-5}$.}
 \label{fig:strong_plume_t50}
 \end{figure}
-\begin{figure}
-\center
-{
-\includegraphics[width=10cm]{t100}
-\caption{Contour of mass fraction and velocity quiver at 100 seconds, it keeps rising up and entraining air. The entrainment of air is clearly shown in zoomed view}
-\label{fig:strong_plume_t100}
-}
-\end{figure}
-\begin{figure}
-\center
-\includegraphics[width=5cm]{t200}
-\caption{Contour of mass fraction at 200 seconds, plume reaches its top height}
-\label{fig:strong_t200}
-\end{figure}
-\begin{figure}
-\center
-\includegraphics[width=5cm]{t300}
-\caption{Contour of mass fraction at 300 seconds, the cloud is spreading}
-\label{fig:strong_t300}
-\end{figure}
-Figure \ref{fig:strong_t15}~\ref{fig:strong_t300} shows the dynamic evolution process of volcano plume.
-The contour level of mass fraction are 0.7, 0.5, 0.3, 0.1, $1\times 10 ^{-1}$, $1\times 10 ^{-2}$, $1\times 10 ^{-3}$, $1\times 10^{-4}$ and $1\times 10^{-5}$.
-The plume keeps rising up and entraining air into the plume until it reaches its top height(around t=200 seconds).
+%\begin{figure}
+%\center
+%{
+%\includegraphics[width=10cm]{t100}
+%\caption{Contour of mass fraction and velocity quiver at 100 seconds. The entrainment of air is clearly shown in the zoomed view (right figure). }
+%\label{fig:strong_plume_t100}
+%}
+%\end{figure}
+%\begin{figure}
+%\center
+%\includegraphics[width=5cm]{t200}
+%\caption{Contour of mass fraction at 200 seconds, plume reaches its top height}
+%\label{fig:strong_t200}
+%\end{figure}
+%\begin{figure}
+%\center
+%\includegraphics[width=5cm]{t300}
+%\caption{Contour of mass fraction at 300 seconds, the cloud is spreading}
+%\label{fig:strong_t300}
+%\end{figure}
+Figure \ref{fig:strong_elevation} shows the overall evolution process of plume.
+Plume keeps rising up and entraining air into the plume until it reaches its top height(around t=300 seconds).
 Afterwards the mixtures of erupted material and air then fall back to neutral height and starts spreading radially.
-Detailed velocity quiver plots show entraining of air into the plume at the edge of column.
+Detailed velocity quiver plot in Fig.
+\ref{fig:strong_plume_t50} shows entraining of air into the plume at the edge of column.
 The evolution of the plume further verifies our model.
@@
 %\appendixtables   %% needs to be added in front of appendix tables in one-column style (\documentclass[acp, manuscript]{copernicus})
-\authorcontribution{TEXT}
+%\authorcontribution{TEXT}
 \competinginterests{Authors have the standard conflict with employers and listed sponsors}
@@
 All developers of Titan-2D, especially Dinesh Kumar who developed the GSPH version of Titan-2D, are greatly appreciated as Plume-SPH is based on their code.
-Advice and data (which is not shown in this paper) provided by Suzuki Yujiro gave us great help at the initial stage of model establishment and therefor are greatly appreciated.
-We appreciate Tomaso Esposti Ongaro for providing simulation data of PDAC (were also provided by Antonio Costa later together with other data) which helped us in doing early verification and improvement.
+Advice and data (which is not shown in this paper) given by Suzuki Yujiro gave us great help at the initial stage of model establishment and therefor are greatly appreciated.
+We appreciate Tomaso Esposti Ongaro for providing simulation data of PDAC (were also provided by Antonio Costa later together with simulation results of other plume models) which helped us in doing early verification and improvement.
 We appreciate Antonio Costa for providing simulation results of existing 3D models, which are used in verification and validation section of this paper.

### 2016-10-18 Manuscripts_GMD.tex
update1
updated according to Dr. Bursik's comments

@@
 The top height of Pinatubo eruption (15 June 1991) simulated by our model is consistent with both observation and existing 3D plume models.
-Profiles of several integrated variables are compared with these by existing 3D plum models and further verifies the correctness of our model.
+Profiles of several integrated variables are compared with these by existing 3D plume models and further verifies the correctness of our model.
 Analysis on the plume evolution process illustrated that this model is able to reproduce the basic physics of plume development.
@@
 \subsection{Volcanic ash hazards}
-The dynamics of explosive eruption clouds has been the central issue of volcanology science for a long time.
-Primary hazards associated with explosive grey volcanoes include pyroclastic density currents (flows and surges), the widespread deposition of airfall tephra, and the threats posed by volcanic ash.
-In addition, there are also other secondary hazards associated with volcanoes \citep{lockwood2013volcanoes}.
+The dynamics of explosive eruption clouds has been the central issue of volcanology science for a long time \citep{settle1978volcanic, wilson1978control}.
+The basic physics of volcanic plume has been understood for more than half a century \citep{morton1956turbulent, settle1978volcanic, wilson1978control}.
+The effects of magma type and vent conditions \citep{woods1988fluid, woods1995decompression}, atmospheric conditions \citep{ woods1993moist, sparks1997volcanic, bursik2001effect}, external surface water \citep{koyaguchi1996formation}, thermal disequilibrium, and particle fallout \citep{woods1991particle} have been assessed in increasing details based on observation, experiments and numerical studies.
+%
+Hazards associated with explosive grey volcanoes include pyroclastic density currents (flows and surges), the widespread deposition of airfall tephra, and the threats posed by volcanic ash. 
+%In addition, there are also other secondary hazards associated with volcanoes \citep{lockwood2013volcanoes}.
+%
 Simulation of all possible hazards with one model is difficult due to diversity among these hazards.
@@
 \subsection{Existing plume models}
-The basic physics of volcanic plume has been understood for more than half a century \citep{morton1956turbulent, settle1978volcanic, wilson1978control}.
-Several volcano plume models have been developed in the past few decades.
-Specific aspects of a plume have been studied by 1D (one dimensional) plume models which describe the steady state solution of a plume under idealized boundary conditions.
-The effects of magma type and vent conditions \citep{woods1988fluid, woods1995decompression}, atmospheric conditions \citep{ woods1993moist, sparks1997volcanic, bursik2001effect}, external surface water \citep{koyaguchi1996formation}, thermal disequilibrium, and particle fallout \citep{woods1991particle} have been assessed in increasing details through 1D numerical simulations.
-Recently developed 1D models (eg.
-\citep{mastin2007user, woodhouse2013interaction}) tend to account for more physics effects in one model.
+Several 1D (one dimensional) volcano plume models have been developed in the past few decades.
+Based on the most basic 1D model \citep{woods1988fluid} which only accounts for basis conservation, recently developed 1D models (eg.
+\citep{bursik2001effect, mastin2007user, degruyter2012improving, woodhouse2013interaction, devenish2013using, de2015plume, folch2016fplume, pouget2016sensitivity}) tend to account for more physics effects.
 FPLUME-1.0 \citep{folch2016fplume} accounts for wind effect, entrainment of moisture, water phase change, particle fallout and re-entrainment and even wet aggregation of ash.
@@
 Different 1D models adopt different entrainment coefficients based on specific formulation or calibration against well-documented case studies.
-The feedback from plume to atmosphere is ignored in 1D models.
+The feedback from plume to atmosphere is also ignored in 1D models.
 Even though determination of essential parameters such as the entrainment is not based on first principles, such simple models nevertheless allow us to investigate the importance of physical mechanisms in a volcanic plume.
@@
 Advanced numerical techniques, such as second order scheme and semi-explicit time upgrading, was also adopted afterwards to improve the accuracy of PDAC \citep{carcano2013semi}.
-Another 3D model, SK-3D \citep{suzuki2005numerical} is a three-dimensional time-dependent fluid dynamics model that can reproduce the entrainment process of eruption clouds with relative simpler physics but high order of accuracy and high spatial resolution, which is applicable to time-dependent phenomena in actual volcanological simulations.
+Another 3D model, SK-3D \citep{suzuki2005numerical} is a 3D time-dependent fluid dynamics model that can reproduce the entrainment process of eruption clouds with relative simpler physics but high order of accuracy and high spatial resolution, which is applicable to time-dependent phenomena in actual volcanological simulations.
 A serials of simulations based on SK-3D was conducted, including determination of the entrainment coefficients of eruption columns as a function of height \citep{suzuki2010numerical}, study the effect of wind field on entrainment coefficient \citep{suzuki20133d}, establishment of the relationship between the observable quantities of the eruption clouds and the eruption conditions at the vent \citep{suzuki2009three} (previously, such relationship is based on 1D models and greatly depends on empirical constants), investigation of the effect of the intensity of turbulence in the umbrella cloud on dispersion and sedimentation of tephra \citep{koyaguchi2009effect}.
@@
 And it currently is, by far the most widely used scheme.
-Specifically, the reason we choose SPH as the numerical method for volcanic plume simulation are as following:
+Implementations of SPH in volcanology are keeping showing up \citep{bursik2003smoothed,herault2010sph,haddad2016smoothed}.\\
+Specifically, the reasons that motivate us to choose SPH as the numerical method for volcanic plume simulation are as following:
 \begin{itemize}
+\item Investigation of mixing phenomena. 
+\item ZFE (Zone of Flow Establishment), ZEF (Zone of Established Flow) investigation and relation to column collapse and the mystery of the development of entrainment and proper characterization through numerical experimentation.
+\item Related to below, easy inclusion of particles of different sizes (phases) and investigation of detailed mechanics of sedimentation and drag force interaction in lower plume.
 \item For mesh based method, either interface tracking (Lagrangian) \citep{harlow1965numerical, wrobel1991computational, cheng1995simplified} or interface capturing (Eulerian) \citep{hirt1981volume, youngs1982time, gerlach2006comparison, gopala2008volume} methods are used to reconstruct the flow interface of free boundary flow.
@@
 The plume keeps rising up until reaches the top height then starts spreading.}
-\label{fig:strong_elevation}
-\end{figure}
-
+\label{fig:strong_evolution}
+\end{figure}
 %\begin{figure}
@@
 %\end{figure}
-Figure \ref{fig:strong_elevation} shows the overall evolution process of plume.
+Figure \ref{fig:strong_evolution} shows the overall evolution process of plume.
 Plume keeps rising up and entraining air into the plume until it reaches its top height(around t=300 seconds).
@@
 and computational techniques).
-The correctness of the code and model was verified and validated by a series of test simulations.\\
+The correctness of the code and model was verified and validated by a series of test simulations.
+Dimensionless velocity and concentration distribution across the cross-section and along the jet axis match well with experimental results of JPUE.
+Top height and integrated local variables simulated by our model are consistent with both observation and simulation results of existing 3D plume models.
+Comparison between our results with that of SK-3D implies that turbulence model plays a significant role in plume modeling.\\
 Currently existing 3D models focus on certain aspect of the volcano plume (PDAC on pyroclastic flow, ATHAM on microphysics, and SK-3D on entrainment with higher accuracy and higher order of accuracy) and hence, naturally, different assumptions were made in these models.

### 2016-12-29 Manuscripts_GMD.tex
update the strong scalability tests

@@
 And it currently is, by far the most widely used scheme.
-Implementations of SPH in volcanology are keeping showing up \citep{bursik2003smoothed,herault2010sph,haddad2016smoothed}.\\
+The first implementations of SPH in volcanology was presented by \citep{bursik2003smoothed} and afterwards, more implementations of SPH in volcanology \citep{mcdougall2004model,
+herault2010sph, prakash2011three,bilotta2015gpusph, haddad2016smoothed} has been keeping showing up. All of these implementations are for imcompressible or weakly compressible on-ground flow.\\
 Specifically, the reasons that motivate us to choose SPH as the numerical method for volcanic plume simulation are as following:

### 2017-02-18 Manuscripts_GMD.tex
Language revised

--- 2016-10-18 Manuscripts_GMD.tex
@@
 \begin{document}
+%<<<<<<< HEAD
+
+\title{Plume-SPH 1.0: A 3D dust-gas volcano plume model based on Smoothed Particle Hydrodynamics}
+
+%=======
 \title{Plume-SPH 1.0: A 3D dust-gas volcano plume model based on smoothed particle hydrodynamics}
+%>>>>>>> 0d96d9682fb484b8b159e8d09a0eb89723d3a0f8
 % \Author[affil]{given_name}{surname}
@@
 \affil[2]{Department of Geology, University at Buffalo, SUNY, New York, United State}
-\affil[3]{Department of Mathematics, University at Buffalo, SUNY, New York, United State}
+\affil[3]{Department of Material Design and Innovation, University at Buffalo, SUNY, New York, United State}
 \affil[4]{Center for Computational Research,
@@
 \runningauthor{Z. Cao et al}
-\correspondence{Abani Patra (abani.patra@gmail.com)}
+%<<<<<<< HEAD
+
+\correspondence{Abani Patra (abani@buffalo.com)}
+
+%=======
+%\correspondence{Abani Patra (abani.patra@gmail.com)}
+%>>>>>>> 0d96d9682fb484b8b159e8d09a0eb89723d3a0f8
 \received{}
@@
 \begin{abstract}
-Volcanic ash, if persists in the atmosphere, will pose great threat on aircrafts and communities.
-VATDs (Volcanic ash transport and dispersion models) usually take output of plume models as source terms.
-The accuracy of source terms is crucial for VATDs.
-All of existing 3D (three dimensional) plume models are using mesh based methods.
+%Volcanic ash, if it persists in the atmosphere, will pose great threat to aviation and communities. Therefore, accurate modeling and forecasts are crucial. 
+VATDs (Volcanic ash transport and dispersion models) used for critical volcanic ash forecasts usually take outputs of eruption plume models as source terms.
+The accuracy of source terms is crucial for forecasts from such VATDs.
+All  existing 3D (three dimensional) plume models described in the literature use mesh based methods.
 SPH (Smoothed particle hydrodynamics), as a mesh free method has several advantages over mesh based methods in modeling of multiphase free boundary flow.
-As an initial effort on exploiting the feasibility and advantage of SPH in volcano plume modeling we adopt a relative simpler model, a 3D dusty-gas dynamic model, targeting at capturing the more basic features of volcano plume.
+As an initial effort on exploiting the feasibility and advantages of SPH in volcano plume modeling we adopt a relatively simpler model, a 3D dusty-gas dynamic model, targeted at capturing the salient features of volcano plume.
 In this model, erupted material is assumed to be well mixed and represented by one phase while air is another phase.
-Immediate dynamic and thermal dynamic equilibrium between air and erupted material are assumed.
-Effect of wind is not considered.
-Several newly developed techniques in SPH are adopted and trimmed to address numerical challenges in simulating multiphase compressible turbulent flow with classical SPH method.
-The $SPH-\varepsilon$ turbulence model is adopted to capture sub-particle mixing.
+We also assume immediate dynamic and thermal dynamic equilibrium between air and erupted material and that the effect of wind is not significant.
+
+Several newly developed techniques in SPH are adopted and extended to address numerical challenges in simulating multiphase compressible turbulent flow with classical SPH method.
+The $SPH-\varepsilon$ turbulence model is adopted to capture mixing at unresolved scales.
 Heat exchange due to turbulence is then calculated with Reynolds analogy.
-Corrected formulism of SPH is used to handle tensile instability and deficiency of particle distribution near the boundaries.
+Corrected formalism of SPH is used to handle tensile instability and deficiency of particle distribution near the boundaries.
 We also propose feasible ways to impose velocity inlet and pressure outlet boundary conditions, both of which are scarce in traditional implementations of SPH.
@@
 \introduction  %% \introduction[modified heading if necessary]
+%<<<<<<< HEAD
 \subsection{Volcanic ash hazards}
-The dynamics of explosive eruption clouds has been the central issue of volcanology science for a long time \citep{settle1978volcanic, wilson1978control}.
-The basic physics of volcanic plume has been understood for more than half a century \citep{morton1956turbulent, settle1978volcanic, wilson1978control}.
+
+%The dynamics of explosive eruption clouds has been a central issue in volcanology  for a long time \citep{settle1978volcanic, wilson1978control}.
+ Primary hazards associated with explosive grey volcanoes include pyroclastic density currents (flows and surges), the widespread deposition of airfall tephra, and the threats to aviation posed by volcanic ash in the upper atmosphere
+% In addition, there are also other secondary hazards associated with volcanoes 
+\citep{lockwood2013volcanoes}.
+Simulation of all possible hazards with one model is difficult due to diversity among these hazards. Our focus here are  the risks that volcanic ash posed on aircrafts and communities. 
+During volcanic eruptions, VATDs (volcanic ash transport and dispersion models) are used to forecast the location and movement of ash clouds over hours to days.
+VATDs use eruption source parameters, such as plume height, mass eruption rate, duration, and the mass fraction distribution of erupted debris finer than about $4 \Phi$ (or $63  \mu$m), which can remain in the cloud for many hours or days.
+Observational data for such parameters are usually unavailable in the first minutes or hours after an eruption is detected.
+Moreover, these input parameters are subject to change during an eruption, requiring rapid re-assignment of new parameters.
+Usually, plume models are used to provide these source terms for VATDs and the forecast accuracy is critically dependent on these models.
+This paper reports on a new 3D (three dimensional) volcano plume model designed to exploit the advantages of mesh-free methods for 3D modeling of such plumes  that involve multi-phase free boundary flows.\\
+%=======
+The dynamics of explosive eruption clouds has been a central issue of volcanology science for a long time \citep{settle1978volcanic, wilson1978control}.
+The basic physics of volcanic plumes has been studied for more than half a century \citep{morton1956turbulent, settle1978volcanic, wilson1978control}.
 The effects of magma type and vent conditions \citep{woods1988fluid, woods1995decompression}, atmospheric conditions \citep{ woods1993moist, sparks1997volcanic, bursik2001effect}, external surface water \citep{koyaguchi1996formation}, thermal disequilibrium, and particle fallout \citep{woods1991particle} have been assessed in increasing details based on observation, experiments and numerical studies.
 %
-Hazards associated with explosive grey volcanoes include pyroclastic density currents (flows and surges), the widespread deposition of airfall tephra, and the threats posed by volcanic ash. 
+%Hazards associated with explosive grey volcanoes include pyroclastic density currents (flows and surges), the widespread deposition of airfall tephra, and the threats posed by volcanic ash. 
 %In addition, there are also other secondary hazards associated with volcanoes \citep{lockwood2013volcanoes}.
-%
-Simulation of all possible hazards with one model is difficult due to diversity among these hazards.
+%>>>>>>> 0d96d9682fb484b8b159e8d09a0eb89723d3a0f8
+%
+%Simulation of all possible hazards with one model is difficult due to diversity among these hazards.
 Our concern in current work is the risks that volcanic ash poses on aircrafts and communities.
-During volcanic eruptions, VATDs (volcanic ash transport and dispersion models) are used to forecast the location and movement of ash clouds over hours to days.
+%During volcanic eruptions, VATDs (volcanic ash transport and dispersion models) are used to forecast the location and movement of ash clouds over hours to days.
 Those VATDs use eruption source parameters, such as plume height, mass eruption rate, duration, and the mass fraction distribution of erupted debris finer than about $4 \Phi$ (or $63  \mu$m), which can remain in the cloud for many hours or days.
@@
 Several 1D (one dimensional) volcano plume models have been developed in the past few decades.
-Based on the most basic 1D model \citep{woods1988fluid} which only accounts for basis conservation, recently developed 1D models (eg.
-\citep{bursik2001effect, mastin2007user, degruyter2012improving, woodhouse2013interaction, devenish2013using, de2015plume, folch2016fplume, pouget2016sensitivity}) tend to account for more physics effects.
+Ranging from the most basic 1D model \citep{woods1988fluid} which only accounts for mass conservation to more recently developed 1D models (eg.
+\citep{bursik2001effect, mastin2007user, degruyter2012improving, woodhouse2013interaction, devenish2013using, de2015plume, folch2016fplume, pouget2016sensitivity}) which tend to account for more complex physics effects.
 FPLUME-1.0 \citep{folch2016fplume} accounts for wind effect, entrainment of moisture, water phase change, particle fallout and re-entrainment and even wet aggregation of ash.
-However, In these 1D models, the entrainment of air is evaluated based on two coefficients: entrainment coefficient for the vertical plume and the entrainment coefficient that describes the effect of wind.
+However, in these 1D models, the entrainment of air is evaluated based on two coefficients: entrainment coefficient for the vertical plume and the entrainment coefficient that describes the effect of wind.
 Different 1D models adopt different entrainment coefficients based on specific formulation or calibration against well-documented case studies.
@@
 Even though determination of essential parameters such as the entrainment is not based on first principles, such simple models nevertheless allow us to investigate the importance of physical mechanisms in a volcanic plume.
-In addition, these simplified models requires little computational resource and can run on standard personal computers or on web sites in very short time.
+In addition, these simplified models require little computational resource and can run on standard personal computers or on web sites in very short time.
 As a result, 1D software for volcano plume development (such as \citep{267, 1194, 3541}), combined with VATDs (such as \citep{114, draxler2015hysplit}) are widely used in research and practice.
 While these 1D models can generate well-matched results with 3D (three dimensional) models for weak plumes, much greater variability is observed for strong plume scenarios, especially for local variables \citep{costa2016results}.
-In addition, the need for greater skills in associated hazards forecasts with 1D models requires 2D (two dimensional) and 3D models.\\
-The development of 2D and 3D, time-dependent, and multiphase numerical models for volcano plume has provided new explanations for many features of explosive volcanism.
-One of these 3D models is PDAC (Pyroclastic Dispersion Analysis Code) \citep{neri2003multiparticle}  which is a non-equilibrium, multiphase, 3D compressible flow model.
+In addition, the need for greater skill in hazards forecasts especially where the plume model is used to generate source conditions for complex VATD models requires 2D (two dimensional) and 3D models.
+
+The development of 2D and 3D, time-dependent, and multiphase numerical models for volcanic plumes has provided new explanations for many features of explosive volcanism.
+One of the earliest of these is the 3D model PDAC (Pyroclastic Dispersion Analysis Code) \citep{neri2003multiparticle}  which is a non-equilibrium, multiphase, 3D compressible flow model.
 Conservation equations for each phase are solved separately with finite volume method.
-And a parallel computing version of PDAC was developed lately \citep{ongaro2007parallel}.
-Advanced numerical techniques, such as second order scheme and semi-explicit time upgrading, was also adopted afterwards to improve the accuracy of PDAC \citep{carcano2013semi}.
-Another 3D model, SK-3D \citep{suzuki2005numerical} is a 3D time-dependent fluid dynamics model that can reproduce the entrainment process of eruption clouds with relative simpler physics but high order of accuracy and high spatial resolution, which is applicable to time-dependent phenomena in actual volcanological simulations.
-A serials of simulations based on SK-3D was conducted, including determination of the entrainment coefficients of eruption columns as a function of height \citep{suzuki2010numerical}, study the effect of wind field on entrainment coefficient \citep{suzuki20133d}, establishment of the relationship between the observable quantities of the eruption clouds and the eruption conditions at the vent \citep{suzuki2009three} (previously, such relationship is based on 1D models and greatly depends on empirical constants), investigation of the effect of the intensity of turbulence in the umbrella cloud on dispersion and sedimentation of tephra \citep{koyaguchi2009effect}.
-While the SK-3D focuses on accurately capturing the entrainment caused by turbulent mixture with higher resolution and numerical method of higher order, PADC takes the disequilibrium between different phases into account and hence is a true multiphase model.
+A parallel computing version of PDAC was developed lately \citep{ongaro2007parallel}.
+Advanced numerical techniques, such as second order scheme and semi-explicit time stepping, was also adopted afterwards to improve the accuracy of PDAC \citep{carcano2013semi}.
+Another 3D model, SK-3D \citep{suzuki2005numerical} is a 3D time-dependent fluid dynamics model that attempts to reproduce the entrainment process of eruption clouds with relatively simpler physics but with high order numerical accuracy and high spatial resolution.
+%which is applicable to time-dependent phenomena in actual volcanological simulations. 
+A series of simulations based on SK-3D was reported, including determination of the entrainment coefficients of eruption columns as a function of height \citep{suzuki2010numerical},  the effect of wind field on entrainment coefficient \citep{suzuki20133d}, establishment of the relationship between the observable quantities of the eruption clouds and the eruption conditions at the vent \citep{suzuki2009three} (previously, such relationship is based on 1D models and greatly depends on empirical constants), investigation of the effect of the intensity of turbulence in the umbrella cloud on dispersion and sedimentation of tephra \citep{koyaguchi2009effect}.
+While the SK-3D focuses on accurately capturing the entrainment caused by turbulent mixture with higher resolution and numerical method of higher order, PDAC takes the disequilibrium between different phases into account and hence is a true multiphase model.
 Another 3D model, ATHAM ( Active Tracer High-Resolution Atmospheric Model) \citep{oberhuber1998volcanic} focuses more on microphysics within the plume.
@@
 However, the resolution of ATHAM is still pretty coarse compared with SK-3D and PDAC.
-Besides strengthening their strength (ATHAM has been adding more and more microphysics, PDAC was extended to consider more phases), these models are learning from each other.
-PDAC is trying to include effect of microphysics into the model while ATHAM wants to extend its ability to modeling pyroclastic flow.
+Besides adding to their special strengths (ATHAM has been adding more and more microphysics, PDAC was extended to consider more phases), these models are learning from each other.
+PDAC developers have moved to include the effect of microphysics into the model while ATHAM developer have extended its ability to modeling pyroclastic flow.
 Both are using finer and finer resolution.
-Recently, a first order, non-equilibrium compressible 3D model, ASHEE \citep{cerminara2016ashee}, is derived based on three dimensional N phases Eulerian transportation equation.
-ASHEE is valid for low concentration and low Strokes number region and much faster than N-phase Eulerian model.
-The model is based an open source numerical solver OpenFOAM adopting unstructured Finite Volume method.\\
-To summarize, each 3D model has its own focus based on problem of interest.
-Different simplification was made based on their scope.
-Improvements have been making since their birth.
-Accuracy (depends on comprehensiveness of the model, resolution of discretization, numerical error, and order of accuracy) and simulation time (depends on number of governing equations, resolution, numerical methods and parallel techniques) are always a couple of conflict considerations in 3D plume simulations.\\
+Recently, a first order, non-equilibrium compressible 3D model, ASHEE \citep{cerminara2016ashee}, was introduced based on three dimensional N phases Eulerian transportation equation.
+ASHEE is valid for low concentration and low Stokes number region and much faster than N-phase Eulerian model.
+The model is based an open source numerical solver OpenFOAM adapting its unstructured Finite Volume solver.\\
+To summarize, each 3D model has its own focus based on problem of interest and modeling/numerical choices made.
+ %Different simplification was made based on their scope. Improvements have been making since their birth. 
+ Accuracy of simulation and forecast skill (depends on comprehensiveness of the model, resolution of discretization, numerical error, and order of accuracy) and simulation time (depends on number of governing equations, resolution, numerical methods and parallel techniques) .%are always a couple of conflict considerations in 3D plume simulations.\\
 %
 \subsection{Features of SPH}
-All of these existing 3D models are based on mesh based methods (Eulerian methods).
-And there is no 3D plume model based on mesh free method (Lagrange methods) yet.
-Actually, Lagrangian methods have several features that are suitable for volcano plume simulation.
-Among such Lagrangian methods, Smoothed Particle Hydrodynamics \citep{gingold1977smoothed,lucy1977numerical}, even though relatively newer compared to well designed mesh based methods such as Finite Difference, has shown good agreement with experiments by many applications in fluid dynamic area.
-And it currently is, by far the most widely used scheme.
-Implementations of SPH in volcanology are keeping showing up \citep{bursik2003smoothed,herault2010sph,haddad2016smoothed}.\\
-Specifically, the reasons that motivate us to choose SPH as the numerical method for volcanic plume simulation are as following:
+All of these existing 3D models are mesh based Eulerian methods.
+To date there are no 3D plume models based on mesh free Lagrangian methods.
+Lagrangian methods have several features that are suitable for volcano plume simulation that we outline below.
+Among such Lagrangian methods, Smoothed Particle Hydrodynamics \citep{gingold1977smoothed,lucy1977numerical}, %even though relatively newer compared to well designed mesh based methods such as Finite Difference,
+ based simulations have shown good agreement with experiments for many applications in fluid dynamics.
+And it currently is, by far the most widely used meshfree scheme.
+Implementations of %SPH in volcanology  may be seen in \citep{bursik2003smoothed,herault2010sph,haddad2016smoothed}.\\
+Specifically, we choose SPH as the numerical method for volcanic plume simulation to enable:
 \begin{itemize}
-\item Investigation of mixing phenomena. 
-\item ZFE (Zone of Flow Establishment), ZEF (Zone of Established Flow) investigation and relation to column collapse and the mystery of the development of entrainment and proper characterization through numerical experimentation.
-\item Related to below, easy inclusion of particles of different sizes (phases) and investigation of detailed mechanics of sedimentation and drag force interaction in lower plume.
+\item better investigation of mixing phenomena;
+\item accurate modeling of the development of ZFE (Zone of Flow Establishment), ZEF (Zone of Established Flow) investigation and relation to column collapse and the questions relating to the development of entrainment; %and proper characterization through numerical experimentation.
+and,
+\item  easy inclusion of particles of different sizes (phases) and investigation of detailed mechanics of sedimentation and drag force interaction in lower plume.
+\end{itemize}
+These  will be enabled by the following features of SPH:
+\begin{itemize}
 \item For mesh based method, either interface tracking (Lagrangian) \citep{harlow1965numerical, wrobel1991computational, cheng1995simplified} or interface capturing (Eulerian) \citep{hirt1981volume, youngs1982time, gerlach2006comparison, gopala2008volume} methods are used to reconstruct the flow interface of free boundary flow.
-High computational cost, a tendency to form numerical instabilities and the inability to track complex topological changes are the significant drawbacks of Lagrangian techniques \citep{hirt1981volume, unverdi1992front, anderson1998diffuse}.
-For interface capturing (Eulerian) method, the surrendering of surface detail before the phase transport calculation means that interface reconstruction is required between time steps to recover the interface information, which need additional numerical effort \citep{hirt1981volume, youngs1982time}.
-Being able to adaptively adjust the discretization and automatically construct the interface, SPH does not require additional numerical effort for interface construction and therefor more suitable for volcano plume simulation.
+High computational cost, a tendency to form numerical instabilities and the inability to track complex topological changes are the significant drawbacks of tracking techniques \citep{hirt1981volume, unverdi1992front, anderson1998diffuse}.
+For interface capturing (Eulerian) method, the surrendering of surface detail before the phase transport calculation means that interface reconstruction is required between time steps to recover the interface information, which needs additional numerical effort \citep{hirt1981volume, youngs1982time}.
+Since SPH is able to adaptively adjust the discretization and automatically construct the interface, SPH does not require additional numerical effort for interface construction and therefore is more suitable for volcano plume simulation.
 \item Advection term in the Navier-Stokes equations does not appear explicitly in discretized formula of SPH (as illustrated in Eq.
@@
 Adding of new phases to the model also does not require a major overhaul and re-tooling.
-As will be shown in later paragraph, adding of new phases will lead to adding of several lines into the source code for new phases and additional interaction terms between existing phases and newly added phases.
+As will be shown in later paragraphs, adding of new phases will lead to adding of several lines into the source code for new phases and additional interaction terms between existing phases and newly added phases.
 \end{itemize}
@@
 As volcanic plume are in nature multiphase and without pre-defined boundary in the atmosphere, SPH is a suitable numerical method for plume modeling.
-The more basic physics, such as entrainment of air and thermal expansion, are essential for all plume modeling while some other physics, such as water condense and aggregation ect., are important in specific scenario.
-As an initial effort on exploiting advantages of SPH in volcano plume modeling, we focus on capturing more basic features in plume development.
-So our current model is far away from comprehensive.
-Fortunately, the easy extension feature of SPH facilitate developing of a more comprehensive model in the future.\\
-%
-\subsection{Our contribution}
-As SPH is a relative new method in computational fluid dynamics, implementation of SPH for compressible multiphase turbulent flow is scarce.
-Several remedies of classical SPH are adopted in our implementation.
-One of the most popular implementations of SPH is simulating of free surface flow, such as breaking-waves and flood.
+The core physics, such as entrainment of air and thermal expansion, are essential for all plume modeling while some other physics, such as water condensation and aggregation etc., are important in specific scenario.
+As an initial effort on exploiting advantages of SPH in volcano plume modeling, we focus on capturing basic features in plume development using a robust numerically and comuptationaly efficient framework with support for scalable parallel computing.
+%So our current model is far away from comprehensive.
+Open source availability and  the easy extension feature of SPH will facilitate developing of a more comprehensive community driven model.\\
+%
+\subsection{Our contributions}
+As SPH is a relatively new method in computational fluid dynamics, implementations of SPH for compressible multiphase turbulent flows are few.
+Several issues endemic to classical SPH necessary for use in this domain are fixed in our implementation.
+The most popular applications of SPH (and their original motivating application) has been in simulating of free surface flow, such as breaking-waves and flood.
 Less attention was paid to velocity inlet and pressure outlet boundary conditions which are required in plume modeling.
-We impose pressure boundary condition by adding extra layers of static ghost particles.
-Additional constrain on time step is used to avoid growing up of numerical fluctuations near the pressure boundary.
+\begin{itemize}
+\item We develop methodology to impose pressure boundary conditions by adding extra layers of static ghost particles.
+Additional constraints on time step is used to avoid the growth of numerical fluctuations near the pressure boundary.
 We impose velocity inlet boundary condition by placing several layers of ghost particles moving with eruption velocity.
-Turbulence model is crucial for reproducing the entrainment of air.
+\item Turbulence model is crucial for reproducing the entrainment of air.
 There are several turbulence models proposed for SPH method \citep{issa2005numerical,violeau2007numerical}.
-We adopt a LANS (Lagrangian Averaged Navier-Stokes) turbulence model \citep{monaghan2011turbulence}, which was originally proposed for incompressible, and extend it for compressible flow accounting for turbulent heat exchange.
-Corrected formulism of SPH is adopted to bypass the well-known tensile instability issues of classical SPH.\\
-%
-Simulation of volcano plume with acceptable accuracy requires fine resolutions (very high particle counts) that cannot be accomplished without parallel computing within given time window.
+We adopt a LANS (Lagrangian Averaged Navier-Stokes) turbulence model \citep{monaghan2011turbulence}, which was originally proposed for incompressible, and is extended here for compressible flow accounting for turbulent heat exchange.
+\item Corrected formulation of SPH \citep{kumar2013parallel} is adapted to bypass the well-known tensile instability issues of classical SPH.\\
+%
+\item Simulation of volcanic plumes with acceptable accuracy requires fine resolution (very high particle counts) that cannot be accomplished without parallel computing.
 The core solver of our model is parallelized by MPI (message passing interface standard) parallelism.
 In addition, a dynamic load balancing strategy is also developed.
-What's more, imposing of some types of boundary conditions (such as eruption boundary condition) requires dynamically adding and removing of particles during simulation.
+\item What's more, imposing of some types of boundary conditions (such as eruption boundary condition) requires dynamically adding and removing of particles during simulation.
 To address this issue, we adopt an efficient data management scheme based on time dependent SFC (space filling curve) and hash table.
-The computational cost is further reduced by adjusting simulation domain adaptively.\\
-The organization of this paper is as following:
-The physics model of the plume is first presented in section \ref{sec:physics-model} and leads to a complete mathematical description of the volcano plume (governing equations and boundary conditions).
+The computational cost is further reduced by adjusting simulation domain adaptively.
+\end{itemize}
+
+The physics model of the plume is first presented in section \ref{sec:physics-model} and leads to a complete mathematical description of the volcanic plume (governing equations and boundary conditions).
 In section \ref{sec:SPH_method} we will briefly introduce the numerical tool, SPH method.
-Both the fundamental discretization formulism and techniques that used to handle specific issues involved in plume modeling will be discussed.
+Both the fundamental discretization formulation and techniques that are used to handle specific issues involved in plume modeling will be discussed.
 Verification and validation with numerical tests is presented in section \ref{sec:verification-validation}.
 In section \ref{sec:conclusion}, a discussion on future work is given following a brief summary.
+
 \section{Physical Model} \label{sec:physics-model}
@@
 Jets that lose their momentum before becoming buoyant collapse back onto the ground and transform into pyroclastic flows, surges and ignimbrites.
-When plume rises up, relative coarse particle might separate with main stream of the plume falling down onto the ground and probably be re-entrained into the plume at a lower height \citep{ernst1996sedimentation}.
+When plume rises up, relatively coarse particle might separate from main stream of the plume falling down onto the ground and probably be re-entrained into the plume at a lower height \citep{ernst1996sedimentation}.
 Within this process, erupted vapor will condense to liquid (droplet) and even further to ice.
@@
 Particle aggregation process \citep{carey1982influence,taddeucci2011aggregation}, either due to presence of liquid water, resulting from particle collision or driven by electrostatic forces might occur inside plume and thereby affect the sedimentation.
-All in all, the process of plume development is  essentially a multiphase turbulent mixing process coupled with heat transfer and many other microphysics and chemical reaction.\\
+All in all, the process of plume development is  essentially a multiphase turbulent mixing process coupled with heat transfer and many other microphysics and chemical reactions.
+
 As an initial effort on exploiting the feasibility and advantage of SPH in plume modeling, our model is designed to describe an injection of well mixed solid and volcanic gas from a circular vent above a flat surface into a stratified stationary atmosphere following SK-3D \citep{suzuki2005numerical}.
@@
 As a result, there is only one vector momentum equation for both phases (drag force term will not show up with this assumption).
-\\
-Because of above assumptions, all other micro-physics process (like phase change of  \texorpdfstring{H\textsubscript{2}O}, aggregation, decomposition, absorption of gas on the surface of solid, solution of gas into the liquid) and chemical process is not be considered in this model.
+
+Because of above assumptions, all other micro-physics processes (like phase change of  \texorpdfstring{H\textsubscript{2}O}, aggregation, decomposition, absorption of gas on the surface of solid, solution of gas into the liquid) and chemical process are not considered in this model.
 These ignored microphysics factors will come to play critical roles under particular eruption conditions.
@@
 Introducing of wind effect in our model will need imposing of pressure boundary conditions in a dynamic way, which requires more numerical effort and algorithm design.
-To summarize, our model is not valid for eruptions where wind effect plays a significant role in its development, usually refereed as weak plume.
-Our model is also lack ability in modeling plumes with large particles or eruptions in which microphysics plays non-ignorable roles, such as a eruption of \textit{El Chich{\'o}n} volcano at April 4th \citep{sigurdsson19841982, folch2016fplume}.
-At present stage, we are more interested in more basic (and more critical) aspects of volcano plume and therefor devote our effort to this relative simpler model.
-It is worthwhile to mention here that because SPH is adopted as our numerical method, adding of these physics into our model would require much less work in terms of programming compared to mesh based methods.\\
+To summarize, our model is not valid for eruptions where wind effect plays a significant role in its development, usually referred as weak plume.
+Our model also lacks the ability in modeling plumes with large particles or eruptions in which microphysics plays non-ignorable roles, such as a eruption of \textit{El Chich{\'o}n} volcano at April 4th \citep{sigurdsson19841982, folch2016fplume}.
+We are focussed here on developing the SPH based methodoloy in the context of the more basic (and more critical) aspects of volcano plume and therefore devote our effort to this relative simpler model.
+It is worthwhile to mention here that because SPH is adopted as our numerical method, adding of these physics into our model would require much less work in terms of programming compared to mesh based methods.
+Since our plan is an open source distribution of the tool we believe some of these enhancements will rapidly ensue with community participation.
+
 \subsection{Governing equations}
@@
 For SPH, governing equations in Lagrange form is needed.
-By deducting Kinetic energy from energy equation, subtracting mass conservation from momentum equation, combining transient term and advection term into material derivative term, the governing equation is put into the final formulism, in which advection term does not show up explicitly.
+By deducting kinetic energy from energy equation, subtracting mass conservation from momentum equation, combining transient term and advection term into material derivative term, the governing equation is put into the final form, in which advection term does not show up explicitly.
 \begin{align}
@@
 Velocity is zero for non-slip wall boundary.
-If assume the boundary to be adiabatic, heat flux should be zero on the boundary.
+If we assume the boundary to be adiabatic, heat flux should be zero on the boundary.
 The flux of mass should also be zero.
 As a result, internal energy flux, which consists of heat flux and energy flux carried by mass flux, will vanish on the wall boundary.
-Equation (\ref{eq:wall_bc_rho}) ~ (\ref{eq:wall_bc_e}) gives no-slip wall boundary condition wrote in terms of primitive variables.
+Equation (\ref{eq:wall_bc_rho}) ~ (\ref{eq:wall_bc_e}) gives no-slip wall boundary condition written in terms of primitive variables.
 \begin{align}
@@
 The pressure of the surrounding atmosphere is given.
-Except for the pressure, boundary conditions for density, velocity, and energy on the outlet should depend on the solution.
-As we ignore the viscosity, the shear stress is gone and norm stress (whose magnitude equals to pressure) will balance with ambient pressure.
+Except for the pressure, boundary values for density, velocity, and energy on the outlet should depend on the solution.
+As we ignore the viscosity, the shear stress is ignored and normal stress (whose magnitude equals to pressure) will balance the ambient pressure.
 \begin{equation}
@@
 SPH is a mesh free Lagragian method.
-In SPH, the domain is discretized as a set of particles or discretization points and the position of each particle (discretization point) is updated at every time step.
+In SPH, the domain (control mass) is discretized by a set of particles or discretization points and the position of each particle (discretization point) is updated at every time step based on the motion computed.
 Approximation of all field variables (velocity, density and pressure, ect.) is obtained by interpolation based on discretization points.
 The physical laws (such as conservation laws of mass, momentum and energy) written in the form of PDEs (partial differential equations)  or ODEs (ordinary differential equations) need to be transformed into the Lagrangian particle formalism of SPH.
-Using a kernel function that provides the weighted estimation of the field variables at any point, the integral equations are evaluated as sums over neighbor particles.
+Using a kernel function that provides the weighted estimation of the field variables at any point, the integral equations are estimated as sums over particles in a compact subdomain defined by the support of the kernel function defined with respect to the discretization points.
 Thus, field variables associated to the particle are updated based on its neighbors.
-The kernel function has a compact support determined by smoothing length of each particle.
+Each kernel function has a compact support determined by smoothing length of each particle.
 There are several review papers \citep{monaghan1992smoothed, monaghan2005smoothed, rosswog2009astrophysical, price2012smoothed, monaghan2012smoothed}, giving a pretty comprehensive view over SPH, we will only refer here to the representation of the constitutive equations in SPH and put more focus on specific numerical techniques for plume modeling.
+
 \subsection{Fundamental principle}
-There are several routines for discretizing governing equations (PDEs or ODEs) with SPH.
-There is slight difference between discretized formulism obtained by different routines.
-We present here one of them following JJ Monaghan \citep{monaghan1992smoothed, monaghan2005smoothed}.
-The start point of approximating a function with SPH is one translation property of Dirac function.
+There are several procedures for discretizing governing equations (PDEs or ODEs) with SPH.
+There is slight difference between discretized formalism obtained by different routes.
+We present here one of them following Monaghan \citep{monaghan1992smoothed, monaghan2005smoothed}.
+The starting point of approximating a function with SPH is the translation property of the Dirac function.
 \begin{equation}
@@
 Recall that $d\textbf{x}\prime = \dfrac{dm \prime}{\rho \prime}$, the integration equation, Eq.
-(\ref{eq:SPH-fundamental-principle}), can be approximated by summation and lead to an approximation of the function at particle (discretization point) a:
+(\ref{eq:SPH-fundamental-principle}), can be approximated by summation and lead to an approximation of the function $A$:
 \begin{equation}
@@
 \end{equation}
-Where the summation is over all the particles within the region of compact support of the weighting function. 
+where the summation is over all the particles within the region of compact support of the weighting function. 
 Gradient terms may be straightforwardly calculated by taking the derivative of Eq. (\ref{eq:SPH-approximation-sum}), giving
@@
 %
+
 \subsection{Weighting function}
@@
 Generally, the accuracy increases with the order of the polynomials of the kernel function, but the computational time also increases as number of interactions increase.
-We are adopting a truncated Gaussian function as the weighting function in our simulation.
+We are adopting a truncated Gaussian function in our simulation.
 \begin{equation}
@@
 \end{equation}
-\subsection{Discretization of governing equations}
+
+\subsection{Discretization of governing equations and extensibility}
 The basic interpolation given in Eq.
 (\ref{eq:SPH-approximation-sum}) to Eq.
-(\ref{eq:SPH-vecctor-function}) provide a general way of obtaining SPH expressions of governing equations.
+(\ref{eq:SPH-vecctor-function}) provides a general way to obtaining SPH expressions of governing equations.
 The problem is that using these expressions ``{as is}" in general leads to quite poor gradient estimates.
@@
 Special treatments are also need for second order derivative terms.
-We only refer here one of these possible discretization forms of compressible Euler equations with SPH:
+We only refer here to one of these possible discretizations of compressible Euler equations with SPH:
 \begin{align}
-<\rho_a> = \sum m_b w_{ab} (h_a) \label{eq:ns-sph-d} \\
+<\rho_a> = \sum_b m_b w_{ab} (h) \label{eq:ns-sph-d} \\
 <\dfrac{d \textbf{v}_a}{d t}>= \sum_b m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}) \nabla_a w_{a b}(h) +\textbf{g} \label{eq:ns-sph-v} \\
@@
 \end{align}
-Where, $\textbf{v}_{a b} = \textbf{v}_a - \textbf{v}_b$. $\Pi$ is artificial viscosity term, which will be discussed in section \ref{sec:artificial-viscosity}.
+where, $a$ is the SPH particle index, $\textbf{v}_{a b} = \textbf{v}_a - \textbf{v}_b$. $\Pi$ is an artificial viscosity term, which will be discussed in section \ref{sec:artificial-viscosity}.
 However, these discretized formulations will not guarantee conservation of entropy.
-We can obtain conservation of entropy by giving up conservation of mass or thermal energy adopting an alternative discretized formulations \citep{monaghan1992smoothed}.
-As a Lagrangian method, particle position will also been update at every time step.
+We can obtain conservation of entropy by giving up on conservation of mass or thermal energy adopting  alternative discretized formulations \citep{monaghan1992smoothed}.
+As a Lagrangian method, particle position will also be updated at every time step.
 \begin{equation}
@@
 \end{equation}
+
+We highlight an important feature of the SPH methodology.
 Adding new physics and new phases into the model is trivial in terms of discretization.
@@
 \begin{align}
-<\rho_a> = \sum m_b w_{ab} (h_a) + \dot{\rho}(\textbf{x},t)\label{eq:ns-source-sph-d} \\
+<\rho_a> = \sum m_b w_{ab} (h) + \dot{\rho}(\textbf{x},t)\label{eq:ns-source-sph-d} \\
 <\dfrac{d \textbf{v}_a}{d t}>= \sum_b m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}) \nabla_a w_{a b}(h) +\textbf{g} + D \sum	_b m_b \dfrac{\textbf{v}_b - \textbf{v}_a}{\rho_b} \label{eq:ns-drag-sph-v} \\
@@
 \end{align}
-Where the source term $\dot{\rho}$ can be a "sink" of erupted vapor due to phase change which is considered in ATHAM \citep{oberhuber1998volcanic}.
-$D$ is drag force coefficient.
+where the source term $\dot{\rho}$ can be a "sink" of erupted vapor due to phase change like in ATHAM \citep{oberhuber1998volcanic}.
+$D$ is a drag force coefficient.
 $\kappa$ is heat conduction coefficient.
 Other physics can be added easily in a similar way.
-Adding of these new terms will lead to modification at several lines in the source code.
-The drag force term should show up only when dynamics dis-equilibrium between different phases is considered.
-In that case, each phase will need one set of governing equations of Navier-Strokes type.
-Adding of new phases into SPH code only needs to add several new lines for the new phase besides adding interaction terms.
+Adding of these new terms will lead to modification of several lines in the source code.
+The drag force term should show up only when dynamic dis-equilibrium between different phases is considered.
+In that case, each phase will need one set of governing equations of Navier-Stokes type.
+Adding of new phases into SPH code only needs few new lines for the new phase besides adding  interaction terms.
+
 \subsection{Artificial viscosity} \label{sec:artificial-viscosity}
@@
 As in the case of first derivatives, second derivatives can be estimated by differentiating a SPH interpolation twice.
-However, such formulism has two disadvantages: First, it is very sensitive to particle disorder.
-Second, the second derivative of the kernel can change sign and will lead to unphysical representation (viscous dissipation cause decrease of the entropy).
-One of the most commonly used form of artificial viscosity is:
+However, such a formulation has two disadvantages: First, it is very sensitive to particle disorder.
+Second, the second derivative of the kernel can change sign and will lead to unphysical representations (viscous dissipation causes decrease of the entropy).
+One of the most commonly used models of artificial viscosity is:
 \begin{equation}
@@
 \end{equation}
-Absolute(kinematic) viscous coefficient $\nu$ is defined as:
+%Absolute(kinematic) viscous%
+Coefficient $\nu$ is defined as:
 \begin{equation}
@@
 $\eta$ is usually taken as 0.1 for preventing singularities when $\textbf{x}_{ab} = 0$.
+
 \subsection{Time step}
@@
 Where $c_a$ is sound speed at particle a, $d$ is number of dimensions. First order Euler time integration, with $CFL = 0.2$ has been used in three dimensional calculations.
+
 \subsection{Tensile instability and corrected derivatives}
 The classical SPH method suffered from tensile instability and boundary deficiency.
-To address these difficulties, Chen \citep{chen1999improvement} proposed a corrective SPH method.
-For 1D case, by employing a Taylor expansion for $A(x)$ about $x_a$, multiplying both side by kernel function and then do intergration over the domain, gives
+To address these difficulties, Chen \citep{chen1999improvement} proposed a corrected SPH method.
+For 1D case, by employing a Taylor expansion for $A(x)$ about $x_a$, multiplying both sides by kernel function and then doing integration over the domain, gives
 \begin{equation}
@@
 \end{equation}
-Ignore derivative terms higher than first order, and write the intergral in particle approximation form:
+Ignoring derivative terms higher than first order, and writing the integral in particle approximation form:
 \begin{equation}
@@
 (\ref{eq:CSP-function-approximation-1d}) even though the derivation is different.
-The first order derivative can be solved explicitly or numerically from:
-\begin{equation}
-\text{\textbf{M}}_{\alpha \beta a} \textbf{A}_{\alpha a} = \textbf{F}_{\beta a}
-\end{equation}
-where
-\begin{align}
-\text{\textbf{M}}_{\alpha \beta a} = \sum_b m_b \dfrac{\textbf{x}_b^{\alpha} - \textbf{x}_a^{\alpha}}{\rho_b} \nabla_a^{\beta} w(\textbf{x}_a -\textbf{x}_b, h) \\
-\textbf{A}_{\alpha a} = \dfrac{\partial A}{\partial x^{\alpha}} (\textbf{x}_a)\\
-\textbf{F}_{\beta a} = \sum_b m_b \dfrac{A_b - A_a}{\rho_b} \nabla_a^{\beta} w(\textbf{x}_a -\textbf{x}_b, h)
-\end{align}
-$\alpha$ and $\beta$ are direcion indices, ranging from 1 to 3 for 3D problem. 
+The first order derivative can be obtained by solving a system of linear equations explicitly or numerically \citep{chen1999improvement}.
+%\begin{equation}
+%\text{\textbf{M}}_{\alpha \beta a} \textbf{A}_{\alpha a} = \textbf{F}_{\beta a}
+%\end{equation}
+%where
+%\begin{align}
+%\text{\textbf{M}}_{\alpha \beta a} = \sum_b m_b \dfrac{\textbf{x}_b^{\alpha} - \textbf{x}_a^{\alpha}}{\rho_b} \nabla_a^{\beta} w(\textbf{x}_a -\textbf{x}_b, h) \\
+%\textbf{A}_{\alpha a} = \dfrac{\partial A}{\partial x^{\alpha}} (\textbf{x}_a)\\
+%\textbf{F}_{\beta a} = \sum_b m_b \dfrac{A_b - A_a}{\rho_b} \nabla_a^{\beta} w(\textbf{x}_a -\textbf{x}_b, h)
+%\end{align}
+%$\alpha$ and $\beta$ are direcion indices, ranging from 1 to 3 for 3D problem. 
 \subsection{Multiphase SPH}
-Numerical simulation of multiphase flows is usually difficult due to the existence of interface between two fluids.
-And the interface is usually moving or deformable.
+Numerical simulation of multiphase flows is usually difficult due to the existence of complex evolving interfaces between phases.
 This present severe challenges to conventional Eulerian grid-based numerical methods.
@@
 What's worse, the engulfment process during plume development makes geometry of interface very complicated.
-SPH is able to handle multiphase problems without additional numerical effort for dealing with interface, which makes it suitable for volcano plume simulation.
+Lagrangian SPH is able to handle multiphase problems without additional numerical effort for dealing with interface by simply tagging particles of each phase, which makes it suitable for volcano plume simulation.
 Take the governing equation adopted in our model as an example.
@@
 (\ref{eq:ns-sph-d}) in spirit.
-Particles of phase 1 will not be counted while evaluate density of phase 2 and vice verse.
-Updating of density is based on the following discretized equations.
-\begin{equation}
-<\rho_{\alpha}^a>=\frac{\sum m_b w_{\alpha b} (h_{\alpha})}{\sum \frac{m_b}{\rho_b} w_{\alpha b} (h_{\alpha}) +\sum \frac{m_j}{\rho_j} w_{\alpha j} (h_{\alpha})} \label{eq:gov-sph-d1}
-\end{equation}
-\begin{equation}
-<\rho_\alpha^{sg}>=\frac{\sum_j m_j w_{\alpha j} (h_\alpha)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} (h_{\alpha}) +\sum \frac{m_j}{\rho_j} w_{\alpha j} (h_{\alpha})} \label{eq:gov-sph-d2}
+Particles of phase 1 will not be counted while evaluating density of phase 2 and vice verse.
+Updating of density is then based on the following discretized equations.
+\begin{equation}
+<\rho_{\alpha}^a>=\frac{\sum m_b w_{\alpha b} (h)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} (h) +\sum \frac{m_j}{\rho_j} w_{\alpha j} (h)} \label{eq:gov-sph-d1}
+\end{equation}
+\begin{equation}
+<\rho_\alpha^{sg}>=\frac{\sum_j m_j w_{\alpha j} (h)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} (h) +\sum \frac{m_j}{\rho_j} w_{\alpha j} (h)} \label{eq:gov-sph-d2}
 \end{equation}
@@
 \begin{equation}
-\rho_{\alpha}=\rho_{\alpha}^a=\frac{\sum m_b w_{\alpha b} (h_{\alpha})}{\sum \frac{m_b}{\rho_b} w_{\alpha b}}
+\rho_{\alpha}=\rho_{\alpha}^a=\frac{\sum m_b w_{\alpha b} (h)}{\sum \frac{m_b}{\rho_b} w_{\alpha b}}
 \end{equation}
@@
 \begin{equation}
-\rho_{\alpha}=\rho_{\alpha}^{sg}=\frac{\sum m_j w_{\alpha j} (h_{\alpha})}{\sum \frac{m_j}{\rho_j} w_{\alpha j}}
+\rho_{\alpha}=\rho_{\alpha}^{sg}=\frac{\sum m_j w_{\alpha j} (h)}{\sum \frac{m_j}{\rho_j} w_{\alpha j}}
 \end{equation}
@@
 \subsection{Turbulence modeling with SPH}
-Typically, turbulence fluctuations in a volcanic column occur on very different length scales, ranging from millimeters up to hundreds of meters.
+Typically, turbulence fluctuations in a volcanic column occurs at very different length scales, ranging from millimeters up to hundreds of meters.
 The entrainment of air and heat transfer are dominated by such turbulent fluctuations as turbulent exchange coefficient would be several magnitudes larger than molecular viscosity (and heat conduction) coefficient.
@@
 %\citep {gotoh2004sph, shao2006sph}. 
-Another way is to use extremely fine resolution (The direct numerical simulation method) at the expense of much higher computational cost that results from both increase in number of particles and decrease in time steps constrained by CFL condition.
+Another way is to use extremely fine resolution ("the direct numerical simulation method") at the expense of much higher computational cost that results from both increase in number of particles and decrease in time steps constrained by CFL condition.
 Here we choose the SPH-SPS method.\\
 Even though free surface flow are in nature turbulent, detailed research in turbulence using SPH method is rather scarce.
-Among the first papers on turbulence are the implementations of the $\alpha$ turbulence model \citep{holm1999fluctuation} by Monaghan \citep{monaghan2002sph} and the RANS (Reynolds averaged Navier-Stokes) approach by Violeau \citep{violeau2007numerical}.
+Among the first papers on turbulence are the implementations of the $\alpha$ turbulence model \citep{holm1999fluctuation}, Monaghan \citep{monaghan2002sph} and the RANS (Reynolds averaged Navier-Stokes) approach by Violeau \citep{violeau2007numerical}.
 A first wall-bounded LES (large eddy simulation) simulation with SPH was performed by Issa \citep{issa2005numerical}.
@@
 We will adopt $SPH-\varepsilon$ method in our model.
-As most of these papers constrain their focus on incompressible or weakly compressible flow.
-To simulate turbulence in plume, we need to extend existing incompressible turbulent modeling methods to compressible flow.\\
+Most of these papers constrain their focus to incompressible or weakly compressible flow.
+To simulate turbulence in plume, we will thus need to extend existing incompressible turbulent modeling methods to compressible flow.
+This implies a consideration of the evolution of densities also and reformulation to avoid any simplifications based on a single density.
+
 $SPH-\alpha$ model \citep{monaghan2002sph} is a powerful extension of the XSPH algorithm which reduces disorder at short length scales and retains the constants of the motion.
-But in that paper, energy conservation equation is discretized in a way that is exactly the same as usual SPH.
+However, in that paper, energy conservation equation is discretized in a way that is exactly the same as usual SPH.
 There is no special treatment on energy conservation equation.
-And the recent improved version \citep{monaghan2011turbulence} of $SPH-\alpha$ method focus on incompressible flow only.
+The recent improved version \citep{monaghan2011turbulence} of $SPH-\alpha$ method focuses on incompressible flow only.
 Actually, there are several additional terms in the averaged compressible Navier-Stokes equations besides the Reynolds stress term \citep{NASACompressibleTurbulence}.
 The averaged momentum equation for compressible flow are in the same form as that for incompressible flow, all of other additional terms except for turbulent stress term are in the energy equation.
-However, most turbulence modeling focuses on the Reynolds stress terms which are either solved directly or defined via a constitutive relation.
+Most turbulence modeling focuses on the Reynolds stress terms which are either solved directly or defined via a constitutive relation.
 Less attention is typically given to the other terms.
@@
 The additional terms associated with molecular diffusion and turbulent transport in the energy equation are either modeled in different ways or neglected sometimes \citep{NASACompressibleTurbulence}.
-We neglect these terms in our simulation.\\
+We neglect these terms in our simulation.
+
 \subsubsection{$SPH-\varepsilon$ turbulence model}
@@
 The discretized form of momentum equation is obtained through lengthy derivation based on Lagrangian equations.
-Derivation details and other discussions, such as conservation of momentum and energy, satisfactory of circulation theory, energy/momentum spectrum analysis based on Founrier transfer are detailed in Monaghan's paper \citep{monaghan2011turbulence}.
-Here we just have a brief summary on key steps.\\
-The smooth that Monaghan adopted is:
+Derivation details and other discussions, such as conservation of momentum and energy, satisfactory of circulation theory, energy/momentum spectrum analysis based on Fourier transfer are detailed in Monaghan's paper \citep{monaghan2011turbulence}.
+Here we just have a brief summary of key steps.\\
+The smoothing that Monaghan adopted is:
 \begin{equation}
@@
 \label{eq:replacement-in-turb-derive}
-\dfrac{G(\vert \textbf{x} _b - \textbf{x} _a \vert, l_a)}{\rho _b} \rightarrow \dfrac{K_{ab}}{M}
+\dfrac{G(\vert \textbf{x} _b - \textbf{x} _a \vert, l)}{\rho _b} \rightarrow \dfrac{K_{ab}}{M}
 \end{equation}
@@
 \label{eq:monaghan-mom-turb}
-\dfrac{d \textbf{v}_a}{dt} = -\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}) \bigtriangledown_aw_{a b}(h_a)] + \sum_b m_b \dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{M} \bigtriangledown_a K_{ab}
-\end{equation}
-Instead of using Eq.
-(\ref{eq:monaghan-mom-turb}) which is from Monaghan's paper, we make inverse replacement of Eq.
-(\ref{eq:replacement-in-turb-derive}) and notice that if $l$ is uniform and keep constant:
+\dfrac{d \textbf{v}_a}{dt} = -\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}) \bigtriangledown_aw_{a b}(h)] + \sum_b m_b \dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{M} \bigtriangledown_a K_{ab}
+\end{equation}
+Instead of using Eq. (\ref{eq:monaghan-mom-turb}) which is from Monaghan's paper, we use Eq. (\ref{eq:replacement-in-turb-derive}) and notice that if $l$ is uniform and keep constant: 
 %Please notice that the above equation is slightly different from the original equation in Monaghan's paper (Eq.
@@
 And the slight modification will not influence following derivation if $l$ is uniform and keep constant because:
+
+%{\bf WE NEED TO SAY SOMETHING ABOUT INCOMPRESSIBILITY HERE}
+
 \begin{equation}
@@
 \label{eq:SPH-mom-epsilon-turb}
-\dfrac{d \textbf{v}_a}{dt} = -\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}) \bigtriangledown_aw_{a b}(h_a)] + \sum_b m_b \Phi_{ab}\bigtriangledown_aG_{ab}(l_a)
+\dfrac{d \textbf{v}_a}{dt} = -\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}) \bigtriangledown_aw_{a b}(h)] + \sum_b m_b \Phi_{ab}\bigtriangledown_aG_{ab}(l)
 \end{equation}
@@
 \end{equation}
-and take $\varepsilon = 0.8$ as recommended by JJ Monaghan.
-\subsubsection{Turbulence heat transfer}
+and take $\varepsilon = 0.8$ as recommended by Monaghan.
+
+\subsubsection{Turbulent heat transfer}
 We adopt Reynolds analogy to get heat transfer coefficient due to turbulence.
@@
 \end{equation}
-Typical value of $Pr_t$ for air is 0.7 $\sim$ 0.9 . We take $Pr_t=0.85$ for gases as recommended by Kays, William M \citep{kays1994turbulent} from summarizing of experimental results. 
-JJ Monaghan \citep{monaghan2005smoothed} summarized the simulation of viscosity and heat conduction in his review on SPH.
+Typical value of $Pr_t$ for air is 0.7 $\sim$ 0.9 . We take $Pr_t=0.85$ for gases as recommended by Kays \citep{kays1994turbulent} from summarizing of experimental results. 
+Monaghan \citep{monaghan2005smoothed} summarized the simulation of viscosity and heat conduction in his review on SPH.
 We will refer to his summary in our following discussion.
@@
 \begin{cases} 
-      \dfrac{1}{6} & for  \quad d=3 \\
+      \dfrac{1}{10} & for  \quad d=3 \\
       \\
-     \dfrac{5}{24}  & for  \quad d=2 
+     \dfrac{1}{8}  & for  \quad d=2 
 \end{cases}
@@
 \begin{equation}
- \sum_b \dfrac{\varepsilon}{2} \dfrac{m_b}{\rho_b} \textbf{v}_{ab} \cdot \textbf{v}_{ab} \nabla_a G_{ab}(l_a)= \sum_b \dfrac{\varepsilon}{2S} m_b \dfrac{\textbf{v}_{ab}}{\rho_b} \dfrac{S \textbf{v}_{ab} \cdot \textbf{x}_{ab}}{x_{ab}^2} \dfrac{x_{ab}^2}{\textbf{x}_{ab}} \nabla_a G_{ab}(l_a)
+ \sum_b \dfrac{\varepsilon}{2} \dfrac{m_b}{\rho_b} \textbf{v}_{ab} \cdot \textbf{v}_{ab} \nabla_a G_{ab}(l)= \sum_b \dfrac{\varepsilon}{2S} m_b \dfrac{\textbf{v}_{ab}}{\rho_b} \dfrac{S \textbf{v}_{ab} \cdot \textbf{x}_{ab}}{x_{ab}^2} \dfrac{x_{ab}^2}{\textbf{x}_{ab}} \nabla_a G_{ab}(l)
 \end{equation}
@@
 \end{equation}
-Please notice that turbulent viscosity term has opposite sigh with molecular viscosity term in discretized momentum equation and there is a minus sign in the expression of $\Pi_{ab}$, they will cancelled out.
+Please note that turbulent viscosity term has opposite sign with molecular viscosity term in discretized momentum equation and there is a minus sign in the expression of $\Pi_{ab}$, they will cancelled out.
 However, the above equation is correct only for 1D situation. For 2D or 3D, it is not easy to get the explicit expression of viscosity. \\
@@
 \begin{equation}
-C_p \dfrac{dT}{dt} = \sum_b \dfrac{m_b}{\rho_a \rho_b} (\kappa_a + \kappa_b) (T_a - T_b) F_{ab} (h_a)
+C_p \dfrac{dT}{dt} = \sum_b \dfrac{m_b}{\rho_a \rho_b} (\kappa_a + \kappa_b) (T_a - T_b) F_{ab} (h)
 \end{equation}
@@
 C_p \dfrac{dT}{dt} 
-& = \sum_b \dfrac{m_b}{\rho_a \rho_b} (\kappa_a + \kappa_b) (T_a - T_b) F_{ab} (h_a) \\
- &= 2 \sum_b \dfrac{m_b}{\rho_a \rho_b} \dfrac{\overline{C_{p,ab}} \overline{\rho_{ab}} \varepsilon x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 Pr_t  S \textbf{v}_{ab} \cdot \textbf{x}_{ab} } (T_a - T_b) F_{ab} (h_a)
+& = \sum_b \dfrac{m_b}{\rho_a \rho_b} (\kappa_a + \kappa_b) (T_a - T_b) F_{ab} (h) \\
+ &= 2 \sum_b \dfrac{m_b}{\rho_a \rho_b} \dfrac{\overline{C_{p,ab}} \overline{\rho_{ab}} \varepsilon x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 Pr_t  S \textbf{v}_{ab} \cdot \textbf{x}_{ab} } (T_a - T_b) F_{ab} (h)
 \end{split}
@@
 C_p \dfrac{dT}{dt}
- =\dfrac{\varepsilon}{S  Pr_t}  \sum_b \dfrac{m_b}{\rho_a \rho_b} \dfrac{\overline{C_{p,ab}} \overline{\rho_{ab}} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\textbf{v}_{ab} \cdot \textbf{x}_{ab}} (T_a - T_b) F_{ab} (h_a)
-\end{equation}
+ =\dfrac{\varepsilon}{S  Pr_t}  \sum_b \dfrac{m_b}{\rho_a \rho_b} \dfrac{\overline{C_{p,ab}} \overline{\rho_{ab}} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\textbf{v}_{ab} \cdot \textbf{x}_{ab}} (T_a - T_b) F_{ab} (h)
+\end{equation}
+
 \subsubsection{Discretized governing equations with $SPH-\varepsilon$ turbulence model}
-Plug discretized turbulent stress term and turbulent heat transfer term into the momentum and energy equation, we got new discretized governing equation:
+Plugging in the discretized turbulent stress term and turbulent heat transfer term into the momentum and energy equation, we get new discretized governing equation:
 \begin{equation}
 <\dfrac{d \textbf{v}_{\alpha}}{d t}>= 
--\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}) \bigtriangledown_{\alpha}w_{\alpha b}(h_{\alpha})]
--\sum_j [m_j (\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j} - \Phi_{\alpha j}) \bigtriangledown_{\alpha}w_{\alpha j}(h_{\alpha})]
+-\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}) \bigtriangledown_{\alpha}w_{\alpha b}(h)]
+-\sum_j [m_j (\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j} - \Phi_{\alpha j}) \bigtriangledown_{\alpha}w_{\alpha j}(h)]
 +\textbf{g} \label{eq:gov-sph-v}
@@
 <\dfrac{d e_{\alpha}}{d t}>
-& = 0.5\sum_b [m_b \widehat{\textbf{v}_{\alpha b}} (\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}) \bigtriangledown_{\alpha}w_{\alpha b}(h_{\alpha})] 
- + 2 \sum_b \dfrac{m_b}{\rho_{\alpha} \rho_b} \kappa_{t,\alpha b} (T_{\alpha} - T_b) F_{\alpha b} (h_{\alpha}) \\
- & +0.5\sum_j [m_j \widehat{\textbf{v}_{\alpha b}}(\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j} - \Phi_{\alpha j}) \bigtriangledown_{\alpha}w_{\alpha j}(h_{\alpha})]
- + 2 \sum_j \dfrac{m_j}{\rho_{\alpha} \rho_j} \kappa_{t,\alpha j} (T_{\alpha} - T_j) F_{\alpha j} (h_{\alpha})
+& = 0.5\sum_b [m_b \widehat{\textbf{v}_{\alpha b}} (\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}) \bigtriangledown_{\alpha}w_{\alpha b}(h)] 
+ + 2 \sum_b \dfrac{m_b}{\rho_{\alpha} \rho_b} \kappa_{t,\alpha b} (T_{\alpha} - T_b) F_{\alpha b} (h) \\
+ & +0.5\sum_j [m_j \widehat{\textbf{v}_{\alpha b}}(\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j} - \Phi_{\alpha j}) \bigtriangledown_{\alpha}w_{\alpha j}(h)]
+ + 2 \sum_j \dfrac{m_j}{\rho_{\alpha} \rho_j} \kappa_{t,\alpha j} (T_{\alpha} - T_j) F_{\alpha j} (h)
 \end{split}
@@
 It should be noted that ghost particles are not used for these interpolations.
-Assign all properties at the corresponding reflect position to the ghost particle except for velocity.
+Assign all properties at the corresponding reflected position to the ghost particle except for velocity.
 We assign a velocity to each wall ghost particles that is the negative of the interpolated velocity at its corresponding reflection inside the fluid domain.
@@
 (\ref{eq:wall_bc_rho}) and (\ref{eq:wall_bc_xi})).
-In addition, as temperature is also symmetric with respect to the boundary, the gradient of temperature will vanish on the wall boundary and hence no internal energy flux on the wall boundary (Eq.
+In addition, as temperature is also symmetric with respect to the boundary, the gradient of temperature will vanish on the wall boundary and hence there is no internal energy flux on the wall boundary (Eq.
 (\ref{eq:wall_bc_e})).
+
 \subsubsection{Eruption boundary condition}
@@
 As soon as they move out from erupt conduit, these ghost particle will be shift to real particles and their physics properties and position will be updated based on discretized governing equations.
-New ghost particle will be added at the bottom of the eruption conduit as these existing ghost particles move upwards.\\
+New ghost particle will be added at the bottom of the eruption conduit as these existing ghost particles move upwards.
 \subsubsection{Pressure boundary condition}
@@
 Real particles that moves out the pressure boundary will be removed.\\
-As simulation progress, change in position and physical properties of real particles near pressure boundary boundaries might corrupt pressure boundary condition that was established initially.
+As simulations progress, change in position and physical properties of real particles near pressure boundary boundaries might corrupt pressure boundary condition that was established initially.
 This shortcoming is relieved by choosing a larger computational domain so that boundaries that might be corrupted will locate far away from turbulent mixing area.
-In addition, to avoid enlarging fluctuation, we add another constrain on time step:
+In addition, to avoiding enlarging fluctuation, we add another constraint on the time step:
 \begin{equation}
@@
 The proper $CFL_p$ is determined by series of simulation tests.
-\\
 In subfigure b of Fig. \ref{fig:bc_and_domain_decomp}, different type of boundary conditons are marked with different color. The red portion are main computational domain. 
@@
 \caption{A cross section view of the simulation domain.
-Subfigure a shows the mass fraction of a plume simulation.
-Subfigure b shows all boundary conditions: the dark blue is for velocity inlet boundary, the light blue is for pressure outlet boundary, the gray is for no-slip wall boundary condition.
-Ghost particle ID of All real particles are set to 100, so is 100 in b.
-Subfigure c shows the cross section view of domain decomposition based on SFC.
-The simulation is conducted on 144 processors (12 nodes with 12 cores each node), so there are 144 subdomains in total.}
+Subfigure $a$ shows the mass fraction of a plume simulation.
+Subfigure $b$ shows all types of boundaries: the dark blue is for velocity inlet boundary, the light blue is for pressure outlet boundary, the gray is for no-slip wall boundary condition.
+"Ghost particle ID" of real particles are set to 100.
+Subfigure $c$ shows the cross section view of domain decomposition based on SFC.
+The simulation is conducted on 360 processors (30 nodes with 12 cores each node), so there are 360 subdomains in total.
+The cut view shows a portion of them.}
 \label{fig:bc_and_domain_decomp}
@@
 One disadvantage of 3D model is that it usually takes much longer time than 1D models to complete one simulation.
-This disadvantage further preventing doing simulation with finer resolution and accounting for more physics in one model.
+This disadvantage further prevented doing simulation with finer resolution and accounting for more physics in one model.
 Non-intrusive uncertainty analysis which is commonly adopted in hazard forecasting requires finishing multiple simulations within a given time window.
 High performance computing provides a powerful routine for relieving the misery.
-Among existing CPU parallel SPH schemes, most of them focus on neighbors searching algorithm and dynamic load balancing (eg.
+Among existing CPU parallel SPH schemes, most of them focus on neighbor searching algorithm and dynamic load balancing (eg.
 \citep {ferrari2009new, crespo2015dualsphysics}).
-Less attention has been paid to developing of more flexible data management schemes for more complicated problems.
-Motivated by techniques developed for mesh based methods, we develop a complete framework for parallelizing SPH program with MPI standard model allowing flexible and efficient data access.\\
+Less attention has been paid to developing more flexible data management schemes for more complicated problems.
+Motivated by techniques developed for mesh based methods, we develop a complete framework for parallelizing SPH program with MPI standard model allowing flexible and efficient data access.
+
 Any implementation of SPH code requires efficient searching and updating of neighbors during simulation.
@@
 In both SPHysics and DualSPHysics \citep {crespo2015dualsphysics}, the physical quantities of each particle (position, velocity, density...) are stored in arrays, and the particles (and the arrays with particle data) are reordered following the order of the cells.
-This has two advantages: 1) access pattern is more regular and more efficient, 2) it is ease to identify the particles that belong to a cell by using a range since the first particle of each cell is known.
+This has two advantages: 1) access pattern is more regular and more efficient, 2) it is easy to identify the particles that belong to a cell by using a range since the first particle of each cell is known.
 But adding, deleting and especially accessing of particles are cumbersome.
@@
 We define C++ classes which contain all data of particles and buckets.
-As for the management of data, we adopt hash tables to store pointers to particles and buckets, which give us not only flexibility of deleting and adding elements, but also quicker access compared with linked list.
-Instead of using the "nature manner" to number particles, we adopt SFC based index to give each particle and background bucket an unique identifier -- a strategy known to preserve some locality at minimal cost.
+As for the management of data, we adopt hash tables to store pointers to particles and buckets, which gives us not only flexibility of deleting and adding elements, but also quicker access compared with linked list.
+Instead of using the "nature manner" to number particles, we adopt space filling curve (SFC) based index to give each particle and background bucket an unique identifier -- a strategy known to preserve data locality at minimal cost.
 The SFC based numbering strategy is further extended to include time step information so that particles added at the same position but different time will have different identifiers.
@@
 Subfigure c of Fig.
-\ref{fig:bc_and_domain_decomp} shows a cross section view of domian decomposition.\\
+\ref{fig:bc_and_domain_decomp} shows a cross section view of domian decomposition.
+
 We will provide a supplementary documentation describing details about the data structure, domain decomposition, load balancing and domain adjusting.\\
@@
 \subsection{Simulation on Plume}
-The development of volcano plume is more complicated than JPUE in several aspects.
-Besides turbulent entrainment of ambient fluids, development of volcano plume also involves heating up of entrained air and expanding in a stratified atmosphere.
+The development of volcanic plume is more complicated than JPUE in several aspects.
+Besides turbulent entrainment of ambient fluids, development of volcanic plume also involves heating up of entrained air and expanding in a stratified atmosphere.
 A strong eruption column without wind is tested in this section for the purpose of further verification and validation.
@@
 \center
-\includegraphics[width=5cm]{Temp}
+\includegraphics[width=7cm]{Temp}
 \caption{Temperature as a function of height}
@@
 \center
-\includegraphics[width=5cm]{density_strong}
+\includegraphics[width=7cm]{density_strong}
 \caption{Density of the strong plume without wind after reaching its top height}
@@
 The profile for gas, which account for both air and vapor, shows a very similar tendency as that of entrained air.
-First of all, vapor condense is not considered in our model.
+First of all, vapor condensation is not considered in our model.
 In addition, as we assume that erupted material behaves like a single phase water, the mass fraction of gas will simply be a function of entrained air (Eq.
 (\ref{eq:gas-frac-express})).
-Among these 3D models, ATHAM takes vapor condense into account and Eq.
+Among these 3D models, ATHAM takes vapor condensation into account and Eq.
 (\ref{eq:gas-frac-express}) does not hold for ATHAM.
@@
 A new plume model was developed based on SPH method.
+Extensions necessary for Lagrangian methodology and compressible flow were made in the formulation of the equations of motion and turbulence models.
 Advanced numerical techniques in SPH were exploited and tailored for this model.
-High performance computing was used to mitigate confliction between accuracy (depends on comprehensiveness of the model, resolution, order of accuracy of numerical methods, scheme for time upgrading.) and simulation time (depends on, comprehensiveness, resolution, order of accuracy of numerical methods, scheme for time upgrading ...
+High performance computing was used to mitigate the tradeoff between accuracy (depends on comprehensiveness of the model, resolution, order of accuracy of numerical methods, scheme for time upgrading.) and simulation time (depends on, comprehensiveness, resolution, order of accuracy of numerical methods, scheme for time upgrading ...
 and computational techniques).
@@
 Top height and integrated local variables simulated by our model are consistent with both observation and simulation results of existing 3D plume models.
-Comparison between our results with that of SK-3D implies that turbulence model plays a significant role in plume modeling.\\
+Comparison between our results with that of SK-3D implies that turbulence model plays a significant role in plume modeling.
+
 Currently existing 3D models focus on certain aspect of the volcano plume (PDAC on pyroclastic flow, ATHAM on microphysics, and SK-3D on entrainment with higher accuracy and higher order of accuracy) and hence, naturally, different assumptions were made in these models.
@@
 For example, a large-particle module (LPM) was added to ATHAM to track the paths of rocky particles (pyroclastic or tephra) within the plume and predicts where these particles fall \citep{kobs2009modeling}.
-We were also motivated by such a tendency of plume modeling to choose SPH as our numerical tool besides its ability in dealing with interface for multiphase flows.
-As mentioned in the introduction section, SPH method has good extension features and adding of new physics and phases requires tiny modification on the code compared with mesh based methods.
+We were also motivated by such an evolution of plume modeling to choose SPH as our numerical tool besides its ability in dealing with interface for multiphase flows.
+As mentioned in the introduction section, SPH method has good extensibility and adding of new physics and phases requires tiny modification on the code compared with mesh based methods.
 The last but not the least, the dramatic development of computational power makes it possible to establish a comprehensive model.
-Even though right now, current computational capacity does not allow us to have too comprehensive model, the easy-extension feature of SPH makes it feasible to keep adding new physics into the model when necessary.
-What presented in this paper is actually our initial effort and results towards developing a first principle based plume model with comprehensive physics, adopting proper numerical tools and power of high performance computing.
-More advanced numerical techniques, such as adaptive particle size, Godnuv-SPH, semi-explicit time advancing scheme and better data management strategies and algorithms are on our list to exploit.
+Even though right now, current computational capacity does not allow us to have too comprehensive a model, the easy-extension feature of SPH makes it feasible to keep adding new physics into the model when necessary and feasible.
+
+We have presented in this paper an initial effort and results towards developing a first principle based plume model with comprehensive physics, adopting proper numerical tools and power of high performance computing.
+More advanced numerical techniques, such as adaptive particle size, Godnuv-SPH, semi-explicit time advancing scheme and better data management strategies and algorithms are on our list to exploit in the future.
 In the near future, effect of wind field will be take into account.
+Code will also be made available in the open source form for the community to enhance.
 %In all existing 3D plume models, numerical calculations are performed on a non-uniform (vertically and horizontally stretched) grid, with different grid resolutions and domain sizes.
 However, implementation of different resolution in SPH methods is not straight forward and requires more work.
-\section{Code availability}
+\subsection{Code availability}
 The Plume-SPH code with a user manual providing instructions for installation, running and visualization are archived in github (https://github.com/Plume-SPH/plume-sph.git).
 The input data for all simulations presented in this work can also be found in the repository.
-\section{Data availability}
+\subsection{Data availability}
 Data for running the test cases, including meteorological data, properties of erupted material and air, and eruption condition data can be found in source code repository.

### 2017-02-21 Manuscripts_GMD.tex
comments from dr. bursik

@@
+%%xx General points
+%%xx 0 Some suggestions in the text itself are shown with "xx" -- so if you just search for "xx" you will find them
+%%xx 1 Make sure tenses match throughout
+%%xx 2 use \citet for citations in-line (text), e.g., "\citet{monaghanpaper} found that SPH is great...", not "Monaghan \citep{monaghanpaper} found that SPH is great..."
+%%xx 3 Need a section on the atmospheric structure, and how atmosphere can be incorporated into model, especially NWP or radiosondes
+%%xx 4 You often seem to forget to put a space between paragraphs.  LaTeX of course does not recognize a <return> as a paragraph delimiter, nor is it proper to end a normal paragraph with \\ or %
+%%xx 5 You are inconsistent with capitalization in section headings.  Either always use an initial capital letter, or only use an initial capital letter on the first word, proper names or acronyms
+%%xx 6 Figures are of generally poor quality -- use PDF and reproduce at larger scale.  
+%%xx 7 Difficult to understand what figure of Pinatubo plume is supposed to be telling us about the model
+
 %% Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
@@
 
-\title{Plume-SPH 1.0: A 3D dust-gas volcano plume model based on Smoothed Particle Hydrodynamics}
-
-%=======
-\title{Plume-SPH 1.0: A 3D dust-gas volcano plume model based on smoothed particle hydrodynamics}
+\title{Plume-SPH 1.0: A three-dimensional, dusty-gas volcanic plume model based on smoothed-particle hydrodynamics}
+
 %>>>>>>> 0d96d9682fb484b8b159e8d09a0eb89723d3a0f8
@@
 
-\correspondence{Abani Patra (abani@buffalo.com)}
+\correspondence{Abani Patra (abani@buffalo.edu)}
 
@@
 %Volcanic ash, if it persists in the atmosphere, will pose great threat to aviation and communities. Therefore, accurate modeling and forecasts are crucial. 
-VATDs (Volcanic ash transport and dispersion models) used for critical volcanic ash forecasts usually take outputs of eruption plume models as source terms.
-The accuracy of source terms is crucial for forecasts from such VATDs.
+VATDs (Volcanic ash transport and dispersal models) used for volcanic ash forecasts sometimes take outputs of eruption plume models as source terms.
+The accuracy of source terms is crucial for forecasts from  VATDs.
 All  existing 3D (three dimensional) plume models described in the literature use mesh based methods.
-SPH (Smoothed particle hydrodynamics), as a mesh free method has several advantages over mesh based methods in modeling of multiphase free boundary flow.
-As an initial effort on exploiting the feasibility and advantages of SPH in volcano plume modeling we adopt a relatively simpler model, a 3D dusty-gas dynamic model, targeted at capturing the salient features of volcano plume.
-In this model, erupted material is assumed to be well mixed and represented by one phase while air is another phase.
-We also assume immediate dynamic and thermal dynamic equilibrium between air and erupted material and that the effect of wind is not significant.
-
-Several newly developed techniques in SPH are adopted and extended to address numerical challenges in simulating multiphase compressible turbulent flow with classical SPH method.
+SPH (smoothed particle hydrodynamics), as a mesh free method, has several advantages over mesh based methods in modeling of multiphase free boundary flows.
+As an initial effort to exploit the feasibility and advantages of SPH in volcanic plume modeling, we adopt a relatively simple model, a 3D dusty-gas dynamic model, targeted at capturing the salient features of a volcanic plume.
+In this model, erupted material is assumed to be well-mixed, and represented by one phase while air is another phase.
+We also assume immediate dynamic and thermodynamic equilibrium between air and erupted material and that the effect of wind is not significant.
+
+Several newly developed techniques in SPH are adopted and extended to address numerical challenges in simulating multiphase compressible turbulent flow with classical SPH methods.
 The $SPH-\varepsilon$ turbulence model is adopted to capture mixing at unresolved scales.
-Heat exchange due to turbulence is then calculated with Reynolds analogy.
-Corrected formalism of SPH is used to handle tensile instability and deficiency of particle distribution near the boundaries.
+Heat exchange due to turbulence is then calculated with a  Reynolds analogy.
+A corrected formalism of SPH is used to handle tensile instability and deficiency of particle distribution near the boundaries.
 We also propose feasible ways to impose velocity inlet and pressure outlet boundary conditions, both of which are scarce in traditional implementations of SPH.
 The core solver of our model is parallelized with MPI (message passing interface) obtaining good weak scalability and strong scalability.
-The model is first verified by comparing velocity and concentration distribution along axis and cross section with experimental results of JPUE (jet or plume which is ejected from a nozzle into a uniform environment).
-The top height of Pinatubo eruption (15 June 1991) simulated by our model is consistent with both observation and existing 3D plume models.
-Profiles of several integrated variables are compared with these by existing 3D plume models and further verifies the correctness of our model.
-Analysis on the plume evolution process illustrated that this model is able to reproduce the basic physics of plume development.
-The comparison also implies that turbulence model plays a significant role in 3D volcano plume modeling.
+The model is first verified by comparing velocity and concentration distribution along axis and in cross section with experimental results of JPUE (jet or plume that is ejected from a nozzle into a uniform environment).
+Next, the top height of the Pinatubo eruption of 15 June 1991, as simulated by our model, is consistent with both observations and existing 3D plume models.
+Profiles of several integrated variables are compared with those calculated in existing 3D plume models, and further verify our model.
+Analysis of the plume evolution process illustrates that this model is able to reproduce the basic physics of plume development.
+The comparison also implies that the turbulence model plays a significant role in 3D volcanic plume modeling.
 \end{abstract}
+
 \introduction  %% \introduction[modified heading if necessary]
@@
 %The dynamics of explosive eruption clouds has been a central issue in volcanology  for a long time \citep{settle1978volcanic, wilson1978control}.
- Primary hazards associated with explosive grey volcanoes include pyroclastic density currents (flows and surges), the widespread deposition of airfall tephra, and the threats to aviation posed by volcanic ash in the upper atmosphere
+ Primary hazards associated with explosive volcanic eruptions include pyroclastic density currents (flows and surges), the widespread deposition of airfall tephra, and the threats to aviation posed by volcanic ash in the atmosphere
 % In addition, there are also other secondary hazards associated with volcanoes 
 \citep{lockwood2013volcanoes}.
-Simulation of all possible hazards with one model is difficult due to diversity among these hazards. Our focus here are  the risks that volcanic ash posed on aircrafts and communities. 
+Simulation of all possible hazards with one model is difficult due to diversity among these hazards. Our focus here is  the hazards that volcanic ash poses to aircraft and communities.
+
 During volcanic eruptions, VATDs (volcanic ash transport and dispersion models) are used to forecast the location and movement of ash clouds over hours to days.
@@
 Usually, plume models are used to provide these source terms for VATDs and the forecast accuracy is critically dependent on these models.
-This paper reports on a new 3D (three dimensional) volcano plume model designed to exploit the advantages of mesh-free methods for 3D modeling of such plumes  that involve multi-phase free boundary flows.\\
+This paper reports on a new 3D (three dimensional) volcano plume model designed to exploit the advantages of mesh-free methods for 3D modeling of such plumes  that involve multi-phase free boundary flows.
+
 %=======
@@
 %which is applicable to time-dependent phenomena in actual volcanological simulations.
+
 A series of simulations based on SK-3D was reported, including determination of the entrainment coefficients of eruption columns as a function of height \citep{suzuki2010numerical},  the effect of wind field on entrainment coefficient \citep{suzuki20133d}, establishment of the relationship between the observable quantities of the eruption clouds and the eruption conditions at the vent \citep{suzuki2009three} (previously, such relationship is based on 1D models and greatly depends on empirical constants), investigation of the effect of the intensity of turbulence in the umbrella cloud on dispersion and sedimentation of tephra \citep{koyaguchi2009effect}.
+
 While the SK-3D focuses on accurately capturing the entrainment caused by turbulent mixture with higher resolution and numerical method of higher order, PDAC takes the disequilibrium between different phases into account and hence is a true multiphase model.
@@
 ASHEE is valid for low concentration and low Stokes number region and much faster than N-phase Eulerian model.
-The model is based an open source numerical solver OpenFOAM adapting its unstructured Finite Volume solver.\\
+The model is based an open source numerical solver OpenFOAM adapting its unstructured finite volume solver.
+
 To summarize, each 3D model has its own focus based on problem of interest and modeling/numerical choices made.
@@
 Lagrangian methods have several features that are suitable for volcano plume simulation that we outline below.
-Among such Lagrangian methods, Smoothed Particle Hydrodynamics \citep{gingold1977smoothed,lucy1977numerical}, %even though relatively newer compared to well designed mesh based methods such as Finite Difference,
+Among such Lagrangian methods, smoothed-particle hydrodynamics \citep{gingold1977smoothed,lucy1977numerical}, %even though relatively newer compared to well designed mesh based methods such as Finite Difference,
  based simulations have shown good agreement with experiments for many applications in fluid dynamics.
-And it currently is, by far the most widely used meshfree scheme.
-Implementations of %SPH in volcanology  may be seen in \citep{bursik2003smoothed,herault2010sph,haddad2016smoothed}.\\
+And it currently is, by far, the most widely used mesh-free scheme.
+Our implementations of SPH in volcanology follows on earlier efforts   \citep{bursik2003smoothed,herault2010sph,haddad2016smoothed}.
+
 Specifically, we choose SPH as the numerical method for volcanic plume simulation to enable:
@@
 \end{itemize}
+
 As discussed in the previous paragraph, existing 3D plume models focus on one or several specific aspects of plume and have been extended to be more comprehensive by accounting for more physics or more phases.
@@
 %So our current model is far away from comprehensive.
-Open source availability and  the easy extension feature of SPH will facilitate developing of a more comprehensive community driven model.\\
-%
+
+Open source availability and  the easy extension features of SPH will facilitate development of a more comprehensive community driven model.
+
 \subsection{Our contributions}
@@
 Less attention was paid to velocity inlet and pressure outlet boundary conditions which are required in plume modeling.
+
 \begin{itemize}
@@
 We adopt a LANS (Lagrangian Averaged Navier-Stokes) turbulence model \citep{monaghan2011turbulence}, which was originally proposed for incompressible, and is extended here for compressible flow accounting for turbulent heat exchange.
-\item Corrected formulation of SPH \citep{kumar2013parallel} is adapted to bypass the well-known tensile instability issues of classical SPH.\\
-%
+\item Corrected formulation of SPH \citep{kumar2013parallel} is adapted to bypass the well-known tensile instability issues of classical SPH.
 \item Simulation of volcanic plumes with acceptable accuracy requires fine resolution (very high particle counts) that cannot be accomplished without parallel computing.
@@
 In addition, a dynamic load balancing strategy is also developed.
-\item What's more, imposing of some types of boundary conditions (such as eruption boundary condition) requires dynamically adding and removing of particles during simulation.
+\item Imposition of some types of boundary conditions (such as eruption boundary condition) requires dynamically adding and removing of particles during simulation.
 To address this issue, we adopt an efficient data management scheme based on time dependent SFC (space filling curve) and hash table.
@@
 
-The physics model of the plume is first presented in section \ref{sec:physics-model} and leads to a complete mathematical description of the volcanic plume (governing equations and boundary conditions).
-In section \ref{sec:SPH_method} we will briefly introduce the numerical tool, SPH method.
-Both the fundamental discretization formulation and techniques that are used to handle specific issues involved in plume modeling will be discussed.
-Verification and validation with numerical tests is presented in section \ref{sec:verification-validation}.
+The physical model of the plume is first presented in section \ref{sec:physics-model} and leads to a complete mathematical description of the volcanic plume (governing equations and boundary conditions).
+In section \ref{sec:SPH_method}, we  briefly introduce the numerical tool -- the SPH method.
+Both the fundamental discretization formulation and techniques that are used to handle specific issues involved in plume modeling are discussed.
+Verification and validation with numerical tests are presented in section \ref{sec:verification-validation}.
 In section \ref{sec:conclusion}, a discussion on future work is given following a brief summary.
@@
 Jets that lose their momentum before becoming buoyant collapse back onto the ground and transform into pyroclastic flows, surges and ignimbrites.
-When plume rises up, relatively coarse particle might separate from main stream of the plume falling down onto the ground and probably be re-entrained into the plume at a lower height \citep{ernst1996sedimentation}.
+When plume rises up, relatively coarse particle might separate with main stream of the plume falling down onto the ground and probably be re-entrained into the plume at a lower height \citep{ernst1996sedimentation}.
 Within this process, erupted vapor will condense to liquid (droplet) and even further to ice.
@@
 Particle aggregation process \citep{carey1982influence,taddeucci2011aggregation}, either due to presence of liquid water, resulting from particle collision or driven by electrostatic forces might occur inside plume and thereby affect the sedimentation.
-All in all, the process of plume development is  essentially a multiphase turbulent mixing process coupled with heat transfer and many other microphysics and chemical reactions.
+
+All in all, the process of plume development is  essentially a multiphase turbulent mixing process coupled with heat transfer and  other microphysical and chemical reactions.
 
@@
 
-Because of above assumptions, all other micro-physics processes (like phase change of  \texorpdfstring{H\textsubscript{2}O}, aggregation, decomposition, absorption of gas on the surface of solid, solution of gas into the liquid) and chemical process are not considered in this model.
+Because of the above assumptions, all other microphysical processes (such as the phase changes of \texorpdfstring{H\textsubscript{2}O}, aggregation, disaggregation, absorption of gas on the surface of solids, solution of gas into a liquid) and chemical processes are not considered in this model.
 These ignored microphysics factors will come to play critical roles under particular eruption conditions.
@@
 Introducing of wind effect in our model will need imposing of pressure boundary conditions in a dynamic way, which requires more numerical effort and algorithm design.
-To summarize, our model is not valid for eruptions where wind effect plays a significant role in its development, usually referred as weak plume.
-Our model also lacks the ability in modeling plumes with large particles or eruptions in which microphysics plays non-ignorable roles, such as a eruption of \textit{El Chich{\'o}n} volcano at April 4th \citep{sigurdsson19841982, folch2016fplume}.
+To summarize, our model is not valid for eruptions where wind effect plays a significant role in its development, usually refereed as weak plume.
+Our model also lacks the ability in modeling plumes with large particles or eruptions in which microphysics plays non-ignorable roles, such as a eruption of \textit{El Chich{\'o}n} volcano on 4 April 1985xx check year xx \citep{sigurdsson19841982, folch2016fplume}.
 We are focussed here on developing the SPH based methodoloy in the context of the more basic (and more critical) aspects of volcano plume and therefore devote our effort to this relative simpler model.
@@
 $E = e + K $ is total energy which is a summation of kinetic energy $K$ and internal energy $e$.
+
 An additional equation is required to close the system. In this model, the equation for closing the system is the EOS (equation of state) for ideal gas.
@@
 $m$ represents mixture of ejected material and air, $s$ represents solid portion in ejected material, $g$ represents gas portion in the ejected material and $a$ represents air.
-In mesh based method, governing equations in Eulerian form, Eq.
+
+In mesh based methods, governing equations in Eulerian form, Eq.
 (\ref{eq:gov-cs-rho}) to Eq.
@@
 \end{align}
-Governing equations in Lagrange form, Eq.
+Governing equations in Lagrangian form, Eq.
 (\ref{eq:gov-nc-rho}) to Eq.(\ref{eq:gov-nc-e}), and Eulerian form, Eq.
 (\ref{eq:gov-cs-rho}) to Eq.
-(\ref{eq:gov-cs-e}), are actually equivalent as deriving of Lagrange form from Eulerian form is purely based on mathematics without any additional physics assumption.
+(\ref{eq:gov-cs-e}), are actually equivalent, as deriving the Lagrangian form from the Eulerian is purely based on mathematics without any additional physical assumptions.
+
 \subsection{Boundary conditions}
-In current model the initial domain is a box.
-The boundaries are categorized into velocity inlet (a circle area at the center of the bottom of the box), wall boundary (box bottom), pressure outlet (other faces of the box).
+In the current model the initial domain is a box.
+The boundaries are categorized as the velocity inlet (a circular area at the center of the bottom of the box), wall boundary (box bottom), pressure outlet (other faces of the box).
+
 \subsubsection{Velocity inlet}
@@
 \end{align}
+
 \subsubsection{Open outlet pressure boundary condition}
@@
 \end{equation}
-%
+
 \section{SPH method} \label{sec:SPH_method}
@@
 
-\subsection{Fundamental principle}
+\subsection{Fundamental principles}
 There are several procedures for discretizing governing equations (PDEs or ODEs) with SPH.
@@
 Recall that $d\textbf{x}\prime = \dfrac{dm \prime}{\rho \prime}$, the integration equation, Eq.
-(\ref{eq:SPH-fundamental-principle}), can be approximated by summation and lead to an approximation of the function $A$:
+(\ref{eq:SPH-fundamental-principle}), can be approximated by summation and lead to an approximation of the function at particle (discretization point) a:
 \begin{equation}
@@
 Generally, the accuracy increases with the order of the polynomials of the kernel function, but the computational time also increases as number of interactions increase.
-We are adopting a truncated Gaussian function in our simulation.
+We are adopting a truncated Gaussian function as the weighting function in our simulation.
 \begin{equation}
@@
 \begin{align}
-<\rho_a> = \sum_b m_b w_{ab} (h) \label{eq:ns-sph-d} \\
+<\rho_a> = \sum_b m_b w_{ab} (h_a) \label{eq:ns-sph-d} \\
 <\dfrac{d \textbf{v}_a}{d t}>= \sum_b m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}) \nabla_a w_{a b}(h) +\textbf{g} \label{eq:ns-sph-v} \\
@@
 \begin{align}
-<\rho_a> = \sum m_b w_{ab} (h) + \dot{\rho}(\textbf{x},t)\label{eq:ns-source-sph-d} \\
+<\rho_a> = \sum m_b w_{ab} (h_a) + \dot{\rho}(\textbf{x},t)\label{eq:ns-source-sph-d} \\
 <\dfrac{d \textbf{v}_a}{d t}>= \sum_b m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}) \nabla_a w_{a b}(h) +\textbf{g} + D \sum	_b m_b \dfrac{\textbf{v}_b - \textbf{v}_a}{\rho_b} \label{eq:ns-drag-sph-v} \\
@@
 Second, the second derivative of the kernel can change sign and will lead to unphysical representations (viscous dissipation causes decrease of the entropy).
+
 One of the most commonly used models of artificial viscosity is:
@@
 \end{equation}
-%Absolute(kinematic) viscous%
-Coefficient $\nu$ is defined as:
+Absolute(kinematic) viscous coefficient $\nu$ is defined as:
 \begin{equation}
@@
 It will produce a repulsive force between two particles when they are approaching each other and an attractive force when they are receding from each other.
-\\
+
 The SPH viscosity can be related to a continuum viscosity by converting the summation to integrals \citep{monaghan2005smoothed}.
@@
 \end{equation}
-For higher dimensions problem, expression for function approximation are exactly the same as Eq.
-(\ref{eq:CSP-function-approximation-1d}) even though the derivation is different.
-The first order derivative can be obtained by solving a system of linear equations explicitly or numerically \citep{chen1999improvement}.
-%\begin{equation}
-%\text{\textbf{M}}_{\alpha \beta a} \textbf{A}_{\alpha a} = \textbf{F}_{\beta a}
-%\end{equation}
-%where
-%\begin{align}
-%\text{\textbf{M}}_{\alpha \beta a} = \sum_b m_b \dfrac{\textbf{x}_b^{\alpha} - \textbf{x}_a^{\alpha}}{\rho_b} \nabla_a^{\beta} w(\textbf{x}_a -\textbf{x}_b, h) \\
-%\textbf{A}_{\alpha a} = \dfrac{\partial A}{\partial x^{\alpha}} (\textbf{x}_a)\\
-%\textbf{F}_{\beta a} = \sum_b m_b \dfrac{A_b - A_a}{\rho_b} \nabla_a^{\beta} w(\textbf{x}_a -\textbf{x}_b, h)
-%\end{align}
-%$\alpha$ and $\beta$ are direcion indices, ranging from 1 to 3 for 3D problem. 
+
+For problems of higher dimension, the expressions for function approximation are exactly the same as Eq.
+(\ref{eq:CSP-function-approximation-1d}), even though the derivation is different.
+The first order derivative can be solved explicitly or numerically from:
+\begin{equation}
+\text{\textbf{M}}_{\alpha \beta a} \textbf{A}_{\alpha a} = \textbf{F}_{\beta a}
+\end{equation}
+where
+\begin{align}
+\text{\textbf{M}}_{\alpha \beta a} = \sum_b m_b \dfrac{\textbf{x}_b^{\alpha} - \textbf{x}_a^{\alpha}}{\rho_b} \nabla_a^{\beta} w(\textbf{x}_a -\textbf{x}_b, h) \\
+\textbf{A}_{\alpha a} = \dfrac{\partial A}{\partial x^{\alpha}} (\textbf{x}_a)\\
+\textbf{F}_{\beta a} = \sum_b m_b \dfrac{A_b - A_a}{\rho_b} \nabla_a^{\beta} w(\textbf{x}_a -\textbf{x}_b, h)
+\end{align}
+$\alpha$ and $\beta$ are direcion indices, ranging from 1 to 3 for 3D problem.
+
 \subsection{Multiphase SPH}
@@
 \begin{equation}
-<\rho_{\alpha}^a>=\frac{\sum m_b w_{\alpha b} (h)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} (h) +\sum \frac{m_j}{\rho_j} w_{\alpha j} (h)} \label{eq:gov-sph-d1}
-\end{equation}
-\begin{equation}
-<\rho_\alpha^{sg}>=\frac{\sum_j m_j w_{\alpha j} (h)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} (h) +\sum \frac{m_j}{\rho_j} w_{\alpha j} (h)} \label{eq:gov-sph-d2}
+<\rho_{\alpha}^a>=\frac{\sum m_b w_{\alpha b} (h_{\alpha})}{\sum \frac{m_b}{\rho_b} w_{\alpha b} (h_{\alpha}) +\sum \frac{m_j}{\rho_j} w_{\alpha j} (h_{\alpha})} \label{eq:gov-sph-d1}
+\end{equation}
+\begin{equation}
+<\rho_\alpha^{sg}>=\frac{\sum_j m_j w_{\alpha j} (h_\alpha)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} (h_{\alpha}) +\sum \frac{m_j}{\rho_j} w_{\alpha j} (h_{\alpha})} \label{eq:gov-sph-d2}
 \end{equation}
@@
 \begin{equation}
-\rho_{\alpha}=\rho_{\alpha}^a=\frac{\sum m_b w_{\alpha b} (h)}{\sum \frac{m_b}{\rho_b} w_{\alpha b}}
+\rho_{\alpha}=\rho_{\alpha}^a=\frac{\sum m_b w_{\alpha b} (h_{\alpha})}{\sum \frac{m_b}{\rho_b} w_{\alpha b}}
 \end{equation}
@@
 \begin{equation}
-\rho_{\alpha}=\rho_{\alpha}^{sg}=\frac{\sum m_j w_{\alpha j} (h)}{\sum \frac{m_j}{\rho_j} w_{\alpha j}}
+\rho_{\alpha}=\rho_{\alpha}^{sg}=\frac{\sum m_j w_{\alpha j} (h_{\alpha})}{\sum \frac{m_j}{\rho_j} w_{\alpha j}}
 \end{equation}
 That is to say, the same density updating equation can be applied for both phases and no additional numerical treatment needed to locate where the interface is.
+
 \begin{figure}
@@
 \end{figure}
+
 \subsection{Turbulence modeling with SPH}
@@
 %\citep {gotoh2004sph, shao2006sph}.
+
 Another way is to use extremely fine resolution ("the direct numerical simulation method") at the expense of much higher computational cost that results from both increase in number of particles and decrease in time steps constrained by CFL condition.
-Here we choose the SPH-SPS method.\\
+Here we choose the SPH-SPS method.
+
 Even though free surface flow are in nature turbulent, detailed research in turbulence using SPH method is rather scarce.
@@
 and is a member of a sequence of functions which tends to the $\delta$ function in the limit where $ l\rightarrow 0$. A typical example is Gaussian.
-The length scale $l$ determines the characteristic width of the kernel and the distance over which the velocity is smoothed.\\
+The length scale $l$ determines the characteristic width of the kernel and the distance over which the velocity is smoothed.
+
 It is common practice in $LANS-\alpha$ to use a differential equation for the smoothing rather than the integral form and finally reach to system of equations that need to be solved implicitly.
@@
 In this way solving of system of equations is avoided and it also makes the method simple to implement and cheap for computation.
-The discretized form of momentum equation is obtained through lengthy derivation based on Lagrangian equations.
+
+The discretized form of the momentum equation is obtained through lengthy derivation based on Lagrangian equations.
 Derivation details and other discussions, such as conservation of momentum and energy, satisfactory of circulation theory, energy/momentum spectrum analysis based on Fourier transfer are detailed in Monaghan's paper \citep{monaghan2011turbulence}.
@@
 \label{eq:replacement-in-turb-derive}
-\dfrac{G(\vert \textbf{x} _b - \textbf{x} _a \vert, l)}{\rho _b} \rightarrow \dfrac{K_{ab}}{M}
+\dfrac{G(\vert \textbf{x} _b - \textbf{x} _a \vert, l_a)}{\rho _b} \rightarrow \dfrac{K_{ab}}{M}
 \end{equation}
@@
 \label{eq:monaghan-mom-turb}
-\dfrac{d \textbf{v}_a}{dt} = -\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}) \bigtriangledown_aw_{a b}(h)] + \sum_b m_b \dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{M} \bigtriangledown_a K_{ab}
-\end{equation}
-Instead of using Eq. (\ref{eq:monaghan-mom-turb}) which is from Monaghan's paper, we use Eq. (\ref{eq:replacement-in-turb-derive}) and notice that if $l$ is uniform and keep constant: 
+\dfrac{d \textbf{v}_a}{dt} = -\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}) \bigtriangledown_aw_{a b}(h_a)] + \sum_b m_b \dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{M} \bigtriangledown_a K_{ab}
+\end{equation}
+Instead of using Eq.
+(\ref{eq:monaghan-mom-turb}) which is from Monaghan's paper xx cite the paper instead of this!!xx, we use Eq.
+(\ref{eq:replacement-in-turb-derive}) and notice that if $l$ is uniform and keep constant:
 %Please notice that the above equation is slightly different from the original equation in Monaghan's paper (Eq.
@@
 \label{eq:SPH-mom-epsilon-turb}
-\dfrac{d \textbf{v}_a}{dt} = -\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}) \bigtriangledown_aw_{a b}(h)] + \sum_b m_b \Phi_{ab}\bigtriangledown_aG_{ab}(l)
+\dfrac{d \textbf{v}_a}{dt} = -\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}) \bigtriangledown_aw_{a b}(h_a)] + \sum_b m_b \Phi_{ab}\bigtriangledown_aG_{ab}(l_a)
 \end{equation}
@@
 Typical value of $Pr_t$ for air is 0.7 $\sim$ 0.9 . We take $Pr_t=0.85$ for gases as recommended by Kays \citep{kays1994turbulent} from summarizing of experimental results.
+
 Monaghan \citep{monaghan2005smoothed} summarized the simulation of viscosity and heat conduction in his review on SPH.
@@
 Recall that viscosity term can be discretized with SPH as shown in Eq. (\ref{eq:art-vis-original}).
-Comparing of this term to physical viscous term for compressible gas shows that it has both bulk viscosity and shear viscosity, where shear viscosity coefficient is \citep{monaghan2005smoothed}:\\
+
+Comparing of this term to physical viscous term for compressible gas shows that it has both bulk viscosity and shear viscosity, where shear viscosity coefficient is \citep{monaghan2005smoothed}:
 \begin{equation}
@@
 \begin{cases} 
-      \dfrac{1}{10} & for  \quad d=3 \\
+      \dfrac{1}{6} & for  \quad d=3 \\
       \\
-     \dfrac{1}{8}  & for  \quad d=2 
+     \dfrac{5}{24}  & for  \quad d=2 
 \end{cases}
@@
 \begin{equation}
- \sum_b \dfrac{\varepsilon}{2} \dfrac{m_b}{\rho_b} \textbf{v}_{ab} \cdot \textbf{v}_{ab} \nabla_a G_{ab}(l)= \sum_b \dfrac{\varepsilon}{2S} m_b \dfrac{\textbf{v}_{ab}}{\rho_b} \dfrac{S \textbf{v}_{ab} \cdot \textbf{x}_{ab}}{x_{ab}^2} \dfrac{x_{ab}^2}{\textbf{x}_{ab}} \nabla_a G_{ab}(l)
+ \sum_b \dfrac{\varepsilon}{2} \dfrac{m_b}{\rho_b} \textbf{v}_{ab} \cdot \textbf{v}_{ab} \nabla_a G_{ab}(l_a)= \sum_b \dfrac{\varepsilon}{2S} m_b \dfrac{\textbf{v}_{ab}}{\rho_b} \dfrac{S \textbf{v}_{ab} \cdot \textbf{x}_{ab}}{x_{ab}^2} \dfrac{x_{ab}^2}{\textbf{x}_{ab}} \nabla_a G_{ab}(l_a)
 \end{equation}
@@
 \end{equation}
-Second spatial derivative can be approximated with SPH by following JJ Monaghan \citep{monaghan2005smoothed}. 
-\begin{equation}
-C_p \dfrac{dT}{dt} = \sum_b \dfrac{m_b}{\rho_a \rho_b} (\kappa_a + \kappa_b) (T_a - T_b) F_{ab} (h)
+Second spatial derivative can be approximated with SPH by following JJxx Monaghan \citet{monaghan2005smoothed}. 
+\begin{equation}
+C_p \dfrac{dT}{dt} = \sum_b \dfrac{m_b}{\rho_a \rho_b} (\kappa_a + \kappa_b) (T_a - T_b) F_{ab} (h_a)
 \end{equation}
@@
 C_p \dfrac{dT}{dt} 
-& = \sum_b \dfrac{m_b}{\rho_a \rho_b} (\kappa_a + \kappa_b) (T_a - T_b) F_{ab} (h) \\
- &= 2 \sum_b \dfrac{m_b}{\rho_a \rho_b} \dfrac{\overline{C_{p,ab}} \overline{\rho_{ab}} \varepsilon x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 Pr_t  S \textbf{v}_{ab} \cdot \textbf{x}_{ab} } (T_a - T_b) F_{ab} (h)
+& = \sum_b \dfrac{m_b}{\rho_a \rho_b} (\kappa_a + \kappa_b) (T_a - T_b) F_{ab} (h_a) \\
+ &= 2 \sum_b \dfrac{m_b}{\rho_a \rho_b} \dfrac{\overline{C_{p,ab}} \overline{\rho_{ab}} \varepsilon x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 Pr_t  S \textbf{v}_{ab} \cdot \textbf{x}_{ab} } (T_a - T_b) F_{ab} (h_a)
 \end{split}
@@
 C_p \dfrac{dT}{dt}
- =\dfrac{\varepsilon}{S  Pr_t}  \sum_b \dfrac{m_b}{\rho_a \rho_b} \dfrac{\overline{C_{p,ab}} \overline{\rho_{ab}} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\textbf{v}_{ab} \cdot \textbf{x}_{ab}} (T_a - T_b) F_{ab} (h)
+ =\dfrac{\varepsilon}{S  Pr_t}  \sum_b \dfrac{m_b}{\rho_a \rho_b} \dfrac{\overline{C_{p,ab}} \overline{\rho_{ab}} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\textbf{v}_{ab} \cdot \textbf{x}_{ab}} (T_a - T_b) F_{ab} (h_a)
 \end{equation}
@@
 <\dfrac{d \textbf{v}_{\alpha}}{d t}>= 
--\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}) \bigtriangledown_{\alpha}w_{\alpha b}(h)]
--\sum_j [m_j (\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j} - \Phi_{\alpha j}) \bigtriangledown_{\alpha}w_{\alpha j}(h)]
+-\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}) \bigtriangledown_{\alpha}w_{\alpha b}(h_{\alpha})]
+-\sum_j [m_j (\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j} - \Phi_{\alpha j}) \bigtriangledown_{\alpha}w_{\alpha j}(h_{\alpha})]
 +\textbf{g} \label{eq:gov-sph-v}
@@
 <\dfrac{d e_{\alpha}}{d t}>
-& = 0.5\sum_b [m_b \widehat{\textbf{v}_{\alpha b}} (\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}) \bigtriangledown_{\alpha}w_{\alpha b}(h)] 
- + 2 \sum_b \dfrac{m_b}{\rho_{\alpha} \rho_b} \kappa_{t,\alpha b} (T_{\alpha} - T_b) F_{\alpha b} (h) \\
- & +0.5\sum_j [m_j \widehat{\textbf{v}_{\alpha b}}(\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j} - \Phi_{\alpha j}) \bigtriangledown_{\alpha}w_{\alpha j}(h)]
- + 2 \sum_j \dfrac{m_j}{\rho_{\alpha} \rho_j} \kappa_{t,\alpha j} (T_{\alpha} - T_j) F_{\alpha j} (h)
+& = 0.5\sum_b [m_b \widehat{\textbf{v}_{\alpha b}} (\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}) \bigtriangledown_{\alpha}w_{\alpha b}(h_{\alpha})] 
+ + 2 \sum_b \dfrac{m_b}{\rho_{\alpha} \rho_b} \kappa_{t,\alpha b} (T_{\alpha} - T_b) F_{\alpha b} (h_{\alpha}) \\
+ & +0.5\sum_j [m_j \widehat{\textbf{v}_{\alpha b}}(\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j} - \Phi_{\alpha j}) \bigtriangledown_{\alpha}w_{\alpha j}(h_{\alpha})]
+ + 2 \sum_j \dfrac{m_j}{\rho_{\alpha} \rho_j} \kappa_{t,\alpha j} (T_{\alpha} - T_j) F_{\alpha j} (h_{\alpha})
 \end{split}
@@
 (\ref{eq:gov-update-pos-turbulence}).
-Smoothed velocity is also used while computing artificial viscosity $\Pi_{ab}$.\\
-\subsection{Boundary conditions in SPH}
-\subsubsection{Wall boundary condtion}
+Smoothed velocity is also used while computing artificial viscosity $\Pi_{ab}$.
+
+\subsection{Boundary conditions}
+\subsubsection{Wall boundary condition}
 Traditionally either ghost particles that mirror SPH particles across the boundary \citep {ferrari2009new}, or boundary forces \citep {monaghan2009sph} have been used to impose the wall boundary conditions.
-One disadvantage of the later one is that the boundary forces tend to corrupt the solution in the neighborhood of such boundary forces.
-In addition, the natural way of imposing eruption boundary condition is using eruption ghost particles.
-To impose boundary conditions in a consistent way we adopt a modified version of ghost particle method \citep {kumar2013parallel} for wall boundary conditions.
-Stationary ghosts particles are deployed the same way as initial deployment of real particles.
-Instead of enforcing the symmetry particle by particle, symmetric field across the boundary is explicitly enforced.
+One disadvantage of the latter  is that the boundary forces tend to corrupt the solution in the local neighborhood.
+In addition, the natural way of imposing an eruption boundary condition is using eruption ghost particles.
+To impose boundary conditions in a consistent way, we adopt a modified version of the ghost particle method \citep {kumar2013parallel} for wall boundary conditions.
+Stationary ghost particles are deployed in the same way as initial  real particles.
+Instead of enforcing symmetry particle by particle, a symmetric field across the boundary is explicitly enforced.
 Ghost positions are reflected into the domain and physical properties are calculated at these reflected positions by SPH interpolations.
@@
 \subsubsection{Eruption boundary condition}
-The natural way of imposing eruption boundary condition is using ghost particles which moves with eruption velocity and bears the temperature of erupt material.
-A parabolic velocity profile which represents a fully developed Hagen-Poiseuille flow is used to determine the particle velocity.
-The detailed shape of the parabolic profile will determined based on averaged eruption velocity (Eq.
+The natural way of imposing an eruption boundary condition is by using ghost particles that move with the eruption velocity and bear the temperature of the erupted material.
+A parabolic velocity profile that represents a fully developed Hagen-Poiseuille flow is used to determine the inlet particle velocity.
+The detailed shape of the parabolic profile is determined based on an averaged eruption velocity (Eq.
 \ref{eq:erupt_bc_v}).
-The correctness of eruption boundary condition for density (Eq.
+The correctness of the eruption boundary condition for density (Eq.
 (\ref{eq:erupt_bc_rho})) can be guaranteed by choosing mass of erupted ghost particles given distribution of these particles.
@@
 New ghost particle will be added at the bottom of the eruption conduit as these existing ghost particles move upwards.
+
 \subsubsection{Pressure boundary condition}
-Another boundary in this model is pressure outlet boundary.
+Another boundary in this model is the pressure outlet boundary.
 For flow in a straight channel, it is possible to treat the exit the same as the entry with a prescribed velocity profile.
@@
 In subfigure b of Fig. \ref{fig:bc_and_domain_decomp}, different type of boundary conditons are marked with different color. The red portion are main computational domain.
+
 \begin{figure}
@@
 \caption{A cross section view of the simulation domain.
-Subfigure $a$ shows the mass fraction of a plume simulation.
-Subfigure $b$ shows all types of boundaries: the dark blue is for velocity inlet boundary, the light blue is for pressure outlet boundary, the gray is for no-slip wall boundary condition.
-"Ghost particle ID" of real particles are set to 100.
-Subfigure $c$ shows the cross section view of domain decomposition based on SFC.
-The simulation is conducted on 360 processors (30 nodes with 12 cores each node), so there are 360 subdomains in total.
-The cut view shows a portion of them.}
+Subfigure a shows the mass fraction of a plume simulation.
+Subfigure b shows all boundary conditions: the dark blue is for velocity inlet boundary, the light blue is for pressure outlet boundary, the gray is for no-slip wall boundary condition.
+Ghost particle ID of All real particles are set to 100, so is 100 in b.
+Subfigure c shows the cross section view of domain decomposition based on SFC.
+The simulation is conducted on 144 processors (12 nodes with 12 cores each node), so there are 144 subdomains in total.}
 \label{fig:bc_and_domain_decomp}
 \end{figure}
+
 \subsection{Parallellism and performance}
@@
 As for the actual storage of data representing the physical quantities associated to each particle, different strategies have been adopted in existing implementations of SPH.
+
 In both SPHysics and DualSPHysics \citep {crespo2015dualsphysics}, the physical quantities of each particle (position, velocity, density...) are stored in arrays, and the particles (and the arrays with particle data) are reordered following the order of the cells.
@@
 The SFC based numbering strategy is further extended to include time step information so that particles added at the same position but different time will have different identifiers.
+
 As for domain decomposition, even though more complicated graph-based partitioning tools \citep {biswas1999experiments} might get higher quality decomposition, they requires much more effort in programming and computation.
@@
 
-We will provide a supplementary documentation describing details about the data structure, domain decomposition, load balancing and domain adjusting.\\
+We will provide a supplementary documentation describing details about the data structure, domain decomposition, load balancing and domain adjusting.
+
 \begin{figure}[!t]
@@
 \end{figure}
-%
+
 \begin{figure}[!t]
@@
 \end{figure}
-%
+
 Experiments have been carried out on CCR (the computational cluster of Center for Computational Research) at Buffalo.
@@
 \ref{fig:strong_scale}).
+
 The weak scalability test is conducted with the same initial domain and various smoothing lengths.
@@
 Figure \ref{fig:adj_vs_no} shows that simulation time of the test problem is greatly reduced when we adopt the domain adjusting strategy in our code.
+
 \section{Verification and validation} \label{sec:verification-validation}
@@
 Analysis on evolution of plume shows detailed air entrainment mechanism.
+
 \subsection{Simulation of JPUE}
@@
 One overall feature of JPUE is "self-similarity" which means that the evolution of the JPUE is determined solely by the local scale of length and velocity which theoretically account for the fact that the rate of entrainment at the edge of JPUE is proportional to a characteristic velocity at each height.
-As a results, physical and numerical experiments do not necessary to have exactly the same setups and are compared on a non-dimensional basis.\\
+As a results, physical and numerical experiments do not necessary to have exactly the same setups and are compared on a non-dimensional basis.
+
 \begin{figure}
@@
 \end{figure}
+
 \begin{figure}
@@
 \end{figure}
+
 A three dimensional axisymmetric JPUE which ejects from a round vent is simulated and results are compared with experiments \citep {george1977turbulence, papanicolaou1988investigations} for verification purpose.
@@
 The coefficient $coef$ for concentration is 80 and 50 respectively according to Panos and George. $coef$ for velocity is 90 and 55 respectively according to Panos and George. 
-$slope$ for jet width based on velocity is 0.104 and for concentration is 0.157. $intercept$ for jet width based on velocity is 2.58 while that for the concentration is 4.35.\\
+$slope$ for jet width based on velocity is 0.104 and for concentration is 0.157. $intercept$ for jet width based on velocity is 2.58 while that for the concentration is 4.35.
+
 Although both velocity and concentration are found to be well matched with experimental results, a small disparity in both velocity and concentration are observed near the boundary of the jet.
@@
 These factors prevent us from numerically reproducing these experiments in an exact way as they were conducted.
-However, the features of JPUE is correctly reproduced with our code.\\
+However, the features of JPUE is correctly reproduced with our code.
+
 We also investigated the mixing due to turbulence in JPUE simulation by checking the mixture of two phases.
@@
 Both advanced SPH techniques like adaptive particles size and better data management and access strategies need to be exploited to address this issue.
+
 \begin{figure}
@@
 \end{figure}
-\subsection{Simulation on Plume}
-The development of volcanic plume is more complicated than JPUE in several aspects.
+
+\subsection{Simulation of a Plume}
+The development of a volcanic plume is more complicated than JPUE in several aspects.
 Besides turbulent entrainment of ambient fluids, development of volcanic plume also involves heating up of entrained air and expanding in a stratified atmosphere.
@@
 Then plume evolution process is investigated.
+
 \subsubsection{Global and local variables}
+
 \begin{figure}
@@
 %\end{figure}
+
 One of the key global quantities of great interest is the altitude to which the plume rises.
@@
 We first project simulation results (on disordered particles) onto a pre-defined grid before doing time average and spatial integration.
-The interpolation method is the basic SPH kernel based interpolation.\\
+The interpolation method is the basic SPH kernel based interpolation.
+
 The profiles of local variables match well with simulation results of existing 3D models in general sense. The basic phenomena in volcano plume development is correctly captured by our model.
 %: the plume first rises up to its top height and then falls down to the neutral buoyancy height and starts expanding.\\
-%
+
 \begin{figure}
@@
 %
-%
+
 \begin{figure}
@@
 \end{figure}
-%
+
 As the height increase, the amount of entrained air will increase.
@@
 This implies that turbulence model plays a critical role in plume simulation.
+
 \subsubsection{Evolution of plume}
+
 \begin{figure}
@@
 %\end{figure}
+
 Figure \ref{fig:strong_evolution} shows the overall evolution process of plume.
@@
 %
+
 \conclusions  \label{sec:conclusion}%% \conclusions[modified heading if necessary]
@@
 Actually, effort has already been put on developing more comprehensive plume models.
-For example, a large-particle module (LPM) was added to ATHAM to track the paths of rocky particles (pyroclastic or tephra) within the plume and predicts where these particles fall \citep{kobs2009modeling}.
+For example, a large-particle module (LPM) was added to ATHAM to track the paths of rocky particles (pyroclastic or tephra) within the plume and predicts where these particles fall \citep{kobs2009modeling}xx actually, I think Cerminara(?) has a paper about this that is published -- probably include as well as the thesis, or instead of the thesisxx.
 We were also motivated by such an evolution of plume modeling to choose SPH as our numerical tool besides its ability in dealing with interface for multiphase flows.
@@
 We have presented in this paper an initial effort and results towards developing a first principle based plume model with comprehensive physics, adopting proper numerical tools and power of high performance computing.
-More advanced numerical techniques, such as adaptive particle size, Godnuv-SPH, semi-explicit time advancing scheme and better data management strategies and algorithms are on our list to exploit in the future.
+More advanced numerical techniques, such as adaptive particle size, Godunov-SPH, semi-explicit time advancing scheme and better data management strategies and algorithms are on our list to exploit in the future.
 In the near future, effect of wind field will be take into account.
@@
 However, implementation of different resolution in SPH methods is not straight forward and requires more work.
+
 \subsection{Code availability}
@@
 The input data for all simulations presented in this work can also be found in the repository.
+
 \subsection{Data availability}
@@
 %\authorcontribution{TEXT}
+
 \competinginterests{Authors have the standard conflict with employers and listed sponsors}
 %\disclaimer{TEXT}
+
 \begin{acknowledgements}

### 2017-02-21 Manuscripts_GMD.tex
before merge

@@
 \begin{abstract}
-%Volcanic ash, if it persists in the atmosphere, will pose great threat to aviation and communities. Therefore, accurate modeling and forecasts are crucial. 
+
 VATDs (Volcanic ash transport and dispersal models) used for volcanic ash forecasts sometimes take outputs of eruption plume models as source terms.
@@
 
-%The dynamics of explosive eruption clouds has been a central issue in volcanology  for a long time \citep{settle1978volcanic, wilson1978control}.
 Primary hazards associated with explosive volcanic eruptions include pyroclastic density currents (flows and surges), the widespread deposition of airfall tephra, and the threats to aviation posed by volcanic ash in the atmosphere
-% In addition, there are also other secondary hazards associated with volcanoes 
-\citep{lockwood2013volcanoes}.
 Simulation of all possible hazards with one model is difficult due to diversity among these hazards. Our focus here is the hazards that volcanic ash poses to aircraft and communities.
 
-During volcanic eruptions, VATDs (volcanic ash transport and dispersion models) are used to forecast the location and movement of ash clouds over hours to days.
+During volcanic eruptions, VATDs (volcanic ash transport and dispersal models) are used to forecast the location and movement of ash clouds over hours to days.
 VATDs use eruption source parameters, such as plume height, mass eruption rate, duration, and the mass fraction distribution of erupted debris finer than about $4 \Phi$ (or $63  \mu$m), which can remain in the cloud for many hours or days.
@@
 Usually, plume models are used to provide these source terms for VATDs and the forecast accuracy is critically dependent on these models.
-This paper reports on a new 3D (three dimensional) volcano plume model designed to exploit the advantages of mesh-free methods for 3D modeling of such plumes  that involve multi-phase free boundary flows.
-
-%=======
+This paper reports on a new 3D (three dimensional) volcanic plume model designed to exploit the advantages of mesh-free methods for 3D modeling of such plumes that involve multi-phase free boundary flows.
+
 The dynamics of explosive eruption clouds has been a central issue of volcanology science for a long time \citep{settle1978volcanic, wilson1978control}.
@@
 The effects of magma type and vent conditions \citep{woods1988fluid, woods1995decompression}, atmospheric conditions \citep{ woods1993moist, sparks1997volcanic, bursik2001effect}, external surface water \citep{koyaguchi1996formation}, thermal disequilibrium, and particle fallout \citep{woods1991particle} have been assessed in increasing details based on observation, experiments and numerical studies.
-%
-%Hazards associated with explosive grey volcanoes include pyroclastic density currents (flows and surges), the widespread deposition of airfall tephra, and the threats posed by volcanic ash. 
-%In addition, there are also other secondary hazards associated with volcanoes \citep{lockwood2013volcanoes}.
-%>>>>>>> 0d96d9682fb484b8b159e8d09a0eb89723d3a0f8
-%
-%Simulation of all possible hazards with one model is difficult due to diversity among these hazards.
-Our concern in current work is the risks that volcanic ash poses on aircrafts and communities.
-%During volcanic eruptions, VATDs (volcanic ash transport and dispersion models) are used to forecast the location and movement of ash clouds over hours to days.
-Those VATDs use eruption source parameters, such as plume height, mass eruption rate, duration, and the mass fraction distribution of erupted debris finer than about $4 \Phi$ (or $63  \mu$m), which can remain in the cloud for many hours or days.
-Observational data for such parameters are usually unavailable in the first minutes or hours after an eruption is detected.
-Moreover, these input parameters are subject to change during an eruption, requiring rapid re-assignment of new parameters.
-Usually, plume models are used to provide source terms for VATDs.
-%This paper reports a new 3D (three dimensional) volcano plume model exploiting advantage of mesh-free methods in multiphase free boundary flow modeling.\\
-%
+
 \subsection{Existing plume models}
-Several 1D (one dimensional) volcano plume models have been developed in the past few decades.
+Several 1D (one dimensional) volcanic plume models have been developed in the past few decades.
 Ranging from the most basic 1D model \citep{woods1988fluid} which only accounts for mass conservation to more recently developed 1D models (eg.
@@
 In addition, these simplified models require little computational resource and can run on standard personal computers or on web sites in very short time.
-As a result, 1D software for volcano plume development (such as \citep{267, 1194, 3541}), combined with VATDs (such as \citep{114, draxler2015hysplit}) are widely used in research and practice.
+As a result, 1D software for volcanic plume development (such as \citep{267, 1194, 3541}), combined with VATDs (such as \citep{114, draxler2015hysplit}) are widely used in research and practice.
 While these 1D models can generate well-matched results with 3D (three dimensional) models for weak plumes, much greater variability is observed for strong plume scenarios, especially for local variables \citep{costa2016results}.
@@
 Advanced numerical techniques, such as second order scheme and semi-explicit time stepping, was also adopted afterwards to improve the accuracy of PDAC \citep{carcano2013semi}.
+
 Another 3D model, SK-3D \citep{suzuki2005numerical} is a 3D time-dependent fluid dynamics model that attempts to reproduce the entrainment process of eruption clouds with relatively simpler physics but with high order numerical accuracy and high spatial resolution.
-%which is applicable to time-dependent phenomena in actual volcanological simulations.
-
 A series of simulations based on SK-3D was reported, including determination of the entrainment coefficients of eruption columns as a function of height \citep{suzuki2010numerical},  the effect of wind field on entrainment coefficient \citep{suzuki20133d}, establishment of the relationship between the observable quantities of the eruption clouds and the eruption conditions at the vent \citep{suzuki2009three} (previously, such relationship is based on 1D models and greatly depends on empirical constants), investigation of the effect of the intensity of turbulence in the umbrella cloud on dispersion and sedimentation of tephra \citep{koyaguchi2009effect}.
@@
 However, the resolution of ATHAM is still pretty coarse compared with SK-3D and PDAC.
+
 Besides adding to their special strengths (ATHAM has been adding more and more microphysics, PDAC was extended to consider more phases), these models are learning from each other.
@@
 Both are using finer and finer resolution.
+
 Recently, a first order, non-equilibrium compressible 3D model, ASHEE \citep{cerminara2016ashee}, was introduced based on three dimensional N phases Eulerian transportation equation.
@@
  %Different simplification was made based on their scope. Improvements have been making since their birth. 
- Accuracy of simulation and forecast skill (depends on comprehensiveness of the model, resolution of discretization, numerical error, and order of accuracy) and simulation time (depends on number of governing equations, resolution, numerical methods and parallel techniques) .%are always a couple of conflict considerations in 3D plume simulations.\\
+ Accuracy of simulation and forecast skill (depends on comprehensiveness of the model, resolution of discretization, numerical error, and order of accuracy) and simulation time (depends on number of governing equations, resolution, numerical methods and parallel techniques).%are always a couple of conflict considerations in 3D plume simulations.\\
 %
@@
 To date there are no 3D plume models based on mesh free Lagrangian methods.
-Lagrangian methods have several features that are suitable for volcano plume simulation that we outline below.
-Among such Lagrangian methods, smoothed-particle hydrodynamics \citep{gingold1977smoothed,lucy1977numerical}, %even though relatively newer compared to well designed mesh based methods such as Finite Difference,
- based simulations have shown good agreement with experiments for many applications in fluid dynamics.
+Lagrangian methods have several features that are suitable for volcanic plume simulation that we outline below.
+Among such Lagrangian methods, smoothed-particle hydrodynamics \citep{gingold1977smoothed,lucy1977numerical} based simulations have shown good agreement with experiments for many applications in fluid dynamics.
 And it currently is, by far, the most widely used mesh-free scheme.
 Our implementations of SPH in volcanology follows on earlier efforts   \citep{bursik2003smoothed,herault2010sph,haddad2016smoothed}.
-
 Specifically, we choose SPH as the numerical method for volcanic plume simulation to enable:
@@
 For interface capturing (Eulerian) method, the surrendering of surface detail before the phase transport calculation means that interface reconstruction is required between time steps to recover the interface information, which needs additional numerical effort \citep{hirt1981volume, youngs1982time}.
-Since SPH is able to adaptively adjust the discretization and automatically construct the interface, SPH does not require additional numerical effort for interface construction and therefore is more suitable for volcano plume simulation.
+Since SPH is able to adaptively adjust the discretization and automatically construct the interface, SPH does not require additional numerical effort for interface construction and therefore is more suitable for volcanic plume simulation.
 \item Advection term in the Navier-Stokes equations does not appear explicitly in discretized formula of SPH (as illustrated in Eq.
@@
 The core physics, such as entrainment of air and thermal expansion, are essential for all plume modeling while some other physics, such as water condensation and aggregation etc., are important in specific scenario.
-As an initial effort on exploiting advantages of SPH in volcano plume modeling, we focus on capturing basic features in plume development using a robust numerically and comuptationaly efficient framework with support for scalable parallel computing.
+As an initial effort on exploiting advantages of SPH in volcanic plume modeling, we focus on capturing basic features in plume development using a robust numerically and comuptationaly efficient framework with support for scalable parallel computing.
 %So our current model is far away from comprehensive.
@@
 Our model also lacks the ability in modeling plumes with large particles or eruptions in which microphysics plays non-ignorable roles, such as a eruption of \textit{El Chich{\'o}n} volcano on 4 April 1985xx check year xx \citep{sigurdsson19841982, folch2016fplume}.
-We are focussed here on developing the SPH based methodoloy in the context of the more basic (and more critical) aspects of volcano plume and therefore devote our effort to this relative simpler model.
+We are focussed here on developing the SPH based methodoloy in the context of the more basic (and more critical) aspects of volcanic plume and therefore devote our effort to this relative simpler model.
 It is worthwhile to mention here that because SPH is adopted as our numerical method, adding of these physics into our model would require much less work in terms of programming compared to mesh based methods.
@@
 What's worse, the engulfment process during plume development makes geometry of interface very complicated.
-Lagrangian SPH is able to handle multiphase problems without additional numerical effort for dealing with interface by simply tagging particles of each phase, which makes it suitable for volcano plume simulation.
+Lagrangian SPH is able to handle multiphase problems without additional numerical effort for dealing with interface by simply tagging particles of each phase, which makes it suitable for volcanic plume simulation.
 Take the governing equation adopted in our model as an example.
@@
 The pattern of ambient particles entrainment is also clearly shown.
-Then a simulation of representative strong volcano plume is conducted.
+Then a simulation of representative strong volcanic plume is conducted.
 Both global variables and local variable are comparable with simulation results from existing 3D plume models.
@@
 
-The profiles of local variables match well with simulation results of existing 3D models in general sense. The basic phenomena in volcano plume development is correctly captured by our model.
+The profiles of local variables match well with simulation results of existing 3D models in general sense. The basic phenomena in volcanic plume development is correctly captured by our model.
 %: the plume first rises up to its top height and then falls down to the neutral buoyancy height and starts expanding.\\
@@
 
-Currently existing 3D models focus on certain aspect of the volcano plume (PDAC on pyroclastic flow, ATHAM on microphysics, and SK-3D on entrainment with higher accuracy and higher order of accuracy) and hence, naturally, different assumptions were made in these models.
-However, these different aspects of volcano plume are not independent and actually are coupled.
+Currently existing 3D models focus on certain aspect of the volcanic plume (PDAC on pyroclastic flow, ATHAM on microphysics, and SK-3D on entrainment with higher accuracy and higher order of accuracy) and hence, naturally, different assumptions were made in these models.
+However, these different aspects of volcanic plume are not independent and actually are coupled.
 In addition, there is no absolute boundary to determine which kind of hazard is dominant in certain eruption.

### 2017-02-22 Manuscripts_GMD.tex

@@
 %%xx General points
-%%xx 0 Some suggestions in the text itself are shown with "xx" -- so if you just search for "xx" you will find them
+%%xx 0 Some suggestions in the text itself are shown with "xx" -- so if you just search for "xx" you will find them --> Done
 %%xx 1 Make sure tenses match throughout
-%%xx 2 use \citet for citations in-line (text), e.g., "\citet{monaghanpaper} found that SPH is great...", not "Monaghan \citep{monaghanpaper} found that SPH is great..."
+%%xx 2 use \citet for citations in-line (text), e.g., "\citet{monaghanpaper} found that SPH is great...", not "Monaghan \citep{monaghanpaper} found that SPH is great..." -->Done
 %%xx 3 Need a section on the atmospheric structure, and how atmosphere can be incorporated into model, especially NWP or radiosondes
-%%xx 4 You often seem to forget to put a space between paragraphs.  LaTeX of course does not recognize a <return> as a paragraph delimiter, nor is it proper to end a normal paragraph with \\ or %
-%%xx 5 You are inconsistent with capitalization in section headings.  Either always use an initial capital letter, or only use an initial capital letter on the first word, proper names or acronyms
+%%xx 4 You often seem to forget to put a space between paragraphs.
+LaTeX of course does not recognize a <return> as a paragraph delimiter, nor is it proper to end a normal paragraph with \\ or % -->Done
+%%xx 5 You are inconsistent with capitalization in section headings.
+Either always use an initial capital letter, or only use an initial capital letter on the first word, proper names or acronyms -->Done
 %%xx 6 Figures are of generally poor quality -- use PDF and reproduce at larger scale.  
@@
 \begin{abstract}
-
 VATDs (Volcanic ash transport and dispersal models) used for volcanic ash forecasts sometimes take outputs of eruption plume models as source terms.
@@
 \subsection{Volcanic ash hazards}
-
 Primary hazards associated with explosive volcanic eruptions include pyroclastic density currents (flows and surges), the widespread deposition of airfall tephra, and the threats to aviation posed by volcanic ash in the atmosphere
@@
  Accuracy of simulation and forecast skill (depends on comprehensiveness of the model, resolution of discretization, numerical error, and order of accuracy) and simulation time (depends on number of governing equations, resolution, numerical methods and parallel techniques).%are always a couple of conflict considerations in 3D plume simulations.\\
-%
+
 \subsection{Features of SPH}
@@
 We adopt a LANS (Lagrangian Averaged Navier-Stokes) turbulence model \citep{monaghan2011turbulence}, which was originally proposed for incompressible, and is extended here for compressible flow accounting for turbulent heat exchange.
-\item Corrected formulation of SPH \citep{kumar2013parallel} is adapted to bypass the well-known tensile instability issues of classical SPH.
+\item Corrected formulation of SPH \citep{chen1999improvement} is adapted to bypass the well-known tensile instability issues of classical SPH.
 \item Simulation of volcanic plumes with acceptable accuracy requires fine resolution (very high particle counts) that cannot be accomplished without parallel computing.
@@
 
-\section{Physical Model} \label{sec:physics-model}
+\section{Physical model} \label{sec:physics-model}
 \subsection{Description of the model}
@@
 Jets that lose their momentum before becoming buoyant collapse back onto the ground and transform into pyroclastic flows, surges and ignimbrites.
-When plume rises up, relatively coarse particle might separate with main stream of the plume falling down onto the ground and probably be re-entrained into the plume at a lower height \citep{ernst1996sedimentation}.
+When plume rises up, relatively coarse particle might separate from main stream of the plume falling down onto the ground and probably be re-entrained into the plume at a lower height \citep{ernst1996sedimentation}.
 Within this process, erupted vapor will condense to liquid (droplet) and even further to ice.
@@
 Introducing of wind effect in our model will need imposing of pressure boundary conditions in a dynamic way, which requires more numerical effort and algorithm design.
-To summarize, our model is not valid for eruptions where wind effect plays a significant role in its development, usually refereed as weak plume.
-Our model also lacks the ability in modeling plumes with large particles or eruptions in which microphysics plays non-ignorable roles, such as a eruption of \textit{El Chich{\'o}n} volcano on 4 April 1985xx check year xx \citep{sigurdsson19841982, folch2016fplume}.
+To summarize, our model is not valid for eruptions where wind effect plays a significant role in its development, usually refered as weak plume.
+Our model also lacks the ability in modeling plumes with large particles or eruptions in which microphysics plays non-ignorable roles, such as a eruption of \textit{El Chich{\'o}n} volcano on April 4th 1982 \citep{sigurdsson19841982, folch2016fplume}.
 We are focussed here on developing the SPH based methodoloy in the context of the more basic (and more critical) aspects of volcanic plume and therefore devote our effort to this relative simpler model.
@@
 $E = e + K $ is total energy which is a summation of kinetic energy $K$ and internal energy $e$.
-
 An additional equation is required to close the system. In this model, the equation for closing the system is the EOS (equation of state) for ideal gas.
@@
 Where $r$ is the radius of the vent.
+
 \subsubsection{Non-slip wall boundary}
@@
 As a result, internal energy flux, which consists of heat flux and energy flux carried by mass flux, will vanish on the wall boundary.
-Equation (\ref{eq:wall_bc_rho}) ~ (\ref{eq:wall_bc_e}) gives no-slip wall boundary condition written in terms of primitive variables.
+Equation (\ref{eq:wall_bc_rho}) to (\ref{eq:wall_bc_e}) gives no-slip wall boundary condition written in terms of primitive variables.
 \begin{align}
@@
 Each kernel function has a compact support determined by smoothing length of each particle.
-There are several review papers \citep{monaghan1992smoothed, monaghan2005smoothed, rosswog2009astrophysical, price2012smoothed, monaghan2012smoothed}, giving a pretty comprehensive view over SPH, we will only refer here to the representation of the constitutive equations in SPH and put more focus on specific numerical techniques for plume modeling.
+There are several review papers by \citet{monaghan1992smoothed, monaghan2005smoothed, rosswog2009astrophysical, price2012smoothed, monaghan2012smoothed}, giving a pretty comprehensive view over SPH, we will only refer here to the representation of the constitutive equations in SPH and put more focus on specific numerical techniques for plume modeling.
 
@@
 There is slight difference between discretized formalism obtained by different routes.
-We present here one of them following Monaghan \citep{monaghan1992smoothed, monaghan2005smoothed}.
+We present here one of them following \citet{monaghan1992smoothed, monaghan2005smoothed, monaghan2012smoothed}.
 The starting point of approximating a function with SPH is the translation property of the Dirac function.
@@
 \end{equation}
+
 There is a wide variety of possible weighting functions, such as spline functions (with different orders) and Gaussian function.
@@
 \begin{align}
-<\rho_a> = \sum_b m_b w_{ab} (h_a) \label{eq:ns-sph-d} \\
+<\rho_a> = \sum_b m_b w_{ab} (h) \label{eq:ns-sph-d} \\
 <\dfrac{d \textbf{v}_a}{d t}>= \sum_b m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}) \nabla_a w_{a b}(h) +\textbf{g} \label{eq:ns-sph-v} \\
@@
 \end{equation}
-Absolute(kinematic) viscous coefficient $\nu$ is defined as:
+The coefficient $\nu$ is defined as:
 \begin{equation}
@@
 (\ref{eq:CSP-function-approximation-1d}), even though the derivation is different.
-The first order derivative can be solved explicitly or numerically from:
-\begin{equation}
-\text{\textbf{M}}_{\alpha \beta a} \textbf{A}_{\alpha a} = \textbf{F}_{\beta a}
-\end{equation}
-where
-\begin{align}
-\text{\textbf{M}}_{\alpha \beta a} = \sum_b m_b \dfrac{\textbf{x}_b^{\alpha} - \textbf{x}_a^{\alpha}}{\rho_b} \nabla_a^{\beta} w(\textbf{x}_a -\textbf{x}_b, h) \\
-\textbf{A}_{\alpha a} = \dfrac{\partial A}{\partial x^{\alpha}} (\textbf{x}_a)\\
-\textbf{F}_{\beta a} = \sum_b m_b \dfrac{A_b - A_a}{\rho_b} \nabla_a^{\beta} w(\textbf{x}_a -\textbf{x}_b, h)
-\end{align}
-$\alpha$ and $\beta$ are direcion indices, ranging from 1 to 3 for 3D problem.
+The first order derivative can be obtained by solving system of equations explicitly or numerically \citep{chen1999improvement}.
 
@@
 \begin{equation}
-<\rho_{\alpha}^a>=\frac{\sum m_b w_{\alpha b} (h_{\alpha})}{\sum \frac{m_b}{\rho_b} w_{\alpha b} (h_{\alpha}) +\sum \frac{m_j}{\rho_j} w_{\alpha j} (h_{\alpha})} \label{eq:gov-sph-d1}
-\end{equation}
-\begin{equation}
-<\rho_\alpha^{sg}>=\frac{\sum_j m_j w_{\alpha j} (h_\alpha)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} (h_{\alpha}) +\sum \frac{m_j}{\rho_j} w_{\alpha j} (h_{\alpha})} \label{eq:gov-sph-d2}
+<\rho_{\alpha}^a>=\frac{\sum m_b w_{\alpha b} (h)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} (h) +\sum \frac{m_j}{\rho_j} w_{\alpha j} (h)} \label{eq:gov-sph-d1}
+\end{equation}
+\begin{equation}
+<\rho_\alpha^{sg}>=\frac{\sum_j m_j w_{\alpha j} (h)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} (h) +\sum \frac{m_j}{\rho_j} w_{\alpha j} (h)} \label{eq:gov-sph-d2}
 \end{equation}
@@
 \end{equation}
+
 In the areas far away from interface, updating of density is exactly the same as that for single phase flow.
@@
 %\citep {gotoh2004sph, shao2006sph}. 
-
 Another way is to use extremely fine resolution ("the direct numerical simulation method") at the expense of much higher computational cost that results from both increase in number of particles and decrease in time steps constrained by CFL condition.
@@
 Even though free surface flow are in nature turbulent, detailed research in turbulence using SPH method is rather scarce.
-Among the first papers on turbulence are the implementations of the $\alpha$ turbulence model \citep{holm1999fluctuation}, Monaghan \citep{monaghan2002sph} and the RANS (Reynolds averaged Navier-Stokes) approach by Violeau \citep{violeau2007numerical}.
-A first wall-bounded LES (large eddy simulation) simulation with SPH was performed by Issa \citep{issa2005numerical}.
-An improved version of $SPH-\alpha$ method is proposed in 2011 by Monaghan \citep{monaghan2011turbulence}, which we refer to as $SPH-\varepsilon$ method.
+Among the first papers on turbulence are the implementations of the $\alpha$ turbulence model \citep{holm1999fluctuation, monaghan2002sph} and the RANS (Reynolds averaged Navier-Stokes) approach \citet{violeau2007numerical}.
+A first wall-bounded LES (large eddy simulation) simulation with SPH was performed by \citet{issa2005numerical}.
+An improved version of $SPH-\alpha$ method is proposed by \citet{monaghan2011turbulence}, which we refer to as $SPH-\varepsilon$ method.
 We will adopt $SPH-\varepsilon$ method in our model.
-Most of these papers constrain their focus to incompressible or weakly compressible flow.
-To simulate turbulence in plume, we will thus need to extend existing incompressible turbulent modeling methods to compressible flow.
-This implies a consideration of the evolution of densities also and reformulation to avoid any simplifications based on a single density.
-
+Most of these papers constrain their focus to incompressible flow.
 $SPH-\alpha$ model \citep{monaghan2002sph} is a powerful extension of the XSPH algorithm which reduces disorder at short length scales and retains the constants of the motion.
@@
 The recent improved version \citep{monaghan2011turbulence} of $SPH-\alpha$ method focuses on incompressible flow only.
+
+To simulate turbulence in plume, we will thus need to extend existing incompressible turbulent modeling methods to compressible flow.
+For compressible flow, energy equation is coupled with momentum equation and mass conservation equation.
+Averaging of thermal energy over space will introduce some additional terms.
 Actually, there are several additional terms in the averaged compressible Navier-Stokes equations besides the Reynolds stress term \citep{NASACompressibleTurbulence}.
@@
 Most commonly, a Reynolds analogy is used to model the turbulent heat flux.
-Simulations of heat transfer or other scalar transfer in turbulent flow simply involve adding transport terms for thermal energy or species concentration, at the expense of greater storage and longer computing times but without other difficulties \citep{cebeci2013analysis}.
+Simulations of heat transfer, or other scalar transfer, in turbulent flow simply involve adding transport terms for thermal energy or species concentration, at the expense of greater storage and longer computing times but without other difficulties \citep{cebeci2013analysis}.
 We adopt this strategy in our simulation of turbulent heat transfer with $SPH-\varepsilon$.
-PDAC \citep{neri2003multiparticle} also adopted a Reynolds analogy in its turbulence modeling.
+%PDAC \citep{neri2003multiparticle} also adopted a Reynolds analogy in its turbulence modeling.
 The additional terms associated with molecular diffusion and turbulent transport in the energy equation are either modeled in different ways or neglected sometimes \citep{NASACompressibleTurbulence}.
@@
 %The overall effect is to produce a redistribution of energy without dissipation.
+
 The typical LANS model uses a smoothed velocity $\widehat{\textbf{v}}$ 
@@
 The discretized form of the momentum equation is obtained through lengthy derivation based on Lagrangian equations.
-Derivation details and other discussions, such as conservation of momentum and energy, satisfactory of circulation theory, energy/momentum spectrum analysis based on Fourier transfer are detailed in Monaghan's paper \citep{monaghan2011turbulence}.
+Derivation and other discussions, such as conservation of momentum and energy, satisfactory of circulation theory, energy/momentum spectrum analysis based on Fourier transfer are detailed in reference \citep{monaghan2011turbulence}.
 Here we just have a brief summary of key steps.\\
@@
 \label{eq:monaghan-mom-turb}
-\dfrac{d \textbf{v}_a}{dt} = -\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}) \bigtriangledown_aw_{a b}(h_a)] + \sum_b m_b \dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{M} \bigtriangledown_a K_{ab}
-\end{equation}
-Instead of using Eq.
-(\ref{eq:monaghan-mom-turb}) which is from Monaghan's paper xx cite the paper instead of this!!xx, we use Eq.
-(\ref{eq:replacement-in-turb-derive}) and notice that if $l$ is uniform and keep constant:
+\dfrac{d \textbf{v}_a}{dt} = -\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}) \bigtriangledown_aw_{a b}(h)] + \sum_b m_b \dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{M} \bigtriangledown_a K_{ab}
+\end{equation}
+Instead of using Eq. (\ref{eq:monaghan-mom-turb}) which is from \citet{monaghan2011turbulence}, we use Eq. (\ref{eq:replacement-in-turb-derive}). Notice that if $l$ is uniform and keep constant: 
 %Please notice that the above equation is slightly different from the original equation in Monaghan's paper (Eq.
@@
 And the slight modification will not influence following derivation if $l$ is uniform and keep constant because:
-
+%
 %{\bf WE NEED TO SAY SOMETHING ABOUT INCOMPRESSIBILITY HERE}
-
+%
 \begin{equation}
@@
 \end{equation}
-Typical value of $Pr_t$ for air is 0.7 $\sim$ 0.9 . We take $Pr_t=0.85$ for gases as recommended by Kays \citep{kays1994turbulent} from summarizing of experimental results.
-
-Monaghan \citep{monaghan2005smoothed} summarized the simulation of viscosity and heat conduction in his review on SPH.
+Typical value of $Pr_t$ for air is 0.7 $\sim$ 0.9 . We take $Pr_t=0.85$ for gases as recommended \citet{kays1994turbulent} from summarizing of experimental results.
+
+\citet{monaghan2005smoothed} summarized the simulation of viscosity and heat conduction in his review on SPH.
 We will refer to his summary in our following discussion.
@@
 Recall that viscosity term can be discretized with SPH as shown in Eq. (\ref{eq:art-vis-original}). 
-
 Comparing of this term to physical viscous term for compressible gas shows that it has both bulk viscosity and shear viscosity, where shear viscosity coefficient is \citep{monaghan2005smoothed}:
@@
 \begin{cases} 
-      \dfrac{1}{6} & for  \quad d=3 \\
+      \dfrac{1}{10} & if  \quad d=3 \\
       \\
-     \dfrac{5}{24}  & for  \quad d=2 
+     \dfrac{1}{8}  & if  \quad d=2 
 \end{cases}
@@
 Please note that turbulent viscosity term has opposite sign with molecular viscosity term in discretized momentum equation and there is a minus sign in the expression of $\Pi_{ab}$, they will cancelled out.
-However, the above equation is correct only for 1D situation. For 2D or 3D, it is not easy to get the explicit expression of viscosity. \\
+
+However, the above equation is correct only for 1D situation. For 2D or 3D, it is not easy to get the explicit expression of viscosity. 
 We adopt an alternative way: obtaining a value for each pair of particles instead of persisting on getting an analytical expression.
@@
 \end{equation}
-The heat conduction equation without source term is:
+
+Heat conduction equation without source term is:
 \begin{equation}
@@
 \end{equation}
-Second spatial derivative can be approximated with SPH by following JJxx Monaghan \citet{monaghan2005smoothed}. 
-\begin{equation}
-C_p \dfrac{dT}{dt} = \sum_b \dfrac{m_b}{\rho_a \rho_b} (\kappa_a + \kappa_b) (T_a - T_b) F_{ab} (h_a)
+Second spatial derivative can be approximated with SPH by following \citet{monaghan2005smoothed}. 
+\begin{equation}
+C_p \dfrac{dT}{dt} = \sum_b \dfrac{m_b}{\rho_a \rho_b} (\kappa_a + \kappa_b) (T_a - T_b) F_{ab} (h)
 \end{equation}
@@
 C_p \dfrac{dT}{dt}
- =\dfrac{\varepsilon}{S  Pr_t}  \sum_b \dfrac{m_b}{\rho_a \rho_b} \dfrac{\overline{C_{p,ab}} \overline{\rho_{ab}} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\textbf{v}_{ab} \cdot \textbf{x}_{ab}} (T_a - T_b) F_{ab} (h_a)
+ =\dfrac{\varepsilon}{S  Pr_t}  \sum_b \dfrac{m_b}{\rho_a \rho_b} \dfrac{\overline{C_{p,ab}} \overline{\rho_{ab}} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\textbf{v}_{ab} \cdot \textbf{x}_{ab}} (T_a - T_b) F_{ab} (h)
 \end{equation}
@@
 <\dfrac{d \textbf{v}_{\alpha}}{d t}>= 
--\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}) \bigtriangledown_{\alpha}w_{\alpha b}(h_{\alpha})]
--\sum_j [m_j (\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j} - \Phi_{\alpha j}) \bigtriangledown_{\alpha}w_{\alpha j}(h_{\alpha})]
+-\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}) \bigtriangledown_{\alpha}w_{\alpha b}(h)]
+-\sum_j [m_j (\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j} - \Phi_{\alpha j}) \bigtriangledown_{\alpha}w_{\alpha j}(h)]
 +\textbf{g} \label{eq:gov-sph-v}
@@
 As position of all pressure ghost particles are keep constant, we are essentially imposing a static pressure boundary condition.
-Real particles that moves out the pressure boundary will be removed.\\
+Real particles that moves out the pressure boundary will be removed.
+
 As simulations progress, change in position and physical properties of real particles near pressure boundary boundaries might corrupt pressure boundary condition that was established initially.
@@
 The proper $CFL_p$ is determined by series of simulation tests.
-In subfigure b of Fig. \ref{fig:bc_and_domain_decomp}, different type of boundary conditons are marked with different color. The red portion are main computational domain.
+
+In subfigure b of Fig. \ref{fig:bc_and_domain_decomp}, different type of boundary conditons are marked with different colors. The red portion are main computational domain.
 
@@
 Subfigure b shows all boundary conditions: the dark blue is for velocity inlet boundary, the light blue is for pressure outlet boundary, the gray is for no-slip wall boundary condition.
-Ghost particle ID of All real particles are set to 100, so is 100 in b.
-Subfigure c shows the cross section view of domain decomposition based on SFC.
-The simulation is conducted on 144 processors (12 nodes with 12 cores each node), so there are 144 subdomains in total.}
+"Ghost particle ID" of all real particles are set to 100, so is 100 in subfigure $b$.
+Subfigure $c$ shows the cross section view of domain decomposition based on SFC.
+The simulation is conducted on 360 processors (30 nodes with 12 cores each node), so there are 360 subdomains in total.
+The cut view shows portion of them}
 \label{fig:bc_and_domain_decomp}
@@
 Any implementation of SPH code requires efficient searching and updating of neighbors during simulation.
-Of the many choices possible for this we adopted a background grid which was proposed by Monaghan and Lattanzio \citep {monaghan1985refined} and is quite popular in parallel SPH.
+Of the many choices possible for this we adopted a background grid which was proposed by \citet {monaghan1985refined} and is quite popular in parallel SPH.
 The background grid is also used for domain decomposition in SPH.
@@
 But adding, deleting and especially accessing of particles are cumbersome.
-Ferrari \citep {ferrari2009new} adopted linked lists using pointers so that particles can be deleted or added during the simulation.
+\citet {ferrari2009new} adopted linked lists using pointers so that particles can be deleted or added during the simulation.
 Storage problems caused by fix-size arrays are thereby also eliminated.
@@
 So we adopt an easy-programming scheme based on SFC \citep {patra1999efficient}.
-Subfigure c of Fig.
+Subfigure $c$ of Fig.
 \ref{fig:bc_and_domain_decomp} shows a cross section view of domian decomposition.
@@
 There exist consistently good experimental data \citep { list1982turbulent,dimotakis1983structure, papanicolaou1988investigations} that describe the JPUE flow field giving an insight into details of JPUE, such as transverse velocity and concentration profile.
-In this section, we will verify that our code and the $SPH-\varepsilon$ turbulence model is able to reproduce feature of turbulent entrainment by a JPUE simulation.\\
+In this section, we will verify that our code and the $SPH-\varepsilon$ turbulence model is able to reproduce feature of turbulent entrainment by a JPUE simulation.
+
 As many of these experiments were conducted with liquid, we replace the original equation of state (Eq.
@@
 This will further simplify the scenario for the convenience of studying turbulent mixing.
+
 One overall feature of JPUE is "self-similarity" which means that the evolution of the JPUE is determined solely by the local scale of length and velocity which theoretically account for the fact that the rate of entrainment at the edge of JPUE is proportional to a characteristic velocity at each height.
@@
 Experimental data of concentration and velocity distribution across the cross-section was fit into a Gaussian profile: $\dfrac{\varphi}{\varphi_c}=exp[-coef (\dfrac{r}{z})^2)]$ (solid line) even though there is no priori reason.
-Panos \citep{papanicolaou1988investigations} also fit concentration distribution and jet width based on velocity along central line into a straight line : $\dfrac{\varphi_0}{\varphi_c}=slope (\dfrac{z}{D} + intercept)$ (dash line).
+\citet{papanicolaou1988investigations} also fit concentration distribution and jet width based on velocity along central line into a straight line : $\dfrac{\varphi_0}{\varphi_c}=slope (\dfrac{z}{D} + intercept)$ (dash line).
 Where $\varphi$ is either velocity or concentration, the subscript $c$ represents the centreline, subscript $0$ represents the cross-sectionally averaged exit value.
@@
 
-\subsection{Simulation of a Plume}
+\subsection{Simulation of a plume}
 The development of a volcanic plume is more complicated than JPUE in several aspects.
@@
 \end{figure}
-%\begin{figure}
-%\center
-%\includegraphics[width=5cm]{velocity_strong}
-%\caption{Velocity of the strong plume without wind after reaching its top height}
-%\label{fig:strong_local_velocity}
-%\end{figure}
 
@@
 \ref{fig:strong_local_temp} ~ \ref{fig:strong_local_radius}.
-To get rid of significant fluctuations in time and space we did time averaging and spatial integration of the dynamic 3D flow fields by following Cerminara \citep {cerminara2016large}.
+To get rid of significant fluctuations in time and space we did time averaging and spatial integration of the dynamic 3D flow fields by following \citet {cerminara2016large}.
 As particles distribute in the space in a disordered manner.
@@
 The profiles of local variables match well with simulation results of existing 3D models in general sense. The basic phenomena in volcanic plume development is correctly captured by our model.
-%: the plume first rises up to its top height and then falls down to the neutral buoyancy height and starts expanding.\\
 
@@
 In the meanwhile, bulk density decrease due to entrainment and expansion (Fig.
-\ref{fig:strong_local_density}).\\
+\ref{fig:strong_local_density}).
+
 Our model adopts the same assumptions and governing equations as SK-3D.
@@
 \end{figure}
-%\begin{figure}
-%\center
-%\includegraphics[width=5cm]{t15}
-%\caption{Contour of mass fraction and velocity quiver at 15 seconds, column rises up quickly due to initial momentum.}
-%\label{fig:strong_t15}
-%\end{figure}
 
@@
 \end{figure}
-%\begin{figure}
-%\center
-%{
-%\includegraphics[width=10cm]{t100}
-%\caption{Contour of mass fraction and velocity quiver at 100 seconds. The entrainment of air is clearly shown in the zoomed view (right figure). }
-%\label{fig:strong_plume_t100}
-%}
-%\end{figure}
-%\begin{figure}
-%\center
-%\includegraphics[width=5cm]{t200}
-%\caption{Contour of mass fraction at 200 seconds, plume reaches its top height}
-%\label{fig:strong_t200}
-%\end{figure}
-%\begin{figure}
-%\center
-%\includegraphics[width=5cm]{t300}
-%\caption{Contour of mass fraction at 300 seconds, the cloud is spreading}
-%\label{fig:strong_t300}
-%\end{figure}
 
@@
 The evolution of the plume further verifies our model.
-%
 
@@
 However, these different aspects of volcanic plume are not independent and actually are coupled.
+For example, it has been illustrated that gas-particle non-equilibrium would introduce a previously unrecognized jet-dragging effect, which imposes great influence on plume development, especially for weak plume \cite{cerminara2016large}.
 In addition, there is no absolute boundary to determine which kind of hazard is dominant in certain eruption.
@@
 Actually, effort has already been put on developing more comprehensive plume models.
-For example, a large-particle module (LPM) was added to ATHAM to track the paths of rocky particles (pyroclastic or tephra) within the plume and predicts where these particles fall \citep{kobs2009modeling}xx actually, I think Cerminara(?) has a paper about this that is published -- probably include as well as the thesis, or instead of the thesisxx.
+For example, a large-particle module (LPM) was added to ATHAM to track the paths of rocky particles (pyroclastic or tephra) within the plume and predicts where these particles fall \citep{kobs2009modeling}.
 We were also motivated by such an evolution of plume modeling to choose SPH as our numerical tool besides its ability in dealing with interface for multiphase flows.
@@
 The last but not the least, the dramatic development of computational power makes it possible to establish a comprehensive model.
-Even though right now, current computational capacity does not allow us to have too comprehensive a model, the easy-extension feature of SPH makes it feasible to keep adding new physics into the model when necessary and feasible.
+Even though right now, current computational capacity does not allow us to have too comprehensive model, the easy-extension feature of SPH makes it convenient to keep adding new physics into the model when necessary and computationally feasible.
 

### 2017-02-24 Manuscripts_GMD.tex
revision according to Dr. Bursik's comments are made

@@
 LaTeX of course does not recognize a <return> as a paragraph delimiter, nor is it proper to end a normal paragraph with \\ or % -->Done
-%%xx 5 You are inconsistent with capitalization in section headings.
-Either always use an initial capital letter, or only use an initial capital letter on the first word, proper names or acronyms -->Done
+%%xx 5 You are inconsistent with capitalization in section headings.  Either always use an initial capital letter, or only use an initial capital letter on the first word, proper names or acronyms
 %%xx 6 Figures are of generally poor quality -- use PDF and reproduce at larger scale.  
@@
 \subsection{Volcanic ash hazards}
-Primary hazards associated with explosive volcanic eruptions include pyroclastic density currents (flows and surges), the widespread deposition of airfall tephra, and the threats to aviation posed by volcanic ash in the atmosphere
+Primary hazards associated with explosive volcanic eruptions include pyroclastic density currents (flows and surges), the widespread deposition of airfall tephra, and the threats to aviation posed by volcanic ash in the atmosphere.
 Simulation of all possible hazards with one model is difficult due to diversity among these hazards. Our focus here is the hazards that volcanic ash poses to aircraft and communities.
@@
 Several 1D (one dimensional) volcanic plume models have been developed in the past few decades.
-Ranging from the most basic 1D model \citep{woods1988fluid} which only accounts for mass conservation to more recently developed 1D models (eg.
-\citep{bursik2001effect, mastin2007user, degruyter2012improving, woodhouse2013interaction, devenish2013using, de2015plume, folch2016fplume, pouget2016sensitivity}) which tend to account for more complex physics effects.
+Ranging from the most basic 1D model \citep{woods1988fluid} which only accounts for mass conservation to more recently developed 1D models  \citep{bursik2001effect, mastin2007user, degruyter2012improving, woodhouse2013interaction, devenish2013using, de2015plume, folch2016fplume, pouget2016sensitivity} which tend to account for more complex physics effects.
 FPLUME-1.0 \citep{folch2016fplume} accounts for wind effect, entrainment of moisture, water phase change, particle fallout and re-entrainment and even wet aggregation of ash.
@@
 In addition, these simplified models require little computational resource and can run on standard personal computers or on web sites in very short time.
-As a result, 1D software for volcanic plume development (such as \citep{267, 1194, 3541}), combined with VATDs (such as \citep{114, draxler2015hysplit}) are widely used in research and practice.
+As a result, 1D software for volcanic plume development \citep[such as][]{267, 1194, 3541}, combined with VATDs \citep[such as][]{114, draxler2015hysplit} are widely used in research and practice.
 While these 1D models can generate well-matched results with 3D (three dimensional) models for weak plumes, much greater variability is observed for strong plume scenarios, especially for local variables \citep{costa2016results}.
@@
 Additional modules include radiation \citep{langmann1997radiative}, gas phase chemistry \citep{trentmann2002simulation} and gas scavenging by hydrometeors \citep{textor2003injection} were added lately.
-A further extension was made to include particle aggregation in 2006 \citep{textor2006volcanic1, textor2006volcanic2}.
-However, the resolution of ATHAM is still pretty coarse compared with SK-3D and PDAC.
+A further extension was made to include particle aggregation \citep{textor2006volcanic1, textor2006volcanic2}.
+However, the resolution of ATHAM is still coarse compared with SK-3D and PDAC.
 
@@
  %Different simplification was made based on their scope. Improvements have been making since their birth. 
- Accuracy of simulation and forecast skill (depends on comprehensiveness of the model, resolution of discretization, numerical error, and order of accuracy) and simulation time (depends on number of governing equations, resolution, numerical methods and parallel techniques).%are always a couple of conflict considerations in 3D plume simulations.\\
+ Accuracy of simulation and forecast skill (depends on comprehensiveness of the model, resolution of discretization, numerical error, and order of accuracy) and simulation time (depends on number of governing equations, resolution, numerical methods and parallel techniques) are always a pair of conflict considerations in 3D plume simulations.\\
 
@@
 \begin{itemize}
-\item better investigation of mixing phenomena;
+\item Better investigation of mixing phenomena;
 \item accurate modeling of the development of ZFE (Zone of Flow Establishment), ZEF (Zone of Established Flow) investigation and relation to column collapse and the questions relating to the development of entrainment; %and proper characterization through numerical experimentation.
-and,
 \item  easy inclusion of particles of different sizes (phases) and investigation of detailed mechanics of sedimentation and drag force interaction in lower plume.
@@
 As an initial effort on exploiting the feasibility and advantage of SPH in plume modeling, our model is designed to describe an injection of well mixed solid and volcanic gas from a circular vent above a flat surface into a stratified stationary atmosphere following SK-3D \citep{suzuki2005numerical}.
-In this model, molecular viscosity and physics heat conduction is neglected since turbulent exchange coefficients are dominant.
+In this model, molecular viscosity and heat conduction is neglected since turbulent exchange coefficients are dominant.
 Erupted material consist of solid with different size and mixture of gases (various constituent) is assumed to be well mixed and behave like a single phase fluid (phase 2) which is valid for eruptions with fine particles and ash.
@@
 The boundaries are categorized as the velocity inlet (a circular area at the center of the bottom of the box), wall boundary (box bottom), pressure outlet (other faces of the box).
+%probably should add a section to list the input parameters in the verification section
 
@@
 \end{equation}
-For vector quantities the expressions are similar, simply replacing $A$ with $\textbf{A}$ in Eq. (\ref{eq:SPH-approximation-sum}) and and Eq. (\ref{eq:SPH-scalar-function-gradient}), giving
+For vector quantities the expressions are similar, simply replacing $A$ with $\textbf{A}$ in Eq. (\ref{eq:SPH-approximation-sum}) and Eq. (\ref{eq:SPH-scalar-function-gradient}), giving
 \begin{align}
@@
 Various tricks are used to conserve linear and angular momentum and thermal energy \citep{monaghan1992smoothed}.
-Special treatments are also need for second order derivative terms.
+Special treatments are also needed for second order derivative terms.
 We only refer here to one of these possible discretizations of compressible Euler equations with SPH:
@@
 (\ref{eq:ns-sph-d}), adding a drag force into Eq.
-(\ref{eq:ns-sph-v})  anding a heat exchange term into Eq.
+(\ref{eq:ns-sph-v})  and adding a heat exchange term into Eq.
 (\ref{eq:ns-sph-e}) will lead to the new discretization form:
@@
 \subsection{Time step}
-The physical properties (velocity, density, position and pressure) change every time step due to particle interactions.
+The physical quantities (velocity, density, position and pressure) change every time step due to particle interactions.
 We have to consider the Courant condition, which is in spirit similar to the Courant condition for the grid-based methods, to determine the amount of the time step $\Delta t$.
@@
 Even though free surface flow are in nature turbulent, detailed research in turbulence using SPH method is rather scarce.
-Among the first papers on turbulence are the implementations of the $\alpha$ turbulence model \citep{holm1999fluctuation, monaghan2002sph} and the RANS (Reynolds averaged Navier-Stokes) approach \citet{violeau2007numerical}.
+Among the first papers on turbulence are the implementations of the $\alpha$ turbulence model \citep{holm1999fluctuation, monaghan2002sph} and the RANS (Reynolds averaged Navier-Stokes) approach \citep{violeau2007numerical}.
 A first wall-bounded LES (large eddy simulation) simulation with SPH was performed by \citet{issa2005numerical}.
@@
 For compressible flow, energy equation is coupled with momentum equation and mass conservation equation.
-Averaging of thermal energy over space will introduce some additional terms.
-Actually, there are several additional terms in the averaged compressible Navier-Stokes equations besides the Reynolds stress term \citep{NASACompressibleTurbulence}.
-The averaged momentum equation for compressible flow are in the same form as that for incompressible flow, all of other additional terms except for turbulent stress term are in the energy equation.
+Averaging of thermal energy over space will introduce some additional terms besides the Reynolds stress term into energy equation \citep{NASACompressibleTurbulence}.
+The averaged momentum equation for compressible flow are in the same form as that for incompressible flow, all of other additional terms besides a corresponding turbulent stress term are in the energy equation.
 Most turbulence modeling focuses on the Reynolds stress terms which are either solved directly or defined via a constitutive relation.
@@
 \end{equation}
-Please note that turbulent viscosity term has opposite sign with molecular viscosity term in discretized momentum equation and there is a minus sign in the expression of $\Pi_{ab}$, they will cancelled out.
+Please note that turbulent viscosity term has opposite sign with molecular viscosity term in discretized momentum equation and there is a minus sign in the expression of $\Pi_{ab}$, they will cancell out.
 
@@
 \subsection{Boundary conditions}
+All boundary conditions are imposed by ghost particles.
+Subfigure $b$ of Fig.
+\ref{fig:bc_and_domain_decomp} shows how boundaries are deployed.
+physical quantities of particles, including real particles and different types of ghost particles, are update differently.
 \subsubsection{Wall boundary condition}
-Traditionally either ghost particles that mirror SPH particles across the boundary \citep {ferrari2009new}, or boundary forces \citep {monaghan2009sph} have been used to impose the wall boundary conditions.
+Traditionally either ghost particles that mirror SPH particles across the boundary \citep {ferrari2009new} or boundary forces \citep {monaghan2009sph} have been used to impose the wall boundary conditions.
 One disadvantage of the latter  is that the boundary forces tend to corrupt the solution in the local neighborhood.
-In addition, the natural way of imposing an eruption boundary condition is using eruption ghost particles.
+In addition, a natural way of imposing eruption boundary condition is using eruption ghost particles.
 To impose boundary conditions in a consistent way, we adopt a modified version of the ghost particle method \citep {kumar2013parallel} for wall boundary conditions.
-Stationary ghost particles are deployed in the same way as initial  real particles.
+Stationary wall ghost particles are deployed in the same way as real particles.
 Instead of enforcing symmetry particle by particle, a symmetric field across the boundary is explicitly enforced.
-Ghost positions are reflected into the domain and physical properties are calculated at these reflected positions by SPH interpolations.
+Ghost particles are reflected into the domain and physical quantities are calculated at these reflected positions by SPH interpolations.
 It should be noted that ghost particles are not used for these interpolations.
-Assign all properties at the corresponding reflected position to the ghost particle except for velocity.
-We assign a velocity to each wall ghost particles that is the negative of the interpolated velocity at its corresponding reflection inside the fluid domain.
+Assign all properties (except for velocity) at the corresponding reflected position to the ghost particle.
+The velocity of each wall ghost particles is set to have the same value but opposite direction of the interpolated velocity at its corresponding reflection.
 By this way, the no-slip wall boundary condition (Eq.
@@
 (\ref{eq:wall_bc_rho}) and (\ref{eq:wall_bc_xi})).
-In addition, as temperature is also symmetric with respect to the boundary, the gradient of temperature will vanish on the wall boundary and hence there is no internal energy flux on the wall boundary (Eq.
+In addition, as temperature is also symmetric with respect to the boundary, the gradient of temperature will vanish and hence there is no internal energy flux on the wall boundary (Eq.
 (\ref{eq:wall_bc_e})).
@@
 \subsubsection{Eruption boundary condition}
-The natural way of imposing an eruption boundary condition is by using ghost particles that move with the eruption velocity and bear the temperature of the erupted material.
+A natural way of imposing eruption boundary condition is using ghost particles that move with the eruption velocity and bear the temperature of the erupted material.
 A parabolic velocity profile that represents a fully developed Hagen-Poiseuille flow is used to determine the inlet particle velocity.
 The detailed shape of the parabolic profile is determined based on an averaged eruption velocity (Eq.
-\ref{eq:erupt_bc_v}).
+(\ref{eq:erupt_bc_v})).
 The correctness of the eruption boundary condition for density (Eq.
@@
 (\ref{eq:erupt_bc_xi})) is automatically satisfied as all particles in the eruption conduit are of phase 2.
-The density, momentum, and internal energy of these eruption ghost particles will not be updated before they move above ground.
-As soon as they move out from erupt conduit, these ghost particle will be shift to real particles and their physics properties and position will be updated based on discretized governing equations.
-New ghost particle will be added at the bottom of the eruption conduit as these existing ghost particles move upwards.
+The density, momentum, and internal energy of these eruption ghost particles are not updated before they move above ground.
+As soon as they move out from erupt conduit, these ghost particles will be shift to real particles and their physical quantities and position will be updated based on discretized governing equations.
+New ghost particles need to be added at the bottom of the eruption conduit as these existing ghost particles move upwards.
 
 \subsubsection{Pressure boundary condition}
-Another boundary in this model is the pressure outlet boundary.
+Another boundary condition in our model is the pressure outlet boundary.
 For flow in a straight channel, it is possible to treat the exit the same as the entry with a prescribed velocity profile.
@@
 However, the natural boundary condition (Eq.
-(\ref{eq:pressure_bc_p})) is more suitable for plume simulation where the outlet is open atmosphere.
+(\ref{eq:pressure_bc_p})) is more suitable for plume simulation as the outlet is open atmosphere.
 The way we impose pressure boundary condition is: adding several layers of pressure ghost particles surrounding the real atmosphere particles.
-Physics properties except for velocity (pressure, density, temperature) are determined based on the elevation of particles.
+Pressure, density and temperature are determined based on the elevation of particles.
 Velocity is set to zero for static atmosphere.
-The physics properties for pressure ghost particles will not be updated while these for real particles will be updated at every time step.
-As position of all pressure ghost particles are keep constant, we are essentially imposing a static pressure boundary condition.
-Real particles that moves out the pressure boundary will be removed.
-
-As simulations progress, change in position and physical properties of real particles near pressure boundary boundaries might corrupt pressure boundary condition that was established initially.
+The physical quantities for pressure ghost particles are not updated while these for real particles are updated at every time step.
+As position of all pressure ghost particles keep constant, we essentially impose a static pressure boundary condition.
+Real particles are removed as soon as they move out the pressure boundary.
+
+As simulations progress, change in position and physical quantities of real particles near pressure boundaries might corrupt pressure boundary condition that was established initially.
 This shortcoming is relieved by choosing a larger computational domain so that boundaries that might be corrupted will locate far away from turbulent mixing area.
-In addition, to avoiding enlarging fluctuation, we add another constraint on the time step:
+In addition, to avoid enlarging fluctuations, we add another constraint on the time step:
 \begin{equation}
@@
 Where $CFL_p$ is a safety coefficient which has similar function as the normal $CFL$ number.
-Too small $CFL_p$ will slow down simulation while too large $CFL_p$ will lose its ability of mitigate numerical fluctuation near the boundary.
+Too small $CFL_p$ would slow down simulation while too large $CFL_p$ would lose its ability of mitigate numerical fluctuation near the boundary.
 The proper $CFL_p$ is determined by series of simulation tests.
-
-In subfigure b of Fig. \ref{fig:bc_and_domain_decomp}, different type of boundary conditons are marked with different colors. The red portion are main computational domain.
 
@@
 \caption{A cross section view of the simulation domain.
-Subfigure a shows the mass fraction of a plume simulation.
-Subfigure b shows all boundary conditions: the dark blue is for velocity inlet boundary, the light blue is for pressure outlet boundary, the gray is for no-slip wall boundary condition.
-"Ghost particle ID" of all real particles are set to 100, so is 100 in subfigure $b$.
+Subfigure $a$ shows the mass fraction.
+Subfigure $b$ shows all boundary conditions: the dark blue region are occupied by eruption ghost particles with "Ghost particle ID" of 0, the light blue area are occupied by pressure ghost particles with "Ghost particle ID" of 1, the gray area is filled with wall ghost particles with "Ghost particle ID" of 2, "Ghost particle ID" of all real particles are set to 100, they occupy the major portion of the domain in subfigure $b$.
 Subfigure $c$ shows the cross section view of domain decomposition based on SFC.
@@
 One disadvantage of 3D model is that it usually takes much longer time than 1D models to complete one simulation.
-This disadvantage further prevented doing simulation with finer resolution and accounting for more physics in one model.
+This disadvantage further prevents doing simulation with finer resolution and accounting for more physics in one model.
 Non-intrusive uncertainty analysis which is commonly adopted in hazard forecasting requires finishing multiple simulations within a given time window.
 High performance computing provides a powerful routine for relieving the misery.
-Among existing CPU parallel SPH schemes, most of them focus on neighbor searching algorithm and dynamic load balancing (eg.
-\citep {ferrari2009new, crespo2015dualsphysics}).
+Among existing CPU parallel SPH schemes, most of them focus on neighbor searching algorithm and dynamic load balancing \citep[eg.
+][] {ferrari2009new, crespo2015dualsphysics}.
 Less attention has been paid to developing more flexible data management schemes for more complicated problems.
@@
 Any implementation of SPH code requires efficient searching and updating of neighbors during simulation.
-Of the many choices possible for this we adopted a background grid which was proposed by \citet {monaghan1985refined} and is quite popular in parallel SPH.
+Of the many choices possible for this we adopt background grid which was proposed by \citet {monaghan1985refined} and is quite popular in parallel SPH.
 The background grid is also used for domain decomposition in SPH.
 We refer to the elements of background grid, namely squares for two dimensions and cubic for three dimensions, as buckets.
-As for the actual storage of data representing the physical quantities associated to each particle, different strategies have been adopted in existing implementations of SPH.
-
-In both SPHysics and DualSPHysics \citep {crespo2015dualsphysics}, the physical quantities of each particle (position, velocity, density...) are stored in arrays, and the particles (and the arrays with particle data) are reordered following the order of the cells.
+As for the actual storage of the physical quantities associated to each particle, different strategies have been adopted in existing implementations of SPH.
+
+In both SPHysics and DualSPHysics \citep {crespo2015dualsphysics}, the physical quantities of each particle (position, velocity, density...) are stored in arrays.
+The particles (and the arrays with particle data) are reordered following the order of the cells.
 This has two advantages: 1) access pattern is more regular and more efficient, 2) it is easy to identify the particles that belong to a cell by using a range since the first particle of each cell is known.
@@
 
-Experiments have been carried out on CCR (the computational cluster of Center for Computational Research) at Buffalo.
+Performance tests have been carried out on CCR (the computational cluster of Center for Computational Research) at Buffalo.
 E5645 CPUs running at 2.40GHz clock rate with 4GB memory per core on a Q-Logic Infiniband is used in these tests.
@@
 \section{Verification and validation} \label{sec:verification-validation}
-We will now present a series of numerical simulations to verify and validate the model.
+We present a series of numerical simulations to verify and validate our model in this section.
 Plume-SPH is first verified by a JPUE simulation.
@@
 Both global variables and local variable are comparable with simulation results from existing 3D plume models.
-Analysis on evolution of plume shows detailed air entrainment mechanism.
 
@@
 In addition, the "atmosphere" is assumed to be uniform and gravity is set to be zero.
-We set the temperature and density of ejected material from the vent the same as its surrounding ambient.
+We set the temperature and density of ejected material the same as surrounding ambient.
 This will further simplify the scenario for the convenience of studying turbulent mixing.
@@
 As a results, physical and numerical experiments do not necessary to have exactly the same setups and are compared on a non-dimensional basis.
+
+\begin{table}[htp]
+	\begin{centering}
+      \caption{List of eruption condition for the test cases}		
+	  \begin{tabular}{lrrr}
+	    \hline
+	    Parameters & Units  & JPUE & Plume \\
+	    \hline
+	    Vent velocity          & $m\cdot s^{-1}$  & 275               & 275 \\
+	    Vent gas mass fraction &             & 0.5               & 0.5 \\
+	    Vent Temperature       & $K$         & 1053              & 1053 \\
+	    Vent height            & $m$         & 1500              & 1500 \\
+	    Mass discharge rate    & $kg\cdot s^{-1}$ & $1.5 \times 10^9$ & $1.5 \times 10^9$\\
+	    \hline
+	  \end{tabular}
+	  \label{tab:input_parameters}
+	\end{centering}
+\end{table}
 
@@
 
-A three dimensional axisymmetric JPUE which ejects from a round vent is simulated and results are compared with experiments \citep {george1977turbulence, papanicolaou1988investigations} for verification purpose.
-Experimental data of concentration and velocity distribution across the cross-section was fit into a Gaussian profile: $\dfrac{\varphi}{\varphi_c}=exp[-coef (\dfrac{r}{z})^2)]$ (solid line) even though there is no priori reason.
+A three dimensional axisymmetric JPUE which ejects from a round vent is simulated with eruption parameters list in Table \ref{tab:input_parameters}.
+Material properties of water are used as material properties for both phases.
+The results are compared with experiments \citep {george1977turbulence, papanicolaou1988investigations} for verification purpose.
+Experimental data of concentration and velocity distribution across the cross-section was fit into a Gaussian profile: $\dfrac{\varphi}{\varphi_c}=exp[-coef (\dfrac{r}{z})^2)]$ (solid line) by \citet{papanicolaou1988investigations} even though there is no priori reason.
 \citet{papanicolaou1988investigations} also fit concentration distribution and jet width based on velocity along central line into a straight line : $\dfrac{\varphi_0}{\varphi_c}=slope (\dfrac{z}{D} + intercept)$ (dash line).
@@
 A strong eruption column without wind is tested in this section for the purpose of further verification and validation.
-Input parameters, material properties and atmosphere are chosen to be the same as the strong plume no wind case in an inter-comparison study on eruptive column models \citep{costa2016results}.
-The strong plume scenario was based on climactic phase of the Pinatubo eruption (Philippines, 15 June 1991).
-Material properties are selected based on properties of Pinatubo and Shinmoe-dake eruption.
 Both global variable and local variables are compared with existing models.
-Then plume evolution process is investigated.
+
+\subsubsection{Input parameters}
+Eruption parameters, material properties and atmosphere are chosen to be the same as the strong plume no wind case in an inter-comparison study on eruptive column models \citep{costa2016results}.
+Eruption condition is list in Table \ref{tab:input_parameters}.
+The density of erupted material at the vent and radius of the vent can be computed from given parameters.
+The eruption pressure is assumed to be as the same as pressure of ambient at the vent and hence these parameters are not given in the table.
+The vertical profiles of atmospheric properties were obtained based on the reanalysis data from ECMWF (European Centre for Medium-Range Weather Forecasts) following \citet{costa2013density} for climactic phase of the Pinatubo eruption (Philippines, 15 June 1991).
+These conditions are more typical of a tropical atmosphere (see Fig.
+1B in \citep{costa2016results}).
+Wind speed is not used in our current model even though it is given (by one component speed from west to east and another component speed from south to north).
+Vertical distribution of temperature, pressure and density is used to establish stratified atmosphere.
+Material properties, shown in Table \ref{tab:material_properties}, are selected based on properties of Pinatubo and Shinmoe-dake eruption.
+Other material properties not given in the table can be infered from these given parameters based on their relationships.
+
+\begin{table}[htp]
+	\begin{centering}
+      \caption{List of material properties}		
+	  \begin{tabular}{lrrr}
+	    \hline
+	    Parameters & Units  & Value \\
+	    \hline
+	    	Specific heat of gas at constant volume     & $J \cdot kg^{-1}\cdot K^{-1}$& 717     \\
+	    Specific heat of air at constant volume     & $J \cdot kg^{-1}\cdot K^{-1}$& 1340    \\
+	    	Specific heat of solid                      & $J \cdot kg^{-1}\cdot K^{-1}$& 1100    \\
+	    	Specific heat of gas at constant pressure   & $J \cdot kg^{-1}\cdot K^{-1}$& 1000    \\
+	    	Specific heat of air at constant pressure   & $J \cdot kg^{-1}\cdot K^{-1}$& 1810    \\
+	    	Density of air at vent height               & $kg \cdot m^{-2}$       & 1.104   \\
+	    Pressure vent height                        & $Pa$              & 84363.4 \\
+	    \hline
+	  \end{tabular}
+	  \label{tab:material_properties}
+	\end{centering}
+\end{table}
 
@@
 One of the key global quantities of great interest is the altitude to which the plume rises.
-The top height predicted by our model is around 40 km which agrees well with other plume models.
+The top height predicted by our model is around 40 km which agrees with other plume models.
 For example, the height predicted by PDAC is 42500 m, by SK-3D is 39920 m, by ATHAM is 33392 m, by AHSEE is 36700 m.
 As for local variables, the profiles of integrated temperature, density, mass fraction of entrained air, gas mass fraction, mass fraction of solid and the radius of the plume as a function of height are compared with existing 1D and 3D models in Fig.
-\ref{fig:strong_local_temp} ~ \ref{fig:strong_local_radius}.
+\ref{fig:strong_local_temp} $\sim$ \ref{fig:strong_local_radius}.
 To get rid of significant fluctuations in time and space we did time averaging and spatial integration of the dynamic 3D flow fields by following \citet {cerminara2016large}.
@@
 Because air occupies a larger portion of the gas and plays a more significant role.
-As for mass fraction of solid, similarly, we have Eq.
-(\ref{eq:solid-frac-express1}) or Eq.
-(\ref{eq:solid-frac-express2}).
-PDAC which treat particles of two different sizes as two separate phases got similar mass fraction profile.
-Which implies that our model is at least valid for eruptions similar to the test case.
+As for mass fraction of solid, similarly, Eq.
+(\ref{eq:solid-frac-express1}) and Eq.
+(\ref{eq:solid-frac-express2}) hold for our model.
+PDAC, which treat particles of two different sizes as two separate phases, got similar mass fraction profile.
+That implies that our model is at least valid for eruptions similar to the test case.
+
 \begin{equation}
@@
 \end{equation}
+
 With more cool air entrained into the plume and mixing with hot erupted material, the temperature of the plume decreases as the height increase as shown in Fig.
@@
 However, there is still a big disparity between the profiles of local variables of our model and 3K-3D.
-One of the big differences between our model and 3K-3D is that we adopt a LANS type of turbulence model while 3K-3D adopts a LES (large eddy simulation) turbulence model.
-This implies that turbulence model plays a critical role in plume simulation.
-
-\subsubsection{Evolution of plume}
-
-\begin{figure}
-\center
-\includegraphics[width=15cm]{strong_elevation}
-\caption{This figure shows the dynamic evolution process of the Pinatubo with time.
+One of the big differences between these two model is that we adopt a LANS type of turbulence model while 3K-3D adopts a LES (large eddy simulation) turbulence model.
+This implies that turbulence model might play a critical role in plume simulation.
+
+%\subsubsection{Evolution of plume} 
+%
+%\begin{figure}
+%\center
+%\includegraphics[width=15cm]{strong_elevation}
+%\caption{This figure shows the dynamic evolution process of the Pinatubo with time.
 At the very beginning, it is driven by initial momentum and rises up quickly, mixing and entraining surround air.
@@
 The plume keeps rising up until reaches the top height then starts spreading.}
-\label{fig:strong_evolution}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=10cm]{t50}
-\caption{Contour of mass fraction and velocity quiver at 50 seconds in plane x=0, the column keeps rising up and at the same time entraining air into the plume.
+%\label{fig:strong_evolution}
+%\end{figure}
+%
+%\begin{figure}
+%\includegraphics[width=10cm]{t50}
+%\caption{Contour of mass fraction and velocity quiver at 50 seconds in plane x=0, the column keeps rising up and at the same time entraining air into the plume.
 The zoomed view (right figure) of the quiver plot clearly shows mechanism of air entrainment.
 The contour levels of mass fraction are 0.7, 0.5, 0.3, 0.1, $1\times 10 ^{-1}$, $1\times 10 ^{-2}$, $1\times 10 ^{-3}$, $1\times 10^{-4}$ and $1\times 10^{-5}$.}
-\label{fig:strong_plume_t50}
-\end{figure}
-
-Figure \ref{fig:strong_evolution} shows the overall evolution process of plume.
+%\label{fig:strong_plume_t50}
+%\end{figure}
+%
+%Figure \ref{fig:strong_evolution} shows the overall evolution process of plume.
 Plume keeps rising up and entraining air into the plume until it reaches its top height(around t=300 seconds).
@@
 Currently existing 3D models focus on certain aspect of the volcanic plume (PDAC on pyroclastic flow, ATHAM on microphysics, and SK-3D on entrainment with higher accuracy and higher order of accuracy) and hence, naturally, different assumptions were made in these models.
-However, these different aspects of volcanic plume are not independent and actually are coupled.
+However, these different aspects of volcanic plume are not independent and actually coupled.
 For example, it has been illustrated that gas-particle non-equilibrium would introduce a previously unrecognized jet-dragging effect, which imposes great influence on plume development, especially for weak plume \cite{cerminara2016large}.

### 2017-03-18 Manuscripts_GMD.tex
modification made based according to dr Pitman's comments

@@
 We also propose feasible ways to impose velocity inlet and pressure outlet boundary conditions, both of which are scarce in traditional implementations of SPH.
-The core solver of our model is parallelized with MPI (message passing interface) obtaining good weak scalability and strong scalability.
+The core solver of our model is parallelized with MPI (message passing interface) obtaining good weak-scalability and strong-scalability.
 The model is first verified by comparing velocity and concentration distribution along axis and in cross section with experimental results of JPUE (jet or plume that is ejected from a nozzle into a uniform environment).
@@
 Primary hazards associated with explosive volcanic eruptions include pyroclastic density currents (flows and surges), the widespread deposition of airfall tephra, and the threats to aviation posed by volcanic ash in the atmosphere.
-Simulation of all possible hazards with one model is difficult due to diversity among these hazards. Our focus here is the hazards that volcanic ash poses to aircraft and communities.
-
-During volcanic eruptions, VATDs (volcanic ash transport and dispersal models) are used to forecast the location and movement of ash clouds over hours to days.
+Simulation of all possible hazards with one model is difficult due the fact that different physics of different length scales are dominant in different hazards.
+Our focus here is the hazards that volcanic ash poses to aircraft.
+
+During volcanic eruptions, VATDs (volcanic ash transport and dispersal models) are used to forecast the location and movement of ash clouds at timescales that range from hours to days.
 VATDs use eruption source parameters, such as plume height, mass eruption rate, duration, and the mass fraction distribution of erupted debris finer than about $4 \Phi$ (or $63  \mu$m), which can remain in the cloud for many hours or days.
@@
 \subsection{Existing plume models}
-Several 1D (one dimensional) volcanic plume models have been developed in the past few decades.
-Ranging from the most basic 1D model \citep{woods1988fluid} which only accounts for mass conservation to more recently developed 1D models  \citep{bursik2001effect, mastin2007user, degruyter2012improving, woodhouse2013interaction, devenish2013using, de2015plume, folch2016fplume, pouget2016sensitivity} which tend to account for more complex physics effects.
+Several 1D (one dimensional) volcanic plume models have been developed in the past few decades, ranging from the most basic 1D model \citep{woods1988fluid} which only accounts for mass conservation to more recently developed 1D models  \citep{bursik2001effect, mastin2007user, degruyter2012improving, woodhouse2013interaction, devenish2013using, de2015plume, folch2016fplume, pouget2016sensitivity} which tend to account for more comprehensive physics effects.
 FPLUME-1.0 \citep{folch2016fplume} accounts for wind effect, entrainment of moisture, water phase change, particle fallout and re-entrainment and even wet aggregation of ash.
@@
 
-Recently, a first order, non-equilibrium compressible 3D model, ASHEE \citep{cerminara2016ashee}, was introduced based on three dimensional N phases Eulerian transportation equation.
+Recently, a first order, non-equilibrium compressible 3D model, ASHEE \citep{cerminara2016ashee}, was introduced based on three dimensional N-phases Eulerian transportation equation.
 ASHEE is valid for low concentration and low Stokes number region and much faster than N-phase Eulerian model.
@@
 \end{itemize}
-These  will be enabled by the following features of SPH:
+These are enabled by the following features of SPH:
 \begin{itemize}
@@
 (\ref{eq:gov-nc-e})).
-So advection term will be treated exactly with SPH \citep{monaghan2005smoothed}.
+So advection term is treated exactly with SPH \citep{monaghan2005smoothed}.
 \item It is easy to include various physics effect (like self gravity, radiative cooling and chemical reaction) into the model.
@@
 Adding of new phases to the model also does not require a major overhaul and re-tooling.
-As will be shown in later paragraphs, adding of new phases will lead to adding of several lines into the source code for new phases and additional interaction terms between existing phases and newly added phases.
+As will be shown in later paragraphs, adding of new phases only leads to adding of several lines into the source code for new phases and additional interaction terms between existing phases and newly added phases.
 \end{itemize}
@@
 
-Open source availability and  the easy extension features of SPH will facilitate development of a more comprehensive community driven model.
+Open source availability and  the easy extension features of SPH facilitate development of a more comprehensive community driven model.
 
 \subsection{Our contributions}
-As SPH is a relatively new method in computational fluid dynamics, implementations of SPH for compressible multiphase turbulent flows are few.
+SPH is still a relatively new method in computational fluid dynamics compared to mesh-based methods, like Finite Volume.
+Implementations of SPH for compressible multiphase turbulent flows are few.
 Several issues endemic to classical SPH necessary for use in this domain are fixed in our implementation.
@@
 Cauliflower-like vortices are generated along jet margins, within which process, air is entrained and heated up, reducing the bulk density of the entire jet, in many cases, to less than that of the surrounding atmosphere.
-Once it becomes buoyant, such a jet develops into a plinian or subplinian plume, rising kilometers to tens of kilometers until its heat is exhausted that buoyancy is lost.
+Once it becomes buoyant, such a jet develops into a plinian or subplinian plume, rising up to several kilometers to tens of kilometers until its heat exhausted.
 Jets that lose their momentum before becoming buoyant collapse back onto the ground and transform into pyroclastic flows, surges and ignimbrites.
-When plume rises up, relatively coarse particle might separate from main stream of the plume falling down onto the ground and probably be re-entrained into the plume at a lower height \citep{ernst1996sedimentation}.
-Within this process, erupted vapor will condense to liquid (droplet) and even further to ice.
-Latent heat released from phase change of erupted vapor will further heat up entrained air and make the bulk density further dilute.
+During the process of plume rising up, relatively larger particles might separate from main stream of the plume, falling down onto the ground and probably be re-entrained into the plume at a lower height \citep{ernst1996sedimentation}.
+Within this process, erupted vapor condenses to liquid (droplet) and even further to ice.
+Latent heat released from phase change of erupted vapor  further heats up entrained air and make the bulk density further dilute.
 The entrained vapor might also experience a similar process and impose impact on plume development.
@@
 Immediate thermodynamics equilibrium is assumed so that no separate energy equation is needed for each phase.
-As a result, there is only one energy equation for both phases (heat exchange term between different phases will not show up under this assumption).
+As a result, there is only one energy equation for both phases (heat exchange term between different phases does not show up under this assumption).
 Immediate dynamics equilibrium is assumed so that no separate momentum equation is needed for each phase.
-As a result, there is only one vector momentum equation for both phases (drag force term will not show up with this assumption).
+As a result, there is only one vector momentum equation for both phases (drag force term does not show up with this assumption).
 
 Because of the above assumptions, all other microphysical processes (such as the phase changes of \texorpdfstring{H\textsubscript{2}O}, aggregation, disaggregation, absorption of gas on the surface of solids, solution of gas into a liquid) and chemical processes are not considered in this model.
-These ignored microphysics factors will come to play critical roles under particular eruption conditions.
+These ignored microphysics factors would play critical roles under particular eruption conditions.
 Capturing of these processes needs a more comprehensive model.
 One critical element in plume development, the effect of wind, is also not considered in this model yet.
-Introducing of wind effect in our model will need imposing of pressure boundary conditions in a dynamic way, which requires more numerical effort and algorithm design.
+Introducing of wind effect in our model need imposing of pressure boundary conditions in a dynamic way, which requires more numerical effort and algorithm design.
 To summarize, our model is not valid for eruptions where wind effect plays a significant role in its development, usually refered as weak plume.
@@
 (\ref{eq:gov-cs-rho}) to Eq.
-(\ref{eq:gov-cs-e}), will be directly discretized.
+(\ref{eq:gov-cs-e}), are directly discretized.
 For SPH, governing equations in Lagrange form is needed.
@@
 \end{align}
-Governing equations in Lagrangian form, Eq.
+%Governing equations in Lagrangian form, Eq.
 (\ref{eq:gov-nc-rho}) to Eq.(\ref{eq:gov-nc-e}), and Eulerian form, Eq.
@@
 The flux of mass should also be zero.
-As a result, internal energy flux, which consists of heat flux and energy flux carried by mass flux, will vanish on the wall boundary.
+As a result, internal energy flux, which consists of heat flux and energy flux carried by mass flux, vanishes on the wall boundary.
 Equation (\ref{eq:wall_bc_rho}) to (\ref{eq:wall_bc_e}) gives no-slip wall boundary condition written in terms of primitive variables.
@@
 Except for the pressure, boundary values for density, velocity, and energy on the outlet should depend on the solution.
-As we ignore the viscosity, the shear stress is ignored and normal stress (whose magnitude equals to pressure) will balance the ambient pressure.
+As we ignore the viscosity, the shear stress is ignored and normal stress (whose magnitude equals to pressure) balances the ambient pressure.
 \begin{equation}
@@
 Each kernel function has a compact support determined by smoothing length of each particle.
-There are several review papers by \citet{monaghan1992smoothed, monaghan2005smoothed, rosswog2009astrophysical, price2012smoothed, monaghan2012smoothed}, giving a pretty comprehensive view over SPH, we will only refer here to the representation of the constitutive equations in SPH and put more focus on specific numerical techniques for plume modeling.
+There are several review papers by \citet{monaghan1992smoothed, monaghan2005smoothed, rosswog2009astrophysical, price2012smoothed, monaghan2012smoothed}, giving a pretty comprehensive view over SPH, we only refer here to the representation of the constitutive equations in SPH and put more focus on specific numerical techniques for plume modeling.
 
@@
 The Dirac function in Eq.
-(\ref{eq:Dirac-translation}) can be approximated by a weighting function $w(\textbf{x}-\textbf{x}\prime, h)$ (or $w(\textbf{x}\prime-\textbf{x}, h)$) which will tend to a Dirac function when $h \rightarrow 0$ :
+(\ref{eq:Dirac-translation}) can be approximated by a weighting function $w(\textbf{x}-\textbf{x}\prime, h)$ (or $w(\textbf{x}\prime-\textbf{x}, h)$) which tends to a Dirac function when $h \rightarrow 0$ :
 \begin{equation}
@@
 \end{equation}
-The function $A(\textbf{x})$ then can be approximated by:
+An arbitrary function $A(\textbf{x})$ can then be approximated by:
 \begin{equation}
@@
 \end{equation}
-Besides normalization, weighting function of particle $a$ has to be symmetric with respect to $a$ to ensure that neighbor particles locate at the same distance away will contribute equally to SPH summation equation, see Eq.
+Besides normalization, weighting function of particle $a$ has to be symmetric with respect to $a$ to ensure that neighbor particles locate at the same distance away from $a$ contribute equally to SPH summation equation, see Eq.
 (\ref{eq:SPH-kernel-symmetric}).
@@
 The problem is that using these expressions ``{as is}" in general leads to quite poor gradient estimates.
-Various tricks are used to conserve linear and angular momentum and thermal energy \citep{monaghan1992smoothed}.
+Various tricks can be used to conserve linear and angular momentum and thermal energy \citep{monaghan1992smoothed}.
 Special treatments are also needed for second order derivative terms.
@@
 <\rho_a> = \sum_b m_b w_{ab} (h) \label{eq:ns-sph-d} \\
-<\dfrac{d \textbf{v}_a}{d t}>= \sum_b m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}) \nabla_a w_{a b}(h) +\textbf{g} \label{eq:ns-sph-v} \\
+<\dfrac{d \textbf{v}_a}{d t}>= \sum_b m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}^{\beta}) \nabla_a w_{a b}(h) +\textbf{g} \label{eq:ns-sph-v} \\
 <\dfrac{d e_a}{d t}>=
- 0.5\sum_b m_b \textbf{v}_{a b}(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}) \cdot \nabla_a w_{a b}(h) \label{eq:ns-sph-e}
+ 0.5\sum_b m_b \textbf{v}_{a b}(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}^{\beta}) \cdot \nabla_a w_{a b}(h) \label{eq:ns-sph-e}
 \end{align}
 where, $a$ is the SPH particle index, $\textbf{v}_{a b} = \textbf{v}_a - \textbf{v}_b$. $\Pi$ is an artificial viscosity term, which will be discussed in section \ref{sec:artificial-viscosity}.
-However, these discretized formulations will not guarantee conservation of entropy.
+%However, these discretized formulations will not guarantee conservation of entropy.
 We can obtain conservation of entropy by giving up on conservation of mass or thermal energy adopting  alternative discretized formulations \citep{monaghan1992smoothed}.
-As a Lagrangian method, particle position will also be updated at every time step.
+As a Lagrangian method, particle position is also updated at every time step.
 \begin{equation}
@@
 (\ref{eq:ns-sph-v})  and adding a heat exchange term into Eq.
-(\ref{eq:ns-sph-e}) will lead to the new discretization form:
+(\ref{eq:ns-sph-e}) leads to the new discretization form:
 \begin{align}
 <\rho_a> = \sum m_b w_{ab} (h) + \dot{\rho}(\textbf{x},t)\label{eq:ns-source-sph-d} \\
-<\dfrac{d \textbf{v}_a}{d t}>= \sum_b m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}) \nabla_a w_{a b}(h) +\textbf{g} + D \sum	_b m_b \dfrac{\textbf{v}_b - \textbf{v}_a}{\rho_b} \label{eq:ns-drag-sph-v} \\
+<\dfrac{d \textbf{v}_a}{d t}>= \sum_b m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}^{\beta}) \nabla_a w_{a b}(h) +\textbf{g} + D \sum	_b m_b \dfrac{\textbf{v}_b - \textbf{v}_a}{\rho_b} \label{eq:ns-drag-sph-v} \\
 <\dfrac{d e_a}{d t}>=
- 0.5\sum_b m_b \textbf{v}_{a b}(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}) \cdot \nabla_a w_{a b}(h) + \sum_b \dfrac{m_b}{\rho_b}(\kappa_a + \kappa_b) \dfrac{(T_a - T_b)}{\textbf{r}_a - \textbf{r}_b} w_{ab}(h) \label{eq:ns-conduction-sph-e}
+ 0.5\sum_b m_b \textbf{v}_{a b}(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}^{\beta}) \cdot \nabla_a w_{a b}(h) + \sum_b \dfrac{m_b}{\rho_b}(\kappa_a + \kappa_b) \dfrac{(T_a - T_b)}{\textbf{r}_a - \textbf{r}_b} w_{ab}(h) \label{eq:ns-conduction-sph-e}
 \end{align}
@@
 Other physics can be added easily in a similar way.
-Adding of these new terms will lead to modification of several lines in the source code.
+Adding of these new terms leads to modification of several lines in the source code.
 The drag force term should show up only when dynamic dis-equilibrium between different phases is considered.
-In that case, each phase will need one set of governing equations of Navier-Stokes type.
+In that case, each phase needs one set of governing equations of Navier-Stokes type.
 Adding of new phases into SPH code only needs few new lines for the new phase besides adding  interaction terms.
@@
 However, such a formulation has two disadvantages: First, it is very sensitive to particle disorder.
-Second, the second derivative of the kernel can change sign and will lead to unphysical representations (viscous dissipation causes decrease of the entropy).
+Second, the second derivative of the kernel can change sign and lead to unphysical representations (viscous dissipation causes decrease of the entropy).
 
@@
 The artificial viscosity term $\Pi_{ab}$ is a Galilean invariant and vanishes for rigid rotation.
-It will produce a repulsive force between two particles when they are approaching each other and an attractive force when they are receding from each other.
+It produces a repulsive force between two particles when they are approaching each other and an attractive force when they are receding from each other.
 
@@
 \begin{equation}
-\Pi_{ab} = 
+\Pi_{ab}^{\beta} = 
 \begin{cases} 
@@
 \end{equation}
-Equation \ref{eq:CSP-function-approximation-1d} implies that $\sum_b m_b \dfrac{1}{\rho_b} w(x-x_b, h) = 1$, which can be viewd as the approximation form of Eq.
+Equation \ref{eq:CSP-function-approximation-1d} implies that $\sum_b m_b \dfrac{1}{\rho_b} w(x-x_b, h) = 1$, which can be viewed as the approximation form of Eq.
 \ref{eq:SPH-kernel-normalization-prop}.
@@
 Take the governing equation adopted in our model as an example.
-Different phases of fluid (air and erupted material) will be represented by two different sets of SPH particles (or discretization points).
+Different phases of fluid (air and erupted material) are represented by two different sets of SPH particles (or discretization points).
 Based on assumption we made in section \ref{sec:physics-model}, only density need to be updated respectively for each phase.
@@
 (\ref{eq:ns-sph-d}) in spirit.
-Particles of phase 1 will not be counted while evaluating density of phase 2 and vice verse.
+Particles of phase 1 are not be counted while evaluating density of phase 2 and vice verse.
 Updating of density is then based on the following discretized equations.
@@
 \ref{fig:SPH-multiple-density}, where there are only air particles, Eq.
-(\ref{eq:gov-sph-d2}) will lead to zero and total density will be:
+(\ref{eq:gov-sph-d2}) evaluates to zero and total density is:
 \begin{equation}
@@
 (\ref{eq:CSP-function-approximation-1d}).
-For these areas occupied by only particles of phase 2 and far away from interface, similarly, the equation for density update will become:
+For these areas occupied by only particles of phase 2 and far away from interface, similarly, the equation for density update becomes:
 \begin{equation}
@@
 An improved version of $SPH-\alpha$ method is proposed by \citet{monaghan2011turbulence}, which we refer to as $SPH-\varepsilon$ method.
-We will adopt $SPH-\varepsilon$ method in our model.
+We adopt $SPH-\varepsilon$ method in our model.
 Most of these papers constrain their focus to incompressible flow.
@@
 
-To simulate turbulence in plume, we will thus need to extend existing incompressible turbulent modeling methods to compressible flow.
+To simulate turbulence in plume, we need thus to extend existing incompressible turbulent modeling methods to compressible flow.
 For compressible flow, energy equation is coupled with momentum equation and mass conservation equation.
-Averaging of thermal energy over space will introduce some additional terms besides the Reynolds stress term into energy equation \citep{NASACompressibleTurbulence}.
+Averaging of thermal energy over space introduces some additional terms besides the Reynolds stress term into energy equation \citep{NASACompressibleTurbulence}.
 The averaged momentum equation for compressible flow are in the same form as that for incompressible flow, all of other additional terms besides a corresponding turbulent stress term are in the energy equation.
@@
 It is common practice in $LANS-\alpha$ to use a differential equation for the smoothing rather than the integral form and finally reach to system of equations that need to be solved implicitly.
-In $SPH-\varepsilon$ method, a XSPH smoothing is adopted which will conserve linear and angular momentum.
+In $SPH-\varepsilon$ method, a XSPH \citep{monaghan1989problem} smoothing is adopted which conserves linear and angular momentum.
 In this way solving of system of equations is avoided and it also makes the method simple to implement and cheap for computation.
@@
 \end{equation}
-As function $G$ has the same feature as kernel function $w$, SPH approximation of the integration will lead to:
+As function $G$ has the same feature as kernel function $w$, SPH approximation of the integration leads to:
 \begin{equation} \label{eq:SPH-epsilon-filtering}
@@
 \end{equation}
-So the discretized momentum equation with $SPH-\varepsilon$ turbulence model in our simulation will be:
+So the discretized momentum equation with $SPH-\varepsilon$ turbulence model in our simulation are:
 \begin{equation}
@@
 \end{equation}
-Then the turbulent viscosity coefficient will be
+Then the turbulent viscosity coefficient becomes
 \begin{equation}
@@
 \end{equation}
-Please note that turbulent viscosity term has opposite sign with molecular viscosity term in discretized momentum equation and there is a minus sign in the expression of $\Pi_{ab}$, they will cancell out.
+Please note that turbulent viscosity term has opposite sign with molecular viscosity term in discretized momentum equation and there is a minus sign in the expression of $\Pi_{ab}$, they cancell out.
 
@@
 \end{equation}
-And  $F_{ab} \leq 0$, which guarantees that heat flux will flow from hot to cold.
+And  $F_{ab} \leq 0$, which guarantees that heat flux flows from hot to cold.
 %JJ Monaghan \citep{cleary1999conduction} showed a simple alteration to the standard SPH formulation and ensures continuity of heat flux across discontinuities in material properties. 
@@
 As the particle-scale movement of flow is based on smoothed velocity, the velocity in the energy equation should also be smoothed.
-The filtering process will be done according to Eq.
+The filtering process is done according to Eq.
 (\ref{eq:SPH-epsilon-filtering}).
-Position of particles will be updated according to Eq.
+Position of particles is updated according to Eq.
 (\ref{eq:gov-update-pos-turbulence}).
-Smoothed velocity is also used while computing artificial viscosity $\Pi_{ab}$.
+Smoothed velocity is also used while computing artificial viscosity.
+% $\Pi_{ab}$.
 
@@
 Ghost particles are reflected into the domain and physical quantities are calculated at these reflected positions by SPH interpolations.
-It should be noted that ghost particles are not used for these interpolations.
-Assign all properties (except for velocity) at the corresponding reflected position to the ghost particle.
+It should be noted that wall ghost particles should not be counted when computing physics properties of these reflected positions.
+Assign all properties, except for velocity, at the corresponding reflected position to the ghost particle.
 The velocity of each wall ghost particles is set to have the same value but opposite direction of the interpolated velocity at its corresponding reflection.
@@
 (\ref{eq:wall_bc_v})) is imposed naturally.
-These wall ghost particles will serve as neighbors in momentum and energy update.
+These wall ghost particles serve as neighbors in momentum and energy update.
 More implementation details about this method can be found in \citep {kumar2013parallel}.
-As these wall ghost particles keep stationary, there will be no mass flux on the boundary (Eq.
+As these wall ghost particles keep stationary, there is no mass flux on the boundary (Eq.
 (\ref{eq:wall_bc_rho}) and (\ref{eq:wall_bc_xi})).
-In addition, as temperature is also symmetric with respect to the boundary, the gradient of temperature will vanish and hence there is no internal energy flux on the wall boundary (Eq.
+In addition, as temperature is also symmetric with respect to the boundary, the gradient of temperature vanishes and hence there is no internal energy flux on the wall boundary (Eq.
 (\ref{eq:wall_bc_e})).
@@
 As simulations progress, change in position and physical quantities of real particles near pressure boundaries might corrupt pressure boundary condition that was established initially.
-This shortcoming is relieved by choosing a larger computational domain so that boundaries that might be corrupted will locate far away from turbulent mixing area.
+This shortcoming is relieved by choosing a larger computational domain so that boundaries that might be corrupted locate far away from turbulent mixing area.
 In addition, to avoid enlarging fluctuations, we add another constraint on the time step:
@@
 Instead of using the "nature manner" to number particles, we adopt space filling curve (SFC) based index to give each particle and background bucket an unique identifier -- a strategy known to preserve data locality at minimal cost.
-The SFC based numbering strategy is further extended to include time step information so that particles added at the same position but different time will have different identifiers.
+The SFC based numbering strategy is further extended to include time step information so that particles added at the same position but different time have different identifiers.
 
@@
 \ref{fig:weak_scale}, simulation times increase around $1/3$ when number of cores increase from 16 to 4208.
-For the test problem in this section, the volcanic plume will finally reach to a region of $[-30km \,\,\, 30km] \times [-30km\,\,\,30km] \times [1.5km\,\,\,40km]$ after around 400 seconds of eruption.
+For the test problem in this section, the volcanic plume finally reaches to a region of $[-30km \,\,\, 30km] \times [-30km\,\,\,30km] \times [1.5km\,\,\,40km]$ after around 400 seconds of eruption.
 When numerical simulation goes up to 90 seconds, the plume is still within a region of $[-10km\,\,\,10km] \times [-10km\,\,\,10km] \times [0km\,\,\,25km]$.
@@
 There exist consistently good experimental data \citep { list1982turbulent,dimotakis1983structure, papanicolaou1988investigations} that describe the JPUE flow field giving an insight into details of JPUE, such as transverse velocity and concentration profile.
-In this section, we will verify that our code and the $SPH-\varepsilon$ turbulence model is able to reproduce feature of turbulent entrainment by a JPUE simulation.
+In this section, we verify that our code and the $SPH-\varepsilon$ turbulence model is able to reproduce feature of turbulent entrainment by a JPUE simulation.
 
@@
 We set the temperature and density of ejected material the same as surrounding ambient.
-This will further simplify the scenario for the convenience of studying turbulent mixing.
+This further simplifies the scenario for the convenience of studying turbulent mixing.
 
@@
 $D$ is the diameter of vent.
-The coefficient $coef$ for concentration is 80 and 50 respectively according to Panos and George. $coef$ for velocity is 90 and 55 respectively according to Panos and George. 
+The coefficient $coef$ for concentration is 80 and 50 respectively according to \citet{george1977turbulence} and \citet{papanicolaou1988investigations}.
+$coef$ for velocity is 90 and 55 respectively according to \citet{george1977turbulence} and \citet{papanicolaou1988investigations}.
 $slope$ for jet width based on velocity is 0.104 and for concentration is 0.157. $intercept$ for jet width based on velocity is 2.58 while that for the concentration is 4.35.
@@
 The density of erupted material at the vent and radius of the vent can be computed from given parameters.
-The eruption pressure is assumed to be as the same as pressure of ambient at the vent and hence these parameters are not given in the table.
-The vertical profiles of atmospheric properties were obtained based on the reanalysis data from ECMWF (European Centre for Medium-Range Weather Forecasts) following \citet{costa2013density} for climactic phase of the Pinatubo eruption (Philippines, 15 June 1991).
+The eruption pressure is assumed to be as the same as pressure of ambient at the vent and hence is not given in the table.
+The vertical profiles of atmospheric properties were obtained based on the reanalysis data from ECMWF (European Centre for Medium-Range Weather Forecasts) for the period corresponding to climactic phase of the Pinatubo eruption (Philippines, 15 June 1991).
 These conditions are more typical of a tropical atmosphere (see Fig.
@@
 
-As the height increase, the amount of entrained air will increase.
+As the height increase, the amount of entrained air also increases.
 Around the neutral height, where the umbrella expanding, the entrainment of air shows a slight decrease due to lack of air surrounding the column at that height.
@@
 First of all, vapor condensation is not considered in our model.
-In addition, as we assume that erupted material behaves like a single phase water, the mass fraction of gas will simply be a function of entrained air (Eq.
+In addition, as we assume that erupted material behaves like a single phase water, the mass fraction of gas is simply a function of entrained air (Eq.
 (\ref{eq:gas-frac-express})).
@@
 Data for running the test cases, including meteorological data, properties of erupted material and air, and eruption condition data can be found in source code repository.
-Raw data of simulation results are archived in google drive.
+Raw data of simulation results are archived in Box.
 Access permission will be given upon request.

### 2017-03-23 Manuscripts_GMD.tex
jonesm changes, mostly small, some larger comments flagged in red

@@
 Primary hazards associated with explosive volcanic eruptions include pyroclastic density currents (flows and surges), the widespread deposition of airfall tephra, and the threats to aviation posed by volcanic ash in the atmosphere.
-Simulation of all possible hazards with one model is difficult due the fact that different physics of different length scales are dominant in different hazards.
-Our focus here is the hazards that volcanic ash poses to aircraft.
+Simulation of all possible hazards with one model is difficult due to the fact that different length scales dominate in different hazards.
+Our focus here is the hazard that volcanic ash poses to aircraft.
 
@@
 One of the earliest of these is the 3D model PDAC (Pyroclastic Dispersion Analysis Code) \citep{neri2003multiparticle}  which is a non-equilibrium, multiphase, 3D compressible flow model.
-Conservation equations for each phase are solved separately with finite volume method.
+Conservation equations for each phase are solved separately with a finite volume method.
 A parallel computing version of PDAC was developed lately \citep{ongaro2007parallel}.
@@
 
-Another 3D model, SK-3D \citep{suzuki2005numerical} is a 3D time-dependent fluid dynamics model that attempts to reproduce the entrainment process of eruption clouds with relatively simpler physics but with high order numerical accuracy and high spatial resolution.
-A series of simulations based on SK-3D was reported, including determination of the entrainment coefficients of eruption columns as a function of height \citep{suzuki2010numerical}, the effect of wind field on entrainment coefficient \citep{suzuki20133d}, establishment of the relationship between the observable quantities of the eruption clouds and the eruption conditions at the vent \citep{suzuki2009three} (previously, such relationship is based on 1D models and greatly depends on empirical constants), investigation of the effect of the intensity of turbulence in the umbrella cloud on dispersion and sedimentation of tephra \citep{koyaguchi2009effect}.
-
-While the SK-3D focuses on accurately capturing the entrainment caused by turbulent mixture with higher resolution and numerical method of higher order, PDAC takes the disequilibrium between different phases into account and hence is a true multiphase model.
-Another 3D model, ATHAM ( Active Tracer High-Resolution Atmospheric Model) \citep{oberhuber1998volcanic} focuses more on microphysics within the plume.
-As pyroclastic flow is not the initial concern of ATHAM, immediately dynamics and thermodynamics equilibrium is assumed in ATHAM.
+Another 3D model, SK-3D \citep{suzuki2005numerical} is a 3D time-dependent fluid dynamics model that attempts to reproduce the entrainment process of eruption clouds with relatively simple physics but with high order numerical accuracy and high spatial resolution.
+A series of simulations based on SK-3D was reported, including determination of the entrainment coefficients of eruption columns as a function of height \citep{suzuki2010numerical}, the effect of wind field on entrainment coefficient \citep{suzuki20133d}, establishment of the relationship between the observable quantities of the eruption clouds and the eruption conditions at the vent \citep{suzuki2009three} (previously, such relationship is based on 1D models and greatly depends on empirical constants), and investigation of the effect of the intensity of turbulence in the umbrella cloud on dispersion and sedimentation of tephra \citep{koyaguchi2009effect}.
+
+While SK-3D focuses on accurately capturing the entrainment caused by turbulent mixture with higher resolution and numerical method of higher order, PDAC takes the disequilibrium between different phases into account and hence is a true multiphase model.
+Another 3D model, ATHAM (Active Tracer High-Resolution Atmospheric Model) \citep{oberhuber1998volcanic} focuses more on microphysics within the plume.
+As pyroclastic flow is not the initial concern of ATHAM, \textcolor{red}{immediately dynamics} and thermodynamic equilibrium is assumed in ATHAM.
 The dynamic core of ATHAM solves the compressible Euler equations for momentum, pressure and temperature of the gas particle mixture \citep{oberhuber1998volcanic}.
@@
 Besides adding to their special strengths (ATHAM has been adding more and more microphysics, PDAC was extended to consider more phases), these models are learning from each other.
-PDAC developers have moved to include the effect of microphysics into the model while ATHAM developer have extended its ability to modeling pyroclastic flow.
+PDAC development has begun to include the effect of microphysics into the model while ATHAM development has extended its ability to modeling pyroclastic flow.
 Both are using finer and finer resolution.
@@
 ASHEE is valid for low concentration and low Stokes number region and much faster than N-phase Eulerian model.
-The model is based an open source numerical solver OpenFOAM adapting its unstructured finite volume solver.
-
-To summarize, each 3D model has its own focus based on problem of interest and modeling/numerical choices made.
+The model is based on the open source numerical solver OpenFOAM \textcolor{red}{cite?}, adapting its unstructured finite volume solver.
+
+To summarize, each 3D model has its own focus based on the problem of interest and modeling/numerical choices made.
  %Different simplification was made based on their scope. Improvements have been making since their birth. 
- Accuracy of simulation and forecast skill (depends on comprehensiveness of the model, resolution of discretization, numerical error, and order of accuracy) and simulation time (depends on number of governing equations, resolution, numerical methods and parallel techniques) are always a pair of conflict considerations in 3D plume simulations.\\
+ Accuracy of simulation and forecast skill (depending on comprehensiveness of the model, resolution of discretization, numerical error, and order of accuracy) and simulation time (depending on number of governing equations, resolution, numerical methods and parallel techniques) are always a pair of conflicting considerations in 3D plume simulations.\\
 
 \subsection{Features of SPH}
-All of these existing 3D models are mesh based Eulerian methods.
-To date there are no 3D plume models based on mesh free Lagrangian methods.
+To the bext of our knowledge, all of the existing 3D plume models are mesh based Eulerian methods, and there are no 3D plume models based on mesh free Lagrangian methods.
 Lagrangian methods have several features that are suitable for volcanic plume simulation that we outline below.
@@
 Easy extensibility and capability of handling multiples phase flow without additional numerical effort greatly facilitate future extension of SPH models.
-As volcanic plume are in nature multiphase and without pre-defined boundary in the atmosphere, SPH is a suitable numerical method for plume modeling.
-The core physics, such as entrainment of air and thermal expansion, are essential for all plume modeling while some other physics, such as water condensation and aggregation etc., are important in specific scenario.
+As volcanic plumes are in nature multiphase and without pre-defined boundary in the atmosphere, SPH is a suitable numerical method for plume modeling.
+The core physics, such as entrainment of air and thermal expansion, are essential for all plume modeling while some other physics, such as water condensation and aggregation etc., are important in specific scenarios.
 As an initial effort on exploiting advantages of SPH in volcanic plume modeling, we focus on capturing basic features in plume development using a robust numerically and comuptationaly efficient framework with support for scalable parallel computing.
@@
 
-Open source availability and  the easy extension features of SPH facilitate development of a more comprehensive community driven model.
+Open source availability and the relatively easy extensibility of SPH will facilitate development of a more comprehensive community driven model.
 
@@
 Several issues endemic to classical SPH necessary for use in this domain are fixed in our implementation.
-The most popular applications of SPH (and their original motivating application) has been in simulating of free surface flow, such as breaking-waves and flood.
+The most popular applications of SPH (and their original motivating application) has been in the simulation of free surface flow, such as breaking-waves and floods.
 Less attention was paid to velocity inlet and pressure outlet boundary conditions which are required in plume modeling.
@@
 \item We develop methodology to impose pressure boundary conditions by adding extra layers of static ghost particles.
-Additional constraints on time step is used to avoid the growth of numerical fluctuations near the pressure boundary.
-We impose velocity inlet boundary condition by placing several layers of ghost particles moving with eruption velocity.
+Additional constraints on the time step is used to avoid the growth of numerical fluctuations near the pressure boundary.
+We impose a velocity inlet boundary condition by placing several layers of ghost particles moving with eruption velocity.
 \item Turbulence model is crucial for reproducing the entrainment of air.
@@
 Cauliflower-like vortices are generated along jet margins, within which process, air is entrained and heated up, reducing the bulk density of the entire jet, in many cases, to less than that of the surrounding atmosphere.
-Once it becomes buoyant, such a jet develops into a plinian or subplinian plume, rising up to several kilometers to tens of kilometers until its heat exhausted.
+Once it becomes buoyant, such a jet develops into a plinian or subplinian plume, rising up to several kilometers to tens of kilometers until its heat is exhausted.
 Jets that lose their momentum before becoming buoyant collapse back onto the ground and transform into pyroclastic flows, surges and ignimbrites.
@@
 Within this process, erupted vapor condenses to liquid (droplet) and even further to ice.
-Latent heat released from phase change of erupted vapor  further heats up entrained air and make the bulk density further dilute.
-The entrained vapor might also experience a similar process and impose impact on plume development.
-Particle aggregation process \citep{carey1982influence,taddeucci2011aggregation}, either due to presence of liquid water, resulting from particle collision or driven by electrostatic forces might occur inside plume and thereby affect the sedimentation.
+Latent heat released from phase change of erupted vapor  further heats up entrained air and further dilutes the bulk density.
+The entrained vapor might also experience a similar process and impact plume development.
+Particle aggregation processes \citep{carey1982influence,taddeucci2011aggregation}, either due to presence of liquid water, resulting from particle collision or driven by electrostatic forces might occur inside plume and thereby affect the sedimentation.
 
@@
 Air (also a mixture of different gases) is assumed to be another phase (phase 1).
-Immediate thermodynamics equilibrium is assumed so that no separate energy equation is needed for each phase.
+Immediate thermodynamic equilibrium is assumed so that no separate energy equation is needed for each phase.
 As a result, there is only one energy equation for both phases (heat exchange term between different phases does not show up under this assumption).
-Immediate dynamics equilibrium is assumed so that no separate momentum equation is needed for each phase.
+Immediate dynamic equilibrium is assumed so that no separate momentum equation is needed for each phase.
 As a result, there is only one vector momentum equation for both phases (drag force term does not show up with this assumption).
@@
 Capturing of these processes needs a more comprehensive model.
-One critical element in plume development, the effect of wind, is also not considered in this model yet.
-Introducing of wind effect in our model need imposing of pressure boundary conditions in a dynamic way, which requires more numerical effort and algorithm design.
-To summarize, our model is not valid for eruptions where wind effect plays a significant role in its development, usually refered as weak plume.
+One critical element in plume development, the effect of wind, is also not yet considered in this model.
+Introducing wind effects in our model requires dynamic pressure boundary conditions, which requires more numerical effort and algorithm design.
+To summarize, our model is not valid for eruptions where wind effect plays a significant role in its development, usually refered as a weak plume.
 Our model also lacks the ability in modeling plumes with large particles or eruptions in which microphysics plays non-ignorable roles, such as a eruption of \textit{El Chich{\'o}n} volcano on April 4th 1982 \citep{sigurdsson19841982, folch2016fplume}.
@@
 \subsection{Governing equations}
-Based above assumptions, the governing equation of our model is given as following, which is the same as governing equations of SK-3D \citep{suzuki2005numerical}.
+Based above assumptions, the governing equation of our model is given as following, which is the same as the governing equations of SK-3D \citep{suzuki2005numerical}.
 \begin{align}
@@
 $E = e + K $ is total energy which is a summation of kinetic energy $K$ and internal energy $e$.
+\textcolor{red}{should define all variates when they first appear, even the ``obvious'' ones.}
 An additional equation is required to close the system. In this model, the equation for closing the system is the EOS (equation of state) for ideal gas.
@@
 \subsubsection{Velocity inlet}
-At the vent, temperature of erupted material $T$, eruption velocity $\textbf{v}$, the mass fraction of vapor in erupted material $n_{g0}$ and mass discharge rate $\dot M$ is given.
+At the vent, temperature of erupted material $T$, eruption velocity $\textbf{v}$, the mass fraction of vapor in erupted material $n_{g0}$ and mass discharge rate $\dot M$ are given.
 The pressure of erupted material $p$ is assumed to be the same as ambient pressure for pressure-balanced eruption.
@@
 \subsection{Weighting function}
-As described in previous section, the weighting function (or kernel function) is used to replace Dirac function and should satisfy Eq.
+As described in the previous section, the weighting function (or kernel function) is used to replace the Dirac function and should satisfy Eq.
 (\ref{eq:SPH_kernel_delta}).
-So it can be viewed as approximation form of Dirac function and satisfy normalization condition:
+So it can be viewed as an approximate form of the Dirac function and satisfy the normalization condition:
 \begin{equation}
@@
 \end{equation}
-Besides normalization, weighting function of particle $a$ has to be symmetric with respect to $a$ to ensure that neighbor particles locate at the same distance away from $a$ contribute equally to SPH summation equation, see Eq.
+Besides normalization, the weighting function of particle $a$ has to be symmetric with respect to $a$ to ensure that neighbor particles located at the same distance away from $a$ contribute equally to SPH summation equation, see Eq.
 (\ref{eq:SPH-kernel-symmetric}).
-Weighting function also needs to satisfy conditions such as positivity and compact support.
-In addition, kernel function must be monotonically decreasing with the distance between particles.\\
+The weighting function also needs to satisfy conditions such as positivity and compact support.
+In addition, the kernel function must be monotonically decreasing with the distance between particles.\\
 \begin{equation}
@@
 
-There is a wide variety of possible weighting functions, such as spline functions (with different orders) and Gaussian function.
+There is a wide variety of possible weighting functions, such as spline functions (with different orders) and Gaussian functions.
 Generally, the accuracy increases with the order of the polynomials of the kernel function, but the computational time also increases as number of interactions increase.
@@
 \begin{cases} 
-      \dfrac{1}{(h \sqrt{\pi})^d} exp [- (\dfrac{\textbf{x} - \textbf{x} \prime}{h})^2 ] &  \vert \textbf{x} - \textbf{x} \prime \vert \leq 3h\\
+      \dfrac{1}{(h \sqrt{\pi})^d} exp \left[- \left(\dfrac{\textbf{x} - \textbf{x} \prime}{h}\right)^2 \right] &  \vert \textbf{x} - \textbf{x} \prime \vert \leq 3h\\
       0 & \text{Otherwise}
@@
 Where $d$ is number of dimensions.
-The derivative of weighting function:
+The derivative of the weighting function:
 \begin{equation}
@@
 \begin{cases} 
-      -2(\dfrac{\textbf{x} - \textbf{x} \prime}{h^2}) \dfrac{1}{(h \sqrt{\pi})^d} exp [- (\dfrac{\textbf{x} - \textbf{x} \prime}{h})^2 ] &  \vert \textbf{x} - \textbf{x} \prime \vert \leq 3h\\
+      -2\left(\dfrac{\textbf{x} - \textbf{x} \prime}{h^2}\right) \dfrac{1}{(h \sqrt{\pi})^d} exp \left[- \left(\dfrac{\textbf{x} - \textbf{x} \prime}{h}\right)^2 \right] &  \vert \textbf{x} - \textbf{x} \prime \vert \leq 3h\\
       0 & \text{Otherwise}
@@
 <\rho_a> = \sum_b m_b w_{ab} (h) \label{eq:ns-sph-d} \\
-<\dfrac{d \textbf{v}_a}{d t}>= \sum_b m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}^{\beta}) \nabla_a w_{a b}(h) +\textbf{g} \label{eq:ns-sph-v} \\
-<\dfrac{d e_a}{d t}>=
- 0.5\sum_b m_b \textbf{v}_{a b}(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}^{\beta}) \cdot \nabla_a w_{a b}(h) \label{eq:ns-sph-e}
+\left\langle\dfrac{d \textbf{v}_a}{d t}\right\rangle = \sum_b m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}^{\beta}\right) \nabla_a w_{a b}(h) +\textbf{g} \label{eq:ns-sph-v} \\
+\left\langle\dfrac{d e_a}{d t}\right\rangle=
+ 0.5\sum_b m_b \textbf{v}_{a b}\left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}^{\beta}\right) \cdot \nabla_a w_{a b}(h) \label{eq:ns-sph-e}
 \end{align}
@@
 \begin{equation}
-<\dfrac{d \textbf{x}_a}{dt}> = \textbf{v}_a \label{eq:SPH-update-pos}
+\left\langle\dfrac{d \textbf{x}_a}{dt}\right\rangle = \textbf{v}_a \label{eq:SPH-update-pos}
 \end{equation}
@@
 <\rho_a> = \sum m_b w_{ab} (h) + \dot{\rho}(\textbf{x},t)\label{eq:ns-source-sph-d} \\
-<\dfrac{d \textbf{v}_a}{d t}>= \sum_b m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}^{\beta}) \nabla_a w_{a b}(h) +\textbf{g} + D \sum	_b m_b \dfrac{\textbf{v}_b - \textbf{v}_a}{\rho_b} \label{eq:ns-drag-sph-v} \\
-<\dfrac{d e_a}{d t}>=
- 0.5\sum_b m_b \textbf{v}_{a b}(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}^{\beta}) \cdot \nabla_a w_{a b}(h) + \sum_b \dfrac{m_b}{\rho_b}(\kappa_a + \kappa_b) \dfrac{(T_a - T_b)}{\textbf{r}_a - \textbf{r}_b} w_{ab}(h) \label{eq:ns-conduction-sph-e}
+\left\langle\dfrac{d \textbf{v}_a}{d t}\right\rangle= \sum_b m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}^{\beta}\right) \nabla_a w_{a b}(h) +\textbf{g} + D \sum	_b m_b \dfrac{\textbf{v}_b - \textbf{v}_a}{\rho_b} \label{eq:ns-drag-sph-v} \\
+\left\langle\dfrac{d e_a}{d t}\right\rangle=
+ 0.5\sum_b m_b \textbf{v}_{a b}\left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}^{\beta}\right) \cdot \nabla_a w_{a b}(h) + \sum_b \dfrac{m_b}{\rho_b}(\kappa_a + \kappa_b) \dfrac{(T_a - T_b)}{\textbf{r}_a - \textbf{r}_b} w_{ab}(h) \label{eq:ns-conduction-sph-e}
 \end{align}
@@
 \begin{equation}
-\Delta t = CFL \min_a \bigg \lbrace \dfrac{[\frac{m_a}{\rho_a}]^{\frac{1}{d}}}{c_a} \bigg \rbrace
-\end{equation}
-Where $c_a$ is sound speed at particle a, $d$ is number of dimensions. First order Euler time integration, with $CFL = 0.2$ has been used in three dimensional calculations.
+\Delta t = \textrm{CFL} \min_a \bigg \lbrace \dfrac{[\frac{m_a}{\rho_a}]^{\frac{1}{d}}}{c_a} \bigg \rbrace
+\end{equation}
+Where $c_a$ is sound speed at particle a, $d$ is number of dimensions. First order Euler time integration, with $\textrm{CFL} = 0.2$ has been used in three dimensional calculations.
 
@@
 Numerical simulation of multiphase flows is usually difficult due to the existence of complex evolving interfaces between phases.
-This present severe challenges to conventional Eulerian grid-based numerical methods.
-As we discussed in the introduction section, special algorithms to treat and track the interface between different phases is required for mesh based method to deal with multiphase problems.
-What's worse, the engulfment process during plume development makes geometry of interface very complicated.
-Lagrangian SPH is able to handle multiphase problems without additional numerical effort for dealing with interface by simply tagging particles of each phase, which makes it suitable for volcanic plume simulation.
+This presents severe challenges to conventional Eulerian grid-based numerical methods.
+As we discussed in the introduction section, special algorithms to treat and track the interface between different phases is required for mesh based methods to deal with multiphase problems.
+Worse, the engulfment process during plume development makes geometry of interfaces very complicated.
+Lagrangian SPH is able to handle multiphase problems without additional numerical effort for dealing with interfaces by simply tagging particles of each phase, which makes it suitable for volcanic plume simulation.
 Take the governing equation adopted in our model as an example.
 Different phases of fluid (air and erupted material) are represented by two different sets of SPH particles (or discretization points).
-Based on assumption we made in section \ref{sec:physics-model}, only density need to be updated respectively for each phase.
+Based on assumptions we made in section \ref{sec:physics-model}, only density needs to be updated respectively for each phase.
 The updating of density is exactly the same as Eq.
 (\ref{eq:ns-sph-d}) in spirit.
-Particles of phase 1 are not be counted while evaluating density of phase 2 and vice verse.
+Particles of phase 1 are not be counted while evaluating density of phase 2 and vice versa.
 Updating of density is then based on the following discretized equations.
@@
 
-In the areas far away from interface, updating of density is exactly the same as that for single phase flow.
+In areas far away from the interface, updating of density is exactly the same as that for single phase flow.
 For example, in the right side and left side (or blue areas) in Fig.
@@
 (\ref{eq:CSP-function-approximation-1d}).
-For these areas occupied by only particles of phase 2 and far away from interface, similarly, the equation for density update becomes:
+For these areas occupied by only particles of phase 2 and far away from the interface, similarly, the equation for density update becomes:
 \begin{equation}
@@
 The entrainment of air and heat transfer are dominated by such turbulent fluctuations as turbulent exchange coefficient would be several magnitudes larger than molecular viscosity (and heat conduction) coefficient.
-One way to include enough turbulence in the model is to use SPH-SPS (SPH sub-particle-scale) method.
+One way to include enough turbulence in the model is to use the SPH-SPS (SPH sub-particle-scale) method.
 %\citep {gotoh2004sph, shao2006sph}. 
-Another way is to use extremely fine resolution ("the direct numerical simulation method") at the expense of much higher computational cost that results from both increase in number of particles and decrease in time steps constrained by CFL condition.
+Another way is to use extremely fine resolution ("the direct numerical simulation method") at the expense of much higher computational cost that results from both increase in number of particles and decrease in time steps constrained by the CFL condition.
 Here we choose the SPH-SPS method.
@@
 
-To simulate turbulence in plume, we need thus to extend existing incompressible turbulent modeling methods to compressible flow.
-For compressible flow, energy equation is coupled with momentum equation and mass conservation equation.
+To simulate turbulence in the plume, we thus need to extend existing incompressible turbulent modeling methods to compressible flow.
+For compressible flow, the energy equation is coupled with the momentum equation and mass conservation equation.
 Averaging of thermal energy over space introduces some additional terms besides the Reynolds stress term into energy equation \citep{NASACompressibleTurbulence}.
@@
 In $SPH-\varepsilon$ method \citep{monaghan2011turbulence}, a turbulence model is constructed within the framework of SPH in such a way that general principles such as conservation of energy, momentum and circulation are satisfied using the ideas associated with the LANS (Lagrangian averaged Navier-Stokes equations).
-The basic idea of $SPH-\varepsilon$ is that determine a smoothed velocity $\widehat{\textbf{v}}$ by a linear operation on the unsmoothed velocity $\textbf{v}$.
+The basic idea of $SPH-\varepsilon$ is to determine a smoothed velocity $\widehat{\textbf{v}}$ by a linear operation on the unsmoothed velocity $\textbf{v}$.
 The SPH particles move with this smoothed velocity and the average motion of the fluid is determined by Eq.
@@
 \label{eq:monaghan-mom-turb}
-\dfrac{d \textbf{v}_a}{dt} = -\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}) \bigtriangledown_aw_{a b}(h)] + \sum_b m_b \dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{M} \bigtriangledown_a K_{ab}
+\dfrac{d \textbf{v}_a}{dt} = -\sum_b [m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}\right) \bigtriangledown_aw_{a b}(h)] + \sum_b m_b \dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{M} \bigtriangledown_a K_{ab}
 \end{equation}
@@
 \label{eq:SPH-mom-epsilon-turb}
-\dfrac{d \textbf{v}_a}{dt} = -\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}) \bigtriangledown_aw_{a b}(h)] + \sum_b m_b \Phi_{ab}\bigtriangledown_aG_{ab}(l)
+\dfrac{d \textbf{v}_a}{dt} = -\sum_b [m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}\right) \bigtriangledown_aw_{a b}(h)] + \sum_b m_b \Phi_{ab}\bigtriangledown_aG_{ab}(l)
 \end{equation}
@@
 
-However, the above equation is correct only for 1D situation. For 2D or 3D, it is not easy to get the explicit expression of viscosity. 
+However, the above equation is correct only for the 1D situation. For 2D or 3D, it is not easy to get the explicit expression of viscosity. 
 We adopt an alternative way: obtaining a value for each pair of particles instead of persisting on getting an analytical expression.
-Choosing smoothing function as the same as SPH kernel and the smoothing scale $l$ as the same as smoothing length $h$, the ratio between turbulent shear stress and physical shear stress is:
+Choosing the smoothing function as the same as SPH kernel and the smoothing scale $l$ as the same as smoothing length $h$, the ratio between turbulent shear stress and physical shear stress is:
 \begin{equation}
@@
 \end{equation}
-$\Upsilon_{ab}$ is essentially equivalent to ratio between turbulent viscous effect attributed by particle b on particle a and molecular viscous effect attributed by particle b on particle a.
+$\Upsilon_{ab}$ is essentially equivalent to the ratio between turbulent viscous effect attributed by particle b on particle a and molecular viscous effect attributed by particle b on particle a.
 Turbulent viscosity can be easily obtained by:
@@
 \begin{equation}
-<\dfrac{d \textbf{v}_{\alpha}}{d t}>= 
--\sum_b [m_b (\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}) \bigtriangledown_{\alpha}w_{\alpha b}(h)]
--\sum_j [m_j (\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j} - \Phi_{\alpha j}) \bigtriangledown_{\alpha}w_{\alpha j}(h)]
+\left\langle\dfrac{d \textbf{v}_{\alpha}}{d t}\right\rangle= 
+-\sum_b \left[ m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}\right) \bigtriangledown_{\alpha}w_{\alpha b}(h)\right]
+-\sum_j \left[m_j \left(\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j} - \Phi_{\alpha j}\right) \bigtriangledown_{\alpha}w_{\alpha j}(h)\right]
 +\textbf{g} \label{eq:gov-sph-v}
@@
 \begin{split}
-<\dfrac{d e_{\alpha}}{d t}>
-& = 0.5\sum_b [m_b \widehat{\textbf{v}_{\alpha b}} (\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}) \bigtriangledown_{\alpha}w_{\alpha b}(h)] 
+\left\langle\dfrac{d e_{\alpha}}{d t}\right\rangle
+& = 0.5\sum_b \left[m_b \widehat{\textbf{v}_{\alpha b}} \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}\right) \bigtriangledown_{\alpha}w_{\alpha b}(h)\right]
  + 2 \sum_b \dfrac{m_b}{\rho_{\alpha} \rho_b} \kappa_{t,\alpha b} (T_{\alpha} - T_b) F_{\alpha b} (h) \\
- & +0.5\sum_j [m_j \widehat{\textbf{v}_{\alpha b}}(\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j} - \Phi_{\alpha j}) \bigtriangledown_{\alpha}w_{\alpha j}(h)]
+ & +0.5\sum_j \left[m_j \widehat{\textbf{v}_{\alpha b}}\left(\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j} - \Phi_{\alpha j}\right) \bigtriangledown_{\alpha}w_{\alpha j}(h)\right]
  + 2 \sum_j \dfrac{m_j}{\rho_{\alpha} \rho_j} \kappa_{t,\alpha j} (T_{\alpha} - T_j) F_{\alpha j} (h)
@@
 \ref{fig:bc_and_domain_decomp} shows how boundaries are deployed.
-physical quantities of particles, including real particles and different types of ghost particles, are update differently.
+Physical quantities of particles, including real particles and different types of ghost particles, are updated differently.
 \subsubsection{Wall boundary condition}
@@
 The density, momentum, and internal energy of these eruption ghost particles are not updated before they move above ground.
-As soon as they move out from erupt conduit, these ghost particles will be shift to real particles and their physical quantities and position will be updated based on discretized governing equations.
+As soon as they move out from erupt conduit, these ghost particles will be shifted to real particles and their physical quantities and position will be updated based on discretized governing equations.
 New ghost particles need to be added at the bottom of the eruption conduit as these existing ghost particles move upwards.
@@
 
-As simulations progress, change in position and physical quantities of real particles near pressure boundaries might corrupt pressure boundary condition that was established initially.
+As simulations progress, changes in position and physical quantities of real particles near pressure boundaries might corrupt pressure boundary condition that was established initially.
 This shortcoming is relieved by choosing a larger computational domain so that boundaries that might be corrupted locate far away from turbulent mixing area.
@@
 The simulation is conducted on 360 processors (30 nodes with 12 cores each node), so there are 360 subdomains in total.
-The cut view shows portion of them}
+The cut view shows a portion of them.
+\textcolor{red}{For showing the domain decomposition in the 3rd figure, it may be better to simply choose a unique color for each one (or just show the boundaries in black and white)} }
 \label{fig:bc_and_domain_decomp}
@@
 \subsection{Parallellism and performance}
-One disadvantage of 3D model is that it usually takes much longer time than 1D models to complete one simulation.
+One disadvantage of the 3D model is that it usually takes a much longer time than 1D models to complete one simulation.
 This disadvantage further prevents doing simulation with finer resolution and accounting for more physics in one model.
-Non-intrusive uncertainty analysis which is commonly adopted in hazard forecasting requires finishing multiple simulations within a given time window.
-High performance computing provides a powerful routine for relieving the misery.
+Non-intrusive uncertainty analysis, which is commonly adopted in hazard forecasting, requires finishing multiple simulations within a given time window.
+High performance computing is therefore essential.
 Among existing CPU parallel SPH schemes, most of them focus on neighbor searching algorithm and dynamic load balancing \citep[eg.
@@
 Less attention has been paid to developing more flexible data management schemes for more complicated problems.
-Motivated by techniques developed for mesh based methods, we develop a complete framework for parallelizing SPH program with MPI standard model allowing flexible and efficient data access.
+Motivated by techniques developed for mesh based methods, we develop a complete framework for parallelizing SPH with distrubted memory parallelism (MPI) allowing flexible and efficient data access.
 
 Any implementation of SPH code requires efficient searching and updating of neighbors during simulation.
-Of the many choices possible for this we adopt background grid which was proposed by \citet {monaghan1985refined} and is quite popular in parallel SPH.
-The background grid is also used for domain decomposition in SPH.
+Of the many choices possible for this we adopt a background grid which was proposed by \citet {monaghan1985refined} and is quite popular in parallel SPH.
+This background grid is also used for domain decomposition in SPH.
 We refer to the elements of background grid, namely squares for two dimensions and cubic for three dimensions, as buckets.
@@
 We define C++ classes which contain all data of particles and buckets.
-As for the management of data, we adopt hash tables to store pointers to particles and buckets, which gives us not only flexibility of deleting and adding elements, but also quicker access compared with linked list.
-Instead of using the "nature manner" to number particles, we adopt space filling curve (SFC) based index to give each particle and background bucket an unique identifier -- a strategy known to preserve data locality at minimal cost.
+As for the management of data, we adopt hash tables to store pointers to particles and buckets, which gives us not only flexibility of deleting and adding elements, but also quicker access compared with linked lists.
+Instead of using the "nature manner" to number particles, we adopt a space filling curve (SFC) based index to give each particle and background bucket an unique identifier -- a strategy known to preserve data locality at minimal cost.
 The SFC based numbering strategy is further extended to include time step information so that particles added at the same position but different time have different identifiers.
@@
 
-We will provide a supplementary documentation describing details about the data structure, domain decomposition, load balancing and domain adjusting.
+We will publish separately further details about the data structure, domain decomposition, load balancing and domain adjusting.
 
@@
 
-Performance tests have been carried out on CCR (the computational cluster of Center for Computational Research) at Buffalo.
-E5645 CPUs running at 2.40GHz clock rate with 4GB memory per core on a Q-Logic Infiniband is used in these tests.
-Each node comprise of two sockets with 6 of these cores.
+Performance tests have been carried out at CCR (the computational cluster of Center for Computational Research) at Buffalo.
+Intel Xeon E5645 CPUs running at 2.40GHz clock rate with 4GB memory per core on a Q-Logic Infiniband is used in these tests.
+Each node is comprised of two sockets with 6 of these cores.
 Memory and level 3 cache are shared on each node.
@@
 Where $c$ is speed of sound in the liquid.
-Energy equation is actually decoupled from momentum and mass conservation equation by using this EOS.
+The energy equation is actually decoupled from momentum and mass conservation equation by using this equation of state (EOS).
 In addition, the "atmosphere" is assumed to be uniform and gravity is set to be zero.
@@
 
-A three dimensional axisymmetric JPUE which ejects from a round vent is simulated with eruption parameters list in Table \ref{tab:input_parameters}.
+A three dimensional axisymmetric JPUE which ejects from a round vent is simulated with eruption parameters listed in Table \ref{tab:input_parameters}.
 Material properties of water are used as material properties for both phases.
-The results are compared with experiments \citep {george1977turbulence, papanicolaou1988investigations} for verification purpose.
-Experimental data of concentration and velocity distribution across the cross-section was fit into a Gaussian profile: $\dfrac{\varphi}{\varphi_c}=exp[-coef (\dfrac{r}{z})^2)]$ (solid line) by \citet{papanicolaou1988investigations} even though there is no priori reason.
+The results are compared with experiments \citep {george1977turbulence, papanicolaou1988investigations} for validation n purposes.
+Experimental data of concentration and velocity distribution across the cross-section were fit into a Gaussian profile: $\dfrac{\varphi}{\varphi_c}=exp[-coef (\dfrac{r}{z})^2)]$ (solid line) by \citet{papanicolaou1988investigations} (even though there is no a priori reason to do so).
 \citet{papanicolaou1988investigations} also fit concentration distribution and jet width based on velocity along central line into a straight line : $\dfrac{\varphi_0}{\varphi_c}=slope (\dfrac{z}{D} + intercept)$ (dash line).
-Where $\varphi$ is either velocity or concentration, the subscript $c$ represents the centreline, subscript $0$ represents the cross-sectionally averaged exit value.
+$\varphi$ is either velocity or concentration, the subscript $c$ represents the centreline, subscript $0$ represents the cross-sectionally averaged exit value.
 $r$ is the distance from the centreline on any cross-section.
@@
 Reynolds number is not reported in many experiments assuming a high enough Reynolds number.
-In addition, some details of the experiments, such as exit velocity profile, viscosity of the experimental liquid, are not reported, neither.
+In addition, some details of the experiments, such as exit velocity profile and viscosity of the experimental liquid, are not reported.
 These factors prevent us from numerically reproducing these experiments in an exact way as they were conducted.
@@
 
-We also investigated the mixing due to turbulence in JPUE simulation by checking the mixture of two phases.
+We also investigated the mixing due to turbulence in JPUE simulation by checking the mixture of the two phases.
 It is shown in Fig.
@@
 \subsubsection{Input parameters}
-Eruption parameters, material properties and atmosphere are chosen to be the same as the strong plume no wind case in an inter-comparison study on eruptive column models \citep{costa2016results}.
-Eruption condition is list in Table \ref{tab:input_parameters}.
-The density of erupted material at the vent and radius of the vent can be computed from given parameters.
+Eruption parameters, material properties and atmosphere are chosen to be the same as the strong plume no wind case in a comparison study on eruptive column models \citep{costa2016results}.
+Eruption conditions are listed in Table \ref{tab:input_parameters}.
+The density of erupted material at the vent and radius of the vent can be computed from the given parameters.
 The eruption pressure is assumed to be as the same as pressure of ambient at the vent and hence is not given in the table.
-The vertical profiles of atmospheric properties were obtained based on the reanalysis data from ECMWF (European Centre for Medium-Range Weather Forecasts) for the period corresponding to climactic phase of the Pinatubo eruption (Philippines, 15 June 1991).
+The vertical profiles of atmospheric properties were obtained based on the reanalysis data from ECMWF (European Centre for Medium-Range Weather Forecasts) for the period corresponding to the climactic phase of the Pinatubo eruption (Philippines, 15 June 1991).
 These conditions are more typical of a tropical atmosphere (see Fig.
@@
 
-The profiles of local variables match well with simulation results of existing 3D models in general sense. The basic phenomena in volcanic plume development is correctly captured by our model.
+The profiles of local variables match well with simulation results of existing 3D models in a general sense. The basic phenomena in volcanic plume development is correctly captured by our model.
 
@@
 \includegraphics[width=5cm]{radius_strong}
-\caption{Radius of the strong plume without wind after reaching its top height}
+\caption{Radius of the strong plume without wind after reaching its top height.}
 \label{fig:strong_local_radius}
@@
 
-As the height increase, the amount of entrained air also increases.
-Around the neutral height, where the umbrella expanding, the entrainment of air shows a slight decrease due to lack of air surrounding the column at that height.
+As the height increases, the amount of entrained air also increases.
+Around the neutral height, where the umbrella expands, the entrainment of air shows a slight decrease due to lack of air surrounding the column at that height.
 The profile for gas, which account for both air and vapor, shows a very similar tendency as that of entrained air.
@@
 (\ref{eq:gas-frac-express}) does not hold for ATHAM.
-However, the profile of entrained air and profile of gas predicted by ATHAM are still very close to each other which implies that ignoring of water phase change is a valid assumption for eruptions similar to this test case (strong plume with erupted water fraction in erupted material less than 5\%).
-Because air occupies a larger portion of the gas and plays a more significant role.
+However, the profile of entrained air and profile of gas predicted by ATHAM are still very close to each other which implies that ignoring of water phase change is a valid assumption for eruptions similar to this test case (strong plume with erupted water fraction in erupted material less than 5\%), because air occupies a larger portion of the gas and plays a more significant role.
 As for mass fraction of solid, similarly, Eq.
@@
 (\ref{eq:solid-frac-express2}) hold for our model.
-PDAC, which treat particles of two different sizes as two separate phases, got similar mass fraction profile.
+PDAC, which treat particles of two different sizes as two separate phases, found a similar mass fraction profile.
 That implies that our model is at least valid for eruptions similar to the test case.
@@
 
-With more cool air entrained into the plume and mixing with hot erupted material, the temperature of the plume decreases as the height increase as shown in Fig.
+With more cool air entrained into the plume and mixing with hot erupted material, the temperature of the plume decreases as the height increases as shown in Fig.
 \ref{fig:strong_local_temp}.
@@
 Our model adopts the same assumptions and governing equations as SK-3D.
-However, there is still a big disparity between the profiles of local variables of our model and 3K-3D.
-One of the big differences between these two model is that we adopt a LANS type of turbulence model while 3K-3D adopts a LES (large eddy simulation) turbulence model.
+However, there is still a big disparity between the profiles of local variables of our model and SK-3D.
+One of the big differences between these two model is that we adopt a LANS type of turbulence model while SK-3D adopts a LES (large eddy simulation) turbulence model.
 This implies that turbulence model might play a critical role in plume simulation.
@@
 \conclusions  \label{sec:conclusion}%% \conclusions[modified heading if necessary]
-A new plume model was developed based on SPH method.
+A new plume model was developed based on the SPH method.
 Extensions necessary for Lagrangian methodology and compressible flow were made in the formulation of the equations of motion and turbulence models.
 Advanced numerical techniques in SPH were exploited and tailored for this model.
-High performance computing was used to mitigate the tradeoff between accuracy (depends on comprehensiveness of the model, resolution, order of accuracy of numerical methods, scheme for time upgrading.) and simulation time (depends on, comprehensiveness, resolution, order of accuracy of numerical methods, scheme for time upgrading ...
+High performance computing was used to mitigate the tradeoff between accuracy (depends on comprehensiveness of the model, resolution, order of accuracy of numerical methods, scheme for time upgrading.) and simulation time (depends on comprehensiveness, resolution, order of accuracy of numerical methods, scheme for time upgrading ...
 and computational techniques).
@@
 Currently existing 3D models focus on certain aspect of the volcanic plume (PDAC on pyroclastic flow, ATHAM on microphysics, and SK-3D on entrainment with higher accuracy and higher order of accuracy) and hence, naturally, different assumptions were made in these models.
-However, these different aspects of volcanic plume are not independent and actually coupled.
-For example, it has been illustrated that gas-particle non-equilibrium would introduce a previously unrecognized jet-dragging effect, which imposes great influence on plume development, especially for weak plume \cite{cerminara2016large}.
-In addition, there is no absolute boundary to determine which kind of hazard is dominant in certain eruption.
+However, these different aspects of volcanic plumes are not independent, but are instead coupled.
+For example, it has been illustrated that gas-particle non-equilibrium would introduce a previously unrecognized jet-dragging effect, which imposes great influence on plume development, especially for weak plumes \cite{cerminara2016large}.
+In addition, there is no absolute boundary to determine which kind of hazard is dominant in certain eruptions.
 So it is necessary to simulate all associated hazards in one model.
 Actually, effort has already been put on developing more comprehensive plume models.
-For example, a large-particle module (LPM) was added to ATHAM to track the paths of rocky particles (pyroclastic or tephra) within the plume and predicts where these particles fall \citep{kobs2009modeling}.
-We were also motivated by such an evolution of plume modeling to choose SPH as our numerical tool besides its ability in dealing with interface for multiphase flows.
-As mentioned in the introduction section, SPH method has good extensibility and adding of new physics and phases requires tiny modification on the code compared with mesh based methods.
-The last but not the least, the dramatic development of computational power makes it possible to establish a comprehensive model.
-Even though right now, current computational capacity does not allow us to have too comprehensive model, the easy-extension feature of SPH makes it convenient to keep adding new physics into the model when necessary and computationally feasible.
-
-We have presented in this paper an initial effort and results towards developing a first principle based plume model with comprehensive physics, adopting proper numerical tools and power of high performance computing.
+For example, a large-particle module (LPM) was added to ATHAM to track the paths of rocky particles (pyroclastic or tephra) within the plume and predict where these particles fall \citep{kobs2009modeling}.
+We were also motivated by such an evolution of plume modeling to choose SPH as our numerical tool besides its ability in dealing with interfaces for multiphase flows.
+As mentioned in the introduction section, SPH method has good extensibility and adding new physics and phases requires much less modification of the code compared with mesh based methods.
+Last but not least, the dramatic development of computational power makes it possible to establish a comprehensive model.
+While current computational capacity may not allow us to have a fully comprehensive model, the easy-extension feature of SPH makes it convenient to keep adding new physics into the model when necessary and computationally feasible.
+
+We have presented in this paper an initial effort and results towards developing a first principle based plume model with comprehensive physics, adopting proper numerical tools and high performance computing.
 More advanced numerical techniques, such as adaptive particle size, Godunov-SPH, semi-explicit time advancing scheme and better data management strategies and algorithms are on our list to exploit in the future.
 In the near future, effect of wind field will be take into account.
-Code will also be made available in the open source form for the community to enhance.
+Our code will also be made available in the open source form for the community to enhance.
 %In all existing 3D plume models, numerical calculations are performed on a non-uniform (vertically and horizontally stretched) grid, with different grid resolutions and domain sizes.
@@
 \subsection{Data availability}
-Data for running the test cases, including meteorological data, properties of erupted material and air, and eruption condition data can be found in source code repository.
+Data for running the test cases, including meteorological data, properties of erupted material and air, and eruption condition data can be found in the source code repository.
 Raw data of simulation results are archived in Box.
 Access permission will be given upon request.
-Post-processed data for plotting figures is provided as supplement along with the manuscript.
+Post-processed data for plotting figures is provided as a supplement along with the manuscript.
 Please note that some of these pictures are generated by visualizing the raw data with visualization tools, there is no separate data for these figures.

### 2017-03-28 Manuscripts_GMD.tex
modification made according to Dr. Jones' comments

@@
 
-\title{Plume-SPH 1.0: A three-dimensional, dusty-gas volcanic plume model based on smoothed-particle hydrodynamics}
+\title{Plume-SPH 1.0: A three-dimensional, dusty-gas volcanic plume model based on smoothed particle hydrodynamics}
 
@@
 %<<<<<<< HEAD
-
 \correspondence{Abani Patra (abani@buffalo.edu)}
-
 %=======
@@
 In this model, erupted material is assumed to be well-mixed, and represented by one phase while air is another phase.
-We also assume immediate dynamic and thermodynamic equilibrium between air and erupted material and that the effect of wind is not significant.
-
-Several newly developed techniques in SPH are adopted and extended to address numerical challenges in simulating multiphase compressible turbulent flow with classical SPH methods.
+We also assume immediately dynamics and thermodynamics equilibrium between air and erupted material and that the effect of wind is not significant.
+
+Several newly developed techniques in SPH are adopted and extended to address numerical challenges in simulating multiphase compressible turbulent flow with classical SPH method.
 The $SPH-\varepsilon$ turbulence model is adopted to capture mixing at unresolved scales.
@@
 Primary hazards associated with explosive volcanic eruptions include pyroclastic density currents (flows and surges), the widespread deposition of airfall tephra, and the threats to aviation posed by volcanic ash in the atmosphere.
-Simulation of all possible hazards with one model is difficult due to the fact that different length scales dominate in different hazards.
+Simulation of all possible hazards with one model is difficult due to the fact that different length scales dominate different hazards.
 Our focus here is the hazard that volcanic ash poses to aircraft.
@@
 Usually, plume models are used to provide these source terms for VATDs and the forecast accuracy is critically dependent on these models.
-This paper reports on a new 3D (three dimensional) volcanic plume model designed to exploit the advantages of mesh-free methods for 3D modeling of such plumes that involve multi-phase free boundary flows.
+This paper reports on a new 3D (three dimensional) volcanic plume model designed to exploit the advantages of mesh-free methods for 3D modeling of such plumes that involve multiphase free boundary flows.
 
@@
 Several 1D (one dimensional) volcanic plume models have been developed in the past few decades, ranging from the most basic 1D model \citep{woods1988fluid} which only accounts for mass conservation to more recently developed 1D models  \citep{bursik2001effect, mastin2007user, degruyter2012improving, woodhouse2013interaction, devenish2013using, de2015plume, folch2016fplume, pouget2016sensitivity} which tend to account for more comprehensive physics effects.
-FPLUME-1.0 \citep{folch2016fplume} accounts for wind effect, entrainment of moisture, water phase change, particle fallout and re-entrainment and even wet aggregation of ash.
+For example, FPLUME-1.0 \citep{folch2016fplume} accounts for wind effect, entrainment of moisture, water phase change, particle fallout and re-entrainment and even wet aggregation of ash.
 However, in these 1D models, the entrainment of air is evaluated based on two coefficients: entrainment coefficient for the vertical plume and the entrainment coefficient that describes the effect of wind.
@@
 One of the earliest of these is the 3D model PDAC (Pyroclastic Dispersion Analysis Code) \citep{neri2003multiparticle}  which is a non-equilibrium, multiphase, 3D compressible flow model.
-Conservation equations for each phase are solved separately with a finite volume method.
+Conservation equations for each phase are solved separately with the finite volume method.
 A parallel computing version of PDAC was developed lately \citep{ongaro2007parallel}.
@@
 Another 3D model, ATHAM (Active Tracer High-Resolution Atmospheric Model) \citep{oberhuber1998volcanic} focuses more on microphysics within the plume.
-As pyroclastic flow is not the initial concern of ATHAM, \textcolor{red}{immediately dynamics} and thermodynamic equilibrium is assumed in ATHAM.
+As pyroclastic flow is not the initial concern of ATHAM, immediately dynamics and thermodynamics equilibrium is assumed in ATHAM.
 The dynamic core of ATHAM solves the compressible Euler equations for momentum, pressure and temperature of the gas particle mixture \citep{oberhuber1998volcanic}.
@@
 
-Recently, a first order, non-equilibrium compressible 3D model, ASHEE \citep{cerminara2016ashee}, was introduced based on three dimensional N-phases Eulerian transportation equation.
-ASHEE is valid for low concentration and low Stokes number region and much faster than N-phase Eulerian model.
-The model is based on the open source numerical solver OpenFOAM \textcolor{red}{cite?}, adapting its unstructured finite volume solver.
+Recently, a first order, nonequilibrium compressible 3D model, ASHEE \citep{cerminara2016ashee}, was introduced based on three dimensional N-phases Eulerian transportation equation.
+ASHEE is valid for low concentration and low Stokes number region and much faster than N-phases Eulerian model.
+The model is based on the open source numerical solver OpenFOAM \citep{weller1998tensorial}, adapting its unstructured finite volume solver.
 
@@
  %Different simplification was made based on their scope. Improvements have been making since their birth. 
- Accuracy of simulation and forecast skill (depending on comprehensiveness of the model, resolution of discretization, numerical error, and order of accuracy) and simulation time (depending on number of governing equations, resolution, numerical methods and parallel techniques) are always a pair of conflicting considerations in 3D plume simulations.\\
+ Accuracy of simulation (depending on comprehensiveness of the model, resolution of discretization, numerical error, and order of accuracy) and simulation time (depending on number of governing equations, resolution, numerical methods and parallel techniques) are always a pair of conflicting considerations in 3D plume simulations.\\
 
 \subsection{Features of SPH}
-To the bext of our knowledge, all of the existing 3D plume models are mesh based Eulerian methods, and there are no 3D plume models based on mesh free Lagrangian methods.
+To the best of our knowledge, all of the existing 3D plume models are mesh based Eulerian methods, and there are no 3D plume models based on mesh free Lagrangian methods.
 Lagrangian methods have several features that are suitable for volcanic plume simulation that we outline below.
-Among such Lagrangian methods, smoothed-particle hydrodynamics \citep{gingold1977smoothed,lucy1977numerical} based simulations have shown good agreement with experiments for many applications in fluid dynamics.
+Among such Lagrangian methods, smoothed particle hydrodynamics \citep{gingold1977smoothed,lucy1977numerical} based simulations have shown good agreement with experiments for many applications in fluid dynamics.
 And it currently is, by far, the most widely used mesh-free scheme.
@@
 \subsection{Our contributions}
-SPH is still a relatively new method in computational fluid dynamics compared to mesh-based methods, like Finite Volume.
+SPH is still a relatively new method in computational fluid dynamics compared to mesh-based methods, such as finite volume method.
 Implementations of SPH for compressible multiphase turbulent flows are few.
-Several issues endemic to classical SPH necessary for use in this domain are fixed in our implementation.
+Several issues endemic to classical SPH are fixed in our implementation.
 The most popular applications of SPH (and their original motivating application) has been in the simulation of free surface flow, such as breaking-waves and floods.
@@
 There are several turbulence models proposed for SPH method \citep{issa2005numerical,violeau2007numerical}.
-We adopt a LANS (Lagrangian Averaged Navier-Stokes) turbulence model \citep{monaghan2011turbulence}, which was originally proposed for incompressible, and is extended here for compressible flow accounting for turbulent heat exchange.
+We adopt a LANS (Lagrangian Averaged Navier-Stokes) turbulence model \citep{monaghan2011turbulence}, which was originally proposed for incompressible flow, and is extended here for compressible flow accounting for turbulent heat exchange.
 \item Corrected formulation of SPH \citep{chen1999improvement} is adapted to bypass the well-known tensile instability issues of classical SPH.
@@
 In this model, molecular viscosity and heat conduction is neglected since turbulent exchange coefficients are dominant.
-Erupted material consist of solid with different size and mixture of gases (various constituent) is assumed to be well mixed and behave like a single phase fluid (phase 2) which is valid for eruptions with fine particles and ash.
+Erupted material consisting of solid with different size and mixture of gases is assumed to be well mixed and behave like a single phase fluid (phase 2) which is valid for eruptions with fine particles and ash.
 Air (also a mixture of different gases) is assumed to be another phase (phase 1).
-Immediate thermodynamic equilibrium is assumed so that no separate energy equation is needed for each phase.
+Immediately thermodynamics equilibrium is assumed so that no separate energy equation is needed for each phase.
 As a result, there is only one energy equation for both phases (heat exchange term between different phases does not show up under this assumption).
-Immediate dynamic equilibrium is assumed so that no separate momentum equation is needed for each phase.
+Immediately dynamics equilibrium is assumed so that no separate momentum equation is needed for each phase.
 As a result, there is only one vector momentum equation for both phases (drag force term does not show up with this assumption).
@@
 \subsection{Governing equations}
-Based above assumptions, the governing equation of our model is given as following, which is the same as the governing equations of SK-3D \citep{suzuki2005numerical}.
+Based on above assumptions, the governing equations of our model are given as following, which is the same as the governing equations of SK-3D \citep{suzuki2005numerical}.
 \begin{align}
-\dfrac{\partial \rho}{\partial t} + \nabla \cdot (\rho \textbf{v}) = 0 \label{eq:gov-cs-rho} \\
-\dfrac{\partial \rho \xi}{\partial t} + \nabla \cdot (\rho \xi \textbf{v}) = 0 \label{eq:gov-cs-ks}\\
-\dfrac{\partial \rho \textbf{v}}{\partial t} + \nabla \cdot (\rho \textbf{v} \textbf{v} + p\textbf{I}) = \rho \textbf{g} \label{eq:gov-cs-v} \\
-\dfrac{\partial \rho E}{\partial t} + \nabla \cdot [(\rho E + p )\textbf{v}] = \rho \textbf{g} \cdot\textbf{v} \label{eq:gov-cs-e}
+\dfrac{\partial \rho}{\partial t} + \nabla \cdot \left(\rho \textbf{v}\right) = 0 \label{eq:gov-cs-rho} \\
+\dfrac{\partial \rho \xi}{\partial t} + \nabla \cdot \left(\rho \xi \textbf{v}\right) = 0 \label{eq:gov-cs-ks}\\
+\dfrac{\partial \rho \textbf{v}}{\partial t} + \nabla \cdot \left(\rho \textbf{v} \textbf{v} + p\textbf{I}\right) = \rho \textbf{g} \label{eq:gov-cs-v} \\
+\dfrac{\partial \rho E}{\partial t} + \nabla \cdot \left[\left(\rho E + p \right)\textbf{v}\right] = \rho \textbf{g} \cdot\textbf{v} \label{eq:gov-cs-e}
 \end{align}
-$\xi$ is the mass fraction of ejected material.
-$E = e + K $ is total energy which is a summation of kinetic energy $K$ and internal energy $e$.
-\textcolor{red}{should define all variates when they first appear, even the ``obvious'' ones.}
-An additional equation is required to close the system. In this model, the equation for closing the system is the EOS (equation of state) for ideal gas.
-\begin{equation}
-p = (\gamma_m - 1)\rho e \label{eq:EOS}
+Where $\rho$ is the density, $\textbf{v}$ is the velocity, $\xi$ is the mass fraction of ejected material, $\textbf{g}$ is the gravitational acceleration, $\textbf{I}$ is a unit tensor.
+$E = e + K $ is the total energy which is a summation of kinetic energy $K$ and internal energy $e$.
+An additional equation is required to close the system. In this model, the equation for closing the system is the following EOS (equation of state).
+\begin{equation}
+p = \left(\gamma_m - 1\right)\rho e \label{eq:EOS}
 \end{equation}
@@
 \begin{equation}
-R_m = n_gR_g + n_aR_a  \label{eq:gov-Rm}
-\end{equation}
-\begin{equation}
-C_{vm} = n_s C_{vs} + n_g C_{vg} + n_a C_{va} \label{eq:gov-Cvm}
-\end{equation}
-\begin{equation}
-n_a = 1 - \xi \label{eq:gov-na}
-\end{equation}
-\begin{equation}
-n_g = \xi n_{g0} \label{eq:gov-ng}
-\end{equation}
-\begin{equation}
-n_s = \xi - n_g \label{eq:gov-ns}
-\end{equation}
-Where, $C_v$ is specific heat with constant volume, $n$ is mass fraction, $R$ is gas constant. The subscript 
-$m$ represents mixture of ejected material and air, $s$ represents solid portion in ejected material, $g$ represents gas portion in the ejected material and $a$ represents air.
+R_m = \xi_g R_g + \xi_a R_a  \label{eq:gov-Rm}
+\end{equation}
+\begin{equation}
+C_{vm} = \xi_s C_{vs} + \xi_g C_{vg} + \xi_a C_{va} \label{eq:gov-Cvm}
+\end{equation}
+\begin{equation}
+\xi_a = 1 - \xi \label{eq:gov-na}
+\end{equation}
+\begin{equation}
+\xi_g = \xi \cdot \xi_{g0} \label{eq:gov-ng}
+\end{equation}
+\begin{equation}
+\xi_s = \xi - \xi_g \label{eq:gov-ns}
+\end{equation}
+Where, $C_v$ is specific heat with constant volume, $R$ is gas constant. $\xi$ with subscript is mass fraction of corresponding constituent. The subscript 
+$m$ represents mixture of ejected material and air, $s$ represents solid portion in the ejected material, $g$ represents gas portion in the ejected material and $a$ represents air.
+$\xi_{g0}$ is the mass fraction of vapor in erupted material.
 
@@
 For SPH, governing equations in Lagrange form is needed.
-By deducting kinetic energy from energy equation, subtracting mass conservation from momentum equation, combining transient term and advection term into material derivative term, the governing equation is put into the final form, in which advection term does not show up explicitly.
+By deducting kinetic energy from energy equation, subtracting mass conservation from momentum equation, combining transient term and advection term into material derivative term, the governing equations are put into the final form, in which advection term does not show up explicitly.
 \begin{align}
@@
 \subsubsection{Velocity inlet}
-At the vent, temperature of erupted material $T$, eruption velocity $\textbf{v}$, the mass fraction of vapor in erupted material $n_{g0}$ and mass discharge rate $\dot M$ are given.
+At the vent, temperature of erupted material $T$, eruption velocity $\textbf{v}$, the mass fraction of vapor in erupted material $\xi_{g0}$ and mass discharge rate $\dot M$ are given.
 The pressure of erupted material $p$ is assumed to be the same as ambient pressure for pressure-balanced eruption.
@@
 \begin{align}
-\rho =const = p/(R_m T) \label{eq:erupt_bc_rho} \\
+\rho =const = p/\left(R_m T\right) \label{eq:erupt_bc_rho} \\
 \xi=const=1 \label{eq:erupt_bc_xi}\\
 \textbf{v} = const =\{u,v,w\}^T \label{eq:erupt_bc_v}\\
-\dfrac{\partial e}{\partial n}=\dot M e /(\pi r^2) \label{eq:erupt_bc_e}
+\dfrac{\partial e}{\partial n}=\dot M e /\left(\pi r^2\right) \label{eq:erupt_bc_e}
 \end{align} 
-Where $r$ is the radius of the vent.
+Where $r$ is the radius of the vent, $n$ is the direction normal to the boundary.
 
@@
 \begin{equation}
-p = p_a(h)\label{eq:pressure_bc_p} 
+p = p_a\left(z\right)\label{eq:pressure_bc_p} 
 \end{equation}
@@
 \section{SPH method} \label{sec:SPH_method}
-SPH is a mesh free Lagragian method.
+SPH is a mesh-free Lagragian method.
 In SPH, the domain (control mass) is discretized by a set of particles or discretization points and the position of each particle (discretization point) is updated at every time step based on the motion computed.
@@
 \begin{equation}
-A(\textbf{x})=\int_{-\infty}^{\infty} A(\textbf{x} \prime) \delta (\textbf{x} \prime - \textbf{x}) d\textbf{x} \prime
+A\left(\textbf{x}\right)=\int_{-\infty}^{\infty} A\left(\textbf{x} \prime\right) \delta \left(\textbf{x} \prime - \textbf{x}\right) d\textbf{x} \prime
 \label{eq:Dirac-translation}
@@
 The Dirac function in Eq.
-(\ref{eq:Dirac-translation}) can be approximated by a weighting function $w(\textbf{x}-\textbf{x}\prime, h)$ (or $w(\textbf{x}\prime-\textbf{x}, h)$) which tends to a Dirac function when $h \rightarrow 0$ :
-\begin{equation}
-\lim _{h \rightarrow 0} w(\textbf{x} \prime-\textbf{x}, h) =  \delta (\textbf{x} \prime - \textbf{x})
+(\ref{eq:Dirac-translation}) can be approximated by a weighting function $w\left(\textbf{x}-\textbf{x}\prime, h\right)$ (or $w\left(\textbf{x}\prime-\textbf{x}, h\right)$)) which tends to a Dirac function when the smoothing length $h \rightarrow 0$ :
+\begin{equation}
+\lim _{h \rightarrow 0} w\left(\textbf{x} \prime-\textbf{x}, h\right) =  \delta \left(\textbf{x} \prime - \textbf{x}\right)
 \label{eq:SPH_kernel_delta}
 \end{equation}
-An arbitrary function $A(\textbf{x})$ can then be approximated by:
-\begin{equation}
-A(\textbf{x}) \approx <A(\textbf{x})> = \int_{\Omega} A(\textbf{x} \prime) w(\textbf{x}-\textbf{x}\prime, h) d\textbf{x}\prime + O(h^2)
+An arbitrary function $A\left(\textbf{x}\right)$ can then be approximated by:
+\begin{equation}
+A\left(\textbf{x}\right) \approx <A\left(\textbf{x}\right)> = \int_{\Omega} A\left(\textbf{x} \prime\right) w\left(\textbf{x}-\textbf{x}\prime, h\right) d\textbf{x}\prime + O\left(h^2\right)
 \label{eq:SPH-fundamental-principle}
@@
 Where $h$ is the smoothing length, determining the interaction distance.
-According to Eq.
-(\ref{eq:SPH-fundamental-principle}), the order of accuracy of SPH method is second order.
-However, in practice, second order can not be achieved because there is no guarantee on the symmetry of particle distribution in real simulation \citep{price2012smoothed}.
+As the weighting function is symmetric (Eq.
+(\ref{eq:SPH-kernel-symmetric})) and satisfies the normalization condition (Eq.
+(\ref{eq:SPH-kernel-normalization-prop})), odd error terms in Eq.
+(\ref{eq:SPH-fundamental-principle}) vanishe leading to a second order approximation.
+However, in practice, second order of accuracy can not be achieved because there is no guarantee on the symmetry of particle distribution in real simulation \citep{price2012smoothed}.
 Recall that $d\textbf{x}\prime = \dfrac{dm \prime}{\rho \prime}$, the integration equation, Eq.
@@
 \begin{equation}
-<A(\textbf{x})> \approx \sum_b m_b \dfrac{A_b}{\rho_b} w(\textbf{x}-\textbf{x}_b, h)
+<A\left(\textbf{x}\right)> \approx \sum_b m_b \dfrac{A_b}{\rho_b} w\left(\textbf{x}-\textbf{x}_b, h\right)
 \label{eq:SPH-approximation-sum}
@@
 \begin{split}
-<\nabla A(\textbf{x})> & = \dfrac{\partial }{\partial \textbf{x}} \int_{\Omega} A(\textbf{x} \prime) w(\textbf{x}-\textbf{x}\prime, h) d\textbf{x}\prime + O(h^2) \\
-& \approx \sum_b m_b \dfrac{A_b}{\rho_b} \nabla w(\textbf{x} - \textbf{x}_b, h)
+<\nabla A\left(\textbf{x}\right)> & = \dfrac{\partial }{\partial \textbf{x}} \int_{\Omega} A\left(\textbf{x} \prime\right) w\left(\textbf{x}-\textbf{x}\prime, h\right) d\textbf{x}\prime + O\left(h^2\right) \\
+& \approx \sum_b m_b \dfrac{A_b}{\rho_b} \nabla w\left(\textbf{x} - \textbf{x}_b, h\right)
 \end{split} 
@@
 \begin{align}
-<\textbf{A}(\textbf{x})> \approx \sum_b m_b \dfrac{\textbf{A}_b}{\rho_b} w(\textbf{x}-\textbf{x}_b, h) \\
-<\nabla \cdot \textbf{A}(\textbf{x})> \approx \sum_b m_b \dfrac{\textbf{A}_b}{\rho_b} \cdot \nabla w(\textbf{x} - \textbf{x}_b, h) \\
-<\nabla \times \textbf{A}(\textbf{x})> \approx \sum_b m_b \dfrac{\textbf{A}_b}{\rho_b} \times \nabla w(\textbf{x} - \textbf{x}_b, h) \\
-<\nabla^j \textbf{A}^i(\textbf{x})> \approx \sum_b m_b \dfrac{\textbf{A}_b^i}{\rho_b} \nabla^j w(\textbf{x} - \textbf{x}_b, h) 
+<\textbf{A}\left(\textbf{x}\right)> \approx \sum_b m_b \dfrac{\textbf{A}_b}{\rho_b} w\left(\textbf{x}-\textbf{x}_b, h\right) \\
+<\nabla \cdot \textbf{A}\left(\textbf{x}\right)> \approx \sum_b m_b \dfrac{\textbf{A}_b}{\rho_b} \cdot \nabla w\left(\textbf{x} - \textbf{x}_b, h\right) \\
+<\nabla \times \textbf{A}\left(\textbf{x}\right)> \approx \sum_b m_b \dfrac{\textbf{A}_b}{\rho_b} \times \nabla w\left(\textbf{x} - \textbf{x}_b, h\right) \\
+<\nabla^j \textbf{A}^i\left(\textbf{x}\right)> \approx \sum_b m_b \dfrac{\textbf{A}_b^i}{\rho_b} \nabla^j w\left(\textbf{x} - \textbf{x}_b, h\right) 
 \label{eq:SPH-vecctor-function}
@@
 (\ref{eq:SPH_kernel_delta}).
-So it can be viewed as an approximate form of the Dirac function and satisfy the normalization condition:
-\begin{equation}
-\int	 w(\textbf{x}-\textbf{x}\prime, h) d\textbf{x}\prime = 1
+So it can be viewed as an approximate form of the Dirac function and satisfies the normalization condition:
+\begin{equation}
+\int	 w\left(\textbf{x}-\textbf{x}\prime, h\right) d\textbf{x}\prime = 1
 \label{eq:SPH-kernel-normalization-prop}
@@
 \begin{equation}
-w(\textbf{x}- \textbf{x} \prime, h) = w(\textbf{x} \prime - \textbf{x}, h)
+w\left(\textbf{x}- \textbf{x} \prime, h\right) = w\left(\textbf{x} \prime - \textbf{x}, h\right)
 \label{eq:SPH-kernel-symmetric}
@@
 There is a wide variety of possible weighting functions, such as spline functions (with different orders) and Gaussian functions.
-Generally, the accuracy increases with the order of the polynomials of the kernel function, but the computational time also increases as number of interactions increase.
+Generally, the accuracy increases with the order of the polynomials of the kernel function, but the computational cost also increases as number of interactions increase.
 We are adopting a truncated Gaussian function as the weighting function in our simulation.
 \begin{equation}
-w(\textbf{x} - \textbf{x}_b) = 
+w\left(\textbf{x} - \textbf{x}_b\right) = 
 \begin{cases} 
-      \dfrac{1}{(h \sqrt{\pi})^d} exp \left[- \left(\dfrac{\textbf{x} - \textbf{x} \prime}{h}\right)^2 \right] &  \vert \textbf{x} - \textbf{x} \prime \vert \leq 3h\\
+      \dfrac{1}{\left(h \sqrt{\pi}\right)^d} exp \left[- \left(\dfrac{\textbf{x} - \textbf{x} \prime}{h} \right)^2 \right] &  \vert \textbf{x} - \textbf{x} \prime \vert \leq 3h\\
       0 & \text{Otherwise}
@@
 \begin{equation}
-\nabla w(\textbf{x} - \textbf{x}_b) = 
+\nabla w\left(\textbf{x} - \textbf{x}_b\right) = 
 \begin{cases} 
-      -2\left(\dfrac{\textbf{x} - \textbf{x} \prime}{h^2}\right) \dfrac{1}{(h \sqrt{\pi})^d} exp \left[- \left(\dfrac{\textbf{x} - \textbf{x} \prime}{h}\right)^2 \right] &  \vert \textbf{x} - \textbf{x} \prime \vert \leq 3h\\
+      -2\left(\dfrac{\textbf{x} - \textbf{x} \prime}{h^2}\right) \dfrac{1}{\left(h \sqrt{\pi}\right)^d} exp \left[- \left(\dfrac{\textbf{x} - \textbf{x} \prime}{h}\right)^2 \right] &  \vert \textbf{x} - \textbf{x} \prime \vert \leq 3h\\
       0 & \text{Otherwise}
@@
 (\ref{eq:SPH-approximation-sum}) to Eq.
-(\ref{eq:SPH-vecctor-function}) provides a general way to obtaining SPH expressions of governing equations.
+(\ref{eq:SPH-vecctor-function}) provides a general way to obtain SPH expressions of governing equations.
 The problem is that using these expressions ``{as is}" in general leads to quite poor gradient estimates.
 Various tricks can be used to conserve linear and angular momentum and thermal energy \citep{monaghan1992smoothed}.
-Special treatments are also needed for second order derivative terms.
+Special treatments are also needed for second order derivative terms \citep{monaghan2005smoothed}.
 We only refer here to one of these possible discretizations of compressible Euler equations with SPH:
 \begin{align}
-<\rho_a> = \sum_b m_b w_{ab} (h) \label{eq:ns-sph-d} \\
-\left\langle\dfrac{d \textbf{v}_a}{d t}\right\rangle = \sum_b m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}^{\beta}\right) \nabla_a w_{a b}(h) +\textbf{g} \label{eq:ns-sph-v} \\
+<\rho_a> = \sum_b m_b w_{ab} \left(h\right) \label{eq:ns-sph-d} \\
+\left\langle\dfrac{d \textbf{v}_a}{d t}\right\rangle = \sum_b m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}\right) \nabla_a w_{a b}\left(h\right) +\textbf{g} \label{eq:ns-sph-v} \\
 \left\langle\dfrac{d e_a}{d t}\right\rangle=
- 0.5\sum_b m_b \textbf{v}_{a b}\left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}^{\beta}\right) \cdot \nabla_a w_{a b}(h) \label{eq:ns-sph-e}
+ 0.5\sum_b m_b \textbf{v}_{a b}\left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}\right) \cdot \nabla_a w_{a b}\left(h\right) \label{eq:ns-sph-e}
 \end{align}
-where, $a$ is the SPH particle index, $\textbf{v}_{a b} = \textbf{v}_a - \textbf{v}_b$. $\Pi$ is an artificial viscosity term, which will be discussed in section \ref{sec:artificial-viscosity}.
+where, $a$ is the SPH particle index, $\textbf{v}_{a b} = \textbf{v}_a - \textbf{v}_b$.
+$\Pi$ is an artificial viscosity term, which will be discussed in section \ref{sec:artificial-viscosity}.
+$w_{a b}\left(h\right)$ is a concise form of $w\left(\textbf{x}_a - \textbf{x}_b, h\right)$ and from here on, we will use this concise form.
 %However, these discretized formulations will not guarantee conservation of entropy.
@@
 \begin{align}
-<\rho_a> = \sum m_b w_{ab} (h) + \dot{\rho}(\textbf{x},t)\label{eq:ns-source-sph-d} \\
-\left\langle\dfrac{d \textbf{v}_a}{d t}\right\rangle= \sum_b m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}^{\beta}\right) \nabla_a w_{a b}(h) +\textbf{g} + D \sum	_b m_b \dfrac{\textbf{v}_b - \textbf{v}_a}{\rho_b} \label{eq:ns-drag-sph-v} \\
+<\rho_a> = \sum m_b w_{ab} \left(h\right) + \dot{\rho}\left(\textbf{x},t\right)\label{eq:ns-source-sph-d} \\
+\left\langle\dfrac{d \textbf{v}_a}{d t}\right\rangle= \sum_b m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}\right) \nabla_a w_{a b}\left(h\right) +\textbf{g} + D \sum	_b m_b \dfrac{\textbf{v}_b - \textbf{v}_a}{\rho_b} \label{eq:ns-drag-sph-v} \\
 \left\langle\dfrac{d e_a}{d t}\right\rangle=
- 0.5\sum_b m_b \textbf{v}_{a b}\left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}^{\beta}\right) \cdot \nabla_a w_{a b}(h) + \sum_b \dfrac{m_b}{\rho_b}(\kappa_a + \kappa_b) \dfrac{(T_a - T_b)}{\textbf{r}_a - \textbf{r}_b} w_{ab}(h) \label{eq:ns-conduction-sph-e}
+ 0.5\sum_b m_b \textbf{v}_{a b}\left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}\right) \cdot \nabla_a w_{a b}\left(h\right) + \sum_b \dfrac{m_b}{\rho_b}\left(\kappa_a + \kappa_b\right) \dfrac{\left(T_a - T_b\right)}{\textbf{r}_a - \textbf{r}_b} w_{ab}\left(h\right) \label{eq:ns-conduction-sph-e}
 \end{align}
-where the source term $\dot{\rho}$ can be a "sink" of erupted vapor due to phase change like in ATHAM \citep{oberhuber1998volcanic}.
+where the source term $\dot{\rho}$ can be a "sink" of erupted vapor due to its phase change.
+%(see microphysics modelling in ATHAM \citep{oberhuber1998volcanic}).
 $D$ is a drag force coefficient.
-$\kappa$ is heat conduction coefficient.
+$\kappa$ is the heat conduction coefficient.
+$T$ is the temperature.
 Other physics can be added easily in a similar way.
 Adding of these new terms leads to modification of several lines in the source code.
-The drag force term should show up only when dynamic dis-equilibrium between different phases is considered.
+The drag force term should show up only when dynamics disequilibrium between different phases is considered.
 In that case, each phase needs one set of governing equations of Navier-Stokes type.
-Adding of new phases into SPH code only needs few new lines for the new phase besides adding  interaction terms.
+Adding of new phase into SPH code only needs adding few new lines for the new phase besides interaction terms introduced by the new phase.
 
@@
 (\ref{eq:ns-sph-e}), represents artificial viscosity term which is essentially a second order derivative term.
-As in the case of first derivatives, second derivatives can be estimated by differentiating a SPH interpolation twice.
+As in the case of first order derivatives, second order derivatives can be estimated by differentiating a SPH interpolation twice.
 However, such a formulation has two disadvantages: First, it is very sensitive to particle disorder.
-Second, the second derivative of the kernel can change sign and lead to unphysical representations (viscous dissipation causes decrease of the entropy).
+Second, the second derivative of the kernel can change sign and lead to unphysical representations (For example, viscous dissipation causes decrease of the entropy).
 
@@
 \begin{equation}
-\Pi_{ab}=- \frac{\nu}{\bar{\rho}_{ab}} \dfrac{ \textbf{v}_{ab} \cdot \textbf{x}_{ab}}{\textbf{x}_{ab}^2 + (\eta h)^2}
+\Pi_{ab}=- \frac{\nu}{\bar{\rho}_{ab}} \dfrac{ \textbf{v}_{ab} \cdot \textbf{x}_{ab}}{\textbf{x}_{ab}^2 + \left(\eta h\right)^2}
 \label{eq:art-vis-original}
@@
 It has been shown that the shear viscosity coefficient $\eta = \frac{\rho \alpha h c}{8} $ and the bulk viscosity coefficient $ \zeta = \frac{5 \eta}{3}$ for two dimensional and $\eta = \frac{\rho \alpha h c}{10} $ , $ \zeta = \frac{5 \eta}{3}$ for three dimensional.
-An extra term was added to $\nu$ considering aspects of the dissipative term in shock solutions based on Riemann solvers and lead to:
+An extra term was added to $\nu$ considering aspects of the dissipative term in shock solutions based on Riemann solvers and lead to a new formulation of artificial viscosity.
+We adopt this new formulation in our simulation:
 \begin{equation}
@@
 \begin{equation}
-\mu_{ab} = \dfrac{h \textbf{v}_{ab} \cdot \textbf{x}_{ab}}{\textbf{x}_{ab}^2 + (\eta h)^2} 
-\end{equation}
-$\alpha$ and $\beta$ are free parameters that must be adjusted for each case.
+\mu_{ab} = \dfrac{h \textbf{v}_{ab} \cdot \textbf{x}_{ab}}{\textbf{x}_{ab}^2 + \left(\eta h\right)^2} 
+\end{equation}
+%$\alpha$ and $\beta$ are two parameters that are free to be adjusted for each case. 
+$\alpha$ and $\beta$ are two parameters that can be adjusted for different cases.
 $\alpha = 1$ and $\beta = 2$ are  recommended by Monaghan for best results.
 In our simulation, these two parameters are calibrated to  $\alpha = 0.3$ and $\beta = 0.6$.
-$\eta$ is usually taken as 0.1 for preventing singularities when $\textbf{x}_{ab} = 0$.
+$\eta$ is usually taken as 0.1 to prevent singularities.
+%when $\textbf{x}_{ab} = 0$.
 
 \subsection{Time step}
-The physical quantities (velocity, density, position and pressure) change every time step due to particle interactions.
-We have to consider the Courant condition, which is in spirit similar to the Courant condition for the grid-based methods, to determine the amount of the time step $\Delta t$.
-\begin{equation}
-\Delta t = \textrm{CFL} \min_a \bigg \lbrace \dfrac{[\frac{m_a}{\rho_a}]^{\frac{1}{d}}}{c_a} \bigg \rbrace
-\end{equation}
-Where $c_a$ is sound speed at particle a, $d$ is number of dimensions. First order Euler time integration, with $\textrm{CFL} = 0.2$ has been used in three dimensional calculations.
+The physical quantities (velocity, density and pressure) and particle position change every time step.
+The Courant condition, which is in spirit similar to the Courant condition for the mesh-based methods, is used to determine the time step $\Delta t$.
+\begin{equation}
+\Delta t = \textrm{CFL} \min_a \bigg \lbrace \dfrac{\left[\frac{m_a}{\rho_a}\right]^{\frac{1}{d}}}{c_a} \bigg \rbrace
+\end{equation}
+Where $c_a$ is sound speed at particle a, $d$ is number of dimensions. First order Euler integration, with $\textrm{CFL} = 0.2$, is used to advance in time.
 
 \subsection{Tensile instability and corrected derivatives}
-The classical SPH method suffered from tensile instability and boundary deficiency.
-To address these difficulties, Chen \citep{chen1999improvement} proposed a corrected SPH method.
-For 1D case, by employing a Taylor expansion for $A(x)$ about $x_a$, multiplying both sides by kernel function and then doing integration over the domain, gives
-\begin{equation}
-\int_{\Omega} A(x) w(x- x_a, h) dx = 
-A_a \int_{\Omega} w(x - x_a, h) dx +A_{xa} \int_{\Omega} (x-x_a) w(x - x_a, h) dx +...
-\end{equation}
-Ignoring derivative terms higher than first order, and writing the integral in particle approximation form:
-\begin{equation}
-A_a = \frac{\sum_b m_b \dfrac{A_b}{\rho_b} w(x_a-x_b, h)}{\sum_b m_b \dfrac{1}{\rho_b} w(x_a-x_b, h)}
+The classical SPH method was known to suffer from tensile instability and boundary deficiency.
+To address these difficulties, Chen \citep{chen1999improvement} proposed a corrected SPH formulation.
+For 1D case, employing a Taylor expansion for $A\left(x\right)$ about $x_a$, multiplying both sides by kernel function and then doing integration over the domain gives
+\begin{equation}
+\int_{\Omega} A\left(x\right) w\left(x- x_a, h\right) dx = 
+A_a \int_{\Omega} w\left(x - x_a, h\right) dx +A_{xa} \int_{\Omega} \left(x-x_a\right) w\left(x - x_a, h\right) dx +...
+\end{equation}
+Ignoring derivative terms higher than first order, and writing the integral in particle approximation form leads to:
+\begin{equation}
+A_a = \frac{\sum_b m_b \dfrac{A_b}{\rho_b} w\left(x_a-x_b, h\right)}{\sum_b m_b \dfrac{1}{\rho_b} w\left(x_a-x_b, h\right)}
 \label{eq:CSP-function-approximation-1d}
 \end{equation}
-Equation \ref{eq:CSP-function-approximation-1d} implies that $\sum_b m_b \dfrac{1}{\rho_b} w(x-x_b, h) = 1$, which can be viewed as the approximation form of Eq.
-\ref{eq:SPH-kernel-normalization-prop}.
-The first order derivative term can be obtained in a similar way and lead to:
-\begin{equation}
-\nabla A_a = \frac{\sum_b m_b \dfrac{A_b - A_a}{\rho_b} \nabla_a w(x_a-x_b, h)}{\sum_b m_b \dfrac{x_b - x_a}{\rho_b} \nabla_a w(x_a-x_b, h)}
+Equation (\ref{eq:CSP-function-approximation-1d}) implies that $\sum_b m_b \dfrac{1}{\rho_b} w\left(x-x_b, h\right) = 1$, which can be viewed as the approximation form of Eq.
+(\ref{eq:SPH-kernel-normalization-prop}).
+The first order derivative term can be obtained in a similar way:
+\begin{equation}
+\nabla A_a = \frac{\sum_b m_b \dfrac{A_b - A_a}{\rho_b} \nabla_a w\left(x_a-x_b, h\right)}{\sum_b m_b \dfrac{x_b - x_a}{\rho_b} \nabla_a w\left(x_a-x_b, h\right)}
 \end{equation}
@@
 As we discussed in the introduction section, special algorithms to treat and track the interface between different phases is required for mesh based methods to deal with multiphase problems.
-Worse, the engulfment process during plume development makes geometry of interfaces very complicated.
-Lagrangian SPH is able to handle multiphase problems without additional numerical effort for dealing with interfaces by simply tagging particles of each phase, which makes it suitable for volcanic plume simulation.
-Take the governing equation adopted in our model as an example.
+What's worse, the engulfment process during plume development makes geometry of interfaces very complicated.
+SPH is able to handle multiphase problems without additional numerical effort for dealing with interfaces by simply tagging particles of each phase, which makes it suitable for volcanic plume simulation.
+Take the governing equations adopted in our model as an example.
 Different phases of fluid (air and erupted material) are represented by two different sets of SPH particles (or discretization points).
@@
 \begin{equation}
-<\rho_{\alpha}^a>=\frac{\sum m_b w_{\alpha b} (h)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} (h) +\sum \frac{m_j}{\rho_j} w_{\alpha j} (h)} \label{eq:gov-sph-d1}
-\end{equation}
-\begin{equation}
-<\rho_\alpha^{sg}>=\frac{\sum_j m_j w_{\alpha j} (h)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} (h) +\sum \frac{m_j}{\rho_j} w_{\alpha j} (h)} \label{eq:gov-sph-d2}
+<\rho_{\alpha}^a>=\frac{\sum m_b w_{\alpha b} \left(h\right)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} \left(h\right) +\sum \frac{m_j}{\rho_j} w_{\alpha j} \left(h\right)} \label{eq:gov-sph-d1}
+\end{equation}
+\begin{equation}
+<\rho_\alpha^{sg}>=\frac{\sum_j m_j w_{\alpha j} \left(h\right)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} \left(h\right) +\sum \frac{m_j}{\rho_j} w_{\alpha j} \left(h\right)} \label{eq:gov-sph-d2}
 \end{equation}
@@
 \begin{equation}
-\rho_{\alpha}=\rho_{\alpha}^a=\frac{\sum m_b w_{\alpha b} (h)}{\sum \frac{m_b}{\rho_b} w_{\alpha b}}
+\rho_{\alpha}=\rho_{\alpha}^a=\frac{\sum m_b w_{\alpha b} \left(h\right)}{\sum \frac{m_b}{\rho_b} w_{\alpha b}}
 \end{equation}
@@
 \begin{equation}
-\rho_{\alpha}=\rho_{\alpha}^{sg}=\frac{\sum m_j w_{\alpha j} (h)}{\sum \frac{m_j}{\rho_j} w_{\alpha j}}
+\rho_{\alpha}=\rho_{\alpha}^{sg}=\frac{\sum m_j w_{\alpha j} \left(h\right)}{\sum \frac{m_j}{\rho_j} w_{\alpha j}}
 \end{equation}
@@
 (\ref{eq:gov-sph-d1}) and Eq.
-(\ref{eq:gov-sph-d1}) without any interface track or capture method.}
+(\ref{eq:gov-sph-d1}) without any other interface track or capture method.}
 \label{fig:SPH-multiple-density}
@@
 Typically, turbulence fluctuations in a volcanic column occurs at very different length scales, ranging from millimeters up to hundreds of meters.
-The entrainment of air and heat transfer are dominated by such turbulent fluctuations as turbulent exchange coefficient would be several magnitudes larger than molecular viscosity (and heat conduction) coefficient.
-One way to include enough turbulence in the model is to use the SPH-SPS (SPH sub-particle-scale) method.
-%\citep {gotoh2004sph, shao2006sph}. 
-Another way is to use extremely fine resolution ("the direct numerical simulation method") at the expense of much higher computational cost that results from both increase in number of particles and decrease in time steps constrained by the CFL condition.
+The entrainment of air and heat transfer are dominated by such turbulent fluctuations as turbulent exchange coefficients are several magnitudes larger than corresponding physical coefficients (molecular viscosity and heat conduction coefficient).
+One way to include enough turbulence in the model is to use the SPH-SPS (SPH sub-particle-scale) turbulence model.
+Another way is to use fine enough resolution (namely, the direct numerical simulation method) at the expense of much higher computational cost which results from both increase in number of particles and decrease in time steps constrained by the CFL (Courant–Friedrichs–Lewy) condition.
 Here we choose the SPH-SPS method.
-
-Even though free surface flow are in nature turbulent, detailed research in turbulence using SPH method is rather scarce.
+Among existing SPH-SPS turbulence models \citep{holm1999fluctuation, monaghan2002sph, violeau2007numerical, monaghan2011turbulence}, we adopt a LANS (Lagrangian Averaged Navier-Stokes) type turbulence model, the $SPH-\varepsilon$ turbulence model \citep{monaghan2011turbulence}.
+However, the $SPH-\varepsilon$ turbulence model is only for incompressible flow.
+In the following section, we will extend it for compressible flow.
+It is necessary to mention that all other existing SPH-SPS turbulence models \citep{holm1999fluctuation, monaghan2002sph, violeau2007numerical} also only focus on incompressible flow.
+
+%Even though free surface flow are in nature turbulent, detailed research on turbulence modeling in SPH method is rather scarce.
 Among the first papers on turbulence are the implementations of the $\alpha$ turbulence model \citep{holm1999fluctuation, monaghan2002sph} and the RANS (Reynolds averaged Navier-Stokes) approach \citep{violeau2007numerical}.
-A first wall-bounded LES (large eddy simulation) simulation with SPH was performed by \citet{issa2005numerical}.
-An improved version of $SPH-\alpha$ method is proposed by \citet{monaghan2011turbulence}, which we refer to as $SPH-\varepsilon$ method.
+The first wall-bounded LES (large eddy simulation) simulation with SPH was performed by \citet{issa2005numerical}.
+An improved version of $SPH-\alpha$ method  \citep{monaghan2002sph} is proposed by \citet{monaghan2011turbulence}, which we refer to as $SPH-\varepsilon$ method.
 We adopt $SPH-\varepsilon$ method in our model.
 Most of these papers constrain their focus to incompressible flow.
-$SPH-\alpha$ model \citep{monaghan2002sph} is a powerful extension of the XSPH algorithm which reduces disorder at short length scales and retains the constants of the motion.
+%$SPH-\alpha$ model \citep{monaghan2002sph} is a powerful extension of the XSPH algorithm which reduces disorder at short length scales and retains the constants of the motion.
 However, in that paper, energy conservation equation is discretized in a way that is exactly the same as usual SPH.
@@
 The recent improved version \citep{monaghan2011turbulence} of $SPH-\alpha$ method focuses on incompressible flow only.
-
-To simulate turbulence in the plume, we thus need to extend existing incompressible turbulent modeling methods to compressible flow.
-For compressible flow, the energy equation is coupled with the momentum equation and mass conservation equation.
-Averaging of thermal energy over space introduces some additional terms besides the Reynolds stress term into energy equation \citep{NASACompressibleTurbulence}.
-The averaged momentum equation for compressible flow are in the same form as that for incompressible flow, all of other additional terms besides a corresponding turbulent stress term are in the energy equation.
-Most turbulence modeling focuses on the Reynolds stress terms which are either solved directly or defined via a constitutive relation.
-Less attention is typically given to the other terms.
-Most commonly, a Reynolds analogy is used to model the turbulent heat flux.
-Simulations of heat transfer, or other scalar transfer, in turbulent flow simply involve adding transport terms for thermal energy or species concentration, at the expense of greater storage and longer computing times but without other difficulties \citep{cebeci2013analysis}.
-We adopt this strategy in our simulation of turbulent heat transfer with $SPH-\varepsilon$.
-%PDAC \citep{neri2003multiparticle} also adopted a Reynolds analogy in its turbulence modeling.
-The additional terms associated with molecular diffusion and turbulent transport in the energy equation are either modeled in different ways or neglected sometimes \citep{NASACompressibleTurbulence}.
-We neglect these terms in our simulation.
-
-\subsubsection{$SPH-\varepsilon$ turbulence model}
-In $SPH-\varepsilon$ method \citep{monaghan2011turbulence}, a turbulence model is constructed within the framework of SPH in such a way that general principles such as conservation of energy, momentum and circulation are satisfied using the ideas associated with the LANS (Lagrangian averaged Navier-Stokes equations).
-The basic idea of $SPH-\varepsilon$ is to determine a smoothed velocity $\widehat{\textbf{v}}$ by a linear operation on the unsmoothed velocity $\textbf{v}$.
-The SPH particles move with this smoothed velocity and the average motion of the fluid is determined by Eq.
-(\ref{eq:gov-update-pos-turbulence}):
+%
+
+\subsubsection{Langrangian average in $SPH-\varepsilon$}
+\citet{monaghan2011turbulence} constructed $SPH-\varepsilon$ turbulence model within the framework of SPH in such a way that general principles such as conservation of energy, momentum and circulation are satisfied using the ideas associated with the LANS turbulence modeling.
+The basic idea of $SPH-\varepsilon$ is to determine a smoothed (averaged) velocity $\widehat{\textbf{v}}$ by a linear operation on the unsmoothed velocity $\textbf{v}$.
+The SPH particles move with this smoothed velocity and hence the average motion of the fluid is determined by the averaged velocity $\widehat{\textbf{v}}$:
 \begin{equation}
@@
 \end{equation}
-Lagrangian equations contain extra term which represents the stresses induced by the smoothing. Once the form of the smoothing is chosen these stresses are determined. 
-%The overall effect is to produce a redistribution of energy without dissipation.
-
+
+Average of physical quantities over space introduces extra terms into the governing equations. Once the form of the smoothing (average) is chosen these extra terms are determined.
 The typical LANS model uses a smoothed velocity $\widehat{\textbf{v}}$ 
@@
 \begin{equation}
-\widehat{\textbf{v}}(\textbf{x})=\int \textbf{v}(\textbf{x} \prime)G(\vert \textbf{x} \prime - \textbf{x} \vert, l) d\textbf{x} \prime
+\widehat{\textbf{v}}\left(\textbf{x}\right)=\int \textbf{v}\left(\textbf{x} \prime\right)G\left(\vert \textbf{x} \prime - \textbf{x} \vert, l\right) d\textbf{x} \prime
 \end{equation}
@@
 \begin{equation}
-\int G(\vert \textbf{x} \prime - \textbf{x} \vert, l) d\textbf{x} \prime =1
-\end{equation}
-and is a member of a sequence of functions which tends to the $\delta$ function in the limit where $ l\rightarrow 0$. A typical example is Gaussian.
+\int G\left(\vert \textbf{x} \prime - \textbf{x} \vert, l\right) d\textbf{x} \prime =1
+\end{equation}
+and is a member of a sequence of functions which tends to the $\delta$ function in the limit when $ l\rightarrow 0$. A typical example is Gaussian.
 The length scale $l$ determines the characteristic width of the kernel and the distance over which the velocity is smoothed.
 
-It is common practice in $LANS-\alpha$ to use a differential equation for the smoothing rather than the integral form and finally reach to system of equations that need to be solved implicitly.
+It is a common practice in LANS to use a differential equation for the smoothing rather than the integral form and finally reach to a system of equations that need to be solved implicitly.
 In $SPH-\varepsilon$ method, a XSPH \citep{monaghan1989problem} smoothing is adopted which conserves linear and angular momentum.
-In this way solving of system of equations is avoided and it also makes the method simple to implement and cheap for computation.
-
-The discretized form of the momentum equation is obtained through lengthy derivation based on Lagrangian equations.
-Derivation and other discussions, such as conservation of momentum and energy, satisfactory of circulation theory, energy/momentum spectrum analysis based on Fourier transfer are detailed in reference \citep{monaghan2011turbulence}.
-Here we just have a brief summary of key steps.\\
-The smoothing that Monaghan adopted is:
-\begin{equation}
-\widehat{\textbf{v}}(\textbf{x})=\textbf{v}(\textbf{x})+ \epsilon \int (\textbf{v}(\textbf{x} \prime)-\textbf{v}(\textbf{x}))G(\vert \textbf{x} \prime - \textbf{x} \vert, l) d\textbf{x} \prime
+In this way, solving of the system of equations is avoided and it also makes the method simple to implement and cheap for computation.
+The discretized form of the momentum equation is obtained through lengthy derivation.
+Derivation and other discussions are detailed in reference \citep{monaghan2011turbulence}.
+Here we just have a brief summary of key steps.
+
+The smoothing adopted by \citet{monaghan2011turbulence}  is:
+\begin{equation}
+\widehat{\textbf{v}}\left(\textbf{x}\right)=\textbf{v}\left(\textbf{x}\right)+ \epsilon \int \left(\textbf{v}\left(\textbf{x} \prime\right)-\textbf{v}\left(\textbf{x}\right)\right)G\left(\vert \textbf{x} \prime - \textbf{x} \vert, l\right) d\textbf{x} \prime
 \end{equation}
@@
 \begin{equation} \label{eq:SPH-epsilon-filtering}
-\widehat{\textbf{v}}(\textbf{x})=\textbf{v}(\textbf{x})+\epsilon \sum_b m_b \dfrac{(\textbf{v}_b -\textbf{v})}{\rho _b} G(\vert \textbf{x} _b - \textbf{x} \vert, l)
+\widehat{\textbf{v}}\left(\textbf{x}\right)=\textbf{v}\left(\textbf{x}\right)+\epsilon \sum_b m_b \dfrac{\left(\textbf{v}_b -\textbf{v}\right)}{\rho _b} G\left(\vert \textbf{x} _b - \textbf{x} \vert, l\right)
 \end{equation}
@@
 \label{eq:replacement-in-turb-derive}
-\dfrac{G(\vert \textbf{x} _b - \textbf{x} _a \vert, l)}{\rho _b} \rightarrow \dfrac{K_{ab}}{M}
+\dfrac{G\left(\vert \textbf{x} _b - \textbf{x} _a \vert, l\right)}{\rho _b} \rightarrow \dfrac{K_{ab}}{M}
 \end{equation}
@@
 \label{eq:monaghan-mom-turb}
-\dfrac{d \textbf{v}_a}{dt} = -\sum_b [m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}\right) \bigtriangledown_aw_{a b}(h)] + \sum_b m_b \dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{M} \bigtriangledown_a K_{ab}
-\end{equation}
-Instead of using Eq. (\ref{eq:monaghan-mom-turb}) which is from \citet{monaghan2011turbulence}, we use Eq. (\ref{eq:replacement-in-turb-derive}). Notice that if $l$ is uniform and keep constant: 
+\dfrac{d \textbf{v}_a}{dt} = -\sum_b \left[ m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}\right) \bigtriangledown_aw_{a b}\left(h\right)\right] + \sum_b m_b \dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{M} \bigtriangledown_a K_{ab}
+\end{equation}
+Notice that if $l$ is uniform: 
 %Please notice that the above equation is slightly different from the original equation in Monaghan's paper (Eq.
@@
 \begin{equation}
-\nabla K_{ab} = \nabla (l^d G_{ab}) = l^d \nabla G_{ab}
-\end{equation}
-So the discretized momentum equation with $SPH-\varepsilon$ turbulence model in our simulation are:
+\nabla K_{ab} = \nabla \left(l^d G_{ab}\right) = l^d \nabla G_{ab}
+\end{equation}
+The discretized momentum equation with $SPH-\varepsilon$ turbulence model can be written in terms of $G_{ab}$ instead of $K_{ab}$:
 \begin{equation}
 \label{eq:SPH-mom-epsilon-turb}
-\dfrac{d \textbf{v}_a}{dt} = -\sum_b [m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}\right) \bigtriangledown_aw_{a b}(h)] + \sum_b m_b \Phi_{ab}\bigtriangledown_aG_{ab}(l)
+\dfrac{d \textbf{v}_a}{dt} = -\sum_b \left[m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}\right) \bigtriangledown_aw_{a b}\left(h\right)\right] + \sum_b m_b \Phi_{ab}\bigtriangledown_aG_{ab}\left(l\right)
 \end{equation}
@@
 \end{equation}
-and take $\varepsilon = 0.8$ as recommended by Monaghan.
+Which is the extra stress term induced by average. We take coefficient $\varepsilon$ as 0.8 following \citet{monaghan2011turbulence}.
+
+For compressible flow, the energy equation is coupled with the momentum equation and mass conservation equation.
+Averaging of thermal energy over space introduces some additional terms besides the stress term induced by velocity average.
+\citep{NASACompressibleTurbulence}.
+The averaged momentum equation for compressible flow are in the same form as that for incompressible flow, all of other additional terms, besides the corresponding velocity average induced stress term, show up in the energy equation.
+Most turbulence modeling focuses on the stress terms induced by average of velocity.
+These stress terms are usually either solved directly (for example, LANS methods) or defined via a constitutive relation (for example, large eddy simulation methods).
+Less attention is typically given to the other terms.
+Most commonly, a Reynolds analogy is used to model the turbulent exchange.
+Simulations of heat transfer, or other scalar transfer, in turbulent flow simply involve adding transport terms for thermal energy or species concentration, at the expense of greater storage and longer computing times but without other difficulties \citep{cebeci2013analysis}.
+We adopt this strategy.
+%PDAC \citep{neri2003multiparticle} also adopted a Reynolds analogy in its turbulence modeling.
+The additional terms associated with molecular diffusion and turbulent transport in the energy equation are either modeled in different ways or neglected sometimes \citep{NASACompressibleTurbulence}.
+We neglect these terms in our simulation.
 
 \subsubsection{Turbulent heat transfer}
-We adopt Reynolds analogy to get heat transfer coefficient due to turbulence.
-Prandtl number is defined as:
+We adopt Reynolds analogy to get the heat transfer coefficient due to turbulence.
+The Prandtl number is defined as:
 \begin{equation}
@@
 \end{equation}
-Where, $\mu$ is dynamic viscosity, $\kappa$ is thermal conductivity. And $\mu$  can be written in term of absolute viscosity (kinematic viscosity) as:
+Where, $\mu$ is the dynamic viscosity, $\kappa$ is the thermal conductivity. And $\mu$  can be written in term of the absolute viscosity (kinematic viscosity) as:
 \begin{align}
@@
 (\ref{eq:SPH-mom-epsilon-turb}), is the turbulent shear stress term.
-Recall that viscosity term can be discretized with SPH as shown in Eq. (\ref{eq:art-vis-original}). 
-Comparing of this term to physical viscous term for compressible gas shows that it has both bulk viscosity and shear viscosity, where shear viscosity coefficient is \citep{monaghan2005smoothed}:
+Recall that molecular viscous term can be discretized with SPH as shown in Eq. (\ref{eq:art-vis-original}). 
+Comparing of this term to molecular viscous term for compressible gas shows that it has both bulk viscosity and shear viscosity, where shear viscosity coefficient is \citep{monaghan2005smoothed}:
 \begin{equation}
@@
 \end{equation}
-The turbulent viscosity coefficient can be inferred from that form if we can reformulate the "turbulence stress" term in a form which is similar to molecular shear viscous term.
-Reformulate the "turbulence stress" term:
-\begin{equation}
- \sum_b \dfrac{\varepsilon}{2} \dfrac{m_b}{\rho_b} \textbf{v}_{ab} \cdot \textbf{v}_{ab} \nabla_a G_{ab}(l_a)= \sum_b \dfrac{\varepsilon}{2S} m_b \dfrac{\textbf{v}_{ab}}{\rho_b} \dfrac{S \textbf{v}_{ab} \cdot \textbf{x}_{ab}}{x_{ab}^2} \dfrac{x_{ab}^2}{\textbf{x}_{ab}} \nabla_a G_{ab}(l_a)
-\end{equation}
-Then the turbulent viscosity coefficient becomes
+The turbulent viscosity coefficient can be inferred from that formulation if we can reformulate the turbulent shear stress term in a form which is similar to molecular shear viscous term.
+Reformulate the turbulent shear stress term:
+\begin{equation}
+\label{eq:turb-stress-reformulate-to-artificial-vis}
+ \sum_b \dfrac{\varepsilon}{2} \dfrac{m_b}{\rho_b} \textbf{v}_{ab} \cdot \textbf{v}_{ab} \nabla_a G_{ab}\left(l_a\right)= \sum_b \dfrac{\varepsilon}{2S} m_b \dfrac{\textbf{v}_{ab}}{\rho_b} \dfrac{S \textbf{v}_{ab} \cdot \textbf{x}_{ab}}{x_{ab}^2} \dfrac{x_{ab}^2}{\textbf{x}_{ab}} \nabla_a G_{ab}\left(l_a\right)
+\end{equation}
+Then the turbulent viscosity coefficient can be inferred from Eq. (\ref{eq:turb-stress-reformulate-to-artificial-vis}).
 \begin{equation}
@@
 
-However, the above equation is correct only for the 1D situation. For 2D or 3D, it is not easy to get the explicit expression of viscosity. 
+However, the above equation is correct only for the 1D situation. For 2D or 3D, it is not easy to get an explicit expression. 
 We adopt an alternative way: obtaining a value for each pair of particles instead of persisting on getting an analytical expression.
-Choosing the smoothing function as the same as SPH kernel and the smoothing scale $l$ as the same as smoothing length $h$, the ratio between turbulent shear stress and physical shear stress is:
+Choosing the smoothing function as the same as SPH kernel and the smoothing length scale $l$ as the same as smoothing length $h$, the ratio between turbulent shear stress and physical shear stress is:
 \begin{equation}
@@
 \Upsilon_{ab} &= \dfrac{\dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\rho_b}}{\dfrac{S \nu}{\rho_{ab}} \frac{\textbf{v}_{ab} \cdot \textbf{x}_{ab}}{x_{ab}^2 + \eta^2 h_{ab}^2}} \\
- & = \dfrac{\varepsilon (x_{ab}^2 + \eta^2 h_{ab}^2)}{2 S \nu} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\textbf{v}_{ab} \cdot \textbf{x}_{ab}}
+ & = \dfrac{\varepsilon \left(x_{ab}^2 + \eta^2 h_{ab}^2\right)}{2 S \nu} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\textbf{v}_{ab} \cdot \textbf{x}_{ab}}
 \end{split}
@@
 \nu_{t,ab} &= \nu \Upsilon_{ab} \\
-&= \dfrac{\varepsilon (x_{ab}^2 + \eta^2 h_{ab}^2)}{2 S} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\textbf{v}_{ab} \cdot \textbf{x}_{ab}}
+&= \dfrac{\varepsilon \left(x_{ab}^2 + \eta^2 h_{ab}^2\right)}{2 S} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\textbf{v}_{ab} \cdot \textbf{x}_{ab}}
 \end{split}
@@
 \begin{equation}
-\kappa_{t,ab}=\dfrac{\varepsilon \overline{C_{p,ab}} \overline{\rho_{ab}} (x_{ab}^2 + \eta^2 h_{ab}^2) \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 S Pr_t\textbf{v}_{ab} \cdot \textbf{x}_{ab}}
-\end{equation}
-The term that used to prevent singularity now can be removed. 
+\kappa_{t,ab}=\dfrac{\varepsilon \overline{C_{p,ab}} \overline{\rho_{ab}} \left(x_{ab}^2 + \eta^2 h_{ab}^2\right) \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 S Pr_t\textbf{v}_{ab} \cdot \textbf{x}_{ab}}
+\end{equation}
+$\overline{C_{p,ab}}$ and $\overline{\rho_{ab}}$ are simply the arithmetic means of specific heat and density. The term that used to prevent singularity now can be removed. 
 \begin{equation}
@@
 \begin{equation}
-C_p \dfrac{dT}{dt} = \dfrac{1}{\rho} \nabla (\kappa \nabla T)
+C_p \dfrac{dT}{dt} = \dfrac{1}{\rho} \nabla \left(\kappa \nabla T\right)
 \end{equation}
@@
 \begin{equation}
-C_p \dfrac{dT}{dt} = \sum_b \dfrac{m_b}{\rho_a \rho_b} (\kappa_a + \kappa_b) (T_a - T_b) F_{ab} (h)
+C_p \dfrac{dT}{dt} = \sum_b \dfrac{m_b}{\rho_a \rho_b} \left(\kappa_a + \kappa_b\right) \left(T_a - T_b\right) F_{ab} \left(h\right)
 \end{equation}
@@
 \end{equation}
-And  $F_{ab} \leq 0$, which guarantees that heat flux flows from hot to cold.
-%JJ Monaghan \citep{cleary1999conduction} showed a simple alteration to the standard SPH formulation and ensures continuity of heat flux across discontinuities in material properties. 
+$F_{ab}$ is nonpositive, which guarantees that heat flux flows from hot to cold. 
 Plug the turbulent thermal conductivity into the heat conduction equation:
@@
 C_p \dfrac{dT}{dt}
-& = \sum_b \dfrac{m_b}{\rho_a \rho_b} (\kappa_a + \kappa_b) (T_a - T_b) F_{ab} (h) \\
- &= 2 \sum_b \dfrac{m_b}{\rho_a \rho_b} \dfrac{\overline{C_{p,ab}} \overline{\rho_{ab}} \varepsilon x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 Pr_t  S \textbf{v}_{ab} \cdot \textbf{x}_{ab} } (T_a - T_b) F_{ab} (h)
+& = \sum_b \dfrac{m_b}{\rho_a \rho_b} \left(\kappa_a + \kappa_b\right) \left(T_a - T_b\right) F_{ab} \left(h\right) \\
+ &= 2 \sum_b \dfrac{m_b}{\rho_a \rho_b} \dfrac{\overline{C_{p,ab}} \overline{\rho_{ab}} \varepsilon x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 Pr_t  S \textbf{v}_{ab} \cdot \textbf{x}_{ab} } \left(T_a - T_b\right) F_{ab} \left(h\right)
 \end{split}
-\end{equation}
-Notice that the number 2 at the front of the equation comes from integration approximation of second order derivative \citep {cleary1999conduction}. By further simplification, we get:
+\label{eq:turb-LANS-heat-conduct1}
+\end{equation}
+Notice that the number "2" in the front of Eq.
+(\ref{eq:turb-LANS-heat-conduct1}) comes from integration approximation of second order derivative \citep {cleary1999conduction}.
+By further simplification, we get:
 \begin{equation}
 C_p \dfrac{dT}{dt}
- =\dfrac{\varepsilon}{S  Pr_t}  \sum_b \dfrac{m_b}{\rho_a \rho_b} \dfrac{\overline{C_{p,ab}} \overline{\rho_{ab}} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\textbf{v}_{ab} \cdot \textbf{x}_{ab}} (T_a - T_b) F_{ab} (h)
+ =\dfrac{\varepsilon}{S  Pr_t}  \sum_b \dfrac{m_b}{\rho_a \rho_b} \dfrac{\overline{C_{p,ab}} \overline{\rho_{ab}} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\textbf{v}_{ab} \cdot \textbf{x}_{ab}} \left(T_a - T_b\right) F_{ab} \left(h\right)
 \end{equation}
@@
 \subsubsection{Discretized governing equations with $SPH-\varepsilon$ turbulence model}
-Plugging in the discretized turbulent stress term and turbulent heat transfer term into the momentum and energy equation, we get new discretized governing equation:
+Plugging in the discretized turbulent stress term and turbulent heat transfer term into the momentum and energy equation, we get new discretized governing equations:
 \begin{equation}
 \left\langle\dfrac{d \textbf{v}_{\alpha}}{d t}\right\rangle= 
--\sum_b \left[ m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}\right) \bigtriangledown_{\alpha}w_{\alpha b}(h)\right]
--\sum_j \left[m_j \left(\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j} - \Phi_{\alpha j}\right) \bigtriangledown_{\alpha}w_{\alpha j}(h)\right]
+-\sum_b \left[ m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}\right) \bigtriangledown_{\alpha}w_{\alpha b}\left(h\right)\right]
+-\sum_j \left[m_j \left(\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j} - \Phi_{\alpha j}\right) \bigtriangledown_{\alpha}w_{\alpha j}\left(h\right)\right]
 +\textbf{g} \label{eq:gov-sph-v}
@@
 \left\langle\dfrac{d e_{\alpha}}{d t}\right\rangle
-& = 0.5\sum_b \left[m_b \widehat{\textbf{v}_{\alpha b}} \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}\right) \bigtriangledown_{\alpha}w_{\alpha b}(h)\right]
- + 2 \sum_b \dfrac{m_b}{\rho_{\alpha} \rho_b} \kappa_{t,\alpha b} (T_{\alpha} - T_b) F_{\alpha b} (h) \\
- & +0.5\sum_j \left[m_j \widehat{\textbf{v}_{\alpha b}}\left(\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j} - \Phi_{\alpha j}\right) \bigtriangledown_{\alpha}w_{\alpha j}(h)\right]
- + 2 \sum_j \dfrac{m_j}{\rho_{\alpha} \rho_j} \kappa_{t,\alpha j} (T_{\alpha} - T_j) F_{\alpha j} (h)
+& = 0.5\sum_b \left[m_b \widehat{\textbf{v}_{\alpha b}}\left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}\right) \bigtriangledown_{\alpha}w_{\alpha b}\left(h\right)\right]
+ + 2 \sum_b \dfrac{m_b}{\rho_{\alpha} \rho_b} \kappa_{t,\alpha b} \left(T_{\alpha} - T_b\right) F_{\alpha b} \left(h\right) \\
+ & +0.5\sum_j \left[m_j \widehat{\textbf{v}_{\alpha b}}\left(\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j} - \Phi_{\alpha j}\right) \bigtriangledown_{\alpha}w_{\alpha j}\left(h\right)\right]
+ + 2 \sum_j \dfrac{m_j}{\rho_{\alpha} \rho_j} \kappa_{t,\alpha j} \left(T_{\alpha} - T_j\right) F_{\alpha j} \left(h\right)
 \end{split}
@@
 \subsection{Boundary conditions}
-All boundary conditions are imposed by ghost particles.
-Subfigure $b$ of Fig.
-\ref{fig:bc_and_domain_decomp} shows how boundaries are deployed.
-Physical quantities of particles, including real particles and different types of ghost particles, are updated differently.
+All boundary conditions are imposed by ghost particles. Subfigure $b$ of Fig. \ref{fig:bc_and_domain_decomp} shows how boundaries are deployed. 
+%Physical quantities of particles, including real particles and different types of ghost particles, are updated differently.
 \subsubsection{Wall boundary condition}
-Traditionally either ghost particles that mirror SPH particles across the boundary \citep {ferrari2009new} or boundary forces \citep {monaghan2009sph} have been used to impose the wall boundary conditions.
+Traditionally either ghost particles that mirror real particles across the boundary \citep {ferrari2009new} or boundary forces \citep {monaghan2009sph} have been used to impose the wall boundary conditions.
 One disadvantage of the latter  is that the boundary forces tend to corrupt the solution in the local neighborhood.
@@
 \begin{figure}
-\includegraphics[width=18cm]{t120_bc_proc.png}
-\caption{A cross section view of the simulation domain.
+\includegraphics[width=18cm]{t120_bc_proc}
+\caption{A cross section view of the simulation domain in $y-z$ plane at 66 seconds.
 Subfigure $a$ shows the mass fraction.
@@
 Subfigure $c$ shows the cross section view of domain decomposition based on SFC.
-The simulation is conducted on 360 processors (30 nodes with 12 cores each node), so there are 360 subdomains in total.
-The cut view shows a portion of them.
-\textcolor{red}{For showing the domain decomposition in the 3rd figure, it may be better to simply choose a unique color for each one (or just show the boundaries in black and white)} }
+The simulation is conducted on 12 processors, so there are 12 subdomains in total.
+The crosss section view shows a portion of them.}
 \label{fig:bc_and_domain_decomp}
@@
 \citet {ferrari2009new} adopted linked lists using pointers so that particles can be deleted or added during the simulation.
-Storage problems caused by fix-size arrays are thereby also eliminated.
+Storage problems caused by fixed-size arrays are thereby also eliminated.
 We define C++ classes which contain all data of particles and buckets.
-As for the management of data, we adopt hash tables to store pointers to particles and buckets, which gives us not only flexibility of deleting and adding elements, but also quicker access compared with linked lists.
+As for the management of data, we adopt hash tables to store pointers to particles and buckets, which gives us not only flexibility of deleting and adding of elements, but also quicker access compared with linked lists.
 Instead of using the "nature manner" to number particles, we adopt a space filling curve (SFC) based index to give each particle and background bucket an unique identifier -- a strategy known to preserve data locality at minimal cost.
@@
 
-We will publish separately further details about the data structure, domain decomposition, load balancing and domain adjusting.
+More details about the data structure, domain decomposition, load balancing and domain adjusting will be published separately.
 
@@
 \caption{The effect of domain adjusting on simulation time.
-The figure on the top shows simulation time without domain adjusting, the figure on the bottom shows simulation time with domain adjusting.
-Different bins represent simulation time up to specific physics time indicated by $x$ axis.}
+The figure on the top shows execution time without domain adjusting, the figure on the bottom shows execution time with domain adjusting.
+Different bins represent execution time up to specific physical time indicated by horizontal axis.}
 \label{fig:adj_vs_no}
@@
 The weak scalability test is conducted with the same initial domain and various smoothing lengths.
-Each simulation is run for 400 steps.
-The average number of real particles of each process is kept constant at $25900$.
+Each simulation runs for 400 time steps.
+The average number of real particles of each process keeps constant at $25900$.
 As shown in Fig.
@@
 \begin{equation}
-p=B[(\dfrac{\rho}{\rho_0})^{\gamma}-1]
+p=B\left[\left(\dfrac{\rho}{\rho_0}\right)^{\gamma}-1\right]
 \label{eq:EOS-Tait}
@@
 \end{equation}
-Where $c$ is speed of sound in the liquid.
-The energy equation is actually decoupled from momentum and mass conservation equation by using this equation of state (EOS).
+Where $c$ is the speed of sound in the liquid.
+The energy equation is actually decoupled from the momentum conservation equation and the mass conservation equation by using this equation of state (EOS).
 In addition, the "atmosphere" is assumed to be uniform and gravity is set to be zero.
@@
 	    \hline
-	    Vent velocity          & $m\cdot s^{-1}$  & 275               & 275 \\
-	    Vent gas mass fraction &             & 0.5               & 0.5 \\
-	    Vent Temperature       & $K$         & 1053              & 1053 \\
-	    Vent height            & $m$         & 1500              & 1500 \\
-	    Mass discharge rate    & $kg\cdot s^{-1}$ & $1.5 \times 10^9$ & $1.5 \times 10^9$\\
+	    Vent velocity          & $m\cdot s^{-1}$  & 500               & 275 \\
+	    Vent gas mass fraction &                  & 1.0               & 0.5 \\
+	    Vent Temperature       & $K$              & 273               & 1053 \\
+	    Vent height            & $m$              & 0                 & 1500 \\
+	    Mass discharge rate    & $kg\cdot s^{-1}$ & $5.47 \times 10^7$ & $1.5 \times 10^9$\\
 	    \hline
@@
 Material properties of water are used as material properties for both phases.
-The results are compared with experiments \citep {george1977turbulence, papanicolaou1988investigations} for validation n purposes.
-Experimental data of concentration and velocity distribution across the cross-section were fit into a Gaussian profile: $\dfrac{\varphi}{\varphi_c}=exp[-coef (\dfrac{r}{z})^2)]$ (solid line) by \citet{papanicolaou1988investigations} (even though there is no a priori reason to do so).
-\citet{papanicolaou1988investigations} also fit concentration distribution and jet width based on velocity along central line into a straight line : $\dfrac{\varphi_0}{\varphi_c}=slope (\dfrac{z}{D} + intercept)$ (dash line).
+The results are compared with experiments \citep {george1977turbulence, papanicolaou1988investigations} for validation purposes.
+Experimental data of concentration and velocity distribution across the cross-section were fit into a Gaussian profile: $\dfrac{\varphi}{\varphi_c}=exp \left[-coef\left( \left(\dfrac{r}{z}\right)^2\right)\right]$ (solid line) by \citet{papanicolaou1988investigations} even though the ctual profile are slightly different from the Gaussian profile.
+\citet{papanicolaou1988investigations} also fit concentration distribution and jet width based on velocity along central line into a straight line : $\dfrac{\varphi_0}{\varphi_c}=slope \left(\dfrac{z}{D} + intercept \right)$ (dash line).
 $\varphi$ is either velocity or concentration, the subscript $c$ represents the centreline, subscript $0$ represents the cross-sectionally averaged exit value.
@@
 \includegraphics[width=8cm]{JPUE_entrainment.png}
-\caption{Left figure shows particle distribution.
-The right figure shows the mass fraction of erupted material at the moment corresponding to left figure.
+\caption{The left figure shows particle distribution.
 Particles of phase 1 (blue) are gradually entrained and mixed with erupted particles (red) as jet flows down stream.
-What need to be noticed is that the mass of particle of phase 1 (blue) is 8 time of that of phase 2 (red).
-It is a trade off between resolution and computational expense to give particle of two phases different particles mass.}
+The right figure shows the mass fraction of erupted material at the moment corresponding to the left figure.}
+%What need to be noticed is that the mass of particle of phase 1 (blue) is 8 time of that of phase 2 (red).
+It is a trade off between resolution and computational expense to give particle of two phases different particles mass.
 \label{fig:Turb_mixing}
@@
 1B in \citep{costa2016results}).
-Wind speed is not used in our current model even though it is given (by one component speed from west to east and another component speed from south to north).
+%Wind speed is not used in our current model even though it is given (by one component speed from west to east and another component speed from south to north). 
 Vertical distribution of temperature, pressure and density is used to establish stratified atmosphere.
@@
 	    	Density of air at vent height               & $kg \cdot m^{-2}$       & 1.104   \\
-	    Pressure vent height                        & $Pa$              & 84363.4 \\
+	    Pressure at vent height                        & $Pa$              & 84363.4 \\
 	    \hline
@@
 \ref{fig:strong_local_temp} $\sim$ \ref{fig:strong_local_radius}.
-To get rid of significant fluctuations in time and space we did time averaging and spatial integration of the dynamic 3D flow fields by following \citet {cerminara2016large}.
-As particles distribute in the space in a disordered manner.
+To get rid of significant fluctuations in time and space we conducted a time averaging and spatial integration of the dynamic 3D flow fields by following \citet {cerminara2016large}.
+As particles distribute in a disordered manner in the space in SPH simulation results.
 We first project simulation results (on disordered particles) onto a pre-defined grid before doing time average and spatial integration.
-The interpolation method is the basic SPH kernel based interpolation.
+The project method is the basic SPH kernel based interpolation.
 
@@
 The profile for gas, which account for both air and vapor, shows a very similar tendency as that of entrained air.
-First of all, vapor condensation is not considered in our model.
-In addition, as we assume that erupted material behaves like a single phase water, the mass fraction of gas is simply a function of entrained air (Eq.
+Recall that vapor condensation is not considered in our model.
+In addition, we assume that erupted material behaves like a single phase fluid.
+So the mass fraction of gas is simply a function of entrained air (Eq.
 (\ref{eq:gas-frac-express})).
@@
 (\ref{eq:gas-frac-express}) does not hold for ATHAM.
-However, the profile of entrained air and profile of gas predicted by ATHAM are still very close to each other which implies that ignoring of water phase change is a valid assumption for eruptions similar to this test case (strong plume with erupted water fraction in erupted material less than 5\%), because air occupies a larger portion of the gas and plays a more significant role.
+However, the profile of entrained air and profile of gas predicted by ATHAM are still very close to each other which implies that ignoring of water phase change is a valid assumption for eruptions similar to this test case (strong plume with erupted water fraction in erupted material less than 5\%).
+This observation can be explained by the fact that air occupies a larger portion of the gas and ignoring of phase change of vapor (which is only a samll portion of gas) causes slight influence on plume development.
 As for mass fraction of solid, similarly, Eq.
@@
 (\ref{eq:solid-frac-express2}) hold for our model.
-PDAC, which treat particles of two different sizes as two separate phases, found a similar mass fraction profile.
-That implies that our model is at least valid for eruptions similar to the test case.
-
-\begin{equation}
-n_a + n_g = n_a + (1-n_a) n_{g0}
+PDAC, which treat particles of two different sizes as two separate phases, predicted a similar mass fraction profile.
+That implies that assumption of immediately dynamics equilibrium in our model is at least valid for eruptions similar to the test case.
+
+\begin{equation}
+\xi_a + \xi_g = \xi_a + \left(1-\xi_a\right) \xi_{g0}
 \label{eq:gas-frac-express}
@@
 \begin{equation}
-n_s = (1 - n_a) (1- n_{g0})
+\xi_s = \left(1 - \xi_a\right) \left(1- \xi_{g0}\right)
 \label{eq:solid-frac-express1}
@@
 \begin{equation}
-n_s = 1 - (n_a + n_g)
+\xi_s = 1 - \left(\xi_a + \xi_g\right)
 \label{eq:solid-frac-express2}
@@
 
-With more cool air entrained into the plume and mixing with hot erupted material, the temperature of the plume decreases as the height increases as shown in Fig.
+With more cool air entrained into the plume and mixing with the hot erupted material, the temperature of the plume decreases as the height increases as shown in Fig.
 \ref{fig:strong_local_temp}.
@@
 However, there is still a big disparity between the profiles of local variables of our model and SK-3D.
-One of the big differences between these two model is that we adopt a LANS type of turbulence model while SK-3D adopts a LES (large eddy simulation) turbulence model.
+One of the big differences between these two models is that we adopt a LANS type of turbulence model while SK-3D adopts a LES (large eddy simulation) turbulence model.
 This implies that turbulence model might play a critical role in plume simulation.
@@
 Advanced numerical techniques in SPH were exploited and tailored for this model.
-High performance computing was used to mitigate the tradeoff between accuracy (depends on comprehensiveness of the model, resolution, order of accuracy of numerical methods, scheme for time upgrading.) and simulation time (depends on comprehensiveness, resolution, order of accuracy of numerical methods, scheme for time upgrading ...
+High performance computing was used to mitigate the tradeoff between accuracy (depends on comprehensiveness of the model, resolution, order of accuracy of numerical methods, scheme for time upgrading.) and simulation time (depends on comprehensiveness of model, resolution, order of accuracy of numerical methods, scheme for time upgrading ...
 and computational techniques).
@@
 Top height and integrated local variables simulated by our model are consistent with both observation and simulation results of existing 3D plume models.
-Comparison between our results with that of SK-3D implies that turbulence model plays a significant role in plume modeling.
+Comparison between our results with these of SK-3D implies that turbulence model plays a significant role in plume modeling.
 
 Currently existing 3D models focus on certain aspect of the volcanic plume (PDAC on pyroclastic flow, ATHAM on microphysics, and SK-3D on entrainment with higher accuracy and higher order of accuracy) and hence, naturally, different assumptions were made in these models.
-However, these different aspects of volcanic plumes are not independent, but are instead coupled.
+However, these different aspects of volcanic plumes are not independent, but are actually coupled.
 For example, it has been illustrated that gas-particle non-equilibrium would introduce a previously unrecognized jet-dragging effect, which imposes great influence on plume development, especially for weak plumes \cite{cerminara2016large}.
@@
 For example, a large-particle module (LPM) was added to ATHAM to track the paths of rocky particles (pyroclastic or tephra) within the plume and predict where these particles fall \citep{kobs2009modeling}.
-We were also motivated by such an evolution of plume modeling to choose SPH as our numerical tool besides its ability in dealing with interfaces for multiphase flows.
-As mentioned in the introduction section, SPH method has good extensibility and adding new physics and phases requires much less modification of the code compared with mesh based methods.
+We were also motivated by such an evolution of plume modeling to choose SPH as our numerical tool.
+Besides its ability in dealing with interfaces for multiphase flows, as mentioned in the introduction section, SPH method has good extensibility and adding new physics and phases requires much less modification of the code compared with mesh based methods.
 Last but not least, the dramatic development of computational power makes it possible to establish a comprehensive model.

### 2017-05-03 Manuscripts_GMD.tex
New modification made, section 3.8 was re-written

@@
 
-The dynamics of explosive eruption clouds has been a central issue of volcanology science for a long time \citep{settle1978volcanic, wilson1978control}.
-The basic physics of volcanic plumes has been studied for more than half a century \citep{morton1956turbulent, settle1978volcanic, wilson1978control}.
+%The dynamics of explosive eruption clouds has been a central issue of volcanology science for a long time \citep{settle1978volcanic, wilson1978control}.
+%The basic physics of volcanic plumes has been studied for more than half a century \citep{morton1956turbulent, settle1978volcanic, wilson1978control}.
 The effects of magma type and vent conditions \citep{woods1988fluid, woods1995decompression}, atmospheric conditions \citep{ woods1993moist, sparks1997volcanic, bursik2001effect}, external surface water \citep{koyaguchi1996formation}, thermal disequilibrium, and particle fallout \citep{woods1991particle} have been assessed in increasing details based on observation, experiments and numerical studies.
@@
 
-Recently, a first order, nonequilibrium compressible 3D model, ASHEE \citep{cerminara2016ashee}, was introduced based on three dimensional N-phases Eulerian transportation equation.
+Recently, a first order, nonequilibrium compressible 3D model, ASHEE \citep{cerminara2016ashee}, was introduced based on three dimensional N-phases Eulerian transportation equations, which are a full set of mass, momentum and energy transport equations for a mixture of gas and dispersed particles.
 ASHEE is valid for low concentration and low Stokes number region and much faster than N-phases Eulerian model.
@@
 Since SPH is able to adaptively adjust the discretization and automatically construct the interface, SPH does not require additional numerical effort for interface construction and therefore is more suitable for volcanic plume simulation.
-\item Advection term in the Navier-Stokes equations does not appear explicitly in discretized formula of SPH (as illustrated in Eq.
-(\ref{eq:gov-nc-rho}) to Eq.
-(\ref{eq:gov-nc-e})).
-So advection term is treated exactly with SPH \citep{monaghan2005smoothed}.
+\item Advection term in the Navier-Stokes equations does not appear explicitly in discretized formula of SPH (as illustrated in Eq. (\ref{eq:gov-nc-rho}) to Eq. (\ref{eq:gov-nc-e})).
 \item It is easy to include various physics effect (like self gravity, radiative cooling and chemical reaction) into the model.
@@
 \subsection{Our contributions}
-SPH is still a relatively new method in computational fluid dynamics compared to mesh-based methods, such as finite volume method.
-Implementations of SPH for compressible multiphase turbulent flows are few.
-Several issues endemic to classical SPH are fixed in our implementation.
+Even though SPH has been known for several decades, implementations of SPH for compressible multiphase turbulent flows are few.
+Several issues endemic to classical SPH, like tensile instabilities,  compressible turbulence modeling and turbulent heat exchange, are fixed in our implementation.
 The most popular applications of SPH (and their original motivating application) has been in the simulation of free surface flow, such as breaking-waves and floods.
@@
 (\ref{eq:erupt_bc_v})).
-The correctness of the eruption boundary condition for density (Eq.
-(\ref{eq:erupt_bc_rho})) can be guaranteed by choosing mass of erupted ghost particles given distribution of these particles.
+The mass of eruption ghost particles is set to a value so that evaluation of Eq.
+(\ref{eq:ns-sph-d}) can return us a density that is consistent with the value given in Eq.
+(\ref{eq:erupt_bc_rho}).
 The internal energy associated with these particles are set to a value so that Eq.

### 2017-05-06 Manuscripts_GMD.tex
update Dr. Patra's modification

@@
 \begin{abstract}
-VATDs (Volcanic ash transport and dispersal models) used for volcanic ash forecasts sometimes take outputs of eruption plume models as source terms.
-The accuracy of source terms is crucial for forecasts from VATDs.
+VATDs (Volcanic ash transport and dispersal models) used for volcanic ash forecasts often take outputs of eruption plume models as source terms.
+The accuracy of these terms is crucial for forecasts from VATDs.
 All existing 3D (three dimensional) plume models described in the literature use mesh based methods.
@@
 In this model, erupted material is assumed to be well-mixed, and represented by one phase while air is another phase.
-We also assume immediately dynamics and thermodynamics equilibrium between air and erupted material and that the effect of wind is not significant.
-
-Several newly developed techniques in SPH are adopted and extended to address numerical challenges in simulating multiphase compressible turbulent flow with classical SPH method.
-The $SPH-\varepsilon$ turbulence model is adopted to capture mixing at unresolved scales.
-Heat exchange due to turbulence is then calculated with a Reynolds analogy.
-A corrected formalism of SPH is used to handle tensile instability and deficiency of particle distribution near the boundaries.
-We also propose feasible ways to impose velocity inlet and pressure outlet boundary conditions, both of which are scarce in traditional implementations of SPH.
-The core solver of our model is parallelized with MPI (message passing interface) obtaining good weak-scalability and strong-scalability.
-The model is first verified by comparing velocity and concentration distribution along axis and in cross section with experimental results of JPUE (jet or plume that is ejected from a nozzle into a uniform environment).
+We also assume dynamic and thermodynamical equilibrium between air and erupted material and that the effect of wind is not significant.
+
+Several newly developed techniques in SPH are adapted and extended to address numerical challenges in simulating multiphase compressible turbulent flow with classical SPH method.
+The $SPH-\varepsilon$ turbulence model is to capture mixing at unresolved scales.
+Heat exchange due to turbulence is then calculated by Reynolds analogy.
+A corrected SPH is used to handle tensile instability and deficiency of particle distribution near the boundaries.
+We also propose to impose velocity inlet and pressure outlet boundary conditions, both of which are scarce in traditional implementations of SPH.
+The core solver of our model is parallelized with MPI (message passing interface) obtaining good weak and strong scalability.
+The model is verified by comparing velocity and concentration distribution along axis and in cross section with experimental results of JPUE (jet or plume that is ejected from a nozzle into a uniform environment).
 Next, the top height of the Pinatubo eruption of 15 June 1991, as simulated by our model, is consistent with both observations and existing 3D plume models.
@@
 Different 1D models adopt different entrainment coefficients based on specific formulation or calibration against well-documented case studies.
-The feedback from plume to atmosphere is also ignored in 1D models.
+The feedback from plume to atmosphere is usually ignored in 1D models.
 Even though determination of essential parameters such as the entrainment is not based on first principles, such simple models nevertheless allow us to investigate the importance of physical mechanisms in a volcanic plume.
@@
 While these 1D models can generate well-matched results with 3D (three dimensional) models for weak plumes, much greater variability is observed for strong plume scenarios, especially for local variables \citep{costa2016results}.
-In addition, the need for greater skill in hazards forecasts especially where the plume model is used to generate source conditions for complex VATD models requires 2D (two dimensional) and 3D models.
+In addition, there is need for greater skill in hazards forecasts especially where the plume model is used to generate source conditions for complex 2D (two dimensional) and 3D VATD models.
 
@@
 Conservation equations for each phase are solved separately with the finite volume method.
-A parallel computing version of PDAC was developed lately \citep{ongaro2007parallel}.
-Advanced numerical techniques, such as second order scheme and semi-explicit time stepping, was also adopted afterwards to improve the accuracy of PDAC \citep{carcano2013semi}.
+A parallel computing version of PDAC was also developed \citep{ongaro2007parallel}.
+Advanced numerical techniques, such as a second order scheme and semi-explicit time stepping, was also adopted afterwards to improve the accuracy of PDAC \citep{carcano2013semi}.
 
@@
 
-Besides adding to their special strengths (ATHAM has been adding more and more microphysics, PDAC was extended to consider more phases), these models are learning from each other.
+Besides adding to their special strengths (ATHAM has been adding more and more microphysics, PDAC was extended to consider more phases), these models are also adding core strengths.
 PDAC development has begun to include the effect of microphysics into the model while ATHAM development has extended its ability to modeling pyroclastic flow.
@@
 \begin{itemize}
-\item Better investigation of mixing phenomena;
+\item better investigation of mixing phenomena;
 \item accurate modeling of the development of ZFE (Zone of Flow Establishment), ZEF (Zone of Established Flow) investigation and relation to column collapse and the questions relating to the development of entrainment; %and proper characterization through numerical experimentation.
@@
 \item Advection term in the Navier-Stokes equations does not appear explicitly in discretized formula of SPH (as illustrated in Eq. (\ref{eq:gov-nc-rho}) to Eq. (\ref{eq:gov-nc-e})).
-\item It is easy to include various physics effect (like self gravity, radiative cooling and chemical reaction) into the model.
+\item It is easy to include various physics effects (like self gravity, radiative cooling and chemical reaction) into the model.
 It does not require a major overhaul and re-tooling every time new physics is introduced \citep{monaghan1995sph}.
@@
 The core physics, such as entrainment of air and thermal expansion, are essential for all plume modeling while some other physics, such as water condensation and aggregation etc., are important in specific scenarios.
-As an initial effort on exploiting advantages of SPH in volcanic plume modeling, we focus on capturing basic features in plume development using a robust numerically and comuptationaly efficient framework with support for scalable parallel computing.
+As an initial effort on exploiting advantages of SPH in volcanic plume modeling, we focus on capturing basic features in plume development using a robust numerically and computationally efficient framework with support for scalable parallel computing.
 %So our current model is far away from comprehensive.
@@
 \item Imposition of some types of boundary conditions (such as eruption boundary condition) requires dynamically adding and removing of particles during simulation.
-To address this issue, we adopt an efficient data management scheme based on time dependent SFC (space filling curve) and hash table.
+To address this issue, we adopt an efficient data management scheme based on time dependent SFC (space filling curve) induced indexing and hash table.
 The computational cost is further reduced by adjusting simulation domain adaptively.
@@
 (\ref{eq:gov-cs-e}), are directly discretized.
-For SPH, governing equations in Lagrange form is needed.
-By deducting kinetic energy from energy equation, subtracting mass conservation from momentum equation, combining transient term and advection term into material derivative term, the governing equations are put into the final form, in which advection term does not show up explicitly.
+For SPH, governing equations in Lagrange form are needed.
+By deducting kinetic energy from energy equation, subtracting mass conservation from momentum equation, combining transient term and advection term into material derivative term, the governing equations are put into the final form, in which advection term does not appear explicitly.
 \begin{align}
@@
 The physical laws (such as conservation laws of mass, momentum and energy) written in the form of PDEs (partial differential equations)  or ODEs (ordinary differential equations) need to be transformed into the Lagrangian particle formalism of SPH.
-Using a kernel function that provides the weighted estimation of the field variables at any point, the integral equations are estimated as sums over particles in a compact subdomain defined by the support of the kernel function defined with respect to the discretization points.
+Using a kernel function that provides the weighted estimation of the field variables at any point, the integral equations are estimated as sums over particles in a compact subdomain defined by the support of the kernel function associated with the discretization points.
 Thus, field variables associated to the particle are updated based on its neighbors.
 Each kernel function has a compact support determined by smoothing length of each particle.
-There are several review papers by \citet{monaghan1992smoothed, monaghan2005smoothed, rosswog2009astrophysical, price2012smoothed, monaghan2012smoothed}, giving a pretty comprehensive view over SPH, we only refer here to the representation of the constitutive equations in SPH and put more focus on specific numerical techniques for plume modeling.
+There are several review papers by \citet{monaghan1992smoothed, monaghan2005smoothed, rosswog2009astrophysical, price2012smoothed, monaghan2012smoothed}, giving a pretty comprehensive view over SPH.
+We only refer here to the representation of the constitutive equations in SPH and put more focus on specific numerical techniques for plume modeling.
 
@@
 There are several procedures for discretizing governing equations (PDEs or ODEs) with SPH.
-There is slight difference between discretized formalism obtained by different routes.
 We present here one of them following \citet{monaghan1992smoothed, monaghan2005smoothed, monaghan2012smoothed}.
@@
 In $SPH-\varepsilon$ method, a XSPH \citep{monaghan1989problem} smoothing is adopted which conserves linear and angular momentum.
-In this way, solving of the system of equations is avoided and it also makes the method simple to implement and cheap for computation.
+In this way, solving of a system of equations is avoided and it also makes the method simple to implement and cheap for computation.
 The discretized form of the momentum equation is obtained through lengthy derivation.
-Derivation and other discussions are detailed in reference \citep{monaghan2011turbulence}.
-Here we just have a brief summary of key steps.
+Derivation and other discussions are available in the literature -- see for e.g.
+\citep{monaghan2011turbulence}.
+Here we provide a brief summary of key steps.
 
@@
 (\ref{eq:SPH-mom-epsilon-turb}), is the turbulent shear stress term.
-Recall that molecular viscous term can be discretized with SPH as shown in Eq. (\ref{eq:art-vis-original}). 
-Comparing of this term to molecular viscous term for compressible gas shows that it has both bulk viscosity and shear viscosity, where shear viscosity coefficient is \citep{monaghan2005smoothed}:
+Recall that molecular viscosity can be discretized with SPH as shown in Eq. (\ref{eq:art-vis-original}). 
+It has been shown that the discretized molecular viscosity has both bulk viscosity and shear viscosity, where shear viscosity coefficient is \citep{monaghan2005smoothed}:
 \begin{equation}
@@
 \end{equation}
-The turbulent viscosity coefficient can be inferred from that formulation if we can reformulate the turbulent shear stress term in a form which is similar to molecular shear viscous term.
+The turbulent viscosity coefficient can be inferred from that formulation if we can reformulate the turbulent shear stress term in a form which is similar to the molecular shear term.
 Reformulate the turbulent shear stress term:
@@
 \end{equation}
-Please note that turbulent viscosity term has opposite sign with molecular viscosity term in discretized momentum equation and there is a minus sign in the expression of $\Pi_{ab}$, they cancell out.
+Please note that turbulent viscosity term has opposite sign with molecular viscosity term in discretized momentum equation and there is a minus sign in the expression of $\Pi_{ab}$, and they cancel out.
 
@@
 \end{equation}
-$\Upsilon_{ab}$ is essentially equivalent to the ratio between turbulent viscous effect attributed by particle b on particle a and molecular viscous effect attributed by particle b on particle a.
+$\Upsilon_{ab}$ is essentially equivalent to the ratio between turbulent viscous effect of particle b on particle a and molecular viscous effect of particle b on particle a.
 Turbulent viscosity can be easily obtained by:
@@
 \end{equation}
-$\overline{C_{p,ab}}$ and $\overline{\rho_{ab}}$ are simply the arithmetic means of specific heat and density. The term that used to prevent singularity now can be removed. 
+$\overline{C_{p,ab}}$ and $\overline{\rho_{ab}}$ are simply the arithmetic means of specific heat and density. The term used to prevent singularity now can be removed. 
 \begin{equation}
@@
 More implementation details about this method can be found in \citep {kumar2013parallel}.
-As these wall ghost particles keep stationary, there is no mass flux on the boundary (Eq.
+As these wall ghost particles are stationary, there is no mass flux on the boundary (Eq.
 (\ref{eq:wall_bc_rho}) and (\ref{eq:wall_bc_xi})).
@@
 Subfigure $c$ of Fig.
-\ref{fig:bc_and_domain_decomp} shows a cross section view of domian decomposition.
+\ref{fig:bc_and_domain_decomp} shows a cross section view of domain decomposition.
 
@@
 However, the profile of entrained air and profile of gas predicted by ATHAM are still very close to each other which implies that ignoring of water phase change is a valid assumption for eruptions similar to this test case (strong plume with erupted water fraction in erupted material less than 5\%).
-This observation can be explained by the fact that air occupies a larger portion of the gas and ignoring of phase change of vapor (which is only a samll portion of gas) causes slight influence on plume development.
+This observation can be explained by the fact that air occupies a larger portion of the gas and ignoring of phase change of vapor (which is only a small portion of gas) causes slight influence on plume development.
 As for mass fraction of solid, similarly, Eq.
@@
 PDAC, which treat particles of two different sizes as two separate phases, predicted a similar mass fraction profile.
-That implies that assumption of immediately dynamics equilibrium in our model is at least valid for eruptions similar to the test case.
+That implies that assumption of immediate dynamic equilibrium in our model is at least valid for eruptions similar to the test case.
 

### 2017-05-07 Manuscripts_GMD.tex
a small edit

@@
 Another 3D model, ATHAM (Active Tracer High-Resolution Atmospheric Model) \citep{oberhuber1998volcanic} focuses more on microphysics within the plume.
-As pyroclastic flow is not the initial concern of ATHAM, immediately dynamics and thermodynamics equilibrium is assumed in ATHAM.
+As pyroclastic flow is not the initial concern of ATHAM, immediate dynamic and thermodynamical equilibrium is assumed in ATHAM.
 The dynamic core of ATHAM solves the compressible Euler equations for momentum, pressure and temperature of the gas particle mixture \citep{oberhuber1998volcanic}.
@@
 Air (also a mixture of different gases) is assumed to be another phase (phase 1).
-Immediately thermodynamics equilibrium is assumed so that no separate energy equation is needed for each phase.
+Immediate thermodynamical equilibrium is assumed so that no separate energy equation is needed for each phase.
 As a result, there is only one energy equation for both phases (heat exchange term between different phases does not show up under this assumption).
-Immediately dynamics equilibrium is assumed so that no separate momentum equation is needed for each phase.
+Immediate dynamic equilibrium is assumed so that no separate momentum equation is needed for each phase.
 As a result, there is only one vector momentum equation for both phases (drag force term does not show up with this assumption).
@@
 \subsection{Governing equations}
-Based on above assumptions, the governing equations of our model are given as following, which is the same as the governing equations of SK-3D \citep{suzuki2005numerical}.
+Based on above assumptions, the governing equations of our model are given as (which is the same as the governing equations of SK-3D \citep{suzuki2005numerical}):
 \begin{align}

### 2017-05-09 Manuscripts_GMD.tex
Ready for submission

@@
 \begin{abstract}
-VATDs (Volcanic ash transport and dispersal models) used for volcanic ash forecasts often take outputs of eruption plume models as source terms.
+VATDs (Volcanic ash transport and dispersion models) used for volcanic ash forecasts often take outputs of eruption plume models as source terms.
 The accuracy of these terms is crucial for forecasts from VATDs.
@@
 In this model, erupted material is assumed to be well-mixed, and represented by one phase while air is another phase.
-We also assume dynamic and thermodynamical equilibrium between air and erupted material and that the effect of wind is not significant.
+We also assume dynamic and thermodynamic equilibrium between air and erupted material and that the effect of wind is not significant.
 
@@
 The core solver of our model is parallelized with MPI (message passing interface) obtaining good weak and strong scalability.
-The model is verified by comparing velocity and concentration distribution along axis and in cross section with experimental results of JPUE (jet or plume that is ejected from a nozzle into a uniform environment).
+The model is verified by comparing velocity and concentration distribution along the central axis and on the transverse cross with experimental results of JPUE (jet or plume that is ejected from a nozzle into a uniform environment).
 Next, the top height of the Pinatubo eruption of 15 June 1991, as simulated by our model, is consistent with both observations and existing 3D plume models.
@@
 
-During volcanic eruptions, VATDs (volcanic ash transport and dispersal models) are used to forecast the location and movement of ash clouds at timescales that range from hours to days.
+During volcanic eruptions, VATDs (volcanic ash transport and dispersion models) are used to forecast the location and movement of ash clouds at timescales that range from hours to days.
 VATDs use eruption source parameters, such as plume height, mass eruption rate, duration, and the mass fraction distribution of erupted debris finer than about $4 \Phi$ (or $63  \mu$m), which can remain in the cloud for many hours or days.
@@
 While these 1D models can generate well-matched results with 3D (three dimensional) models for weak plumes, much greater variability is observed for strong plume scenarios, especially for local variables \citep{costa2016results}.
-In addition, there is need for greater skill in hazards forecasts especially where the plume model is used to generate source conditions for complex 2D (two dimensional) and 3D VATD models.
+In addition, there is need for greater skills in hazards forecasts especially where the plume model is used to generate source conditions for complex 2D (two dimensional) and 3D VATD models.
 
@@
 Another 3D model, SK-3D \citep{suzuki2005numerical} is a 3D time-dependent fluid dynamics model that attempts to reproduce the entrainment process of eruption clouds with relatively simple physics but with high order numerical accuracy and high spatial resolution.
-A series of simulations based on SK-3D was reported, including determination of the entrainment coefficients of eruption columns as a function of height \citep{suzuki2010numerical}, the effect of wind field on entrainment coefficient \citep{suzuki20133d}, establishment of the relationship between the observable quantities of the eruption clouds and the eruption conditions at the vent \citep{suzuki2009three} (previously, such relationship is based on 1D models and greatly depends on empirical constants), and investigation of the effect of the intensity of turbulence in the umbrella cloud on dispersion and sedimentation of tephra \citep{koyaguchi2009effect}.
+A series of simulations based on SK-3D was reported, including establishment of the relationship between the observable quantities of the eruption clouds and the eruption conditions at the vent \citep{suzuki2009three}, investigation of the effect of the intensity of turbulence in the umbrella cloud on dispersion and sedimentation of tephra \citep{koyaguchi2009effect}, determination of the entrainment coefficients of eruption columns as a function of height \citep{suzuki2010numerical} and investigation of the effect of wind field on entrainment coefficient \citep{suzuki20133d}.
 
@@
 Another 3D model, ATHAM (Active Tracer High-Resolution Atmospheric Model) \citep{oberhuber1998volcanic} focuses more on microphysics within the plume.
-As pyroclastic flow is not the initial concern of ATHAM, immediate dynamic and thermodynamical equilibrium is assumed in ATHAM.
+As pyroclastic flow is not the initial concern of ATHAM, dynamic and thermodynamical equilibrium is assumed in ATHAM.
 The dynamic core of ATHAM solves the compressible Euler equations for momentum, pressure and temperature of the gas particle mixture \citep{oberhuber1998volcanic}.
-The subgrid-scale turbulence closure scheme differentiates between the horizontal and vertical directions \citep{herzog2003prognostic} to capture turbulent mixing.
+The subgrid-scale turbulence closure scheme differentiates between the horizontal and vertical directions \citep{herzog2003prognostic} captures turbulent mixing.
 The cloud microphysics predicts the mass of hydrometeors in liquid and ice phase \citep{herzog1998effect}.
-Additional modules include radiation \citep{langmann1997radiative}, gas phase chemistry \citep{trentmann2002simulation} and gas scavenging by hydrometeors \citep{textor2003injection} were added lately.
+Additional modules, including gas phase chemistry \citep{trentmann2002simulation} and gas scavenging by hydrometeors \citep{textor2003injection} were added lately.
 A further extension was made to include particle aggregation \citep{textor2006volcanic1, textor2006volcanic2}.
@@
 \subsection{Features of SPH}
-To the best of our knowledge, all of the existing 3D plume models are mesh based Eulerian methods, and there are no 3D plume models based on mesh free Lagrangian methods.
+To the best of our knowledge, all of the existing 3D plume models use mesh based Eulerian methods, and there are no 3D plume models based on mesh free Lagrangian methods.
 Lagrangian methods have several features that are suitable for volcanic plume simulation that we outline below.
@@
 Air (also a mixture of different gases) is assumed to be another phase (phase 1).
-Immediate thermodynamical equilibrium is assumed so that no separate energy equation is needed for each phase.
+Thermodynamic equilibrium is assumed so that no separate energy equation is needed for each phase.
 As a result, there is only one energy equation for both phases (heat exchange term between different phases does not show up under this assumption).
-Immediate dynamic equilibrium is assumed so that no separate momentum equation is needed for each phase.
+Dynamic equilibrium is assumed so that no separate momentum equation is needed for each phase.
 As a result, there is only one vector momentum equation for both phases (drag force term does not show up with this assumption).
@@
 \end{equation}
-Where, $C_v$ is specific heat with constant volume, $R$ is gas constant. $\xi$ with subscript is mass fraction of corresponding constituent. The subscript 
+Where, $C_v$ is the specific heat with constant volume, $R$ is the gas constant. $\xi$ with subscript is the mass fraction of corresponding constituent. The subscript 
 $m$ represents mixture of ejected material and air, $s$ represents solid portion in the ejected material, $g$ represents gas portion in the ejected material and $a$ represents air.
-$\xi_{g0}$ is the mass fraction of vapor in erupted material.
+$\xi_{g0}$ is the mass fraction of vapor in the erupted material.
 
@@
 (\ref{eq:gov-cs-rho}) to Eq.
-(\ref{eq:gov-cs-e}), are actually equivalent, as deriving the Lagrangian form from the Eulerian is purely based on mathematics without any additional physical assumptions.
+(\ref{eq:gov-cs-e}), are a\dfrac{•}{•}ly equivalent, as deriving the Lagrangian form from the Eulerian is purely based on mathematics without any additional physical assumptions.
 
@@
 The radius of vent is determined from $\rho$, $\dot M$ and $\textbf{v}$.
-Equation (\ref{eq:erupt_bc_rho}) ~ (\ref{eq:erupt_bc_e}) gives velocity outlet boundary condition wrote in terms of primitive variables.
+Equation (\ref{eq:erupt_bc_rho}) to (\ref{eq:erupt_bc_e}) gives velocity outlet boundary condition wrote in terms of primitive variables.
 \begin{align}
@@
 SPH is a mesh-free Lagragian method.
-In SPH, the domain (control mass) is discretized by a set of particles or discretization points and the position of each particle (discretization point) is updated at every time step based on the motion computed.
+In SPH, the domain is discretized by a set of particles or discretization points and the position of each particle is updated at every time step based on the motion computed.
 Approximation of all field variables (velocity, density and pressure, ect.) is obtained by interpolation based on discretization points.
@@
 <\rho_a> = \sum_b m_b w_{ab} \left(h\right) \label{eq:ns-sph-d} \\
-\left\langle\dfrac{d \textbf{v}_a}{d t}\right\rangle = \sum_b m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}\right) \nabla_a w_{a b}\left(h\right) +\textbf{g} \label{eq:ns-sph-v} \\
+\left\langle\dfrac{d \textbf{v}_a}{d t}\right\rangle = -\sum_b m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}\right) \nabla_a w_{a b}\left(h\right) +\textbf{g} \label{eq:ns-sph-v} \\
 \left\langle\dfrac{d e_a}{d t}\right\rangle=
@@
 <\rho_a> = \sum m_b w_{ab} \left(h\right) + \dot{\rho}\left(\textbf{x},t\right)\label{eq:ns-source-sph-d} \\
-\left\langle\dfrac{d \textbf{v}_a}{d t}\right\rangle= \sum_b m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}\right) \nabla_a w_{a b}\left(h\right) +\textbf{g} + D \sum	_b m_b \dfrac{\textbf{v}_b - \textbf{v}_a}{\rho_b} \label{eq:ns-drag-sph-v} \\
+\left\langle\dfrac{d \textbf{v}_a}{d t}\right\rangle= -\sum_b m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}\right) \nabla_a w_{a b}\left(h\right) +\textbf{g} + D \sum	_b m_b \dfrac{\textbf{v}_b - \textbf{v}_a}{\rho_b} \label{eq:ns-drag-sph-v} \\
 \left\langle\dfrac{d e_a}{d t}\right\rangle=
@@
 Here we choose the SPH-SPS method.
-Among existing SPH-SPS turbulence models \citep{holm1999fluctuation, monaghan2002sph, violeau2007numerical, monaghan2011turbulence}, we adopt a LANS (Lagrangian Averaged Navier-Stokes) type turbulence model, the $SPH-\varepsilon$ turbulence model \citep{monaghan2011turbulence}.
-However, the $SPH-\varepsilon$ turbulence model is only for incompressible flow.
+Among existing SPH-SPS turbulence models \citep{holm1999fluctuation, monaghan2002sph, violeau2007numerical, monaghan2011turbulence}, we adopt a LANS type turbulence model, the $SPH-\varepsilon$ turbulence model \citep{monaghan2011turbulence}.
+However, the $SPH-\varepsilon$ turbulence model was proposed only for incompressible flow.
 In the following section, we will extend it for compressible flow.
@@
 Most turbulence modeling focuses on the stress terms induced by average of velocity.
-These stress terms are usually either solved directly (for example, LANS methods) or defined via a constitutive relation (for example, large eddy simulation methods).
+These stress terms are usually either solved directly (for example, LANS methods) or defined via a constitutive relation (for example, large eddy simulation method).
 Less attention is typically given to the other terms.
@@
 \left\langle\dfrac{d \textbf{v}_{\alpha}}{d t}\right\rangle= 
--\sum_b \left[ m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}\right) \bigtriangledown_{\alpha}w_{\alpha b}\left(h\right)\right]
--\sum_j \left[m_j \left(\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j} - \Phi_{\alpha j}\right) \bigtriangledown_{\alpha}w_{\alpha j}\left(h\right)\right]
+-\sum_b \left[ m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b}^{\beta} - \Phi_{\alpha b}\right) \bigtriangledown_{\alpha}w_{\alpha b}\left(h\right)\right]
+-\sum_j \left[m_j \left(\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j}^{\beta} - \Phi_{\alpha j}\right) \bigtriangledown_{\alpha}w_{\alpha j}\left(h\right)\right]
 +\textbf{g} \label{eq:gov-sph-v}
@@
 \left\langle\dfrac{d e_{\alpha}}{d t}\right\rangle
-& = 0.5\sum_b \left[m_b \widehat{\textbf{v}_{\alpha b}}\left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b} - \Phi_{\alpha b}\right) \bigtriangledown_{\alpha}w_{\alpha b}\left(h\right)\right]
+& = 0.5\sum_b \left[m_b \widehat{\textbf{v}_{\alpha b}}\left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b}^{\beta} - \Phi_{\alpha b}\right) \bigtriangledown_{\alpha}w_{\alpha b}\left(h\right)\right]
  + 2 \sum_b \dfrac{m_b}{\rho_{\alpha} \rho_b} \kappa_{t,\alpha b} \left(T_{\alpha} - T_b\right) F_{\alpha b} \left(h\right) \\
- & +0.5\sum_j \left[m_j \widehat{\textbf{v}_{\alpha b}}\left(\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j} - \Phi_{\alpha j}\right) \bigtriangledown_{\alpha}w_{\alpha j}\left(h\right)\right]
+ & +0.5\sum_j \left[m_j \widehat{\textbf{v}_{\alpha b}}\left(\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j}^{\beta} - \Phi_{\alpha j}\right) \bigtriangledown_{\alpha}w_{\alpha j}\left(h\right)\right]
  + 2 \sum_j \dfrac{m_j}{\rho_{\alpha} \rho_j} \kappa_{t,\alpha j} \left(T_{\alpha} - T_j\right) F_{\alpha j} \left(h\right)
@@
 Ghost particles are reflected into the domain and physical quantities are calculated at these reflected positions by SPH interpolations.
-It should be noted that wall ghost particles should not be counted when computing physics properties of these reflected positions.
+It should be noted that wall ghost particles should not be counted when computing physical properties of these reflected positions.
 Assign all properties, except for velocity, at the corresponding reflected position to the ghost particle.
@@
 Any implementation of SPH code requires efficient searching and updating of neighbors during simulation.
-Of the many choices possible for this we adopt a background grid which was proposed by \citet {monaghan1985refined} and is quite popular in parallel SPH.
+Of the many possible choices we adopt a background grid which was proposed by \citet {monaghan1985refined} and is quite popular in parallel SPH.
 This background grid is also used for domain decomposition in SPH.
@@
 
-In both SPHysics and DualSPHysics \citep {crespo2015dualsphysics}, the physical quantities of each particle (position, velocity, density...) are stored in arrays.
+In DualSPHysics \citep {crespo2015dualsphysics}, the physical quantities of each particle (position, velocity, density...) are stored in arrays.
 The particles (and the arrays with particle data) are reordered following the order of the cells.
@@
 \includegraphics[scale=0.35]{strong}
-\caption{Strong scalability test results}
+\caption{Strong scalability test result}
 \label{fig:strong_scale}
@@
 \includegraphics[scale=0.35]{weak_scale}
-\caption{Weak scalability test results}
+\caption{Weak scalability test result}
 \label{fig:weak_scale}
@@
 
-Performance tests have been carried out at CCR (the computational cluster of Center for Computational Research) at Buffalo.
+Performance tests have been carried out on the computational cluster of CCR (Center for Computational Research) at Buffalo.
 Intel Xeon E5645 CPUs running at 2.40GHz clock rate with 4GB memory per core on a Q-Logic Infiniband is used in these tests.
@@
 A domain adjusting algorithm is designed and implemented in our code.
-Figure \ref{fig:adj_vs_no} shows that simulation time of the test problem is greatly reduced when we adopt the domain adjusting strategy in our code.
+Figure \ref{fig:adj_vs_no} shows that simulation time of the test problem is greatly reduced when we adopt the domain adjusting strategy.
 
@@
 We present a series of numerical simulations to verify and validate our model in this section.
-Plume-SPH is first verified by a JPUE simulation.
+Plume-SPH is first verified by a JPUE (jet or plume that is ejected from a nozzle into a uniform environment) simulation.
 Velocity and mass fraction distribution both along the central axis and cross transverse are compared with experimental results.
@@
 Both global variables and local variable are comparable with simulation results from existing 3D plume models.
-
 \subsection{Simulation of JPUE}
@@
 The results are compared with experiments \citep {george1977turbulence, papanicolaou1988investigations} for validation purposes.
-Experimental data of concentration and velocity distribution across the cross-section were fit into a Gaussian profile: $\dfrac{\varphi}{\varphi_c}=exp \left[-coef\left( \left(\dfrac{r}{z}\right)^2\right)\right]$ (solid line) by \citet{papanicolaou1988investigations} even though the ctual profile are slightly different from the Gaussian profile.
+Experimental data of concentration and velocity distribution across the cross-section were fit into a Gaussian profile: $\dfrac{\varphi}{\varphi_c}=exp \left[-coef\left( \left(\dfrac{r}{z}\right)^2\right)\right]$ (solid line) by \citet{papanicolaou1988investigations} even though the actual profile are slightly different from the Gaussian profile.
 \citet{papanicolaou1988investigations} also fit concentration distribution and jet width based on velocity along central line into a straight line : $\dfrac{\varphi_0}{\varphi_c}=slope \left(\dfrac{z}{D} + intercept \right)$ (dash line).
@@
 For example, the height predicted by PDAC is 42500 m, by SK-3D is 39920 m, by ATHAM is 33392 m, by AHSEE is 36700 m.
-As for local variables, the profiles of integrated temperature, density, mass fraction of entrained air, gas mass fraction, mass fraction of solid and the radius of the plume as a function of height are compared with existing 1D and 3D models in Fig.
+As for local variables, the profiles of integrated temperature, density, mass fraction of entrained air, gas mass fraction, mass fraction of solid and the radius of the plume as a function of height are compared with existing 3D models in Fig.
 \ref{fig:strong_local_temp} $\sim$ \ref{fig:strong_local_radius}.
@@
 PDAC, which treat particles of two different sizes as two separate phases, predicted a similar mass fraction profile.
-That implies that assumption of immediate dynamic equilibrium in our model is at least valid for eruptions similar to the test case.
+That implies that assumption of dynamic equilibrium in our model is at least valid for eruptions similar to the test case.
 
@@
 Dimensionless velocity and concentration distribution across the cross-section and along the jet axis match well with experimental results of JPUE.
-Top height and integrated local variables simulated by our model are consistent with both observation and simulation results of existing 3D plume models.
+Top height and integrated local variables simulated by our model are consistent with simulation results of existing 3D plume models.
 Comparison between our results with these of SK-3D implies that turbulence model plays a significant role in plume modeling.
@@
 However, these different aspects of volcanic plumes are not independent, but are actually coupled.
-For example, it has been illustrated that gas-particle non-equilibrium would introduce a previously unrecognized jet-dragging effect, which imposes great influence on plume development, especially for weak plumes \cite{cerminara2016large}.
+For example, it has been illustrated by \cite{cerminara2016large} that gas-particle non-equilibrium would introduce a previously unrecognized jet-dragging effect, which imposes great influence on plume development, especially for weak plumes.
 In addition, there is no absolute boundary to determine which kind of hazard is dominant in certain eruptions.
@@
 \subsection{Code availability}
-The Plume-SPH code with a user manual providing instructions for installation, running and visualization are archived in github (https://github.com/Plume-SPH/plume-sph.git).
-The input data for all simulations presented in this work can also be found in the repository.
-
-\subsection{Data availability}
-Data for running the test cases, including meteorological data, properties of erupted material and air, and eruption condition data can be found in the source code repository.
+The Plume-SPH code, together with a user manual providing instructions for installation, running and visualization are archived at \url{ https://zenodo.org/record/572819#.WRCy7xiZORs} (DOI: 10.5281/zenodo.572819).
+The input data for all simulations presented in this work are archived in the same repository.
+Output of simulations presented in this paper are archived in Box.
+Access permission will be given upon request.
+
+%\subsection{Data availability}
+%Data for running the test cases, including meteorological data, properties of erupted material and air, and eruption condition data are archived in the source code repository.
 Raw data of simulation results are archived in Box.

### 2017-05-15 Manuscripts_GMD.tex
slight modification on turbulence modeling section

@@
 \begin{equation}
-\nu = \alpha \bar{h}_{ab} \bar{c}_{ab}
+\nu = \alpha \bar{h}_{ab} \bar{c}_{ab} \label{eq:def_nu_in_av}
 \end{equation}
@@
 Typically, turbulence fluctuations in a volcanic column occurs at very different length scales, ranging from millimeters up to hundreds of meters.
-The entrainment of air and heat transfer are dominated by such turbulent fluctuations as turbulent exchange coefficients are several magnitudes larger than corresponding physical coefficients (molecular viscosity and heat conduction coefficient).
+The entrainment of air and heat transfer are dominated by such turbulent fluctuations as turbulent exchange coefficients are several magnitudes larger than corresponding physical exchange coefficients (molecular viscosity and heat conduction coefficient).
 One way to include enough turbulence in the model is to use the SPH-SPS (SPH sub-particle-scale) turbulence model.
@@
 
-%Even though free surface flow are in nature turbulent, detailed research on turbulence modeling in SPH method is rather scarce.
-Among the first papers on turbulence are the implementations of the $\alpha$ turbulence model \citep{holm1999fluctuation, monaghan2002sph} and the RANS (Reynolds averaged Navier-Stokes) approach \citep{violeau2007numerical}.
-The first wall-bounded LES (large eddy simulation) simulation with SPH was performed by \citet{issa2005numerical}.
-An improved version of $SPH-\alpha$ method  \citep{monaghan2002sph} is proposed by \citet{monaghan2011turbulence}, which we refer to as $SPH-\varepsilon$ method.
-We adopt $SPH-\varepsilon$ method in our model.
-Most of these papers constrain their focus to incompressible flow.
-%$SPH-\alpha$ model \citep{monaghan2002sph} is a powerful extension of the XSPH algorithm which reduces disorder at short length scales and retains the constants of the motion.
-However, in that paper, energy conservation equation is discretized in a way that is exactly the same as usual SPH.
-There is no special treatment on energy conservation equation.
-The recent improved version \citep{monaghan2011turbulence} of $SPH-\alpha$ method focuses on incompressible flow only.
-%
-
 \subsubsection{Langrangian average in $SPH-\varepsilon$}
-\citet{monaghan2011turbulence} constructed $SPH-\varepsilon$ turbulence model within the framework of SPH in such a way that general principles such as conservation of energy, momentum and circulation are satisfied using the ideas associated with the LANS turbulence modeling.
+\citet{monaghan2011turbulence} constructed the $SPH-\varepsilon$ turbulence model within the framework of SPH in such a way that general principles such as conservation of energy, momentum and circulation are satisfied using the ideas associated with the LANS turbulence modeling.
 The basic idea of $SPH-\varepsilon$ is to determine a smoothed (averaged) velocity $\widehat{\textbf{v}}$ by a linear operation on the unsmoothed velocity $\textbf{v}$.
@@
 In $SPH-\varepsilon$ method, a XSPH \citep{monaghan1989problem} smoothing is adopted which conserves linear and angular momentum.
-In this way, solving of a system of equations is avoided and it also makes the method simple to implement and cheap for computation.
+In this way, solving of a system of equations is avoided and hence makes the method simple to implement and cheap for computation.
 The discretized form of the momentum equation is obtained through lengthy derivation.
@@
 \end{equation}
-Where $K_{ab} = l^d G_{ab}$, $M = \rho_0 l^d$ in which $d$ is the dimension and $\rho_0$ is initial density. $SPH-\varepsilon$ turbulence model is obtained after lengthy derivation:
+Where $K_{ab} = l^d G_{ab}$, $M = \rho_b l^d$ in which $d$ is the dimension and $\rho_0$ is initial density. $SPH-\varepsilon$ turbulence model is obtained after lengthy derivation:
 \begin{equation}
@@
 
-For compressible flow, the energy equation is coupled with the momentum equation and mass conservation equation.
+For compressible flow, the energy conservation equation is coupled with the momentum conservation equation and mass conservation equation.
 Averaging of thermal energy over space introduces some additional terms besides the stress term induced by velocity average.
@@
 These stress terms are usually either solved directly (for example, LANS methods) or defined via a constitutive relation (for example, large eddy simulation method).
-Less attention is typically given to the other terms.
+Less attention is typically given to other terms.
 Most commonly, a Reynolds analogy is used to model the turbulent exchange.
-Simulations of heat transfer, or other scalar transfer, in turbulent flow simply involve adding transport terms for thermal energy or species concentration, at the expense of greater storage and longer computing times but without other difficulties \citep{cebeci2013analysis}.
+Simulations of heat transfer, or other scalar transfer due to turbulence simply involve adding transport terms for thermal energy or species concentration, at the expense of greater storage and longer computing times but without other difficulties \citep{cebeci2013analysis}.
 We adopt this strategy.
@@
 \end{equation}
-Where, $\mu$ is the dynamic viscosity, $\kappa$ is the thermal conductivity. And $\mu$  can be written in term of the absolute viscosity (kinematic viscosity) as:
+Where, $\mu$ is the dynamic viscosity, $\kappa$ is the thermal conductivity. And $\mu$  can be written in term of the absolute viscosity (kinematic viscosity).
 \begin{align}
@@
 \end{equation}
-The corresponding turbulent thermal conductivity should be
-\begin{equation}
-\kappa_{t,ab}=\dfrac{\varepsilon \overline{C_{p,ab}} \overline{\rho_{ab}} \left(x_{ab}^2 + \eta^2 h_{ab}^2\right) \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 S Pr_t\textbf{v}_{ab} \cdot \textbf{x}_{ab}}
-\end{equation}
-$\overline{C_{p,ab}}$ and $\overline{\rho_{ab}}$ are simply the arithmetic means of specific heat and density. The term used to prevent singularity now can be removed. 
-\begin{equation}
-\kappa_{t,ab}=\dfrac{\varepsilon \overline{C_{p,ab}} \overline{\rho_{ab}} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 S Pr_t\textbf{v}_{ab} \cdot \textbf{x}_{ab} }
-\end{equation}
-We also need to prevent singularity, so: 
+Recall that in Eq. (\ref{eq:def_nu_in_av}), $\nu$ is scaled by $\bar{h}_{ab} \bar{c}_{ab}$. Then the corresponding turbulent thermal conductivity should be:
+\begin{equation}
+\kappa_{t,ab}=\dfrac{\varepsilon \bar{C}_{p,ab} \bar{\rho}_{ab} \left(x_{ab}^2 + \eta^2 h_{ab}^2\right) \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 S Pr_t \bar{h}_{ab} \bar{c}_{ab} \textbf{v}_{ab} \cdot \textbf{x}_{ab}}
+\end{equation}
+$\bar{C}_{p,ab}$ and $\bar{\rho}_{ab}$ are simply the arithmetic means of specific heat and density. The term used to prevent singularity can be removed. 
+\begin{equation}
+\kappa_{t,ab}=\dfrac{\varepsilon \bar{C}_{p,ab} \bar{\rho}_{ab} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 S Pr_t \bar{h}_{ab} \bar{c}_{ab} \textbf{v}_{ab} \cdot \textbf{x}_{ab} }
+\end{equation}
+We also need to prevent singularity and guarantee positive conductivity coefficient, so: 
 \begin{equation}
@@
 \begin{cases} 
-      0 & if  \quad \textbf{v}_{ab}=0 \quad or \quad \textbf{x}_{ab}=0 \\
-      \dfrac{\varepsilon \overline{C_{p,ab}} \overline{\rho_{ab}} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 S Pr_t\textbf{v}_{ab} \cdot \textbf{x}_{ab} } & \text{otherwise}
+      0 & if \quad \textbf{v}_{ab} \cdot \textbf{x}_{ab} \leq 0 \\
+      \dfrac{\varepsilon \bar{C}_{p,ab} \bar{\rho}_{ab} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 S Pr_t \bar{h}_{ab} \bar{c}_{ab} \textbf{v}_{ab} \cdot \textbf{x}_{ab} } & \text{otherwise}
 \end{cases}
@@
 & = \sum_b \dfrac{m_b}{\rho_a \rho_b} \left(\kappa_a + \kappa_b\right) \left(T_a - T_b\right) F_{ab} \left(h\right) \\
- &= 2 \sum_b \dfrac{m_b}{\rho_a \rho_b} \dfrac{\overline{C_{p,ab}} \overline{\rho_{ab}} \varepsilon x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 Pr_t  S \textbf{v}_{ab} \cdot \textbf{x}_{ab} } \left(T_a - T_b\right) F_{ab} \left(h\right)
+ &= 2 \sum_b \dfrac{m_b}{\rho_a \rho_b} \kappa_{t,ab} \left(T_a - T_b\right) F_{ab} \left(h\right)
 \end{split}
@@
 C_p \dfrac{dT}{dt}
- =\dfrac{\varepsilon}{S  Pr_t}  \sum_b \dfrac{m_b}{\rho_a \rho_b} \dfrac{\overline{C_{p,ab}} \overline{\rho_{ab}} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\textbf{v}_{ab} \cdot \textbf{x}_{ab}} \left(T_a - T_b\right) F_{ab} \left(h\right)
+ =\dfrac{\varepsilon}{S  Pr_t}  \sum_b \dfrac{m_b}{\rho_a \rho_b} \dfrac{\bar{C}_{p,ab} \bar{\rho}_{ab} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\bar{h}_{ab} \bar{c}_{ab} \textbf{v}_{ab} \cdot \textbf{x}_{ab}} \left(T_a - T_b\right) F_{ab} \left(h\right)
 \end{equation}

### 2017-08-29 Abstract.tex
abani adds

--- 2017-05-15 Manuscripts_GMD.tex
@@
 \begin{abstract}
-VATDs (Volcanic ash transport and dispersion models) used for volcanic ash forecasts often take outputs of eruption plume models as source terms.
+
+Plume-SPH provides the the first particle based simulation of volcanic plumes.
+SPH (smoothed particle hydrodynamics) has several advantages over currently used mesh based methods in modeling of multiphase free boundary flows like volcanic plumes.
+This tool will provide more accurate eruption source terms to users of VATDs (Volcanic ash transport and dispersion models) greatly improving volcanic ash forecasts.
+The accuracy of these terms is crucial for forecasts from VATDs and the 3D SPH model presented here will provide better numerical accuracy.
+As an initial effort to exploit the feasibility and advantages of SPH in volcanic plume modeling, we adopt a relatively simple physics model  (3D dusty-gas dynamic model assuming well mixed eruption materiel and dynamic and thermodynamic equilibrium between air and erupted material and minimal effect of winds) targeted at capturing the salient features of a volcanic plume.
+The documented open source code  is easily obtained and extended to incorporate other models of physics of interest to the large community of researchers investigating multiphase free boundary flows of volcanic or other origins.
+
+The Plume-SPH code also incorporates  several newly developed techniques in SPH  needed to address numerical challenges in simulating multiphase compressible turbulent flow.
+The code should thus be also of general interest to the much larger community of researchers using and developing SPH based tools.
+In particular, the $SPH-\varepsilon$ turbulence model is to capture mixing at unresolved scales, heat exchange due to turbulence is calculated by a Reynolds analogy and a corrected SPH is used to handle tensile instability and deficiency of particle distribution near the boundaries.
+We also developed methodology to impose velocity inlet and pressure outlet boundary conditions, both of which are scarce in traditional implementations of SPH.
+
+The core solver of our model is parallelized with MPI (message passing interface) obtaining good weak and strong scalability using novel techniques for data management using a SFCs (space-filling curves) and object creation time based indexing and hash table based storage scheme.
+These techniques are of  interest to  researchers engaged in developing particle in cell type methods.
+The model is verified by comparing velocity and concentration distribution along the central axis and on the transverse cross with experimental results of JPUE (jet or plume that is ejected from a nozzle into a uniform environment) and the top height of the Pinatubo eruption of 15 June 1991.
+Our results are consistent with both observations and existing 3D plume models.
+Profiles of several integrated variables are compared with those calculated in existing 3D plume models, and further verify our model.
+Analysis of the plume evolution process illustrates that this model is able to reproduce the physics of plume development.
+
+%
+%
+%VATDs (Volcanic ash transport and dispersion models) used for volcanic ash forecasts often take outputs of eruption plume models as source terms.
 The accuracy of these terms is crucial for forecasts from VATDs.
@@
 We also assume dynamic and thermodynamic equilibrium between air and erupted material and that the effect of wind is not significant.
-
-Several newly developed techniques in SPH are adapted and extended to address numerical challenges in simulating multiphase compressible turbulent flow with classical SPH method.
+%
+%Several newly developed techniques in SPH are adapted and extended to address numerical challenges in simulating multiphase compressible turbulent flow with classical SPH method.
 The $SPH-\varepsilon$ turbulence model is to capture mixing at unresolved scales.
@@
 \end{abstract}
-
-\introduction  %% \introduction[modified heading if necessary]
-%<<<<<<< HEAD
-\subsection{Volcanic ash hazards}
-Primary hazards associated with explosive volcanic eruptions include pyroclastic density currents (flows and surges), the widespread deposition of airfall tephra, and the threats to aviation posed by volcanic ash in the atmosphere.
-Simulation of all possible hazards with one model is difficult due to the fact that different length scales dominate different hazards.
-Our focus here is the hazard that volcanic ash poses to aircraft.
-
-During volcanic eruptions, VATDs (volcanic ash transport and dispersion models) are used to forecast the location and movement of ash clouds at timescales that range from hours to days.
-VATDs use eruption source parameters, such as plume height, mass eruption rate, duration, and the mass fraction distribution of erupted debris finer than about $4 \Phi$ (or $63  \mu$m), which can remain in the cloud for many hours or days.
-Observational data for such parameters are usually unavailable in the first minutes or hours after an eruption is detected.
-Moreover, these input parameters are subject to change during an eruption, requiring rapid re-assignment of new parameters.
-Usually, plume models are used to provide these source terms for VATDs and the forecast accuracy is critically dependent on these models.
-This paper reports on a new 3D (three dimensional) volcanic plume model designed to exploit the advantages of mesh-free methods for 3D modeling of such plumes that involve multiphase free boundary flows.
-
-%The dynamics of explosive eruption clouds has been a central issue of volcanology science for a long time \citep{settle1978volcanic, wilson1978control}.
-%The basic physics of volcanic plumes has been studied for more than half a century \citep{morton1956turbulent, settle1978volcanic, wilson1978control}.
-The effects of magma type and vent conditions \citep{woods1988fluid, woods1995decompression}, atmospheric conditions \citep{ woods1993moist, sparks1997volcanic, bursik2001effect}, external surface water \citep{koyaguchi1996formation}, thermal disequilibrium, and particle fallout \citep{woods1991particle} have been assessed in increasing details based on observation, experiments and numerical studies.
-
-\subsection{Existing plume models}
-Several 1D (one dimensional) volcanic plume models have been developed in the past few decades, ranging from the most basic 1D model \citep{woods1988fluid} which only accounts for mass conservation to more recently developed 1D models  \citep{bursik2001effect, mastin2007user, degruyter2012improving, woodhouse2013interaction, devenish2013using, de2015plume, folch2016fplume, pouget2016sensitivity} which tend to account for more comprehensive physics effects.
-For example, FPLUME-1.0 \citep{folch2016fplume} accounts for wind effect, entrainment of moisture, water phase change, particle fallout and re-entrainment and even wet aggregation of ash.
-However, in these 1D models, the entrainment of air is evaluated based on two coefficients: entrainment coefficient for the vertical plume and the entrainment coefficient that describes the effect of wind.
-Different 1D models adopt different entrainment coefficients based on specific formulation or calibration against well-documented case studies.
-The feedback from plume to atmosphere is usually ignored in 1D models.
-Even though determination of essential parameters such as the entrainment is not based on first principles, such simple models nevertheless allow us to investigate the importance of physical mechanisms in a volcanic plume.
-In addition, these simplified models require little computational resource and can run on standard personal computers or on web sites in very short time.
-As a result, 1D software for volcanic plume development \citep[such as][]{267, 1194, 3541}, combined with VATDs \citep[such as][]{114, draxler2015hysplit} are widely used in research and practice.
-While these 1D models can generate well-matched results with 3D (three dimensional) models for weak plumes, much greater variability is observed for strong plume scenarios, especially for local variables \citep{costa2016results}.
-In addition, there is need for greater skills in hazards forecasts especially where the plume model is used to generate source conditions for complex 2D (two dimensional) and 3D VATD models.
-
-The development of 2D and 3D, time-dependent, and multiphase numerical models for volcanic plumes has provided new explanations for many features of explosive volcanism.
-One of the earliest of these is the 3D model PDAC (Pyroclastic Dispersion Analysis Code) \citep{neri2003multiparticle}  which is a non-equilibrium, multiphase, 3D compressible flow model.
-Conservation equations for each phase are solved separately with the finite volume method.
-A parallel computing version of PDAC was also developed \citep{ongaro2007parallel}.
-Advanced numerical techniques, such as a second order scheme and semi-explicit time stepping, was also adopted afterwards to improve the accuracy of PDAC \citep{carcano2013semi}.
-
-Another 3D model, SK-3D \citep{suzuki2005numerical} is a 3D time-dependent fluid dynamics model that attempts to reproduce the entrainment process of eruption clouds with relatively simple physics but with high order numerical accuracy and high spatial resolution.
-A series of simulations based on SK-3D was reported, including establishment of the relationship between the observable quantities of the eruption clouds and the eruption conditions at the vent \citep{suzuki2009three}, investigation of the effect of the intensity of turbulence in the umbrella cloud on dispersion and sedimentation of tephra \citep{koyaguchi2009effect}, determination of the entrainment coefficients of eruption columns as a function of height \citep{suzuki2010numerical} and investigation of the effect of wind field on entrainment coefficient \citep{suzuki20133d}.
-
-While SK-3D focuses on accurately capturing the entrainment caused by turbulent mixture with higher resolution and numerical method of higher order, PDAC takes the disequilibrium between different phases into account and hence is a true multiphase model.
-Another 3D model, ATHAM (Active Tracer High-Resolution Atmospheric Model) \citep{oberhuber1998volcanic} focuses more on microphysics within the plume.
-As pyroclastic flow is not the initial concern of ATHAM, dynamic and thermodynamical equilibrium is assumed in ATHAM.
-The dynamic core of ATHAM solves the compressible Euler equations for momentum, pressure and temperature of the gas particle mixture \citep{oberhuber1998volcanic}.
-The subgrid-scale turbulence closure scheme differentiates between the horizontal and vertical directions \citep{herzog2003prognostic} captures turbulent mixing.
-The cloud microphysics predicts the mass of hydrometeors in liquid and ice phase \citep{herzog1998effect}.
-Additional modules, including gas phase chemistry \citep{trentmann2002simulation} and gas scavenging by hydrometeors \citep{textor2003injection} were added lately.
-A further extension was made to include particle aggregation \citep{textor2006volcanic1, textor2006volcanic2}.
-However, the resolution of ATHAM is still coarse compared with SK-3D and PDAC.
-
-Besides adding to their special strengths (ATHAM has been adding more and more microphysics, PDAC was extended to consider more phases), these models are also adding core strengths.
-PDAC development has begun to include the effect of microphysics into the model while ATHAM development has extended its ability to modeling pyroclastic flow.
-Both are using finer and finer resolution.
-
-Recently, a first order, nonequilibrium compressible 3D model, ASHEE \citep{cerminara2016ashee}, was introduced based on three dimensional N-phases Eulerian transportation equations, which are a full set of mass, momentum and energy transport equations for a mixture of gas and dispersed particles.
-ASHEE is valid for low concentration and low Stokes number region and much faster than N-phases Eulerian model.
-The model is based on the open source numerical solver OpenFOAM \citep{weller1998tensorial}, adapting its unstructured finite volume solver.
-
-To summarize, each 3D model has its own focus based on the problem of interest and modeling/numerical choices made.
- %Different simplification was made based on their scope. Improvements have been making since their birth. 
- Accuracy of simulation (depending on comprehensiveness of the model, resolution of discretization, numerical error, and order of accuracy) and simulation time (depending on number of governing equations, resolution, numerical methods and parallel techniques) are always a pair of conflicting considerations in 3D plume simulations.\\
-
-\subsection{Features of SPH}
-To the best of our knowledge, all of the existing 3D plume models use mesh based Eulerian methods, and there are no 3D plume models based on mesh free Lagrangian methods.
-Lagrangian methods have several features that are suitable for volcanic plume simulation that we outline below.
-Among such Lagrangian methods, smoothed particle hydrodynamics \citep{gingold1977smoothed,lucy1977numerical} based simulations have shown good agreement with experiments for many applications in fluid dynamics.
-And it currently is, by far, the most widely used mesh-free scheme.
-Our implementations of SPH in volcanology follows on earlier efforts   \citep{bursik2003smoothed,herault2010sph,haddad2016smoothed}.
-Specifically, we choose SPH as the numerical method for volcanic plume simulation to enable:
-\begin{itemize}
-\item better investigation of mixing phenomena;
-\item accurate modeling of the development of ZFE (Zone of Flow Establishment), ZEF (Zone of Established Flow) investigation and relation to column collapse and the questions relating to the development of entrainment; %and proper characterization through numerical experimentation.
-\item  easy inclusion of particles of different sizes (phases) and investigation of detailed mechanics of sedimentation and drag force interaction in lower plume.
-\end{itemize}
-These are enabled by the following features of SPH:
-\begin{itemize}
-\item For mesh based method, either interface tracking (Lagrangian) \citep{harlow1965numerical, wrobel1991computational, cheng1995simplified} or interface capturing (Eulerian) \citep{hirt1981volume, youngs1982time, gerlach2006comparison, gopala2008volume} methods are used to reconstruct the flow interface of free boundary flow.
-High computational cost, a tendency to form numerical instabilities and the inability to track complex topological changes are the significant drawbacks of tracking techniques \citep{hirt1981volume, unverdi1992front, anderson1998diffuse}.
-For interface capturing (Eulerian) method, the surrendering of surface detail before the phase transport calculation means that interface reconstruction is required between time steps to recover the interface information, which needs additional numerical effort \citep{hirt1981volume, youngs1982time}.
-Since SPH is able to adaptively adjust the discretization and automatically construct the interface, SPH does not require additional numerical effort for interface construction and therefore is more suitable for volcanic plume simulation.
-\item Advection term in the Navier-Stokes equations does not appear explicitly in discretized formula of SPH (as illustrated in Eq. (\ref{eq:gov-nc-rho}) to Eq. (\ref{eq:gov-nc-e})).
-\item It is easy to include various physics effects (like self gravity, radiative cooling and chemical reaction) into the model.
-It does not require a major overhaul and re-tooling every time new physics is introduced \citep{monaghan1995sph}.
-This implies that accounting for more physics is easier for SPH model.
-\item With more than one material, each described by its own set of particles, interface problems between phases are often trivial for SPH but difficult for mesh based schemes.
-So multiphase flow can be easily handled by SPH.
-Adding of new phases to the model also does not require a major overhaul and re-tooling.
-As will be shown in later paragraphs, adding of new phases only leads to adding of several lines into the source code for new phases and additional interaction terms between existing phases and newly added phases.
-\end{itemize}
-
-As discussed in the previous paragraph, existing 3D plume models focus on one or several specific aspects of plume and have been extended to be more comprehensive by accounting for more physics or more phases.
-Easy extensibility and capability of handling multiples phase flow without additional numerical effort greatly facilitate future extension of SPH models.
-As volcanic plumes are in nature multiphase and without pre-defined boundary in the atmosphere, SPH is a suitable numerical method for plume modeling.
-The core physics, such as entrainment of air and thermal expansion, are essential for all plume modeling while some other physics, such as water condensation and aggregation etc., are important in specific scenarios.
-As an initial effort on exploiting advantages of SPH in volcanic plume modeling, we focus on capturing basic features in plume development using a robust numerically and computationally efficient framework with support for scalable parallel computing.
-%So our current model is far away from comprehensive.
-
-Open source availability and the relatively easy extensibility of SPH will facilitate development of a more comprehensive community driven model.
-
-\subsection{Our contributions}
-Even though SPH has been known for several decades, implementations of SPH for compressible multiphase turbulent flows are few.
-Several issues endemic to classical SPH, like tensile instabilities,  compressible turbulence modeling and turbulent heat exchange, are fixed in our implementation.
-The most popular applications of SPH (and their original motivating application) has been in the simulation of free surface flow, such as breaking-waves and floods.
-Less attention was paid to velocity inlet and pressure outlet boundary conditions which are required in plume modeling.
-
-\begin{itemize}
-\item We develop methodology to impose pressure boundary conditions by adding extra layers of static ghost particles.
-Additional constraints on the time step is used to avoid the growth of numerical fluctuations near the pressure boundary.
-We impose a velocity inlet boundary condition by placing several layers of ghost particles moving with eruption velocity.
-\item Turbulence model is crucial for reproducing the entrainment of air.
-There are several turbulence models proposed for SPH method \citep{issa2005numerical,violeau2007numerical}.
-We adopt a LANS (Lagrangian Averaged Navier-Stokes) turbulence model \citep{monaghan2011turbulence}, which was originally proposed for incompressible flow, and is extended here for compressible flow accounting for turbulent heat exchange.
-\item Corrected formulation of SPH \citep{chen1999improvement} is adapted to bypass the well-known tensile instability issues of classical SPH.
-\item Simulation of volcanic plumes with acceptable accuracy requires fine resolution (very high particle counts) that cannot be accomplished without parallel computing.
-The core solver of our model is parallelized by MPI (message passing interface standard) parallelism.
-In addition, a dynamic load balancing strategy is also developed.
-\item Imposition of some types of boundary conditions (such as eruption boundary condition) requires dynamically adding and removing of particles during simulation.
-To address this issue, we adopt an efficient data management scheme based on time dependent SFC (space filling curve) induced indexing and hash table.
-The computational cost is further reduced by adjusting simulation domain adaptively.
-\end{itemize}
-
-The physical model of the plume is first presented in section \ref{sec:physics-model} and leads to a complete mathematical description of the volcanic plume (governing equations and boundary conditions).
-In section \ref{sec:SPH_method}, we  briefly introduce the numerical tool -- the SPH method.
-Both the fundamental discretization formulation and techniques that are used to handle specific issues involved in plume modeling are discussed.
-Verification and validation with numerical tests are presented in section \ref{sec:verification-validation}.
-In section \ref{sec:conclusion}, a discussion on future work is given following a brief summary.
-
-\section{Physical model} \label{sec:physics-model}
-\subsection{Description of the model}
-During an explosive eruption, a volcanic jet erupts out from a vent with a speed of several tens to more than 150 meters per second, driven by expanding gas.
-The jet is initially denser than the surrounding atmosphere and begins to decelerate through negative buoyancy and turbulent interaction with surrounding air.
-Cauliflower-like vortices are generated along jet margins, within which process, air is entrained and heated up, reducing the bulk density of the entire jet, in many cases, to less than that of the surrounding atmosphere.
-Once it becomes buoyant, such a jet develops into a plinian or subplinian plume, rising up to several kilometers to tens of kilometers until its heat is exhausted.
-Jets that lose their momentum before becoming buoyant collapse back onto the ground and transform into pyroclastic flows, surges and ignimbrites.
-During the process of plume rising up, relatively larger particles might separate from main stream of the plume, falling down onto the ground and probably be re-entrained into the plume at a lower height \citep{ernst1996sedimentation}.
-Within this process, erupted vapor condenses to liquid (droplet) and even further to ice.
-Latent heat released from phase change of erupted vapor  further heats up entrained air and further dilutes the bulk density.
-The entrained vapor might also experience a similar process and impact plume development.
-Particle aggregation processes \citep{carey1982influence,taddeucci2011aggregation}, either due to presence of liquid water, resulting from particle collision or driven by electrostatic forces might occur inside plume and thereby affect the sedimentation.
-
-All in all, the process of plume development is  essentially a multiphase turbulent mixing process coupled with heat transfer and other microphysical and chemical reactions.
-
-As an initial effort on exploiting the feasibility and advantage of SPH in plume modeling, our model is designed to describe an injection of well mixed solid and volcanic gas from a circular vent above a flat surface into a stratified stationary atmosphere following SK-3D \citep{suzuki2005numerical}.
-In this model, molecular viscosity and heat conduction is neglected since turbulent exchange coefficients are dominant.
-Erupted material consisting of solid with different size and mixture of gases is assumed to be well mixed and behave like a single phase fluid (phase 2) which is valid for eruptions with fine particles and ash.
-Air (also a mixture of different gases) is assumed to be another phase (phase 1).
-Thermodynamic equilibrium is assumed so that no separate energy equation is needed for each phase.
-As a result, there is only one energy equation for both phases (heat exchange term between different phases does not show up under this assumption).
-Dynamic equilibrium is assumed so that no separate momentum equation is needed for each phase.
-As a result, there is only one vector momentum equation for both phases (drag force term does not show up with this assumption).
-
-Because of the above assumptions, all other microphysical processes (such as the phase changes of \texorpdfstring{H\textsubscript{2}O}, aggregation, disaggregation, absorption of gas on the surface of solids, solution of gas into a liquid) and chemical processes are not considered in this model.
-These ignored microphysics factors would play critical roles under particular eruption conditions.
-Capturing of these processes needs a more comprehensive model.
-One critical element in plume development, the effect of wind, is also not yet considered in this model.
-Introducing wind effects in our model requires dynamic pressure boundary conditions, which requires more numerical effort and algorithm design.
-To summarize, our model is not valid for eruptions where wind effect plays a significant role in its development, usually refered as a weak plume.
-Our model also lacks the ability in modeling plumes with large particles or eruptions in which microphysics plays non-ignorable roles, such as a eruption of \textit{El Chich{\'o}n} volcano on April 4th 1982 \citep{sigurdsson19841982, folch2016fplume}.
-We are focussed here on developing the SPH based methodoloy in the context of the more basic (and more critical) aspects of volcanic plume and therefore devote our effort to this relative simpler model.
-It is worthwhile to mention here that because SPH is adopted as our numerical method, adding of these physics into our model would require much less work in terms of programming compared to mesh based methods.
-Since our plan is an open source distribution of the tool we believe some of these enhancements will rapidly ensue with community participation.
-
-\subsection{Governing equations}
-Based on above assumptions, the governing equations of our model are given as (which is the same as the governing equations of SK-3D \citep{suzuki2005numerical}):
-\begin{align}
-\dfrac{\partial \rho}{\partial t} + \nabla \cdot \left(\rho \textbf{v}\right) = 0 \label{eq:gov-cs-rho} \\
-\dfrac{\partial \rho \xi}{\partial t} + \nabla \cdot \left(\rho \xi \textbf{v}\right) = 0 \label{eq:gov-cs-ks}\\
-\dfrac{\partial \rho \textbf{v}}{\partial t} + \nabla \cdot \left(\rho \textbf{v} \textbf{v} + p\textbf{I}\right) = \rho \textbf{g} \label{eq:gov-cs-v} \\
-\dfrac{\partial \rho E}{\partial t} + \nabla \cdot \left[\left(\rho E + p \right)\textbf{v}\right] = \rho \textbf{g} \cdot\textbf{v} \label{eq:gov-cs-e}
-\end{align}
-Where $\rho$ is the density, $\textbf{v}$ is the velocity, $\xi$ is the mass fraction of ejected material, $\textbf{g}$ is the gravitational acceleration, $\textbf{I}$ is a unit tensor.
-$E = e + K $ is the total energy which is a summation of kinetic energy $K$ and internal energy $e$.
-An additional equation is required to close the system. In this model, the equation for closing the system is the following EOS (equation of state).
-\begin{equation}
-p = \left(\gamma_m - 1\right)\rho e \label{eq:EOS}
-\end{equation}
-Where 
-\begin{equation}
-\gamma_m = R_m/C_{vm} + 1 \label{eq:gov-gm}
-\end{equation}
-\begin{equation}
-R_m = \xi_g R_g + \xi_a R_a  \label{eq:gov-Rm}
-\end{equation}
-\begin{equation}
-C_{vm} = \xi_s C_{vs} + \xi_g C_{vg} + \xi_a C_{va} \label{eq:gov-Cvm}
-\end{equation}
-\begin{equation}
-\xi_a = 1 - \xi \label{eq:gov-na}
-\end{equation}
-\begin{equation}
-\xi_g = \xi \cdot \xi_{g0} \label{eq:gov-ng}
-\end{equation}
-\begin{equation}
-\xi_s = \xi - \xi_g \label{eq:gov-ns}
-\end{equation}
-Where, $C_v$ is the specific heat with constant volume, $R$ is the gas constant. $\xi$ with subscript is the mass fraction of corresponding constituent. The subscript 
-$m$ represents mixture of ejected material and air, $s$ represents solid portion in the ejected material, $g$ represents gas portion in the ejected material and $a$ represents air.
-$\xi_{g0}$ is the mass fraction of vapor in the erupted material.
-
-In mesh based methods, governing equations in Eulerian form, Eq.
-(\ref{eq:gov-cs-rho}) to Eq.
-(\ref{eq:gov-cs-e}), are directly discretized.
-For SPH, governing equations in Lagrange form are needed.
-By deducting kinetic energy from energy equation, subtracting mass conservation from momentum equation, combining transient term and advection term into material derivative term, the governing equations are put into the final form, in which advection term does not appear explicitly.
-\begin{align}
-\dfrac{D \rho}{D t} + \rho \nabla \cdot \textbf{v} = 0 \label{eq:gov-nc-rho}\\
-\dfrac{D \rho \xi}{D t} + \rho \xi \nabla \cdot \textbf{v} = 0 \label{eq:gov-nc-ks}\\
-\dfrac{D \textbf{v}}{D t} + \dfrac{\nabla P}{\rho} =\textbf{g} \label{eq:gov-nc-v}\\
-\dfrac{D e}{D t} + \dfrac{P \nabla \cdot \textbf{v}}{\rho} = 0 \label{eq:gov-nc-e}
-\end{align}
-%Governing equations in Lagrangian form, Eq.
-(\ref{eq:gov-nc-rho}) to Eq.(\ref{eq:gov-nc-e}), and Eulerian form, Eq.
-(\ref{eq:gov-cs-rho}) to Eq.
-(\ref{eq:gov-cs-e}), are a\dfrac{•}{•}ly equivalent, as deriving the Lagrangian form from the Eulerian is purely based on mathematics without any additional physical assumptions.
-
-\subsection{Boundary conditions}
-In the current model the initial domain is a box.
-The boundaries are categorized as the velocity inlet (a circular area at the center of the bottom of the box), wall boundary (box bottom), pressure outlet (other faces of the box).
-%probably should add a section to list the input parameters in the verification section
-
-\subsubsection{Velocity inlet}
-At the vent, temperature of erupted material $T$, eruption velocity $\textbf{v}$, the mass fraction of vapor in erupted material $\xi_{g0}$ and mass discharge rate $\dot M$ are given.
-The pressure of erupted material $p$ is assumed to be the same as ambient pressure for pressure-balanced eruption.
-The radius of vent is determined from $\rho$, $\dot M$ and $\textbf{v}$.
-Equation (\ref{eq:erupt_bc_rho}) to (\ref{eq:erupt_bc_e}) gives velocity outlet boundary condition wrote in terms of primitive variables.
-\begin{align}
-\rho =const = p/\left(R_m T\right) \label{eq:erupt_bc_rho} \\
-\xi=const=1 \label{eq:erupt_bc_xi}\\
-\textbf{v} = const =\{u,v,w\}^T \label{eq:erupt_bc_v}\\
-\dfrac{\partial e}{\partial n}=\dot M e /\left(\pi r^2\right) \label{eq:erupt_bc_e}
-\end{align} 
-Where $r$ is the radius of the vent, $n$ is the direction normal to the boundary.
-
-\subsubsection{Non-slip wall boundary}
-Velocity is zero for non-slip wall boundary.
-If we assume the boundary to be adiabatic, heat flux should be zero on the boundary.
-The flux of mass should also be zero.
-As a result, internal energy flux, which consists of heat flux and energy flux carried by mass flux, vanishes on the wall boundary.
-Equation (\ref{eq:wall_bc_rho}) to (\ref{eq:wall_bc_e}) gives no-slip wall boundary condition written in terms of primitive variables.
-\begin{align}
-\dfrac{\partial \rho}{\partial n} = const = 0\label{eq:wall_bc_rho} \\
-\dfrac{\partial \xi}{\partial n} = const = 0 \label{eq:wall_bc_xi}\\ 
-\textbf{v} = const =\{0,0,0\}^T \label{eq:wall_bc_v}\\
-\dfrac{\partial e }{\partial n} = 0\label{eq:wall_bc_e}
-\end{align}
-
-\subsubsection{Open outlet pressure boundary condition}
-The pressure of the surrounding atmosphere is given.
-Except for the pressure, boundary values for density, velocity, and energy on the outlet should depend on the solution.
-As we ignore the viscosity, the shear stress is ignored and normal stress (whose magnitude equals to pressure) balances the ambient pressure.
-\begin{equation}
-p = p_a\left(z\right)\label{eq:pressure_bc_p} 
-\end{equation}
-
-\section{SPH method} \label{sec:SPH_method}
-SPH is a mesh-free Lagragian method.
-In SPH, the domain is discretized by a set of particles or discretization points and the position of each particle is updated at every time step based on the motion computed.
-Approximation of all field variables (velocity, density and pressure, ect.) is obtained by interpolation based on discretization points.
-The physical laws (such as conservation laws of mass, momentum and energy) written in the form of PDEs (partial differential equations)  or ODEs (ordinary differential equations) need to be transformed into the Lagrangian particle formalism of SPH.
-Using a kernel function that provides the weighted estimation of the field variables at any point, the integral equations are estimated as sums over particles in a compact subdomain defined by the support of the kernel function associated with the discretization points.
-Thus, field variables associated to the particle are updated based on its neighbors.
-Each kernel function has a compact support determined by smoothing length of each particle.
-There are several review papers by \citet{monaghan1992smoothed, monaghan2005smoothed, rosswog2009astrophysical, price2012smoothed, monaghan2012smoothed}, giving a pretty comprehensive view over SPH.
-We only refer here to the representation of the constitutive equations in SPH and put more focus on specific numerical techniques for plume modeling.
-
-\subsection{Fundamental principles}
-There are several procedures for discretizing governing equations (PDEs or ODEs) with SPH.
-We present here one of them following \citet{monaghan1992smoothed, monaghan2005smoothed, monaghan2012smoothed}.
-The starting point of approximating a function with SPH is the translation property of the Dirac function.
-\begin{equation}
-A\left(\textbf{x}\right)=\int_{-\infty}^{\infty} A\left(\textbf{x} \prime\right) \delta \left(\textbf{x} \prime - \textbf{x}\right) d\textbf{x} \prime
-\label{eq:Dirac-translation}
-\end{equation}
-The Dirac function in Eq.
-(\ref{eq:Dirac-translation}) can be approximated by a weighting function $w\left(\textbf{x}-\textbf{x}\prime, h\right)$ (or $w\left(\textbf{x}\prime-\textbf{x}, h\right)$)) which tends to a Dirac function when the smoothing length $h \rightarrow 0$ :
-\begin{equation}
-\lim _{h \rightarrow 0} w\left(\textbf{x} \prime-\textbf{x}, h\right) =  \delta \left(\textbf{x} \prime - \textbf{x}\right)
-\label{eq:SPH_kernel_delta}
-\end{equation}
-An arbitrary function $A\left(\textbf{x}\right)$ can then be approximated by:
-\begin{equation}
-A\left(\textbf{x}\right) \approx <A\left(\textbf{x}\right)> = \int_{\Omega} A\left(\textbf{x} \prime\right) w\left(\textbf{x}-\textbf{x}\prime, h\right) d\textbf{x}\prime + O\left(h^2\right)
-\label{eq:SPH-fundamental-principle}
-\end{equation}
-Where $h$ is the smoothing length, determining the interaction distance.
-As the weighting function is symmetric (Eq.
-(\ref{eq:SPH-kernel-symmetric})) and satisfies the normalization condition (Eq.
-(\ref{eq:SPH-kernel-normalization-prop})), odd error terms in Eq.
-(\ref{eq:SPH-fundamental-principle}) vanishe leading to a second order approximation.
-However, in practice, second order of accuracy can not be achieved because there is no guarantee on the symmetry of particle distribution in real simulation \citep{price2012smoothed}.
-Recall that $d\textbf{x}\prime = \dfrac{dm \prime}{\rho \prime}$, the integration equation, Eq.
-(\ref{eq:SPH-fundamental-principle}), can be approximated by summation and lead to an approximation of the function $A$:
-\begin{equation}
-<A\left(\textbf{x}\right)> \approx \sum_b m_b \dfrac{A_b}{\rho_b} w\left(\textbf{x}-\textbf{x}_b, h\right)
-\label{eq:SPH-approximation-sum}
-\end{equation}
-where the summation is over all the particles within the region of compact support of the weighting function. 
-Gradient terms may be straightforwardly calculated by taking the derivative of Eq. (\ref{eq:SPH-approximation-sum}), giving
-\begin{equation}
-\begin{split}
-<\nabla A\left(\textbf{x}\right)> & = \dfrac{\partial }{\partial \textbf{x}} \int_{\Omega} A\left(\textbf{x} \prime\right) w\left(\textbf{x}-\textbf{x}\prime, h\right) d\textbf{x}\prime + O\left(h^2\right) \\
-& \approx \sum_b m_b \dfrac{A_b}{\rho_b} \nabla w\left(\textbf{x} - \textbf{x}_b, h\right)
-\end{split} 
-\label{eq:SPH-scalar-function-gradient}
-\end{equation}
-For vector quantities the expressions are similar, simply replacing $A$ with $\textbf{A}$ in Eq. (\ref{eq:SPH-approximation-sum}) and Eq. (\ref{eq:SPH-scalar-function-gradient}), giving
-\begin{align}
-<\textbf{A}\left(\textbf{x}\right)> \approx \sum_b m_b \dfrac{\textbf{A}_b}{\rho_b} w\left(\textbf{x}-\textbf{x}_b, h\right) \\
-<\nabla \cdot \textbf{A}\left(\textbf{x}\right)> \approx \sum_b m_b \dfrac{\textbf{A}_b}{\rho_b} \cdot \nabla w\left(\textbf{x} - \textbf{x}_b, h\right) \\
-<\nabla \times \textbf{A}\left(\textbf{x}\right)> \approx \sum_b m_b \dfrac{\textbf{A}_b}{\rho_b} \times \nabla w\left(\textbf{x} - \textbf{x}_b, h\right) \\
-<\nabla^j \textbf{A}^i\left(\textbf{x}\right)> \approx \sum_b m_b \dfrac{\textbf{A}_b^i}{\rho_b} \nabla^j w\left(\textbf{x} - \textbf{x}_b, h\right) 
-\label{eq:SPH-vecctor-function}
-\end{align}
-%
-
-\subsection{Weighting function}
-As described in the previous section, the weighting function (or kernel function) is used to replace the Dirac function and should satisfy Eq.
-(\ref{eq:SPH_kernel_delta}).
-So it can be viewed as an approximate form of the Dirac function and satisfies the normalization condition:
-\begin{equation}
-\int	 w\left(\textbf{x}-\textbf{x}\prime, h\right) d\textbf{x}\prime = 1
-\label{eq:SPH-kernel-normalization-prop}
-\end{equation}
-Besides normalization, the weighting function of particle $a$ has to be symmetric with respect to $a$ to ensure that neighbor particles located at the same distance away from $a$ contribute equally to SPH summation equation, see Eq.
-(\ref{eq:SPH-kernel-symmetric}).
-The weighting function also needs to satisfy conditions such as positivity and compact support.
-In addition, the kernel function must be monotonically decreasing with the distance between particles.\\
-\begin{equation}
-w\left(\textbf{x}- \textbf{x} \prime, h\right) = w\left(\textbf{x} \prime - \textbf{x}, h\right)
-\label{eq:SPH-kernel-symmetric}
-\end{equation}
-
-There is a wide variety of possible weighting functions, such as spline functions (with different orders) and Gaussian functions.
-Generally, the accuracy increases with the order of the polynomials of the kernel function, but the computational cost also increases as number of interactions increase.
-We are adopting a truncated Gaussian function as the weighting function in our simulation.
-\begin{equation}
-w\left(\textbf{x} - \textbf{x}_b\right) = 
-\begin{cases} 
-      \dfrac{1}{\left(h \sqrt{\pi}\right)^d} exp \left[- \left(\dfrac{\textbf{x} - \textbf{x} \prime}{h} \right)^2 \right] &  \vert \textbf{x} - \textbf{x} \prime \vert \leq 3h\\
-      0 & \text{Otherwise}
-\end{cases}
-\label{eq:SPH-kernel}
-\end{equation}
-Where $d$ is number of dimensions.
-The derivative of the weighting function:
-\begin{equation}
-\nabla w\left(\textbf{x} - \textbf{x}_b\right) = 
-\begin{cases} 
-      -2\left(\dfrac{\textbf{x} - \textbf{x} \prime}{h^2}\right) \dfrac{1}{\left(h \sqrt{\pi}\right)^d} exp \left[- \left(\dfrac{\textbf{x} - \textbf{x} \prime}{h}\right)^2 \right] &  \vert \textbf{x} - \textbf{x} \prime \vert \leq 3h\\
-      0 & \text{Otherwise}
-\end{cases}
-\end{equation}
-
-\subsection{Discretization of governing equations and extensibility}
-The basic interpolation given in Eq.
-(\ref{eq:SPH-approximation-sum}) to Eq.
-(\ref{eq:SPH-vecctor-function}) provides a general way to obtain SPH expressions of governing equations.
-The problem is that using these expressions ``{as is}" in general leads to quite poor gradient estimates.
-Various tricks can be used to conserve linear and angular momentum and thermal energy \citep{monaghan1992smoothed}.
-Special treatments are also needed for second order derivative terms \citep{monaghan2005smoothed}.
-We only refer here to one of these possible discretizations of compressible Euler equations with SPH:
-\begin{align}
-<\rho_a> = \sum_b m_b w_{ab} \left(h\right) \label{eq:ns-sph-d} \\
-\left\langle\dfrac{d \textbf{v}_a}{d t}\right\rangle = -\sum_b m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}\right) \nabla_a w_{a b}\left(h\right) +\textbf{g} \label{eq:ns-sph-v} \\
-\left\langle\dfrac{d e_a}{d t}\right\rangle=
- 0.5\sum_b m_b \textbf{v}_{a b}\left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}\right) \cdot \nabla_a w_{a b}\left(h\right) \label{eq:ns-sph-e}
-\end{align}
-where, $a$ is the SPH particle index, $\textbf{v}_{a b} = \textbf{v}_a - \textbf{v}_b$.
-$\Pi$ is an artificial viscosity term, which will be discussed in section \ref{sec:artificial-viscosity}.
-$w_{a b}\left(h\right)$ is a concise form of $w\left(\textbf{x}_a - \textbf{x}_b, h\right)$ and from here on, we will use this concise form.
-%However, these discretized formulations will not guarantee conservation of entropy.
-We can obtain conservation of entropy by giving up on conservation of mass or thermal energy adopting  alternative discretized formulations \citep{monaghan1992smoothed}.
-As a Lagrangian method, particle position is also updated at every time step.
-\begin{equation}
-\left\langle\dfrac{d \textbf{x}_a}{dt}\right\rangle = \textbf{v}_a \label{eq:SPH-update-pos}
-\end{equation}
-
-We highlight an important feature of the SPH methodology.
-Adding new physics and new phases into the model is trivial in terms of discretization.
-For example, adding of new source (or sink) into Eq.
-(\ref{eq:ns-sph-d}), adding a drag force into Eq.
-(\ref{eq:ns-sph-v})  and adding a heat exchange term into Eq.
-(\ref{eq:ns-sph-e}) leads to the new discretization form:
-\begin{align}
-<\rho_a> = \sum m_b w_{ab} \left(h\right) + \dot{\rho}\left(\textbf{x},t\right)\label{eq:ns-source-sph-d} \\
-\left\langle\dfrac{d \textbf{v}_a}{d t}\right\rangle= -\sum_b m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}\right) \nabla_a w_{a b}\left(h\right) +\textbf{g} + D \sum	_b m_b \dfrac{\textbf{v}_b - \textbf{v}_a}{\rho_b} \label{eq:ns-drag-sph-v} \\
-\left\langle\dfrac{d e_a}{d t}\right\rangle=
- 0.5\sum_b m_b \textbf{v}_{a b}\left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2} + \Pi_{ab}\right) \cdot \nabla_a w_{a b}\left(h\right) + \sum_b \dfrac{m_b}{\rho_b}\left(\kappa_a + \kappa_b\right) \dfrac{\left(T_a - T_b\right)}{\textbf{r}_a - \textbf{r}_b} w_{ab}\left(h\right) \label{eq:ns-conduction-sph-e}
-\end{align}
-where the source term $\dot{\rho}$ can be a "sink" of erupted vapor due to its phase change.
-%(see microphysics modelling in ATHAM \citep{oberhuber1998volcanic}).
-$D$ is a drag force coefficient.
-$\kappa$ is the heat conduction coefficient.
-$T$ is the temperature.
-Other physics can be added easily in a similar way.
-Adding of these new terms leads to modification of several lines in the source code.
-The drag force term should show up only when dynamics disequilibrium between different phases is considered.
-In that case, each phase needs one set of governing equations of Navier-Stokes type.
-Adding of new phase into SPH code only needs adding few new lines for the new phase besides interaction terms introduced by the new phase.
-
-\subsection{Artificial viscosity} \label{sec:artificial-viscosity}
-In classical SPH, shock waves are handled by introducing artificial viscosity to smear out discontinuities.
-$\Pi$, in discretized momentum equation, Eq.
-(\ref{eq:ns-sph-v}), and energy equation, Eq.
-(\ref{eq:ns-sph-e}), represents artificial viscosity term which is essentially a second order derivative term.
-As in the case of first order derivatives, second order derivatives can be estimated by differentiating a SPH interpolation twice.
-However, such a formulation has two disadvantages: First, it is very sensitive to particle disorder.
-Second, the second derivative of the kernel can change sign and lead to unphysical representations (For example, viscous dissipation causes decrease of the entropy).
-
-One of the most commonly used models of artificial viscosity is:
-\begin{equation}
-\Pi_{ab}=- \frac{\nu}{\bar{\rho}_{ab}} \dfrac{ \textbf{v}_{ab} \cdot \textbf{x}_{ab}}{\textbf{x}_{ab}^2 + \left(\eta h\right)^2}
-\label{eq:art-vis-original}
-\end{equation}
-The coefficient $\nu$ is defined as:
-\begin{equation}
-\nu = \alpha \bar{h}_{ab} \bar{c}_{ab} \label{eq:def_nu_in_av}
-\end{equation}
-Where 
-\begin{align}
-\bar{c}_{ab} = \dfrac{c_a + c_b}{2} \\
-\bar{\rho}_{ab} = \dfrac{\rho_a + \rho_b}{2} \\
-\textbf{v}_{ab}=\textbf{v}_a-\textbf{v}_b \\
-\textbf{x}_{ab}=\textbf{x}_a-\textbf{x}_b
-\end{align}
-The artificial viscosity term $\Pi_{ab}$ is a Galilean invariant and vanishes for rigid rotation.
-It produces a repulsive force between two particles when they are approaching each other and an attractive force when they are receding from each other.
-
-The SPH viscosity can be related to a continuum viscosity by converting the summation to integrals \citep{monaghan2005smoothed}.
-It has been shown that the shear viscosity coefficient $\eta = \frac{\rho \alpha h c}{8} $ and the bulk viscosity coefficient $ \zeta = \frac{5 \eta}{3}$ for two dimensional and $\eta = \frac{\rho \alpha h c}{10} $ , $ \zeta = \frac{5 \eta}{3}$ for three dimensional.
-An extra term was added to $\nu$ considering aspects of the dissipative term in shock solutions based on Riemann solvers and lead to a new formulation of artificial viscosity.
-We adopt this new formulation in our simulation:
-\begin{equation}
-\Pi_{ab}^{\beta} = 
-\begin{cases} 
-      \dfrac{- \alpha \mu_{ab} \bar{c}_{ab} + \beta \mu_{ab}^2} {\bar{\rho}_{ab}} & \textbf{v}_{ab} \cdot \textbf{x}_{ab} < 0\\
-      0 & \textbf{v}_{ab} \cdot \textbf{x}_{ab} > 0
-\end{cases}
-\label{eq:art-vis-shock}
-\end{equation}
-Where
-\begin{equation}
-\mu_{ab} = \dfrac{h \textbf{v}_{ab} \cdot \textbf{x}_{ab}}{\textbf{x}_{ab}^2 + \left(\eta h\right)^2} 
-\end{equation}
-%$\alpha$ and $\beta$ are two parameters that are free to be adjusted for each case. 
-$\alpha$ and $\beta$ are two parameters that can be adjusted for different cases.
-$\alpha = 1$ and $\beta = 2$ are  recommended by Monaghan for best results.
-In our simulation, these two parameters are calibrated to  $\alpha = 0.3$ and $\beta = 0.6$.
-$\eta$ is usually taken as 0.1 to prevent singularities.
-%when $\textbf{x}_{ab} = 0$.
-
-\subsection{Time step}
-The physical quantities (velocity, density and pressure) and particle position change every time step.
-The Courant condition, which is in spirit similar to the Courant condition for the mesh-based methods, is used to determine the time step $\Delta t$.
-\begin{equation}
-\Delta t = \textrm{CFL} \min_a \bigg \lbrace \dfrac{\left[\frac{m_a}{\rho_a}\right]^{\frac{1}{d}}}{c_a} \bigg \rbrace
-\end{equation}
-Where $c_a$ is sound speed at particle a, $d$ is number of dimensions. First order Euler integration, with $\textrm{CFL} = 0.2$, is used to advance in time.
-
-\subsection{Tensile instability and corrected derivatives}
-The classical SPH method was known to suffer from tensile instability and boundary deficiency.
-To address these difficulties, Chen \citep{chen1999improvement} proposed a corrected SPH formulation.
-For 1D case, employing a Taylor expansion for $A\left(x\right)$ about $x_a$, multiplying both sides by kernel function and then doing integration over the domain gives
-\begin{equation}
-\int_{\Omega} A\left(x\right) w\left(x- x_a, h\right) dx = 
-A_a \int_{\Omega} w\left(x - x_a, h\right) dx +A_{xa} \int_{\Omega} \left(x-x_a\right) w\left(x - x_a, h\right) dx +...
-\end{equation}
-Ignoring derivative terms higher than first order, and writing the integral in particle approximation form leads to:
-\begin{equation}
-A_a = \frac{\sum_b m_b \dfrac{A_b}{\rho_b} w\left(x_a-x_b, h\right)}{\sum_b m_b \dfrac{1}{\rho_b} w\left(x_a-x_b, h\right)}
-\label{eq:CSP-function-approximation-1d}
-\end{equation}
-Equation (\ref{eq:CSP-function-approximation-1d}) implies that $\sum_b m_b \dfrac{1}{\rho_b} w\left(x-x_b, h\right) = 1$, which can be viewed as the approximation form of Eq.
-(\ref{eq:SPH-kernel-normalization-prop}).
-The first order derivative term can be obtained in a similar way:
-\begin{equation}
-\nabla A_a = \frac{\sum_b m_b \dfrac{A_b - A_a}{\rho_b} \nabla_a w\left(x_a-x_b, h\right)}{\sum_b m_b \dfrac{x_b - x_a}{\rho_b} \nabla_a w\left(x_a-x_b, h\right)}
-\end{equation}
-
-For problems of higher dimension, the expressions for function approximation are exactly the same as Eq.
-(\ref{eq:CSP-function-approximation-1d}), even though the derivation is different.
-The first order derivative can be obtained by solving system of equations explicitly or numerically \citep{chen1999improvement}.
-
-\subsection{Multiphase SPH}
-Numerical simulation of multiphase flows is usually difficult due to the existence of complex evolving interfaces between phases.
-This presents severe challenges to conventional Eulerian grid-based numerical methods.
-As we discussed in the introduction section, special algorithms to treat and track the interface between different phases is required for mesh based methods to deal with multiphase problems.
-What's worse, the engulfment process during plume development makes geometry of interfaces very complicated.
-SPH is able to handle multiphase problems without additional numerical effort for dealing with interfaces by simply tagging particles of each phase, which makes it suitable for volcanic plume simulation.
-Take the governing equations adopted in our model as an example.
-Different phases of fluid (air and erupted material) are represented by two different sets of SPH particles (or discretization points).
-Based on assumptions we made in section \ref{sec:physics-model}, only density needs to be updated respectively for each phase.
-The updating of density is exactly the same as Eq.
-(\ref{eq:ns-sph-d}) in spirit.
-Particles of phase 1 are not be counted while evaluating density of phase 2 and vice versa.
-Updating of density is then based on the following discretized equations.
-\begin{equation}
-<\rho_{\alpha}^a>=\frac{\sum m_b w_{\alpha b} \left(h\right)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} \left(h\right) +\sum \frac{m_j}{\rho_j} w_{\alpha j} \left(h\right)} \label{eq:gov-sph-d1}
-\end{equation}
-\begin{equation}
-<\rho_\alpha^{sg}>=\frac{\sum_j m_j w_{\alpha j} \left(h\right)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} \left(h\right) +\sum \frac{m_j}{\rho_j} w_{\alpha j} \left(h\right)} \label{eq:gov-sph-d2}
-\end{equation}
-Where the subscript $a$ and $b$ represents air particles (phase 1) while $i$ and $j$ represents erupted material. $\beta$ and $\alpha$ represents either erupted material or air.
-$\rho_a^a$ is density of phase 1 (air). 
- $\rho_i^{sg}$ is density of phase 2 (erupted material).
-$\rho=\rho^a + \rho^{sg}$ is density of mixture of air and erupted material. By definition, the mass fraction is updated according to Eq. (\ref{eq:gov-sph-xi}).
-\begin{equation}
-<\xi_{\alpha}> = \dfrac{\rho^{sg}_{\alpha}}{\rho_{\alpha}}
-\label{eq:gov-sph-xi}
-\end{equation}
-
-In areas far away from the interface, updating of density is exactly the same as that for single phase flow.
-For example, in the right side and left side (or blue areas) in Fig.
-\ref{fig:SPH-multiple-density}, where there are only air particles, Eq.
-(\ref{eq:gov-sph-d2}) evaluates to zero and total density is:
-\begin{equation}
-\rho_{\alpha}=\rho_{\alpha}^a=\frac{\sum m_b w_{\alpha b} \left(h\right)}{\sum \frac{m_b}{\rho_b} w_{\alpha b}}
-\end{equation}
-Which is a special case of Eq.
-(\ref{eq:CSP-function-approximation-1d}).
-For these areas occupied by only particles of phase 2 and far away from the interface, similarly, the equation for density update becomes:
-\begin{equation}
-\rho_{\alpha}=\rho_{\alpha}^{sg}=\frac{\sum m_j w_{\alpha j} \left(h\right)}{\sum \frac{m_j}{\rho_j} w_{\alpha j}}
-\end{equation}
-That is to say, the same density updating equation can be applied for both phases and no additional numerical treatment needed to locate where the interface is.
-
-\begin{figure}
-\includegraphics[width=12cm]{Interface.png}
-\caption{In the left figure, the blue particles (phaseID=1) represent air particles the red ones (phaseID=2) represent erupted material.
-The right figure shows corresponding mass fraction.
-Mass fraction are evaluated based on Eq.
-(\ref{eq:gov-sph-d1}) and Eq.
-(\ref{eq:gov-sph-d1}) without any other interface track or capture method.}
-\label{fig:SPH-multiple-density}
-\end{figure}
-
-\subsection{Turbulence modeling with SPH}
-Typically, turbulence fluctuations in a volcanic column occurs at very different length scales, ranging from millimeters up to hundreds of meters.
-The entrainment of air and heat transfer are dominated by such turbulent fluctuations as turbulent exchange coefficients are several magnitudes larger than corresponding physical exchange coefficients (molecular viscosity and heat conduction coefficient).
-One way to include enough turbulence in the model is to use the SPH-SPS (SPH sub-particle-scale) turbulence model.
-Another way is to use fine enough resolution (namely, the direct numerical simulation method) at the expense of much higher computational cost which results from both increase in number of particles and decrease in time steps constrained by the CFL (Courant–Friedrichs–Lewy) condition.
-Here we choose the SPH-SPS method.
-Among existing SPH-SPS turbulence models \citep{holm1999fluctuation, monaghan2002sph, violeau2007numerical, monaghan2011turbulence}, we adopt a LANS type turbulence model, the $SPH-\varepsilon$ turbulence model \citep{monaghan2011turbulence}.
-However, the $SPH-\varepsilon$ turbulence model was proposed only for incompressible flow.
-In the following section, we will extend it for compressible flow.
-It is necessary to mention that all other existing SPH-SPS turbulence models \citep{holm1999fluctuation, monaghan2002sph, violeau2007numerical} also only focus on incompressible flow.
-
-\subsubsection{Langrangian average in $SPH-\varepsilon$}
-\citet{monaghan2011turbulence} constructed the $SPH-\varepsilon$ turbulence model within the framework of SPH in such a way that general principles such as conservation of energy, momentum and circulation are satisfied using the ideas associated with the LANS turbulence modeling.
-The basic idea of $SPH-\varepsilon$ is to determine a smoothed (averaged) velocity $\widehat{\textbf{v}}$ by a linear operation on the unsmoothed velocity $\textbf{v}$.
-The SPH particles move with this smoothed velocity and hence the average motion of the fluid is determined by the averaged velocity $\widehat{\textbf{v}}$:
-\begin{equation}
-\dfrac{d \textbf{x}_a}{dt} = \widehat{\textbf{v}}_a \label{eq:gov-update-pos-turbulence}
-\end{equation}
-
-Average of physical quantities over space introduces extra terms into the governing equations. Once the form of the smoothing (average) is chosen these extra terms are determined.
-The typical LANS model uses a smoothed velocity $\widehat{\textbf{v}}$ 
-defined in terms of the unsmoothed velocity $\textbf{v}$ by:
-\begin{equation}
-\widehat{\textbf{v}}\left(\textbf{x}\right)=\int \textbf{v}\left(\textbf{x} \prime\right)G\left(\vert \textbf{x} \prime - \textbf{x} \vert, l\right) d\textbf{x} \prime
-\end{equation}
-Where $G$ satisfies:
-\begin{equation}
-\int G\left(\vert \textbf{x} \prime - \textbf{x} \vert, l\right) d\textbf{x} \prime =1
-\end{equation}
-and is a member of a sequence of functions which tends to the $\delta$ function in the limit when $ l\rightarrow 0$. A typical example is Gaussian.
-The length scale $l$ determines the characteristic width of the kernel and the distance over which the velocity is smoothed.
-
-It is a common practice in LANS to use a differential equation for the smoothing rather than the integral form and finally reach to a system of equations that need to be solved implicitly.
-In $SPH-\varepsilon$ method, a XSPH \citep{monaghan1989problem} smoothing is adopted which conserves linear and angular momentum.
-In this way, solving of a system of equations is avoided and hence makes the method simple to implement and cheap for computation.
-The discretized form of the momentum equation is obtained through lengthy derivation.
-Derivation and other discussions are available in the literature -- see for e.g.
-\citep{monaghan2011turbulence}.
-Here we provide a brief summary of key steps.
-
-The smoothing adopted by \citet{monaghan2011turbulence}  is:
-\begin{equation}
-\widehat{\textbf{v}}\left(\textbf{x}\right)=\textbf{v}\left(\textbf{x}\right)+ \epsilon \int \left(\textbf{v}\left(\textbf{x} \prime\right)-\textbf{v}\left(\textbf{x}\right)\right)G\left(\vert \textbf{x} \prime - \textbf{x} \vert, l\right) d\textbf{x} \prime
-\end{equation}
-As function $G$ has the same feature as kernel function $w$, SPH approximation of the integration leads to:
-\begin{equation} \label{eq:SPH-epsilon-filtering}
-\widehat{\textbf{v}}\left(\textbf{x}\right)=\textbf{v}\left(\textbf{x}\right)+\epsilon \sum_b m_b \dfrac{\left(\textbf{v}_b -\textbf{v}\right)}{\rho _b} G\left(\vert \textbf{x} _b - \textbf{x} \vert, l\right)
-\end{equation}
-By making the replacement:
-\begin{equation}
-\label{eq:replacement-in-turb-derive}
-\dfrac{G\left(\vert \textbf{x} _b - \textbf{x} _a \vert, l\right)}{\rho _b} \rightarrow \dfrac{K_{ab}}{M}
-\end{equation}
-Where $K_{ab} = l^d G_{ab}$, $M = \rho_b l^d$ in which $d$ is the dimension and $\rho_0$ is initial density. $SPH-\varepsilon$ turbulence model is obtained after lengthy derivation:
-\begin{equation}
-\label{eq:monaghan-mom-turb}
-\dfrac{d \textbf{v}_a}{dt} = -\sum_b \left[ m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}\right) \bigtriangledown_aw_{a b}\left(h\right)\right] + \sum_b m_b \dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{M} \bigtriangledown_a K_{ab}
-\end{equation}
-Notice that if $l$ is uniform: 
-%Please notice that the above equation is slightly different from the original equation in Monaghan's paper (Eq.
-(2.17)).
-The original equation is based on the assumption that density is uniform initially.
-This assumption is not general and atmosphere density is stratified in our plume simulation.
-And the slight modification will not influence following derivation if $l$ is uniform and keep constant because:
-%
-%{\bf WE NEED TO SAY SOMETHING ABOUT INCOMPRESSIBILITY HERE}
-%
-\begin{equation}
-\nabla K_{ab} = \nabla \left(l^d G_{ab}\right) = l^d \nabla G_{ab}
-\end{equation}
-The discretized momentum equation with $SPH-\varepsilon$ turbulence model can be written in terms of $G_{ab}$ instead of $K_{ab}$:
-\begin{equation}
-\label{eq:SPH-mom-epsilon-turb}
-\dfrac{d \textbf{v}_a}{dt} = -\sum_b \left[m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_a}{\rho_a^2}\right) \bigtriangledown_aw_{a b}\left(h\right)\right] + \sum_b m_b \Phi_{ab}\bigtriangledown_aG_{ab}\left(l\right)
-\end{equation}
-where 
-\begin{equation}
-\Phi_{ab}=\dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\rho_b} 
-\end{equation}
-Which is the extra stress term induced by average. We take coefficient $\varepsilon$ as 0.8 following \citet{monaghan2011turbulence}.
-
-For compressible flow, the energy conservation equation is coupled with the momentum conservation equation and mass conservation equation.
-Averaging of thermal energy over space introduces some additional terms besides the stress term induced by velocity average.
-\citep{NASACompressibleTurbulence}.
-The averaged momentum equation for compressible flow are in the same form as that for incompressible flow, all of other additional terms, besides the corresponding velocity average induced stress term, show up in the energy equation.
-Most turbulence modeling focuses on the stress terms induced by average of velocity.
-These stress terms are usually either solved directly (for example, LANS methods) or defined via a constitutive relation (for example, large eddy simulation method).
-Less attention is typically given to other terms.
-Most commonly, a Reynolds analogy is used to model the turbulent exchange.
-Simulations of heat transfer, or other scalar transfer due to turbulence simply involve adding transport terms for thermal energy or species concentration, at the expense of greater storage and longer computing times but without other difficulties \citep{cebeci2013analysis}.
-We adopt this strategy.
-%PDAC \citep{neri2003multiparticle} also adopted a Reynolds analogy in its turbulence modeling.
-The additional terms associated with molecular diffusion and turbulent transport in the energy equation are either modeled in different ways or neglected sometimes \citep{NASACompressibleTurbulence}.
-We neglect these terms in our simulation.
-
-\subsubsection{Turbulent heat transfer}
-We adopt Reynolds analogy to get the heat transfer coefficient due to turbulence.
-The Prandtl number is defined as:
-\begin{equation}
-Pr=\dfrac{C_p \mu}{\kappa}
-\end{equation}
-Where, $\mu$ is the dynamic viscosity, $\kappa$ is the thermal conductivity. And $\mu$  can be written in term of the absolute viscosity (kinematic viscosity).
-\begin{align}
-\mu=\rho \nu
-\end{align}
-Then
-\begin{equation}
-\kappa=\dfrac{C_p \mu}{Pr}
-\end{equation}
-Typical value of $Pr_t$ for air is 0.7 $\sim$ 0.9 . We take $Pr_t=0.85$ for gases as recommended \citet{kays1994turbulent} from summarizing of experimental results.
-
-\citet{monaghan2005smoothed} summarized the simulation of viscosity and heat conduction in his review on SPH.
-We will refer to his summary in our following discussion.
-The additional term in discretized momentum equation, Eq.
-(\ref{eq:SPH-mom-epsilon-turb}), is the turbulent shear stress term.
-Recall that molecular viscosity can be discretized with SPH as shown in Eq. (\ref{eq:art-vis-original}). 
-It has been shown that the discretized molecular viscosity has both bulk viscosity and shear viscosity, where shear viscosity coefficient is \citep{monaghan2005smoothed}:
-\begin{equation}
-\nu_t = S \nu
-\end{equation}
-with
-\begin{equation}
-S= 
-\begin{cases} 
-      \dfrac{1}{10} & if  \quad d=3 \\
-      \\
-     \dfrac{1}{8}  & if  \quad d=2 
-\end{cases}
-\end{equation}
-The turbulent viscosity coefficient can be inferred from that formulation if we can reformulate the turbulent shear stress term in a form which is similar to the molecular shear term.
-Reformulate the turbulent shear stress term:
-\begin{equation}
-\label{eq:turb-stress-reformulate-to-artificial-vis}
- \sum_b \dfrac{\varepsilon}{2} \dfrac{m_b}{\rho_b} \textbf{v}_{ab} \cdot \textbf{v}_{ab} \nabla_a G_{ab}\left(l_a\right)= \sum_b \dfrac{\varepsilon}{2S} m_b \dfrac{\textbf{v}_{ab}}{\rho_b} \dfrac{S \textbf{v}_{ab} \cdot \textbf{x}_{ab}}{x_{ab}^2} \dfrac{x_{ab}^2}{\textbf{x}_{ab}} \nabla_a G_{ab}\left(l_a\right)
-\end{equation}
-Then the turbulent viscosity coefficient can be inferred from Eq. (\ref{eq:turb-stress-reformulate-to-artificial-vis}).
-\begin{equation}
-\nu_t = \dfrac{\varepsilon}{2S} \dfrac{\textbf{v}_{ab} \cdot \textbf{x}_{ab}}{\rho_b}
-\end{equation}
-Please note that turbulent viscosity term has opposite sign with molecular viscosity term in discretized momentum equation and there is a minus sign in the expression of $\Pi_{ab}$, and they cancel out.
-
-However, the above equation is correct only for the 1D situation. For 2D or 3D, it is not easy to get an explicit expression. 
-We adopt an alternative way: obtaining a value for each pair of particles instead of persisting on getting an analytical expression.
-Choosing the smoothing function as the same as SPH kernel and the smoothing length scale $l$ as the same as smoothing length $h$, the ratio between turbulent shear stress and physical shear stress is:
-\begin{equation}
-\begin{split}
-\Upsilon_{ab} &= \dfrac{\dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\rho_b}}{\dfrac{S \nu}{\rho_{ab}} \frac{\textbf{v}_{ab} \cdot \textbf{x}_{ab}}{x_{ab}^2 + \eta^2 h_{ab}^2}} \\
- & = \dfrac{\varepsilon \left(x_{ab}^2 + \eta^2 h_{ab}^2\right)}{2 S \nu} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\textbf{v}_{ab} \cdot \textbf{x}_{ab}}
-\end{split}
-\end{equation}
-$\Upsilon_{ab}$ is essentially equivalent to the ratio between turbulent viscous effect of particle b on particle a and molecular viscous effect of particle b on particle a.
-Turbulent viscosity can be easily obtained by:
-\begin{equation}
-\begin{split}
-\nu_{t,ab} &= \nu \Upsilon_{ab} \\
-&= \dfrac{\varepsilon \left(x_{ab}^2 + \eta^2 h_{ab}^2\right)}{2 S} \dfrac{\textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\textbf{v}_{ab} \cdot \textbf{x}_{ab}}
-\end{split}
-\end{equation}
-Recall that in Eq. (\ref{eq:def_nu_in_av}), $\nu$ is scaled by $\bar{h}_{ab} \bar{c}_{ab}$. Then the corresponding turbulent thermal conductivity should be:
-\begin{equation}
-\kappa_{t,ab}=\dfrac{\varepsilon \bar{C}_{p,ab} \bar{\rho}_{ab} \left(x_{ab}^2 + \eta^2 h_{ab}^2\right) \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 S Pr_t \bar{h}_{ab} \bar{c}_{ab} \textbf{v}_{ab} \cdot \textbf{x}_{ab}}
-\end{equation}
-$\bar{C}_{p,ab}$ and $\bar{\rho}_{ab}$ are simply the arithmetic means of specific heat and density. The term used to prevent singularity can be removed. 
-\begin{equation}
-\kappa_{t,ab}=\dfrac{\varepsilon \bar{C}_{p,ab} \bar{\rho}_{ab} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 S Pr_t \bar{h}_{ab} \bar{c}_{ab} \textbf{v}_{ab} \cdot \textbf{x}_{ab} }
-\end{equation}
-We also need to prevent singularity and guarantee positive conductivity coefficient, so: 
-\begin{equation}
-\kappa_{t,ab}= 
-\begin{cases} 
-      0 & if \quad \textbf{v}_{ab} \cdot \textbf{x}_{ab} \leq 0 \\
-      \dfrac{\varepsilon \bar{C}_{p,ab} \bar{\rho}_{ab} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{2 S Pr_t \bar{h}_{ab} \bar{c}_{ab} \textbf{v}_{ab} \cdot \textbf{x}_{ab} } & \text{otherwise}
-\end{cases}
-\label{eq:SPH-LANS-heat-conductivity}
-\end{equation}
-
-Heat conduction equation without source term is:
-\begin{equation}
-C_p \dfrac{dT}{dt} = \dfrac{1}{\rho} \nabla \left(\kappa \nabla T\right)
-\end{equation}
-Second spatial derivative can be approximated with SPH by following \citet{monaghan2005smoothed}. 
-\begin{equation}
-C_p \dfrac{dT}{dt} = \sum_b \dfrac{m_b}{\rho_a \rho_b} \left(\kappa_a + \kappa_b\right) \left(T_a - T_b\right) F_{ab} \left(h\right)
-\end{equation}
-Where
-\begin{equation}
-F_{ab} \textbf{x}_{ab} = \nabla _a w_{ab}
-\end{equation}
-$F_{ab}$ is nonpositive, which guarantees that heat flux flows from hot to cold. 
-Plug the turbulent thermal conductivity into the heat conduction equation:
-\begin{equation}
-\begin{split}
-C_p \dfrac{dT}{dt}
-& = \sum_b \dfrac{m_b}{\rho_a \rho_b} \left(\kappa_a + \kappa_b\right) \left(T_a - T_b\right) F_{ab} \left(h\right) \\
- &= 2 \sum_b \dfrac{m_b}{\rho_a \rho_b} \kappa_{t,ab} \left(T_a - T_b\right) F_{ab} \left(h\right)
-\end{split}
-\label{eq:turb-LANS-heat-conduct1}
-\end{equation}
-Notice that the number "2" in the front of Eq.
-(\ref{eq:turb-LANS-heat-conduct1}) comes from integration approximation of second order derivative \citep {cleary1999conduction}.
-By further simplification, we get:
-\begin{equation}
-C_p \dfrac{dT}{dt}
- =\dfrac{\varepsilon}{S  Pr_t}  \sum_b \dfrac{m_b}{\rho_a \rho_b} \dfrac{\bar{C}_{p,ab} \bar{\rho}_{ab} x_{ab}^2 \textbf{v}_{ab} \cdot \textbf{v}_{ab}}{\bar{h}_{ab} \bar{c}_{ab} \textbf{v}_{ab} \cdot \textbf{x}_{ab}} \left(T_a - T_b\right) F_{ab} \left(h\right)
-\end{equation}
-
-\subsubsection{Discretized governing equations with $SPH-\varepsilon$ turbulence model}
-Plugging in the discretized turbulent stress term and turbulent heat transfer term into the momentum and energy equation, we get new discretized governing equations:
-\begin{equation}
-\left\langle\dfrac{d \textbf{v}_{\alpha}}{d t}\right\rangle= 
--\sum_b \left[ m_b \left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b}^{\beta} - \Phi_{\alpha b}\right) \bigtriangledown_{\alpha}w_{\alpha b}\left(h\right)\right]
--\sum_j \left[m_j \left(\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j}^{\beta} - \Phi_{\alpha j}\right) \bigtriangledown_{\alpha}w_{\alpha j}\left(h\right)\right]
-+\textbf{g} \label{eq:gov-sph-v}
-\end{equation}
-With 
-\begin{equation}
-\Phi_{\alpha \beta}=\dfrac{\varepsilon}{2} \dfrac{\textbf{v}_{\alpha \beta} \cdot \textbf{v}_{\alpha \beta}} {\rho_{\beta}} 
-\end{equation}
-\begin{equation}
-\begin{split}
-\left\langle\dfrac{d e_{\alpha}}{d t}\right\rangle
-& = 0.5\sum_b \left[m_b \widehat{\textbf{v}_{\alpha b}}\left(\dfrac{p_b}{\rho_b^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha b}^{\beta} - \Phi_{\alpha b}\right) \bigtriangledown_{\alpha}w_{\alpha b}\left(h\right)\right]
- + 2 \sum_b \dfrac{m_b}{\rho_{\alpha} \rho_b} \kappa_{t,\alpha b} \left(T_{\alpha} - T_b\right) F_{\alpha b} \left(h\right) \\
- & +0.5\sum_j \left[m_j \widehat{\textbf{v}_{\alpha b}}\left(\dfrac{p_j}{\rho_j^2} + \dfrac{p_{\alpha}}{\rho_{\alpha}^2} + \Pi_{\alpha j}^{\beta} - \Phi_{\alpha j}\right) \bigtriangledown_{\alpha}w_{\alpha j}\left(h\right)\right]
- + 2 \sum_j \dfrac{m_j}{\rho_{\alpha} \rho_j} \kappa_{t,\alpha j} \left(T_{\alpha} - T_j\right) F_{\alpha j} \left(h\right)
-\end{split}
-\label{eq:gov-sph-e}
-\end{equation}
-with $\kappa_{t,\alpha \beta}$ given by Eq. (\ref{eq:SPH-LANS-heat-conductivity}). 
-As the particle-scale movement of flow is based on smoothed velocity, the velocity in the energy equation should also be smoothed.
-The filtering process is done according to Eq.
-(\ref{eq:SPH-epsilon-filtering}).
-Position of particles is updated according to Eq.
-(\ref{eq:gov-update-pos-turbulence}).
-Smoothed velocity is also used while computing artificial viscosity.
-% $\Pi_{ab}$.
-
-\subsection{Boundary conditions}
-All boundary conditions are imposed by ghost particles. Subfigure $b$ of Fig. \ref{fig:bc_and_domain_decomp} shows how boundaries are deployed. 
-%Physical quantities of particles, including real particles and different types of ghost particles, are updated differently.
-\subsubsection{Wall boundary condition}
-Traditionally either ghost particles that mirror real particles across the boundary \citep {ferrari2009new} or boundary forces \citep {monaghan2009sph} have been used to impose the wall boundary conditions.
-One disadvantage of the latter  is that the boundary forces tend to corrupt the solution in the local neighborhood.
-In addition, a natural way of imposing eruption boundary condition is using eruption ghost particles.
-To impose boundary conditions in a consistent way, we adopt a modified version of the ghost particle method \citep {kumar2013parallel} for wall boundary conditions.
-Stationary wall ghost particles are deployed in the same way as real particles.
-Instead of enforcing symmetry particle by particle, a symmetric field across the boundary is explicitly enforced.
-Ghost particles are reflected into the domain and physical quantities are calculated at these reflected positions by SPH interpolations.
-It should be noted that wall ghost particles should not be counted when computing physical properties of these reflected positions.
-Assign all properties, except for velocity, at the corresponding reflected position to the ghost particle.
-The velocity of each wall ghost particles is set to have the same value but opposite direction of the interpolated velocity at its corresponding reflection.
-By this way, the no-slip wall boundary condition (Eq.
-(\ref{eq:wall_bc_v})) is imposed naturally.
-These wall ghost particles serve as neighbors in momentum and energy update.
-More implementation details about this method can be found in \citep {kumar2013parallel}.
-As these wall ghost particles are stationary, there is no mass flux on the boundary (Eq.
-(\ref{eq:wall_bc_rho}) and (\ref{eq:wall_bc_xi})).
-In addition, as temperature is also symmetric with respect to the boundary, the gradient of temperature vanishes and hence there is no internal energy flux on the wall boundary (Eq.
-(\ref{eq:wall_bc_e})).
-
-\subsubsection{Eruption boundary condition}
-A natural way of imposing eruption boundary condition is using ghost particles that move with the eruption velocity and bear the temperature of the erupted material.
-A parabolic velocity profile that represents a fully developed Hagen-Poiseuille flow is used to determine the inlet particle velocity.
-The detailed shape of the parabolic profile is determined based on an averaged eruption velocity (Eq.
-(\ref{eq:erupt_bc_v})).
-The mass of eruption ghost particles is set to a value so that evaluation of Eq.
-(\ref{eq:ns-sph-d}) can return us a density that is consistent with the value given in Eq.
-(\ref{eq:erupt_bc_rho}).
-The internal energy associated with these particles are set to a value so that Eq.
-(\ref{eq:erupt_bc_e}) is satisfied.
-The mass fraction of erupted material (Eq.
-(\ref{eq:erupt_bc_xi})) is automatically satisfied as all particles in the eruption conduit are of phase 2.
-The density, momentum, and internal energy of these eruption ghost particles are not updated before they move above ground.
-As soon as they move out from erupt conduit, these ghost particles will be shifted to real particles and their physical quantities and position will be updated based on discretized governing equations.
-New ghost particles need to be added at the bottom of the eruption conduit as these existing ghost particles move upwards.
-
-\subsubsection{Pressure boundary condition}
-Another boundary condition in our model is the pressure outlet boundary.
-For flow in a straight channel, it is possible to treat the exit the same as the entry with a prescribed velocity profile.
-For flow with more complex channel, an exit far downstream of the flow disturbance is also feasible.
-However, the natural boundary condition (Eq.
-(\ref{eq:pressure_bc_p})) is more suitable for plume simulation as the outlet is open atmosphere.
-The way we impose pressure boundary condition is: adding several layers of pressure ghost particles surrounding the real atmosphere particles.
-Pressure, density and temperature are determined based on the elevation of particles.
-Velocity is set to zero for static atmosphere.
-The physical quantities for pressure ghost particles are not updated while these for real particles are updated at every time step.
-As position of all pressure ghost particles keep constant, we essentially impose a static pressure boundary condition.
-Real particles are removed as soon as they move out the pressure boundary.
-
-As simulations progress, changes in position and physical quantities of real particles near pressure boundaries might corrupt pressure boundary condition that was established initially.
-This shortcoming is relieved by choosing a larger computational domain so that boundaries that might be corrupted locate far away from turbulent mixing area.
-In addition, to avoid enlarging fluctuations, we add another constraint on the time step:
-\begin{equation}
-\Delta t \leq CFL_p \dfrac{h}{v}
-\end{equation}
-Where $CFL_p$ is a safety coefficient which has similar function as the normal $CFL$ number.
-Too small $CFL_p$ would slow down simulation while too large $CFL_p$ would lose its ability of mitigate numerical fluctuation near the boundary.
-The proper $CFL_p$ is determined by series of simulation tests.
-
-\begin{figure}
-\includegraphics[width=18cm]{t120_bc_proc}
-\caption{A cross section view of the simulation domain in $y-z$ plane at 66 seconds.
-Subfigure $a$ shows the mass fraction.
-Subfigure $b$ shows all boundary conditions: the dark blue region are occupied by eruption ghost particles with "Ghost particle ID" of 0, the light blue area are occupied by pressure ghost particles with "Ghost particle ID" of 1, the gray area is filled with wall ghost particles with "Ghost particle ID" of 2, "Ghost particle ID" of all real particles are set to 100, they occupy the major portion of the domain in subfigure $b$.
-Subfigure $c$ shows the cross section view of domain decomposition based on SFC.
-The simulation is conducted on 12 processors, so there are 12 subdomains in total.
-The crosss section view shows a portion of them.}
-\label{fig:bc_and_domain_decomp}
-\end{figure}
-
-\subsection{Parallellism and performance}
-One disadvantage of the 3D model is that it usually takes a much longer time than 1D models to complete one simulation.
-This disadvantage further prevents doing simulation with finer resolution and accounting for more physics in one model.
-Non-intrusive uncertainty analysis, which is commonly adopted in hazard forecasting, requires finishing multiple simulations within a given time window.
-High performance computing is therefore essential.
-Among existing CPU parallel SPH schemes, most of them focus on neighbor searching algorithm and dynamic load balancing \citep[eg.
-][] {ferrari2009new, crespo2015dualsphysics}.
-Less attention has been paid to developing more flexible data management schemes for more complicated problems.
-Motivated by techniques developed for mesh based methods, we develop a complete framework for parallelizing SPH with distrubted memory parallelism (MPI) allowing flexible and efficient data access.
-
-Any implementation of SPH code requires efficient searching and updating of neighbors during simulation.
-Of the many possible choices we adopt a background grid which was proposed by \citet {monaghan1985refined} and is quite popular in parallel SPH.
-This background grid is also used for domain decomposition in SPH.
-We refer to the elements of background grid, namely squares for two dimensions and cubic for three dimensions, as buckets.
-As for the actual storage of the physical quantities associated to each particle, different strategies have been adopted in existing implementations of SPH.
-
-In DualSPHysics \citep {crespo2015dualsphysics}, the physical quantities of each particle (position, velocity, density...) are stored in arrays.
-The particles (and the arrays with particle data) are reordered following the order of the cells.
-This has two advantages: 1) access pattern is more regular and more efficient, 2) it is easy to identify the particles that belong to a cell by using a range since the first particle of each cell is known.
-But adding, deleting and especially accessing of particles are cumbersome.
-\citet {ferrari2009new} adopted linked lists using pointers so that particles can be deleted or added during the simulation.
-Storage problems caused by fixed-size arrays are thereby also eliminated.
-We define C++ classes which contain all data of particles and buckets.
-As for the management of data, we adopt hash tables to store pointers to particles and buckets, which gives us not only flexibility of deleting and adding of elements, but also quicker access compared with linked lists.
-Instead of using the "nature manner" to number particles, we adopt a space filling curve (SFC) based index to give each particle and background bucket an unique identifier -- a strategy known to preserve data locality at minimal cost.
-The SFC based numbering strategy is further extended to include time step information so that particles added at the same position but different time have different identifiers.
-
-As for domain decomposition, even though more complicated graph-based partitioning tools \citep {biswas1999experiments} might get higher quality decomposition, they requires much more effort in programming and computation.
-So we adopt an easy-programming scheme based on SFC \citep {patra1999efficient}.
-Subfigure $c$ of Fig.
-\ref{fig:bc_and_domain_decomp} shows a cross section view of domain decomposition.
-
-More details about the data structure, domain decomposition, load balancing and domain adjusting will be published separately.
-
-\begin{figure}[!t]
-\centering
-\includegraphics[scale=0.35]{strong}
-\caption{Strong scalability test result}
-\label{fig:strong_scale}
-\end{figure}
-
-\begin{figure}[!t]
-\centering
-\includegraphics[scale=0.35]{weak_scale}
-\caption{Weak scalability test result}
-\label{fig:weak_scale}
-\end{figure}
-
-\begin{figure}[!t]
-\centering
-\includegraphics[scale=0.35]{adj_vs_no}
-\caption{The effect of domain adjusting on simulation time.
-The figure on the top shows execution time without domain adjusting, the figure on the bottom shows execution time with domain adjusting.
-Different bins represent execution time up to specific physical time indicated by horizontal axis.}
-\label{fig:adj_vs_no}
-\end{figure}
-
-Performance tests have been carried out on the computational cluster of CCR (Center for Computational Research) at Buffalo.
-Intel Xeon E5645 CPUs running at 2.40GHz clock rate with 4GB memory per core on a Q-Logic Infiniband is used in these tests.
-Each node is comprised of two sockets with 6 of these cores.
-Memory and level 3 cache are shared on each node.
-The initial domain is $[-4.8km,4.8km] \times [-4.8km,4.8km] \times [0km, 6km]$.
-Almost linear speed up is observed in our strong scalability test (Fig.
-\ref{fig:strong_scale}).
-
-The weak scalability test is conducted with the same initial domain and various smoothing lengths.
-Each simulation runs for 400 time steps.
-The average number of real particles of each process keeps constant at $25900$.
-As shown in Fig.
-\ref{fig:weak_scale}, simulation times increase around $1/3$ when number of cores increase from 16 to 4208.
-For the test problem in this section, the volcanic plume finally reaches to a region of $[-30km \,\,\, 30km] \times [-30km\,\,\,30km] \times [1.5km\,\,\,40km]$ after around 400 seconds of eruption.
-When numerical simulation goes up to 90 seconds, the plume is still within a region of $[-10km\,\,\,10km] \times [-10km\,\,\,10km] \times [0km\,\,\,25km]$.
-This implies that adjusting of domain can avoid computing large number of uninfluenced air particles, especially for the beginning stage of simulation.
-A domain adjusting algorithm is designed and implemented in our code.
-Figure \ref{fig:adj_vs_no} shows that simulation time of the test problem is greatly reduced when we adopt the domain adjusting strategy.
-
-\section{Verification and validation} \label{sec:verification-validation}
-We present a series of numerical simulations to verify and validate our model in this section.
-Plume-SPH is first verified by a JPUE (jet or plume that is ejected from a nozzle into a uniform environment) simulation.
-Velocity and mass fraction distribution both along the central axis and cross transverse are compared with experimental results.
-The pattern of ambient particles entrainment is also clearly shown.
-Then a simulation of representative strong volcanic plume is conducted.
-Both global variables and local variable are comparable with simulation results from existing 3D plume models.
-\subsection{Simulation of JPUE}
-JPUE can be considered as a simplified volcanic plume.
-While the effect of stratified atmosphere and the effect of expansion due to high temperature in volcanic plume are not represented, JPUE reproduces the entrainment due to turbulent mixing which is one of the key elements in volcanic plume development.
-There exist consistently good experimental data \citep { list1982turbulent,dimotakis1983structure, papanicolaou1988investigations} that describe the JPUE flow field giving an insight into details of JPUE, such as transverse velocity and concentration profile.
-In this section, we verify that our code and the $SPH-\varepsilon$ turbulence model is able to reproduce feature of turbulent entrainment by a JPUE simulation.
-
-As many of these experiments were conducted with liquid, we replace the original equation of state (Eq.
-(\ref{eq:EOS})) with a weakly compressible Tait equation of state \citep {becker2007weakly} (see Eq.
-(\ref{eq:EOS-Tait})) to avoid solving the Poisson equation.
-\begin{equation}
-p=B\left[\left(\dfrac{\rho}{\rho_0}\right)^{\gamma}-1\right]
-\label{eq:EOS-Tait}
-\end{equation}
-with $\gamma=7$ and $B$ is evaluated by:
-\begin{equation}
-B=\dfrac{\rho_0 c^2}{\gamma}
-\end{equation}
-Where $c$ is the speed of sound in the liquid.
-The energy equation is actually decoupled from the momentum conservation equation and the mass conservation equation by using this equation of state (EOS).
-In addition, the "atmosphere" is assumed to be uniform and gravity is set to be zero.
-We set the temperature and density of ejected material the same as surrounding ambient.
-This further simplifies the scenario for the convenience of studying turbulent mixing.
-
-One overall feature of JPUE is "self-similarity" which means that the evolution of the JPUE is determined solely by the local scale of length and velocity which theoretically account for the fact that the rate of entrainment at the edge of JPUE is proportional to a characteristic velocity at each height.
-As a results, physical and numerical experiments do not necessary to have exactly the same setups and are compared on a non-dimensional basis.
-
-\begin{table}[htp]
-	\begin{centering}
-      \caption{List of eruption condition for the test cases}		
-	  \begin{tabular}{lrrr}
-	    \hline
-	    Parameters & Units  & JPUE & Plume \\
-	    \hline
-	    Vent velocity          & $m\cdot s^{-1}$  & 500               & 275 \\
-	    Vent gas mass fraction &                  & 1.0               & 0.5 \\
-	    Vent Temperature       & $K$              & 273               & 1053 \\
-	    Vent height            & $m$              & 0                 & 1500 \\
-	    Mass discharge rate    & $kg\cdot s^{-1}$ & $5.47 \times 10^7$ & $1.5 \times 10^9$\\
-	    \hline
-	  \end{tabular}
-	  \label{tab:input_parameters}
-	\end{centering}
-\end{table}
-
-\begin{figure}
-\includegraphics[width=8cm]{vel_cross}
-\caption{Dimensionless velocity distribution across the cross-section.}
-\label{fig:JPUE_cross-section_vel}
-\end{figure}
-\begin{figure}
-\includegraphics[width=8cm]{conc_cross}
-\caption{Dimensionless concentration distribution across the cross-section.}
-\label{fig:JPUE_cross-section_conc}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=8cm]{velo_along_axis}
-\caption{Dimensionless velocity distribution along centraline.}
-\label{fig:JPUE_along-axis_vel}
-\end{figure}
-\begin{figure}
-\includegraphics[width=8cm]{conc_along_axis}
-\caption{Dimensionless concetration distribution along centraline.}
-\label{fig:JPUE_along-axis_conc}
-\end{figure}
-
-A three dimensional axisymmetric JPUE which ejects from a round vent is simulated with eruption parameters listed in Table \ref{tab:input_parameters}.
-Material properties of water are used as material properties for both phases.
-The results are compared with experiments \citep {george1977turbulence, papanicolaou1988investigations} for validation purposes.
-Experimental data of concentration and velocity distribution across the cross-section were fit into a Gaussian profile: $\dfrac{\varphi}{\varphi_c}=exp \left[-coef\left( \left(\dfrac{r}{z}\right)^2\right)\right]$ (solid line) by \citet{papanicolaou1988investigations} even though the actual profile are slightly different from the Gaussian profile.
-\citet{papanicolaou1988investigations} also fit concentration distribution and jet width based on velocity along central line into a straight line : $\dfrac{\varphi_0}{\varphi_c}=slope \left(\dfrac{z}{D} + intercept \right)$ (dash line).
-$\varphi$ is either velocity or concentration, the subscript $c$ represents the centreline, subscript $0$ represents the cross-sectionally averaged exit value.
-$r$ is the distance from the centreline on any cross-section.
-$z$ is the axial distance from the origin of the jet transverse section under consideration.
-$D$ is the diameter of vent.
-The coefficient $coef$ for concentration is 80 and 50 respectively according to \citet{george1977turbulence} and \citet{papanicolaou1988investigations}.
-$coef$ for velocity is 90 and 55 respectively according to \citet{george1977turbulence} and \citet{papanicolaou1988investigations}.
-$slope$ for jet width based on velocity is 0.104 and for concentration is 0.157. $intercept$ for jet width based on velocity is 2.58 while that for the concentration is 4.35.
-
-Although both velocity and concentration are found to be well matched with experimental results, a small disparity in both velocity and concentration are observed near the boundary of the jet.
-There are several factors that would attribute to such disparity.
-Reynolds number is not reported in many experiments assuming a high enough Reynolds number.
-In addition, some details of the experiments, such as exit velocity profile and viscosity of the experimental liquid, are not reported.
-These factors prevent us from numerically reproducing these experiments in an exact way as they were conducted.
-However, the features of JPUE is correctly reproduced with our code.
-
-We also investigated the mixing due to turbulence in JPUE simulation by checking the mixture of the two phases.
-It is shown in Fig.
-\ref{fig:Turb_mixing} that the ejected material and ambient fluids are mixed through eddies at the outer shear region.
-And the inner dense core dispersed gradually due to erosion of the outer shear region.
-Hence, our confidence in the numerical correctness of our code  is greatly reinforced.
-%What need to be noticed is that the mass of particle of phase 1 (blue) is 8 time of that of phase 2 (red).
-It is a trade off between resolution and computational expense to give particle of two phases different particles mass.
-Both advanced SPH techniques like adaptive particles size and better data management and access strategies need to be exploited to address this issue.
-
-\begin{figure}
-\includegraphics[width=8cm]{JPUE_entrainment.png}
-\caption{The left figure shows particle distribution.
-Particles of phase 1 (blue) are gradually entrained and mixed with erupted particles (red) as jet flows down stream.
-The right figure shows the mass fraction of erupted material at the moment corresponding to the left figure.}
-%What need to be noticed is that the mass of particle of phase 1 (blue) is 8 time of that of phase 2 (red).
-It is a trade off between resolution and computational expense to give particle of two phases different particles mass.
-\label{fig:Turb_mixing}
-\end{figure}
-
-\subsection{Simulation of a plume}
-The development of a volcanic plume is more complicated than JPUE in several aspects.
-Besides turbulent entrainment of ambient fluids, development of volcanic plume also involves heating up of entrained air and expanding in a stratified atmosphere.
-A strong eruption column without wind is tested in this section for the purpose of further verification and validation.
-Both global variable and local variables are compared with existing models.
-
-\subsubsection{Input parameters}
-Eruption parameters, material properties and atmosphere are chosen to be the same as the strong plume no wind case in a comparison study on eruptive column models \citep{costa2016results}.
-Eruption conditions are listed in Table \ref{tab:input_parameters}.
-The density of erupted material at the vent and radius of the vent can be computed from the given parameters.
-The eruption pressure is assumed to be as the same as pressure of ambient at the vent and hence is not given in the table.
-The vertical profiles of atmospheric properties were obtained based on the reanalysis data from ECMWF (European Centre for Medium-Range Weather Forecasts) for the period corresponding to the climactic phase of the Pinatubo eruption (Philippines, 15 June 1991).
-These conditions are more typical of a tropical atmosphere (see Fig.
-1B in \citep{costa2016results}).
-%Wind speed is not used in our current model even though it is given (by one component speed from west to east and another component speed from south to north). 
-Vertical distribution of temperature, pressure and density is used to establish stratified atmosphere.
-Material properties, shown in Table \ref{tab:material_properties}, are selected based on properties of Pinatubo and Shinmoe-dake eruption.
-Other material properties not given in the table can be infered from these given parameters based on their relationships.
-
-\begin{table}[htp]
-	\begin{centering}
-      \caption{List of material properties}		
-	  \begin{tabular}{lrrr}
-	    \hline
-	    Parameters & Units  & Value \\
-	    \hline
-	    	Specific heat of gas at constant volume     & $J \cdot kg^{-1}\cdot K^{-1}$& 717     \\
-	    Specific heat of air at constant volume     & $J \cdot kg^{-1}\cdot K^{-1}$& 1340    \\
-	    	Specific heat of solid                      & $J \cdot kg^{-1}\cdot K^{-1}$& 1100    \\
-	    	Specific heat of gas at constant pressure   & $J \cdot kg^{-1}\cdot K^{-1}$& 1000    \\
-	    	Specific heat of air at constant pressure   & $J \cdot kg^{-1}\cdot K^{-1}$& 1810    \\
-	    	Density of air at vent height               & $kg \cdot m^{-2}$       & 1.104   \\
-	    Pressure at vent height                        & $Pa$              & 84363.4 \\
-	    \hline
-	  \end{tabular}
-	  \label{tab:material_properties}
-	\end{centering}
-\end{table}
-
-\subsubsection{Global and local variables}
-
-\begin{figure}
-\center
-\includegraphics[width=7cm]{Temp}
-\caption{Temperature as a function of height}
-\label{fig:strong_local_temp}
-\end{figure}
-\begin{figure}
-\center
-\includegraphics[width=7cm]{density_strong}
-\caption{Density of the strong plume without wind after reaching its top height}
-\label{fig:strong_local_density}
-\end{figure}
-
-One of the key global quantities of great interest is the altitude to which the plume rises.
-The top height predicted by our model is around 40 km which agrees with other plume models.
-For example, the height predicted by PDAC is 42500 m, by SK-3D is 39920 m, by ATHAM is 33392 m, by AHSEE is 36700 m.
-As for local variables, the profiles of integrated temperature, density, mass fraction of entrained air, gas mass fraction, mass fraction of solid and the radius of the plume as a function of height are compared with existing 3D models in Fig.
-\ref{fig:strong_local_temp} $\sim$ \ref{fig:strong_local_radius}.
-To get rid of significant fluctuations in time and space we conducted a time averaging and spatial integration of the dynamic 3D flow fields by following \citet {cerminara2016large}.
-As particles distribute in a disordered manner in the space in SPH simulation results.
-We first project simulation results (on disordered particles) onto a pre-defined grid before doing time average and spatial integration.
-The project method is the basic SPH kernel based interpolation.
-
-The profiles of local variables match well with simulation results of existing 3D models in a general sense. The basic phenomena in volcanic plume development is correctly captured by our model.
-
-\begin{figure}
-\includegraphics[width=15cm]{msfrac}
-\caption{The mass fraction of entrained air, gas, and solid as a function of height.}
-\label{fig:strong_plume_mass_fraction}
-\end{figure}
-%
-
-\begin{figure}
-\center
-\includegraphics[width=5cm]{radius_strong}
-\caption{Radius of the strong plume without wind after reaching its top height.}
-\label{fig:strong_local_radius}
-\end{figure}
-
-As the height increases, the amount of entrained air also increases.
-Around the neutral height, where the umbrella expands, the entrainment of air shows a slight decrease due to lack of air surrounding the column at that height.
-The profile for gas, which account for both air and vapor, shows a very similar tendency as that of entrained air.
-Recall that vapor condensation is not considered in our model.
-In addition, we assume that erupted material behaves like a single phase fluid.
-So the mass fraction of gas is simply a function of entrained air (Eq.
-(\ref{eq:gas-frac-express})).
-Among these 3D models, ATHAM takes vapor condensation into account and Eq.
-(\ref{eq:gas-frac-express}) does not hold for ATHAM.
-However, the profile of entrained air and profile of gas predicted by ATHAM are still very close to each other which implies that ignoring of water phase change is a valid assumption for eruptions similar to this test case (strong plume with erupted water fraction in erupted material less than 5\%).
-This observation can be explained by the fact that air occupies a larger portion of the gas and ignoring of phase change of vapor (which is only a small portion of gas) causes slight influence on plume development.
-As for mass fraction of solid, similarly, Eq.
-(\ref{eq:solid-frac-express1}) and Eq.
-(\ref{eq:solid-frac-express2}) hold for our model.
-PDAC, which treat particles of two different sizes as two separate phases, predicted a similar mass fraction profile.
-That implies that assumption of dynamic equilibrium in our model is at least valid for eruptions similar to the test case.
-
-\begin{equation}
-\xi_a + \xi_g = \xi_a + \left(1-\xi_a\right) \xi_{g0}
-\label{eq:gas-frac-express}
-\end{equation}
-\begin{equation}
-\xi_s = \left(1 - \xi_a\right) \left(1- \xi_{g0}\right)
-\label{eq:solid-frac-express1}
-\end{equation}
-\begin{equation}
-\xi_s = 1 - \left(\xi_a + \xi_g\right)
-\label{eq:solid-frac-express2}
-\end{equation}
-
-With more cool air entrained into the plume and mixing with the hot erupted material, the temperature of the plume decreases as the height increases as shown in Fig.
-\ref{fig:strong_local_temp}.
-In the meanwhile, bulk density decrease due to entrainment and expansion (Fig.
-\ref{fig:strong_local_density}).
-
-Our model adopts the same assumptions and governing equations as SK-3D.
-However, there is still a big disparity between the profiles of local variables of our model and SK-3D.
-One of the big differences between these two models is that we adopt a LANS type of turbulence model while SK-3D adopts a LES (large eddy simulation) turbulence model.
-This implies that turbulence model might play a critical role in plume simulation.
-
-%\subsubsection{Evolution of plume} 
-%
-%\begin{figure}
-%\center
-%\includegraphics[width=15cm]{strong_elevation}
-%\caption{This figure shows the dynamic evolution process of the Pinatubo with time.
-At the very beginning, it is driven by initial momentum and rises up quickly, mixing and entraining surround air.
-Around 100s, the jet stage finishes and it becomes a plume as there were enough air entrained.
-The plume keeps rising up until reaches the top height then starts spreading.}
-%\label{fig:strong_evolution}
-%\end{figure}
-%
-%\begin{figure}
-%\includegraphics[width=10cm]{t50}
-%\caption{Contour of mass fraction and velocity quiver at 50 seconds in plane x=0, the column keeps rising up and at the same time entraining air into the plume.
-The zoomed view (right figure) of the quiver plot clearly shows mechanism of air entrainment.
-The contour levels of mass fraction are 0.7, 0.5, 0.3, 0.1, $1\times 10 ^{-1}$, $1\times 10 ^{-2}$, $1\times 10 ^{-3}$, $1\times 10^{-4}$ and $1\times 10^{-5}$.}
-%\label{fig:strong_plume_t50}
-%\end{figure}
-%
-%Figure \ref{fig:strong_evolution} shows the overall evolution process of plume.
-Plume keeps rising up and entraining air into the plume until it reaches its top height(around t=300 seconds).
-Afterwards the mixtures of erupted material and air then fall back to neutral height and starts spreading radially.
-Detailed velocity quiver plot in Fig.
-\ref{fig:strong_plume_t50} shows entraining of air into the plume at the edge of column.
-The evolution of the plume further verifies our model.
-
-\conclusions  \label{sec:conclusion}%% \conclusions[modified heading if necessary]
-A new plume model was developed based on the SPH method.
-Extensions necessary for Lagrangian methodology and compressible flow were made in the formulation of the equations of motion and turbulence models.
-Advanced numerical techniques in SPH were exploited and tailored for this model.
-High performance computing was used to mitigate the tradeoff between accuracy (depends on comprehensiveness of the model, resolution, order of accuracy of numerical methods, scheme for time upgrading.) and simulation time (depends on comprehensiveness of model, resolution, order of accuracy of numerical methods, scheme for time upgrading ...
-and computational techniques).
-The correctness of the code and model was verified and validated by a series of test simulations.
-Dimensionless velocity and concentration distribution across the cross-section and along the jet axis match well with experimental results of JPUE.
-Top height and integrated local variables simulated by our model are consistent with simulation results of existing 3D plume models.
-Comparison between our results with these of SK-3D implies that turbulence model plays a significant role in plume modeling.
-
-Currently existing 3D models focus on certain aspect of the volcanic plume (PDAC on pyroclastic flow, ATHAM on microphysics, and SK-3D on entrainment with higher accuracy and higher order of accuracy) and hence, naturally, different assumptions were made in these models.
-However, these different aspects of volcanic plumes are not independent, but are actually coupled.
-For example, it has been illustrated by \cite{cerminara2016large} that gas-particle non-equilibrium would introduce a previously unrecognized jet-dragging effect, which imposes great influence on plume development, especially for weak plumes.
-In addition, there is no absolute boundary to determine which kind of hazard is dominant in certain eruptions.
-So it is necessary to simulate all associated hazards in one model.
-Actually, effort has already been put on developing more comprehensive plume models.
-For example, a large-particle module (LPM) was added to ATHAM to track the paths of rocky particles (pyroclastic or tephra) within the plume and predict where these particles fall \citep{kobs2009modeling}.
-We were also motivated by such an evolution of plume modeling to choose SPH as our numerical tool.
-Besides its ability in dealing with interfaces for multiphase flows, as mentioned in the introduction section, SPH method has good extensibility and adding new physics and phases requires much less modification of the code compared with mesh based methods.
-Last but not least, the dramatic development of computational power makes it possible to establish a comprehensive model.
-While current computational capacity may not allow us to have a fully comprehensive model, the easy-extension feature of SPH makes it convenient to keep adding new physics into the model when necessary and computationally feasible.
-
-We have presented in this paper an initial effort and results towards developing a first principle based plume model with comprehensive physics, adopting proper numerical tools and high performance computing.
-More advanced numerical techniques, such as adaptive particle size, Godunov-SPH, semi-explicit time advancing scheme and better data management strategies and algorithms are on our list to exploit in the future.
-In the near future, effect of wind field will be take into account.
-Our code will also be made available in the open source form for the community to enhance.
-%In all existing 3D plume models, numerical calculations are performed on a non-uniform (vertically and horizontally stretched) grid, with different grid resolutions and domain sizes.
-However, implementation of different resolution in SPH methods is not straight forward and requires more work.
-
-\subsection{Code availability}
-The Plume-SPH code, together with a user manual providing instructions for installation, running and visualization are archived at \url{ https://zenodo.org/record/572819#.WRCy7xiZORs} (DOI: 10.5281/zenodo.572819).
-The input data for all simulations presented in this work are archived in the same repository.
-Output of simulations presented in this paper are archived in Box.
-Access permission will be given upon request.
-
-%\subsection{Data availability}
-%Data for running the test cases, including meteorological data, properties of erupted material and air, and eruption condition data are archived in the source code repository.
-Raw data of simulation results are archived in Box.
-Access permission will be given upon request.
-Post-processed data for plotting figures is provided as a supplement along with the manuscript.
-Please note that some of these pictures are generated by visualizing the raw data with visualization tools, there is no separate data for these figures.
-%\appendix
-%\section{}    %% Appendix A
-%
-%\subsection{}     %% Appendix A1, A2, etc.
-%
-%\appendixfigures  %% needs to be added in front of appendix figures in one-column style (\documentclass[acp, manuscript]{copernicus})
-%
-%\appendixtables   %% needs to be added in front of appendix tables in one-column style (\documentclass[acp, manuscript]{copernicus})
-%\authorcontribution{TEXT}
-
-\competinginterests{Authors have the standard conflict with employers and listed sponsors}
-%\disclaimer{TEXT}
-
-\begin{acknowledgements}
-All developers of Titan-2D, especially Dinesh Kumar who developed the GSPH version of Titan-2D, are greatly appreciated as Plume-SPH is based on their code.
-Advice and data (which is not shown in this paper) given by Suzuki Yujiro gave us great help at the initial stage of model establishment and therefor are greatly appreciated.
-We appreciate Tomaso Esposti Ongaro for providing simulation data of PDAC (were also provided by Antonio Costa later together with simulation results of other plume models) which helped us in doing early verification and improvement.
-We appreciate Antonio Costa for providing simulation results of existing 3D models, which are used in verification and validation section of this paper.
-We thank Matteo Cerminara for his helping on post-processing of plume simulation results.
-Computational results reported here were performed at the Center for Computational Research at the University at Buffalo.
-This project is supported by Grants No.
-NSF 1131074 from the National Science Foundation.
-\end{acknowledgements}
-%% REFERENCES
-%% The reference list is compiled as follows:
-\bibliographystyle{copernicus}
-\bibliography{Reference.bib}
-%% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,
-%% authors experienced with BibTeX only have to include the following two lines:
-%%
-%% \bibliographystyle{copernicus}
-%% \bibliography{example.bib}
-%%
-%% URLs and DOIs can be entered in your BibTeX file as:
-%%
-%% URL = {http://www.xyz.org/~jones/idx_g.htm}
-%% DOI = {10.5194/xyz}
-%% LITERATURE CITATIONS
-%%
-%% command                        & example result
-%% \citet{jones90}|               & Jones et al. (1990)
-%% \citep{jones90}|               & (Jones et al., 1990)
-%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
-%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
-%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
-%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
-%% \citeauthor{jones90}|          & Jones et al.
-%% \citeyear{jones90}|            & 1990
-%% FIGURES
-%% When figures and tables are placed at the end of the MS (article in one-column style), please add \clearpage
-%% between bibliography and first table and/or figure as well as between each table and/or figure.
-%% ONE-COLUMN FIGURES
-%%f
-%\begin{figure}[t]
-%\includegraphics[width=8.3cm]{FILE NAME}
-%\caption{TEXT}
-%\end{figure}
-%
-%%% TWO-COLUMN FIGURES
-%
-%%f
-%\begin{figure*}[t]
-%\includegraphics[width=12cm]{FILE NAME}
-%\caption{TEXT}
-%\end{figure*}
-%
-%
-%%% TABLES
-%%%
-%%% The different columns must be seperated with a & command and should
-%%% end with \\ to identify the column brake.
-%
-%%% ONE-COLUMN TABLE
-%
-%%t
-%\begin{table}[t]
-%\caption{TEXT}
-%\begin{tabular}{column = lcr}
-%\tophline
-%
-%\middlehline
-%
-%\bottomhline
-%\end{tabular}
-%\belowtable{} % Table Footnotes
-%\end{table}
-%
-%%% TWO-COLUMN TABLE
-%
-%%t
-%\begin{table*}[t]
-%\caption{TEXT}
-%\begin{tabular}{column = lcr}
-%\tophline
-%
-%\middlehline
-%
-%\bottomhline
-%\end{tabular}
-%\belowtable{} % Table Footnotes
-%\end{table*}
-%
-%
-%%% MATHEMATICAL EXPRESSIONS
-%
-%%% All papers typeset by Copernicus Publications follow the math typesetting regulations
-%%% given by the IUPAC Green Book (IUPAC: Quantities, Units and Symbols in Physical Chemistry,
-%%% 2nd Edn., Blackwell Science, available at: http://old.iupac.org/publications/books/gbook/green_book_2ed.pdf, 1993).
-%%%
-%%% Physical quantities/variables are typeset in italic font (t for time, T for Temperature)
-%%% Indices which are not defined are typeset in italic font (x, y, z, a, b, c)
-%%% Items/objects which are defined are typeset in roman font (Car A, Car B)
-%%% Descriptions/specifications which are defined by itself are typeset in roman font (abs, rel, ref, tot, net, ice)
-%%% Abbreviations from 2 letters are typeset in roman font (RH, LAI)
-%%% Vectors are identified in bold italic font using \vec{x}
-%%% Matrices are identified in bold roman font
-%%% Multiplication signs are typeset using the LaTeX commands \times (for vector products, grids, and exponential notations) or \cdot
-%%% The character * should not be applied as mutliplication sign
-%
-%
-%%% EQUATIONS
-%
-%%% Single-row equation
-%
-%\begin{equation}
-%
-%\end{equation}
-%
-%%% Multiline equation
-%
-%\begin{align}
-%& 3 + 5 = 8\\
-%& 3 + 5 = 8\\
-%& 3 + 5 = 8
-%\end{align}
-%
-%
-%%% MATRICES
-%
-%\begin{matrix}
-%x & y & z\\
-%x & y & z\\
-%x & y & z\\
-%\end{matrix}
-%
-%
-%%% ALGORITHM
-%
-%\begin{algorithm}
-%\caption{_}
-%\label{a1}
-%\begin{algorithmic}
-%_
-%\end{algorithmic}
-%\end{algorithm}
-%
-%
-%%% CHEMICAL FORMULAS AND REACTIONS
-%
-%%% For formulas embedded in the text, please use \chem{}
-%
-%%% The reaction environment creates labels including the letter R, i.e. (R1), (R2), etc.
-%
-%\begin{reaction}
-%%% \rightarrow should be used for normal (one-way) chemical reactions
-%%% \rightleftharpoons should be used for equilibria
-%%% \leftrightarrow should be used for resonance structures
-%\end{reaction}
-%
-%
-%%% PHYSICAL UNITS
-%%%
-%%% Please use \unit{} and apply the exponential notation
 \end{document}

### 2017-08-29 reposnse_ed.tex
abani adds

\documentclass[11pt]{article}
\usepackage{times}
%\usepackage[a4paper, total={6in, 8in}]{geometry}
%\usepackage{fullpage} % Package to use full page
%\usepackage{parskip} % Package to tweak paragraph skipping
\setlength{\parskip}{3em}%
%\setlength{\parindent}{4em}%

\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{float}
\usepackage[section]{placeins}
%\usepackage{tikz} % Package for drawing
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage[numbers]{natbib}
\setlength\parindent{0pt}
\setcounter{section}{0}
\renewcommand{\baselinestretch}{1.5}

\begin{document}

\begin{itemize}

\item {\it Please correct the missing references (see ?? in the citations in the text).}

Fixed -- sorry for the oversight.

\item {\it Also, please clarify briefly the statement "Access permission will be given upon request." What license governs the distribution and use of your code? Who will be given access, and who may be denied access? }

The Plume-SPH code, together with a user manual providing instructions for installation, running and visualization are archived at \url{ https://zenodo.org/record/572819#.WRCy7xiZORs} (DOI: 10.5281/zenodo.572819). The input data for all simulations presented in this work are archived in the same repository. The MPI license governs the distribution and use of the code and associated documentation files. Permission is granted, free of charge, to any person to deal in the software without restriction.

Output of simulations presented in this paper are archived in Box. These are the ones for which ``Access permission will be given upon request."
Access will be provided to all requestors. Code will eventually migrate to the vhub.org platform also.

\item {\it More generally, please sharpen the abstract by clarifying what elements does this work bring that are important to other modelers (i.e., why should the paper be published in GMD and not elsewhere) - the software? new mathematical equations? novel computational techniques that other modelers can implement in their own code? etc.}

Modified abstract to highlight contributions of value to other modelers. Changes highlighted below.

\end{itemize}

{\bf  Plume-SPH provides the the first particle based simulation of volcanic plumes.  SPH (smoothed particle hydrodynamics) has several advantages over currently used mesh based methods in modeling of multiphase free boundary flows like volcanic plumes. }This tool will provide more accurate eruption source terms to users of VATDs (Volcanic ash transport and dispersion models) greatly improving volcanic ash forecasts.  The accuracy of these terms is crucial for forecasts from VATDs and the 3D SPH model presented here will provide better numerical accuracy. As an initial effort to exploit the feasibility and advantages of SPH in volcanic plume modeling, we adopt a relatively simple physics model  (3D dusty-gas dynamic model assuming well mixed eruption materiel and dynamic and thermodynamic equilibrium between air and erupted material and minimal effect of winds) targeted at capturing the salient features of a volcanic plume.  {\bf The documented open source code (MPI License)  is easily obtained and extended to incorporate other models of physics of interest to the large community of researchers investigating multiphase free boundary flows of volcanic or other origins. }

The Plume-SPH code also incorporates  several newly developed techniques in SPH  needed to address numerical challenges in simulating multiphase compressible turbulent flow.  {\bf The code should thus be also of general interest to the much larger community of researchers using and developing SPH based tools.} In particular, the $SPH-\varepsilon$ turbulence model is to capture mixing at unresolved scales, heat exchange due to turbulence is calculated by a Reynolds analogy and a corrected SPH is used to handle tensile instability and deficiency of particle distribution near the boundaries. We also developed methodology to impose velocity inlet and pressure outlet boundary conditions, both of which are scarce in traditional implementations of SPH.

{\bf The core solver of our model is parallelized with MPI (message passing interface) obtaining good weak and strong scalability using novel techniques for data management using a SFCs (space-filling curves) and object creation time based indexing and hash table based storage scheme. These techniques are of  interest to  researchers engaged in developing particle in cell type methods. } The model is verified by comparing velocity and concentration distribution along the central axis and on the transverse cross with experimental results of JPUE (jet or plume that is ejected from a nozzle into a uniform environment) and the top height of the Pinatubo eruption of 15 June 1991. Our results are consistent with both observations and existing 3D plume models. Profiles of several integrated variables are compared with those calculated in existing 3D plume models, and further verify our model. Analysis of the plume evolution process illustrates that this model is able to reproduce the physics of plume development.

\end{document}

### 2017-08-29 Manuscripts_GMD_ap.tex
new master tex

--- 2017-05-09 Manuscripts_GMD.tex
@@
 \Author[4]{Matthew}{Jones}
-\affil[1]{Department of Mechanical and Aerospace Engineering, University at Buffalo, SUNY, New York, United State}
-\affil[2]{Department of Geology, University at Buffalo, SUNY, New York, United State}
-\affil[3]{Department of Material Design and Innovation, University at Buffalo, SUNY, New York, United State}
+\affil[1]{Department of Mechanical and Aerospace Engineering and Computational Data Science and Engineering, University at Buffalo, SUNY, New York, USA}
+\affil[2]{Department of Geology, University at Buffalo, SUNY, New York, USA}
+\affil[3]{Department of Material Design and Innovation, University at Buffalo, SUNY, New York, USA}
 \affil[4]{Center for Computational Research,
-University at Buffalo, SUNY, New York, United State}
+University at Buffalo, SUNY, New York, USA}
 %% The [] brackets identify the author with the corresponding affiliation. 1, 2, 3, etc. should be inserted.
@@
 \begin{abstract}
-VATDs (Volcanic ash transport and dispersion models) used for volcanic ash forecasts often take outputs of eruption plume models as source terms.
-The accuracy of these terms is crucial for forecasts from VATDs.
-All existing 3D (three dimensional) plume models described in the literature use mesh based methods.
-SPH (smoothed particle hydrodynamics), as a mesh free method, has several advantages over mesh based methods in modeling of multiphase free boundary flows.
-As an initial effort to exploit the feasibility and advantages of SPH in volcanic plume modeling, we adopt a relatively simple model, a 3D dusty-gas dynamic model, targeted at capturing the salient features of a volcanic plume.
-In this model, erupted material is assumed to be well-mixed, and represented by one phase while air is another phase.
-We also assume dynamic and thermodynamic equilibrium between air and erupted material and that the effect of wind is not significant.
-
-Several newly developed techniques in SPH are adapted and extended to address numerical challenges in simulating multiphase compressible turbulent flow with classical SPH method.
-The $SPH-\varepsilon$ turbulence model is to capture mixing at unresolved scales.
-Heat exchange due to turbulence is then calculated by Reynolds analogy.
-A corrected SPH is used to handle tensile instability and deficiency of particle distribution near the boundaries.
-We also propose to impose velocity inlet and pressure outlet boundary conditions, both of which are scarce in traditional implementations of SPH.
-The core solver of our model is parallelized with MPI (message passing interface) obtaining good weak and strong scalability.
-The model is verified by comparing velocity and concentration distribution along the central axis and on the transverse cross with experimental results of JPUE (jet or plume that is ejected from a nozzle into a uniform environment).
-Next, the top height of the Pinatubo eruption of 15 June 1991, as simulated by our model, is consistent with both observations and existing 3D plume models.
+
+Plume-SPH provides the the first particle based simulation of volcanic plumes.
+SPH (smoothed particle hydrodynamics) has several advantages over currently used mesh based methods in modeling of multiphase free boundary flows like volcanic plumes.
+This tool will provide more accurate eruption source terms to users of VATDs (Volcanic ash transport and dispersion models) greatly improving volcanic ash forecasts.
+The accuracy of these terms is crucial for forecasts from VATDs and the 3D SPH model presented here will provide better numerical accuracy.
+As an initial effort to exploit the feasibility and advantages of SPH in volcanic plume modeling, we adopt a relatively simple physics model  (3D dusty-gas dynamic model assuming well mixed eruption materiel and dynamic and thermodynamic equilibrium between air and erupted material and minimal effect of winds) targeted at capturing the salient features of a volcanic plume.
+The documented open source code  is easily obtained and extended to incorporate other models of physics of interest to the large community of researchers investigating multiphase free boundary flows of volcanic or other origins.
+
+The Plume-SPH code also incorporates  several newly developed techniques in SPH  needed to address numerical challenges in simulating multiphase compressible turbulent flow.
+The code should thus be also of general interest to the much larger community of researchers using and developing SPH based tools.
+In particular, the $SPH-\varepsilon$ turbulence model is to capture mixing at unresolved scales, heat exchange due to turbulence is calculated by a Reynolds analogy and a corrected SPH is used to handle tensile instability and deficiency of particle distribution near the boundaries.
+We also developed methodology to impose velocity inlet and pressure outlet boundary conditions, both of which are scarce in traditional implementations of SPH.
+
+The core solver of our model is parallelized with MPI (message passing interface) obtaining good weak and strong scalability using novel techniques for data management using a SFCs (space-filling curves) and object creation time based indexing and hash table based storage scheme.
+These techniques are of  interest to  researchers engaged in developing particle in cell type methods.
+The model is verified by comparing velocity and concentration distribution along the central axis and on the transverse cross with experimental results of JPUE (jet or plume that is ejected from a nozzle into a uniform environment) and the top height of the Pinatubo eruption of 15 June 1991.
+Our results are consistent with both observations and existing 3D plume models.
 Profiles of several integrated variables are compared with those calculated in existing 3D plume models, and further verify our model.
-Analysis of the plume evolution process illustrates that this model is able to reproduce the basic physics of plume development.
-The comparison also implies that the turbulence model plays a significant role in 3D volcanic plume modeling.
+Analysis of the plume evolution process illustrates that this model is able to reproduce the physics of plume development.
+
 \end{abstract}

### 2017-09-08 Manuscripts_GMD_ap_ref1.tex
Response to 1st Ref added

--- 2017-08-29 Manuscripts_GMD_ap.tex
@@
-%%xx General points
-%%xx 0 Some suggestions in the text itself are shown with "xx" -- so if you just search for "xx" you will find them --> Done
-%%xx 1 Make sure tenses match throughout
-%%xx 2 use \citet for citations in-line (text), e.g., "\citet{monaghanpaper} found that SPH is great...", not "Monaghan \citep{monaghanpaper} found that SPH is great..." -->Done
-%%xx 3 Need a section on the atmospheric structure, and how atmosphere can be incorporated into model, especially NWP or radiosondes
-%%xx 4 You often seem to forget to put a space between paragraphs.
-LaTeX of course does not recognize a <return> as a paragraph delimiter, nor is it proper to end a normal paragraph with \\ or % -->Done
-%%xx 5 You are inconsistent with capitalization in section headings.  Either always use an initial capital letter, or only use an initial capital letter on the first word, proper names or acronyms
-%%xx 6 Figures are of generally poor quality -- use PDF and reproduce at larger scale.  
-%%xx 7 Difficult to understand what figure of Pinatubo plume is supposed to be telling us about the model
-
 %% Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
@@
 %% 2-column papers and discussion papers
-\documentclass[journal abbreviation, manuscript]{copernicus}
+\documentclass[gmd, manuscript]{copernicus}
 %% Journal abbreviations (please use the same for discussion papers and final revised papers)
@@
 
-Plume-SPH provides the the first particle based simulation of volcanic plumes.
+Plume-SPH provides the first particle based simulation of volcanic plumes.
 SPH (smoothed particle hydrodynamics) has several advantages over currently used mesh based methods in modeling of multiphase free boundary flows like volcanic plumes.
@@
 The accuracy of these terms is crucial for forecasts from VATDs and the 3D SPH model presented here will provide better numerical accuracy.
-As an initial effort to exploit the feasibility and advantages of SPH in volcanic plume modeling, we adopt a relatively simple physics model  (3D dusty-gas dynamic model assuming well mixed eruption materiel and dynamic and thermodynamic equilibrium between air and erupted material and minimal effect of winds) targeted at capturing the salient features of a volcanic plume.
+As an initial effort to exploit the feasibility and advantages of SPH in volcanic plume modeling, we adopt a relatively simple physics model (3D dusty-gas dynamic model assuming well mixed eruption material, dynamic equilibrium and thermodynamic equilibrium between erupted material and air that entrained into the plume and minimal effect of winds) targeted at capturing the salient features of a volcanic plume.
 The documented open source code is easily obtained and extended to incorporate other models of physics of interest to the large community of researchers investigating multiphase free boundary flows of volcanic or other origins.
@@
 These techniques are of interest to  researchers engaged in developing particle in cell type methods.
-The model is verified by comparing velocity and concentration distribution along the central axis and on the transverse cross with experimental results of JPUE (jet or plume that is ejected from a nozzle into a uniform environment) and the top height of the Pinatubo eruption of 15 June 1991.
-Our results are consistent with both observations and existing 3D plume models.
-Profiles of several integrated variables are compared with those calculated in existing 3D plume models, and further verify our model.
+The code is first verified by 1D shock tube tests, then by comparing velocity and concentration distribution along the central axis and on the transverse
+cross with experimental results of JPUE (jet or plume that is ejected from a nozzle into a uniform environment). 
+%The model is validated by comparing against simulation results from other 3D plume models.
+Profiles of several integrated variables are compared with those calculated by existing 3D plume models for an eruption with the same MER (mass eruption rate) estimated for the Pinatubo eruption of June 15 1991.
+Our results are consistent with existing 3D plume models.
 Analysis of the plume evolution process illustrates that this model is able to reproduce the physics of plume development.
@@
 During volcanic eruptions, VATDs (volcanic ash transport and dispersion models) are used to forecast the location and movement of ash clouds at timescales that range from hours to days.
-VATDs use eruption source parameters, such as plume height, mass eruption rate, duration, and the mass fraction distribution of erupted debris finer than about $4 \Phi$ (or $63  \mu$m), which can remain in the cloud for many hours or days.
+VATDs use eruption source parameters, such as plume height, mass eruption rate, duration, and the mass fraction distribution of erupted particles finer than about $4 \Phi$ (or $62.5 \mu$m), which can remain in the cloud for many hours or days.
 Observational data for such parameters are usually unavailable in the first minutes or hours after an eruption is detected.
@@
 For example, FPLUME-1.0 \citep{folch2016fplume} accounts for wind effect, entrainment of moisture, water phase change, particle fallout and re-entrainment and even wet aggregation of ash.
-However, in these 1D models, the entrainment of air is evaluated based on two coefficients: entrainment coefficient for the vertical plume and the entrainment coefficient that describes the effect of wind.
+However, in these 1D models, the entrainment of air is evaluated based on two coefficients: entrainment coefficient due to turbulence in the rising buoyant jet and the crosswind field.
 Different 1D models adopt different entrainment coefficients based on specific formulation or calibration against well-documented case studies.
@@
  %Different simplification was made based on their scope. Improvements have been making since their birth. 
- Accuracy of simulation (depending on comprehensiveness of the model, resolution of discretization, numerical error, and order of accuracy) and simulation time (depending on number of governing equations, resolution, numerical methods and parallel techniques) are always a pair of conflicting considerations in 3D plume simulations.\\
+ Accuracy of simulation (depending on comprehensiveness of the model, resolution of discretization, numerical error, and order of accuracy) and simulation time (depending on number of governing equations, resolution, numerical methods and parallel techniques) are always a pair of conflicting considerations in 3D plume simulations.
 
@@
 \item better investigation of mixing phenomena;
-\item accurate modeling of the development of ZFE (Zone of Flow Establishment), ZEF (Zone of Established Flow) investigation and relation to column collapse and the questions relating to the development of entrainment; %and proper characterization through numerical experimentation.
+\item accurate modeling of the development of ZFE (Zone of Flow Establishment), ZEF (Zone of Established Flow) investigation and relation to column collapse and the questions relating to the development of entrainment;
 \item  easy inclusion of particles of different sizes (phases) and investigation of detailed mechanics of sedimentation and drag force interaction in lower plume.
 \end{itemize}
+
 These are enabled by the following features of SPH:
 \begin{itemize}
-\item For mesh based method, either interface tracking (Lagrangian) \citep{harlow1965numerical, wrobel1991computational, cheng1995simplified} or interface capturing (Eulerian) \citep{hirt1981volume, youngs1982time, gerlach2006comparison, gopala2008volume} methods are used to reconstruct the flow interface of free boundary flow.
+%\item For mesh based method, either interface tracking (Lagrangian) \citep{harlow1965numerical, wrobel1991computational, cheng1995simplified} or interface capturing (Eulerian) \citep{hirt1981volume, youngs1982time, gerlach2006comparison, gopala2008volume} methods are used to reconstruct the flow interface of free boundary flow.
 High computational cost, a tendency to form numerical instabilities and the inability to track complex topological changes are the significant drawbacks of tracking techniques \citep{hirt1981volume, unverdi1992front, anderson1998diffuse}.
@@
 This implies that accounting for more physics is easier for SPH model.
-\item With more than one material, each described by its own set of particles, interface problems between phases are often trivial for SPH but difficult for mesh based schemes.
+\item With more than one phase, each described by its own set of particles, interface problems between phases are often trivial for SPH but difficult for mesh based schemes.
 So multiphase flow can be easily handled by SPH.
@@
 As will be shown in later paragraphs, adding of new phases only leads to adding of several lines into the source code for new phases and additional interaction terms between existing phases and newly added phases.
+\item Interface tracking is explicit in SPH through capturing of the locations of the particles.
+Less numerical effort is required for interface construction when we attempt to include the effects of mixing by resolving the detailed interface structure and dynamics of turbulence.
 \end{itemize}
@@
 Even though SPH has been known for several decades, implementations of SPH for compressible multiphase turbulent flows are few.
+\citet{colagrossi2003numerical} proposed a multiphase SPH model for numerical simulation of air entrapment in violent fluid-structure interactions.
+\citet{hu2007incompressible} and \citet{adami2010new} proposed weakly compressible and incompressible multiphase SPH solvers in their papers.
+\citet{monaghan2013simple} also proposed new formulations for weakly compressible fluid with speed of sound sufficiently large to guarantee that the relative density variations are typically $1\%$.
+\citet {chen2015sph} recently presented a new SPH model for weakly compressible multiphase flows with complex interfaces and large density differences.
+All of them focus on incompressible or weakly compressible flow.
+
 Several issues endemic to classical SPH, like tensile instabilities,  compressible turbulence modeling and turbulent heat exchange, are fixed in our implementation.
@@
 Jets that lose their momentum before becoming buoyant collapse back onto the ground and transform into pyroclastic flows, surges and ignimbrites.
-During the process of plume rising up, relatively larger particles might separate from main stream of the plume, falling down onto the ground and probably be re-entrained into the plume at a lower height \citep{ernst1996sedimentation}.
+During the process of plume rising up, relatively larger particles might separate from main stream of the plume, falling down onto the ground and possibly be re-entrained into the plume at a lower height \citep{ernst1996sedimentation}.
 Within this process, erupted vapor condenses to liquid (droplet) and even further to ice.
@@
 \end{align}
-Where $\rho$ is the density, $\textbf{v}$ is the velocity, $\xi$ is the mass fraction of ejected material, $\textbf{g}$ is the gravitational acceleration, $\textbf{I}$ is a unit tensor.
+where $\rho$ is the density, $\textbf{v}$ is the velocity, $\xi$ is the mass fraction of ejected material, $\textbf{g}$ is the gravitational acceleration, $\textbf{I}$ is a unit tensor.
 $E = e + K $ is the total energy which is a summation of kinetic energy $K$ and internal energy $e$.
@@
 \end{equation}
-Where 
+where 
 \begin{equation}
@@
 \end{equation}
-Where, $C_v$ is the specific heat with constant volume, $R$ is the gas constant. $\xi$ with subscript is the mass fraction of corresponding constituent. The subscript 
+where, $C_v$ is the specific heat with constant volume, $R$ is the gas constant. $\xi$ with subscript is the mass fraction of corresponding constituent. The subscript 
 $m$ represents mixture of ejected material and air, $s$ represents solid portion in the ejected material, $g$ represents gas portion in the ejected material and $a$ represents air.
@@
 For SPH, governing equations in Lagrange form are needed.
-By deducting kinetic energy from energy equation, subtracting mass conservation from momentum equation, combining transient term and advection term into material derivative term, the governing equations are put into the final form, in which advection term does not appear explicitly.
+By deducting kinetic energy from energy equation, subtracting mass conservation from momentum equation, combining transient term and advection term into material derivative term (For any function $A$, material derivative is defined as $\frac{D A}{Dt} = \frac{\partial A}{\partial t} + \textbf{v} \cdot \nabla A$), the governing equations are put into the final form, in which advection term does not appear explicitly.
 \begin{align}
@@
 The radius of vent is determined from $\rho$, $\dot M$ and $\textbf{v}$.
-Equation (\ref{eq:erupt_bc_rho}) to (\ref{eq:erupt_bc_e}) gives velocity outlet boundary condition wrote in terms of primitive variables.
+Equation (\ref{eq:erupt_bc_rho}) to (\ref{eq:erupt_bc_e}) gives velocity inlet boundary condition wrote in terms of primitive variables.
 \begin{align}
@@
 \end{align} 
-Where $r$ is the radius of the vent, $n$ is the direction normal to the boundary.
+where $r$ is the radius of the vent, $n$ is the direction normal to the boundary.
 
@@
 We present here one of them following \citet{monaghan1992smoothed, monaghan2005smoothed, monaghan2012smoothed}.
-The starting point of approximating a function with SPH is the translation property of the Dirac function.
+The starting point of approximating a function with SPH is the translation property of the Dirac function $\delta(\textbf{x})$, for an arbitary function $A(\textbf{x})$, the following equation holds.
 \begin{equation}
@@
 The Dirac function in Eq.
-(\ref{eq:Dirac-translation}) can be approximated by a weighting function $w\left(\textbf{x}-\textbf{x}\prime, h\right)$ (or $w\left(\textbf{x}\prime-\textbf{x}, h\right)$)) which tends to a Dirac function when the smoothing length $h \rightarrow 0$ :
+(\ref{eq:Dirac-translation}) can be approximated by a weighting function $w\left(\textbf{x}-\textbf{x}\prime, h\right)$ (or $w\left(\textbf{x}\prime-\textbf{x}, h\right)$) which tends to a Dirac function when the smoothing length $h \rightarrow 0$ :
 \begin{equation}
@@
 \end{equation}
-An arbitrary function $A\left(\textbf{x}\right)$ can then be approximated by:
+So it can be viewed as an approximate form of the Dirac function and should satisfy the normalization condition:
+\begin{equation}
+\int	 w\left(\textbf{x}-\textbf{x}\prime, h\right) d\textbf{x}\prime = 1
+\label{eq:SPH-kernel-normalization-prop}
+\end{equation}
+Besides normalization, the weighting function of particle $a$ has to be symmetric with respect to $a$ to ensure that neighbor particles located at the same distance away from $a$ contribute equally to SPH summation equation, see Eq.
+(\ref{eq:SPH-kernel-symmetric}).
+\begin{equation}
+w\left(\textbf{x}- \textbf{x} \prime, h\right) = w\left(\textbf{x} \prime - \textbf{x}, h\right)
+\label{eq:SPH-kernel-symmetric}
+\end{equation}
+The weighting function also needs to satisfy conditions such as positivity and compact support. In addition, the kernel function must be monotonically decreasing with the distance between particles.
+
+There is a wide variety of possible weighting functions that can satisfy these requirements, such as spline functions (with different orders) and Gaussian functions.
+Generally, the accuracy increases with the order of the polynomials of the kernel function, but the computational cost also increases as number of interactions increase.
+We are adopting a truncated Gaussian function as the weighting function in our simulation.
+\begin{equation}
+w\left(\textbf{x} - \textbf{x} \prime \right) = 
+\begin{cases} 
+      \dfrac{1}{\left(h \sqrt{\pi}\right)^d} exp \left[- \left(\dfrac{\textbf{x} - \textbf{x} \prime}{h} \right)^2 \right] &  \vert \textbf{x} - \textbf{x} \prime \vert \leq 3h\\
+      0 & \text{Otherwise}
+\end{cases}
+\label{eq:SPH-kernel}
+\end{equation}
+where $d$ is number of dimensions.
+The derivative of the weighting function is:
+\begin{equation}
+\nabla w\left(\textbf{x} - \textbf{x} \prime \right) = 
+\begin{cases} 
+      -2\left(\dfrac{\textbf{x} - \textbf{x} \prime}{h}\right) \dfrac{1}{\left(h \sqrt{\pi}\right)^d} exp \left[- \left(\dfrac{\textbf{x} - \textbf{x} \prime}{h}\right)^2 \right] &  \vert \textbf{x} - \textbf{x} \prime \vert \leq 3h\\
+      0 & \text{Otherwise}
+\end{cases}
+\label{eq:SPH-kernel-gradient}
+\end{equation}
+
+By replacing $\delta$ function in Eq. (\ref{eq:SPH_kernel_delta}) with the kernel function $w$, an arbitrary function $A\left(\textbf{x}\right)$ can then be approximated by:
 \begin{equation}
@@
 \end{equation}
-Where $h$ is the smoothing length, determining the interaction distance.
 As the weighting function is symmetric (Eq.
@@
 However, in practice, second order of accuracy can not be achieved because there is no guarantee on the symmetry of particle distribution in real simulation \citep{price2012smoothed}.
-Recall that $d\textbf{x}\prime = \dfrac{dm \prime}{\rho \prime}$, the integration equation, Eq.
+Recall that $d\textbf{x}\prime = \dfrac{dm (\textbf{x} \prime)}{\rho (\textbf{x} \prime)}$, the integration equation, Eq.
 (\ref{eq:SPH-fundamental-principle}), can be approximated by summation and lead to an approximation of the function $A$:
@@
 \end{equation}
-where the summation is over all the particles within the region of compact support of the weighting function. 
+where the summation is over all the particles within the region of compact support (See Eq. (\ref{eq:SPH-kernel})) of the weighting function. 
 Gradient terms may be straightforwardly calculated by taking the derivative of Eq. (\ref{eq:SPH-approximation-sum}), giving
@@
 \end{align}
-%
-
-\subsection{Weighting function}
-As described in the previous section, the weighting function (or kernel function) is used to replace the Dirac function and should satisfy Eq.
-(\ref{eq:SPH_kernel_delta}).
-So it can be viewed as an approximate form of the Dirac function and satisfies the normalization condition:
-\begin{equation}
-\int	 w\left(\textbf{x}-\textbf{x}\prime, h\right) d\textbf{x}\prime = 1
-\label{eq:SPH-kernel-normalization-prop}
-\end{equation}
-Besides normalization, the weighting function of particle $a$ has to be symmetric with respect to $a$ to ensure that neighbor particles located at the same distance away from $a$ contribute equally to SPH summation equation, see Eq.
-(\ref{eq:SPH-kernel-symmetric}).
-The weighting function also needs to satisfy conditions such as positivity and compact support.
-In addition, the kernel function must be monotonically decreasing with the distance between particles.\\
-\begin{equation}
-w\left(\textbf{x}- \textbf{x} \prime, h\right) = w\left(\textbf{x} \prime - \textbf{x}, h\right)
-\label{eq:SPH-kernel-symmetric}
-\end{equation}
-
-There is a wide variety of possible weighting functions, such as spline functions (with different orders) and Gaussian functions.
-Generally, the accuracy increases with the order of the polynomials of the kernel function, but the computational cost also increases as number of interactions increase.
-We are adopting a truncated Gaussian function as the weighting function in our simulation.
-\begin{equation}
-w\left(\textbf{x} - \textbf{x}_b\right) = 
+
+\subsection{Artificial viscosity} \label{sec:artificial-viscosity}
+In classical SPH, shock waves are handled by introducing artificial viscosity, a term that defined based on second derivatives of velocity, to smear out discontinuities.
+As in the case of first order derivatives, second order derivatives can be estimated by differentiating a SPH interpolation twice.
+However, such a formulation has two disadvantages: first, it is very sensitive to irregular distribution of particles, second, the second derivative of the kernel can change sign and lead to unphysical representations (for example, viscous dissipation causes decrease of the entropy).
+
+One of the most commonly used models of artificial viscosity \citep{monaghan1983shock} is:
+\begin{equation}
+\Pi_{ab}=- \frac{\nu}{\bar{\rho}_{ab}} \dfrac{ \textbf{v}_{ab} \cdot \textbf{x}_{ab}}{\textbf{x}_{ab}^2 + \left(\eta h\right)^2}
+\label{eq:art-vis-original}
+\end{equation}
+The coefficient $\nu$ is defined as:
+\begin{equation}
+\nu = \alpha \bar{h}_{ab} \bar{c}_{ab}
+\end{equation}
+where 
+\begin{align}
+\bar{c}_{ab} = \dfrac{c_a + c_b}{2} \\
+\bar{\rho}_{ab} = \dfrac{\rho_a + \rho_b}{2} \\
+\textbf{v}_{ab}=\textbf{v}_a-\textbf{v}_b \\
+\textbf{x}_{ab}=\textbf{x}_a-\textbf{x}_b
+\end{align}
+The artificial viscosity term $\Pi_{ab}$ is a Galilean invariant and vanishes for rigid rotation.
+It produces a repulsive force between two particles when they are approaching each other and an attractive force when they are receding from each other.
+
+The SPH viscosity can be related to a continuum viscosity by converting the summation to integrals \citep{monaghan2005smoothed}.
+It has been shown that the shear viscosity coefficient $\eta = \frac{\rho \alpha h c}{8} $ and the bulk viscosity coefficient $ \zeta = \frac{5 \eta}{3}$ for two dimensional and $\eta = \frac{\rho \alpha h c}{10} $ , $ \zeta = \frac{5 \eta}{3}$ for three dimensional.
+An extra term was added to $\nu$ considering aspects of the dissipative term in shock solutions based on Riemann solvers and lead to a new formulation of artificial viscosity.
+We adopt this new formulation in our simulation:
+\begin{equation}
+\Pi_{ab}^{\beta} = 
 \begin{cases} 
-      \dfrac{1}{\left(h \sqrt{\pi}\right)^d} exp \left[- \left(\dfrac{\textbf{x} - \textbf{x} \prime}{h} \right)^2 \right] &  \vert \textbf{x} - \textbf{x} \prime \vert \leq 3h\\
-      0 & \text{Otherwise}
+      \dfrac{- \alpha \mu_{ab} \bar{c}_{ab} + \beta \mu_{ab}^2} {\bar{\rho}_{ab}} & \textbf{v}_{ab} \cdot \textbf{x}_{ab} < 0\\
+      0 & \textbf{v}_{ab} \cdot \textbf{x}_{ab} > 0
 \end{cases}
-\label{eq:SPH-kernel}
-\end{equation}
-Where $d$ is number of dimensions.
-The derivative of the weighting function:
-\begin{equation}
-\nabla w\left(\textbf{x} - \textbf{x}_b\right) = 
-\begin{cases} 
-      -2\left(\dfrac{\textbf{x} - \textbf{x} \prime}{h^2}\right) \dfrac{1}{\left(h \sqrt{\pi}\right)^d} exp \left[- \left(\dfrac{\textbf{x} - \textbf{x} \prime}{h}\right)^2 \right] &  \vert \textbf{x} - \textbf{x} \prime \vert \leq 3h\\
-      0 & \text{Otherwise}
-\end{cases}
-\end{equation}
+\label{eq:art-vis-shock}
+\end{equation}
+where
+\begin{equation}
+\mu_{ab} = \dfrac{h \textbf{v}_{ab} \cdot \textbf{x}_{ab}}{\textbf{x}_{ab}^2 + \left(\eta h\right)^2} 
+\end{equation}
+%$\alpha$ and $\beta$ are two parameters that are free to be adjusted for each case. 
+$\alpha$ and $\beta$ are two parameters that can be adjusted for different cases.
+$\alpha = 1$ and $\beta = 2$ are  recommended by Monaghan for best results.
+In our simulation, these two parameters are calibrated to  $\alpha = 0.3$ and $\beta = 0.6$.
+$\eta$ is usually taken as 0.1 to prevent singularities.
+%when $\textbf{x}_{ab} = 0$.
 
@@
 where, $a$ is the SPH particle index, $\textbf{v}_{a b} = \textbf{v}_a - \textbf{v}_b$.
-$\Pi$ is an artificial viscosity term, which will be discussed in section \ref{sec:artificial-viscosity}.
+$\Pi$ is an artificial viscosity term, which is discussed in section \ref{sec:artificial-viscosity}.
 $w_{a b}\left(h\right)$ is a concise form of $w\left(\textbf{x}_a - \textbf{x}_b, h\right)$ and from here on, we will use this concise form.
@@
 Adding of these new terms leads to modification of several lines in the source code.
-The drag force term should show up only when dynamics disequilibrium between different phases is considered.
+The drag force term should show up when dynamics disequilibrium between different phases is considered.
 In that case, each phase needs one set of governing equations of Navier-Stokes type.
 Adding of new phase into SPH code only needs adding few new lines for the new phase besides interaction terms introduced by the new phase.
-
-\subsection{Artificial viscosity} \label{sec:artificial-viscosity}
-In classical SPH, shock waves are handled by introducing artificial viscosity to smear out discontinuities.
-$\Pi$, in discretized momentum equation, Eq.
-(\ref{eq:ns-sph-v}), and energy equation, Eq.
-(\ref{eq:ns-sph-e}), represents artificial viscosity term which is essentially a second order derivative term.
-As in the case of first order derivatives, second order derivatives can be estimated by differentiating a SPH interpolation twice.
-However, such a formulation has two disadvantages: First, it is very sensitive to particle disorder.
-Second, the second derivative of the kernel can change sign and lead to unphysical representations (For example, viscous dissipation causes decrease of the entropy).
-
-One of the most commonly used models of artificial viscosity is:
-\begin{equation}
-\Pi_{ab}=- \frac{\nu}{\bar{\rho}_{ab}} \dfrac{ \textbf{v}_{ab} \cdot \textbf{x}_{ab}}{\textbf{x}_{ab}^2 + \left(\eta h\right)^2}
-\label{eq:art-vis-original}
-\end{equation}
-The coefficient $\nu$ is defined as:
-\begin{equation}
-\nu = \alpha \bar{h}_{ab} \bar{c}_{ab}
-\end{equation}
-Where 
-\begin{align}
-\bar{c}_{ab} = \dfrac{c_a + c_b}{2} \\
-\bar{\rho}_{ab} = \dfrac{\rho_a + \rho_b}{2} \\
-\textbf{v}_{ab}=\textbf{v}_a-\textbf{v}_b \\
-\textbf{x}_{ab}=\textbf{x}_a-\textbf{x}_b
-\end{align}
-The artificial viscosity term $\Pi_{ab}$ is a Galilean invariant and vanishes for rigid rotation.
-It produces a repulsive force between two particles when they are approaching each other and an attractive force when they are receding from each other.
-
-The SPH viscosity can be related to a continuum viscosity by converting the summation to integrals \citep{monaghan2005smoothed}.
-It has been shown that the shear viscosity coefficient $\eta = \frac{\rho \alpha h c}{8} $ and the bulk viscosity coefficient $ \zeta = \frac{5 \eta}{3}$ for two dimensional and $\eta = \frac{\rho \alpha h c}{10} $ , $ \zeta = \frac{5 \eta}{3}$ for three dimensional.
-An extra term was added to $\nu$ considering aspects of the dissipative term in shock solutions based on Riemann solvers and lead to a new formulation of artificial viscosity.
-We adopt this new formulation in our simulation:
-\begin{equation}
-\Pi_{ab}^{\beta} = 
-\begin{cases} 
-      \dfrac{- \alpha \mu_{ab} \bar{c}_{ab} + \beta \mu_{ab}^2} {\bar{\rho}_{ab}} & \textbf{v}_{ab} \cdot \textbf{x}_{ab} < 0\\
-      0 & \textbf{v}_{ab} \cdot \textbf{x}_{ab} > 0
-\end{cases}
-\label{eq:art-vis-shock}
-\end{equation}
-Where
-\begin{equation}
-\mu_{ab} = \dfrac{h \textbf{v}_{ab} \cdot \textbf{x}_{ab}}{\textbf{x}_{ab}^2 + \left(\eta h\right)^2} 
-\end{equation}
-%$\alpha$ and $\beta$ are two parameters that are free to be adjusted for each case. 
-$\alpha$ and $\beta$ are two parameters that can be adjusted for different cases.
-$\alpha = 1$ and $\beta = 2$ are  recommended by Monaghan for best results.
-In our simulation, these two parameters are calibrated to  $\alpha = 0.3$ and $\beta = 0.6$.
-$\eta$ is usually taken as 0.1 to prevent singularities.
-%when $\textbf{x}_{ab} = 0$.
 
@@
 \end{equation}
-Where $c_a$ is sound speed at particle a, $d$ is number of dimensions. First order Euler integration, with $\textrm{CFL} = 0.2$, is used to advance in time.
+where $c_a$ is sound speed at particle $a$ and calculated based on heat specific ration of the mixture $\gamma_m$ (See Eq. (\ref{eq:sound-speed-mixture})). 
+$d$ is number of dimensions. 
+\begin{equation}
+c_a = \left( \gamma_m \frac{p}{\rho} \right)^{0.5}
+\label{eq:sound-speed-mixture}
+\end{equation}
+
+First order Euler integration, with $\textrm{CFL} = 0.2$, is used to advance in time.
 
@@
 The classical SPH method was known to suffer from tensile instability and boundary deficiency.
-To address these difficulties, Chen \citep{chen1999improvement} proposed a corrected SPH formulation.
+Tests of the standard SPH method indicate an instability in the tensile regime, while the calculations are stable in compression.
+A simplest example of such test calculation exhibiting the instability involves a body which is subjected to an uniform initial stress, either compressive or tensile.
+If the initial stress is tensile, a very small velocity perturbation on a single particle can lead to particles clumping together, forming large voids and seriously corrupting density distribution.
+But if the initial stress is compressive, the small velocity perturbation on a single particle can not lead to any changes in particle distribution \citep{swegle1995smoothed}.
+To address these difficulties, \citep{chen1999improvement} proposed a corrected SPH formulation.
 For 1D case, employing a Taylor expansion for $A\left(x\right)$ about $x_a$, multiplying both sides by kernel function and then doing integration over the domain gives
@@
 \int_{\Omega} A\left(x\right) w\left(x- x_a, h\right) dx = 
-A_a \int_{\Omega} w\left(x - x_a, h\right) dx +A_{xa} \int_{\Omega} \left(x-x_a\right) w\left(x - x_a, h\right) dx +...
+A_a \int_{\Omega} w\left(x - x_a, h\right) dx +\dfrac{\partial A}{\partial x}(x_a) \int_{\Omega} \left(x-x_a\right) w\left(x - x_a, h\right) dx +...
 \end{equation}
@@
 \end{equation}
-Equation (\ref{eq:CSP-function-approximation-1d}) implies that $\sum_b m_b \dfrac{1}{\rho_b} w\left(x-x_b, h\right) = 1$, which can be viewed as the approximation form of Eq.
+Notice that the denominator in Eq.
+(\ref{eq:CSP-function-approximation-1d}) is actually summation approximation of Eq.
 (\ref{eq:SPH-kernel-normalization-prop}).
+That is to say, Eq.
+(\ref{eq:CSP-function-approximation-1d}) and Eq.
+(\ref{eq:SPH-approximation-sum}) are the same for particles far away from boundaries as the denominator in Eq.
+(\ref{eq:CSP-function-approximation-1d}) becomes $1$ in that case.
 The first order derivative term can be obtained in a similar way:
@@
 
-\subsection{Multiphase SPH}
-Numerical simulation of multiphase flows is usually difficult due to the existence of complex evolving interfaces between phases.
-This presents severe challenges to conventional Eulerian grid-based numerical methods.
-As we discussed in the introduction section, special algorithms to treat and track the interface between different phases is required for mesh based methods to deal with multiphase problems.
-What's worse, the engulfment process during plume development makes geometry of interfaces very complicated.
-SPH is able to handle multiphase problems without additional numerical effort for dealing with interfaces by simply tagging particles of each phase, which makes it suitable for volcanic plume simulation.
-Take the governing equations adopted in our model as an example.
-Different phases of fluid (air and erupted material) are represented by two different sets of SPH particles (or discretization points).
+\subsection{Mass fraction update}
+Air and erupted material are represented by two different sets of SPH particles (or discretization points) in the model.
 Based on assumptions we made in section \ref{sec:physics-model}, only density needs to be updated respectively for each phase.
@@
 \end{equation}
-Where the subscript $a$ and $b$ represents air particles (phase 1) while $i$ and $j$ represents erupted material. $\beta$ and $\alpha$ represents either erupted material or air.
+where the subscript $a$ and $b$ represents air particles (phase 1) while $i$ and $j$ represents erupted material. $\beta$ and $\alpha$ represents either erupted material or air.
 $\rho_a^a$ is density of phase 1 (air). 
  $\rho_i^{sg}$ is density of phase 2 (erupted material).
-$\rho=\rho^a + \rho^{sg}$ is density of mixture of air and erupted material. By definition, the mass fraction is updated according to Eq. (\ref{eq:gov-sph-xi}).
+$\rho=\rho^a + \rho^{sg}$ is density of mixture of air and erupted material. 
+By definition, the mass fraction is updated according to Eq. (\ref{eq:gov-sph-xi}).
 \begin{equation}
@@
 \begin{figure}
-\includegraphics[width=12cm]{Interface.png}
+\includegraphics[width=12cm]{Interface}
 \caption{In the left figure, the blue particles (phaseID=1) represent air particles the red ones (phaseID=2) represent erupted material.
@@
 
+Interface construction will become necessary and important when attempt to include the effects of mixing by resolving the detailed interface structure and dynamics of turbulence. 
+As a Lagrangian method, interface tracking in SPH is explicit through capturing of the locations of the particles, much simplier than Euleruan Methods.
+The existence of complex evolving interfaces between phases presents severe challenges to conventional Eulerian grid-based numerical methods.
+Either interface tracking (Lagrangian) \citep{harlow1965numerical, wrobel1991computational, cheng1995simplified} or interface capturing (Eulerian) \citep{hirt1981volume, youngs1982time, gerlach2006comparison, gopala2008volume} methods are used to reconstruct the flow interface of free boundary flow.
+High computational cost, a tendency to form numerical instabilities and the inability to track complex topological changes are the significant drawbacks of tracking techniques \citep{hirt1981volume, unverdi1992front, anderson1998diffuse}.
+For interface capturing (Eulerian) method, the surrendering of surface detail before the phase transport calculation means that interface reconstruction is required between time steps to recover the interface information, which needs additional numerical effort \citep{hirt1981volume, youngs1982time}.
+Since SPH is able to adaptively adjust the discretization and automatically construct the interface.
+SPH requires less additional numerical effort for interface construction and therefore is more suitable for volcanic plume simulation.
+
+%SPH is able to handle multiphase problems with less additional numerical effort for dealing with interfaces by simply tagging particles of each phase.
+
 \subsection{Turbulence modeling with SPH}
-Typically, turbulence fluctuations in a volcanic column occurs at very different length scales, ranging from millimeters up to hundreds of meters.
-The entrainment of air and heat transfer are dominated by such turbulent fluctuations as turbulent exchange coefficients are several magnitudes larger than corresponding physical coefficients (molecular viscosity and heat conduction coefficient).
+For high speed shearing flow, the momentum exchange and heat transfer are dominated by turbulent fluctuations as turbulent exchange coefficients are several magnitudes larger than corresponding physical coefficients (molecular viscosity and heat conduction coefficient).
+In addition to momentum and enrgy exchange, mixing between plume and air is important in plume modeling.
+Quantifying these mixing processes in real implementation is challenging because of the scale disparity between the large-scale fluid motion and the diffusion processes on interface that ultimately lead to mixing.
+%The energy transfer between these scales occurs through turbulent motion, created either by fluid instabilities or by breaking internal waves.
+For volcanic plume, a shearing flow, the fluid instabilities dominated the mixing process.
 One way to include enough turbulence in the model is to use the SPH-SPS (SPH sub-particle-scale) turbulence model. 
-Another way is to use fine enough resolution (namely, the direct numerical simulation method) at the expense of much higher computational cost which results from both increase in number of particles and decrease in time steps constrained by the CFL (Courant–Friedrichs–Lewy) condition.
+Another way is to use fine enough resolution (namely, the direct numerical simulation method) at the expense of much higher computational cost which results from both increase in number of particles and decrease in time steps constrained by the CFL (Courant-Friedrichs-Lewy) condition.
 Here we choose the SPH-SPS method.
@@
 \citet{monaghan2011turbulence} constructed $SPH-\varepsilon$ turbulence model within the framework of SPH in such a way that general principles such as conservation of energy, momentum and circulation are satisfied using the ideas associated with the LANS turbulence modeling.
-The basic idea of $SPH-\varepsilon$ is to determine a smoothed (averaged) velocity $\widehat{\textbf{v}}$ by a linear operation on the unsmoothed velocity $\textbf{v}$.
+The basic idea of $SPH-\varepsilon$ is to determine a smoothed (averaged in space) velocity $\widehat{\textbf{v}}$ by a linear operation on the unsmoothed velocity $\textbf{v}$.
 The SPH particles move with this smoothed velocity and hence the average motion of the fluid is determined by the averaged velocity $\widehat{\textbf{v}}$:
@@
 \end{equation}
-Where $G$ satisfies:
+where $G$ satisfies:
 \begin{equation}
@@
 \end{equation}
-Where $K_{ab} = l^d G_{ab}$, $M = \rho_0 l^d$ in which $d$ is the dimension and $\rho_0$ is initial density. $SPH-\varepsilon$ turbulence model is obtained after lengthy derivation:
+where $K_{ab} = l^d G_{ab}$, $M = \rho_0 l^d$ in which $d$ is the dimension and $\rho_0$ is initial density. $SPH-\varepsilon$ turbulence model is obtained after lengthy derivation:
 \begin{equation}
@@
 \end{equation}
-Notice that if $l$ is uniform: 
+Notice that if $l$ is constant: 
 %Please notice that the above equation is slightly different from the original equation in Monaghan's paper (Eq.
@@
 \end{equation}
-Where, $\mu$ is the dynamic viscosity, $\kappa$ is the thermal conductivity. And $\mu$  can be written in term of the absolute viscosity (kinematic viscosity) as:
+where, $\mu$ is the dynamic viscosity, $\kappa$ is the thermal conductivity. And $\mu$  can be written in term of the absolute viscosity (kinematic viscosity) as:
 \begin{align}
@@
 The turbulent viscosity coefficient can be inferred from that formulation if we can reformulate the turbulent shear stress term in a form which is similar to the molecular shear term.
-Reformulate the turbulent shear stress term:
+Reformulating the turbulent shear stress term:
 \begin{equation}
@@
 \end{equation}
-Then the turbulent viscosity coefficient can be inferred from Eq. (\ref{eq:turb-stress-reformulate-to-artificial-vis}).
+then the turbulent viscosity coefficient can be inferred from Eq. (\ref{eq:turb-stress-reformulate-to-artificial-vis}).
 \begin{equation}
@@
 
-However, the above equation is correct only for the 1D situation. For 2D or 3D, it is not easy to get an explicit expression. 
+However, the above equation is correct only for the 1D situations. For 2D or 3D, it is not easy to get an explicit expression. 
 We adopt an alternative way: obtaining a value for each pair of particles instead of persisting on getting an analytical expression.
@@
 \end{equation}
-Second spatial derivative can be approximated with SPH by following \citet{monaghan2005smoothed}. 
+Second spatial derivative can be approximated with SPH by following \citet{monaghan2005smoothed}: 
 \begin{equation}
@@
 \end{equation}
-Where
-\begin{equation}
-F_{ab} \textbf{x}_{ab} = \nabla _a w_{ab}
-\end{equation}
-$F_{ab}$ is nonpositive, which guarantees that heat flux flows from hot to cold. 
+where $F_{ab} (h)$ is short for $F \left( \textbf{x}_a - \textbf{x}_b, h \right)$, whose definition is:
+\begin{equation}
+F_{ab}(h) \textbf{x}_{ab} = \nabla _a w_{ab}
+\end{equation}
+$F_{ab}$ is always nonpositive, which guarantees that heat flux flows from hot to cold. 
 Plug the turbulent thermal conductivity into the heat conduction equation:
@@
 (\ref{eq:wall_bc_e})).
-
+In our current model, the ground is assumed to be a flat ground.
+For more complicated topography, it has been shown in other paper \citep {kumar2013parallel} that it works as well.
+One of our concern regarding potential limitation of such method is deployment of ghost particles for more complicated boundaries which is not described in detail in the paper.
 \subsubsection{Eruption boundary condition}
@@
 \end{equation}
-Where $CFL_p$ is a safety coefficient which has similar function as the normal $CFL$ number.
+where $CFL_p$ is a safety coefficient which has similar function as the normal $CFL$ number.
 Too small $CFL_p$ would slow down simulation while too large $CFL_p$ would lose its ability of mitigate numerical fluctuation near the boundary.
@@
 
-More details about the data structure, domain decomposition, load balancing and domain adjusting will be published separately.
-
-\begin{figure}[!t]
-\centering
-\includegraphics[scale=0.35]{strong}
+More details about the data structure, domain decomposition, load balancing, domain adjusting and performance testing has been published separately \citep{cao2017data}.
+
+\begin{figure}[!htb]
+    \centering
+    \begin{minipage}{.425\textwidth}
+		\centering
+		\includegraphics[scale=0.45]{strong}
 		\caption{Strong scalability test result}
 		\label{fig:strong_scale}
-\end{figure}
-
-\begin{figure}[!t]
-\centering
-\includegraphics[scale=0.35]{weak_scale}
+    \end{minipage}%
+    \begin{minipage}{.425 \textwidth}
+		\centering
+		\includegraphics[scale=0.45]{weak_scale}
 		\caption{Weak scalability test result}
 		\label{fig:weak_scale}
+    \end{minipage}% 
 \end{figure}
 
+%\begin{figure}[!t]
+%\centering
+%\includegraphics[scale=0.35]{strong}
+%\caption{Strong scalability test result}
+%\label{fig:strong_scale}
+%\end{figure}
+%
+%\begin{figure}[!t]
+%\centering
+%\includegraphics[scale=0.35]{weak_scale}
+%\caption{Weak scalability test result}
+%\label{fig:weak_scale}
+%\end{figure}
+%
 \begin{figure}[!t]
@@
 
-Performance tests have been carried out on the computational cluster of CCR (Center for Computational Research) at Buffalo.
+Performance tests have been carried out on the computational cluster of CCR (Center for Computational Research) at University at Buffalo.
 Intel Xeon E5645 CPUs running at 2.40GHz clock rate with 4GB memory per core on a Q-Logic Infiniband is used in these tests.
@@
 Memory and level 3 cache are shared on each node.
-The initial domain is $[-4.8km,4.8km] \times [-4.8km,4.8km] \times [0km, 6km]$.
+The initial domain is $[-4.8km, 4.8km] \times [-4.8km,4.8km] \times [0km, 6km]$.
 Almost linear speed up is observed in our strong scalability test (Fig.
@@
 \ref{fig:weak_scale}, simulation times increase around $1/3$ when number of cores increase from 16 to 4208.
-For the test problem in this section, the volcanic plume finally reaches to a region of $[-30km \,\,\, 30km] \times [-30km\,\,\,30km] \times [1.5km\,\,\,40km]$ after around 400 seconds of eruption.
-When numerical simulation goes up to 90 seconds, the plume is still within a region of $[-10km\,\,\,10km] \times [-10km\,\,\,10km] \times [0km\,\,\,25km]$.
+For the test problem in this section, the volcanic plume finally reaches to a region of $[-30km , 30km] \times [-30km, 30km] \times [1.5km, 40km]$ after around 400 seconds of eruption.
+When numerical simulation goes up to 90 seconds, the plume is still within a region of $[-10km, 10km] \times [-10km, 10km] \times [0km,25km]$.
 This implies that adjusting of domain can avoid computing large number of uninfluenced air particles, especially for the beginning stage of simulation.
@@
 We present a series of numerical simulations to verify and validate our model in this section.
-Plume-SPH is first verified by a JPUE (jet or plume that is ejected from a nozzle into a uniform environment) simulation.
+Plume-SPH is first verified by 1D shock tube tests then by a JPUE (jet or plume that is ejected from a nozzle into a uniform environment) simulation.
 Velocity and mass fraction distribution both along the central axis and cross transverse are compared with experimental results.
@@
 Then a simulation of representative strong volcanic plume is conducted.
-Both global variables and local variable are comparable with simulation results from existing 3D plume models.
+Intergrated local variable are comparable with simulation results from existing 3D plume models.
+\subsection{1D Shock Tube Tests}
+1D shock tube tests are first conducted to verify our code. Input parameter of each tests can be found in Table \ref{tab:1D-shock-input_parameters}. These tests can represent typical cases in 1D.
+Test 1 consists of a left rarefaction, a right travelling contact and a right shock. Density decreases at down wind of contact wave.
+Test 2 also consists of a left rarefaction, a right travelling contact and a right shock. Density increases at down wind of contact wave.
+Test 3 is a double expansion test with different initial dnesity.
+
+\begin{table}[htp]
+\centering
+      \caption{Input parameters of 1D shock tube tests}		
+	  \begin{tabular}{lrrrrrrr}
+	    \hline
+	          & $\rho_L$ & $p_L$ &$v_L$ & $\rho_R$ & $p_R$ &$v_R$ & $t_f$\\
+	    \hline
+	    Test 1 & $1.0$ & $1.0$ &$0$ & $0.5$ & $0.2$ &$0$ & $0.2$\\
+	    Test 2 & $1.0$ & $1.0$ &$0$ & $0.25$ & $0.1795$ &$0$ & $0.17$\\
+	    	Test 3 & $2.0$ & $1.95$ &$1.0$ & $1.0$ & $1.95$ &$-1.0$ & $0.13$\\
+	    \hline
+	  \end{tabular}
+	  \label{tab:1D-shock-input_parameters}
+\end{table}
+
+In the table \ref{tab:1D-shock-input_parameters},  subscript $L$ refers left side and $R$ for right side, $t_f$ is the total simulation time.
+The initial interval between two adjcent particles is 0.03.
+The computational domain is $[-0.4, 0.4]$.
+The specific internal energy is is compared against exact solutions.
+As shown in Fig.
+\ref{fig:1D-shock-tests-verification}, the position and magnitude of the waves are correctly predicted.
+The fluctuations near the contact discontinuity are caused by sharp change of smoothing length.
+
+\begin{figure}[!htb]
+    \centering
+    \begin{minipage}{.325\textwidth}
+        \centering
+        \includegraphics[width=0.99 \textwidth]{./GSPH-Sod-verification-e}
+    \end{minipage}%
+    \begin{minipage}{.325 \textwidth}
+        \centering
+        \includegraphics[width=0.99 \textwidth]{./Sod-verification-e}
+    \end{minipage}%
+    \begin{minipage}{.325 \textwidth}
+        \centering
+        \includegraphics[width=0.99 \textwidth]{./d-exp-verification-e}
+    \end{minipage}%   
+    \caption{Comparison of specific internal energy of simulation results agaist analytical results for shock tube tests.
+The plots from left to right are corresponding to test 1, test 2 and test 3 respectively.}
+    \label{fig:1D-shock-tests-verification}
+\end{figure}
+
 \subsection{Simulation of JPUE}
@@
 While the effect of stratified atmosphere and the effect of expansion due to high temperature in volcanic plume are not represented, JPUE reproduces the entrainment due to turbulent mixing which is one of the key elements in volcanic plume development.
-There exist consistently good experimental data \citep { list1982turbulent,dimotakis1983structure, papanicolaou1988investigations} that describe the JPUE flow field giving an insight into details of JPUE, such as transverse velocity and concentration profile.
+There exist consistently good experimental data \citep { list1982turbulent,dimotakis1983structure, papanicolaou1988investigations, ezzamel2015dynamical} that describe the JPUE flow field giving an insight into details of JPUE, such as transverse velocity and concentration profile.
 In this section, we verify that our code and the $SPH-\varepsilon$ turbulence model is able to reproduce feature of turbulent entrainment by a JPUE simulation.
@@
 \end{equation}
-Where $c$ is the speed of sound in the liquid.
+where $c$ is the speed of sound in the liquid.
 The energy equation is actually decoupled from the momentum conservation equation and the mass conservation equation by using this equation of state (EOS).
@@
 	    Vent velocity          & $m\cdot s^{-1}$  & 500               & 275 \\
-	    Vent gas mass fraction &                  & 1.0               & 0.5 \\
+	    Vent gas mass fraction &                  & 1.0               & 0.05 \\
 	    Vent Temperature       & $K$              & 273               & 1053 \\
@@
 
-\begin{figure}
+\begin{figure}[!htb]
+    \centering
+    \begin{minipage}{.425\textwidth}
+        \centering
         \includegraphics[width=8cm]{vel_cross}
-\caption{Dimensionless velocity distribution across the cross-section.}
-\label{fig:JPUE_cross-section_vel}
+    \end{minipage}%
+    \begin{minipage}{.425 \textwidth}
+        \centering
+        \includegraphics[width=8cm]{conc_cross}]
+    \end{minipage}%  
+    \caption{Dimensionless concentration and velocity distribution across the cross-section.}
+    \label{fig:JPUE_cross-section}
 \end{figure}
-\begin{figure}
-\includegraphics[width=8cm]{conc_cross}
-\caption{Dimensionless concentration distribution across the cross-section.}
-\label{fig:JPUE_cross-section_conc}
+
+\begin{figure}[!htb]
+    \centering
+    \begin{minipage}{.425\textwidth}
+        \centering
+        \includegraphics[width=8cm]{velo_along_axis}
+    \end{minipage}%
+    \begin{minipage}{.425 \textwidth}
+        \centering
+        \includegraphics[width=8cm]{conc_along_axis}
+    \end{minipage}%  
+    \caption{The left plot is normalized jet width (which determined based on velocity) along the centerline. The right plot shows normalized concentration along the centerline.}
+    \label{fig:JPUE_along-axis}
 \end{figure}
-
-\begin{figure}
-\includegraphics[width=8cm]{velo_along_axis}
-\caption{Dimensionless velocity distribution along centraline.}
-\label{fig:JPUE_along-axis_vel}
-\end{figure}
-\begin{figure}
-\includegraphics[width=8cm]{conc_along_axis}
-\caption{Dimensionless concetration distribution along centraline.}
-\label{fig:JPUE_along-axis_conc}
-\end{figure}
+%
+%\begin{figure}
+%\includegraphics[width=8cm]{vel_cross}
+%\caption{Dimensionless velocity distribution across the cross-section.}
+%\label{fig:JPUE_cross-section_vel}
+%\end{figure}
+%
+%\begin{figure}
+%\includegraphics[width=8cm]{conc_cross}
+%\caption{Dimensionless concentration distribution across the cross-section.}
+%\label{fig:JPUE_cross-section_conc}
+%\end{figure}
+
+%\begin{figure}
+%\includegraphics[width=8cm]{velo_along_axis}
+%\caption{Dimensionless velocity and concentration distribution along centraline.}
+%\label{fig:JPUE_along-axis_vel}
+%\end{figure}
+%\begin{figure}
+%\includegraphics[width=8cm]{conc_along_axis}
+%\caption{Dimensionless concetration distribution along centraline.}
+%\label{fig:JPUE_along-axis_conc}
+%\end{figure}
 
@@
 Material properties of water are used as material properties for both phases.
-The results are compared with experiments \citep {george1977turbulence, papanicolaou1988investigations} for validation purposes.
-Experimental data of concentration and velocity distribution across the cross-section were fit into a Gaussian profile: $\dfrac{\varphi}{\varphi_c}=exp \left[-coef\left( \left(\dfrac{r}{z}\right)^2\right)\right]$ (solid line) by \citet{papanicolaou1988investigations} even though the actual profile are slightly different from the Gaussian profile.
-\citet{papanicolaou1988investigations} also fit concentration distribution and jet width based on velocity along central line into a straight line : $\dfrac{\varphi_0}{\varphi_c}=slope \left(\dfrac{z}{D} + intercept \right)$ (dash line).
-$\varphi$ is either velocity or concentration, the subscript $c$ represents the centreline, subscript $0$ represents the cross-sectionally averaged exit value.
+The results are compared with experiments \citep {george1977turbulence, papanicolaou1988investigations} for verification purposes.
+Experimental data of concentration and velocity distribution across the cross-section were fit into a Gaussian profile (See Eq.
+(\ref{eq:JPUE-experiment-fit-corss})) by \citet{papanicolaou1988investigations} and  \citet{ george1977turbulence} even though the actual profile are slightly different from the Gaussian profile.
+
+\begin{equation}
+\dfrac{\varphi}{\varphi_c}=exp \left[-coef\left( \left(\dfrac{r}{z}\right)^2\right)\right]
+\label{eq:JPUE-experiment-fit-corss}
+\end{equation}
+
+where $\varphi$ is either velocity or concentration, the subscript $c$ represents the centreline.
 $r$ is the distance from the centreline on any cross-section.
 $z$ is the axial distance from the origin of the jet transverse section under consideration.
-$D$ is the diameter of vent.
 The coefficient $coef$ for concentration is 80 and 50 respectively according to \citet{george1977turbulence} and \citet{papanicolaou1988investigations}.
 $coef$ for velocity is 90 and 55 respectively according to \citet{george1977turbulence} and \citet{papanicolaou1988investigations}.
-$slope$ for jet width based on velocity is 0.104 and for concentration is 0.157. $intercept$ for jet width based on velocity is 2.58 while that for the concentration is 4.35.
+
+\citet{papanicolaou1988investigations} also fit concentration distribution and jet width based on velocity along centerline into a straight line (See Eq. (\ref{eq:JPUE-experiment-fit-along-axis}))..
+
+\begin{equation}
+\dfrac{\varphi_0}{\varphi_c}=slope \left(\dfrac{z}{D} + intercept \right)
+\label{eq:JPUE-experiment-fit-along-axis}
+\end{equation}
+
+where subscript $0$ represents the cross-sectionally averaged exit value, $D$ is the diameter of vent. 
+$slope$ for jet width based on velocity is 0.104 and for concentration is 0.157. 
+$intercept$ for jet width based on velocity is 2.58 while that for the concentration is 4.35.
 
 Although both velocity and concentration are found to be well matched with experimental results, a small disparity in both velocity and concentration are observed near the boundary of the jet.
-There are several factors that would attribute to such disparity.
+Which is possibly caused by overestimating of the drag effect by standard SPH \citep {ritchie2001multiphase}.
+\citet {ritchie2001multiphase} also proposed an alternative way for density update which releived the overestimating of the drag effect.
+However, how well does his method conserve mass is not clear.
+There are several other factors that might also attribute to such disparity.
 Reynolds number is not reported in many experiments assuming a high enough Reynolds number.
@@
 
-\subsection{Simulation of a plume}
+\subsection{Simulation of a volcanic plume}
 The development of a volcanic plume is more complicated than JPUE in several aspects.
@@
 \subsubsection{Input parameters}
-Eruption parameters, material properties and atmosphere are chosen to be the same as the strong plume no wind case in a comparison study on eruptive column models \citep{costa2016results}.
+Eruption parameters, material properties and atmosphere are chosen to be the same as the strong plume no wind case in a comparison study on eruptive column models by \citet {costa2016results}.
 Eruption conditions are listed in Table \ref{tab:input_parameters}.
+As our model does not distinguish solid particles of different sizes, only mass fraction of solids of all size is used in simulation even though two size classes were provided by \citet {costa2016results}.
 The density of erupted material at the vent and radius of the vent can be computed from the given parameters.
@@
 Vertical distribution of temperature, pressure and density is used to establish stratified atmosphere.
+Wind velocity and specific humidity are not used in our simulation even though they were also provided in Fig.
+1B \citep{costa2016results}.
 Material properties, shown in Table \ref{tab:material_properties}, are selected based on properties of Pinatubo and Shinmoe-dake eruption.
@@
 	    	Specific heat of air at constant pressure   & $J \cdot kg^{-1}\cdot K^{-1}$& 1810    \\
-	    	Density of air at vent height               & $kg \cdot m^{-2}$       & 1.104   \\
+	    	Density of air at vent height               & $kg \cdot m^{-3}$       & 1.104   \\
 	    Pressure at vent height                        & $Pa$              & 84363.4 \\
@@
 
+Fig.
+\ref{{fig:pinatubo-simulation-results-vis}} shows the mass fraction of simulated volcanic plume at 300s after eruption, at which time the plume starts spreading radially.
+A contour plot of the mass fraction on the vertical cross-section (y-z plane) were also provided.
+The zoomed view of the quivier plot shows detailed entrainment of air at the margin of the plume.
+
+\begin{figure}[!htb]
+    \centering
+    \begin{minipage}{.45\textwidth}
+        \centering
+        \includegraphics[width=0.99 \textwidth]{./mssfrc-Diverging}
+    \end{minipage}%
+    \begin{minipage}{.45 \textwidth}
+        \centering
+        \includegraphics[width=0.99 \textwidth]{./mssfrc-Diverging-cut}
+    \end{minipage}%
+    \\
+    \begin{minipage}{.45 \textwidth}
+        \centering
+        \includegraphics[width=0.99 \textwidth]{./gmd-mssfrc-yz}
+    \end{minipage}% 
+    \begin{minipage}{.45 \textwidth}
+        \centering
+        \includegraphics[width=0.99 \textwidth]{./gmd-mssfrc-yz-zoomed}
+    \end{minipage}%  
+    \caption{Mass fraction for $t=300s$ after eruption.
+Figures on the first row are visualization of SPH simulation results.
+The figure on the up right corner is visualization of a slice of the computational, whose thickness is around $10000m$.
+The lowest portion of the plume represents erupted material in eruption vent (the underground portition).
+Figures at the second row show contour of mass fraction and velocity quiver on y-z plane.
+These figures are plot utilizing post-processed data (see Appendix \ref{app:project-SPH-grid}).
+The contour level in the plot are $0.00001, 0.0001, 0.001, 0.1, 0.3, 0.75, 0.95$.
+The figure on the lower right corner is a zoomed view of velocity quivier showing plume expansion and entrainment of air.}
+    \label{fig:pinatubo-simulation-results-vis}
+\end{figure}
+
 \subsubsection{Global and local variables}
+One of the key global quantities of great interest is the altitude to which the plume rises.
+The top height predicted by our model is around 40 km which agrees with other plume models.
+For example, the height predicted by PDAC is 42500 m, by SK-3D is 39920 m, by ATHAM is 33392 m, by AHSEE is 36700 m.
+As for local variables, the profiles of integrated temperature, density, mass fraction of entrained air, gas mass fraction, mass fraction of solid and the radius of the plume as a function of height are compared with existing 3D models in Fig.
+\ref{fig:strong_local_temp} $\sim$ \ref{fig:strong_local_radius}.
+To get rid of significant fluctuations in time and space we conducted a time averaging and spatial integration of the dynamic 3D flow fields by following \citet {cerminara2016large}.
+
+As particles distribute irregularly in the space in SPH simulation results.
+We need to project simulation results (on irregular particles) onto a pre-defined grid before doing time average and spatial integration.
+See appendix A for more details of post processing.
+
+The profiles of local variables match well with simulation results of existing 3D models in a general sense. The basic phenomena in volcanic plume development is correctly captured by our model.
 
@@
 \end{figure}
-\begin{figure}
-\center
-\includegraphics[width=7cm]{density_strong}
-\caption{Density of the strong plume without wind after reaching its top height}
-\label{fig:strong_local_density}
-\end{figure}
-
-One of the key global quantities of great interest is the altitude to which the plume rises.
-The top height predicted by our model is around 40 km which agrees with other plume models.
-For example, the height predicted by PDAC is 42500 m, by SK-3D is 39920 m, by ATHAM is 33392 m, by AHSEE is 36700 m.
-As for local variables, the profiles of integrated temperature, density, mass fraction of entrained air, gas mass fraction, mass fraction of solid and the radius of the plume as a function of height are compared with existing 3D models in Fig.
-\ref{fig:strong_local_temp} $\sim$ \ref{fig:strong_local_radius}.
-To get rid of significant fluctuations in time and space we conducted a time averaging and spatial integration of the dynamic 3D flow fields by following \citet {cerminara2016large}.
-As particles distribute in a disordered manner in the space in SPH simulation results.
-We first project simulation results (on disordered particles) onto a pre-defined grid before doing time average and spatial integration.
-The project method is the basic SPH kernel based interpolation.
-
-The profiles of local variables match well with simulation results of existing 3D models in a general sense. The basic phenomena in volcanic plume development is correctly captured by our model.
 
@@
 \label{fig:strong_plume_mass_fraction}
-\end{figure}
-%
-
-\begin{figure}
-\center
-\includegraphics[width=5cm]{radius_strong}
-\caption{Radius of the strong plume without wind after reaching its top height.}
-\label{fig:strong_local_radius}
 \end{figure}
@@
 (\ref{eq:gas-frac-express})).
+\begin{equation}
+\xi_a + \xi_g = \xi_a + \left(1-\xi_a\right) \xi_{g0}
+\label{eq:gas-frac-express}
+\end{equation}
+
 Among these 3D models, ATHAM takes vapor condensation into account and Eq.
@@
 (\ref{eq:solid-frac-express2}) hold for our model.
+\begin{equation}
+\xi_s = \left(1 - \xi_a\right) \left(1- \xi_{g0}\right)
+\label{eq:solid-frac-express1}
+\end{equation}
+\begin{equation}
+\xi_s = 1 - \left(\xi_a + \xi_g\right)
+\label{eq:solid-frac-express2}
+\end{equation}
+
 PDAC, which treat particles of two different sizes as two separate phases, predicted a similar mass fraction profile.
 That implies that assumption of dynamic equilibrium in our model is at least valid for eruptions similar to the test case.
-
-\begin{equation}
-\xi_a + \xi_g = \xi_a + \left(1-\xi_a\right) \xi_{g0}
-\label{eq:gas-frac-express}
-\end{equation}
-\begin{equation}
-\xi_s = \left(1 - \xi_a\right) \left(1- \xi_{g0}\right)
-\label{eq:solid-frac-express1}
-\end{equation}
-\begin{equation}
-\xi_s = 1 - \left(\xi_a + \xi_g\right)
-\label{eq:solid-frac-express2}
-\end{equation}
 
@@
 \ref{fig:strong_local_density}).
-
+\begin{figure}
+\center
+\includegraphics[width=7cm]{density_strong}
+\caption{Density of the strong plume without wind after reaching its top height}
+\label{fig:strong_local_density}
+\end{figure}
+\begin{figure}
+\center
+\includegraphics[width=7cm]{radius_strong}
+\caption{Radius of the strong plume without wind after reaching its top height.}
+\label{fig:strong_local_radius}
+\end{figure}
 Our model adopts the same assumptions and governing equations as SK-3D.
-However, there is still a big disparity between the profiles of local variables of our model and SK-3D.
+However, there is still a obvious disparity between the profiles of local variables of our model and SK-3D.
 One of the big differences between these two models is that we adopt a LANS type of turbulence model while SK-3D adopts a LES (large eddy simulation) turbulence model.
@@
 %\subsection{Data availability}
-%Data for running the test cases, including meteorological data, properties of erupted material and air, and eruption condition data are archived in the source code repository.
-Raw data of simulation results are archived in Box.
-Access permission will be given upon request.
-Post-processed data for plotting figures is provided as a supplement along with the manuscript.
-Please note that some of these pictures are generated by visualizing the raw data with visualization tools, there is no separate data for these figures.
-%\appendix
-%\section{}    %% Appendix A
+
+\appendix
+\section{Post process of particle data} \label{app:project-SPH-grid}    %% Appendix A
+\appendixfigures
+\begin{figure}[!htb]
+    \centering
+    \begin{minipage}{.325\textwidth}
+        \centering
+        \includegraphics[width=0.90 \textwidth]{./Project1}
+    \end{minipage}%
+    \begin{minipage}{.325 \textwidth}
+        \centering
+        \includegraphics[width=0.90 \textwidth]{./Project2}
+    \end{minipage}%
+    \begin{minipage}{.325 \textwidth}
+        \centering
+        \includegraphics[width=0.90 \textwidth]{./Project3}
+    \end{minipage}%   
+    \caption{Procedure of projection of simulation results carried by particles onto regular grids.
+as shown in these figures from left to right: 1) raw data of SPH simulation results (irregularly distributed in space), 2) Add regular mesh, 3) searching for neighbours of each node (the blue SPH particles within the green circle around the red dot).
+The last step is not shown in these pictures, which is treating each node on the regular mesh as a SPH particle and projecting data on particles onto nodes utilizing "SPH interpolation" ( see Eq.
+(\ref{eq:SPH-fundamental-principle})).}
+    \label{fig:1D-shock-tests-verification}
+\end{figure}
+
+Particles distribute irregularly in SPH simulation results.
+To adopt post process originally proposed for mesh-based method, we need to project simulation results onto a pre-defined regular mesh.
+Fig.
+\ref{fig:1D-shock-tests-verification} shows the procedure for posting process.
+
 %
@@
 \end{acknowledgements}
+
+\clearpage
+
 %% REFERENCES
@@
 \bibliographystyle{copernicus}
-\bibliography{Reference.bib}
+\bibliography{Reference}
 %% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,

### 2017-09-08 Reply-to-RC1.tex
Response to 1st Ref added

\documentclass[10pt,a4paper]{article}
%\documentclass[10pt,a4paper]{beamer}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\usepackage{kpfonts}
\usepackage[numbers]{natbib}
%other options include: round, curly, angle, or default options

\usepackage{media9}

\author{\Large \textbf{by Zhixuan Cao et al.}}

\title{ \Large{ Authors' reply to the comments by M. de’ Michieli Vitturi (Referee \#1) of the manuscript gmd-2017-119} \\
\LARGE \textbf{“Plume-SPH 1.0: A three-dimensional, dusty-gas volcanic plume model based on smoothed particle hydrodynamics”}}

\date{\vspace{-5ex}}

\begin{document}
\maketitle
The comments by the reviewer are recited in italics, followed by our reply in upright font. Equation, section, and figure numbers that we refered in our response are corresponding to these in the original varsion of GMD discussion paper. \\[12pt]

\textbf{\large General Comments}

\textit{This paper presents an initial effort towards developing a plume model with comprehensive physics based on the smoothed particles hydrodynamics method. To my knowledge, this is the first numerical model of volcanic plumes adopting this technique and this is a good addition to the existing models. While the model, the derivation of the discretised equations and the computational techniques are satisfactory for a reader expert in SPH, a few more details would help in the comprehension other readers. The scientific content of the paper could be improved by a better and extended description of the applications. Overall, the manuscript represent a substantial contribution to modelling Âascience within the scope of Geoscientific Model Development and it is suitable for publication after some improvements and corrections are made. These are described below.}

We would like to thank the reviewer for carefully reading our manuscript and giving positive review and constructive remarks.
We have revised the manuscript as shown in the supplemental PDF file, and we hope that we have dealt with all suggestions in an adequate manner. The revised manuscript is also attached.

On behalf of all co-authors.

The following are our response to reviewer's specific comments. Modifications are made accordingly in the manuscript. \\[12pt]

\textbf{\large Specific comments}

\textbf{Abstract}

\textit{Line 8. What does dynamic and thermodynamic equilibrium with surrounding air mean? No relative velocity? Same temperature? I think it is an equilibrium between volcanic gas and particles.}

Yes, dynamic equilibrium means same velocity while thermodynamic equilibrium means same temperature. We treat erupted material (includes volcanic gas and particles) as one phase, so assumption of dynamic equilibrium and thermodynamic equilibrium are made for erupted material. Such assumption is made implicitly when we assuming erupted material behaves like a single phase fluid.
What we intended to say by "dynamic and thermodynamic equilibrium between air and erupted material" is actually "dynamic equilibrium and thermodynamic equilibrium between erupted material and air that is entrained into the plume". That is to say, we assume that surrounding air and erupted material will reach an equilibrium status (same velocity and temperature) as soon as they get mixed.

We are aware of that the original statement is misleading and modified it as "dynamic equilibrium and thermodynamic equilibrium between erupted material and air that is entrained into the plume". We thanks reviewer for pointing this out. \\[3pt]

\textit{Line 23. The model is not really compared with the top height of the Pinatubo eruption, because atmospheric conditions are changed from real ones (no wind in imposed,
weak vs strong in Costa et al 2016); in addition, MER is fixed. So, it is more a comparison with results from other 3D models, for an eruption with the same MER estimated for Pinatubo 1991.}

Thank you for your clarification on this. We revised original statement "The model is verified by comparing velocity and concentration distribution along the central axis and on the transverse cross with experimental results of JPUE (jet or plume that is ejected from a nozzle into a uniform environment) and the top height of the Pinatubo eruption of 15 June 1991." to "The code is first verified by 1D shock tube tests, then by comparing velocity and concentration distribution along the central axis and on the transverse
cross with experimental results of JPUE (jet or plume that is ejected from a nozzle into a uniform environment). 
%The model is validated by comparing against simulation results from other 3D plume models.
Profiles of several integrated variables are compared with those calculated by existing 3D plume models for an eruption with the same MER (mass eruption rate) estimated for the Pinatubo eruption of June 15 1991. Our results are consistent with existing 3D plume models." \\[6pt]

\textbf{Section 1.3}

\textit{Page 4, lines 23-25. It is not clear to me why interface tracking or interface capturing are mentioned here for mesh based method. I think all the existing 3D Eulerian models (based on mesh based method) do not model the interface between the plume and the
atmosphere. While I understand that this represent a big advantage for other applications of SPH, for example for dam-break problems, because it allows to solve only for the “flow” region and not for the surrounding, for a volcanic plume mixing between plume and air is important and I don’t really see a clear interface between them. Please clarify and expand the point.}

We first thank reviewer for asking this question. This question pushed us to do more investigation on "interface construction" in mesh based methods. More importantly, we got an deeper understanding on why we need turbulence model in plume modeling. Our original motivation for using a turbulence model was simply based on the following logic: The Reynolds number is large enough hence the flow is turbulent, we can not resolve all turbulent exchange at all different scales with a coarse resolution, so we need turbulence model to resolve sub-particle scale turbulence. Now we can see the necessary of including a turbulence model in purpose of resolving mixing which happens in a much smaller scale.

Because interface acts as portion of boundaries.
It is critical to capturing or tracking the interface during simulation for immiscible flow. This is one of major motivations for using interface tracking and interface capturing method. However, this is not the case for volcanic plume modeling. As "fluids" in our plume model are miscible.

Even though, clear transient interface do exist in miscible flow. There are experiments \citep {papantoniou1989large} showing clear interface between air and plume. Another experiment with higher resolution showing more detailed and more complicated interface by \citet{crimaldi2001high}. 
It is through the interface (or boundary) where mxing happens. \citet{jacobson2008mixing} study detailed mixing mechanism at a flat interface.

However, quantifying these mixing processes in real implementation is challenging because of the scale disparity between the large–scale fluid motion and the diffusion processes on interface that ultimately lead to mixing. The energy transfer between these scales occurs through turbulent motion, created either by fluid instabilities or by breaking internal waves. For shearing flow, like volcanic plume, the fluid instabilities dominated the mixing process. Ideally, one would like to be able to include the effects of mixing on the large scale dynamics without resolving the detailed interface structure and dynamics of turbulence to reduce computational cost. Such strategy was also used in all other mesh-based 3D plume models (though different turbulence closures were adopted by different models). 
By resolving sub-particle scale (or sub-mesh scale) turbulent mixing with turbulence closure, relative coarse resolution is used in all 3D plume models (both particle based model and mesh-based models). Such solution is a trade-off between computational accuracy and computational cost. And in this case, interface might not be clearly observed in simulation due to coarse resolution.

Interface construction will become necessary and important when we attempt to include the effects of mixing by resolving the detailed interface structure and dynamics of turbulence. This option is not attractive at current stage of plume modeling as turbulence closure can give an acceptable accuracy with affordable computational cost.

Remedies of standard SPH is needed if detailed mixing process needs to be resolved, for example \citep{cha2010kelvin, zhuimproved}.
Remedies, such as adding of surface tension \citep{behjati2017modelling}, are needed for tracking interface while simulating immiscible flow using SPH. Even though, as a Lagrangian method, interface constructing is explicit through capturing of the locations of the particles and much simplier than Eulerian Methods.

Modification in the manuscripts:
1) Since the exact location of interface is less important for plume modeling when turbulence model is adopted, we changed this paragraph to: 
"Interface tracking is explicit in SPH through capturing of the locations of the particles. Less numerical effort is required for interface construction when we attempt to include the effects of mixing by resolving the detailed interface structure and dynamics of turbulence." and put it as the last item among SPH features.
2) Added a short paragraph (Basically the fourth paragraph of this response) at the beginning of section 3.8 explaining motivation of using turbulence model. 
3) Changing the title of section 3.7 to "Mass fraction update". Re-organize this section and focus more on mass fraction updating. Make a brief remark on interface constructing using SPH and mesh-based methods at the end of this section. \\[6pt]

\textbf{Section 3. SPH method}

\textit{Sometime equation are referenced before they have been introduced. See for example lines 23 and 24 at page 10 where there is a reference to Eqs. 34 and 35, which are in introduced in the next subsection. Again, at page 10, sentence at lines 25-26 should be referred to the discretised version of the equations, while discretisation is presented later in the section. Page 11, line 1. It is written “region of compact support” before having stated the
properties of the weighting function. Maybe add an equation that define this property: $w(x-xb,h) = 0 if |x-xb|>kh$ for some $k$, i.e. the support is proportional to the smoothing length. A suggestion. . . Probably the issues raised above could be solved moving the content of section 3.2 to section 3.1, after equation 26.}

As suggested by the reviewer, we reorganized the order of equations and merged section 3.1 and  section 3.2 into one section. We also put section for "Artificial visosity" before the section for "governing equation discretization". \\[6pt]

\textbf{Section 3.3}

\textit{Page 12, line 22. The “sink” term has been added to eq. 42, representing the discretised form of mixture equation. If the term is associated with a phase change, it would be better add it in the equation for a phase and not for the mixture with density $\rho$.}

First of all we would like to calrify that the purpose of showing Eq. (42) ~ (44) is to show that only slight modifications is needed in discretized governing equations when new physics is considered. In addition, we do not have separate equation for mass conservation of volcanic gas (due to the assumption that erupted material behaves like a single phase). Appologize if reviewer has already been aware of this.

We completely agree with the reviewer that for “sink” associated with phase change should be in the equation of the corresponding phase. For any future extending of current model, such as the example mentioned by the reviewer, considering phase change of volcanic gas, such sink term should be added to mass conservation equation of volcanic gas. The make the article more readable, we prefer to avoiding discussing new physics models, which is easily understandable for readers working on plume modeling but might cause confusion for readers without such background.
%To avoid confusing readers, we emphasized at the end of section 3.3:  “Eq. (42) - Eq. (44) show examples of adding more physics into SPH model based on Euler equations to show easy extensibility of SPH models. The mass source (and sink), drag force effect, and heat exchange effect are not actually added to our model (Eq. (1) - (4)) "
\\[3pt]

\textit {Page 12, lines 24-25. “ The drag force term should show up only when dynamics disequilibrium between different phases is considered”. In this case is the summation in the last term of eq. 43 only extended to particles of a different phase?}

The purpose of showing Eq. (43)  is to show that only slight modifications is needed in discretized governing equations when new physics (drag force term) is considered. The drag force term is not considered in current model, as we treat erupted material as one phase and assume immediate dynamic equilibrium between entrained air and erupted material. 
So dynamics disequilibrium between different phases (for example, solid particles of different size and volcanic gas) are currently not considered in the model.

If such disequilibrium need to be considered in future extension of current model, drag force effect should be considered. How many drag force effects should be considered depends on new assumptions and the new physics model. For example, if the physics model is a four phases model, one phase for atmosphere air, one phase for volcanic gas, one phase for large size solid particles and another for small size ash (maybe immediate equilibrium assumption between air and gas should be made). Drag effects between gas (air) and particles should be considered. But there should be no drag force term within gas phase itself. As for the small size ash, we can assume them behaves like single phase fluid (possibly not a reasonable assumption) then there should not drag force effect within the ash phase. Otherwise. drag effect within fine ash phase need to be considered.

We changed the original sentence “The drag force term should show up only when dynamics disequilibrium between different phases is considered” to “The drag force term should show up when dynamics disequilibrium between different phases is considered”. By deleting the word “only”, we change original necessary and sufficient statement to a sufficient statement. \\[6pt]

\textbf{Section 3.4}

\textit{Page 13, line 5. Please define “particle disorder”.}

We replaced “particle disorder” with “irregular distribution of particles” \\[6pt]

\textbf{Section 3.5}

\textit{Page 14, line 5. how sound speed of a particle is defined for the gas-solid mixture? Please add an equation}

It is calculated by $c=(gamma_m* \frac{p}{/rho•} )^0.5$, we added an equation in our manuscript.  \\[3pt]

\textit{Page 14, line 5. What is the order of magnitude of the time step for the tests presented, with a fully explicit scheme?}

I am supposing that you are asking about how much is each time step.

It depends on many factors: the value of smoothing length, particle mass, CFL number et al.. Time step also changes during simulation. For the simulation of Pinatubo eruption in this paper ($CFL=0.2$ was used, smoothing length is $170m$),  The time step is around 0.01s. \\[6pt]

\textbf{Section 3.6}

\textit{Page 14, line 8. “The classical SPH method was known to suffer from tensile instability and boundary deficiency.” Please describe there problems with some example, in order
to help the readers not expert in SPH.}

We added explanations and references on tensile instability: "Tests of the standard SPH method indicate an instability in the tensile regime, while the calculations are stable in compression.  A simplest example of such test calculation exhibiting the instability involves a body which is subjected to an uniform initial stress, either compressive or tensile. If the initial stress is tensile, a very small velocity perturbation on a single particle can lead to particles clumping together, forming large voids and seriously corrupting density distribution. But if the initial stress is compressive, the small velocity perturbation on a single particle can not lead to any changes in particle distribution. See paper by \citet{swegle1995smoothed} for more details." \\[3pt]

\textit{Page 14, line 14. “Equation (55) implies that $. . ..=1$”. Why? It is not equation 55 which implies the term is equal to 1, but the properties of the kernel. Furthermore, this is
true for the integral (not discretised) formulation, and only for points far away from the boundary (see Chen et al., 1999).}

The sentence is a misleading expression. Sorry for that. Here is the logic behind: The new approximation of function $A$, Eq. (55), should be consistent with original formulation, Eq. (28). Compare Eq. (55) and Eq. (28), if the denominator is 1, then these two equations will be the same. so it implies that the denominator in Eq. (55) should be close to 1. 
We completely agree with you that the normalization condition originally come from definition of kernel. By the statement "implies that $...=1$" we were actually trying to show kind of consistent connection between this new approximation formulation and property of kernel function. In other words, in the derivation based on Taylor series expansion, we never take the property of kernel into account but finally reach to an equation that consistent with one property of kernel function.

In addition, the summation (Eq. (28)) is usually a good approximation of integration in the area far away from boundaries. Hence the denominator in Eq. (55) is close to 1. it has ignorable effect on evaluating of $A_a$. That is to say, Eq. (55) and Eq. (28) are the same for particles far away from boundaries. But for particles close to boundaries, the denominator will definitely not be close to 1. For these cases, the non-unit denominator helps normalize the summation and overcomes boundary deficiency.

We changed the original sentence "Equation (55) implies that $\sum_b	m_b \rho_b w (x-x_b, h) = 1$ , which can be viewed as the approximation form of Eq. (34)." to  "Notice that the denominator in Eq. (55) is actually summation approximation of Eq. (34). That is to say, Eq. (55) and Eq. (28) are the same for particles far away from boundaries as the denominator in Eq. (55) becomes $1$ in that case." \\[6pt]

\textbf{Section 3.7}

\textit{Page 14, line 21-22. “Numerical simulation of multiphase flows is usually difficult due to the existence of complex evolving interfaces between phases.”
This is true when the different phases are immiscible. But in the application investigated, the volcanic plume, phases are not immiscible and mixing is very important. So, there is no need to track the interface between phases Eulerian grid-based numerical methods. Conversely, standard formulations of SPH cannot resolve fluid mixing and instabilities at flow boundaries. IT should be discussed in the paper if, and how, this is has been addressed in the model presented. Please take a look at: J. I. Read, T. Hayfield, O. Agertz; Resolving mixing in smoothed particle hydrodynamics, Monthly Notices of the Royal Astronomical Society, Volume 405, Issue 3, 1 July 2010, Pages 1513–1530}

As we have mentioned in our response to reviewer's previous question regarding interface capturing/tracking in SPH, interface construction will become necessary and important when we attempt to include the effects of mixing by resolving the detailed interface structure and dynamics of turbulence. This option is not attractive at current stage of plume modeling as turbulence closure can give an acceptable accuracy with affordable computational cost.

The following is our response to reviewer's concern on how did we handle the issue of classical SPH: "Classical SPH has problems correctly integrating fluid instabilities and mixing at boundaries". I will call it "mixing challenge" in later paragraph for short.

We were not aware of Read's paper previously. But we adopted several techniques (corrected formulation for tensile instability issue, smaller artificial viscosity coefficients, turbulence model with heat transfer considered) that probably have similar affect as the methods proposed in Read's paper. In Read's paper, one of his remedies of SPH is based on fixing the clumping instability which is actually tensile instability. We just adopted a different method (Chen's method) to handle this issue.

In addition, before we adopting the turbulence model, we got a "mushroom like" plume and almost no mixing happened (exactly due to the "mixing challenge"). The simulation results became much more realistic after adding turbulence model. So the turbulence model helped us to resolve sub-particle scale mixing. As has been mentioned, it is a common practice in CFD to adopt turbulence closure to resolve instabilities (which dominate the mixing process for shearing flow).

\citet{agertz2007fundamental} did a Kelvin-Helmholtz (KH) test showing that using smaller artificial viscosity coefficient can help get more mixing (Fig. 15 and discussion in Section 6.1 in his paper). In our model we also use a much smaller artificial viscosity (See the last paragraph in section 3.4, we use $\alpha =0.3$, while traditionally $\alpha$ is taken as 1, our original motivation was to avoid excessive artificial viscosity).

\citet{price2008modelling} believes that the mixing issue is due to the fact that entropy is discontinuous at the boundaries while density is continuous. He found that adding thermal conductivity at boundaries can improve mixing in SPH as thermal conductivity can smooth the entropy. The thermal conductivity (heat transfer) due to turbulence is considered in our model. What's more the thermal conductivity coefficient is much larger at the interfaces due to large shearing effect at the interface. Which can definitely help to releive the "mixing challenge" if Price is correct. Of course, we were not motivated by Price's paper to adding a turbulent heat transfer term into our model. It is natural to have a turbulent heat transfer term for compressible flow in which energy conservation equation is coupled with momentum conservation and mass conservation equation. As you can see we actually and luckly "adopted" Price's method.

\citet{wadsley2008treatment} and \citet{ritchie2001multiphase} made similar argument as Price. \citet{borgani2012hydrodynamic} believes that the "mixing challenge" of SPH is due to its poor ability of capturing contact discontinuity. They show that using GSPH can avoid appearance of spurious pressure force and help to to follow the kelvin instability.

Since people still have different opinions on the sources of "mixing issue", it would be interesting to see the connections between these different opinions. But such work is obviously out of our topic in this paper. These remedies of traditional SPH we adopted and extended in our model helped us to overcome, at least relieve the "mixing challenge".

At the end, we would thank reviewer for asking this critical question and pointing out other possible ways for remedying traditional SPH.\\[6pt]

\textbf{Section 3.8.1}

\textit{Page 16, line 10. For a reader like me who is more familiar with Eulerian and meshbased formulation, it would be good to state at the beginning that, differently from RANS (Reynolds Averaged Navier-Stokes) equations, which are time-averaged equations of motion, here the Lagrangian average is in space and not in time.}

Thank you for your suggestion for improving readability of the article. We totally agree with you and have made revision as you suggested. \\[6pt]

\textbf{Section 3.8.2}

\textit{Page 20, line 8. Is this equation the definition of $F_{ab}$ or a property? If this is the definition, where is the argument of the function $F$?}

It is the definition of $F_{ab}$. By the way, $F_{ab}(h)$ is short for $F(\textbf{x}_a -\textbf{x}_b, h)$, we added clarification in our revised manuscript. \\[6pt]

\textbf{Section 3.9.1 Wall boundary condition}

\textit{Please discuss the limitation of this approach when a complex topography is considered.
In particular, how could you deal with convex geometries?}

In our current model, the ground is assumed to be a flat ground. Since the topography is pretty simple, we did not experience any trouble. it has been shown in some other papers that it works as well for more complex topography \citep{kumar2013parallel}. One potential issue for complicated topography might be deployment of wall ghost particles. For flat boundary, we simply deploy several layers of ghost particles with equal interval and each ghost particle has the same mass. It has not been investigated yet how different deployment of wall ghost particle will influence the accuracy of imposed boundary conditions. A more basic question is how many degree of freedom should we allow for ghost particle deployment: only allows changing of particle position? Or allows using different particle mass size?

For non-collapse plume, the flat ground assumption made in our current model should be good enough. For other phenomena, such as PDCs, a more realistic topography is required. As we mentioned in the manuscripts, current plume models are still not comprehensive enough, each model has its own problem of interest and assumptions were made accordingly. Fortunately, we’ve been seeing efforts on making these models to be more general. As the first trial of using SPH in plume modeling, we prefer to narrowing down the problem of interesting and focus on more fundamental aspects.

We appreciate such discussion proposed by the reviewer which are constructive for future model improvement. \\[6pt]

\textbf{Section 3.9.3}

\textit{Keeping the pressure constant at atmospheric boundary can represent a problem when particles exit from the domain and choosing a larger computational domain cannot be the solution for long simulations. A better way to implement the boundary conditions at the exit would be to impose a condition on the total pressure ($p_{\rho}-0.5 \rho u^2$, where $p_{rho}=p-\rho g h$). This allow to have pressure changes at the boundaries associated with outflow.}

I have two questions regarding your suggestion: 
1) When we assign pressure value to these static pressure ghost particles, the decrease of pressure due to increase of height ($h$) has already been took into account. That is to say, pressure of our pressure ghost particles is consistent with meteorological data when establishing the initial condition. I am wondering should I still deduce $\rho g h$ from pressure?  
2) Should the total pressure equals to a pressure \textbf{plus} kinetic energy, or minus kinetic energy? I have similar question for definition of $p_{\rho}$. Should $\rho g h$ been added to or deduced from pressure? Because total pressure in other fields, such as thermal dynamic analysis of flowing process in engines are definied as pressure plus kinetic energy. Just to make sure.

In our current model, pressure ghost particles are assumed to be stationary. It does not make difference whether we use pressure or total pressure since velocity of pressure ghost particles are zero. The next important improvement of our current model under planning is taking the effect of wind field into account. In that case, the pressure ghost particles will not be static any more. The idea of using total pressure can be naturally implemented for that case. Thank you very much for pointing out an potential way for improving pressure outlet boundary condition. \\[6pt]

\textbf{Section 4. Verification and validation}

\textit{deals with the numerical resolution of the equations of the model, not with the agreement between the model and reality. It checks no code errors have been introduced
in the code and one way to do it is a comparison with analytical solutions. I think it is better to simply rename the section "Applications” or “Results”.}

Since JPUE is not the phenomena that we are targeted to model. So that is not validation (validation is to make sure assumptions in physics model is consistent with reality). Meanwhile, we compared our results against experimental results. We might call such comparison (simulation against measurements) as verification in a more general sense (?).

We added simulation results of shock tube tests which were compared against analytic solutions. By adding such real verification tests, we keep the title of this section unchanged.
\\[6pt]

\textbf{Section 4.1}

\textit{Page 25, line 24. Please look also at the results from this paper: Ezzamel, Adam, Pietro Salizzoni, and Gary R. Hunt. "Dynamical variability of axisymmetric
buoyant plumes." Journal of Fluid Mechanics 765 (2015): 576-611.}

Thank you for pointing out a source for accessing more recent experimental results. We’ve added the paper in the reference. \citet{ezzamel2015dynamical} did not provide coefficients of their fit equations. So we did not add plots corresponding to  Ezzamel's results in our figures. \\[3pt]

\textit{Page 26. Please put figure 6 and 7 together as to subplots
Page 26, line 12. The expression for the Gaussian profile should be written as an equation (on a single line and numbered) and the terms should be defined after it is introduced, before writing the expression for the straight line. Page 27. Please put figure 8 and 9 together as to subplots Page 28, line 1. Please write the straight line expression as an equation on a single line.}

We revised manuscripts as your suggestions for Fig. 6 - Fig. 9 \\[3pt]

\textit{Page 28, lines 8-9. “results, a small disparity in both velocity and concentration are observed near the boundary of the jet.” Looking at figure 10, it seems that near the boundary of the jet there are regions without particles, and this could be the reason of the differences with experimental results. The emptying of particles in region with steep density gradients is described in section 3.4 of this paper: Ritchie, Benedict W., and Peter A. Thomas. "Multiphase smoothed-particle hydrodynamics."
Monthly Notices of the Royal Astronomical Society 323.3 (2001): 743-756. Look for example at their figure 9.}

Thank you very much for pointing this out. The low particle density (or the emptying of particles as what you call it) near the vent should be at least one reason of the disparity. We added an explanation of the disparity based on that paper.

Ritchie’s paper proposed two remedies of standard SPH: 1) Keep a density-weighted quantity constant. This might prevent the smoothing length changing in a proper way. One potential issue with such method is that there is no guarantee on enough number of neighbor particles when cavity forms in the fluid field. Probably his method could help us cure the disparity but might also introduce other problems. 2) The second remedy he proposed is assuming a constant pressure within the kernel (not sure whether such assumption can reflect real physics or not). The density is then updated based on equation of state. One good property of density updating equation in standard SPH that adopted in our model is that the total mass is strictly conserved. All other alternative methods for density updating, for example, density updating based on mass conservation PDE, are not able to be not able to guarantee strict conservation of total mass. I did not see any investigation in that paper about how well does his method conserve mass.

Quite a lot variations of SPH has been proposed since its born. Each has advantages and disadvantages. Making a proper trade-off is one of the challenging thing for implementing SPH in simulating of complicated phenomena. And this is actually where we spent a lot of effort. Our choice were made either based on our preference (for example, we prefer strict conservation of mass, momentum and energy) or relevance of test simulations shown in thoese papers.

We thank the reviewer for proposing such discussion. \\[3pt]

\textit{Page 28. End of section. The paper presents also details about the numerical implementation and the parallelisation of the code, so it would be interesting to have more info for this test about computational cost, number of particles, core/cpu used, …}

The performance benchmark results are reported in 3.10. More Details has been published in another paper \citep{cao2017data}. We added the reference in section 3.10.

The performace bechmarking should be done exclusively while computational nodes are usually shared among users when doing daily simulation. The difference between exclusive testing and no-exclusive testing might be large. That's why we put performace testing in a separate section and did not mention any thing, like total simulation time, in this section.
\\[6pt]

\textbf{Section 4.2.1}

\textit{ Page 29, lines 7-8. From Costa et al. 2016 "For the erupted particles, only two size classes were considered, representing coarse ash ($\Phi c$) and fine ash ($\Phi f$), each comprising 50 wt. $\&$ of the erupted particles " Here I cannot find any info about particles ...}

We treat all erupted material including particles of different size and volcanic gas as a single phase. So our model is not able to make use of such detailed information. In our model, only the mass fraction of total solid particles is used, and dose not treat particles of different size separately. That's why we did not mention anything about detailed particle size distribution in the input parameter.

Details about assumptions on which our model relies can be found in section 2. These assumptions are also briefly summarized in abstract. To avoid confusing readers, we added emphasize on which portion of Costa’s data are used in our simulation. \\[3pt]

\textit{Page 29, line 13. In Fig. 1B it is plotted the meteo profile where it is also shown the presence of wind. Please clarify.}

We use a static pressure boundary condition (pressure ghost particles have zero velocity) and do not consider the effect of wind field. Even though wind field data and specific humidity are also in that Fig. 1B in \citet{costa2016results}, we only used density, temperature and pressure but did not use wind velocity and specific humidity in our simulation. 
We added clarifications on this in the revised manuscripts. \\[3pt]

\textit{Page 30. Why there are no figures showing a 2d section of the plume, as done for the previous test? It would be interesting to see a vertical section of some variables (for example instantaneous value of mixture velocity modulus) and velocity streamlines, as shown in Cerminara et al.}

Visualization of the simulated plume is added.

\textit{Page 30, line 1. The averaging technique of Cerminara et al. is defined for grid-based eulerian models. It would be interesting to know the details of the implementation of the technique for the SPH code (perhaps in an appendix).}
\textit{Page 30, lines 1-4. “ As particles distribute in a disordered manner in the space in SPH simulation results. We first project simulation results (on disordered particles) onto a pre-defined grid before doing time average and spatial integration. The project method is the basic SPH kernel based interpolation.” These lines are confused, please check.}

For these two questions above. We added an appendix describing the posting process in detail. Then we replaced the sentence “ As particles distribute in a disordered manner in the space in SPH
simulation results. We first project simulation results (on disordered particles) onto a pre-defined grid before doing time average and spatial integration. The project method is the basic SPH kernel based interpolation.”  with “As particles distribute irregularly in the space in SPH simulation results. We need to project simulation results (on irregular particles) onto a pre-defined grid before doing time average and spatial integration. See appendix A for more details of post processing."\\[3pt]

\textit{Page 30, line 11. Eq. 95 is referenced here, so it should be written immediately after this sentence.}

Yes, we adjusted the position of the equation and other two equations.\\[3pt]

\textit{Page 32, end of section. In this section it would be good to have more details on the simulation (number of particles, computational cost, number of cores/cpu), and also to make a comparison with the other models regarding the time needed for a simulation.}

%The computational efficiency of our current model is OK for plume simulation but too slow for multiple scale problem simulation, such as simulating of plume raising up combined with umbrellas spreading of volcanic ash in one run. The length scale and time scale of umbrella spreading of volcanic ash is much larger than the raising up of volcanic plume. If the spatial resolution can be adjusted adaptively, the computational efficiency could be increased a lot. In addition, if different time step can be used for different stages (the raising up stage and the spreading stage), the computational efficiency would be increased further. However, as far as I can see, such method would cause some challenges for parallelism. The dispersion and dissipation of volcanic ash in a regional (or even global) atmosphere has another level of length scale and time scale, and hence are not suitable for current model. All these challenges need to be addressed with more careful and solid work.

Performance bechmarking of our solver is repoorted in section 3.10. More details has been published in a separate paper. We added an reference to paper.  As the simulation of Pinatubo reported here was done on computational cluster on CCR in a " \textbf{no-exclusive}" manner, the computational performance given by statement like "spend XXX hours on XXX cpus" might be misleading.

To give an intuitive impression on the "computational performance" of our solver: it took around 12 days on 240 cores for the Pinatubo simulation. The smoothing length is 170m, $CLF=0.2$, the total duration of simulation is 550 s after eruption. The size of computational domain is $[-40800, 40800] \times [-40800, 40800] \times [1500, 55000] $;
%Even though, we added a paragraph briefly describing the total simulation time that we spend on this volcanic plume simulation.

A comparison between all existing 3D models in terms of computational cost would be interesting. 
considering many factors could influence the total simulation time, such as resolution, size of the computational domain, duration of simulation, type of hardwares on which the simulation is done, number of CPUs, et al, more careful benchmark testing are required for doing good comparison.
\\[12pt]

%Another question is: for models adopting higher order methods, they can get as good solution as these model using lower order method, with a lower resolution. It is "not fair" to ask model using higher order method to have the same resolution as models using lower order method.

\textbf{\large Technical corrections}

\textit{See pdf attached. Please also note the supplement to this comment: https://www.geosci-model-dev-discuss.net/gmd-2017-119/gmd-2017-119-RC1-supplement.pdf}

Other revisions related to languages and typos have been made according to reviewer's suggestions in the supplement PDF that he attached.

The only one suggestion that we did not follow is his suggestion for replacing "heat" with "positive buoyancy". on page 6 line 24. 
The reason for using "heat" is: Since the erupted material are hot, they keep heating up entrained air until its heat exhausts. And due to such heating up process, the plume can keep expanding and reducing its bulk density and keep a positive buoyancy. 
So we prefer to use "heat" here. We are definitely open and happy to discuss with reviewers about which word is more proper.

We thanks him again for his careful reading and many constructive comments.\\[12pt]

\textbf{\large Major adjustments of the manuscripts}

Here is a summary on major adjustments made in the revised manuscript: 
\begin{itemize}
\item In section 3, moved contents in section 3.2 right behind Eq. (26) and merged section 3.1 and 3.2  into one section. Moved section 3.4 before 3.3. Changed the tile of section 3.7 and adjusted the content in that section.
\item Added a new section in section 4 showing some 1D shock tube tests against analytical results.
\item In section 4, added 4 figures showing mass fraction and velocity of simulated volcanic plume.
\item Added an Appendix on post processing of SPH simulation results.
\end{itemize}

\bibliographystyle{plainnat}
\bibliography{Reference}
\end{document}

### 2017-09-21 Manuscripts_GMD_ap_ref1.tex
9/20 edits ap

@@
 While these 1D models can generate well-matched results with 3D (three dimensional) models for weak plumes, much greater variability is observed for strong plume scenarios, especially for local variables \citep{costa2016results}.
-In addition, there is need for greater skills in hazards forecasts especially where the plume model is used to generate source conditions for complex 2D (two dimensional) and 3D VATD models.
+In addition, there is need for greater skill in hazards forecasts especially where the plume model is used to generate source conditions for complex 2D (two dimensional) and 3D VATD models.
 
@@
  %Different simplification was made based on their scope. Improvements have been making since their birth. 
- Accuracy of simulation (depending on comprehensiveness of the model, resolution of discretization, numerical error, and order of accuracy) and simulation time (depending on number of governing equations, resolution, numerical methods and parallel techniques) are always a pair of conflicting considerations in 3D plume simulations.
+ Accuracy of simulation (depending on comprehensiveness of the model, resolution of discretization, numerical error, and order of accuracy) and simulation time (depending on number of governing equations, resolution, numerical methods and parallel techniques) are always conflicting considerations in 3D plume simulations.
 
@@
 Among such Lagrangian methods, smoothed particle hydrodynamics \citep{gingold1977smoothed,lucy1977numerical} based simulations have shown good agreement with experiments for many applications in fluid dynamics.
-And it currently is, by far, the most widely used mesh-free scheme.
+Currently it is, by far, the most widely used mesh-free scheme.
 Our implementations of SPH in volcanology follows on earlier efforts   \citep{bursik2003smoothed,herault2010sph,haddad2016smoothed}.
@@
 As discussed in the previous paragraph, existing 3D plume models focus on one or several specific aspects of plume and have been extended to be more comprehensive by accounting for more physics or more phases.
-Easy extensibility and capability of handling multiples phase flow without additional numerical effort greatly facilitate future extension of SPH models.
+Easy extensibility and capability of handling multiple phase flow without additional numerical effort greatly facilitate future extension of SPH models.
 As volcanic plumes are in nature multiphase and without pre-defined boundary in the atmosphere, SPH is a suitable numerical method for plume modeling.
@@
 \item We develop methodology to impose pressure boundary conditions by adding extra layers of static ghost particles.
-Additional constraints on the time step is used to avoid the growth of numerical fluctuations near the pressure boundary.
+Additional constraints on the time step are used to avoid the growth of numerical fluctuations near the pressure boundary.
 We impose a velocity inlet boundary condition by placing several layers of ghost particles moving with eruption velocity.
@@
 We adopt a LANS (Lagrangian Averaged Navier-Stokes) turbulence model \citep{monaghan2011turbulence}, which was originally proposed for incompressible flow, and is extended here for compressible flow accounting for turbulent heat exchange.
-\item Corrected formulation of SPH \citep{chen1999improvement} is adapted to bypass the well-known tensile instability issues of classical SPH.
-\item Simulation of volcanic plumes with acceptable accuracy requires fine resolution (very high particle counts) that cannot be accomplished without parallel computing.
-The core solver of our model is parallelized by MPI (message passing interface standard) parallelism.
+\item Corrected formulation of SPH \citep{chen1999improvement} is adopted to bypass the well-known tensile instability issues of classical SPH.
+\item Simulation of volcanic plumes with acceptable accuracy requires fine resolution (very high particle counts) that cannot be accomplished without parallel computing large process counts.
+The core solver of our model is parallelized by distributed memory MPI (message passing interface standard) parallelism.
 In addition, a dynamic load balancing strategy is also developed.
@@
 To summarize, our model is not valid for eruptions where wind effect plays a significant role in its development, usually refered as a weak plume.
-Our model also lacks the ability in modeling plumes with large particles or eruptions in which microphysics plays non-ignorable roles, such as a eruption of \textit{El Chich{\'o}n} volcano on April 4th 1982 \citep{sigurdsson19841982, folch2016fplume}.
+Our model also lacks the ability in modeling plumes with large particles or eruptions in which microphysics plays non-ignorable roles, such as an eruption of \textit{El Chich{\'o}n} volcano on April 4th 1982 \citep{sigurdsson19841982, folch2016fplume}.
 We are focussed here on developing the SPH based methodoloy in the context of the more basic (and more critical) aspects of volcanic plume and therefore devote our effort to this relative simpler model.
@@
 \begin{equation}
-p = p_a\left(z\right)\label{eq:pressure_bc_p} 
+p = p_a\left(z\right)  \label{eq:pressure_bc_p} 
 \end{equation}
@@
 
-\subsection{Artificial viscosity} \label{sec:artificial-viscosity}
-In classical SPH, shock waves are handled by introducing artificial viscosity, a term that defined based on second derivatives of velocity, to smear out discontinuities.
+\subsection{Artificial Viscosity} \label{sec:artificial-viscosity}
+In classical SPH, shock waves are handled by introducing artificial viscosity, a term that is defined based on second derivatives of velocity, to smear out discontinuities.
 As in the case of first order derivatives, second order derivatives can be estimated by differentiating a SPH interpolation twice.
@@
 The SPH viscosity can be related to a continuum viscosity by converting the summation to integrals \citep{monaghan2005smoothed}.
-It has been shown that the shear viscosity coefficient $\eta = \frac{\rho \alpha h c}{8} $ and the bulk viscosity coefficient $ \zeta = \frac{5 \eta}{3}$ for two dimensional and $\eta = \frac{\rho \alpha h c}{10} $ , $ \zeta = \frac{5 \eta}{3}$ for three dimensional.
+It has been shown that  shear viscosity coefficient $\eta = \frac{\rho \alpha h c}{8} $ and bulk viscosity coefficient $ \zeta = \frac{5 \eta}{3}$ are appropriate for two dimensional flows and modified to $\eta = \frac{\rho \alpha h c}{10} $ , $ \zeta = \frac{5 \eta}{3}$ for three dimensional flows.
 An extra term was added to $\nu$ considering aspects of the dissipative term in shock solutions based on Riemann solvers and lead to a new formulation of artificial viscosity.
@@
 Other physics can be added easily in a similar way.
-Adding of these new terms leads to modification of several lines in the source code.
-The drag force term should show up when dynamics disequilibrium between different phases is considered.
+Adding of these new terms leads to modification of only a few lines in the source code.
+The drag force term should show up when dynamic disequilibrium between different phases is considered.
 In that case, each phase needs one set of governing equations of Navier-Stokes type.
@@
 Tests of the standard SPH method indicate an instability in the tensile regime, while the calculations are stable in compression.
-A simplest example of such test calculation exhibiting the instability involves a body which is subjected to an uniform initial stress, either compressive or tensile.
+A simple example of such a test calculation exhibiting the instability involves a body which is subjected to an uniform initial stress, either compressive or tensile.
 If the initial stress is tensile, a very small velocity perturbation on a single particle can lead to particles clumping together, forming large voids and seriously corrupting density distribution.
@@
 (\ref{eq:wall_bc_e})).
-In our current model, the ground is assumed to be a flat ground.
-For more complicated topography, it has been shown in other paper \citep {kumar2013parallel} that it works as well.
-One of our concern regarding potential limitation of such method is deployment of ghost particles for more complicated boundaries which is not described in detail in the paper.
+In our current model, the ground is assumed to be  flat.
+For more complicated topography, it has been shown in other work \citep {kumar2013parallel} that it also works as well.
+We do have some concern regarding potential limitation of this method of deployment of ghost particles for more complicated boundaries in three dimensions but that is one that is not related to the plume simulations that are the target here.
+
 \subsubsection{Eruption boundary condition}
@@
 The density, momentum, and internal energy of these eruption ghost particles are not updated before they move above ground.
-As soon as they move out from erupt conduit, these ghost particles will be shifted to real particles and their physical quantities and position will be updated based on discretized governing equations.
+As soon as they move out from eruption conduit, these ghost particles will be shifted to real particles and their physical quantities and position will be updated based on discretized governing equations.
 New ghost particles need to be added at the bottom of the eruption conduit as these existing ghost particles move upwards.
@@
 The physical quantities for pressure ghost particles are not updated while these for real particles are updated at every time step.
-As position of all pressure ghost particles keep constant, we essentially impose a static pressure boundary condition.
-Real particles are removed as soon as they move out the pressure boundary.
+As position of all pressure ghost particles keeps constant, we essentially impose a static pressure boundary condition.
+Real particles are removed as soon as they move out of the pressure boundary.
 
@@
 Too small $CFL_p$ would slow down simulation while too large $CFL_p$ would lose its ability of mitigate numerical fluctuation near the boundary.
-The proper $CFL_p$ is determined by series of simulation tests.
+The proper $CFL_p$ is determined by a series of simulation tests.
 
@@
 High performance computing is therefore essential.
-Among existing CPU parallel SPH schemes, most of them focus on neighbor searching algorithm and dynamic load balancing \citep[eg.
-][] {ferrari2009new, crespo2015dualsphysics}.
+Among existing CPU parallel SPH schemes, most of them focus on neighbor search algorithm and dynamic load balancing \citep   {ferrari2009new, crespo2015dualsphysics}.
 Less attention has been paid to developing more flexible data management schemes for more complicated problems.
@@
 Then a simulation of representative strong volcanic plume is conducted.
-Intergrated local variable are comparable with simulation results from existing 3D plume models.
+Integrated local variable are comparable with simulation results from existing 3D plume models.
 \subsection{1D Shock Tube Tests}
@@
 
-In the table \ref{tab:1D-shock-input_parameters},  subscript $L$ refers left side and $R$ for right side, $t_f$ is the total simulation time.
+In  table \ref{tab:1D-shock-input_parameters},  subscript $L$ refers left side and $R$ for right side, $t_f$ is the total simulation time.
 The initial interval between two adjcent particles is 0.03.
@@
 These factors prevent us from numerically reproducing these experiments in an exact way as they were conducted.
-However, the features of JPUE is correctly reproduced with our code.
+However, the features of JPUE are correctly reproduced with our code.
 
@@
 Material properties, shown in Table \ref{tab:material_properties}, are selected based on properties of Pinatubo and Shinmoe-dake eruption.
-Other material properties not given in the table can be infered from these given parameters based on their relationships.
+Other material properties not given in the table can be inferred from these given parameters based on their relationships.
 
@@
 Fig.
-\ref{{fig:pinatubo-simulation-results-vis}} shows the mass fraction of simulated volcanic plume at 300s after eruption, at which time the plume starts spreading radially.
+\ref{fig:pinatubo-simulation-results-vis} shows the mass fraction of simulated volcanic plume at 300s after eruption, at which time the plume starts spreading radially.
 A contour plot of the mass fraction on the vertical cross-section (y-z plane) were also provided.
@@
 Figures at the second row show contour of mass fraction and velocity quiver on y-z plane.
-These figures are plot utilizing post-processed data (see Appendix \ref{app:project-SPH-grid}).
+These figures are plotted utilizing post-processed data (see Appendix \ref{app:project-SPH-grid}).
 The contour level in the plot are $0.00001, 0.0001, 0.001, 0.1, 0.3, 0.75, 0.95$.
@@
 
-PDAC, which treat particles of two different sizes as two separate phases, predicted a similar mass fraction profile.
+PDAC, which treats particles of two different sizes as two separate phases, predicted a similar mass fraction profile.
 That implies that assumption of dynamic equilibrium in our model is at least valid for eruptions similar to the test case.
@@
 One of the big differences between these two models is that we adopt a LANS type of turbulence model while SK-3D adopts a LES (large eddy simulation) turbulence model.
-This implies that turbulence model might play a critical role in plume simulation.
+This implies that choice of  turbulence model might play a critical role in plume simulation.
 
@@
 Top height and integrated local variables simulated by our model are consistent with simulation results of existing 3D plume models.
-Comparison between our results with these of SK-3D implies that turbulence model plays a significant role in plume modeling.
+Comparison between our results with those of SK-3D implies that turbulence model plays a significant role in plume modeling.
 
@@
 \appendix
-\section{Post process of particle data} \label{app:project-SPH-grid}    %% Appendix A
+\section{Post processing of particle data} \label{app:project-SPH-grid}    %% Appendix A
 \appendixfigures
@@
 (\ref{eq:SPH-fundamental-principle})).}
-    \label{fig:1D-shock-tests-verification}
+    \label{appfig:1D-shock-tests-verification}
 \end{figure}
@@
 Particles distribute irregularly in SPH simulation results.
-To adopt post process originally proposed for mesh-based method, we need to project simulation results onto a pre-defined regular mesh.
+To adapt post processing originally proposed for mesh-based method, we need to project simulation results onto a pre-defined regular mesh.
 Fig.
-\ref{fig:1D-shock-tests-verification} shows the procedure for posting process.
+\ref{appfig:1D-shock-tests-verification} shows the procedure for posting process.
 
@@
 This project is supported by Grants No.
-NSF 1131074 from the National Science Foundation.
+NSF ACI/1131074 from the National Science Foundation.
 \end{acknowledgements}

### 2017-09-22 Manuscripts_GMD_ap_ref1.tex
done proof reading

@@
 The accuracy of these terms is crucial for forecasts from VATDs and the 3D SPH model presented here will provide better numerical accuracy.
-As an initial effort to exploit the feasibility and advantages of SPH in volcanic plume modeling, we adopt a relatively simple physics model (3D dusty-gas dynamic model assuming well mixed eruption material, dynamic equilibrium and thermodynamic equilibrium between erupted material and air that entrained into the plume and minimal effect of winds) targeted at capturing the salient features of a volcanic plume.
+As an initial effort to exploit the feasibility and advantages of SPH in volcanic plume modeling, we adopt a relatively simple physics model (3D dusty-gas dynamic model assuming well mixed eruption material, dynamic equilibrium and thermodynamic equilibrium between erupted material and air that entrained into the plume, and minimal effect of winds) targeted at capturing the salient features of a volcanic plume.
 The documented open source code is easily obtained and extended to incorporate other models of physics of interest to the large community of researchers investigating multiphase free boundary flows of volcanic or other origins.
@@
 Our results are consistent with existing 3D plume models.
-Analysis of the plume evolution process illustrates that this model is able to reproduce the physics of plume development.
+Analysis of the plume evolution process demonstrates that this model is able to reproduce the physics of plume development.
 
@@
 The dynamic core of ATHAM solves the compressible Euler equations for momentum, pressure and temperature of the gas particle mixture \citep{oberhuber1998volcanic}.
-The subgrid-scale turbulence closure scheme differentiates between the horizontal and vertical directions \citep{herzog2003prognostic} captures turbulent mixing.
+The subgrid-scale turbulence closure scheme that differentiates between the horizontal and vertical directions \citep{herzog2003prognostic} is adopted to capture turbulent mixing.
 The cloud microphysics predicts the mass of hydrometeors in liquid and ice phase \citep{herzog1998effect}.
@@
 As discussed in the previous paragraph, existing 3D plume models focus on one or several specific aspects of plume and have been extended to be more comprehensive by accounting for more physics or more phases.
-Easy extensibility and capability of handling multiple phase flow without additional numerical effort greatly facilitate future extension of SPH models.
+Easy extensibility and capability of handling multiple phase flow with less additional numerical effort greatly facilitate future extension of SPH models.
 As volcanic plumes are in nature multiphase and without pre-defined boundary in the atmosphere, SPH is a suitable numerical method for plume modeling.
 The core physics, such as entrainment of air and thermal expansion, are essential for all plume modeling while some other physics, such as water condensation and aggregation etc., are important in specific scenarios.
-As an initial effort on exploiting advantages of SPH in volcanic plume modeling, we focus on capturing basic features in plume development using a robust numerically and computationally efficient framework with support for scalable parallel computing.
+As an initial effort on exploiting advantages of SPH in volcanic plume modeling, we focus on capturing basic features in plume development using a numerically robust and computationally efficient framework with support for scalable parallel computing.
 %So our current model is far away from comprehensive.
@@
 \item Corrected formulation of SPH \citep{chen1999improvement} is adopted to bypass the well-known tensile instability issues of classical SPH.
-\item Simulation of volcanic plumes with acceptable accuracy requires fine resolution (very high particle counts) that cannot be accomplished without parallel computing large process counts.
+\item Simulation of volcanic plumes with acceptable accuracy requires fine resolution (very high particle counts) that cannot be accomplished without parallel computing using large process counts.
 The core solver of our model is parallelized by distributed memory MPI (message passing interface standard) parallelism.
@@
 In the current model the initial domain is a box.
-The boundaries are categorized as the velocity inlet (a circular area at the center of the bottom of the box), wall boundary (box bottom), pressure outlet (other faces of the box).
+The boundaries are categorized as the velocity inlet (a circular area at the center of the bottom of the box), wall boundary (box bottom) and pressure outlet (other faces of the box).
 %probably should add a section to list the input parameters in the verification section
@@
 \end{equation}
-The Dirac function in Eq.
+The Dirac function $\delta$ in Eq.
 (\ref{eq:Dirac-translation}) can be approximated by a weighting function $w\left(\textbf{x}-\textbf{x}\prime, h\right)$ (or $w\left(\textbf{x}\prime-\textbf{x}, h\right)$) which tends to a Dirac function when the smoothing length $h \rightarrow 0$ :
@@
 \end{equation}
-So it can be viewed as an approximate form of the Dirac function and should satisfy the normalization condition:
+The weighting function, as an approximate form of the Dirac function, should satisfy the normalization condition:
 \begin{equation}
@@
 The SPH viscosity can be related to a continuum viscosity by converting the summation to integrals \citep{monaghan2005smoothed}.
-It has been shown that  shear viscosity coefficient $\eta = \frac{\rho \alpha h c}{8} $ and bulk viscosity coefficient $ \zeta = \frac{5 \eta}{3}$ are appropriate for two dimensional flows and modified to $\eta = \frac{\rho \alpha h c}{10} $ , $ \zeta = \frac{5 \eta}{3}$ for three dimensional flows.
+It has been shown that  shear viscosity coefficient $\eta = \frac{\rho \alpha h c}{8} $ and bulk viscosity coefficient $ \zeta = \frac{5 \eta}{3}$ are appropriate for two dimensional flows.
+$\eta = \frac{\rho \alpha h c}{10} $ and $ \zeta = \frac{5 \eta}{3}$ are appropriate for three dimensional flows.
 An extra term was added to $\nu$ considering aspects of the dissipative term in shock solutions based on Riemann solvers and lead to a new formulation of artificial viscosity.
@@
 \end{align}
-where, $a$ is the SPH particle index, $\textbf{v}_{a b} = \textbf{v}_a - \textbf{v}_b$.
+where, $a$ is the SPH particle index.
 $\Pi$ is an artificial viscosity term, which is discussed in section \ref{sec:artificial-viscosity}.
@@
 where $c_a$ is sound speed at particle $a$ and calculated based on heat specific ration of the mixture $\gamma_m$ (See Eq. (\ref{eq:sound-speed-mixture})). 
-$d$ is number of dimensions. 
 \begin{equation}
@@
 Tests of the standard SPH method indicate an instability in the tensile regime, while the calculations are stable in compression.
-A simple example of such a test calculation exhibiting the instability involves a body which is subjected to an uniform initial stress, either compressive or tensile.
+A simple example of such a test calculation exhibiting the instability involves a body which is subject to an uniform initial stress, either compressive or tensile.
 If the initial stress is tensile, a very small velocity perturbation on a single particle can lead to particles clumping together, forming large voids and seriously corrupting density distribution.
 But if the initial stress is compressive, the small velocity perturbation on a single particle can not lead to any changes in particle distribution \citep{swegle1995smoothed}.
-To address these difficulties, \citep{chen1999improvement} proposed a corrected SPH formulation.
+To address these difficulties, \citet{chen1999improvement} proposed a corrected SPH formulation.
 For 1D case, employing a Taylor expansion for $A\left(x\right)$ about $x_a$, multiplying both sides by kernel function and then doing integration over the domain gives
@@
 \begin{equation}
-A_a = \frac{\sum_b m_b \dfrac{A_b}{\rho_b} w\left(x_a-x_b, h\right)}{\sum_b m_b \dfrac{1}{\rho_b} w\left(x_a-x_b, h\right)}
+A_a = \frac{\sum_b m_b \dfrac{A_b}{\rho_b} w_{ab}\left(h\right)}{\sum_b m_b \dfrac{1}{\rho_b} w_{ab}\left(h\right)}
 \label{eq:CSP-function-approximation-1d}
@@
 Notice that the denominator in Eq.
-(\ref{eq:CSP-function-approximation-1d}) is actually summation approximation of Eq.
+(\ref{eq:CSP-function-approximation-1d}) is actually summation approximation of the left side in Eq.
 (\ref{eq:SPH-kernel-normalization-prop}).
@@
 \begin{equation}
-\nabla A_a = \frac{\sum_b m_b \dfrac{A_b - A_a}{\rho_b} \nabla_a w\left(x_a-x_b, h\right)}{\sum_b m_b \dfrac{x_b - x_a}{\rho_b} \nabla_a w\left(x_a-x_b, h\right)}
+\nabla A_a = \frac{\sum_b m_b \dfrac{A_b - A_a}{\rho_b} \nabla_a w_{ab}\left(h\right)}{\sum_b m_b \dfrac{x_b - x_a}{\rho_b} \nabla_a w_{ab}\left( h\right)}
 \end{equation}
@@
 (\ref{eq:ns-sph-d}) in spirit.
-Particles of phase 1 are not be counted while evaluating density of phase 2 and vice versa.
+Particles of phase 1 are not counted while evaluating density of phase 2 and vice versa.
 Updating of density is then based on the following discretized equations.
@@
 \begin{equation}
-<\rho_\alpha^{sg}>=\frac{\sum_j m_j w_{\alpha j} \left(h\right)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} \left(h\right) +\sum \frac{m_j}{\rho_j} w_{\alpha j} \left(h\right)} \label{eq:gov-sph-d2}
+<\rho_\alpha^{sg}>=\frac{\sum m_j w_{\alpha j} \left(h\right)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} \left(h\right) +\sum \frac{m_j}{\rho_j} w_{\alpha j} \left(h\right)} \label{eq:gov-sph-d2}
 \end{equation}
@@
 \begin{equation}
-\rho_{\alpha}=\rho_{\alpha}^a=\frac{\sum m_b w_{\alpha b} \left(h\right)}{\sum \frac{m_b}{\rho_b} w_{\alpha b}}
+\rho_{\alpha}=\rho_{\alpha}^a=\frac{\sum m_b w_{\alpha b} \left(h\right)}{\sum \frac{m_b}{\rho_b} w_{\alpha b} \left(h\right)}
 \end{equation}
@@
 \begin{equation}
-\rho_{\alpha}=\rho_{\alpha}^{sg}=\frac{\sum m_j w_{\alpha j} \left(h\right)}{\sum \frac{m_j}{\rho_j} w_{\alpha j}}
+\rho_{\alpha}=\rho_{\alpha}^{sg}=\frac{\sum m_j w_{\alpha j} \left(h\right)}{\sum \frac{m_j}{\rho_j} w_{\alpha j} \left(h\right)}
 \end{equation}
@@
 (\ref{eq:gov-sph-d1}) and Eq.
-(\ref{eq:gov-sph-d1}) without any other interface track or capture method.}
+(\ref{eq:gov-sph-d2}) without any other interface track or capture method.}
 \label{fig:SPH-multiple-density}
@@
 \end{equation}
-where $K_{ab} = l^d G_{ab}$, $M = \rho_0 l^d$ in which $d$ is the dimension and $\rho_0$ is initial density. $SPH-\varepsilon$ turbulence model is obtained after lengthy derivation:
+where $K_{ab} = l^d G_{ab}$, $M = \rho_0 l^d$ in which $\rho_0$ is initial density. $SPH-\varepsilon$ turbulence model is obtained after lengthy derivation:
 \begin{equation}
@@
 \begin{equation}
-F_{ab}(h) \textbf{x}_{ab} = \nabla _a w_{ab}
+F_{ab}(h) \textbf{x}_{ab} = \nabla_a w_{ab} \left( h \right)
 \end{equation}
@@
 
-More details about the data structure, domain decomposition, load balancing, domain adjusting and performance testing has been published separately \citep{cao2017data}.
+More details about the data structure, domain decomposition, load balancing, domain adjusting and performance benchmarking has been published separately \citep{cao2017data}.
 
@@
 This implies that adjusting of domain can avoid computing large number of uninfluenced air particles, especially for the beginning stage of simulation.
-A domain adjusting algorithm is designed and implemented in our code.
+A domain adjusting algorithm \citep{cao2017data} is designed and implemented in our code.
 Figure \ref{fig:adj_vs_no} shows that simulation time of the test problem is greatly reduced when we adopt the domain adjusting strategy.
@@
 where $c$ is the speed of sound in the liquid.
-The energy equation is actually decoupled from the momentum conservation equation and the mass conservation equation by using this equation of state (EOS).
+The energy equation is actually decoupled from the momentum conservation equation and the mass conservation equation by using this EOS.
 In addition, the "atmosphere" is assumed to be uniform and gravity is set to be zero.
@@
 
-One overall feature of JPUE is "self-similarity" which means that the evolution of the JPUE is determined solely by the local scale of length and velocity which theoretically account for the fact that the rate of entrainment at the edge of JPUE is proportional to a characteristic velocity at each height.
+One overall feature of JPUE is "self-similarity" which means that the evolution of the JPUE is determined solely by the local scale of length and velocity, which theoretically account for the fact that the rate of entrainment at the edge of JPUE is proportional to a characteristic velocity at each height.
 As a results, physical and numerical experiments do not necessary to have exactly the same setups and are compared on a non-dimensional basis.
@@
 Vertical distribution of temperature, pressure and density is used to establish stratified atmosphere.
-Wind velocity and specific humidity are not used in our simulation even though they were also provided in Fig.
-1B \citep{costa2016results}.
+Wind velocity and specific humidity are not used in our simulation even though they were also provided by \citet{costa2016results} (see Fig.
+1B).
 Material properties, shown in Table \ref{tab:material_properties}, are selected based on properties of Pinatubo and Shinmoe-dake eruption.
@@
 
-Fig.
-\ref{fig:pinatubo-simulation-results-vis} shows the mass fraction of simulated volcanic plume at 300s after eruption, at which time the plume starts spreading radially.
+Figure \ref{fig:pinatubo-simulation-results-vis} shows the mass fraction of simulated volcanic plume at 300s after eruption, at which time the plume starts spreading radially.
 A contour plot of the mass fraction on the vertical cross-section (y-z plane) were also provided.
@@
 Our model adopts the same assumptions and governing equations as SK-3D.
-However, there is still a obvious disparity between the profiles of local variables of our model and SK-3D.
+However, there is still an obvious disparity between the profiles of local variables of our model and SK-3D.
 One of the big differences between these two models is that we adopt a LANS type of turbulence model while SK-3D adopts a LES (large eddy simulation) turbulence model.
@@
 Advanced numerical techniques in SPH were exploited and tailored for this model.
-High performance computing was used to mitigate the tradeoff between accuracy (depends on comprehensiveness of the model, resolution, order of accuracy of numerical methods, scheme for time upgrading.) and simulation time (depends on comprehensiveness of model, resolution, order of accuracy of numerical methods, scheme for time upgrading ...
+High performance computing was used to mitigate the tradeoff between accuracy (depends on comprehensiveness of the model, resolution, order of accuracy of numerical methods, scheme for time upgrading) and simulation time (depends on comprehensiveness of model, resolution, order of accuracy of numerical methods, scheme for time upgrading ...
 and computational techniques).
 The correctness of the code and model was verified and validated by a series of test simulations.
+Typical 1D shock tube problems were simulated and compared against analytical results showing good agreement.
 Dimensionless velocity and concentration distribution across the cross-section and along the jet axis match well with experimental results of JPUE.
@@
 
-Currently existing 3D models focus on certain aspect of the volcanic plume (PDAC on pyroclastic flow, ATHAM on microphysics, and SK-3D on entrainment with higher accuracy and higher order of accuracy) and hence, naturally, different assumptions were made in these models.
+Currently existing 3D models focus on certain aspect of the volcanic plume (PDAC on pyroclastic flow, ATHAM on microphysics, and SK-3D on entrainment with higher resolution and higher order of accuracy) and hence, naturally, different assumptions were made in these models.
 However, these different aspects of volcanic plumes are not independent, but are actually coupled.
@@
 To adapt post processing originally proposed for mesh-based method, we need to project simulation results onto a pre-defined regular mesh.
-Fig.
-\ref{appfig:1D-shock-tests-verification} shows the procedure for posting process.
+As shown in Fig.
+\ref{appfig:1D-shock-tests-verification}, the basic steps for such projection are:
+\begin{itemize}
+\item obtain raw simulation results carried by particles that irregularly distribute in the space
+\item create regular grids
+\item search for neighbour paricles for each node of the regular grids.
+\item interpolate physical quanties from neighbour particle onto corresponding node of regular grids according to Eq. (\ref{eq:SPH-fundamental-principle})
+\end{itemize}
 

### 2018-03-16 Reply-to-RC2.tex
Reply to second reviewer is added

\documentclass[10pt,a4paper]{article}
%\documentclass[10pt,a4paper]{beamer}
%\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\usepackage{kpfonts}
\usepackage[numbers]{natbib}
%other options include: round, curly, angle, or default options

\usepackage{media9}
\begin{document}
\author{\Large \textbf{by Zhixuan Cao et al.}}

\title{ \Large{ Authors' reply to the comments by Anonymous Referee (Referee \#2) of the manuscript gmd-2017-119} \\
\LARGE \textbf{Plume-SPH 1.0: A three-dimensional, dusty-gas volcanic plume model based on smoothed particle hydrodynamics}}

\date{\vspace{-5ex}}

\maketitle
The comments by the reviewer are recited in italics, followed by our reply in upright font. Equation, section, and figure numbers that  in our response are corresponding to these in the original 
version of GMD discussion paper. \\[12pt]

\textbf{\large General Comments}

\textit{After reading the paper, I think the proposed method is highly detailed and the type of discretization properly described.}

We would like to thank the reviewer for carefully reading our work and giving constructive comments.
We have revised the manuscript as shown in the supplemental PDF file, and we hope that we have dealt with all suggestions in an adequate manner. The revised manuscript is also attached.

On behalf of all co-authors. \\[1pt]

The following are our responses to reviewer's specific comments. Modifications are made accordingly in the revised manuscript. \\[12pt]

\textbf{\large Specific comments}

\textit{1. Please add the recent papers, which I consider, are related to your research: }

\textit{(a) Costa, A., Suzuki, Y., $\&$ Koyaguchi, T. (2018). Understanding the plume dynamics of explosive super-eruptions. Nature communications, 9(1), 654.}

\textit{(b) Terray, L., Gauthier, P. J., Salerno, G., Caltabiano, T., Spina, A. L., Sellitto, P., $\&$ Briole, P. (2018). A New Degassing Model to Infer Magma Dynamics from Radioactive Disequilibria in Volcanic Plumes. Geosciences, 8(1), 27.}

Thanks for pointing out the most recent progress in volcanic plume modeling. 
Work by \citet{costa2018understanding} is a strong backup for the necessarity of developing more comprehensive 3D computational model. We also came up with some application ideas for Plume-SPH after reading this paper. Citation of this paper is added in "Introduction" section to emphasize the advantage of developing more comprehensive plume model based on SPH.

We have been considering coupling volcanic plume model with magma reservoir model. It is critical for any plume simulation to use more accurate eruption conditions, which might be obtained from magma reservoir model. Thanks for pointing out a new degassing model to infer magma dynamics in volcanic plumes \cite{terray2018new}. Comments regarding coupling volcanic plume model with magma reservoir model is added in conclusion section of this paper along with proper citation.
\\[3pt]

\textit{It would be nice to provide a theoretical bound for the computational effort of your method for a single simulation step and the numerical-grid resolution. You can make use of, for instance, the number of long-computations (e.g., matrix- vector products). Then, please provide a theoretical bound for such value when computations are performed across different processors.
}

My understanding on what does the reviewer mean by "theoretical bound for computational effort" is the time complexity. Apologize if we misunderstood that term.

Without accounting for detailed algorithm, the time complexity for raw SPH method is $O(NN)$. Where, $N$ is the total number of particles (SPH particles are essentially discretization points and equivalent to grids in mesh-based methods). By adopting background mesh and compact-supported kernel function, the time complexity will be reduced to $O(MN)$. Where $M$ is the average of number of particles within the compact support of kernel function. Please be noted that number of particles within the compact support of kernel function is not constant in our case.

When computations are performed across different processors, theoretically speaking, the total computational time will be $(s+\frac{p}{n}+\bar{p})$. Where $(s+p)\sim O(MN)$ is sequential computational time with $p$ representing computation that parallelizable and $s$ representing computation that can not be parallelized. $n$ is total number of processors. $\bar{p}$ is extra computation introduced by parallelization, for example, 
unecessary duplicating work, communication overhead, time to split and combine.

It requires analyzing on detailed algorithms to obtain more accurate theorical bound. We admit that such theoretical time complexity would be very helpful for future performance optimiztion. However, considering HPC is not the major focus of this paper, we decided to reserve this topic for future research. We thanks reviewer again for proposing this idea.\\[3pt]

\textit{It is not clear for me how the parallelization is performed, for instance, for a given time, do you split the domain across different processors? in such case, what constraints must be satisfied at each local domain in order to guaranty a consistent numerical solution of your equations?. Other possibility is to speed-up matrix computations, is this your case? or both?}

The parallelization is not covered with enough details in this paper as it has been addressed in a separate paper \cite {cao2017data}. Sorry for confusing you. The parallelization is achieved only by splitting computational domain (Fig. 2(c) in GMD discussion paper shows a typical domain decomposition). No matrix computation is involved in SPH scheme, so there is no parallelization regarding solving matrix in our case.
For any subdomain, information from its neighbouring subdomains is required when updating physical quantities.   To guarantee consistency, data is synchronized after each updating of physical quantity.

To address your questions and clean up confusions, we made major revision in section 3.10 (is section 3.9 in the revised manuscript). In addition, portion of our response to reviewer's second comments is also added in section 3.10.

We thanks reviewer again for the careful reading and constructive comments.\\[12pt]

\textbf{\large Major adjustments of the manuscripts}

Here is a summary on major adjustments made in the revised manuscript: 
Major revision is made in "Parallellism and performance" section (sectiom 3.10 in original manuscript and 3.9 in the revised manuscript).

\bibliographystyle{plainnat}
\bibliography{Reference}
\end{document}

### 2018-03-17 Manuscripts_GMD_ap_ref2.tex
updated the manuscripts

--- 2017-09-22 Manuscripts_GMD_ap_ref1.tex
@@
 The code should thus be also of general interest to the much larger community of researchers using and developing SPH based tools.
-In particular, the $SPH-\varepsilon$ turbulence model is to capture mixing at unresolved scales, heat exchange due to turbulence is calculated by a Reynolds analogy and a corrected SPH is used to handle tensile instability and deficiency of particle distribution near the boundaries.
+In particular, the $SPH-\varepsilon$ turbulence model is to capture mixing at unresolved scales.
+Heat exchange due to turbulence is calculated by a Reynolds analogy and a corrected SPH is used to handle tensile instability and deficiency of particle distribution near the boundaries.
 We also developed methodology to impose velocity inlet and pressure outlet boundary conditions, both of which are scarce in traditional implementations of SPH.
@@
 The development of 2D and 3D, time-dependent, and multiphase numerical models for volcanic plumes has provided new explanations for many features of explosive volcanism.
+For example, a recent study based on 3D fluid dynamical model \citep{costa2018understanding} shows that simple extrapolations of integral models for Plinian columns to those of super-eruption plumes are not valid and their dynamics diverge from current ideas of how volcanic plumes operate.
 One of the earliest of these is the 3D model PDAC (Pyroclastic Dispersion Analysis Code) \citep{neri2003multiparticle}  which is a non-equilibrium, multiphase, 3D compressible flow model.
@@
 \end{equation}
-where, $C_v$ is the specific heat with constant volume, $R$ is the gas constant. $\xi$ with subscript is the mass fraction of corresponding constituent. The subscript 
-$m$ represents mixture of ejected material and air, $s$ represents solid portion in the ejected material, $g$ represents gas portion in the ejected material and $a$ represents air.
+where, $C_v$ is the specific heat with constant volume, $R$ is the gas constant.
+$\xi$ is the mass fraction of.
+The subscript $m$ represents mixture of ejected material and air, $s$ represents solid portion in the ejected material, $g$ represents gas portion in the ejected material, $a$ represents air, $0$ represents phyiscal properties of erupted material.
 $\xi_{g0}$ is the mass fraction of vapor in the erupted material.
@@
 \textbf{v} = const =\{u,v,w\}^T \label{eq:erupt_bc_v}\\
-\dfrac{\partial e}{\partial n}=\dot M e /\left(\pi r^2\right) \label{eq:erupt_bc_e}
+\dfrac{\partial e}{\partial n}=\dot M e_0 /\left(\pi r^2\right) \label{eq:erupt_bc_e}
 \end{align} 
-where $r$ is the radius of the vent, $n$ is the direction normal to the boundary.
+where $r$ is the radius of the vent, $n$ is the direction normal to the boundary, $e_0$ is internal energy of erupted material which can be calculated based on temperature and specific heat of erupted material.
 
@@
 \end{equation}
-where the subscript $a$ and $b$ represents air particles (phase 1) while $i$ and $j$ represents erupted material. $\beta$ and $\alpha$ represents either erupted material or air.
+where the subscript $a$ and $b$ represents air particles (phase 1) while $i$ and $j$ represents particles of erupted material.
+$\beta$ and $\alpha$ represents either erupted material particles or air particles.
 $\rho_a^a$ is density of phase 1 (air). 
@@
 Interface construction will become necessary and important when attempt to include the effects of mixing by resolving the detailed interface structure and dynamics of turbulence. 
-As a Lagrangian method, interface tracking in SPH is explicit through capturing of the locations of the particles, much simplier than Euleruan Methods.
+As a Lagrangian method, interface tracking in SPH is explicit through capturing of the locations of the particles, much simplier than Euleruan methods.
 The existence of complex evolving interfaces between phases presents severe challenges to conventional Eulerian grid-based numerical methods.
@@
 Quantifying these mixing processes in real implementation is challenging because of the scale disparity between the large-scale fluid motion and the diffusion processes on interface that ultimately lead to mixing.
-%The energy transfer between these scales occurs through turbulent motion, created either by fluid instabilities or by breaking internal waves.
-For volcanic plume, a shearing flow, the fluid instabilities dominated the mixing process.
-One way to include enough turbulence in the model is to use the SPH-SPS (SPH sub-particle-scale) turbulence model. 
-Another way is to use fine enough resolution (namely, the direct numerical simulation method) at the expense of much higher computational cost which results from both increase in number of particles and decrease in time steps constrained by the CFL (Courant-Friedrichs-Lewy) condition.
-Here we choose the SPH-SPS method.
-Among existing SPH-SPS turbulence models \citep{holm1999fluctuation, monaghan2002sph, violeau2007numerical, monaghan2011turbulence}, we adopt a LANS type turbulence model, the $SPH-\varepsilon$ turbulence model \citep{monaghan2011turbulence}.
+Ideally, one would like to be able to include the effects of mixing on the large scale dynamics without resolving the detailed interface structure and dynamics of turbulence to reduce computational cost.
+To resolve all turbulent exchange at all different scales and sub-particle scale mixing with relative coarse resolution, a SPH-SPS turbulence models should be included.
+Among existing SPH-SPS (SPH sub-particle-scale) turbulence models \citep{holm1999fluctuation, monaghan2002sph, violeau2007numerical, monaghan2011turbulence}, we adopt a LANS type turbulence model, the $SPH-\varepsilon$ turbulence model \citep{monaghan2011turbulence}.
 However, the $SPH-\varepsilon$ turbulence model was proposed only for incompressible flow.
@@
 For more complicated topography, it has been shown in other work \citep {kumar2013parallel} that it also works as well.
-We do have some concern regarding potential limitation of this method of deployment of ghost particles for more complicated boundaries in three dimensions but that is one that is not related to the plume simulations that are the target here.
+We do have concern regarding potential limitation of this method of deployment of ghost particles for more complicated boundaries in three dimensions.
+But that is not related to the plume simulations in current plume model.
 
@@
 
-Any implementation of SPH code requires efficient searching and updating of neighbors during simulation.
+The time complexity for raw SPH method is $O(NN)$.
+Where, $N$ is the total number of particles (SPH particles are essentially discretization points and equivalent to grids in mesh-based methods).
+Efficient neighbour searching and compact-supported kernel function helps reduce computational cost.
 Of the many possible choices we adopt a background grid which was proposed by \citet {monaghan1985refined} and is quite popular in parallel SPH.
-This background grid is also used for domain decomposition in SPH.
+Then the time complexity is reduced to $O(MN+mN)$.
+Where $m$ is the average of number of particles within the compact support of kernel function, $M$ is average number of particles in sub-domain within which neighbour searching is carried out.
+This background grid is also used for domain decomposition.
 We refer to the elements of background grid, namely squares for two dimensions and cubic for three dimensions, as buckets.
@@
 
-As for domain decomposition, even though more complicated graph-based partitioning tools \citep {biswas1999experiments} might get higher quality decomposition, they requires much more effort in programming and computation.
-So we adopt an easy-programming scheme based on SFC \citep {patra1999efficient}.
+The parallelization is achieved by splitting computational domain into subdomains.
+Each subdomain is computed by a single processor.
+For any subdomain, information from its neighbouring subdomains is required when updating physical quantities.
+To guarantee consistency, data is synchronized after each updating of physical quantity.
+Even though more complicated graph-based partitioning tools \citep {biswas1999experiments} might get higher quality decomposition, they requires much more effort in programming and computation.
+So we adopt an easy-programming scheme based on SFC \citep {patra1999efficient} to decompose the computational domain.
 Subfigure $c$ of Fig.
@@
 
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[scale=0.35]{strong}
-%\caption{Strong scalability test result}
-%\label{fig:strong_scale}
-%\end{figure}
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[scale=0.35]{weak_scale}
-%\caption{Weak scalability test result}
-%\label{fig:weak_scale}
-%\end{figure}
-%
 \begin{figure}[!t]
@@
         \centering
-        \includegraphics[width=0.99 \textwidth]{./gmd-mssfrc-yz}
+        \includegraphics[width=0.99 \textwidth]{./gmd-mssfrc-xz}
     \end{minipage}% 
@@
         \centering
-        \includegraphics[width=0.99 \textwidth]{./gmd-mssfrc-yz-zoomed}
+        \includegraphics[width=0.99 \textwidth]{./gmd-mssfrc-xz-zoomed}
     \end{minipage}%  
-    \caption{Mass fraction for $t=300s$ after eruption.
+    \caption{Mass fraction for $t=500s$ after eruption.
 Figures on the first row are visualization of SPH simulation results.
@@
 The lowest portion of the plume represents erupted material in eruption vent (the underground portition).
-Figures at the second row show contour of mass fraction and velocity quiver on y-z plane.
+Figures at the second row show contour of mass fraction and velocity quiver on x-z plane.
 These figures are plotted utilizing post-processed data (see Appendix \ref{app:project-SPH-grid}).
@@
 
-\conclusions  \label{sec:conclusion}%% \conclusions[modified heading if necessary]
+\conclusions  \label{sec:conclusion}%% 
 A new plume model was developed based on the SPH method.
@@
 Our code will also be made available in the open source form for the community to enhance.
-%In all existing 3D plume models, numerical calculations are performed on a non-uniform (vertically and horizontally stretched) grid, with different grid resolutions and domain sizes.
-However, implementation of different resolution in SPH methods is not straight forward and requires more work.
+Besides improving plume model, coupling volcanic plume model with magma reservoir models \citep[e.g.][]{terray2018new}, which could provide more accurate eruption conditions, might improve accuracy of volcanic plume simulation.
 

### 2018-03-17 Reply-to-RC2.tex
updated the manuscripts

@@
 Where, $N$ is the total number of particles (SPH particles are essentially discretization points and equivalent to grids in mesh-based methods).
-By adopting background mesh and compact-supported kernel function, the time complexity will be reduced to $O(MN)$.
-Where $M$ is the average of number of particles within the compact support of kernel function.
+By adopting background mesh and compact-supported kernel function, the time complexity will be reduced to $O(MN+mN)$.
+Where $m$ is the average of number of particles within the compact support of kernel function, $M$ is number of particles among which neighbour searching is carried out.
 Please be noted that number of particles within the compact support of kernel function is not constant in our case.
@@
 When computations are performed across different processors, theoretically speaking, the total computational time will be $(s+\frac{p}{n}+\bar{p})$.
-Where $(s+p)\sim O(MN)$ is sequential computational time with $p$ representing computation that parallelizable and $s$ representing computation that can not be parallelized.
+Where $(s+p)\sim O(MN+mN)$ is sequential computational time with $p$ representing computation that parallelizable and $s$ representing computation that can not be parallelized.
 $n$ is total number of processors.

### 2018-03-17 Manuscripts_GMD_ap_ref2.tex
add diff PDF, removed useless files

@@
 where, $C_v$ is the specific heat with constant volume, $R$ is the gas constant.
-$\xi$ is the mass fraction of.
+$\xi$ is the mass fraction of erupted material.
 The subscript $m$ represents mixture of ejected material and air, $s$ represents solid portion in the ejected material, $g$ represents gas portion in the ejected material, $a$ represents air, $0$ represents phyiscal properties of erupted material.
@@
 \end{align} 
-where $r$ is the radius of the vent, $n$ is the direction normal to the boundary, $e_0$ is internal energy of erupted material which can be calculated based on temperature and specific heat of erupted material.
+where $r$ is the radius of the vent, $n$ is the direction normal to the boundary, $e_0$ is specific internal energy of erupted material which can be calculated based on temperature and specific heat of erupted material.
 
@@
 In our current model, the ground is assumed to be  flat.
-For more complicated topography, it has been shown in other work \citep {kumar2013parallel} that it also works as well.
+For more complicated topography, it has been shown in other work \citep {kumar2013parallel} that this method works as well.
 We do have concern regarding potential limitation of this method of deployment of ghost particles for more complicated boundaries in three dimensions.
-But that is not related to the plume simulations in current plume model.
+Fortunately, current model does not involve complicated wall boundary.
 
@@
 The time complexity for raw SPH method is $O(NN)$.
-Where, $N$ is the total number of particles (SPH particles are essentially discretization points and equivalent to grids in mesh-based methods).
-Efficient neighbour searching and compact-supported kernel function helps reduce computational cost.
+Where, $N$ is the total number of SPH particles.
+Efficient neighbour searching and compact-supported kernel function could help to reduce computational cost.
 Of the many possible choices we adopt a background grid which was proposed by \citet {monaghan1985refined} and is quite popular in parallel SPH.
@@
 For any subdomain, information from its neighbouring subdomains is required when updating physical quantities.
-To guarantee consistency, data is synchronized after each updating of physical quantity.
+To guarantee consistency, data is synchronized if physical quantities are updated.
 Even though more complicated graph-based partitioning tools \citep {biswas1999experiments} might get higher quality decomposition, they requires much more effort in programming and computation.
@@
 
-Figure \ref{fig:pinatubo-simulation-results-vis} shows the mass fraction of simulated volcanic plume at 300s after eruption, at which time the plume starts spreading radially.
-A contour plot of the mass fraction on the vertical cross-section (y-z plane) were also provided.
+Figure \ref{fig:pinatubo-simulation-results-vis} shows the mass fraction of simulated volcanic plume at 500s after eruption, at which time the plume starts spreading radially.
+A contour plot of the mass fraction on the vertical cross-section (x-z plane) were also provided.
 The zoomed view of the quivier plot shows detailed entrainment of air at the margin of the plume.

### 2018-03-19 Manuscripts_GMD_ap_ref2.tex
cleanups

@@
 Less attention has been paid to developing more flexible data management schemes for more complicated problems.
-Motivated by techniques developed for mesh based methods, we develop a complete framework for parallelizing SPH with distrubted memory parallelism (MPI) allowing flexible and efficient data access.
-
-The time complexity for raw SPH method is $O(NN)$.
-Where, $N$ is the total number of SPH particles.
-Efficient neighbour searching and compact-supported kernel function could help to reduce computational cost.
+Motivated by techniques developed for mesh based methods, we develop a complete framework for parallelizing SPH with distriubted memory parallelism (MPI) allowing flexible and efficient data access.
+
+The time complexity for an SPH method is $O(N^2)$, where, $N$ is the total number of SPH particles.
+Efficient neighbor searches and compact-supported kernel functions can help to reduce computational cost.
 Of the many possible choices we adopt a background grid  which was proposed by \citet {monaghan1985refined} and is quite popular in parallel SPH.
-Then the time complexity is reduced to $O(MN+mN)$.
-Where $m$ is the average of number of particles within the compact support of kernel function, $M$ is average number of particles in sub-domain within which neighbour searching is carried out.
+Then the time complexity is reduced to $O(MN+mN)$, where $m$ is the average of number of particles within the compact support of the kernel function, $M$ is the average number of particles in sub-domain within which neighbor searches are carried out.
 This background grid is also used for domain decomposition.
 We refer to the elements of background grid, namely squares for two dimensions and cubic for three dimensions, as buckets.
-As for the actual storage of the physical quantities associated to each particle, different strategies have been adopted in existing implementations of SPH.
+As for the actual storage of the physical quantities  different strategies have been adopted in existing implementations of SPH.
 
@@
 As for the management of data, we adopt hash tables to store pointers to particles and buckets, which gives us not only flexibility of deleting and adding of elements, but also quicker access compared with linked lists.
-Instead of using the "nature manner" to number particles, we adopt a space filling curve (SFC) based index to give each particle and background bucket an unique identifier -- a strategy known to preserve data locality at minimal cost.
+Instead of using the "natural order" to number particles, we adopt a space filling curve (SFC) based index to give each particle and background bucket an unique identifier -- a strategy known to preserve data locality at minimal cost.
 The SFC based numbering strategy is further extended to include time step information so that particles added at the same position but different time have different identifiers.

### 2018-03-19 Reply-to-RC2.tex
cleanups

--- 2018-03-16 Reply-to-RC2.tex
@@
-\documentclass[10pt,a4paper]{article}
+\documentclass{article}
 %\documentclass[10pt,a4paper]{beamer}
@@
 \maketitle
-The comments by the reviewer are recited in italics, followed by our reply in upright font. Equation, section, and figure numbers that  in our response are corresponding to these in the original 
+The comments by the reviewer are recited in italics, followed by our reply in upright font. Equation, section, and figure numbers   in our response are corresponding to these in the original 
 version of GMD discussion paper. \\[12pt]
@@
 
-Thanks for pointing out the most recent progress in volcanic plume modeling. 
-Work by \citet{costa2018understanding} is a strong backup for the necessarity of developing more comprehensive 3D computational model.
-We also came up with some application ideas for Plume-SPH after reading this paper.
+Thanks for pointing out some of the more recent work on volcanic plume modeling. 
+Work by \citet{costa2018understanding} motivates the development of more comprehensive 3D computational models like those in this paper.
+Reading this paper has inspired new application ideas for Plume-SPH.
 Citation of this paper is added in "Introduction" section to emphasize the advantage of developing more comprehensive plume model based on SPH.
 
-We have been considering coupling volcanic plume model with magma reservoir model.
-It is critical for any plume simulation to use more accurate eruption conditions, which might be obtained from magma reservoir model.
-Thanks for pointing out a new degassing model to infer magma dynamics in volcanic plumes \cite{terray2018new}.
-Comments regarding coupling volcanic plume model with magma reservoir model is added in conclusion section of this paper along with proper citation.
+We have considered coupling a volcanic plume model with a magma reservoir model.
+It is critical for any plume simulation to use more accurate eruption conditions, which might be obtained from a magma reservoir model.
+Thank you for pointing out the new degassing model to infer magma dynamics in volcanic plumes \cite{terray2018new}.
+Comments regarding coupling volcanic plume model with magma reservoir model are added in conclusion section of this paper along with proper citation.
 \\[3pt]
@@
 
-My understanding on what does the reviewer mean by "theoretical bound for computational effort" is the time complexity. Apologize if we misunderstood that term.
+Our interpretation of  the reviewer's request for "theoretical bound for computational effort" is the operational complexity of the algorithms used. Apologize if we  have misunderstood .
 
-Without accounting for detailed algorithm, the time complexity for raw SPH method is $O(NN)$.
-Where, $N$ is the total number of particles (SPH particles are essentially discretization points and equivalent to grids in mesh-based methods).
-By adopting background mesh and compact-supported kernel function, the time complexity will be reduced to $O(MN)$.
-Where $M$ is the average of number of particles within the compact support of kernel function.
-Please be noted that number of particles within the compact support of kernel function is not constant in our case.
+Without accounting for the detailed algorithm, the complexity for raw SPH method is $O(N*N)$ operations.
+Where, $N$ is the total number of particles (SPH particles are essentially equivalent to discretization points  in mesh-based methods).
+By adopting a background mesh and a compact-supported kernel function, the  complexity will be reduced to $O(MN+mN)$.
+Where $m$ is the average of number of particles within the compact support of the kernel function, $M$ is number of particles among which neighbor searches are carried out.
+Please note that number of particles within the compact support of the kernel function is not constant in our case.
 
-When computations are performed across different processors, theoretically speaking, the total computational time will be $(s+\frac{p}{n}+\bar{p})$.
-Where $(s+p)\sim O(MN)$ is sequential computational time with $p$ representing computation that parallelizable and $s$ representing computation that can not be parallelized.
-$n$ is total number of processors.
-$\bar{p}$ is extra computation introduced by parallelization, for example,
-unecessary duplicating work, communication overhead, time to split and combine.
+When computations are performed across different processors, theoretically speaking, the total computational time will be $(s+\frac{p}{n}+\bar{p})$,  where $(s+p)$(time needed for$ \sim O(MN+mN)$ operations) is total sequential computational time with $p$ representing computation that is parallelizable and $s$ represents computation that can not be parallelized,  $n$ is total number of processors and $\bar{p}$ is extra computation and communication introduced by parallelization, for example,
+unnecessary duplication of work, communication overhead, time to split and combine.
 
-It requires analyzing on detailed algorithms to obtain more accurate theorical bound.
-We admit that such theoretical time complexity would be very helpful for future performance optimiztion.
-However, considering HPC is not the major focus of this paper, we decided to reserve this topic for future research.
-We thanks reviewer again for proposing this idea.\\[3pt]
+It requires complete analysis of the detailed algorithms and instrumented data gathering to obtain  more accurate theoretical and achieved bounds.
+We agree that such  time complexity analysis would be very helpful for future performance optimization.
+However, considering HPC is not the major focus of this paper, we would like to reserve this topic for future research.
+We thank the reviewer again for proposing this idea.\\[3pt]
 
 \textit{It is not clear for me how the parallelization is performed, for instance, for a given time, do you split the domain across different processors?
-in such case, what constraints must be satisfied at each local domain in order to guaranty a consistent numerical solution of your equations?.
+in such case, what constraints must be satisfied at each local domain in order to guarantee a consistent numerical solution of your equations?.
 Other possibility is to speed-up matrix computations, is this your case?
@@
 The parallelization is not covered with enough details in this paper as it has been addressed in a separate paper \cite {cao2017data}.
-Sorry for confusing you.
-The parallelization is achieved only by splitting computational domain (Fig.
+The parallelization is achieved only by splitting the computational domain (Fig.
 2(c) in GMD discussion paper shows a typical domain decomposition).
-No matrix computation is involved in SPH scheme, so there is no parallelization regarding solving matrix in our case.
-For any subdomain, information from its neighbouring subdomains is required when updating physical quantities.
-To guarantee consistency, data is synchronized after each updating of physical quantity.
+No matrix computation is involved in the SPH scheme, so there is no parallelization regarding solving the matrix in our case.
+For any subdomain, information from its neighboring subdomains is required when updating physical quantities.
+To guarantee consistency, data is synchronized after each updates of physical quantities in the shared (or "halo") regions.
 
-To address your questions and clean up confusions, we made major revision in section 3.10 (is section 3.9 in the revised manuscript).
-In addition, portion of our response to reviewer's second comments is also added in section 3.10.
+To address your questions and clean up confusion, we made a major revision in section 3.10 (is section 3.9 in the revised manuscript).
+In addition, portion of our response to reviewer's second comments are also added in section 3.10.
 
-We thanks reviewer again for the careful reading and constructive comments.\\[12pt]
+We thank reviewer again for the careful reading and constructive comments.\\[12pt]
 
@@
 Here is a summary on major adjustments made in the revised manuscript: 
-Major revision is made in "Parallellism and performance" section (sectiom 3.10 in original manuscript and 3.9 in the revised manuscript).
+Major revision is made in "Parallelism and performance" section (section 3.10 in original manuscript and 3.9 in the revised manuscript). 
+\begin{thebibliography}{3}
+\providecommand{\natexlab}[1]{#1}
+\providecommand{\url}[1]{\texttt{#1}}
+\expandafter\ifx\csname urlstyle\endcsname\relax
+  \providecommand{\doi}[1]{doi: #1}\else
+  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi
 
-\bibliographystyle{plainnat}
-\bibliography{Reference}
+\bibitem[Cao et~al.(2017)Cao, Patra, and Jones]{cao2017data}
+Zhixuan Cao, Abani Patra, and Matthew Jones.
+\newblock Data management and volcano plume simulation with parallel sph method
+  and dynamic halo domains.
+\newblock \emph{Procedia Comput. Sci.,}, 108:\penalty0 786--795, 2017.
+
+\bibitem[Costa et~al.(2018)Costa, Suzuki, and
+  Koyaguchi]{costa2018understanding}
+Antonio Costa, Yujiro Suzuki, and Takehiro Koyaguchi.
+\newblock Understanding the plume dynamics of explosive super-eruptions.
+\newblock \emph{Nature communications}, 9\penalty0 (1):\penalty0 654, 2018.
+
+\bibitem[Terray et~al.(2018)Terray, Gauthier, Salerno, Caltabiano, Spina,
+  Sellitto, and Briole]{terray2018new}
+Luca Terray, Pierre-J Gauthier, Giuseppe Salerno, Tommaso Caltabiano,
+  Alessandro~La Spina, Pasquale Sellitto, and Pierre Briole.
+\newblock A new degassing model to infer magma dynamics from radioactive
+  disequilibria in volcanic plumes.
+\newblock \emph{Geosciences}, 8\penalty0 (1):\penalty0 27, 2018.
+
+\end{thebibliography}
+%\bibliographystyle{plainnat}
+%\bibliography{Reference}
 \end{document}

### 2018-03-21 Reply-to-RC2.tex
Final Response, Figures

@@
 \maketitle
-The comments by the reviewer are recited in italics, followed by our reply in upright font. Equation, section, and figure numbers   in our response are corresponding to these in the original 
+The comments by the reviewer are recited in italics, followed by our reply in upright font.
+Revision are highlighted with blue.
+Equation, section, and figure numbers   in our response are corresponding to these in the original
 version of GMD discussion paper. \\[12pt]
@@
 Reading this paper has inspired new application ideas for Plume-SPH.
-Citation of this paper is added in "Introduction" section to emphasize the advantage of developing more comprehensive plume model based on SPH.
+{\color{blue} Citation of this paper is added in "Introduction" section} to emphasize the advantage of developing more comprehensive plume model based on SPH.
 
@@
 Thank you for pointing out the new degassing model to infer magma dynamics in volcanic plumes \cite{terray2018new}.
-Comments regarding coupling volcanic plume model with magma reservoir model are added in conclusion section of this paper along with proper citation.
+{\color{blue}Comments regarding coupling volcanic plume model with magma reservoir model are added in conclusion section of this paper along with proper citation.}
 \\[3pt]
@@
 Please note that number of particles within the compact support of the kernel function is not constant in our case.
+{\color{blue} Basic analysis on time complexity is added in section ``Parallellism and Performance".}
 
@@
 
-To address your questions and clean up confusion, we made a major revision in section 3.10 (is section 3.9 in the revised manuscript).
-In addition, portion of our response to reviewer's second comments are also added in section 3.10.
+To address your questions and clean up confusion, {\color{blue}we made a major revision in section 3.10 (is section 3.9 in the revised manuscript).
+In addition, portion of our response to reviewer's second comments are also added in section 3.10.}
 
@@
 
-Here is a summary on major adjustments made in the revised manuscript: 
-Major revision is made in "Parallelism and performance" section (section 3.10 in original manuscript and 3.9 in the revised manuscript). 
+{\color{blue}Here is a summary on major adjustments made in the revised manuscript: 
+Major revision is made in "Parallelism and performance" section (section 3.10 in original manuscript and 3.9 in the revised manuscript).}
+
 \begin{thebibliography}{3}

### 2018-06-20 FinalSubmit.tex
ready for final submit

--- 2018-03-19 Manuscripts_GMD_ap_ref2.tex
@@
 
-%>>>>>>> 0d96d9682fb484b8b159e8d09a0eb89723d3a0f8
 % \Author[affil]{given_name}{surname}
@@
 \runningauthor{Z. Cao et al}
-%<<<<<<< HEAD
 \correspondence{Abani Patra (abani@buffalo.edu)}
-%=======
-%\correspondence{Abani Patra (abani.patra@gmail.com)}
-%>>>>>>> 0d96d9682fb484b8b159e8d09a0eb89723d3a0f8
+
 \received{}
@@
 cross with experimental results of JPUE (jet or plume that is ejected from a nozzle into a uniform environment). 
-%The model is validated by comparing against simulation results from other 3D plume models.
 Profiles of several integrated variables are compared with those calculated by existing 3D plume models for an eruption with the same MER (mass eruption rate) estimated for the Pinatubo eruption of June 15 1991.
@@
 \introduction  %% \introduction[modified heading if necessary]
-%<<<<<<< HEAD
 \subsection{Volcanic ash hazards}
@@
 This paper reports on a new 3D (three dimensional) volcanic plume model designed to exploit the advantages of mesh-free methods for 3D modeling of such plumes that involve multiphase free boundary flows.
-
-%The dynamics of explosive eruption clouds has been a central issue of volcanology science for a long time \citep{settle1978volcanic, wilson1978control}.
-%The basic physics of volcanic plumes has been studied for more than half a century \citep{morton1956turbulent, settle1978volcanic, wilson1978control}.
-The effects of magma type and vent conditions \citep{woods1988fluid, woods1995decompression}, atmospheric conditions \citep{ woods1993moist, sparks1997volcanic, bursik2001effect}, external surface water \citep{koyaguchi1996formation}, thermal disequilibrium, and particle fallout \citep{woods1991particle} have been assessed in increasing details based on observation, experiments and numerical studies.
 
@@
 To summarize, each 3D model has its own focus based on the problem of interest and modeling/numerical choices made. 
- %Different simplification was made based on their scope. Improvements have been making since their birth. 
 Accuracy of simulation (depending on comprehensiveness of the model, resolution of discretization, numerical error, and order of accuracy) and simulation time (depending on number of governing equations, resolution, numerical methods and parallel techniques) are always conflicting considerations in 3D plume simulations.
@@
 \begin{itemize}
-%\item For mesh based method, either interface tracking (Lagrangian) \citep{harlow1965numerical, wrobel1991computational, cheng1995simplified} or interface capturing (Eulerian) \citep{hirt1981volume, youngs1982time, gerlach2006comparison, gopala2008volume} methods are used to reconstruct the flow interface of free boundary flow.
-High computational cost, a tendency to form numerical instabilities and the inability to track complex topological changes are the significant drawbacks of tracking techniques \citep{hirt1981volume, unverdi1992front, anderson1998diffuse}.
-For interface capturing (Eulerian) method, the surrendering of surface detail before the phase transport calculation means that interface reconstruction is required between time steps to recover the interface information, which needs additional numerical effort \citep{hirt1981volume, youngs1982time}.
-Since SPH is able to adaptively adjust the discretization and automatically construct the interface, SPH does not require additional numerical effort for interface construction and therefore is more suitable for volcanic plume simulation.
 \item Advection term in the Navier-Stokes equations does not appear explicitly in discretized formula of SPH (as illustrated in Eq. (\ref{eq:gov-nc-rho}) to Eq. (\ref{eq:gov-nc-e})).
@@
 \end{align}
-%Governing equations in Lagrangian form, Eq.
-(\ref{eq:gov-nc-rho}) to Eq.(\ref{eq:gov-nc-e}), and Eulerian form, Eq.
-(\ref{eq:gov-cs-rho}) to Eq.
-(\ref{eq:gov-cs-e}), are a\dfrac{•}{•}ly equivalent, as deriving the Lagrangian form from the Eulerian is purely based on mathematics without any additional physical assumptions.
 
@@
 The boundaries are categorized as the velocity inlet (a circular area at the center of the bottom of the box), wall boundary (box bottom) and pressure outlet (other faces of the box).
-%probably should add a section to list the input parameters in the verification section
 
@@
 \end{equation}
-%$\alpha$ and $\beta$ are two parameters that are free to be adjusted for each case. 
 $\alpha$ and $\beta$ are two parameters that can be adjusted for different cases.
@@
 $\eta$ is usually taken as 0.1 to prevent singularities.
-%when $\textbf{x}_{ab} = 0$.
 
@@
 $w_{a b}\left(h\right)$ is a concise form of $w\left(\textbf{x}_a - \textbf{x}_b, h\right)$ and from here on, we will use this concise form.
-%However, these discretized formulations will not guarantee conservation of entropy.
-We can obtain conservation of entropy by giving up on conservation of mass or thermal energy adopting  alternative discretized formulations \citep{monaghan1992smoothed}.
 As a Lagrangian method, particle position is also updated at every time step.
@@
 where the source term $\dot{\rho}$ can be a "sink" of erupted vapor due to its phase change.
-%(see microphysics modelling in ATHAM \citep{oberhuber1998volcanic}).
 $D$ is a drag force coefficient.
@@
 \begin{figure}
-\includegraphics[width=12cm]{Interface}
+\includegraphics[width=12cm]{Fig1}
 \caption{In the left figure, the blue particles (phaseID=1) represent air particles the red ones (phaseID=2) represent erupted material.
@@
 SPH requires less additional numerical effort for interface construction and therefore is more suitable for volcanic plume simulation.
-
-%SPH is able to handle multiphase problems with less additional numerical effort for dealing with interfaces by simply tagging particles of each phase.
 
@@
 
-%Even though free surface flow are in nature turbulent, detailed research on turbulence modeling in SPH method is rather scarce.
-Among the first papers on turbulence are the implementations of the $\alpha$ turbulence model \citep{holm1999fluctuation, monaghan2002sph} and the RANS (Reynolds averaged Navier-Stokes) approach \citep{violeau2007numerical}.
-The first wall-bounded LES (large eddy simulation) simulation with SPH was performed by \citet{issa2005numerical}.
-An improved version of $SPH-\alpha$ method  \citep{monaghan2002sph} is proposed by \citet{monaghan2011turbulence}, which we refer to as $SPH-\varepsilon$ method.
-We adopt $SPH-\varepsilon$ method in our model.
-Most of these papers constrain their focus to incompressible flow.
-%$SPH-\alpha$ model \citep{monaghan2002sph} is a powerful extension of the XSPH algorithm which reduces disorder at short length scales and retains the constants of the motion.
-However, in that paper, energy conservation equation is discretized in a way that is exactly the same as usual SPH.
-There is no special treatment on energy conservation equation.
-The recent improved version \citep{monaghan2011turbulence} of $SPH-\alpha$ method focuses on incompressible flow only.
-%
-
 \subsubsection{Langrangian average in $SPH-\varepsilon$}
@@
 Notice that if $l$ is constant: 
-%Please notice that the above equation is slightly different from the original equation in Monaghan's paper (Eq.
-(2.17)).
-The original equation is based on the assumption that density is uniform initially.
-This assumption is not general and atmosphere density is stratified in our plume simulation.
-And the slight modification will not influence following derivation if $l$ is uniform and keep constant because:
-%
-%{\bf WE NEED TO SAY SOMETHING ABOUT INCOMPRESSIBILITY HERE}
-%
 \begin{equation}
@@
 We adopt this strategy.
-%PDAC \citep{neri2003multiparticle} also adopted a Reynolds analogy in its turbulence modeling.
 The additional terms associated with molecular diffusion and turbulent transport in the energy equation are either modeled in different ways or neglected sometimes \citep{NASACompressibleTurbulence}.
@@
 All boundary conditions are imposed by ghost particles. Subfigure $b$ of Fig. \ref{fig:bc_and_domain_decomp} shows how boundaries are deployed.
-%Physical quantities of particles, including real particles and different types of ghost particles, are updated differently.
 \subsubsection{Wall boundary condition}
@@
 \begin{figure}
-\includegraphics[width=18cm]{t120_bc_proc}
+\includegraphics[width=18cm]{Fig2}
 \caption{A cross section view of the simulation domain in $y-z$ plane at 66 seconds.
@@
 		\centering
-		\includegraphics[scale=0.45]{strong}
+		\includegraphics[scale=0.45]{Fig3}
 		\caption{Strong scalability test result}
@@
 		\centering
-		\includegraphics[scale=0.45]{weak_scale}
+		\includegraphics[scale=0.45]{Fig4}
 		\caption{Weak scalability test result}
@@
 \centering
-\includegraphics[scale=0.35]{adj_vs_no}
+\includegraphics[scale=0.35]{Fig5}
 \caption{The effect of domain adjusting on simulation time.
@@
     \centering
-    \begin{minipage}{.325\textwidth}
-        \centering
-        \includegraphics[width=0.99 \textwidth]{./GSPH-Sod-verification-e}
-    \end{minipage}%
-    \begin{minipage}{.325 \textwidth}
-        \centering
-        \includegraphics[width=0.99 \textwidth]{./Sod-verification-e}
-    \end{minipage}%
-    \begin{minipage}{.325 \textwidth}
-        \centering
-        \includegraphics[width=0.99 \textwidth]{./d-exp-verification-e}
-    \end{minipage}%   
+	  \includegraphics[width=18cm]{Fig6} 
     \caption{Comparison of specific internal energy of simulation results agaist analytical results for shock tube tests.
@@
     \centering
-    \begin{minipage}{.425\textwidth}
-        \centering
-        \includegraphics[width=8cm]{vel_cross}
-    \end{minipage}%
-    \begin{minipage}{.425 \textwidth}
-        \centering
-        \includegraphics[width=8cm]{conc_cross}]
-    \end{minipage}%  
+	  \includegraphics[width=16cm]{Fig7}
     \caption{Dimensionless concentration and velocity distribution across the cross-section.}
@@
     \centering
-    \begin{minipage}{.425\textwidth}
-        \centering
-        \includegraphics[width=8cm]{velo_along_axis}
-    \end{minipage}%
-    \begin{minipage}{.425 \textwidth}
-        \centering
-        \includegraphics[width=8cm]{conc_along_axis}
-    \end{minipage}%  
+	\includegraphics[width=16cm]{Fig8} 
     \caption{The left plot is normalized jet width (which determined based on velocity) along the centerline. The right plot shows normalized concentration along the centerline.}
@@
 \end{figure}
-%
-%\begin{figure}
-%\includegraphics[width=8cm]{vel_cross}
-%\caption{Dimensionless velocity distribution across the cross-section.}
-%\label{fig:JPUE_cross-section_vel}
-%\end{figure}
-%
-%\begin{figure}
-%\includegraphics[width=8cm]{conc_cross}
-%\caption{Dimensionless concentration distribution across the cross-section.}
-%\label{fig:JPUE_cross-section_conc}
-%\end{figure}
-
-%\begin{figure}
-%\includegraphics[width=8cm]{velo_along_axis}
-%\caption{Dimensionless velocity and concentration distribution along centraline.}
-%\label{fig:JPUE_along-axis_vel}
-%\end{figure}
-%\begin{figure}
-%\includegraphics[width=8cm]{conc_along_axis}
-%\caption{Dimensionless concetration distribution along centraline.}
-%\label{fig:JPUE_along-axis_conc}
-%\end{figure}
 
@@
 Hence, our confidence in the numerical correctness of our code is greatly reinforced.
-%What need to be noticed is that the mass of particle of phase 1 (blue) is 8 time of that of phase 2 (red).
-It is a trade off between resolution and computational expense to give particle of two phases different particles mass.
-Both advanced SPH techniques like adaptive particles size and better data management and access strategies need to be exploited to address this issue.
 
 \begin{figure}
-\includegraphics[width=8cm]{JPUE_entrainment.png}
+\includegraphics[width=8cm]{Fig9.png}
 \caption{The left figure shows particle distribution.
@@
 The right figure shows the mass fraction of erupted material at the moment corresponding to the left figure.}
-%What need to be noticed is that the mass of particle of phase 1 (blue) is 8 time of that of phase 2 (red).
-It is a trade off between resolution and computational expense to give particle of two phases different particles mass.
 \label{fig:Turb_mixing}
@@
 1B in \citep{costa2016results}).
-%Wind speed is not used in our current model even though it is given (by one component speed from west to east and another component speed from south to north). 
 Vertical distribution of temperature, pressure and density is used to establish stratified atmosphere.
@@
     \centering
-    \begin{minipage}{.45\textwidth}
-        \centering
-        \includegraphics[width=0.99 \textwidth]{./mssfrc-Diverging}
-    \end{minipage}%
-    \begin{minipage}{.45 \textwidth}
-        \centering
-        \includegraphics[width=0.99 \textwidth]{./mssfrc-Diverging-cut}
-    \end{minipage}%
-    \\
-    \begin{minipage}{.45 \textwidth}
-        \centering
-        \includegraphics[width=0.99 \textwidth]{./gmd-mssfrc-xz}
-    \end{minipage}% 
-    \begin{minipage}{.45 \textwidth}
-        \centering
-        \includegraphics[width=0.99 \textwidth]{./gmd-mssfrc-xz-zoomed}
-    \end{minipage}%  
+	  \includegraphics[width=16cm]{Fig10}
     \caption{Mass fraction for $t=500s$ after eruption.
@@
 \center
-\includegraphics[width=7cm]{Temp}
+\includegraphics[width=7cm]{Fig11}
 \caption{Temperature as a function of height}
@@
 \begin{figure}
-\includegraphics[width=15cm]{msfrac}
+\includegraphics[width=15cm]{Fig12}
 \caption{The mass fraction of entrained air, gas, and solid as a function of height.}
@@
 \center
-\includegraphics[width=7cm]{density_strong}
+\includegraphics[width=7cm]{Fig13}
 \caption{Density of the strong plume without wind after reaching its top height}
@@
 \center
-\includegraphics[width=7cm]{radius_strong}
+\includegraphics[width=7cm]{Fig14}
 \caption{Radius of the strong plume without wind after reaching its top height.}
@@
 This implies that choice of  turbulence model might play a critical role in plume simulation.
-
-%\subsubsection{Evolution of plume} 
-%
-%\begin{figure}
-%\center
-%\includegraphics[width=15cm]{strong_elevation}
-%\caption{This figure shows the dynamic evolution process of the Pinatubo with time.
-At the very beginning, it is driven by initial momentum and rises up quickly, mixing and entraining surround air.
-Around 100s, the jet stage finishes and it becomes a plume as there were enough air entrained.
-The plume keeps rising up until reaches the top height then starts spreading.}
-%\label{fig:strong_evolution}
-%\end{figure}
-%
-%\begin{figure}
-%\includegraphics[width=10cm]{t50}
-%\caption{Contour of mass fraction and velocity quiver at 50 seconds in plane x=0, the column keeps rising up and at the same time entraining air into the plume.
-The zoomed view (right figure) of the quiver plot clearly shows mechanism of air entrainment.
-The contour levels of mass fraction are 0.7, 0.5, 0.3, 0.1, $1\times 10 ^{-1}$, $1\times 10 ^{-2}$, $1\times 10 ^{-3}$, $1\times 10^{-4}$ and $1\times 10^{-5}$.}
-%\label{fig:strong_plume_t50}
-%\end{figure}
-%
-%Figure \ref{fig:strong_evolution} shows the overall evolution process of plume.
-Plume keeps rising up and entraining air into the plume until it reaches its top height(around t=300 seconds).
-Afterwards the mixtures of erupted material and air then fall back to neutral height and starts spreading radially.
-Detailed velocity quiver plot in Fig.
-\ref{fig:strong_plume_t50} shows entraining of air into the plume at the edge of column.
-The evolution of the plume further verifies our model.
 
@@
 
-%\subsection{Data availability}
-
 \appendix
@@
     \centering
-    \begin{minipage}{.325\textwidth}
-        \centering
-        \includegraphics[width=0.90 \textwidth]{./Project1}
-    \end{minipage}%
-    \begin{minipage}{.325 \textwidth}
-        \centering
-        \includegraphics[width=0.90 \textwidth]{./Project2}
-    \end{minipage}%
-    \begin{minipage}{.325 \textwidth}
-        \centering
-        \includegraphics[width=0.90 \textwidth]{./Project3}
-    \end{minipage}%   
+	\includegraphics[width=18cm]{FigA1} 
     \caption{Procedure of projection of simulation results carried by particles onto regular grids.
@@
 
-%
-%\subsection{}     %% Appendix A1, A2, etc.
-%
-%\appendixfigures  %% needs to be added in front of appendix figures in one-column style (\documentclass[acp, manuscript]{copernicus})
-%
-%\appendixtables   %% needs to be added in front of appendix tables in one-column style (\documentclass[acp, manuscript]{copernicus})
-%\authorcontribution{TEXT}
-
 \competinginterests{Authors have the standard conflict with employers and listed sponsors}
-%\disclaimer{TEXT}
 
@@
 We appreciate Antonio Costa for providing simulation results of existing 3D models, which are used in verification and validation section of this paper.
-We thank Matteo Cerminara for his helping on post-processing of plume simulation results.
+We thank Matteo Cerminara for his help on post-processing of plume simulation results and his careful review of our manuscript and constructive remarks.
+Comments by another anonymous reviewer are appreciated as well.
 Computational results reported here were performed at the Center for Computational Research at the University at Buffalo.
@@
 \bibliographystyle{copernicus}
-\bibliography{Reference}
-%% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,
-%% authors experienced with BibTeX only have to include the following two lines:
-%%
-%% \bibliographystyle{copernicus}
-%% \bibliography{example.bib}
-%%
-%% URLs and DOIs can be entered in your BibTeX file as:
-%%
-%% URL = {http://www.xyz.org/~jones/idx_g.htm}
-%% DOI = {10.5194/xyz}
-%% LITERATURE CITATIONS
-%%
-%% command                        & example result
-%% \citet{jones90}|               & Jones et al. (1990)
-%% \citep{jones90}|               & (Jones et al., 1990)
-%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
-%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
-%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
-%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
-%% \citeauthor{jones90}|          & Jones et al.
-%% \citeyear{jones90}|            & 1990
-%% FIGURES
-%% When figures and tables are placed at the end of the MS (article in one-column style), please add \clearpage
-%% between bibliography and first table and/or figure as well as between each table and/or figure.
-%% ONE-COLUMN FIGURES
-%%f
-%\begin{figure}[t]
-%\includegraphics[width=8.3cm]{FILE NAME}
-%\caption{TEXT}
-%\end{figure}
-%
-%%% TWO-COLUMN FIGURES
-%
-%%f
-%\begin{figure*}[t]
-%\includegraphics[width=12cm]{FILE NAME}
-%\caption{TEXT}
-%\end{figure*}
-%
-%
-%%% TABLES
-%%%
-%%% The different columns must be seperated with a & command and should
-%%% end with \\ to identify the column brake.
-%
-%%% ONE-COLUMN TABLE
-%
-%%t
-%\begin{table}[t]
-%\caption{TEXT}
-%\begin{tabular}{column = lcr}
-%\tophline
-%
-%\middlehline
-%
-%\bottomhline
-%\end{tabular}
-%\belowtable{} % Table Footnotes
-%\end{table}
-%
-%%% TWO-COLUMN TABLE
-%
-%%t
-%\begin{table*}[t]
-%\caption{TEXT}
-%\begin{tabular}{column = lcr}
-%\tophline
-%
-%\middlehline
-%
-%\bottomhline
-%\end{tabular}
-%\belowtable{} % Table Footnotes
-%\end{table*}
-%
-%
-%%% MATHEMATICAL EXPRESSIONS
-%
-%%% All papers typeset by Copernicus Publications follow the math typesetting regulations
-%%% given by the IUPAC Green Book (IUPAC: Quantities, Units and Symbols in Physical Chemistry,
-%%% 2nd Edn., Blackwell Science, available at: http://old.iupac.org/publications/books/gbook/green_book_2ed.pdf, 1993).
-%%%
-%%% Physical quantities/variables are typeset in italic font (t for time, T for Temperature)
-%%% Indices which are not defined are typeset in italic font (x, y, z, a, b, c)
-%%% Items/objects which are defined are typeset in roman font (Car A, Car B)
-%%% Descriptions/specifications which are defined by itself are typeset in roman font (abs, rel, ref, tot, net, ice)
-%%% Abbreviations from 2 letters are typeset in roman font (RH, LAI)
-%%% Vectors are identified in bold italic font using \vec{x}
-%%% Matrices are identified in bold roman font
-%%% Multiplication signs are typeset using the LaTeX commands \times (for vector products, grids, and exponential notations) or \cdot
-%%% The character * should not be applied as mutliplication sign
-%
-%
-%%% EQUATIONS
-%
-%%% Single-row equation
-%
-%\begin{equation}
-%
-%\end{equation}
-%
-%%% Multiline equation
-%
-%\begin{align}
-%& 3 + 5 = 8\\
-%& 3 + 5 = 8\\
-%& 3 + 5 = 8
-%\end{align}
-%
-%
-%%% MATRICES
-%
-%\begin{matrix}
-%x & y & z\\
-%x & y & z\\
-%x & y & z\\
-%\end{matrix}
-%
-%
-%%% ALGORITHM
-%
-%\begin{algorithm}
-%\caption{_}
-%\label{a1}
-%\begin{algorithmic}
-%_
-%\end{algorithmic}
-%\end{algorithm}
-%
-%
-%%% CHEMICAL FORMULAS AND REACTIONS
-%
-%%% For formulas embedded in the text, please use \chem{}
-%
-%%% The reaction environment creates labels including the letter R, i.e. (R1), (R2), etc.
-%
-%\begin{reaction}
-%%% \rightarrow should be used for normal (one-way) chemical reactions
-%%% \rightleftharpoons should be used for equilibria
-%%% \leftrightarrow should be used for resonance structures
-%\end{reaction}
-%
-%
-%%% PHYSICAL UNITS
-%%%
-%%% Please use \unit{} and apply the exponential notation
+%\bibliography{Reference}
+\bibliography{Ref}
 \end{document}
